### 2015-01-20 analyses/ps1.Rmd
Computed percent changes

---
title: 'Psych 254 W15 PS #1'
author: "Mike Frank"
date: "January 11, 2015"
---

This is problem set #1, in which we hope you will practice the packages tidyr and dplyr. There are some great cheat sheets from [RStudio](http://www.rstudio.com/resources/cheatsheets/).

The data set
------------

This data set comes from a replication of [Janiszewski and Uy (2008)](http://dx.doi.org/10.1111/j.1467-9280.2008.02057.x), who investigated whether the precision of the anchor for a price influences the amount of adjustment.

In the data frame, the `Input.condition` variable represents the experimental condition (under the rounded anchor, the rounded anchor, over the rounded anchor). `Input.price1`, `Input.price2`, and `Input.price3` are the anchors for the `Answer.dog_cost`, `Answer.plasma_cost`, and `Answer.sushi_cost` items.

Preliminaries
-------------

I pretty much always clear the workspace and load the same basic helper functions before starting an analysis.

```{r prelims}
rm(list=ls())
setwd("/Users/rxdh/Repos/psych254/analyses/")
source("../helper/useful.R")
```

Note that I'm using a "relative" path (the "../helper") rather than an absolute path (e.g. "/Users/mcfrank/code/projects/etc..."). The relative path means that someone else can run your code by changing to the right directory, while the absolute path will force someone else to make trivial changes every time they want to run it.

Part 1: Data cleaning
---------------------

The first part of this exercise actually just consists of getting the data in a format usable for analysis. This is not trivial. Let's try it:

```{r data1}
d <- read.csv("../data/janiszewski_rep_exercise.csv")
str(d)
```

Fine, right? Why can't we go forward with the analysis?

**There are several ways in which the data need to be cleaned. One tip-off about this is that a number of variables that should be numeric end up as factors... In particular, we note:**
1. **Some participants appear multiple times (90 rows, but only 87 worker ids)**
2. **There are commas in some of the input.price numbers**
3. **someone entered their response for "dog cost" as a character string: "five hundred"**

HINT: try computing some summary statistics for the different items. Also, are there any participants that did the task more than once?

Fix the data file so that it looks reasonable.

**To fix these issues, we just open the .csv in another program (e.g. emacs), fix these little problems by hand (or using find/replace), and then save it under a different name, as "cleaned"**

Part 2: Making these data tidy
------------------------------

Now let's start with the cleaned data, so that we are all beginning from the same place.

```{r data2}
d <- read.csv("../data/janiszewski_rep_cleaned.csv")
```

This data frame is in *wide* format - that means that each row is a participant and there are multiple observations per participant. This data is not *tidy*.

To make this data tidy, we'll do some cleanup. First, remove the columns you don't need, using the verb `select`.

HINT: `?select` and the examples of helper functions will help you be efficient.

```{r select}
d.tidy <- select(d, WorkerId, Input.condition, 
                 starts_with("Answer"))
```

Try renaming some variables using `rename`. A good naming scheme is:

* consistent with case
* consistent with "." or "_" ( "_" is usually preferred)
* concise as will be comprehensible to others

Try using the `%>%` operator as well. So you will be "piping" `d %>% rename(...)`.

```{r rename}
d.tidy <- d.tidy %>% rename(condition = Input.condition,
                            dog_response = Answer.dog_cost,
                            plasma_response = Answer.plasma_cost,
                            sushi_response = Answer.sushi_cost)
```

OK, now for the tricky part. Use the verb *gather* to turn this into a *tidy* data frame.

HINT: look for online examples!

```{r gather}
d.tidy <- d.tidy %>% gather(name, cost, 
                 dog_response, plasma_response, sushi_response) 
```

Bonus problem: *spread* these data back into a wide format data frame.

```{r spread}
d.wide <- d.tidy %>% spread(name, cost)
```

Part 3: Manipulating the data using dplyr
-----------------------------------------

Try also using the dplyr `distinct` function to remove the duplicate participants from the raw csv file that you discovered in part 1.

```{r}
d.raw <- read.csv("../data/janiszewski_rep_exercise.csv")
d.unique.subs <- distinct(d.raw, WorkerId)
```

As we said in class, a good thing to do is always to check histograms of the response variable. Do that now, using either regular base graphics or ggplot. What can you conclude?

```{r}
g <- (ggplot(d.tidy, aes(x = cost))
      + geom_histogram())
g
```

**We can conclude that** 
1. **there are three clumps of data, corresponding to the three different objects (sushi, dog, plasma)**, 
2. **recalling that the anchors were set at 9, 2500, and 5000 (remember, we're not yet separating out classes of objects), we see that there seems to be a general trend of under-estimation**

OK, now we turn to the actual data anlysis. We'll be using dplyr verbs to `filter`, `group`,`mutate`, and `summarise` the data.

Start by using `summarise` on `d.tidy` to compute the mean bet across all participants. Note that this is simply taking the grand mean. Ultimately, we would like to compute the mean for different conditions and items, but this will come later. Right now we're just learning the syntax of `summarise`.

```{r}
# Note that worker A3S99S9NJ6X08J left "sushi_response" blank, so we 
# must exclude NAs
summarise(d.tidy, m = mean(cost, na.rm = TRUE))
```

This is a great time to get comfortable with the `%>%` operator. In brief, `%>%` allows you to pipe data from one function to another. So if you would have written:

```{r}
# d <- function(d, other_stuff)
```

you can now write:

```{r}
# d <- d %>% function(other_stufF)
```

That doesn't seem like much, but it's cool when you can replace:

```{r}
# d <- function1(d, other_stuff)
# d <- function2(d, lots_of_other_stuff, more_stuff)
# d <- function3(d, yet_more_stuff)
```

with

```{r}
# d <- d %>% 
#   function1(other_stuff) %>%
#   function2(lots_of_other_stuff, more_stuff) %>%
#   function3(yet_more_stuff)
```

In other words, you get to make a clean list of the things you want to do and chain them together without a lot of intermediate assignments.

Let's use that capacity to combine `summarise` with `group_by`, which allows us to break up our summary into groups. Try grouping by item and condition and taking means using `summarise`, chaining these two verbs with `%>%`.

```{r}
d.tidy %>% 
  group_by(name, condition) %>% 
  summarise(mean(cost, na.rm = TRUE))
```

OK, it's looking like there are maybe some differences between conditions, but how are we going to plot these? They are fundamentally different magnitudes from one another.

Really we need the size of the deviation from the anchor, which means we need the anchor value (the `Input.price` variables that we've ignored up until now). Let's go back to the data and add that in.

Take a look at this complex expression. You don't have to modify it, but see what is being done here with gather, separate and spread. Run each part (e.g. the first verb, the first two verbs, etc.) and after doing each, look at `head(d.tidy)` to see what they do.

```{r}
d.tidy <- d %>%
  select(WorkerId, Input.condition, 
         starts_with("Answer"), 
         starts_with("Input")) %>%
  rename(workerid = WorkerId,
         condition = Input.condition,
         plasma_anchor = Input.price1,
         dog_anchor = Input.price2,
         sushi_anchor = Input.price3,
         dog_cost = Answer.dog_cost,
         plasma_cost = Answer.plasma_cost, 
         sushi_cost = Answer.sushi_cost) %>%
  gather(name, cost, 
         dog_anchor, plasma_anchor, sushi_anchor, 
         dog_cost, plasma_cost, sushi_cost) %>%
  separate(name, c("item", "type"), sep = "_") %>%
  spread(type, cost)
```

Now we can do the same thing as before but look at the relative difference between anchor and estimate. Let's do this two ways:

* By computing absolute value of percentage change in price, and 
* By computing z-scores over items.

To do the first, use the `mutate` verb to add a percent change column, then compute the same summary as before.

```{r}
pcts <- d.tidy %>%
  mutate(pct_change = anchor / cost) %>%
  group_by(name, condition) %>% 
  summarise(mean(cost, na.rm = TRUE))  
```

To do the second, you will need to `group` first by item, compute z-scores with respect to items, then further group by condition.

HINT: `scale(x)` returns a complicated data structure that doesn't play nicely with dplyr. try `scale(x)[,1]` to get what you need.

HINT: by default, `group_by` undoes any previous groupings. If you want to add new grouping variables *on top* of pre-existing ones, specify `add = TRUE`, e.g., `d %>% group_by(var1) %>% group_by(var2, add = TRUE)`.

```{r}
z.scores <- d.tidy %>% ...
```

OK, now here comes the end: we're going to plot the differences and see if anything happened. First the percent change:

```{r}
qplot(item, pct, fill=condition,
      position="dodge",
      stat="identity", geom="bar",
      data=pcts)
```

and the z-scores:

```{r}
qplot(item, z, fill=condition,
      position="dodge",
      stat="identity", geom="bar",
      data=z.scores)
```

Oh well. This replication didn't seem to work out straightforwardly.

### 2015-01-25 analyses/ps1.Rmd
Completed assignment1

@@
 **There are several ways in which the data need to be cleaned. One tip-off about this is that a number of variables that should be numeric end up as factors... In particular, we note:**
+
 1. **Some participants appear multiple times (90 rows, but only 87 worker ids)**
+
 2. **There are commas in some of the input.price numbers**
+
 3. **someone entered their response for "dog cost" as a character string: "five hundred"**
@@
 **We can conclude that**
+
 1. **there are three clumps of data, corresponding to the three different objects (sushi, dog, plasma)**,
-2. **recalling that the anchors were set at 9, 2500, and 5000 (remember, we're not yet separating out classes of objects), we see that there seems to be a general trend of under-estimation**
+
+2. **recalling that the anchors were set at 9, 2500, and 5000 (remember, we're not yet separating out classes of objects), we see that there seems to be a general trend of under-estimation. **
+
+**However, because these three classes of objects are at different scales, it's hard to compare them all on the same axis...
+In the following plot, we separate out 'dog', 'plasma', and 'sushi', while still lumping together the 'over', 'under', and 'rounded' anchor conditions.
+This supports our observation that people tend to be under-estimating prices, overall.**
+
+```{r}
+g <- (ggplot(d.tidy, aes(x = cost))
+      + geom_histogram()
+      + facet_wrap(~ name, scales = 'free'))
+g
+```
 
@@
 
-This is a great time to get comfortable with the `%>%` operator. In brief, `%>%` allows you to pipe data from one function to another. So if you would have written:
-
-```{r}
-# d <- function(d, other_stuff)
-```
-
-you can now write:
-
-```{r}
-# d <- d %>% function(other_stufF)
-```
-
-That doesn't seem like much, but it's cool when you can replace:
-
-```{r}
-# d <- function1(d, other_stuff)
-# d <- function2(d, lots_of_other_stuff, more_stuff)
-# d <- function3(d, yet_more_stuff)
-```
-
-with
-
-```{r}
-# d <- d %>% 
-#   function1(other_stuff) %>%
-#   function2(lots_of_other_stuff, more_stuff) %>%
-#   function3(yet_more_stuff)
-```
-
+This is a great time to get comfortable with the `%>%` operator.
+In brief, `%>%` allows you to pipe data from one function to another.
 In other words, you get to make a clean list of the things you want to do and chain them together without a lot of intermediate assignments.
@@
   group_by(name, condition) %>% 
-  summarise(mean(cost, na.rm = TRUE))
+  summarise(mean_cost = mean(cost, na.rm = TRUE))
 ```
@@
 pcts <- d.tidy %>%
-  mutate(pct_change = anchor / cost) %>%
-  group_by(name, condition) %>% 
-  summarise(mean(cost, na.rm = TRUE))  
+  mutate(pct_change = abs((anchor - cost) / anchor)) %>%
+  group_by(item, condition) %>% 
+  summarise(pct = mean(pct_change, na.rm = TRUE))  
+pcts
 ```
@@
 ```{r}
-z.scores <- d.tidy %>% ...
+z.scores <- d.tidy %>% 
+  group_by(item) %>%
+  mutate(z_item = scale(cost)[,1]) %>%
+  group_by(condition, add = TRUE) %>%
+  summarise(z = mean(z_item, na.rm = TRUE))
+z.scores
 ```

### 2015-02-04 analyses/ps2.Rmd
Add ps2 (Sklar stuff isn't finished yet)
(This is not a merge with the instructor branch/repo - I did it manually)

---
title: 'Psych 254 W15 PS #2'
author: "Mike Frank"
date: "February 4, 2015"
output: html_document
---

This is problem set #2, in which we hope you will practice the visualization package ggplot2, as well as hone your knowledge of the packages tidyr and dplyr.

Part 1: Basic intro to ggplot
=============================

Part 1A: Exploring ggplot2 using qplot
--------------------------------------

Note, that this example is from the_grammar.R on http://had.co.nz/ggplot2 
I've adapted this for psych 254 purposes

First install and load the package.

```{r}
install.packages("ggplot2")
library(ggplot2)
```

Now we're going to use qplot. qplot is the easy interface, meant to replace plot. You can give it simple `qplot(x,y)` examples, or slightly more complex examples like `qplot(x, y, col=grp, data=d)`.

We're going to be using the diamonds dataset. This is a set of measurements of diamonds, along with their price etc.

```{r}
head(diamonds)
qplot(diamonds$carat, diamonds$price)
```

Scatter plots are trivial, and easy to add features to. Modify this plot so that it uses the dataframe rather than working from variables in the general namespace (good to get away from retyping `diamonds$` every time you reference a variable).

```{r}
```

Try adding clarity and cut, using shape and color as your visual variables.

```{r}
```

One of the primary benefits of `ggplot2` is the use of facets - also known as small multiples in the Tufte vocabulary. That last plot was probably hard to read. Facets could make it better. Try adding a `facets = x ~ y` argument. `x ~ y` means row facets are by x, column facets by y.

```{r}
```

But facets can also get overwhelming. Try to strike a good balance between color, shape, and faceting.

HINT: `facets = . ~ x` puts x on the columns, but `facets = ~ x` (no dot) *wraps* the facets. These are underlying calls to different functions, `facet_wrap` (no dot) and `facet_grid` (two arguments).

```{r}
```

The basic unit of a ggplot plot is a "geom" - a mapping between data (via an "aesthetic") and a particular geometric configuration on coordinate axes.

Let's try some other geoms and manipulate their parameters. First, try a histogram (`geom="hist"`).

```{r}
```

Now facet your histogram by clarity and cut.

```{r}
```

I like a slightly cleaner look to my plots. Luckily, ggplot allows you to add "themes" to your plots. Try doing the same plot but adding `+ theme_bw()` or `+ theme_classic()`. Different themes work better for different applications, in my experience.

```{r}
```

Part 1B: Exploring ggplot2 using ggplot
---------------------------------------

`ggplot` is just a way of building `qplot` calls up more systematically. It's
sometimes easier to use and sometimes a bit more complicated. What I want to show off here is the functionality of being able to build up complex plots with multiple elements. You can actually do this using qplot pretty easily, but there are a few things that are hard to do.

`ggplot` is the basic call, where you specify A) a dataframe and B) an aesthetic mapping from variables in the plot space to variables in the dataset.

```{r}
d <- ggplot(diamonds, aes(x=carat, y=price)) # first you set the aesthetic and dataset
d + geom_point() # then you add geoms
d + geom_point(aes(colour = carat)) # and you can keep doing this to add layers to the plot
```

Try writing this as a single set of additions (e.g. one line of R code, though you can put in linebreaks). This is the most common workflow for me.

```{r}
```

You can also set the aesthetic separately for each geom, and make some great plots this way. Though this can get complicated. Try using `ggplot` to build a histogram of prices.

```{r}
```

Part 2: Diving into real data: Sklar et al. (2012)
==================================================

Sklar et al. (2012) claims evidence for unconscious arithmetic processing. We're going to do a reanalysis of their Experiment 6, which is the primary piece of evidence for that claim. The data are generously contributed by Asael Sklar.

First let's set up a few preliminaries.

```{r}
library(tidyr)
library(dplyr)

sem <- function(x) {sd(x) / sqrt(length(x))}
ci95 <- function(x) {sem(x) * 1.96}
```

Data Prep
---------

First read in two data files and subject info. A and B refer to different trial order counterbalances.

```{r}
subinfo <- read.csv("../data/sklar_expt6_subinfo_corrected.csv")
d.a <- read.csv("../data/sklar_expt6a_corrected.csv")
d.b <- read.csv("../data/sklar_expt6b_corrected.csv")
```

Gather these datasets into long form and get rid of the Xs in the headers.

```{r}
```

Bind these together. Check out `bind_rows`.

```{r}
```

Merge these with subject info. You will need to look into merge and its relatives, `left_join` and `right_join`. Call this dataframe `d`, by convention.

```{r}
```

Clean up the factor structure.

```{r}
d$presentation.time <- factor(d$presentation.time)
levels(d$operand) <- c("addition","subtraction")
```

Data Analysis
-------------

### 2015-02-07 replication_project/mixed_model_analysis.Rmd
added couzin replication code

---
title: "couzin_lmer"
output: html_document
---

```{r}
library(lme4)
d = read.csv("~/Repos/couzin_replication/individual-data.csv")
```

```{r}
mod0 = lm(score ~ n_players, data = d)
summary(mod0)
mod0b = lm(score ~ n_players * difficulty, data = d)
summary(mod0b)
(ggplot(d, aes(x = n_players, y = score, col = difficulty, fill = difficulty))#, fill = difficulty))
 + geom_point()
 + stat_summary(fun.y=mean,geom="line", position= 'dodge')
 )
mod1 = lmer(score ~ n_players + (1 | game), data = d)
summary(mod1)
mod2 = lmer(score ~ n_players + (1 | game) + (1 + n_players | noise), data = d)
summary(mod2)
anova(mod1, mod2)
dev_diff <- -2*c(logLik(mod0)-logLik(mod1))
print(dev_diff)
pchisq(dev_diff,1,lower.tail=FALSE) 
```

```{r}
easy_d = subset(d, difficulty == "1en01")
easy_d$pid = factor(easy_d$pid)
easy_d$game = factor(easy_d$game)
easy_d$noise = factor(easy_d$noise)
str(easy_d)
mod0 = lm(score ~ n_players, data = easy_d)
summary(mod0)
mod1 = lmer(score ~ n_players + (1 | game), data = easy_d)
summary(mod1)
mod2 = lmer(score ~ n_players + (1 | game) + (1 + n_players | noise), data = easy_d)
summary(mod2)
anova(mod1, mod2)
dev_diff <- -2*c(logLik(mod0)-logLik(mod1))
print(dev_diff)
pchisq(dev_diff,1,lower.tail=FALSE)

qplot(x = n_players, y = score, col = noise, data = easy_d)
```

```{r}
med_d = subset(d, difficulty == "2en01")
med_d$pid = factor(med_d$pid)
med_d$game = factor(med_d$game)
med_d$noise = factor(med_d$noise)
mod0 = lm(score ~ n_players, data = med_d)
summary(mod0)
mod1 = lmer(score ~ n_players + (1 | game) + (1 | noise), data = med_d)
summary(mod1)
dev_diff <- -2*c(logLik(mod0)-logLik(mod1))
pchisq(dev_diff,1,lower.tail=FALSE) 
```

### 2015-02-07 replication_project/app.js

@@ -1,0 +1,108 @@
+/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 2013 Robert XD Hawkins
+
+    originally written for: http://buildnewgames.com/real-time-multiplayer/
+    
+    substantially modified for collective behavior experiments 
+
+    MIT Licensed.
+*/
+
+var 
+    use_db          = false,
+    gameport        = 8888,
+    app             = require('express')(),
+    server          = app.listen(gameport),
+    io              = require('socket.io')(server),
+    _               = require('underscore'),
+    fs              = require('fs');
+
+if (use_db) {
+    database        = require(__dirname + "/database"),
+    connection      = database.getConnection();
+}
+
+game_server = require('./game.server.js');
+utils = require('./utils.js');
+
+global_player_set = {}
+
+// Log something so we know that server-side setup succeeded
+console.log("info  - socket.io started");
+console.log('\t :: Express :: Listening on port ' + gameport );
+
+//  This handler will listen for requests on /*, any file from the
+//  root of our server. See expressjs documentation for more info 
+app.get( '/*' , function( req, res ) {
+    // this is the current file they have requested
+    var file = req.params[0]; 
+    console.log('\t :: Express :: file requested: ' + file);    
+    
+    if(req.query.id && !valid_id(req.query.id)) {
+	res.redirect('http://projects.csail.mit.edu/ci/turk/forms/invalid.html')
+    } else {
+	if(req.query.id && req.query.id in global_player_set) {
+	    res.redirect('http://projects.csail.mit.edu/ci/turk/forms/duplicate.html')
+	} else {
+	    // give them what they want
+	    res.sendfile("./" + file);
+	}
+    }
+}); 
+
+// Socket.io will call this function when a client connects. We check
+// to see if the client supplied a id. If so, we distinguish them by
+// that, otherwise we assign them one at random
+io.on('connection', function (client) {
+    // Recover query string information and set condition
+    var hs = client.handshake;    
+    var query = require('url').parse(client.handshake.headers.referer, true).query;
+    if( !(query.id && query.id in global_player_set) ) {
+	if(query.id) {
+	    global_player_set[query.id] = true
+	    var id = query.id; // use id from query string if exists
+	} else {
+	    var id = utils.UUID();
+	}
+	if(valid_id(id)) {
+	    console.log("user connecting...")
+	    initialize(query, client, id);
+	}
+    }
+});
+
+var valid_id = function(id) {
[…]

### 2015-02-07 replication_project/clean_and_run.sh

@@ -1,0 +1,3 @@
+rm data/*/*
+rm data/log
+node app.js >> data/log 2>&1

### 2015-02-07 replication_project/drawing.js

@@ -1,0 +1,73 @@
+// Draw players as triangles using HTML5 canvas
+draw_player = function(game, player){
+    // Draw avatar as triangle
+    var v = [[0,-player.size.x],
+	     [-player.size.hx,player.size.y],
+	     [player.size.hy,player.size.x]];
+    game.ctx.save();
+    game.ctx.translate(player.pos.x, player.pos.y);
+    game.ctx.rotate((player.angle * Math.PI) / 180);
+
+    // This draws the triangle
+    game.ctx.fillStyle = player.color;
+    game.ctx.strokeStyle = player.color;
+    game.ctx.beginPath();
+    game.ctx.moveTo(v[0][0],v[0][1]);
+    game.ctx.lineTo(v[1][0],v[1][1]);
+    game.ctx.lineTo(v[2][0],v[2][1]);
+    game.ctx.closePath();
+    game.ctx.stroke();
+    game.ctx.fill();
+
+    game.ctx.beginPath();
+    game.ctx.restore();
+
+    // Draw message in center (for countdown, e.g.)
+    game.ctx.font = "bold 12pt Helvetica";
+    game.ctx.fillStyle = 'red';
+    game.ctx.textAlign = 'center';
+    game.ctx.fillText(player.message, game.world.width/2, game.world.height/5);
+}; 
+
+draw_label = function(game, player, label) {
+    game.ctx.font = "8pt Helvetica";
+    game.ctx.fillStyle = 'white';
+    game.ctx.fillText(label, player.pos.x+10, player.pos.y + 20); 
+}
+
+draw_visibility_radius = function(game, player) {
+    var x = player.pos.x
+    var y = player.pos.y
+    var r = game.visibility_radius
+    game.ctx.beginPath();
+    game.ctx.strokeStyle = 'gray';
+    game.ctx.arc(x, y, r, 0, 2 * Math.PI, false);
+    game.ctx.stroke();
+}
+
+draw_other_dot = function(game, player, other) {
+    var scaling_factor = game.distance_between(player.pos, other.pos);
+    var X = (other.pos.x - player.pos.x);
+    var Y = (player.pos.y - other.pos.y);
+    var theta = Math.atan2(Y,X);
+    game.ctx.beginPath();
+    game.ctx.arc(player.pos.x + game.visibility_radius*Math.cos(theta),
+		 player.pos.y - game.visibility_radius*Math.sin(theta), 
+		 5, 0, 2 * Math.PI, false);
[…]

### 2015-02-07 replication_project/forms/validate-form.js

@@ -1,0 +1,27 @@
+function validateForm() {
+    
+    var completed = document.forms["quiz"]["completed"].value;
+    if (completed == "yes") {
+        location.replace("./notwice.html");
+    } else {
+	var count = 0
+	var qs = Array('dir', 'speed', 'none', 'loc', 'comp', 'diff', 'goal');
+	var as = Array(qs.length);
+	for (i = 0; i < qs.length; i++) { 
+	    as[i] = document.forms["quiz"][qs[i]].value;
+	    if(as[i] != "yes" && as[i] != "no") {
+		alert("Please enter responses to all questions to continue.");
+		return false
+	    }
+	    if(as[i] == "yes") {
+		count += 1
+	    }
+	}
+	if (count == qs.length) {
+	    location.replace("./pass.html");
+	} else {
+	    location.replace("./fail.html");
+	}
+    }
+    return false
+}

### 2015-02-07 replication_project/game.client.js

@@ -1,0 +1,437 @@
+/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 
+                  2013 Robert XD Hawkins
+    
+    written by : http://underscorediscovery.com
+    written for : http://buildnewgames.com/real-time-multiplayer/
+    
+    modified for collective behavior experiments on Amazon Mechanical Turk
+
+    MIT Licensed.
+*/
+
+/* 
+   THE FOLLOWING FUNCTIONS MAY NEED TO BE CHANGED
+*/
+
+// A window global for our game root variable.
+var game = {};
+// A window global for our id, which we can use to look ourselves up
+var my_id = null;
+// Keeps track of whether player is paying attention...
+var visible;
+var active_keys = []; 
+var speed_change = "none";
+var started = false;
+var ending = false;
+
+// what happens when you press 'left'?
+left_turn = function() {
+    var self = game.get_player(my_id);
+    self.angle = (Number(self.angle) - 5) % 360;
+};
+
+// what happens when you press 'left'?
+right_turn = function() {
+    var self = game.get_player(my_id);
+    self.angle = (Number(self.angle) + 5) % 360;
+};
+
+client_ondisconnect = function(data) {
+    // Redirect to exit survey
+    console.log("server booted")
+    if(game.get_player(my_id).kicked) {
+        var URL = 'http://projects.csail.mit.edu/ci/turk/forms/away.html';
+    } else if(game.get_player(my_id).inactive) {
+        var URL = 'http://projects.csail.mit.edu/ci/turk/forms/error.html?id=' + my_id;
+    } else if(game.get_player(my_id).lagging) {
+        var URL = 'http://projects.csail.mit.edu/ci/turk/forms/latency.html?id=' + my_id;
+    } else {
+	var URL = 'http://projects.csail.mit.edu/ci/turk/forms/end.html?id=' + my_id;
+    }
+    window.location.replace(URL);
+};
+
+
+/* 
+Note: If you add some new variable to your game that must be shared
+  across server and client, add it both here and the server_send_update
+  function in game.core.js to make sure it syncs 
+
+Explanation: This function is at the center of the problem of
+  networking -- everybody has different INSTANCES of the game. The
+  server has its own, and both players have theirs too. This can get
+  confusing because the server will update a variable, and the variable
+  of the same name won't change in the clients (because they have a
+  different instance of it). To make sure everybody's on the same page,
[…]

### 2015-02-07 replication_project/game.core.js

@@ -1,0 +1,598 @@
+/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 
+                  2013 Robert XD Hawkins
+    
+    written by : http://underscorediscovery.com
+    written for : http://buildnewgames.com/real-time-multiplayer/
+    
+    substantially modified for collective behavior experiments on the web
+
+    MIT Licensed.
+*/
+
+/*
+  The main game class. This gets created on both server and
+  client. Server creates one for each game that is hosted, and each
+  client creates one for itself to play the game. When you set a
+  variable, remember that it's only set in that instance.
+*/
+var has_require = typeof require !== 'undefined'
+
+if( typeof _ === 'undefined' ) {
+    if( has_require ) {
+        _ = require('underscore')
+    }
+    else throw new Error('mymodule requires underscore, see http://underscorejs.org');
+}
+
+var game_core = function(game_instance){
+
+    this.debug = false
+
+    // Define some variables specific to our game to avoid
+    // 'magic numbers' elsewhere
+    this.self_color = '#2288cc';
+    this.other_color = 'white';
+    
+    //Store the instance, if any (passed from game.server.js)
+    this.instance = game_instance;
+
+    //Store a flag if we are the server instance
+    this.server = this.instance !== undefined;
+
+    //Dimensions of world -- Used in collision detection, etc.
+    this.world = {width : 485, height : 280};  // 160cm * 3
+
+    //How often the players move forward <global_speed>px in ms.
+    this.tick_frequency = 125;     //Update 8 times per second
+    this.ticks_per_sec = 1000/125
+
+    if(this.debug) {
+	this.waiting_room_limit = 0.5 // set maximum waiting room time (in minutes)
+	this.round_length = 0.5; // set how long each round will last (in minutes)
+	this.max_bonus = 1.25*6/this.round_length; // total $ players can make in bonuses 
+	this.booting = false;
+    } else {
+	this.waiting_room_limit = 5 // set maximum waiting room time (in minutes)
+	this.round_length = 6 // set how long each round will last (in minutes)
+	this.max_bonus = 1.25; // total $ players can make in bonuses 
+	this.booting = true;
+    }
+
+    // game and waiting length in seconds
+    this.game_length = this.round_length*60*this.ticks_per_sec;
+
+    //The speed at which the clients move (e.g. # px/tick)
[…]

### 2015-02-07 replication_project/game.server.js

@@ -1,0 +1,287 @@
+/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 2013 Robert XD Hawkins
+    
+    written by : http://underscorediscovery.com
+    written for : http://buildnewgames.com/real-time-multiplayer/
+    
+    modified for collective behavior experiments on Amazon Mechanical Turk
+
+    MIT Licensed.
+*/
+
+//require('look').start()
+
+    var
+        use_db      = false,
+        game_server = module.exports = { games : {}, game_count:0, assignment:0},
+        fs          = require('fs');
+	    
+    if (use_db) {
+	    database    = require(__dirname + "/database"),
+	    connection  = database.getConnection();
+    }
+
+global.window = global.document = global;
+require('./game.core.js');
+utils = require('./utils.js');
+
+counter = 0
+
+// This is the function where the server parses and acts on messages
+// sent from 'clients' aka the browsers of people playing the
+// game. For example, if someone clicks on the map, they send a packet
+// to the server (check the client_on_click function in game.client.js)
+// with the coordinates of the click, which this function reads and
+// applies.
+game_server.server_onMessage = function(client,message) {
+    //Cut the message up into sub components
+    var message_parts = message.split('.');
+
+    //The first is always the type of message
+    var message_type = message_parts[0];
+    //console.log("received message: " + message)
+    //Extract important variables
+    var target = client.game.gamecore.get_player(client.userid);
+    var others = client.game.gamecore.get_others(client.userid);
+    if(message_type == 'a') {    // Client is changing angle
+        // Set their (server) angle 
+        target.angle = message_parts[1];
+	
+    } else if (message_type == 's') {
+        target.speed = message_parts[1].replace(/-/g,'.');;
+    } else if (message_type == "h") { // Receive message when browser focus shifts
+        target.visible = message_parts[1];
+    } else if (message_type == 'pong') {
+	var latency = (Date.now() - message_parts[1])/2;
+	target.latency = latency;
+	if(client.game.gamecore.game_started) {
+	    client.game.gamecore.latencyStream.write(String(client.userid)+","+message_parts[2]+","+latency+"\n",
+						     function(err) { if(err) throw err; });
+	} else {
+	    client.game.gamecore.waitingLatencyStream.write(String(client.userid)+","+message_parts[2]+","+latency+"\n",
+						     function(err) { if(err) throw err; });
+	}
+    }
+};
+
+/* 
[…]

### 2015-02-09 analyses/ps2.Rmd
Add rest of Sklar analysis to ps2

@@
 
-Data Analysis
--------------
+Data Analysis Preliminaries
+---------------------------
+
+Examine the basic properties of the dataset. First, take a histogram.
+
+```{r}
+```
+
+Challenge question: what is the sample rate of the input device they are using to gather RTs?
+
+```{r}
+```
+
+Sklar et al.
+did two manipulation checks.
+Subjective - asking participants whether they saw the primes - and objective - asking them to report the parity of the primes (even or odd) to find out if they could actually read the primes when they tried.
+Examine both the unconscious and conscious manipulation checks (this information is stored in subinfo).
+What do you see?
+Are they related to one another?
+
+```{r} 
+```
+
+OK, let's turn back to the measure and implement Sklar et al.'s exclusion criterion.
+You need to have said you couldn't see (subjective test) and also be not significantly above chance on the objective test (< .6 correct).
+Call your new data frame `ds`.
+
+```{r}
+```
+
+Sklar et al.'s analysis
+-----------------------
+
+Sklar et al.
+show a plot of a "facilitation effect" - the time to respond to incongruent primes minus the time to respond to congruent primes.
+They then show plot this difference score for the subtraction condition and for the two presentation times they tested.
+Try to reproduce this analysis.
+
+HINT: first take averages within subjects, then compute your error bars across participants, using the `sem` function (defined above).
+
+```{r}
+```
+
+Now plot this summary, giving more or less the bar plot that Sklar et al.
+gave (though I would keep operation as a variable here.
+Make sure you get some error bars on there (e.g.
+`geom_errorbar` or `geom_linerange`).
+
+```{r}
+```
+
+What do you see here? How close is it to what Sklar et al. report? Do the error bars match? How do you interpret these data?
+
+Challenge problem: verify Sklar et al.'s claim about the relationship between RT and the objective manipulation check.
+
+```{r}
+```
+
+Your own analysis
+-----------------
+
+Show us what you would do with these data, operating from first principles.
+What's the fairest plot showing a test of Sklar et al.'s original hypothesis that people can do arithmetic "non-consciously"?
+
+```{r}
+```
+
+Challenge problem: Do you find any statistical support for Sklar et al.'s findings?
+
+```{r}
+```

### 2015-02-09 replication_project/drawing.js
removed some redundant code: broken

@@ -1,3 +1,5 @@
+
+
 // Draw players as triangles using HTML5 canvas
 draw_player = function(game, player){
     // Draw avatar as triangle

### 2015-02-09 replication_project/game.client.js

@@ -228,42 +228,45 @@ client_update = function() {
 
 };
 
-var timeRemaining = function(remaining, limit) {
-    var time_remaining = limit - Math.floor(remaining / (1000*60));
-    if(time_remaining > 1) {
-	return {t: time_remaining, unit: 'minutes', actual: (limit - remaining/1000)}
-    } else {
-	time_remaining = limit - Math.floor(remaining / (1000 * 60)*6)/6
-	time_remaining = Math.max(Math.floor(time_remaining*6)*10, 10)
-	return {t: time_remaining, unit: 'seconds', actual: (limit - remaining/1000)}
-    }
+function mouseDownListener(evt) {
+    var i;
+    //We are going to pay attention to the layering order of the objects so that if a mouse down occurs over more than object,
+    //only the topmost one will be dragged.
+    var highestIndex = -1;
+        
+    //getting mouse position correctly, being mindful of resizing that may have occured in the browser:
+    var bRect = game.viewport.getBoundingClientRect();
+    mouseX = (evt.clientX - bRect.left)*(game.viewport.width/bRect.width);
+    mouseY = (evt.clientY - bRect.top)*(game.viewport.height/bRect.height);
 
-}
-
-var percentColors = [
-//    { pct: 0.0, color: { r: 0xff, g: 0x00, b: 0 } },
-    { pct: 0.0, color: { r: 0xff, g: 0xff, b: 0xff } },
-    { pct: 1.0, color: { r: 0x00, g: 0xff, b: 0 } } ];
- 
-var getColorForPercentage = function(pct) {
-    for (var i = 0; i < percentColors.length; i++) {
-        if (pct <= percentColors[i].pct) {
-            var lower = percentColors[i - 1] || { pct: 0.1, color: { r: 0x0, g: 0x00, b: 0 } };
-            var upper = percentColors[i];
-            var range = upper.pct - lower.pct;
-            var rangePct = (pct - lower.pct) / range;
-            var pctLower = 1 - rangePct;
-            var pctUpper = rangePct;
-            var color = {
-                r: Math.floor(lower.color.r * pctLower + upper.color.r * pctUpper),
-                g: Math.floor(lower.color.g * pctLower + upper.color.g * pctUpper),
-                b: Math.floor(lower.color.b * pctLower + upper.color.b * pctUpper)
-            };
-            return 'rgb(' + [color.r, color.g, color.b].join(',') + ')';
+    //find which shape was clicked
[…]

### 2015-02-09 replication_project/game.core.js

@@ -30,231 +30,55 @@ var game_core = function(game_instance){
 
     // Define some variables specific to our game to avoid
     // 'magic numbers' elsewhere
-    this.self_color = '#2288cc';
-    this.other_color = 'white';
-    
-    //Store the instance, if any (passed from game.server.js)
     this.instance = game_instance;
 
     //Store a flag if we are the server instance
     this.server = this.instance !== undefined;
 
     //Dimensions of world -- Used in collision detection, etc.
-    this.world = {width : 485, height : 280};  // 160cm * 3
-
-    //How often the players move forward <global_speed>px in ms.
-    this.tick_frequency = 125;     //Update 8 times per second
-    this.ticks_per_sec = 1000/125
-
-    if(this.debug) {
-	this.waiting_room_limit = 0.5 // set maximum waiting room time (in minutes)
-	this.round_length = 0.5; // set how long each round will last (in minutes)
-	this.max_bonus = 1.25*6/this.round_length; // total $ players can make in bonuses 
-	this.booting = false;
-    } else {
-	this.waiting_room_limit = 5 // set maximum waiting room time (in minutes)
-	this.round_length = 6 // set how long each round will last (in minutes)
-	this.max_bonus = 1.25; // total $ players can make in bonuses 
-	this.booting = true;
-    }
-
-    // game and waiting length in seconds
-    this.game_length = this.round_length*60*this.ticks_per_sec;
-
-    //The speed at which the clients move (e.g. # px/tick)
-    this.min_speed = 17 /this.ticks_per_sec; // 7.5cm * 3 * .5s 
-    this.max_speed = 57 /this.ticks_per_sec; // 7.5cm * 3 * .5s 
-
-    // minimun wage per tick
-    var us_min_wage_per_tick = 7.25 / (60*60*(1000 / this.tick_frequency))
-    this.waiting_background = us_min_wage_per_tick * this.game_length / this.max_bonus
-    this.waiting_start = new Date(); 
-
-    this.game_clock = 0;
-
-    //We create a player set, passing them the game that is running
-    //them, as well. Both the server and the clients need separate
-    //instances of both players, but the server has more information
-    //about who is who. Clients will be given this info later.
-    if(this.server) {
-        this.players = [{
-            id: this.instance.player_instances[0].id, 
-            player: new game_player(this,this.instance.player_instances[0].player)
-        }];
-    } else {
-        this.players = [{
-            id: null, 
-            player: new game_player(this)
[…]

### 2015-02-16 analyses/ps2.Rmd
Some progress on ps2
Finished ggplot assignment

@@
 title: 'Psych 254 W15 PS #2'
-author: "Mike Frank"
-date: "February 4, 2015"
+author: "Robert Hawkins"
+date: "February 16, 2015"
 output: html_document
@@
 ```{r}
-install.packages("ggplot2")
 library(ggplot2)
+setwd("~/Repos/psych254/analyses/")
 ```
@@
 ```{r}
+qplot(carat, price, data = diamonds)
 ```
@@
 ```{r}
+qplot(carat, price, col=clarity, shape=cut, data = diamonds)
 ```
@@
 ```{r}
+qplot(carat, price, facets = clarity ~ cut, data = diamonds)
 ```
@@
 ```{r}
+qplot(carat, price, col = clarity, facets = ~ cut, data = diamonds)
 ```
@@
 ```{r}
+qplot(price, geom="histogram", data = diamonds)
 ```
+
+**We see that most diamonds are between \$0 and \$1000 (with a mode around \$725), but that there is a very long tail, dragging the mean up to \$4000**
 
@@
 ```{r}
+qplot(price, geom="histogram", facets =clarity ~ cut , data = diamonds)
 ```
+
+**Unsurprisingly, the tail gets longer as the clarity gets bettter, although the VVS2 cut appears to have fairly long tails regardless of cut**
 
@@
 ```{r}
+qplot(price, geom="histogram", facets =clarity ~ cut , data = diamonds) + theme_bw()
 ```
@@
 ```{r}
+d <- (ggplot(diamonds, aes(x = carat, y=price))
+      + geom_point(aes(colour = carat)))
+d
 ```
@@
 ```{r}
+d <- (ggplot(diamonds, aes(x = price))
+      + geom_histogram())
+d
 ```
@@
 
-sem <- function(x) {sd(x) / sqrt(length(x))}
+sem <- function(x) {sd(x, na.rm = T) / sqrt(length(x))}
 ci95 <- function(x) {sem(x) * 1.96}
@@
 ```{r}
+d.a.long = gather(d.a, subid, RT, starts_with("X")) %>% 
+  mutate(subid = as.numeric(substr(subid, 2,3))) 
+d.b.long = gather(d.b, subid, RT, starts_with("X")) %>% 
+  mutate(subid = as.numeric(substr(subid, 2,3)))
 ```
@@
 ```{r}
+d.data <- bind_rows(d.a.long, d.b.long)
+cat(length(d.a.long$subid), length(d.b.long$subid), length(d.data$subid))
 ```
@@
 ```{r}
+d <- right_join(d.data, subinfo, by = c('subid'))
 ```
@@
 ```{r}
+library(ggplot2)
+qplot(RT, data = d)
 ```
+
+**We see that the data is a bit skewed to the left but otherwise passably normal, centered at about 600ms. We also note a few outliers toward the top and bottom.**
 
@@
 ```{r}
+qplot(RT, subid, geom = 'point', data = d)
 ```
+
+**First, we note that if we plot individual RTs for each participant, they seem to fall in bins rather than continuously**
+
+```{r}
+unique_RTs = sort(unique(subset(d, RT <= 1000 & RT >= 500)$RT))
+qplot(1:length(unique_RTs), unique_RTs)
+cat("sample rate is approximately", (length(unique_RTs)/3)*2, "samples per second")
+```
+
+**To make this observation precise, we look at the interval between 500ms and 1000ms where the most data was collected.
+We pool together all RTs from all participants in all conditions, record the unique values, and sort them from lowest to highest.
+This reveals a clear 'binning' pattern, with rounding error or tolerance leading to three values in each bin (1ms above and below the central value).
+There were 14 bins (with 3 unique values each) between 500ms and 100ms, which we extrapolate to a sampling rate of 28 samples per second.**
 
@@
 ```{r} 
+qplot(objective.test, subjective.test,  data = d)
+log_fit = glm(subjective.test ~ objective.test, family = binomial, data = d)
+summary(log_fit)
+xr = seq(.3, 1, .05)    # possible values of Responsible
+yc = predict(log_fit, data.frame(objective.test = xr), type = 'response')
+logit.data <- data.frame(xr,yc)
+ggplot(d, aes(x=objective.test, y=subjective.test)) +
+    geom_rect(xmin = -0.1, ymin = -0.1, xmax = .6, ymax = .1, fill = 'red') +
+  geom_point() +
+  geom_line(data = logit.data, aes(x = xr, y = yc)) 
 ```
+
+**We ran a logistic regression predicting the probability of subjectively seeing the primes from performance on an objective 2AFC task about their parity.
+We find that people who reported subjectively seeing the primes were also more likely to objectively do well on the 2AFC test about primes.
+$b = 9.36, t(6467) = 37.1, p < 0.001$ **
 
@@
 ```{r}
+ds <- d %>% 
+  filter(subjective.test == 0 & objective.test < .6)
+cat("excluded", length(unique(d$subid)) - length(unique(ds$subid)), "out of ",
+    length(unique(d$subid)))
 ```
@@
 show a plot of a "facilitation effect" - the time to respond to incongruent primes minus the time to respond to congruent primes.
-They then show plot this difference score for the subtraction condition and for the two presentation times they tested.
+They then plot this difference score for the subtraction condition and for the two presentation times they tested.
 Try to reproduce this analysis.
@@
 ```{r}
+RT_congruent <- ds %>% 
+  filter(congruent == "yes") %>% 
+  group_by(subid, operand, presentation.time) %>% 
+  summarise(congruent_mean = mean(RT, na.rm = T))
+RT_incongruent <- ds %>% 
+  filter(congruent == "no") %>%
+  group_by(subid, operand, presentation.time) %>% 
+  summarise(incongruent_mean = mean(RT, na.rm = T))
+RT_sum <- right_join(RT_congruent, RT_incongruent) %>%
+  mutate(diff_mean = incongruent_mean - congruent_mean) %>%
+  group_by(presentation.time, operand) %>%
+  mutate(error = sem(diff_mean)) %>%
+  select(subid, presentation.time, operand, error, diff_mean) %>%
+  summarize(error = mean(error), 
+            overall_mean = mean(diff_mean, na.rm = T))
+head(RT_sum)
 ```
@@
 ```{r}
+d <- (ggplot(RT_sum, aes(x = operand, y = overall_mean, 
+                         fill = presentation.time))
+      + geom_bar(stat = "identity", position = "dodge")
+      + geom_errorbar(aes(ymax = overall_mean + error, 
+                          ymin=overall_mean - error),
+                      position = position_dodge(width=0.9),
+                      width = 0.25)
+      + ylab("Facilitation (ms)")
+      + scale_fill_discrete(name="Presentation Duration"))
+d
 ```
@@
 
+**The "subtraction" part of our graph is numerically identical to the one presented by Sklar et al, and our interpretation of this result is the same: there is a greater facilitation effect for shorter display durations than longer display durations in the subtraction condition.
+We also note some differences in presentation:**
+
+1) **Note that our error bars appear much larger.
+Upon further inspection, this can be attributed to what they mean by "SEM." Typically, this means the "standard error of the mean" (i.e.
+$\bar{x} \pm 1 SE$, which is what we plotted), but they appear to have plotted the interval containing exactly one standard error.**
+
+2) **The tick mark labels on their $y$-axis show three (redundant) decimal places, making the numbers look bigger.**
+
+3) **They decided not to plot results for the 'addition' condition rather than including it in the plot.**
+
+**Overall, following remarks made by Andrew Gelman and others on experimenter degrees of freedom, we must express some skepticism about the interpretation offered by the authors.
+If an effect was found in the addition but not the subtraction condition, or in both conditions, would other justifications have been offered?
+If the effect were reversed such that longer presentation durations led to greater facilitation, would the authors be able to justify this as well?
+What if only one or the other of the manipulation checks were used, or the threshold in the objective test were moved down?
+Since the actual size of the effect is quite small (as seen from the error bars), further work seems to be needed in justifying these decisions.**
+
 Challenge problem: verify Sklar et al.'s claim about the relationship between RT and the objective manipulation check.
 
-```{r}
-```
+**Didn't get around to this**
 
@@
 
+**If arithmetic could be done 'non-consciously', we wouldn't necessarily expect presentation time or operand to matter.
+We would just want to know whether or not there was a significant facilitation effect across the board.**
+
 ```{r}
+RT_sum2 <- right_join(RT_congruent, RT_incongruent) %>%
+  mutate(diff_mean = incongruent_mean - congruent_mean) %>%
+  group_by(subid) %>%
+  summarize(overall_mean = mean(diff_mean, na.rm = T)) %>%
+  mutate(error = sem(overall_mean))
+
+pop_mean = mean(RT_sum2$overall_mean)
+error = mean(RT_sum2$error)
+d <- (ggplot(RT_sum2, aes(x = overall_mean))
+      + geom_rect(aes(xmin=pop_mean - 2*error, 
+                xmax=pop_mean + 2*error, 
+                ymin=0, ymax=Inf), 
+            fill = 'red', alpha = 0.1)
+      + geom_histogram()
+      + geom_vline(xintercept = pop_mean)
+)
+d
 ```
+
+**We plot a histogram of each participant's average facilitation effect across conditions, and show the approximate 95% confidence interval ($\bar{x} \pm 2SE$) for the facilitation effect estimate.
+It is not clear from the graph whether it is significant -- the confidence interval includes 0 -- but it is certainly trending in the direction of the claim for non-conscious arithmetic**
 
@@
 
+**We simply run a t-test to check whether the distribution of individual facilitation effects is different from 0**
+
 ```{r}
+t.test(RT_sum2$overall_mean)
 ```
+
+**We find that the facilitation effect is marginally greater than 0 ($m = 5.37, t(16) = 1.84, p = 0.084$), but not significant.
+Sklar et al.'s finding are based on various ways of slicing these data into subsets which are significantly different from other subsets, but the fact that there is the overall facilitation effect is small (5ms) and only marginally significant shows that whatever non-conscious processes contribute to the way we do arithmetic are sensitive to a variety of other factors (including at least presentation time and operand).
+This same conclusion is given by the authors at the end of their results: "What the specific conditions are under which addition and subtraction equations are nonconsciously solved turns out to be a complex issue that must be left for future investigations."**

### 2015-02-23 analyses/ps3.Rmd

---
title: 'Psych 254 W15 PS #3'
author: "Mike Frank"
date: "February 22, 2015"
output: html_document
---

This is problem set #3, in which we want you to integrate your knowledge of data wrangling with some basic simulation skills and some linear modeling.

```{r}
library(dplyr)
library(ggplot2)
```

Part 1: Basic simulation and NHST
=================================

Let's start by convincing ourselves that t-tests have the appropriate false positive rate. Run 10,000 t-tests with standard, normally-distributed data from a made up 30-person, single-measurement experiment (the command for sampling from a normal distribution is `rnorm`). What's the mean number of "significant" results?

First do this using a `for` loop.

```{r}
```

Next, do this using the `replicate` function:

```{r}
```

Ok, that was a bit boring. Let's try something moderately more interesting - let's implement a p-value sniffing simulation, in the style of Simons, Nelson, & Simonsohn (2011).

Consider this scenario: you have done an experiment, again with 30 participants (one observation each, just for simplicity). The question is whether their performance is above chance. You aren't going to check the p-value every trial, but let's say you run 30 - then if the p-value is within the range p < .25 and p > .05, you optionally run 30 more and add those data, then test again. But if the original p value is < .05, you call it a day, and if the original is > .25, you also stop.

First, write a function that implements this sampling regime.

```{r}
double.sample <- function () {
}
```

Now call this function 10k times and find out what happens.

```{r}
```

Is there an inflation of false positives? How bad is it?

Now modify this code so that you can investigate this "double the sample" rule in a bit more depth. Let's see what happens when you double the sample ANY time p > .05 (not just when p < .25), or when you do it only if p < .5 or < .75. How do these choices affect the false positive rate?

HINT: Try to do this by making the function `double.sample` take the upper p value as an argument, so that you can pass this through dplyr.

HINT 2: You may need more samples. Find out by looking at how the results change from run to run.

```{r}
```

What do you conclude on the basis of this simulation? How bad is this kind of data-dependent policy?

Part 2: The Linear Model
========================

2A: Basic Linear Modeling
-------------------------

Let's use the `ToothGrowth` dataset, on guineapig teeth based on orange juice
and vitamin C. This is super simple. (Try `?ToothGrowth`).

First plot the data, we'll use `qplot` to understand how `len` (tooth length) depends on `dose` (amount of Vitamin C) and `supp` (delivery method).

```{r}
```

So now you see what's going on.

Next, always make a histogram of the DV before making a linear model! This reveals the distribution and can be helpful in choosing your model type.

```{r}
```

Now make a linear model of tooth lengths using `lm`. Try making one with main effects and interactions and another with just main  effects. Make sure to assign them to variables so that you can get them later.

```{r}
```

Now try taking out the intercept, using a -1 term in the formula. what does this do?

```{r}
```

Thought question: Take a moment to interpret the coefficients of the model. 
Q1 - What are the units?
Q2 - How does the interaction relate to the plot?
Q3 - Should there be an interaction in the model? What does it mean? How important is it?

Now make predictions from the model you like the best. What should happen with
doses of 0, 1.5, 2.5, and 10 under both supplements?

HINT: use the `predict` function ...

HINT 2: you will have to make a dataframe to do the prediction with, so use something like `data.frame(dose=...)`.

```{r}
```

Now plot the residuals from the original model. How do they look?
HINT: `?resid`

```{r}
```

BONUS: test them for normality of distribution using a quantile-quantile plot.

HINT: `?qqplot` and `?qqnorm`

```{r}
```

2B: Exploratory Linear Modeling
-------------------------------

What the heck is going on? Load data from Frank, Vul, Saxe (2011, Infancy), a study in which we measured infants' looking to hands in moving scenes. There were infants from 3 months all the way to about two years, and there were two movie conditions (`Faces_Medium`, in which kids played on a white background, and `Faces_Plus`, in which the backgrounds were more complex and the people in the videos were both kids and adults). Forgive our bad naming conventions.

Try to figure out what the most reasonable linear model of the data is.

```{r}
d <- read.csv("../data/FVS2011-hands.csv")

```

Plot that model on the same plot as the data.

HINT: you can do this either using `predict` or (if you are feeling confident of your understanding of the models) using the built-in linear models in `ggplot`'s `geom_smooth`.

```{r}
```

What do you conclude from this pattern of data?

### 2015-03-02 analyses/ps3.Rmd
Add rest of PS3

@@
 What do you conclude from this pattern of data?
+
+3: Linear Mixed Effect Models
+=============================
+
+The goal here is to learn to use LMEMs using `lme4` and to compare them to
+standard by subject, by item LMs, as well as the standard (no repeated measures) fixed effects GLM.
+
+The dataset here is from Stiller, Goodman, & Frank (2014), a paper on children's pragmatic inferences.
+We saw the paradigm in the counterbalancing lecture: it's three faces: a smiley, a smiley with glasses, and a smiley with a hat and glasses.
+When told "my friend has glasses" do kids pick the one with the glasses and no hat?
+`age.group` is the kids' age group, `condition` is either "label," described above, or "no label," which was a control condition in which kids picked without hearing the term "glasses" at all.
+
+```{r}
+d <- read.csv("../data/scales.csv")
+d$age.group <- factor(d$age.group)
+```
+
+Always begin with a histogram!
+
+```{r}
+```
+
+Brief Confidence Interval Digression
+------------------------------------
+
+Start out by setting up a function for a 95% CI using the normal approximation.
+
+```{r}
+ci95.norm <- function(x) {
+}
+```
+
+But the number of participants in a group is likely to be < 30, so let's also compute this with a t distribution.
+
+```{r}
+ci95.t <- function(x) {
+}
+```
+
+On the other hand, maybe we should use bootstrap CIs because these are actually  proportions, and the normal/t approximations don't know that they are 0/1 bounded.
+
+```{r}
+library(boot)
+library(bootstrap)
+```
+
+Take a look at `?boot` and `?bootci`. Note that the syntax for the `boot` library is terrible, so we're going to use it to check some code that I use:
+
+```{r}
+theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
+ci.low <- function(x,na.rm=T) {
+  mean(x,na.rm=na.rm) - 
+    quantile(bootstrap(1:length(x),
+                       10000,theta,x,na.rm=na.rm)$thetastar,.025,na.rm=na.rm)}
+ci.high <- function(x,na.rm=T) {
+  quantile(bootstrap(1:length(x),
+                     10000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - 
+    mean(x,na.rm=na.rm)}
+```
+
+Now make 95% confidence intervals across participants using all the methods above:
+
+- Normal
+- t
+- Bootstrap percentile using `boot.ci`
+- Bootstrap percentile using my code
+
+```{r}
+```
+
+Now plot the data by age and condition using `dplyr` and `ggplot2`.
+Plot some CIs on here - extra credit if you plot all of them and compare visually (you'll need `position = position_dodge()` or some other way to offset them).
+
+```{r}
+```
+
+What do you conclude about confidence interval computation?
+
+Back to LMEMs
+-------------
+
+```{r}
+library(lme4)
+```
+
+OK, now do a basic GLM over the entire data frame, using `age.group`, `condition`, and their interaction to predict correctness.
+(If we were focusing on developmental issues, I would ask you to think about how to model age here, but let's treat it as three discrete groups for now).
+
+NOTE: this model is not appropriate, because it assumes that each subject's observations are independent from one another.
+It's still fine to do the analysis, though: it can tell you a lot about the data and is easy and fast to fit, as long as you know that you can't trust the p-values!
+
+```{r}
+```
+
+Let's now use `dplyr` to get data frames for by-items (`msi`) and by-subjects (`mss`) analyses.
+`msi` should contain the mean ratings for every item and `mss` should contain the mean ratings for every subject.
+
+```{r}
+```
+
+Now do standard linear models on each of these.
+
+NOTE: These are not strictly correct either because of the normal approximation on percent correct (model doesn't know it's 0 - 1 bounded and could give you standard error that goes above 1).
+Again, useful to do and see what happens.
+
+```{r}
+```
+
+Do ANOVA on these. Note that ANOVA doesn't let you figure out what is going on with individual levels of age.
+
+```{r}
+```
+
+On to linear mixed effect models.
+Create the maximal random effects model a la Barr et al.
+(2013).
+Does it converge?
+If not, what will you do to make it converge?
+(The internet can be your friend here).
+
+HINT: try simplifying your model to a "semi-maximal" model. Bonus: try using a different fitting procedure on the maximal model.
+
+HINT: make sure that you consider which random effects are appropriate.
+Consider which observations are within/between subjects.
+E.g.
+having a random coefficient for age by subject doesn't make sense, because each subject has only one age.
+
+```{r}
+```
+
+How do these coefficients compare with the independent coefficients linear model? What do you conclude?
+
+Which random effects make the most difference? Find out using `ranef`. Plot the random effects for subject and item.
+
+```{r}
+```
+
+Make the minimal random effects model with just a subject intecept. How does this compare?
+
+```{r}
+```
+
+Get an estimate of the significance value for the coefficient on the `age*condition` interaction by using anova to compare between your semi-maximal model and the model without an intercept.
+
+```{r}
+```

### 2015-03-05 replication_project/drawing.js
Add scripted instructions

@@ -31,13 +31,13 @@ var drawGrid = function(game){
     // occluded cells...
     for(var i = 0; i < occludedList.length; i++) {
       var cell = occludedList[i]
-      console.log(cell)
+//      console.log(cell)
       drawOccludedCell(cell[0],cell[1])
     }
 }
 
 var drawOccludedCell = function(x, y) {
-  console.log("occluding")
+//  console.log("occluding")
     var cell = game.getPixelFromCell(y,x)
     var topLeft = [cell.centerX - cell.width/2, cell.centerY - cell.height/2]
     var topRight = [cell.centerX + cell.width/2, topLeft[1]]
@@ -79,6 +79,10 @@ var drawInstructions = function(game) {
     var dest = game.getPixelFromCell(game.currentDestination[0], game.currentDestination[1])
     drawArrow(game, origin.centerX, origin.centerY, 
               dest.centerX, dest.centerY, 50)
+    if(game.scriptedInstruction != "none")
+      $('#chatbox').val(game.scriptedInstruction);
+    else 
+      $('#chatbox').val("")
 }
 
 var drawScreen = function(game, player) {

### 2015-03-05 replication_project/game.client.js

@@ -25,7 +25,7 @@ var dragging;
 client_ondisconnect = function(data) {
     // Redirect to exit survey
     console.log("server booted")
-    var URL = 'http://projects.csail.mit.edu/ci/turk/forms/end.html?id=' + my_id;
+    var URL = 'http://web.stanford.edu/~rxdh/psych254/replication_project/forms/end.html?id=' + my_id;
     window.location.replace(URL);
 };
 
@@ -59,10 +59,13 @@ client_onserverupdate_received = function(data){
         function(e){ return e.name})
     var localNames = _.map(game.objects,
         function(e){return e.name})
-    if (game.objects.length == 0 || !_.isEqual(dataNames, localNames)) { //(game.objects != data.objects && my_role == 'director') || 
+
+    // If your objects are out-of-date (i.e. if there's a new round), update them
+    if (game.objects.length == 0 || !_.isEqual(dataNames, localNames)) { 
         game.objects = _.map(data.objects, function(obj) {
             var imgObj = new Image()
             imgObj.src = obj.url
+            // Set it up to load properly
             imgObj.onload = function(){
                 game.ctx.drawImage(imgObj, parseInt(obj.trueX), parseInt(obj.trueY), obj.width, obj.height)
                if(my_role == "director") 
@@ -72,25 +75,26 @@ client_onserverupdate_received = function(data){
                 {img: imgObj, trueX : obj.trueX, trueY : obj.trueY})
         })
     }
-//    } else {
-        _.map(game.objects, function(obj) {
-            data_obj = _.find(data.objects, function(o) {return o.name == obj.name})
-            obj.trueX = data_obj.trueX;
-            obj.trueY = data_obj.trueY;
-        })
-    // }
+
+    // Update local object positions
+    _.map(game.objects, function(obj) {
+        data_obj = _.find(data.objects, function(o) {return o.name == obj.name})
+        obj.trueX = data_obj.trueX;
+        obj.trueY = data_obj.trueY;
+    })
 
     game.currentDestination = data.curr_dest;
+    game.scriptedInstruction = data.scriptedInstruction;
+
+    console.log(game.scriptedInstruction)
+
     game.instructions = data.instructions
     game.instructionNum = data.instructionNum;
     game.game_started = data.gs;
     game.players_threshold = data.pt;
     game.player_count = data.pc;
-    console.log("official objects data")
-    console.log(data.objects)
-    console.log("updated objects info to...")
-    console.log(game.objects)
 
+    // Draw all this new stuff
[…]

### 2015-03-05 replication_project/game.core.js

@@ -115,10 +115,14 @@ game_core.prototype.newRound = function() {
 game_core.prototype.newInstruction = function() {
     this.instructionNum += 1;
     var instruction = this.instructions[this.instructionNum]
-    console.log(instruction)
     var item = instruction.split(' ')[0]
     var dir = instruction.split(' ')[1]
     var object = _.find(this.objects, function(obj) { return obj.name == item })
+    this.scriptedInstruction = (object.hasOwnProperty('scriptedInstruction') ?
+        object.scriptedInstruction :
+        "none")
+    console.log(this.scriptedInstruction)
+
     var dest;
     switch(dir) {
         case "down" :
@@ -218,7 +222,6 @@ game_core.prototype.makeTrialList = function () {
             obj.trueY = gridCell.centerY - obj.height/2
         })
     })
-    console.log(trialList[1].objects)
     return trialList
 }
 
@@ -251,13 +254,17 @@ game_core.prototype.server_send_update = function(){
             player: null}
         })
 
+    console.log("sending:", this.scriptedInstruction)
+
     var state = {
             gs : this.game_started,                      // true when game's started
             pt : this.players_threshold,
             pc : this.player_count,
             curr_dest : this.currentDestination,
+            scriptedInstruction : this.scriptedInstruction,
             instructionNum : this.instructionNum,
         };
+
     _.extend(state, {players: player_packet})
     _.extend(state, {instructions: this.instructions})
     if(player_packet.length == 2) {

### 2015-03-05 replication_project/game.server.js

@@ -33,7 +33,7 @@ counter = 0
 // with the coordinates of the click, which this function reads and
 // applies.
 game_server.server_onMessage = function(client,message) {
-    console.log("received message: " + message)
+//    console.log("received message: " + message)
     //Cut the message up into sub components
     var message_parts = message.split('.');
 
@@ -47,7 +47,7 @@ game_server.server_onMessage = function(client,message) {
     switch(message_type) {
         case 'objMove' :    // Client is changing angle
             var obj = client.game.gamecore.objects[message_parts[1]]
-            console.log("moving " + obj.name)
+//            console.log("moving " + obj.name)
             obj.trueX = parseInt(message_parts[2])
             obj.trueY = parseInt(message_parts[3])
             _.map(others, function(p) {
@@ -56,7 +56,7 @@ game_server.server_onMessage = function(client,message) {
             break;
         case 'correctDrop' :
             var obj = client.game.gamecore.objects[message_parts[1]]
-            console.log("dropping " + obj.name)
+//            console.log("dropping " + obj.name)
             obj.trueX = parseInt(message_parts[2])
             obj.trueY = parseInt(message_parts[3])
             _.map(others, function(p) {
@@ -130,7 +130,7 @@ game_server.findGame = function(player) {
                 game.gamecore.mouseDataStream = fs.createWriteStream(mouse_f, {'flags' : 'a'});
                 fs.writeFile(message_f, "time, sender, contents\n", function (err) {if(err) throw err;})
                 game.gamecore.messageStream = fs.createWriteStream(message_f, {'flags' : 'a'});
-                console.log('game ' + game.id + ' starting with ' + game.player_count + ' players...')
+//                console.log('game ' + game.id + ' starting with ' + game.player_count + ' players...')
     
                 // Attach game to player so server can look at it later
                 player.game = game;

### 2015-03-05 replication_project/stimuli/objectSet.js

@@ -6,7 +6,7 @@
 var sunGlasses = {
 	url: 'stimuli/sunglasses.png', name: "sunGlasses", width: 130, height: 65,
 	instruction: "sunGlasses right", initialLoc : [3,2],
-	scriptedInstruction: "the glasses"}
+	scriptedInstruction: "move the glasses to the right"}
 var glassesCase = {
 	url: 'stimuli/glassesCase.png', name: "glassesCase", width : 100, height: 107,
 	instruction: "", initialLoc: [2,2]}
@@ -18,7 +18,8 @@ var saxophone = {
 	instruction: "saxophone left", initialLoc: [4,2]}
 var airplane = {
 	url: 'stimuli/airplane.jpg', name: 'airplane', width: 130, height: 60,
-	instruction: "airplane right", initialLoc : [1,2]}
+	instruction: "airplane right", initialLoc : [1,2],
+	scriptedInstruction: "move the airplane to the right"}
 var barrel = {
 	url: 'stimuli/barrel.png', name: 'barrel', width: 87, height: 130,
 	instruction: "barrel down", initialLoc : [2,4]}
@@ -30,7 +31,7 @@ var watch = {
 var middleBlock = {
 	url: 'stimuli/middleBlock.png', name: "middleBlock", width: 80, height: 80,
 	instruction: "middleBlock left", initialLoc: [3,2],
-	scriptedInstruction: "the bottom block", }
+	scriptedInstruction: "move the bottom block to the left", }
 var bottomBlock = {
 	url: 'stimuli/bottomBlock.jpg', name: "bottomBlock", width: 80, height: 80,
 	instruction: "", initialLoc: [4,4]}
@@ -42,19 +43,21 @@ var topBlock = {
 	instruction: "", initialLoc: [1,1]}
 var binoculars = {
 	url: 'stimuli/binoculars.jpg', name: 'binoculars', width: 100, height: 100,
-	instruction: "binoculars right", initialLoc: [3,3]}
+	instruction: "binoculars right", initialLoc: [3,3],
+	scriptedInstruction: "move the binoculars to the right"}
 var wrench = {
 	url: 'stimuli/wrench.png', name: 'wrench', width: 130, height: 33,
 	instruction: "wrench left", initialLoc: [1,3]}
 var coffeeMug = {
 	url: 'stimuli/coffeeMug.png', name: 'coffeeMug', width: 130, height: 122,
[…]

### 2015-03-05 replication_project/drawing.js
Flash title when started

@@ -31,13 +31,11 @@ var drawGrid = function(game){
     // occluded cells...
     for(var i = 0; i < occludedList.length; i++) {
       var cell = occludedList[i]
-//      console.log(cell)
       drawOccludedCell(cell[0],cell[1])
     }
 }
 
 var drawOccludedCell = function(x, y) {
-//  console.log("occluding")
     var cell = game.getPixelFromCell(y,x)
     var topLeft = [cell.centerX - cell.width/2, cell.centerY - cell.height/2]
     var topRight = [cell.centerX + cell.width/2, topLeft[1]]

### 2015-03-05 replication_project/game.client.js

@@ -86,8 +86,6 @@ client_onserverupdate_received = function(data){
     game.currentDestination = data.curr_dest;
     game.scriptedInstruction = data.scriptedInstruction;
 
-    console.log(game.scriptedInstruction)
-
     game.instructions = data.instructions
     game.instructionNum = data.instructionNum;
     game.game_started = data.gs;
@@ -129,6 +127,9 @@ client_onMessage = function(data) {
             client_onjoingame(num_players, commands[3]); break;
         case 'add_player' : // New player joined... Need to add them to our list.
             console.log("adding player" + commanddata)
+            if(hidden === 'hidden') {
+                flashTitle("GO!")
+            }
             game.players.push({id: commanddata, player: new game_player(game)}); break;
         case 'begin_game' :
             client_newgame(); break;
@@ -381,3 +382,27 @@ function onchange (evt) {
     game.socket.send("h." + document.body.className);
 };
 
+(function () {
+
+    var original = document.title;
+    var timeout;
+
+    window.flashTitle = function (newMsg, howManyTimes) {
+        function step() {
+            document.title = (document.title == original) ? newMsg : original;
+            if (visible === "hidden") {
+                timeout = setTimeout(step, 500);
+            } else {
+                document.title = original;
+            }
+        };
+        cancelFlashTitle(timeout);
+        step();
+    };
+
+    window.cancelFlashTitle = function (timeout) {
+        clearTimeout(timeout);
+        document.title = original;
+    };
+
+}());

### 2015-03-05 replication_project/game.core.js

@@ -254,8 +254,6 @@ game_core.prototype.server_send_update = function(){
             player: null}
         })
 
-    console.log("sending:", this.scriptedInstruction)
-
     var state = {
             gs : this.game_started,                      // true when game's started
             pt : this.players_threshold,

### 2015-03-07 replication_project/game.client.js
Closes #5 -- writing full info

@@ -250,12 +250,14 @@ client_onjoingame = function(num_players, role) {
         $('#instructs').append("Click and drag objects to follow the director's instructions.")
     }
 
+    // set role locally
     my_role = role;
     game.get_player(my_id).role = my_role;
 
     if(num_players == 1)
         game.get_player(my_id).message = 'Waiting for other player to connect...';
 
+    // set mouse-tracking event handler
     if(role === "matcher") {
         $('#viewport').mousemove(function(event){
             var bRect = game.viewport.getBoundingClientRect();
@@ -330,6 +332,7 @@ function mouseUpListener(evt) {
     game.viewport.addEventListener("mousedown", mouseDownListener, false);
     window.removeEventListener("mouseup", mouseUpListener, false);
     if (dragging) {
+        // Set up the right variables
         var bRect = game.viewport.getBoundingClientRect();
         dropX = (evt.clientX - bRect.left)*(game.viewport.width/bRect.width);
         dropY = (evt.clientY - bRect.top)*(game.viewport.height/bRect.height);
@@ -337,6 +340,7 @@ function mouseUpListener(evt) {
         var cell = game.getCellFromPixel(dropX, dropY)
         console.log(cell)
         console.log([obj.gridX, obj.gridY])
+        
         // If you were dragging the correct object... And dragged it to the correct location...
         if (_.isEqual(obj.name, game.instructions[game.instructionNum].split(' ')[0])
             && _.isEqual(cell, game.currentDestination)) {
@@ -346,12 +350,14 @@ function mouseUpListener(evt) {
             obj.trueX = game.getPixelFromCell(cell[0], cell[1]).centerX - obj.width/2
             obj.trueY = game.getPixelFromCell(cell[0], cell[1]).centerY - obj.height/2
             game.socket.send("correctDrop." + dragIndex + "." + Math.round(obj.trueX) + "." + Math.round(obj.trueY))
+        
         // If you didn't drag it beyond cell bounds, snap it back w/o comment
         } else if (obj.gridX == cell[0] && obj.gridY == cell[1]) {
             console.log("here!")
             obj.trueX = game.getPixelFromCell(obj.gridX, obj.gridY).centerX - obj.width/2
             obj.trueY = game.getPixelFromCell(obj.gridX, obj.gridY).centerY - obj.height/2
             game.socket.send("objMove." + dragIndex + "." + Math.round(obj.trueX) + "." + Math.round(obj.trueY))
+        
[…]

### 2015-03-07 replication_project/game.server.js

@@ -50,6 +50,7 @@ game_server.server_onMessage = function(client,message) {
 
     //Extract important variables
     var gc = client.game.gamecore
+    var id = gc.instance.id.slice(0,6)
     var all = gc.get_active_players();
     var target = gc.get_player(client.userid);
     var others = gc.get_others(client.userid);
@@ -65,11 +66,18 @@ game_server.server_onMessage = function(client,message) {
             break;
 
         case 'incorrectDrop' :
+            var condition = gc.trialList[gc.roundNum].condition
             moveObject(client, message_parts[1], message_parts[2], message_parts[3])
+            var trueItem = gc.instructions[gc.instructionNum].split(' ')[0]
+            var line = (id + ',' + String(message_parts[6]) + ',' + condition + ',' 
+                        + trueItem + ',' 
+                        + gc.objects[message_parts[1]].name + ','
+                        + parseInt(gc.currentDestination[0]) + ',' 
+                        + parseInt(gc.currentDestination[1]) + ','
+                        + parseInt(message_parts[4]) + ',' + parseInt(message_parts[5]) + '\n')
             // write error to file
-            gc.errorStream.write(message_parts[6] + 
-                + ',' + gc.objects[message_parts[1]].name + "," 
-                + parseInt(message_parts[4]) + ',' + parseInt(message_parts[5]) + '\n')
+            console.log("incorrect: ", line);
+            gc.errorStream.write(line)
             gc.paused = true;
             _.map(all, function(p) {p.player.instance.send("s.waiting.incorrect") })
             break;
@@ -88,26 +96,37 @@ game_server.server_onMessage = function(client,message) {
             var date = message_parts[1]
             var msg = message_parts[2].replace(/-/g,'.')
             if(client.game.player_count == 2)
-                gc.messageStream.write(date + ',' + client.role + ',"' + msg + '"\n')
+                gc.messageStream.write(id + ',' + date + ',' + client.role + ',"' + msg + '"\n')
             _.map(all, function(p){
                 p.player.instance.emit( 'chatMessage', {user: client.userid, msg: msg})})
             break;
 
         case 'update_mouse' :
-            var date = message_parts[1]
-            var x = message_parts[2]
-            var y = message_parts[3]
-            var roundNum = gc.roundNum
-            var instructionNum = gc.instructionNum
             if(!gc.paused) {
-                console.log(gc.currentDestination)
-                var destinationPixel = gc.getPixelFromCell(gc.currentDestination[0],
[…]

### 2015-03-07 replication_project/stimuli/objectSet.js

@@ -5,185 +5,183 @@
 // BLOCK 1
 var sunGlasses = {
 	url: 'stimuli/sunGlasses.png', name: "sunGlasses", width: 130, height: 65,
-	instruction: "sunGlasses right", initialLoc : [3,2],
+	instruction: "sunGlasses right", initialLoc : [3,2], critical : "target",
 	scriptedInstruction: "move the glasses to the right"}
 var glassesCase = {
 	url: 'stimuli/glassesCase.png', name: "glassesCase", width : 100, height: 107,
-	instruction: "", initialLoc: [2,2]}
+	instruction: "", initialLoc: [2,2], critical : "distractor",}
 var soccerBall = {
 	url: 'stimuli/soccerBall.jpg', name: "soccerBall", width: 100, height: 100,
-	instruction: "", initialLoc: glassesCase.initialLoc}
+	instruction: "", initialLoc: glassesCase.initialLoc, critical : "distractor",}
 var saxophone = {
 	url: 'stimuli/saxophone.png', name: "saxophone", width: 97, height: 130,
-	instruction: "saxophone left", initialLoc: [4,2]}
+	instruction: "saxophone left", initialLoc: [4,2], critical : "filler",}
 var airplane = {
 	url: 'stimuli/airplane.jpg', name: 'airplane', width: 130, height: 60,
-	instruction: "airplane right", initialLoc : [1,2],
+	instruction: "airplane right", initialLoc : [1,2],critical : "filler",
 	scriptedInstruction: "move the airplane to the right"}
 var barrel = {
 	url: 'stimuli/barrel.png', name: 'barrel', width: 87, height: 130,
-	instruction: "barrel down", initialLoc : [2,4]}
+	instruction: "barrel down", initialLoc : [2,4], critical : "filler",}
 var watch = {
 	url: 'stimuli/watch.png', name: 'watch', width: 130, height: 84,
-	instruction: "", initialLoc: [3,1]}
+	instruction: "", initialLoc: [3,1], critical : "filler",}
 
 // BLOCK 2
 var middleBlock = {
 	url: 'stimuli/middleBlock.png', name: "middleBlock", width: 80, height: 80,
 	instruction: "middleBlock left", initialLoc: [3,2],
-	scriptedInstruction: "move the bottom block to the left", }
+	scriptedInstruction: "move the bottom block to the left", critical : "target"}
 var bottomBlock = {
[…]

### 2015-03-08 replication_project/game.server.js
Write correct distractor location

@@ -115,8 +115,8 @@ game_server.server_onMessage = function(client,message) {
                 } else {
                     var critical = 1;
                     var distractor = _.find(gc.objects, function(obj) { return obj.critical == "distractor" })
-                    var distractorX = distractor.trueX
-                    var distractorY = distractor.trueY
+                    var distractorX = distractor.trueX + distractor.width/2
+                    var distractorY = distractor.trueY + distractor.height/2
                 }
 
                 var objX = object.trueX + object.width/2

### 2015-03-08 replication_project/game.server.js
Write more data (refactor)

@@ -66,18 +66,8 @@ game_server.server_onMessage = function(client,message) {
             break;
 
         case 'incorrectDrop' :
-            var condition = gc.trialList[gc.roundNum].condition
+            writeData(client, "error", message_parts)
             moveObject(client, message_parts[1], message_parts[2], message_parts[3])
-            var trueItem = gc.instructions[gc.instructionNum].split(' ')[0]
-            var line = (id + ',' + String(message_parts[6]) + ',' + condition + ',' 
-                        + trueItem + ',' 
-                        + gc.objects[message_parts[1]].name + ','
-                        + parseInt(gc.currentDestination[0]) + ',' 
-                        + parseInt(gc.currentDestination[1]) + ','
-                        + parseInt(message_parts[4]) + ',' + parseInt(message_parts[5]) + '\n')
-            // write error to file
-            console.log("incorrect: ", line);
-            gc.errorStream.write(line)
             gc.paused = true;
             _.map(all, function(p) {p.player.instance.send("s.waiting.incorrect") })
             break;
@@ -93,42 +83,19 @@ game_server.server_onMessage = function(client,message) {
             break;
 
         case 'chatMessage' :
-            var date = message_parts[1]
+            //write data to file
+            if(client.game.player_count == 2) 
+                writeData(client, "message", message_parts)
+            // Update others
             var msg = message_parts[2].replace(/-/g,'.')
-            if(client.game.player_count == 2)
-                gc.messageStream.write(id + ',' + date + ',' + client.role + ',"' + msg + '"\n')
             _.map(all, function(p){
                 p.player.instance.emit( 'chatMessage', {user: client.userid, msg: msg})})
             break;
 
         case 'update_mouse' :
-            if(!gc.paused) {
-                var date = message_parts[1]
-                var x = message_parts[2]
-                var y = message_parts[3]
-                var object_name = gc.instructions[gc.instructionNum].split(' ')[0]
-                var object = _.find(gc.objects, function(obj) { return obj.name == object_name })
-                if(object.critical === "filler") {
-                    var critical = 0;
-                    var distractorX = "none"
-                    var distractorY = "none"
-                } else {
-                    var critical = 1;
-                    var distractor = _.find(gc.objects, function(obj) { return obj.critical == "distractor" })
-                    var distractorX = distractor.trueX + distractor.width/2
[…]

### 2015-03-09 replication_project/stimuli/objectSet.js
Moved critical items to outside rim

@@ -5,17 +5,17 @@
 // BLOCK 1
 var sunGlasses = {
 	url: 'stimuli/sunGlasses.png', name: "sunGlasses", width: 130, height: 65,
-	instruction: "sunGlasses right", initialLoc : [3,2], critical : "target",
+	instruction: "sunGlasses right", initialLoc : [4,3], critical : "target",
 	scriptedInstruction: "move the glasses to the right"}
 var glassesCase = {
 	url: 'stimuli/glassesCase.png', name: "glassesCase", width : 100, height: 107,
-	instruction: "", initialLoc: [2,2], critical : "distractor",}
+	instruction: "", initialLoc: [3,1], critical : "distractor",}
 var soccerBall = {
 	url: 'stimuli/soccerBall.jpg', name: "soccerBall", width: 100, height: 100,
 	instruction: "", initialLoc: glassesCase.initialLoc, critical : "distractor",}
 var saxophone = {
 	url: 'stimuli/saxophone.png', name: "saxophone", width: 97, height: 130,
-	instruction: "saxophone left", initialLoc: [4,2], critical : "filler",}
+	instruction: "saxophone up", initialLoc: [4,2], critical : "filler",}
 var airplane = {
 	url: 'stimuli/airplane.jpg', name: 'airplane', width: 130, height: 60,
 	instruction: "airplane right", initialLoc : [1,2],critical : "filler",}
@@ -25,12 +25,12 @@ var barrel = {
 	scriptedInstruction: "move the barrel down"}
 var watch = {
 	url: 'stimuli/watch.png', name: 'watch', width: 130, height: 84,
-	instruction: "", initialLoc: [3,1], critical : "filler",}
+	instruction: "", initialLoc: [2,2], critical : "filler",}
 
 // BLOCK 2
 var middleBlock = {
 	url: 'stimuli/middleBlock.png', name: "middleBlock", width: 80, height: 80,
-	instruction: "middleBlock left", initialLoc: [3,2],
+	instruction: "middleBlock left", initialLoc: [3,4],
 	scriptedInstruction: "move the bottom block to the left", critical : "target"}
 var bottomBlock = {
 	url: 'stimuli/bottomBlock.jpg', name: "bottomBlock", width: 80, height: 80,
@@ -43,7 +43,7 @@ var topBlock = {
 	instruction: "", initialLoc: [1,1], critical : "filler",}
 var binoculars = {
[…]

### 2015-03-09 replication_project/game.core.js
Fixed little bug disconnecting people

@@ -104,7 +104,7 @@ game_core.prototype.newRound = function() {
     if(this.roundNum == this.numRounds - 1) {
         var local_game = this;
         _.map(local_game.get_active_players(), function(p){
-            p.player.instance.send('s.end')})
+            p.player.instance.disconnect()})//send('s.end')})
     } else {
         this.roundNum += 1;
         this.objects = this.trialList[this.roundNum].objects

### 2015-03-11 replication_project/game.core.js
Final changes before full exp run

@@ -40,6 +40,7 @@ var game_core = function(game_instance){
     this.roundNum = -1;
     this.instructionNum = -1;
     this.numRounds = 8;
+    this.attemptNum = 0; // Increments whenever someone makes a mistake
     this.paused = true;
     this.objects = [];
     this.instructions = []

### 2015-03-11 replication_project/game.server.js

@@ -61,6 +61,7 @@ game_server.server_onMessage = function(client,message) {
 
         case 'correctDrop' :
             moveObject(client, message_parts[1], message_parts[2], message_parts[3])
+            gc.attemptNum = 0;
             _.map(all, function(p) {p.player.instance.send("s.waiting.correct")})
             gc.paused = true;
             break;
@@ -69,6 +70,7 @@ game_server.server_onMessage = function(client,message) {
             writeData(client, "error", message_parts)
             moveObject(client, message_parts[1], message_parts[2], message_parts[3])
             gc.paused = true;
+            gc.attemptNum += 1;
             _.map(all, function(p) {p.player.instance.send("s.waiting.incorrect") })
             break;
 
@@ -106,8 +108,10 @@ game_server.server_onMessage = function(client,message) {
 
 var writeData = function(client, type, message_parts) {
     var gc = client.game.gamecore
+    var attemptNum = gc.attemptNum;
     var condition = gc.trialList[gc.roundNum].condition
     var objectSet = gc.trialList[gc.roundNum].objectSet
+    var instructionNum = gc.instructionNum
     var object_name = gc.instructions[gc.instructionNum].split(' ')[0]
     var object = _.find(gc.objects, function(obj) { return obj.name == object_name })
     var critical = object.critical === "filler" ? 0 : 1
@@ -130,7 +134,7 @@ var writeData = function(client, type, message_parts) {
             var objY = object.trueY + object.height/2
 
             var line = String(id + ',' + date + ',' + condition + ',' + critical + ',' + 
-                objectSet + ',' + objX + ',' + objY + ',' +
+                objectSet + ',' + instructionNum + ',' + attemptNum + ',' + objX + ',' + objY + ',' +
                 distractorX + ',' + distractorY + ',' + x + ',' + y ) + "\n"
             console.log("mouse:" + line)
             gc.mouseDataStream.write(line, function (err) {if(err) throw err;}); 
@@ -139,14 +143,15 @@ var writeData = function(client, type, message_parts) {
             var date = message_parts[1]
             var msg = message_parts[2].replace(/-/g,'.')
             var line = (id + ',' + date + ',' + condition + ',' + critical + ',' +
-                objectSet + ',' + client.role + ',"' + msg + '"\n')
+                objectSet + ',' + instructionNum + ',' + attemptNum + ',' + client.role + ',"' + msg + '"\n')
             console.log("message:" + line)
[…]

### 2015-03-19 replication_project/stimuli/objectSet.js
Cleaned up directory

@@ -142,7 +142,7 @@ var pencilEraser = {
 	instruction: "", initialLoc: [3,1],critical : "distractor",}
 var brain = {
 	url: 'stimuli/brain.png', name: "brain", width: 130, height:104,
-	instruction: "", initialLoc: [2,2],critical : "distractor",}
+	instruction: "", initialLoc: [3,1],critical : "distractor",}
 var dollar = {
 	url: 'stimuli/dollar.jpg', name: 'dollar', width: 130, height: 54,
 	instruction: "dollar right", initialLoc: [3,3],critical : "filler",}

### 2015-03-21 analyses/ps3.Rmd
Started on ps3
Worked on ps3

@@
 ```{r}
+sigCount = 0
+numSamples = 10000
+for(i in 1:numSamples) {
+  fakeSample = rnorm(30)
+  stat = t.test(fakeSample)
+  if(stat$p.value < 0.05) {
+    sigCount = sigCount + 1;
+  }
+}
+print(sigCount / numSamples)
 ```
@@
 ```{r}
+numSamples = 10000
+testRes <- replicate(numSamples, t.test(rnorm(30))$p.value < .05)
+cat("false positive rate:", sum(testRes)/numSamples)
 ```
+
+**Using both methods, we find a false positive rate of about 0.05!
+In other words, if the null hypothesis is true and there is no effect, we'll still report one about 5% of the time.
+This is our alpha value (in decision-theoretic terms).
+As the number of samples increases, we'll converge to this rate exactly.
+**
 
@@
 ```{r}
-double.sample <- function () {
+double.sample <- function (maxP) {
+  fakeSample1 = rnorm(30)
+  test1p = t.test(fakeSample1)$p.value
+  if(test1p >= maxP || test1p < .05) {
+    return (test1p < 0.05)
+  } else {
+    fakeSample2 = rnorm(30)
+    test2p = t.test(c(fakeSample1, fakeSample2))$p.value
+    return (test2p < 0.05)
+  }
 }
@@
 ```{r}
+numSamples = 10000
+sum(replicate(numSamples, double.sample(.25)))/numSamples
 ```
@@
 Is there an inflation of false positives? How bad is it?
+
+**There is indeed an inflation of false positives: we find false positives about 7% of the times instead of 5% as before**
 
@@
 ```{r}
+numSamples = 10000
+d <- data.frame(caps = c(.05, .25, .5, .75, 1)) %>%
+  group_by(caps) %>%
+  mutate(falseAlarm = sum(replicate(numSamples, double.sample(caps)))/numSamples)
+d
 ```
@@
 What do you conclude on the basis of this simulation? How bad is this kind of data-dependent policy?
+
+**Inflation continues to grow up to about 9% in the worse-case (when you always run 30 more trials after observing $p>0.05$).
+This means you will be about twice as likely to incorrectly report a significant finding when no effect exists.
+This is pretty bad.
+**
 
@@
 ```{r}
+data(ToothGrowth)
+qplot(y = len, x = dose, color = supp, data = ToothGrowth)
 ```
@@
 ```{r}
+qplot(len, data = ToothGrowth)
 ```
+
+** We see that length has a rather wide distribution, possibly multimodal. It is likely a mixture of different distributions, which we can investigate in our linear model.**
 
@@
 ```{r}
+mainEffectModel <- lm(len ~ dose + supp, data = ToothGrowth)
+summary(mainEffectModel)
+interactionModel <- lm(len ~ dose * supp, data = ToothGrowth)
+summary(interactionModel)
+anova(mainEffectModel, interactionModel, test="Chisq")
 ```
+
+**We see that both dose and supplement are significant in the main effects model, and that the interaction between them is significant in the interactive model.**
 
@@
 ```{r}
+mainEffectModel <- lm(len ~ -1 + dose + supp, data = ToothGrowth)
+summary(mainEffectModel)
+interactionModel <- lm(len ~ -1 + dose * supp, data = ToothGrowth)
+summary(interactionModel)
 ```
+
+**This did not affect the interaction coefficient, but it added a level of the supplement factor**
 
@@
 Q1 - What are the units?
+**The coefficient in front of dose is the average slope of the line if the two supplement groups were collapsed together.
+The coefficients in front of the dose groups are the numbers added to this baseline dose at each level to get their respective group means.**
 Q2 - How does the interaction relate to the plot?
+**In the plot, we saw that the slope across doses is steeper for VC supplement than OJ supplement. This is the interaction our model is picking up on.** 
 Q3 - Should there be an interaction in the model? What does it mean? How important is it?
-
+**Yes, a nested model comparison shows that the interactive model accounts for a significantly higher proportion of the variance than the additive model, $\chi^2(1) =  88.92, p = 0.02$.**
 Now make predictions from the model you like the best. What should happen with
@@
 ```{r}
+newD = data.frame(dose=rep(c(0,1.5,2.5,10),2), supp = c(rep('OJ', 4), rep('VC', 4)))
+newD
+newD$predictions = predict(interactionModel, newD)
+newD
+ggplot(newD, aes(x = dose, y = predictions, color = supp)) + geom_line()
 ```
@@
 ```{r}
+plot(resid(interactionModel))
 ```
+
+**Not bad! No clear trends, but also quite a bit of variability**
 

### 2015-03-29 analysis/keysarMouseTracking.Rmd
Removed non-keysar materials

---
title: "KeysarMouseTracking.csv"
output: html_document
---

```{r}
setwd('~/Repos/psych254/replication_project/analysis/')
mouseD = read.csv('fullSampleMouse.csv')
errorD = read.csv('fullSampleError.csv')
messageD = read.csv('fullSampleMessage.csv')
str(mouseD)
library(tidyr)
library(dplyr)
library(ggplot2)
```

Set up error data

```{r}
nonCriticalMistakes <- errorD %>%
  # only look at mistakes on noncritical (filler) items
  filter(critical != 1) %>%
  # don't want to double-count people for messing the same thing up multiple times
  filter(attemptNum == 0) %>% 
  group_by(gameid) %>%
  tally() %>%
  # implement exclusion criteria of errors on >~10% of non-critical trials
  filter(n >= 2)

bannedIDs <- nonCriticalMistakes$gameid
```

Make error table

```{r}
eD <- errorD %>%
  filter(!(gameid %in% bannedIDs)) %>%
  filter(critical == 1) %>%
  filter(attemptNum == 0) %>%
  group_by(gameid, condition) %>%
  tally() %>%
  mutate(cond2 = as.numeric(n >= 2)) %>%
  group_by(condition) %>%
  summarise(atLeastOnce = length(n)/14, 
            atLeastTwice = sum(cond2)/14)
eD

# To illustrate how much a few items are driving the effect, we exclude three items 
# where more than 60% of participants made errors.
excludedED <- errorD %>%
  filter(!(gameid %in% bannedIDs)) %>%
  filter(critical == 1) %>%
  filter(!(objectSet %in% c(2, 6, 7))) %>%
  filter(attemptNum == 0) %>%
  group_by(gameid, condition) %>%
  tally() %>%
  mutate(cond2 = as.numeric(n >= 2)) %>%
  group_by(condition) %>%
  summarise(atLeastOnce = length(n)/14, 
            atLeastTwice = sum(cond2)/14)
excludedED
```

When we look at errors per item, we have to remember that not all items had the same number of participants in the experimental condition.

```{r}
numPlayersPerCondition = mouseD %>% 
  filter(!(gameid %in% bannedIDs)) %>% 
  filter(critical == 1) %>% 
  filter(attemptNum == 0) %>% 
  group_by(gameid, condition, objectSet) %>% 
  summarize() %>% 
  group_by(condition, objectSet) %>% 
  summarize(total = n()) %>% 
  filter(condition == "exp")
```

Follow-up: are all critical items equal?

```{r}
itemInequalities <- errorD %>%
  filter(!(gameid %in% bannedIDs)) %>%
  filter(critical == 1) %>%
  filter(attemptNum == 0) %>%
  filter(condition == "exp") %>%
  group_by(objectSet) %>%
  tally() %>%
  right_join(numPlayersPerCondition, by = c('objectSet')) %>%
  transmute(errorCount = n, correctCount = total - n)

t(itemInequalities)
t <- chisq.test(itemInequalities)
print(t)
t$expected
t$observed
```

Set up mouse/message data

```{r}

mouData <- mouseD %>%
  filter(!(gameid %in% bannedIDs)) %>%
  filter(attemptNum == 0)

mD <- messageD %>% 
  filter(!(gameid %in% bannedIDs)) %>%
  filter(attemptNum == 0) %>%
  filter(sender == "director") %>%
  group_by(gameid, objectSet, instructionNum) %>%
  mutate(messageStamp = first(time)) %>%
  select(gameid, condition, attemptNum, instructionNum, critical, objectSet,sender,contents,messageStamp) %>%
  distinct(gameid,contents)

joined <- (right_join(mouData, mD, by = c("gameid", "condition", "objectSet", 
                                         "instructionNum", "attemptNum", "critical")))
```

```{r}
dUnfilt <- joined %>%
  filter(critical == 1) %>%
  group_by(objectSet)

d <- joined %>% 
  filter(critical == 1) %>% 
  group_by(objectSet, condition) %>%
#  filter(time > messageStamp) #%>%
  mutate(mouseY = 600 - mouseY) %>%
  mutate(targetY = 600 - targetY) %>%
  mutate(begTargetX = first(targetX),
         begTargetY = first(targetY),
         distractorX = as.numeric(levels(distractorX))[distractorX],
         distractorY = 600 - as.numeric(levels(distractorY))[distractorY]) %>%
   mutate(begDistrX = first(distractorX),
          begDistrY = first(distractorY)) %>%
   filter(targetX == begTargetX) %>%
   filter(targetY == begTargetY) %>%
   filter(distractorX == begDistrX) %>%
   filter(distractorY == begDistrY)

```

Compute hover time statistics

```{r}

sem <- function(x) {sd(x, na.rm = T) / sqrt(length(x))}
ci95 <- function(x) {sem(x) * 1.96}

heatMapForLM <- d %>%
  group_by(gameid, objectSet) %>%
  mutate(inTargetSquare = as.numeric(mouseX > targetX - 75 
                                     & mouseX < targetX + 75
                                     & mouseY > targetY - 75
                                    & mouseY < targetY + 75)) %>%
  mutate(inDistractorSquare = as.numeric(mouseX > distractorX - 75
                                         & mouseX < distractorX + 75
                                         & mouseY > distractorY - 75
                                         & mouseY < distractorY + 75)) %>%
  group_by(gameid, condition) %>%
  summarise(distractor = sum(inDistractorSquare) / n(),
         target = sum(inTargetSquare) / n()) %>%
  gather(cellType, percent, distractor, target)

heatMapForLM

t1 <- lm(percent ~ condition * cellType, data = heatMapForLM)
summary(t1)
```

Plot it?

```{r}
heatMapForPlot <- d %>%
  group_by(gameid, objectSet) %>%
  mutate(inTargetSquare = as.numeric(mouseX > targetX - 75 
                                     & mouseX < targetX + 75
                                     & mouseY > targetY - 75
                                    & mouseY < targetY + 75)) %>%
  mutate(inDistractorSquare = as.numeric(mouseX > distractorX - 75
                                         & mouseX < distractorX + 75
                                         & mouseY > distractorY - 75
                                         & mouseY < distractorY + 75)) %>%
  group_by(gameid, condition) %>%
  summarize(distractor = sum(inDistractorSquare) / n(),
         target = sum(inTargetSquare) / n()) %>%
  gather(cellType, percent, distractor, target) %>%
  group_by(condition, cellType) %>%
  summarize(error = sem(percent),
            percent = mean(percent))

heatMapForPlot

g <- (ggplot(heatMapForPlot, aes(x = condition, y = percent, group = cellType,
                                  color = cellType)) 
      + geom_line(aes(linetype = cellType), size = 2)
      + geom_errorbar(aes(ymax = percent + error, 
                          ymin = percent - error), size =2 ,
                      width = 0.1))
g
```

Very noisy (unfinished) plot of all mouse-movements

```{r}
# heatMapForPlot$new_labels = as.factor(sapply(X = heatMapForPlot$objectSet, 
#                               FUN = function(v) {return(paste("Item", v))}))
# 
# g <- (ggplot(heatMapForPlot, aes(x = mouseX, y = mouseY, color = condition)) +
#       geom_vline(xintercept = c(0, 150,300,450, 600)) +
#       geom_hline(yintercept = c(0, 150,300,450, 600)) +
#       geom_point() +
#       geom_point(aes(x = begTargetX, y = begTargetY), size = 20,
#                  shape = 4, color = "black", show_guide = FALSE) +
#       geom_point(aes(x = begDistrX, y = begDistrY), size = 20,
#                  shape = 4, color = "grey50", show_guide = FALSE) + 
#       theme(panel.grid.major = element_blank(), 
#             panel.grid.minor = element_blank()) +
#       xlim(0, 600) +
#       ylim(0, 600) +
#       theme(aspect.ratio = 1) +
#       facet_wrap(~ new_labels, nrow = 2))
# g
```

### 2015-03-29 drawing.js

@@ -1,0 +1,226 @@
+var occludedList = [[4,4], [2,2], [2,3], [3,1], [1,4]];
+
+var drawGrid = function(game){
+    //size of canvas
+    var cw = game.viewport.width;
+    var ch = game.viewport.height;
+
+    //padding around grid
+    var p = 25;
+
+    //grid width and height
+    var bw = cw - (p*2) ;
+    var bh = ch - (p*2) ;
+
+    game.ctx.beginPath();
+
+    // vertical lines
+    for (var x = 0; x <= bw; x += Math.floor((cw - 2*p) / 4)) {
+        game.ctx.moveTo(0.5 + x + p, p);
+        game.ctx.lineTo(0.5 + x + p, bh + p);}
+
+    // horizontal lines
+    for (var x = 0; x <= bh; x += Math.floor((ch - 2*p) / 4)) {
+        game.ctx.moveTo(p, 0.5 + x + p);
+        game.ctx.lineTo(bw + p, 0.5 + x + p);}
+
+    game.ctx.lineWidth = 1;
+    game.ctx.strokeStyle = "black";
+    game.ctx.stroke();
+
+    // occluded cells...
+    for(var i = 0; i < occludedList.length; i++) {
+      var cell = occludedList[i]
+      drawOccludedCell(cell[0],cell[1])
+    }
+}
+
+var drawClickPoint = function(game) {
+  var centerX = game.viewport.width / 2;
+  var centerY = game.viewport.height / 2;
+  var radius = 8;
+
+  game.ctx.beginPath();
+  game.ctx.arc(centerX, centerY, radius, 0, 2 * Math.PI, false);
+  game.ctx.fillStyle = 'green';
+  game.ctx.fill();
+  game.ctx.lineWidth = 3;
+  game.ctx.strokeStyle = '#003300';
+  game.ctx.stroke();
+}
+
+var drawOccludedCell = function(x, y) {
+    var cell = game.getPixelFromCell(y,x)
+    var topLeft = [cell.centerX - cell.width/2, cell.centerY - cell.height/2]
+    var topRight = [cell.centerX + cell.width/2, topLeft[1]]
+    var bottomLeft = [topLeft[0], cell.centerY + cell.height/2]
+    var bottomRight = [topRight[0], bottomLeft[1]]
+    game.ctx.beginPath();
+    game.ctx.moveTo(topLeft[0], topLeft[1])
[…]

### 2015-03-29 examples/sims/one_vs_two_expts.R

@@ -1,40 +1,0 @@
-rm(list=ls())
-library(dplyr)
-library(ggplot2)
-
-## ----------------- ONE EXPT vs. TWO EXPTS ----------------- 
-# simulations from class 2/2 on whether you should run one experiment or two 
-# given a fixed sample size
-
-sim <- function (samp, es) {
-  full.samp <- rnorm(samp, mean = es)
-  full.p <- t.test(full.samp)$p.val < .05
-  
-  half.samp1 <- full.samp[1:floor(samp/2)]
-  half.samp2 <- full.samp[floor(samp/2)+1:samp]
-  halves.p <- t.test(half.samp1)$p.val < .05 |
-    t.test(half.samp2)$p.val < .05
-  
-  return(list(full=full.p, halves=halves.p))
-}
-
-ns <- seq(20,200,20)
-es <- seq(0,.5,.1)
-  
-d <- data.frame()
-for (n in ns) {
-  for (e in es) { 
-    rs <- replicate(1000, sim(n, e))
-    ms <- rowMeans(rs==TRUE)
-    d <- bind_rows(d, data.frame(power = ms,
-                                 regime = c("full","halves"), 
-                                 n=n, 
-                                 effect = e))
-                   
-  }
-}
-
-qplot(n, power, facets=~effect,
-      col=regime, geom="line", group=regime,
-      data=d)
-

### 2015-03-29 examples/sims/pval_sniffing.R

@@ -1,83 +1,0 @@
-rm(list = ls())
-
-#### pvalue sniffing ####
-n.samps <- 1000
-n.obs <- 50
-
-## first no sniffing
-no.sniff <- array(dim=n.samps)
-for (i in 1:n.samps) {
-	x <- rnorm(n.obs)
-	no.sniff[i] <- t.test(x)$p.val < .05
-}
-mean(no.sniff)
-
-## next with sniffing every subject, starting at 20
-sniff <- array(dim=n.samps)
-for (i in 1:n.samps) {
-	x <- rnorm(n.obs)
-	
-	j <- 20
-	while (j < n.obs) {
-		samp <- x[1:j]
-		sniff[i] <- t.test(samp)$p.val < .05
-		if (sniff[i]) {
-			break
-		} else {	
-			j <- j + 1
-		}
-	}
-}
-mean(sniff)
-
-## next with sniffing every subject, starting at 20
-intervals <- c(1,5,10,20)
-sniff.int <- matrix(nrow=n.samps,ncol=length(intervals))
-
-for (k in 1:length(intervals)) {
-	for (i in 1:n.samps) {
-		x <- rnorm(n.obs)
-		
-		j <- 20
-		while (j <= n.obs) {
-			samp <- x[1:j]
-			sniff.int[i,k] <- t.test(samp)$p.val < .05
-			if (sniff.int[i,k]) {
-				break
-			} else {	
-				j <- j + intervals[k]
-			}
-		}
-	}
-}
-
-
-colMeans(sniff.int)
-plot(intervals,colMeans(sniff.int),type="b",bty="n",
-	ylim=c(0,.2),xlab="interval between sniffs",xlim=c(0,20),
-	ylab="false positive rate")
-lines(c(0,30),c(.05,.05),lty=2,col=2)
-text(intervals,colMeans(sniff.int)+.02,colMeans(sniff.int),xpd="n")
-
-
-## try sniffing with a doubling rule 
-## (interesting - has a .08 from .05 inflation every sample) 
-n.obs <- 192
-sniff <- array(dim=n.samps)
-n.success <- array(dim=n.samps)
-for (i in 1:n.samps) {
-  x <- rnorm(n.obs)
-  
-  j <- 96
-  while (j <= n.obs) {
-    samp <- x[1:j]
-    sniff[i] <- t.test(samp)$p.val < .05
-    if (sniff[i]) {
-      n.success[i] <- j
-      break
[…]

### 2015-03-29 examples/sims/simple_t_power.R

@@ -1,67 +1,0 @@
-## Simple power analysis simulations
-
-### ----------------- ONE EFFECT SIZE ----------------- 
-true.mean <- 1
-true.sd <- 3
-ns <- seq(5,200,5)
-num.samps <- 1000
-
-ps <- matrix(nrow=num.samps,ncol=length(ns))
-for (i in 1:length(ns)) {
-	for (j in 1:num.samps) {
-		d <- rnorm(ns[i],true.mean,true.sd)
-		ps[j,i] <- t.test(d)$p.val < .05
-	}	
-}
-
-power <- colMeans(ps)
-
-plot(ns,power,xlab="number of observations",ylim=c(0,1),
-	ylab="probability of rejecting H0",bty="n")
-	
-
-
-### ----------------- VARY EFFECT SIZE ----------------- 
-true.mean <- 1
-ns <- seq(10,120,10)
-num.samps <- 100
-es <- seq(0.1,1,.1) # effect size
-
-ps <- list()
-power <- matrix(nrow=length(es),ncol=length(ns))
-for (e in 1:length(es)) {
-	true.sd <- true.mean / es[e]  # effect size is SD scaled by mean
-	
-	ps[[e]] <- matrix(nrow=num.samps,ncol=length(ns))
-	
-	for (i in 1:length(ns)) {
-		for (j in 1:num.samps) {
-			d <- rnorm(ns[i],true.mean,true.sd)
-			ps[[e]][j,i] <- t.test(d)$p.val < .05
-		}	
-		power[e,i] <- mean(ps[[e]][,i])
-	}
-	
-}
-
-plot(NA,xlab="number of observations",
-	ylab="probability of rejecting H0",bty="n",type="l",ylim=c(0,1),xlim=c(0,120))
-
-for (e in 1:length(es)) {
-	lines(ns,power[e,],col=e)	
-	text(ends[i] + (ends[i]/15),power[e,ns==ends[i]],es[e],col=e,xpd="n")
-}	
-
-
-### pretty layout
-layout(matrix(c(1,2,3,4), 2, 2, byrow = TRUE))
-ends <- c(30,60,90,120)
-for (i in 1:4) {
-	plot(NA,xlab="number of observations",
-		ylab="probability of rejecting H0",bty="n",type="l",ylim=c(0,1),xlim=c(10,ends[i]))
-
-	for (e in 1:length(es)) {
[…]

### 2015-03-29 examples/viz/level_ordering.R

@@ -1,35 +1,0 @@
-## from http://stackoverflow.com/questions/12064007/factor-order-within-faceted-dotplot-using-ggplot2
-rm(list=ls())
-library(ggplot2)
-library(dplyr)
-d <- read.csv("data/medals.csv")
-
-## no ordering
-quartz()
-qplot(x = count, y = country, data = d, 
-      geom = "point", 
-      facets = medal.type ~.)
-
-## ordering across facets
-quartz()
-ms <- d %>% group_by(country) %>% summarise(count = sum(count))
-d$country <- factor(d$country, 
-                    levels = levels(ms$country)[order(ms$count, decreasing=FALSE)])
-                                   
-qplot(x = count, y = country, data = d, 
-      geom = "point", 
-      facets = medal.type ~.)
-
-## ordering within facets
-d$medal.type <- factor(d$medal.type, 
-                       levels = c("gold","silver","bronze"))
-
-d$country_l <- paste(d$country, d$medal.type, sep = "_") # Make every country unique
-
-quartz()
-qplot(x = count, y = reorder(country_l, count), 
-      data = d, geom = "point") + 
-  facet_grid(medal.type ~., scales = "free_y") + 
-  scale_y_discrete("Country", breaks = d$country_l, 
-                   label = d$country)
-

### 2015-03-29 game.client.js

@@ -1,0 +1,474 @@
+/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 
+                  2013 Robert XD Hawkins
+    
+    written by : http://underscorediscovery.com
+    written for : http://buildnewgames.com/real-time-multiplayer/
+    
+    modified for collective behavior experiments on Amazon Mechanical Turk
+
+    MIT Licensed.
+*/
+
+/* 
+   THE FOLLOWING FUNCTIONS MAY NEED TO BE CHANGED
+*/
+
+// A window global for our game root variable.
+var game = {};
+// A window global for our id, which we can use to look ourselves up
+var my_id = null;
+var my_role = null;
+// Keeps track of whether player is paying attention...
+var visible;
+var incorrect;
+var dragging;
+var waiting;
+
+client_ondisconnect = function(data) {
+    // Redirect to exit survey
+    console.log("server booted")
+    var URL = 'http://web.stanford.edu/~rxdh/psych254/replication_project/forms/end.html?id=' + my_id;
+    window.location.replace(URL);
+};
+
+
+/* 
+Note: If you add some new variable to your game that must be shared
+  across server and client, add it both here and the server_send_update
+  function in game.core.js to make sure it syncs 
+
+Explanation: This function is at the center of the problem of
+  networking -- everybody has different INSTANCES of the game. The
+  server has its own, and both players have theirs too. This can get
+  confusing because the server will update a variable, and the variable
+  of the same name won't change in the clients (because they have a
+  different instance of it). To make sure everybody's on the same page,
+  the server regularly sends news about its variables to the clients so
+  that they can update their variables to reflect changes.
+*/
+client_onserverupdate_received = function(data){
+
+    // Update client versions of variables with data received from
+    // server_send_update function in game.core.js
+    if(data.players) {
+        _.map(_.zip(data.players, game.players),
+            function(z){
+                z[1].id = z[0].id;
+            })
+    }
+
+    var dataNames = _.map(data.objects, 
+        function(e){ return e.name})
+    var localNames = _.map(game.objects,
+        function(e){return e.name})
+
+    // If your objects are out-of-date (i.e. if there's a new round), update them
+    if (game.objects.length == 0 || !_.isEqual(dataNames, localNames)) { 
+        game.objects = _.map(data.objects, function(obj) {
+            var imgObj = new Image()
+            imgObj.src = obj.url
+            // Set it up to load properly
[…]

### 2015-03-29 game.server.js

@@ -1,0 +1,303 @@
+/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 2013 Robert XD Hawkins
+    
+    written by : http://underscorediscovery.com
+    written for : http://buildnewgames.com/real-time-multiplayer/
+    
+    modified for collective behavior experiments on Amazon Mechanical Turk
+
+    MIT Licensed.
+*/
+
+//require('look').start()
+
+    var
+        use_db      = false,
+        game_server = module.exports = { games : {}, game_count:0, assignment:0},
+        fs          = require('fs');
+	    
+    if (use_db) {
+	    database    = require(__dirname + "/database"),
+	    connection  = database.getConnection();
+    }
+
+global.window = global.document = global;
+require('./game.core.js');
+utils = require('./utils.js');
+
+var moveObject = function(client, i, x, y) {
+    var obj = client.game.gamecore.objects[i]
+    var others = client.game.gamecore.get_others(client.userid);
+    obj.trueX = parseInt(x)
+    obj.trueY = parseInt(y)
+    _.map(others, function(p) {
+      p.player.instance.emit('objMove', {i: i, x: x, y: y})
+  })
+}
+
+// This is the function where the server parses and acts on messages
+// sent from 'clients' aka the browsers of people playing the
+// game. For example, if someone clicks on the map, they send a packet
+// to the server (check the client_on_click function in game.client.js)
+// with the coordinates of the click, which this function reads and
+// applies.
+game_server.server_onMessage = function(client,message) {
+//    console.log("received message: " + message)
+    //Cut the message up into sub components
+    var message_parts = message.split('.');
+
+    //The first is always the type of message
+    var message_type = message_parts[0];
+
+    //Extract important variables
+    var gc = client.game.gamecore
+    var id = gc.instance.id.slice(0,6)
+    var all = gc.get_active_players();
+    var target = gc.get_player(client.userid);
+    var others = gc.get_others(client.userid);
+    switch(message_type) {
+        case 'objMove' :    // Client is changing angle
+            moveObject(client, message_parts[1], message_parts[2], message_parts[3])
+            break;
+
+        case 'correctDrop' :
+            moveObject(client, message_parts[1], message_parts[2], message_parts[3])
+            gc.attemptNum = 0;
+            _.map(all, function(p) {p.player.instance.send("s.waiting.correct")})
+            gc.paused = true;
+            break;
+
+        case 'incorrectDrop' :
+            writeData(client, "error", message_parts)
[…]

### 2015-03-29 replication_project/app.js

@@ -1,109 +1,0 @@
-/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 2013 Robert XD Hawkins
-
-    originally written for: http://buildnewgames.com/real-time-multiplayer/
-    
-    substantially modified for collective behavior experiments 
-
-    MIT Licensed.
-*/
-
-var 
-    use_db          = false,
-    gameport        = 8888,
-    app             = require('express')(),
-    server          = app.listen(gameport),
-    io              = require('socket.io')(server),
-    _               = require('underscore'),
-    fs              = require('fs');
-
-if (use_db) {
-    database        = require(__dirname + "/database"),
-    connection      = database.getConnection();
-}
-
-game_server = require('./game.server.js');
-utils = require('./utils.js');
-
-global_player_set = {}
-
-// Log something so we know that server-side setup succeeded
-console.log("info  - socket.io started");
-console.log('\t :: Express :: Listening on port ' + gameport );
-
-//  This handler will listen for requests on /*, any file from the
-//  root of our server. See expressjs documentation for more info 
-app.get( '/*' , function( req, res ) {
-    // this is the current file they have requested
-    var file = req.params[0]; 
-    console.log('\t :: Express :: file requested: ' + file);    
-    
-    if(req.query.id && !valid_id(req.query.id)) {
-       res.redirect('http://projects.csail.mit.edu/ci/turk/forms/invalid.html')
-    } else {
-	if(req.query.id && req.query.id in global_player_set) {
-	    res.redirect('http://projects.csail.mit.edu/ci/turk/forms/duplicate.html')
-	} else {
-	    // give them what they want
-	    res.sendfile("./" + file);
-	}
-    }
-}); 
-
-// Socket.io will call this function when a client connects. We check
-// to see if the client supplied a id. If so, we distinguish them by
-// that, otherwise we assign them one at random
-io.on('connection', function (client) {
-    // Recover query string information and set condition
-    var hs = client.handshake;    
-    var query = require('url').parse(client.handshake.headers.referer, true).query;
-    if( !(query.id && query.id in global_player_set) ) {
-        if(query.id) {
-            console.log("got id")
-            global_player_set[query.id] = true;
-            // use id from query string if exists
-            var id = query.id; 
-        } else {
-            var id = utils.UUID();
-        }
-        if(valid_id(id)) {
-            console.log("user connecting...")
-            initialize(query, client, id);
-        }
-    }
-});
-
-var valid_id = function(id) {
-    return id.length == 41;
-}
-
[…]

### 2015-03-29 replication_project/drawing.js

@@ -1,226 +1,0 @@
-var occludedList = [[4,4], [2,2], [2,3], [3,1], [1,4]];
-
-var drawGrid = function(game){
-    //size of canvas
-    var cw = game.viewport.width;
-    var ch = game.viewport.height;
-
-    //padding around grid
-    var p = 25;
-
-    //grid width and height
-    var bw = cw - (p*2) ;
-    var bh = ch - (p*2) ;
-
-    game.ctx.beginPath();
-
-    // vertical lines
-    for (var x = 0; x <= bw; x += Math.floor((cw - 2*p) / 4)) {
-        game.ctx.moveTo(0.5 + x + p, p);
-        game.ctx.lineTo(0.5 + x + p, bh + p);}
-
-    // horizontal lines
-    for (var x = 0; x <= bh; x += Math.floor((ch - 2*p) / 4)) {
-        game.ctx.moveTo(p, 0.5 + x + p);
-        game.ctx.lineTo(bw + p, 0.5 + x + p);}
-
-    game.ctx.lineWidth = 1;
-    game.ctx.strokeStyle = "black";
-    game.ctx.stroke();
-
-    // occluded cells...
-    for(var i = 0; i < occludedList.length; i++) {
-      var cell = occludedList[i]
-      drawOccludedCell(cell[0],cell[1])
-    }
-}
-
-var drawClickPoint = function(game) {
-  var centerX = game.viewport.width / 2;
-  var centerY = game.viewport.height / 2;
-  var radius = 8;
-
-  game.ctx.beginPath();
-  game.ctx.arc(centerX, centerY, radius, 0, 2 * Math.PI, false);
-  game.ctx.fillStyle = 'green';
-  game.ctx.fill();
-  game.ctx.lineWidth = 3;
-  game.ctx.strokeStyle = '#003300';
-  game.ctx.stroke();
-}
-
-var drawOccludedCell = function(x, y) {
-    var cell = game.getPixelFromCell(y,x)
-    var topLeft = [cell.centerX - cell.width/2, cell.centerY - cell.height/2]
-    var topRight = [cell.centerX + cell.width/2, topLeft[1]]
-    var bottomLeft = [topLeft[0], cell.centerY + cell.height/2]
-    var bottomRight = [topRight[0], bottomLeft[1]]
-    game.ctx.beginPath();
-    game.ctx.moveTo(topLeft[0], topLeft[1])
[…]

### 2015-03-29 replication_project/game.client.js

@@ -1,474 +1,0 @@
-/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 
-                  2013 Robert XD Hawkins
-    
-    written by : http://underscorediscovery.com
-    written for : http://buildnewgames.com/real-time-multiplayer/
-    
-    modified for collective behavior experiments on Amazon Mechanical Turk
-
-    MIT Licensed.
-*/
-
-/* 
-   THE FOLLOWING FUNCTIONS MAY NEED TO BE CHANGED
-*/
-
-// A window global for our game root variable.
-var game = {};
-// A window global for our id, which we can use to look ourselves up
-var my_id = null;
-var my_role = null;
-// Keeps track of whether player is paying attention...
-var visible;
-var incorrect;
-var dragging;
-var waiting;
-
-client_ondisconnect = function(data) {
-    // Redirect to exit survey
-    console.log("server booted")
-    var URL = 'http://web.stanford.edu/~rxdh/psych254/replication_project/forms/end.html?id=' + my_id;
-    window.location.replace(URL);
-};
-
-
-/* 
-Note: If you add some new variable to your game that must be shared
-  across server and client, add it both here and the server_send_update
-  function in game.core.js to make sure it syncs 
-
-Explanation: This function is at the center of the problem of
-  networking -- everybody has different INSTANCES of the game. The
-  server has its own, and both players have theirs too. This can get
-  confusing because the server will update a variable, and the variable
-  of the same name won't change in the clients (because they have a
-  different instance of it). To make sure everybody's on the same page,
-  the server regularly sends news about its variables to the clients so
-  that they can update their variables to reflect changes.
-*/
-client_onserverupdate_received = function(data){
-
-    // Update client versions of variables with data received from
-    // server_send_update function in game.core.js
-    if(data.players) {
-        _.map(_.zip(data.players, game.players),
-            function(z){
-                z[1].id = z[0].id;
-            })
-    }
-
-    var dataNames = _.map(data.objects, 
-        function(e){ return e.name})
-    var localNames = _.map(game.objects,
-        function(e){return e.name})
-
-    // If your objects are out-of-date (i.e. if there's a new round), update them
-    if (game.objects.length == 0 || !_.isEqual(dataNames, localNames)) { 
-        game.objects = _.map(data.objects, function(obj) {
-            var imgObj = new Image()
-            imgObj.src = obj.url
-            // Set it up to load properly
[…]

### 2015-03-29 replication_project/game.core.js

@@ -1,287 +1,0 @@
-/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 
-                  2013 Robert XD Hawkins
-    
-    written by : http://underscorediscovery.com
-    written for : http://buildnewgames.com/real-time-multiplayer/
-    
-    substantially modified for collective behavior experiments on the web
-
-    MIT Licensed.
-*/
-
-/*
-  The main game class. This gets created on both server and
-  client. Server creates one for each game that is hosted, and each
-  client creates one for itself to play the game. When you set a
-  variable, remember that it's only set in that instance.
-*/
-var has_require = typeof require !== 'undefined'
-
-if( typeof _ === 'undefined' ) {
-    if( has_require ) {
-        _ = require('underscore')
-    }
-    else throw new ('mymodule requires underscore, see http://underscorejs.org');
-}
-
-var game_core = function(game_instance){
-
-    this.debug = false
-
-    // Define some variables specific to our game to avoid
-    // 'magic numbers' elsewhere
-    this.instance = game_instance;
-
-    //Store a flag if we are the server instance
-    this.server = this.instance !== undefined;
-
-    //Dimensions of world -- Used in collision detection, etc.
-    this.world = {width : 600, height : 600};  // 160cm * 3
-    this.roundNum = -1;
-    this.instructionNum = -1;
-    this.numRounds = 8;
-    this.attemptNum = 0; // Increments whenever someone makes a mistake
-    this.paused = true;
-    this.objects = [];
-    this.instructions = []
-    this.currentDestination = [];
-    if(this.server) {
-        this.trialList = this.makeTrialList()
-        this.players = [{
-            id: this.instance.player_instances[0].id, 
-            player: new game_player(this,this.instance.player_instances[0].player)
-        }];
-        this.server_send_update()
-    } else {
-        this.players = [{
-            id: null, 
-            player: new game_player(this)
-        }]
-    }
-}; 
-
-var game_player = function( game_instance, player_instance) {
-    //Store the instance, if any
-    this.instance = player_instance;
-    this.game = game_instance;
-    this.role = ''
-    //Set up initial values for our state information
-    this.message = '';
-    this.id = '';
-}; 
-
-// server side we set some classes to global types, so that
-// it can use them in other files (specifically, game.server.js)
-if('undefined' != typeof global) {
[…]

### 2015-03-29 replication_project/game.server.js

@@ -1,303 +1,0 @@
-/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 2013 Robert XD Hawkins
-    
-    written by : http://underscorediscovery.com
-    written for : http://buildnewgames.com/real-time-multiplayer/
-    
-    modified for collective behavior experiments on Amazon Mechanical Turk
-
-    MIT Licensed.
-*/
-
-//require('look').start()
-
-    var
-        use_db      = false,
-        game_server = module.exports = { games : {}, game_count:0, assignment:0},
-        fs          = require('fs');
-	    
-    if (use_db) {
-	    database    = require(__dirname + "/database"),
-	    connection  = database.getConnection();
-    }
-
-global.window = global.document = global;
-require('./game.core.js');
-utils = require('./utils.js');
-
-var moveObject = function(client, i, x, y) {
-    var obj = client.game.gamecore.objects[i]
-    var others = client.game.gamecore.get_others(client.userid);
-    obj.trueX = parseInt(x)
-    obj.trueY = parseInt(y)
-    _.map(others, function(p) {
-      p.player.instance.emit('objMove', {i: i, x: x, y: y})
-  })
-}
-
-// This is the function where the server parses and acts on messages
-// sent from 'clients' aka the browsers of people playing the
-// game. For example, if someone clicks on the map, they send a packet
-// to the server (check the client_on_click function in game.client.js)
-// with the coordinates of the click, which this function reads and
-// applies.
-game_server.server_onMessage = function(client,message) {
-//    console.log("received message: " + message)
-    //Cut the message up into sub components
-    var message_parts = message.split('.');
-
-    //The first is always the type of message
-    var message_type = message_parts[0];
-
-    //Extract important variables
-    var gc = client.game.gamecore
-    var id = gc.instance.id.slice(0,6)
-    var all = gc.get_active_players();
-    var target = gc.get_player(client.userid);
-    var others = gc.get_others(client.userid);
-    switch(message_type) {
-        case 'objMove' :    // Client is changing angle
-            moveObject(client, message_parts[1], message_parts[2], message_parts[3])
-            break;
-
-        case 'correctDrop' :
-            moveObject(client, message_parts[1], message_parts[2], message_parts[3])
-            gc.attemptNum = 0;
-            _.map(all, function(p) {p.player.instance.send("s.waiting.correct")})
-            gc.paused = true;
-            break;
-
-        case 'incorrectDrop' :
-            writeData(client, "error", message_parts)
[…]

### 2015-03-29 replication_project/stimuli/objectSet.js

@@ -1,289 +1,0 @@
-// Indexed by object set ID
-
-
-
-// BLOCK 1
-var sunGlasses = {
-	url: 'stimuli/sunGlasses.png', name: "sunGlasses", width: 130, height: 65,
-	instruction: "sunGlasses right", initialLoc : [4,3], critical : "target",
-	scriptedInstruction: "move the glasses one space to the right"}
-var glassesCase = {
-	url: 'stimuli/glassesCase.png', name: "glassesCase", width : 100, height: 107,
-	instruction: "", initialLoc: [3,1], critical : "distractor",}
-var soccerBall = {
-	url: 'stimuli/soccerBall.jpg', name: "soccerBall", width: 100, height: 100,
-	instruction: "", initialLoc: glassesCase.initialLoc, critical : "distractor",}
-var saxophone = {
-	url: 'stimuli/saxophone.png', name: "saxophone", width: 97, height: 130,
-	instruction: "saxophone up", initialLoc: [4,2], critical : "filler",}
-var airplane = {
-	url: 'stimuli/airplane.jpg', name: 'airplane', width: 130, height: 60,
-	instruction: "airplane right", initialLoc : [1,2],critical : "filler",}
-var barrel = {
-	url: 'stimuli/barrel.png', name: 'barrel', width: 87, height: 130,
-	instruction: "barrel down", initialLoc : [2,4], critical : "filler",
-	scriptedInstruction: "move the barrel down by one space"}
-var watch = {
-	url: 'stimuli/watch.png', name: 'watch', width: 130, height: 84,
-	instruction: "", initialLoc: [2,2], critical : "filler",}
-
-// BLOCK 2
-var middleBlock = {
-	url: 'stimuli/middleBlock.png', name: "middleBlock", width: 80, height: 80,
-	instruction: "middleBlock left", initialLoc: [3,4],
-	scriptedInstruction: "move the bottom block one space to the left", critical : "target"}
-var bottomBlock = {
-	url: 'stimuli/bottomBlock.jpg', name: "bottomBlock", width: 80, height: 80,
-	instruction: "", initialLoc: [4,4], critical : "distractor"}
-var stapler = {
-	url: 'stimuli/stapler.jpg', name: "stapler", width: 80, height: 80,
-	instruction: "", initialLoc: [4,4], critical : "distractor",}
-var topBlock = {
-	url: 'stimuli/topBlock.jpg', name: "topBlock", width: 80, height: 80,
[…]

### 2016-01-17 writing/cogsci-submission/cogsci_keysar.tex
Checked in initial draft of cog sci paper

% arara: pdflatex
% arara: bibtex
% arara: pdflatex
% arara: pdflatex
% 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified : Niels Taatgen (taatgen@cmu.edu)         10/24/2006
% Modified : David Noelle (dnoelle@ucmerced.edu)     11/19/2014

%% Change "letterpaper" in the following line to "a4paper" if you must.

\documentclass[10pt,letterpaper]{article}

\usepackage{hyperref}
\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{apacite}
\usepackage{color}
\usepackage{todonotes}
\usepackage{dsfont}
\usepackage{textcomp}

\definecolor{Red}{RGB}{255,0,0}
\newcommand{\red}[1]{\textcolor{Red}{#1}}

\newcommand{\jd}[1]{\green{$^*$}\marginpar{\footnotesize{JD: \green{#1}}}}

\newcommand{\subsubsubsection}[1]{{\em #1}}
\newcommand{\eref}[1]{(\ref{#1})}
\newcommand{\tableref}[1]{Table \ref{#1}}
\newcommand{\figref}[1]{Figure \ref{#1}}
\newcommand{\appref}[1]{Appendix \ref{#1}}
\newcommand{\sectionref}[1]{Section \ref{#1}}

\title{?}

\author{{\large \bf Robert X.~D.~Hawkins, Noah D.~Goodman}\\
  \{rxdh,ngoodman\}@stanford.edu\\
  Department of Psychology, 450 Serra Mall \\
  Stanford, CA 94305 USA}

\begin{document}

\maketitle

\begin{abstract}
?

\textbf{Keywords:} 
?
\end{abstract}

\section{Introduction}
\label{sec:intro}

Humans can accurately and intelligently reason about the mental states of other humans. Among other things, this ability -- called \emph{theory of mind} \cite{PremackWoodruff78_ChimpanzeeToM} -- allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}. Most children acquire this ability by age six \cite{WimmerPerner83_BeliefsAboutBeliefs, WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theories \cite{GopnikWellman12_ReconstructingConstructivism}.

While theory of mind use often appears to be automatic and effortless, Keysar and colleagues \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind} have argued that it is actually the opposite, even for adults: we are ``mindblind" by default and only overcome our egocentric biases through an effortful process of perspective-taking. In other words, while adults are \emph{capable} of applying theory of mind reasoning, we do not always apply it reliably. This argument is based on a simple experimental paradigm, where participants played a simple communication game with a confederate. The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects. The confederate played the role of `director,' giving instructions about how to move objects around a grid, and the participant played the role of `matcher,' attempting to follow these instructions. For example, the objects in one trial included a medium-sized measuring cup in one location and a smaller measuring cup in another location. The director gave an instruction like `move the large cup up one square,' referring to the medium-sized one.

Some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge. To perform accurately, the matcher would need to apply theory of mind and reason about the director?s beliefs. For example, if a much larger measuring cup were placed in an occluded slot and agents failed to account for the director?s beliefs, they might interpret `the large cup' to mean this occluded cup even though the director couldn't possibly know about it. Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent. 46\% of participants reached for the hidden object at least twice. Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.

While these results are compelling, the paradigm has been criticized from several different angles. First, in all experiments, the privileged object was a better fit for the referring expression than the one in common ground (e.g. the largest measuring cup vs. the medium measuring cup for ``the large measuring cup"), making the privileged object \emph{a priori} more likely to be the referent. It would be fairer to compare two objects that fit the referring expression equally well \cite{HellerGrodnerTanenhaus08_Perspective}. Second, the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility \cite{HannaTanenhausTrueswell03_CommonGroundPerspective}. Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g. the largest measuring cup) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}. It would be useful to know whether people referred to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results. %Note that none of these criticisms invalidate the paradoxical result that listeners move an occluded object, they just suggest an alternative explanation. Reference disambiguation is probabilistic: common ground and goodness of fit are weighed against one another.

In this paper, we offer a novel pragmatic account of Keysar's results. When both players know that objects are occluded in the display, we note that the speaker is likely to be overinformative. If the listener \emph{expects} the speaker to be overinformative, they will pragmatically pick the \emph{a priori} more likely referent of the referring expression, which in critical trials will be the occluded object. In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item. We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment, recruiting participants to be both director and matcher instead of using a confederate. Instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study. We then ran the same experiment without scripted items, observing unconstrained director behavior. This minimal pair of experiments demonstrates that listener mistakes are in fact due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to theory of mind.

\section{Exp.~1: Scripted Replication}
\label{sec:Exp1}

\subsection{Materials:}

Participants were  4 x 4 slots. Five of the slots were occluded from the director?s perspective, and the remaining 11 were visible to both participant and director. One of these unoccluded slots included an object that was the target, such as a cassette tape box. Another object, such as a roll of tape, was hidden by the participant in a small brown paper bag and placed in an occluded slot. In addition to the intended object and the object in the bag, each array had several unrelated objects. For each grid, there was one ?critical instruction? (e.g. ?move the tape?) in which the director gave an instruction to move a mutually-visible object that could also potentially refer to the object hidden in the bag.

The experiment had eight items, each with a different set of objects and critical instruction. Each item consisted of a series of instructions to move objects around and included one critical instruction. Each item also included one critical pair of objects, one of them the intended object and the other hidden in the bag, such as the cassette tape and the roll of tape.

To collect baseline performance information for each item we added a condition in which the hidden object (e.g. roll of tape) was replaced with an object that did not fit the critical instruction (e.g. a battery). Thus, the `move the tape? instruction appeared after the participant had hidden in the bag either a roll of tape (experimental condition) or a battery (baseline condition). Each participant received half the items in the experimental condition and half in the baseline condition. Items and conditions were counterbalanced across participants. Order of presentation was random, with the provision that no more than two items in the same condition would appear consecutively.

A trained female confederate played the role of the director in order to ensure uniformity of critical instructions across conditions and participants. The confederate was well practiced in playing the role of a naive participant. To create a realistic situation, she indicated having some difficulty with the task, interjected her instructions with hesitations, and made occasional errors with non-critical objects. In addition, the director improvised most of the instructions, except that critical instructions for the target objects were scripted. Indeed, with the exception of one person, none of the participants later reported that they suspected during the experiment that the director was a confederate.?

We will precisely replicate all objects sets, instruction sets, and randomization schemes. However, given the constraints of the web experiment format, there will be several exceptions as well. First, participants cannot be seated across from each other at a table: they will each see their respective views of the 4 x 4 grid on a screen and will communicate via a text box. Second, we will not use a trained confederate, and will instead randomly assign one of the players to the role of the instruction giver. To make sure the critical instructions are still scripted, we will provide a ?send instructions? button for a subset of trials (including some non-critical ones).

\subsection{Procedure}

The original article states:

?The participant and the confederate arrived at the laboratory and the experimenter explained that they would be playing two different roles in a communication game. She then assigned roles, ostensibly randomly, and the participant received the role of the `addressee? while the confederate was assigned the director?s role. At the beginning of each item, the director received a picture of the array of objects with arrows indicating where each object should be moved. The arrows were numbered to specify the order of object movement. The director then used this picture and instructed the participant to rearrange the objects accordingly.

The director?s picture showed her perspective, meaning that only mutually-visible objects appeared on the picture, with the remainder of the slots clearly occluded. This was demonstrated to the participant, and the experimenter also pointed out that the objects in the occluded slots were not part of the game. In addition, before each item began, the experimenter put a large cardboard wall between the confederate and the participant as a visual barrier. Then she handed the participant an object and a brown paper bag and asked the participant to hide the object in the bag and place the bag in one of the occluded slots. The experimenter did not name the object but simply handed it to the participant and referred to it only as `this?. After the participant had hidden the object in the bag and put the bag in the slot, the experimenter removed the barrier and the director started with the instructions.

The experiment began with a practice item to familiarize the participant with the task and to correct any misunderstanding. In order to make absolutely sure that the participant fully appreciated the director?s difference in perspective, the participant and the director switched roles and the participant gave instructions for a second practice item. In this manner there would be no question that the participant understood the information provided in the picture of the array, appreciated that the director could not see hidden objects, and knew that the only objects relevant to the game were the mutually-visible ones. After the role reversal, the participant and the confederate resumed their original roles and the experimenter presented the first item. The experiment proceeded through all eight items, with the director providing instructions and the participant moving the objects. Before each instruction the director said `ready?? at which point the participant looked at the center of the array and answered `ready?. The participant was free to converse with the director, to ask questions and so on.?

We will follow these procedures precisely, but all the practice items will be moved to a general set of instructions, before players enter the real-time, multi-player environment. Because there will be no experimenter actively administrating the practice trials, we will give a comprehension test to participants before they are assigned to the role of ?director? or ?agent.? In this way, both participants will understand both tasks and appreciate the information asymmetry.

\subsection{Results}

Same or stronger than Keysar, except used mouse-tracking instead of eye-tracking...

Item-wise analysis

\subsection{Discussion}

\begin{itemize}
\item quibble with ?moved at least one thing? measure (allows one difficult item to dominate if item-wise differences)
\item in non-scripted instructions, noticed that people were being overinformative
\end{itemize}

\section{Exp.~2: Unscripted Replication}
\label{sec:Exp2}

\subsection{Methods}

 same as exp 1, except no scripted instructions (i.e. we don?t force the director to say ?tape?)

\subsection{Results}

\begin{itemize}
\item Listener makes fewer mistakes?
\item Director more overinformative on critical items?
\end{itemize}

\subsection{Discussion}

\begin{itemize}
\item scripted instruction unnatural, 
\item violates listener expectations of overinformativity
\end{itemize}

\section{General Discussion}

\begin{itemize}
\item Keysar's results are ironically accounted for by a ToM argument, via pragmatics
\item Running multi-player control w/o confederate is important to know whether confederate behavior is valid and cooperative
\end{itemize}

\section{Acknowledgements}

\small We're grateful to Boaz Keysar for providing select materials for our replication. Exp.~1 was originally conducted under the supervision of Michael Frank, with early input from Desmond Ong. This work was supported by ONR grants N00014-13-1-0788 and N00014-13-1-0287,  and a James S. McDonnell Foundation Scholar Award to NDG. RXDH was supported by the Stanford Graduate Fellowship and the National Science Foundation Graduate Research Fellowship under Grant No. DGE-114747.

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{bibs}

\end{document}

### 2016-01-21 experiment2/analysis/exp2Analysis.Rmd
Renamed analysis scripts

--- 2015-03-29 analysis/keysarMouseTracking.Rmd
@@
 ```{r}
-setwd('~/Repos/psych254/replication_project/analysis/')
-mouseD = read.csv('fullSampleMouse.csv')
-errorD = read.csv('fullSampleError.csv')
-messageD = read.csv('fullSampleMessage.csv')
+setwd('~/Repos/keysar_replication/experiment2/data/')
+mouseD = read.csv('mouse/exp2Mouse.csv')
+errorD = read.csv('error/exp2Error.csv')
+messageD = read.csv('message/exp2Message.csv')
 str(mouseD)
@@
 ```{r}
+nonNativeEnglishIDs <- c("0366-3", "8676-f")
+numParticipants <- length(unique(mouseD$gameid))
 nonCriticalMistakes <- errorD %>%
@@
 
-bannedIDs <- nonCriticalMistakes$gameid
+mistakeIDs <- as.character(nonCriticalMistakes$gameid)
+bannedIDs <- c(mistakeIDs, nonNativeEnglishIDs)
 ```
@@
 eD <- errorD %>%
-  filter(!(gameid %in% bannedIDs)) %>%
+  filter(!(gameid %in% mistakeIDs)) %>%
+  filter(!(gameid %in% nonNativeEnglishIDs)) %>%
   filter(critical == 1) %>%
@@
   group_by(condition) %>%
-  summarise(atLeastOnce = length(n)/14, 
-            atLeastTwice = sum(cond2)/14)
+  summarise(atLeastOnce = length(n)/numParticipants, 
+            atLeastTwice = sum(cond2)/numParticipants,
+            total = sum(n) / (4 * numParticipants))
 eD
@@
   filter(critical == 1) %>%
-  filter(!(objectSet %in% c(2, 6, 7))) %>%
+  filter(!(objectSet %in% c(5, 6, 8))) %>% # Just 5 & 6?
   filter(attemptNum == 0) %>%
@@
   group_by(condition) %>%
-  summarise(atLeastOnce = length(n)/14, 
-            atLeastTwice = sum(cond2)/14)
+  summarise(atLeastOnce = length(n)/numParticipants, 
+            atLeastTwice = sum(cond2)/numParticipants,
+            total = sum(n) / (4 * numParticipants))
 excludedED

### 2016-01-21 experiment1/analysis/exp1Analysis.Rmd
Updated analysis scripts

--- 2015-03-29 analysis/keysarMouseTracking.Rmd
@@
 ```{r}
-setwd('~/Repos/psych254/replication_project/analysis/')
+setwd('~/Repos/keysar_replication/experiment1/analysis/')
 mouseD = read.csv('fullSampleMouse.csv')
@@
 ```{r}
+numParticipants <- length(unique(mouseD$gameid))
 nonCriticalMistakes <- errorD %>%
@@
   summarise(atLeastOnce = length(n)/14, 
-            atLeastTwice = sum(cond2)/14)
+            atLeastTwice = sum(cond2)/14,
+            total = sum(n) / (4 * numParticipants))
 eD
@@
                           ymin = percent - error), size =2 ,
-                      width = 0.1))
-g
+                      width = 0.1)
+      + ggtitle("Experiment 1 Hover-Time"))
+ggsave("../../writing/cogsci-submission/images/exp1MouseTracking.pdf")
 ```

### 2016-01-21 writing/cogsci-submission/cogsci_keysar.tex
Filled in exp1 methods & results

@@
 
-\title{?}
+\title{Pragmatics account for limits on theory of mind use in adults}
 
@@
 In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item.
-We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment, recruiting participants to be both director and matcher instead of using a confederate.
+We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} (in a multi-player web experiment, recruiting participants to be both director and matcher instead of using a confederate.
 Instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study.
 We then ran the same experiment without scripted items, observing unconstrained director behavior.
-This minimal pair of experiments demonstrates that listener mistakes are in fact due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to theory of mind.
+This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to theory of mind.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[scale = .35]{images/overall_view.jpg}
+\end{center}
+\vspace{-.25cm}
+\caption{Interface used in the reported experiments. Objects behind the black squares were hidden from the director.}
+\label{fig:interface}
+\end{figure*}
 
@@
 
+\subsection{Participants:}
+
+We recruited 34 participants (17 pairs) from Amazon Mechanical Turk. All participants were from the U.S. Three pairs were excluded for making 2 or more errors on non-critical items.
+
 \subsection{Materials:}
 
-Participants were  4 x 4 slots.
-Five of the slots were occluded from the director?s perspective, and the remaining 11 were visible to both participant and director.
-One of these unoccluded slots included an object that was the target, such as a cassette tape box.
-Another object, such as a roll of tape, was hidden by the participant in a small brown paper bag and placed in an occluded slot.
-In addition to the intended object and the object in the bag, each array had several unrelated objects.
-For each grid, there was one ?critical instruction?
-(e.g.
-?move the tape?) in which the director gave an instruction to move a mutually-visible object that could also potentially refer to the object hidden in the bag.
+Pairs of participants -- assigned randomly to `director' and `matcher' roles -- interacted with one another through a real-time web interface, shown in Fig.
+\ref{fig:interface}.
+On the left side of the screen, participants could freely type messages to one another; on the right side the screen, players could view a set of objects placed in a 4 x 4 grid.
+Five of the grid cells were occluded from the director's perspective, and the remaining 11 were visible to both participant and director.
+Six or seven objects were displayed in the grid at a given time.
+One of these objects was a `target', such as a cassette tape, placed in an unoccluded cell such that both participants could see it.
+Another object, such as a roll of tape, was placed in an occluded slot such that it was only visible to the matcher.
+The rest of the objects were unrelated `fillers' placed in random locations.
+We used the same set of targets and occluded alternatives as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, but we were unable to obtain the filler objects from the original experiment and created our own.
 
-The experiment had eight items, each with a different set of objects and critical instruction.
-Each item consisted of a series of instructions to move objects around and included one critical instruction.
-Each item also included one critical pair of objects, one of them the intended object and the other hidden in the bag, such as the cassette tape and the roll of tape.
+The experiment was composed of eight items, with each item using a different set of objects.
+Each item included one `critical pair' of objects, one of them the target and the other hidden, such as the cassette tape and the roll of tape.
+For each item, we gave the director a series of four instructions to move objects around, which were displayed as a series of arrows pointing from some object to an unoccupied cell.
+One of these instructions was a `critical instruction,' which referred to the target object.
+For half the instructions, directors were free to communicate however they wished.
+For the other half, including the critical instruction, their messages to the matcher were pre-scripted such that we could use the precise wording from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square."
 
-To collect baseline performance information for each item we added a condition in which the hidden object (e.g.
-roll of tape) was replaced with an object that did not fit the critical instruction (e.g.
+We collected baseline performance for each condition by replacing the hidden alternative (e.g.
+a roll of tape) with an object that did not fit the critical instruction (e.g.
 a battery).
-Thus, the `move the tape?
-instruction appeared after the participant had hidden in the bag either a roll of tape (experimental condition) or a battery (baseline condition).
+We also used the same unambiguous replacements as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 Each participant received half the items in the experimental condition and half in the baseline condition.
-Items and conditions were counterbalanced across participants.
-Order of presentation was random, with the provision that no more than two items in the same condition would appear consecutively.
+The assignment of items to conditions was randomized across participants, and the order of conditions was randomized under the constraint that the same condition would not be used on more than two consecutive items.
+All object sets, object placements, and corresponding instruction sets were the same for all participants.
 
-A trained female confederate played the role of the director in order to ensure uniformity of critical instructions across conditions and participants.
-The confederate was well practiced in playing the role of a naive participant.
-To create a realistic situation, she indicated having some difficulty with the task, interjected her instructions with hesitations, and made occasional errors with non-critical objects.
-In addition, the director improvised most of the instructions, except that critical instructions for the target objects were scripted.
-Indeed, with the exception of one person, none of the participants later reported that they suspected during the experiment that the director was a confederate.?
-
-We will precisely replicate all objects sets, instruction sets, and randomization schemes.
-However, given the constraints of the web experiment format, there will be several exceptions as well.
-First, participants cannot be seated across from each other at a table: they will each see their respective views of the 4 x 4 grid on a screen and will communicate via a text box.
-Second, we will not use a trained confederate, and will instead randomly assign one of the players to the role of the instruction giver.
-To make sure the critical instructions are still scripted, we will provide a ?send instructions?
-button for a subset of trials (including some non-critical ones).
+These materials differ from those used by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a few ways.
+First, participants were not seated across from each other at a table: they each saw a views of the 4 x 4 grid on screen and communicated via a text box.
+Second, we did not use a trained confederate.
+We randomly assigned one of the players to the role of the instruction giver, and maintained the original wording by scripting a subset of their instructions.
+Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \citeA{KeysarBarr___Brauner00_TakingPerspective}.
+\begin{table*}
+\begin{center}
+\begin{tabular}{ p{2cm} | r | r |  r | r | r | r}
+& \multicolumn{2}{c|}{\% attempted at least once} & \multicolumn{2}{c|}{\% attempted at least twice} & \multicolumn{2}{c}{\% of total cases}\\
+\hline
+& Original & Replication & Original & Replication & Original & Replication \\
+Experimental & 71 & 93 & 46 & 64 & 30 & 37 \\
+Baseline & 0 & 14 & 0 & 0 & 0 & 3 \\
+\end{tabular}
+\caption{Side-by-side comparison of error rates in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our replication.
+The first column shows the percentage of \emph{participants} attempting to move the occluded distractor at least one of the four possible cases.
+The second column shows the percentage attempting to move it at least twice.
+The third column shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object.
+\todo[inline]{Big caption maybe unnecessary}}
+\label{table:exp1}
+\end{center}
+\end{table*}
 
@@
 
-The original article states:
+Participants interacted in a real-time, multi-player environment on the web \cite{Hawkins15_RealTimeWebExperiments}.
+Before entering this environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface.
+Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
+To deal with participants arriving at unpredictable times, we implemented a waiting room mechanism, where the first person to enter the game environment would see a screen saying ``Waiting for another player..." and the second person would trigger the game to start.
 
-?The participant and the confederate arrived at the laboratory and the experimenter explained that they would be playing two different roles in a communication game.
-She then assigned roles, ostensibly randomly, and the participant received the role of the `addressee?
-while the confederate was assigned the director?s role.
-At the beginning of each item, the director received a picture of the array of objects with arrows indicating where each object should be moved.
-The arrows were numbered to specify the order of object movement.
-The director then used this picture and instructed the participant to rearrange the objects accordingly.
-
-The director?s picture showed her perspective, meaning that only mutually-visible objects appeared on the picture, with the remainder of the slots clearly occluded.
-This was demonstrated to the participant, and the experimenter also pointed out that the objects in the occluded slots were not part of the game.
-In addition, before each item began, the experimenter put a large cardboard wall between the confederate and the participant as a visual barrier.
-Then she handed the participant an object and a brown paper bag and asked the participant to hide the object in the bag and place the bag in one of the occluded slots.
-The experimenter did not name the object but simply handed it to the participant and referred to it only as `this?.
-After the participant had hidden the object in the bag and put the bag in the slot, the experimenter removed the barrier and the director started with the instructions.
-
-The experiment began with a practice item to familiarize the participant with the task and to correct any misunderstanding.
-In order to make absolutely sure that the participant fully appreciated the director?s difference in perspective, the participant and the director switched roles and the participant gave instructions for a second practice item.
-In this manner there would be no question that the participant understood the information provided in the picture of the array, appreciated that the director could not see hidden objects, and knew that the only objects relevant to the game were the mutually-visible ones.
-After the role reversal, the participant and the confederate resumed their original roles and the experimenter presented the first item.
+To collect clean mouse-tracking data, we began every instruction by asking the matcher to click a small circle in the center of the grid.
+After this small circle was clicked, the director was allowed to communicate the next instruction and we started recording from the matcher's mouse.
 The experiment proceeded through all eight items, with the director providing instructions and the participant moving the objects.
-Before each instruction the director said `ready??
-at which point the participant looked at the center of the array and answered `ready?.
-The participant was free to converse with the director, to ask questions and so on.?
-
-We will follow these procedures precisely, but all the practice items will be moved to a general set of instructions, before players enter the real-time, multi-player environment.
-Because there will be no experimenter actively administrating the practice trials, we will give a comprehension test to participants before they are assigned to the role of ?director?
-or ?agent.?
-In this way, both participants will understand both tasks and appreciate the information asymmetry.
+Except for instructions with scripted messages, the participants were free to use the chat box for questions, small talk, or whatever else.
 
@@
 
-Same or stronger than Keysar, except used mouse-tracking instead of eye-tracking...
+Using the error rates on critical items, we compare our behavioral results with those from from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in Table \ref{table:exp1}.
+We find that 93\% of participants attempted to move a hidden object at least once in the Experiment condition, out of four possible items, compared to only 14\% in the baseline condition.
+This is similar to the effect observed by the authors in the original study, which found 71\% and 0\%.
+Our errors were larger across the board, perhaps due to the interface and the Turk population, but the gap between the two is roughly the same size.
 
-Item-wise analysis
+Next, we examine whether there are differences in error rates across items.
+In Table \ref{table:itemWiseExp1}, we break down the pattern of errors by item.
+Because our randomization procedure undersampled the experimental condition for several items, a chi-squared test is inappropriate.
+Using Fisher's exact test, we cannot reject the null hypothesis of independence ($p = 0.13$).
+Still, we note that several conditions have much higher error rates than others -- for example, 75\% of participants in the experimental condition of item 6 made an error (the ``whiteboard eraser'' vs.
+the ``pencil eraser'') while only 17\% of participants in item 8 made an error (the ``computer mouse'' vs.
+the ``toy mouse'').
+
+Finally, as a proxy for the eye-tracking analyses reported by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
+We define the decision window as the span of time between the point when the matcher received their instruction message and when they started moving an object.
+If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt.
+We took the set of all mouse position samples in the decision window (collected approximately every 15ms), and measured what proportion were located inside the cell containing the target vs.
+the cell containing the distractor.
+We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
+
+We found that there is an interaction between cell type and condition on the percent of hover time spent in that cell, $b = -0.21, t(11) = -3.1, p = 0.003$.
+In baseline trials, people spent much more time in the target cell and much less time in the distractor cell than in the experimental condition (see Fig.
+\ref{fig:exp1hover}).
+Note that this analysis includes participants who actually made errors, since the data is too sparse to exclude them.
+%accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
+
+\begin{table*}
+\begin{center}
+\begin{tabular}{ p{2cm} | r | r |  r | r | r | r | r | r |}
+& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
+\# incorrect & 1 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
+\# correct & 5 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
+\end{tabular}
+\caption{Item-wise error analysis}
+\label{table:exp1}
+\end{center}
+\end{table*}
+
+\begin{figure}[b!]
+\begin{center}
+\includegraphics[scale=.8]{images/exp1MouseTracking.pdf}
+\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
+\label{fig:exp1hover}
+\end{center}
+\end{figure}
 
@@
 \end{itemize}
+
+We noticed several difference between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} that prevented a direct replication.
+The primary difference between our study and the original is the fact that it will be run online with participants connected via a virtual environment, instead of face-to-face in a room.
+A related difference is our decision not to use a confederate.
+In general, studies using confederates are difficult to replicate exactly and are impractical to run online.
+We are aware that this change (while keeping scripted instructions, which were the primary justification for the confederate) may lead to additional variation across groups.
+It is also known that textual communication (via our instant messaging interface) can differ from verbal communication (as in the original study).
+Finally, the way people attempt to move things in our virtual environment may be different from how people attempt to move things in the lab, and we cannot eye-tracking data to back up our analysis, as the original paper did.
 
@@
 
-\begin{itemize}
-\item Keysar's results are ironically accounted for by a ToM argument, via pragmatics
-\item Running multi-player control w/o confederate is important to know whether confederate behavior is valid and cooperative
-\end{itemize}
+First, our analysis of item-level differences shows that the dependent variable used in the original study (i.e.
+?percentage of participants who moved the critical item at least once?) is somewhat problematic.
+It could look like 100% of participants made the errors even if they all made those errors on one particularly difficult item (e.g.
+the ?bottom?
+block, where they actually mean the one on the second-to-bottom row).
+If we exclude the three ?hard?
+items (2, 6, and 7) where over 60% of participants in the experimental condition made errors, our results look much less strong than Keysar?s.
+
+At least once
+At least twice
+Experiment
+43
+14
+Baseline
+7
+0
+
+This suggests that (1) the ?at least one error?
+DV is not appropriate in settings with high variability across items and (2) we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
+
+Second, it?s worth thinking about some ways our online version may differ from the original.
+The fact that we didn?t use a confederate didn?t seem to be a problem, given our use of randomly scripted instructions.
+Similarly, because players could type messages directly to one another, and since the director could watch the matcher moving objects in real-time, it?s also fair to say that participants truly believed they were playing with another human being.
+
+One major concern is our graphical representation of occluded cells.
+We simply told players that items behind black cells were hidden to the director, and this was a critical question in the quiz that players had to pass before playing, so we know that they were aware of it.
+But they may forget part-way through the experiment, or don?t fully internalize it.
+In the in-lab version, there was an actual divider between players, so it would have been more salient.
+This would mean that some errors in our version game aren?t due to a failure of perspective-taking, but instead because of a misunderstanding of what the other person?s perspective is to begin with.
+However, this makes our task
+
+Running multi-player control w/o confederate is important to know whether confederate behavior is valid and cooperative
 

### 2016-01-21 experiment2/analysis/exp2Analysis.Rmd
Fix number of participants to include in denominator

@@
 nonNativeEnglishIDs <- c("0366-3", "8676-f")
-numParticipants <- length(unique(mouseD$gameid))
 nonCriticalMistakes <- errorD %>%
@@
 mistakeIDs <- as.character(nonCriticalMistakes$gameid)
-bannedIDs <- c(mistakeIDs, nonNativeEnglishIDs)
+bannedIDs <- unique(c(mistakeIDs, nonNativeEnglishIDs))
+numParticipants <- length(unique(mouseD$gameid)) - length(bannedIDs)
 ```
@@
 eD <- errorD %>%
-  filter(!(gameid %in% mistakeIDs)) %>%
-  filter(!(gameid %in% nonNativeEnglishIDs)) %>%
+  filter(!(gameid %in% bannedIDs)) %>%
   filter(critical == 1) %>%
@@
   filter(critical == 1) %>%
-  filter(!(objectSet %in% c(5, 6, 8))) %>% # Just 5 & 6?
+  filter(!(objectSet %in% c(5, 6))) %>% # Just 5 & 6?
   filter(attemptNum == 0) %>%

### 2016-01-21 writing/cogsci-submission/cogsci_keysar.tex
Fill in most of Exp. 2, main figures

@@
 Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \citeA{KeysarBarr___Brauner00_TakingPerspective}.
+
 \begin{table*}
 \begin{center}
-\begin{tabular}{ p{2cm} | r | r |  r | r | r | r}
-& \multicolumn{2}{c|}{\% attempted at least once} & \multicolumn{2}{c|}{\% attempted at least twice} & \multicolumn{2}{c}{\% of total cases}\\
+\begin{tabular}{ p{2cm} | r | r |  r || r | r | r || r | r | r}
+& \multicolumn{3}{c||}{\% attempted at least once} & \multicolumn{3}{c||}{\% attempted at least twice} & \multicolumn{3}{c}{\% of total cases}\\
 \hline
-& Original & Replication & Original & Replication & Original & Replication \\
-Experimental & 71 & 93 & 46 & 64 & 30 & 37 \\
-Baseline & 0 & 14 & 0 & 0 & 0 & 3 \\
+& Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2  \\
+Experimental & 71 & 93 & 60 & 46 & 64 & 36 & 30 & 37 & 26\\
+Baseline & 0 & 14 & 14 & 0 & 0 & 0 & 0 & 3 & 4\\
 \end{tabular}
-\caption{Side-by-side comparison of error rates in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our replication.
-The first column shows the percentage of \emph{participants} attempting to move the occluded distractor at least one of the four possible cases.
-The second column shows the percentage attempting to move it at least twice.
-The third column shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object.
-\todo[inline]{Big caption maybe unnecessary}}
-\label{table:exp1}
+\caption{Side-by-side comparison of error rates in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications.
+The first section shows the percentage of \emph{participants} attempting to move the occluded distractor at least one of the four possible cases.
+The second section shows the percentage attempting to move it at least twice.
+The third section shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object.
+\todo[inline]{Big caption maybe unnecessary.
+Anything else to include in this table?
+Maybe Keysar et al, 2000, which reported \% of total cases, or our Exp.
+2 results if you remove the two most difficult items?
+There, errors drop quite a lot.}}
+\label{table:mainResults}
 \end{center}
@@
 
-Using the error rates on critical items, we compare our behavioral results with those from from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in Table \ref{table:exp1}.
+Using the error rates on critical items, we compare our behavioral results with those from from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in Table \ref{table:mainResults}.
 We find that 93\% of participants attempted to move a hidden object at least once in the Experiment condition, out of four possible items, compared to only 14\% in the baseline condition.
@@
 Next, we examine whether there are differences in error rates across items.
-In Table \ref{table:itemWiseExp1}, we break down the pattern of errors by item.
+In Table \ref{table:exp1ItemWise}, we break down the pattern of errors by item.
 Because our randomization procedure undersampled the experimental condition for several items, a chi-squared test is inappropriate.
@@
 \caption{Item-wise error analysis}
-\label{table:exp1}
+\label{table:exp1ItemWise}
 \end{center}
@@
 
+In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it.
+Furthermore, the matcher tended to spend more time hovering their cursor over the occluded cell, and less time hovering over the target object, in ambiguous experimental trials than in baseline trials with no ambiguity.
+Both of these results support Keysar's claim that adults do not reliably deploy theory of mind reasoning when it is called for.
+
+Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data.
+First, our analysis of item-level differences shows that the dependent variable highlighted in the original study (i.e.
+``percentage of participants who moved the critical item at least once'') is somewhat problematic.
+It could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item.
+Indeed, if we exclude the three `hard' items where over 60\% of participants in the experimental condition made errors, this dependent variable drops from 93\% to only 43\% of participants.
+This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
+
+Second, because half of the instructions were unscripted, we were able to observe natural production of referential descriptions for a number of fillers.
+Informally, we noted a tendency toward over-informativity.
+Instead of ``move the stuffed animal down'', they'll say ``move the stuffed pa'nda bear down.'' Or, instead of saying ``move the plane to the right,'' they'll say ``move the red airplane to the right,'' even though there is only one plane.
+This suggested that the confederate's scripted instructions were in some sense uncooperative and counter to the matcher's expectations.
+Perhaps directors were taking the time to make more costly over-informative descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment.
+If the matcher reasoned in this way and expected the director to be overinformative, then they would be justified in picking the first object that meets the description rather than worrying about occluded cells.
+We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
+If speakers really are being overinformative, and listeners are correctly expecting this, we should see fewer errors after removing uncooperative scripts.
+
+\begin{table*}
+\begin{center}
+\begin{tabular}{ p{2cm} | r | r |  r | r | r | r | r | r |}
+& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
+ \# incorrect & 1 & 1 & 2 & 3 & 9 & 7 & 1 & 5 \\
+ \# correct   & 11&14&10&13&7 &8   &12& 8 \\
+\end{tabular}
+\caption{Item-wise error analysis for experiment 2}
+\label{table:exp2ItemWise}
+\end{center}
+\end{table*}
+
+\begin{figure}[b!]
+\begin{center}
+\includegraphics[scale=.9]{images/exp2MouseTracking.pdf}
+\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
+\label{fig:exp2hover}
+\end{center}
+\end{figure}
+\section{Exp.~2: Unscripted Replication}
+\label{sec:Exp2}
+
+\subsection{Participants}
+
+We recruited 64 participants (32 pairs) from Amazon Mechanical Turk, roughly doubling the sample size from Exp.~1.
+All participants were from the U.S.
+Three participants were excluded for making 2 or more errors on non-critical items, and one additional participant was excluded because they were not a native English speaker.
+
+\subsection{Materials \& Procedures}
+
+Everything was the same as Exp.~1, except we did not use scripted messages for critical instructions.
+
+\subsection{Results}
+
+Error rates are reported in Table \ref{table:mainResults}, alongside the results from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our direct replication in Exp.~1.
+Note that error rates for baseline trials are the same as Exp.~1, confirming that participants tend to make more errors in the web version of the experiment.
+Error rates for experimental trials, however, are lower than the rates reported in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and in Exp.~1 \todo[inline]{rdh: need to run a comparison of estimated binomial proportions here}.
+
+More dramatically, we find that patterns of errors in Exp.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 20.1, p = 0.005$.
+Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the costly over-informative utterance remains ambiguous.
+For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
+``Hair brush'' is already a label at the sub-class level.
+If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 26\% to 12\% and the percentage of participants making at least one error drops from 60\% to 36\%.
+
+\todo[inline]{rdh: should merge item-wise analyses for Exp. 1 and Exp. 2 into a single table, and actually write out what the targets and distractors are for each items}
+
+When we conducted a mouse-tracking analysis identical to the one reported for Exp.~1, we again found a significant interaction between condition and cell type on hover time, $b = -0.18, t(108) = -5.2, p < 0.001$.
+However, the effect is attenuated: in experimental trials, participants spend significantly more time hovering over the target than the distractor (see Figure \ref{fig:exp2hover}) whereas in Exp.~1, participants spent roughly the same amount of time hovering over each.
+
+\todo[inline]{rdh: make this analysis more rigorous by explicitly testing the size of the interaction?
+By testing the diff.
+in hover-time in target vs.
+distractor cells after restricting to experimental trials?}
+
+\todo[inline]{rdh: should do a basic analysis of over-informativity...
+Don't forget about the person who always gave directions like ``start at the third square of the first row and move down one square''}
+
+\subsection{Discussion}
+
 \begin{itemize}
-\item quibble with ?moved at least one thing? measure (allows one difficult item to dominate if item-wise differences)
-\item in non-scripted instructions, noticed that people were being overinformative
+\item people still try to move occluded objects, but they do it less, and that behavior is more confined to a few particularly difficult items.
+\item this happens because the scripted instructions we forced directors to use in Exp.~1 were unnatural and violated listener expectations of overinformativity
+\item the fact that the effect is attenuated in exp.
+2 is actually evidence of some theory of mind use (insofar as the director choosing to be overinformative, and the matcher expecting overinformativity constitutes ToM
 \end{itemize}
 
-We noticed several difference between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} that prevented a direct replication.
+\section{General Discussion}
+
+We noticed several differences between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} that prevented a direct replication.
 The primary difference between our study and the original is the fact that it will be run online with participants connected via a virtual environment, instead of face-to-face in a room.
@@
 
-\section{Exp.~2: Unscripted Replication}
-\label{sec:Exp2}
+Second, it's worth thinking about some ways our online version may differ from the original.
+The fact that we didn't use a confederate didn't seem to be a problem, given our use of randomly scripted instructions.
+Similarly, because players could type messages directly to one another, and since the director could watch the matcher moving objects in real-time, it's also fair to say that participants truly believed they were playing with another human being.
 
-\subsection{Methods}
-
- same as exp 1, except no scripted instructions (i.e. we don?t force the director to say ?tape?)
-
-\subsection{Results}
-
-\begin{itemize}
-\item Listener makes fewer mistakes?
-\item Director more overinformative on critical items?
-\end{itemize}
-
-\subsection{Discussion}
-
-\begin{itemize}
-\item scripted instruction unnatural, 
-\item violates listener expectations of overinformativity
-\end{itemize}
-
-\section{General Discussion}
-
-First, our analysis of item-level differences shows that the dependent variable used in the original study (i.e.
-?percentage of participants who moved the critical item at least once?) is somewhat problematic.
-It could look like 100% of participants made the errors even if they all made those errors on one particularly difficult item (e.g.
-the ?bottom?
-block, where they actually mean the one on the second-to-bottom row).
-If we exclude the three ?hard?
-items (2, 6, and 7) where over 60% of participants in the experimental condition made errors, our results look much less strong than Keysar?s.
-
-At least once
-At least twice
-Experiment
-43
-14
-Baseline
-7
-0
-
-This suggests that (1) the ?at least one error?
-DV is not appropriate in settings with high variability across items and (2) we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
-
-Second, it?s worth thinking about some ways our online version may differ from the original.
-The fact that we didn?t use a confederate didn?t seem to be a problem, given our use of randomly scripted instructions.
-Similarly, because players could type messages directly to one another, and since the director could watch the matcher moving objects in real-time, it?s also fair to say that participants truly believed they were playing with another human being.
+\todo[inline]{rdh: still have some worries about whether people really understood the point of the black squares, even after the quiz (e.g.
+one person said ``This seems to work wether the space is black or not.'' Note that this would just inflate our errors, so our analysis here is conservative.)}
 
@@
 We simply told players that items behind black cells were hidden to the director, and this was a critical question in the quiz that players had to pass before playing, so we know that they were aware of it.
-But they may forget part-way through the experiment, or don?t fully internalize it.
+But they may forget part-way through the experiment, or don't fully internalize it.
 In the in-lab version, there was an actual divider between players, so it would have been more salient.
-This would mean that some errors in our version game aren?t due to a failure of perspective-taking, but instead because of a misunderstanding of what the other person?s perspective is to begin with.
-However, this makes our task
+This would mean that some errors in our version game aren't due to a failure of perspective-taking, but instead because of a misunderstanding of what the other person's perspective is to begin with.
+However, this would only bias our estimates of errors toward supporting Keysar's results, making our analysis conservative if anything.
 
-Running multi-player control w/o confederate is important to know whether confederate behavior is valid and cooperative
+\todo[inline]{rdh: methodological note that running multi-player control w/o confederate is important to know whether confederate behavior is valid and cooperative}
 

### 2016-01-21 writing/cogsci-submission/cogsci_keysar.tex
Updated title; fixed some typos

@@
 
-\title{Pragmatics account for limits on theory of mind use in adults}
+\title{Pragmatics account for apparent limits on theory of mind use in adults}
 
@@
 Some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
-To perform accurately, the matcher would need to apply theory of mind and reason about the director?s beliefs.
-For example, if a much larger measuring cup were placed in an occluded slot and agents failed to account for the director?s beliefs, they might interpret `the large cup' to mean this occluded cup even though the director couldn't possibly know about it.
+To perform accurately, the matcher would need to apply theory of mind and reason about the director's beliefs.
+For example, if a much larger measuring cup were placed in an occluded slot and agents failed to account for the director's beliefs, they might interpret `the large cup' to mean this occluded cup even though the director couldn't possibly know about it.
 Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.

### 2016-01-21 writing/cogsci-submission/cogsci_keysar.tex
edits to intro

--- 2016-01-21 writing/cogsci-submission/cogsci_keysar.tex
@@
 \newcommand{\jd}[1]{\green{$^*$}\marginpar{\footnotesize{JD: \green{#1}}}}
+\definecolor{Green}{RGB}{10,200,100}
+\newcommand{\ndg}[1]{\textcolor{Green}{[ndg: #1]}}
 
@@
 Among other things, this ability -- called \emph{theory of mind} \cite{PremackWoodruff78_ChimpanzeeToM} -- allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}.
-Most children acquire this ability by age six \cite{WimmerPerner83_BeliefsAboutBeliefs, WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theories \cite{GopnikWellman12_ReconstructingConstructivism}.
-
+Children acquire this ability by at least age six \cite{WimmerPerner83_BeliefsAboutBeliefs, WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
+%
 While theory of mind use often appears to be automatic and effortless, Keysar and colleagues \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind} have argued that it is actually the opposite, even for adults: we are ``mindblind" by default and only overcome our egocentric biases through an effortful process of perspective-taking.
 In other words, while adults are \emph{capable} of applying theory of mind reasoning, we do not always apply it reliably.
-This argument is based on a simple experimental paradigm, where participants played a simple communication game with a confederate.
+In this paper we argue that the apparent failures this view is based on are in fact successes for sophisticated social reasoning. 
+In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners expect these more informative utterances, and interpret the actual utterances accordingly---resulting in ``errors''.
+
+The argument of Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate.
 The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects.
+\ndg{it'd be nice to have a grid illustrating one of the conditions to refer to here...}
 The confederate played the role of `director,' giving instructions about how to move objects around a grid, and the participant played the role of `matcher,' attempting to follow these instructions.
@@
 The director gave an instruction like `move the large cup up one square,' referring to the medium-sized one.
-
 Some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
-To perform accurately, the matcher would need to apply theory of mind and reason about the director?s beliefs.
-For example, if a much larger measuring cup were placed in an occluded slot and agents failed to account for the director?s beliefs, they might interpret `the large cup' to mean this occluded cup even though the director couldn't possibly know about it.
+To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about the director's beliefs.
+For example, imagine a much larger measuring cup were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `the large cup' to mean this occluded cup (which the director couldn't possibly know about).
 Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.
-46\% of participants reached for the hidden object at least twice.
+%46\% of participants reached for the hidden object at least twice. 
 Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.
@@
 Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g.
-the largest measuring cup) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}.
-It would be useful to know whether people referred to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results.
+the largest measuring cup) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}---perhaps people would refer to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results.
 %Note that none of these criticisms invalidate the paradoxical result that listeners move an occluded object, they just suggest an alternative explanation.
@@
 
-In this paper, we offer a novel pragmatic account of Keysar's results.
-When both players know that objects are occluded in the display, we note that the speaker is likely to be overinformative.
-If the listener \emph{expects} the speaker to be overinformative, they will pragmatically pick the \emph{a priori} more likely referent of the referring expression, which in critical trials will be the occluded object.
+In this paper, we offer an additional factor that helps account for Keysar's results. 
+Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations; given an utterance, a listener can then reason backward to the most plausible situation.
+This suggests that we consider whether the utterances used by the confederate in Keysar's critical conditions were actually what a speaker would be expected to say. 
+If not, then perhaps participants are making choices that are in fact consistent with correct pragmatic interpretation on the confederate's (uncooperative) utterance.
+More precisely, when both players know that objects are occluded in the display, the speaker may tend to add additional precision to references in order to avoid confusion.
+If the listener \emph{expects} the speaker to do this, they will pragmatically pick the \emph{a priori} more likely referent of the referring expression, which in critical trials will be the occluded object.
 In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item.
-We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} (in a multi-player web experiment, recruiting participants to be both director and matcher instead of using a confederate.
-Instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study.
-We then ran the same experiment without scripted items, observing unconstrained director behavior.
-This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to theory of mind.
+
+We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment.
+We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study.
+We replicated the original finding, but noted a tendency of directors to be ove-rinformative in unscripted trials.
+We then ran the same experiment without scripting the directions for critical trials, observing unconstrained director behavior.
+We observe much greater precision in unconstrained director utterances and better performance of the matchers.
+This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectatuions about speaker behavior---that is, to theory of mind.
+
+\ndg{intro can probably be compressed some if needed.}
 

### 2016-01-21 writing/cogsci-submission/cogsci_keysar.tex
massage results arrangement for expt 1

@@
 In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item.
+%=======
+%Some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
+To perform accurately, the matcher would need to apply theory of mind and reason about the director's beliefs.
+For example, if a much larger measuring cup were placed in an occluded slot and agents failed to account for the director's beliefs, they might interpret `the large cup' to mean this occluded cup even though the director couldn't possibly know about it.
+Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.
+46\% of participants reached for the hidden object at least twice.
+Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.
+%>>>>>>> 179ee1de2933b596af109cc10f157c31fdcea761
 
@@
 On the left side of the screen, participants could freely type messages to one another; on the right side the screen, players could view a set of objects placed in a 4 x 4 grid.
-Five of the grid cells were occluded from the director's perspective, and the remaining 11 were visible to both participant and director.
+Five of the grid cells were occluded from only the director's perspective, and the remaining 11 were visible to both matcher and director.
 Six or seven objects were displayed in the grid at a given time.
@@
 For half the instructions, directors were free to communicate however they wished.
-For the other half, including the critical instruction, their messages to the matcher were pre-scripted such that we could use the precise wording from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square."
+For the other half, including all the critical instructions, their messages to the matcher were pre-scripted such that we could use the precise wording from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat window, and they would have to click `send' for the experiment to continue.)
 
-We collected baseline performance for each condition by replacing the hidden alternative (e.g.
-a roll of tape) with an object that did not fit the critical instruction (e.g.
-a battery).
-We also used the same unambiguous replacements as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.
+a battery); we used the same unambiguous replacements as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 Each participant received half the items in the experimental condition and half in the baseline condition.
@@
 
-These materials differ from those used by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a few ways.
+These materials differ from those used by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three ways.
 First, participants were not seated across from each other at a table: they each saw a views of the 4 x 4 grid on screen and communicated via a text box.
@@
 
+\ndg{methods and procedures can be merged into one section...}
+
 Participants interacted in a real-time, multi-player environment on the web \cite{Hawkins15_RealTimeWebExperiments}.
@@
 
-\subsection{Results}
+\subsection{Results and discussion}
 
-Using the error rates on critical items, we compare our behavioral results with those from from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in Table \ref{table:mainResults}.
+In Table \ref{table:mainResults} we show the error rates on critical items, and compare to the data from
+\citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 We find that 93\% of participants attempted to move a hidden object at least once in the Experiment condition, out of four possible items, compared to only 14\% in the baseline condition.
 This is similar to the effect observed by the authors in the original study, which found 71\% and 0\%.
-Our errors were larger across the board, perhaps due to the interface and the Turk population, but the gap between the two is roughly the same size.
-
-Next, we examine whether there are differences in error rates across items.
+Our errors were larger across the board, perhaps due to the interface or the population, but the gap between the two conditions is roughly the same size.
+%Next, we examine whether there are differences in error rates across items. 
 In Table \ref{table:exp1ItemWise}, we break down the pattern of errors by item. 
-Because our randomization procedure undersampled the experimental condition for several items, a chi-squared test is inappropriate.
+%Because our randomization procedure undersampled the experimental condition for several items, a chi-squared test is inappropriate.
 Using Fisher's exact test, we cannot reject the null hypothesis of independence ($p = 0.13$).
-Still, we note that several conditions have much higher error rates than others -- for example, 75\% of participants in the experimental condition of item 6 made an error (the ``whiteboard eraser'' vs.
+Still,
+We note that several conditions have much higher error rates than others -- for example, 75\% of participants in the experimental condition of item 6 made an error (the ``whiteboard eraser'' vs.
 the ``pencil eraser'') while only 17\% of participants in item 8 made an error (the ``computer mouse'' vs.
 the ``toy mouse'').
+This suggests that the dependent variable highlighted in the original study (i.e.
+``percentage of participants who moved the critical item at least once'') is somewhat problematic.
+It could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item.
+Indeed, if we exclude the three `hard' items where over 60\% of participants in the experimental condition made errors, this dependent variable drops from 93\% to only 43\% of participants.
+%This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
 
-Finally, as a proxy for the eye-tracking analyses reported by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
+As a proxy for the eye-tracking analyses reported by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
 We define the decision window as the span of time between the point when the matcher received their instruction message and when they started moving an object.
@@
 In baseline trials, people spent much more time in the target cell and much less time in the distractor cell than in the experimental condition (see Fig.
-\ref{fig:exp1hover}).
-Note that this analysis includes participants who actually made errors, since the data is too sparse to exclude them.
-%accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
+\ref{fig:exp1hover}).\footnote{Note that this analysis includes participants who actually made errors, since the data is too sparse to exclude them.} %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
+
+An additional source of data came from running this replication as a multi-player web experiment: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items.
+Informally, we noted a tendency toward over-informativity.
+Instead of ``move the stuffed animal down'', participants say ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right,'' participants say ``move the red airplane to the right,'' even though there is only one plane.
+\ndg{cite work on overinformative ref expr...}
+This suggested that the confederate's scripted instructions were in some sense uncooperative and counter to the matcher's expectations.
+Perhaps directors were taking the time to make more costly over-informative descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment.
+If the matcher reasoned in this way and expected the director to be overinformative, then they would be justified in picking the first object that meets the description rather than worrying about occluded cells.
+We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
+If speakers really are being overinformative, and listeners are correctly expecting this, we should see fewer errors after removing uncooperative scripts.
 
@@
 
-\subsection{Discussion}
-
-In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+%\subsection{Discussion}
+%
+%In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it.
@@
 Both of these results support Keysar's claim that adults do not reliably deploy theory of mind reasoning when it is called for.
-
-Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data.
-First, our analysis of item-level differences shows that the dependent variable highlighted in the original study (i.e.
+%
+%Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data. First, 
+%
+%our analysis of item-level differences shows that the dependent variable highlighted in the original study (i.e.
 ``percentage of participants who moved the critical item at least once'') is somewhat problematic.
@@
 
-Second, because half of the instructions were unscripted, we were able to observe natural production of referential descriptions for a number of fillers.
+%Second, because half of the instructions were unscripted, we were able to observe natural production of referential descriptions for a number of fillers.
 Informally, we noted a tendency toward over-informativity.

### 2016-01-22 writing/cogsci-submission/cogsci_keysar.tex
minor edit
editing transition between expts

@@
 We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
-
 We found that there is an interaction between cell type and condition on the percent of hover time spent in that cell, $b = -0.21, t(11) = -3.1, p = 0.003$.
@@
 
-An additional source of data came from running this replication as a multi-player web experiment: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items.
-Informally, we noted a tendency toward over-informativity.
-Instead of ``move the stuffed animal down'', participants say ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right,'' participants say ``move the red airplane to the right,'' even though there is only one plane.
-\ndg{cite work on overinformative ref expr...}
-This suggested that the confederate's scripted instructions were in some sense uncooperative and counter to the matcher's expectations.
-Perhaps directors were taking the time to make more costly over-informative descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment.
-If the matcher reasoned in this way and expected the director to be overinformative, then they would be justified in picking the first object that meets the description rather than worrying about occluded cells.
+By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items.
+Informally, we noted a tendency toward, possibly unnecessary, precision in descriptions.
+Instead of ``move the stuffed animal down'', participants say ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants say ``move the red airplane to the right.''
+Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions.
+If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells).
+That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative}, and thus led matchers astray. 
 We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
-If speakers really are being overinformative, and listeners are correctly expecting this, we should see fewer errors after removing uncooperative scripts.
+By the reasoning above we expect to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
 
@@
 
+\ndg{mention overinformativity in refexp somewhere in here...}
+
 \section{Acknowledgements}

### 2016-01-23 writing/cogsci-submission/cogsci_keysar.tex

@@
 
-\subsection{Participants:}
+\subsection{Participants}
 
@@
 
-\subsection{Materials:}
+\subsection{Materials \& Procedures}
 
-Pairs of participants -- assigned randomly to `director' and `matcher' roles -- interacted with one another through a real-time web interface, shown in Fig.
+Participants interacted in a real-time, multi-player environment on the web \cite{Hawkins15_RealTimeWebExperiments}. 
+%To deal with participants arriving at unpredictable times, we implemented a waiting room mechanism, where the first person to enter the game environment would see a screen saying ``Waiting for another player..." and the second person would trigger the game to start.
+Pairs of participants---assigned randomly to `director' and `matcher' roles---interacted with one another through a web interface, shown in Fig.
 \ref{fig:interface}.
@@
 We used the same set of targets and occluded alternatives as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, but we were unable to obtain the filler objects from the original experiment and created our own.
+Before entering the game environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface.
+Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
 
@@
 For each item, we gave the director a series of four instructions to move objects around, which were displayed as a series of arrows pointing from some object to an unoccupied cell.
-One of these instructions was a `critical instruction,' which referred to the target object.
+To collect clean mouse-tracking data, we began every instruction by asking the matcher to click a small circle in the center of the grid.
+After this small circle was clicked, the director was allowed to communicate the next instruction and we started recording from the matcher's mouse.
+One of the instructions was a `critical instruction,' which referred to the target object.
 For half the instructions, directors were free to communicate however they wished.
@@
 
-These materials differ from those used by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three ways.
+This paradigm differs from those used by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways.
 First, participants were not seated across from each other at a table: they each saw a views of the 4 x 4 grid on screen and communicated via a text box.
@@
 Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \citeA{KeysarBarr___Brauner00_TakingPerspective}.
+
+%The experiment proceeded through all eight items, with the director providing instructions and the participant moving the objects.
+Except for instructions with scripted messages, the participants were free to use the chat box for questions, small talk, or whatever else.
 
@@
 \end{table*}
-
-\subsection{Procedure}
-
-\ndg{methods and procedures can be merged into one section...}
-
-Participants interacted in a real-time, multi-player environment on the web \cite{Hawkins15_RealTimeWebExperiments}.
-Before entering this environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface.
-Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
-To deal with participants arriving at unpredictable times, we implemented a waiting room mechanism, where the first person to enter the game environment would see a screen saying ``Waiting for another player..." and the second person would trigger the game to start.
-
-To collect clean mouse-tracking data, we began every instruction by asking the matcher to click a small circle in the center of the grid.
-After this small circle was clicked, the director was allowed to communicate the next instruction and we started recording from the matcher's mouse.
-The experiment proceeded through all eight items, with the director providing instructions and the participant moving the objects.
-Except for instructions with scripted messages, the participants were free to use the chat box for questions, small talk, or whatever else.
 

### 2016-01-24 writing/cogsci-submission/cogsci_keysar.tex
Add exp. 2 overinf results and update item-wise table

@@
 \usepackage{dsfont}
+\usepackage{array}
 \usepackage{textcomp}
@@
 
+\begin{figure*}[t!]
+\begin{center}
+\includegraphics[scale = .35]{images/overall_view.jpg}
+\end{center}
+\vspace{-.25cm}
+\caption{Interface used in the reported experiments.
+Objects behind the black squares were hidden from the director.
+\todo[inline]{TODO: Use images from tape condition, add extra panel to show matcher view.}}
+\label{fig:interface}
+\end{figure*}
+
 Humans can accurately and intelligently reason about the mental states of other humans.
@@
 In other words, while adults are \emph{capable} of applying theory of mind reasoning, we do not always apply it reliably.
+
 In this paper we argue that the apparent failures this view is based on are in fact successes for sophisticated social reasoning. 
@@
 The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects.
-\ndg{it'd be nice to have a grid illustrating one of the conditions to refer to here...}
+\todo[inline]{Refer to updated Fig.~1 with both director screen and matcher screen }
 The confederate played the role of `director,' giving instructions about how to move objects around a grid, and the participant played the role of `matcher,' attempting to follow these instructions.
@@
 
-\begin{figure*}
+\begin{table*}
 \begin{center}
-\includegraphics[scale = .35]{images/overall_view.jpg}
+\begin{tabular}{ p{2cm} | r | r |  r || r | r | r || r | r | r}
+& \multicolumn{3}{c||}{\% attempted at least once} & \multicolumn{3}{c||}{\% attempted at least twice} & \multicolumn{3}{c}{\% of total cases}\\
+\hline
+& Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2  \\
+Experimental & 71 & 93 & 60 & 46 & 64 & 36 & 30 & 37 & 26\\
+Baseline & 0 & 14 & 14 & 0 & 0 & 0 & 0 & 3 & 4\\
+\end{tabular}
+\caption{Side-by-side comparison of error rates in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications.
+The first section shows the percentage of \emph{participants} attempting to move the occluded distractor at least one of the four possible cases.
+The second section shows the percentage attempting to move it at least twice.
+The third section shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object.
+\todo[inline]{Big caption maybe unnecessary.
+Anything else to include in this table?
+Maybe Keysar et al, 2000, which reported \% of total cases, or our Exp.
+2 results if you remove the two most difficult items?
+There, errors drop quite a lot.}}
+\label{table:mainResults}
 \end{center}
-\vspace{-.25cm}
-\caption{Interface used in the reported experiments. Objects behind the black squares were hidden from the director.}
-\label{fig:interface}
-\end{figure*}
+\end{table*}
 
@@
 
-We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.
-a battery); we used the same unambiguous replacements as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 Each participant received half the items in the experimental condition and half in the baseline condition.
@@
 Except for instructions with scripted messages, the participants were free to use the chat box for questions, small talk, or whatever else.
+\begin{table*}
+\begin{center}
+\begin{tabular}{ m{1.5cm} | m{1.5cm} | m{1.5cm} |  r | m{1.6cm} | m{1.4cm} | m{1.3cm} | m{1.3cm} | m{1.5cm} |}
+& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
+instruction & ``glasses'' & ``bottom block'' & ``tape'' & ``large measuring cup'' & ``brush'' & ``eraser'' & ``small candle'' & ``mouse'' \\\hline
+target & sunglasses & block (3rd row) & cassette & medium cup & round hairbrush & board eraser & medium candle & computer mouse \\\hline
+hidden distractor & glasses case & block (4th row) & scotch-tape & large cup & flat \,\,\,hairbrush & pencil eraser & small candle & toy mouse \\\hline
+ \# incorrect & 1 & 1 & 2 & 3 & 9 & 7 & 1 & 5 \\
+ \# correct   & 11&14&10&13&7 &8   &12& 8 \\
+\end{tabular}
+\caption{Item-wise error analysis for experiment 2}
+\label{table:exp2ItemWise}
+\end{center}
+\end{table*}
 
@@
 \begin{center}
-\begin{tabular}{ p{2cm} | r | r |  r || r | r | r || r | r | r}
-& \multicolumn{3}{c||}{\% attempted at least once} & \multicolumn{3}{c||}{\% attempted at least twice} & \multicolumn{3}{c}{\% of total cases}\\
-\hline
-& Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2  \\
-Experimental & 71 & 93 & 60 & 46 & 64 & 36 & 30 & 37 & 26\\
-Baseline & 0 & 14 & 14 & 0 & 0 & 0 & 0 & 3 & 4\\
+\begin{tabular}{ p{2cm} | r | r |  r | r | r | r | r | r |}
+& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
+\# incorrect & 1 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
+\# correct & 5 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
 \end{tabular}
-\caption{Side-by-side comparison of error rates in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications.
-The first section shows the percentage of \emph{participants} attempting to move the occluded distractor at least one of the four possible cases.
-The second section shows the percentage attempting to move it at least twice.
-The third section shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object.
-\todo[inline]{Big caption maybe unnecessary.
-Anything else to include in this table?
-Maybe Keysar et al, 2000, which reported \% of total cases, or our Exp.
-2 results if you remove the two most difficult items?
-There, errors drop quite a lot.}}
-\label{table:mainResults}
+\caption{Item-wise error analysis}
+\label{table:exp1ItemWise}
 \end{center}
@@
 
-\begin{table*}
-\begin{center}
-\begin{tabular}{ p{2cm} | r | r |  r | r | r | r | r | r |}
-& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
-\# incorrect & 1 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
-\# correct & 5 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
-\end{tabular}
-\caption{Item-wise error analysis}
-\label{table:exp1ItemWise}
-\end{center}
-\end{table*}
-
-\begin{figure}[b!]
-\begin{center}
-\includegraphics[scale=.8]{images/exp1MouseTracking.pdf}
-\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
-\label{fig:exp1hover}
-\end{center}
-\end{figure}
-
 %\subsection{Discussion}
@@
 If speakers really are being overinformative, and listeners are correctly expecting this, we should see fewer errors after removing uncooperative scripts.
-
-\begin{table*}
-\begin{center}
-\begin{tabular}{ p{2cm} | r | r |  r | r | r | r | r | r |}
-& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
- \# incorrect & 1 & 1 & 2 & 3 & 9 & 7 & 1 & 5 \\
- \# correct   & 11&14&10&13&7 &8   &12& 8 \\
-\end{tabular}
-\caption{Item-wise error analysis for experiment 2}
-\label{table:exp2ItemWise}
-\end{center}
-\end{table*}
-
-\begin{figure}[b!]
-\begin{center}
-\includegraphics[scale=.9]{images/exp2MouseTracking.pdf}
-\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
-\label{fig:exp2hover}
-\end{center}
-\end{figure}
-\section{Exp.~2: Unscripted Replication}
-\label{sec:Exp2}
 
@@
 
-\todo[inline]{rdh: should do a basic analysis of over-informativity...
-Don't forget about the person who always gave directions like ``start at the third square of the first row and move down one square''}
+Next, we test whether these reductions in error rate and hover-time interference are in fact due to different speaker behavior.
+Two coders were given the list of messages that directors sent to convey instructions for critical items, blind to condition.
+The coders determined (1) whether or not the label used by the speaker fit the hidden distractor and (2) whether or not the label used by the director was more or less precise than the scripted label used in Exp.~1.
+Inter-rater reliability was relatively high, with $r=0.82$ for distractor fit and $r=0.69$ for the precision comparison.
+Binary rater decisions were averaged to form a single distractor fit score for each message, and a single precision score.
 
+We found that the referring expressions speakers used to convey critical instructions only fit the distractor 53\% of the time, compared to the scripted message in Exp.~1, which fit the distractor 100\% of the time.
+Furthermore, 47\% of these referring expressions were rated to be more precise than the corresponding scripted instruction used in Exp.\~1.
+As expected, there's a strong negative relationship between the precision of the label and whether it might apply to the hidden distractor, $r = -0.73$.
+Finally, we test whether the speaker's choice of label accounts for variability in error rates across different items.
+We find that error rates are higher when the speaker's label fits the hidden distractor $b = 0.79, t(5) = 2.8, p = 0.36$ \todo[inline]{rdh: this isn't the right regression...
+I did it on the aggregated scores, and it should really be done on the raw subject-by-subject data.
+It's just a bit tricky to join the error data and the message data.}
+
+\begin{figure}[b!]
+\begin{center}
+\includegraphics[scale=.75]{images/exp2MouseTracking.pdf}
+\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
+\label{fig:exp2hover}
+\end{center}
+\end{figure}
+\section{Exp.~2: Unscripted Replication}
+\label{sec:Exp2}
+
+\begin{figure}[b!]
+\begin{center}
+\includegraphics[scale=.75]{images/exp1MouseTracking.pdf}
+\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
+\label{fig:exp1hover}
+\end{center}
+\end{figure}
 \subsection{Discussion}

### 2016-01-24 writing/cogsci-submission/cogsci_keysar.tex
comments on exp2 analysis

@@
 
+\section{Exp.~2: Unscripted Replication}
+\label{sec:Exp2}
+
 \subsection{Participants}
@@
 Inter-rater reliability was relatively high, with $r=0.82$ for distractor fit and $r=0.69$ for the precision comparison.
-Binary rater decisions were averaged to form a single distractor fit score for each message, and a single precision score.
+\ndg{r is not right measure of agreement for binary coding -- \%agreement or kappa?} Binary rater decisions were averaged to form a single distractor fit score for each message, and a single precision score.
 
 We found that the referring expressions speakers used to convey critical instructions only fit the distractor 53\% of the time, compared to the scripted message in Exp.~1, which fit the distractor 100\% of the time.
-Furthermore, 47\% of these referring expressions were rated to be more precise than the corresponding scripted instruction used in Exp.\~1.
-As expected, there's a strong negative relationship between the precision of the label and whether it might apply to the hidden distractor, $r = -0.73$.
+Furthermore, 47\% of these referring expressions were rated to be more precise than the corresponding scripted instruction used in Exp.~1.
+\ndg{was remaining 53\% rated as scripted more precise or mostly `can't tell'?} As expected, there's a strong negative relationship between the precision of the label and whether it might apply to the hidden distractor, $r = -0.73$.
 Finally, we test whether the speaker's choice of label accounts for variability in error rates across different items.
@@
 It's just a bit tricky to join the error data and the message data.}
+\ndg{also it's not clear what that analysis is, or how it tells us about the item effects...}
 
@@
 \end{figure}
-\section{Exp.~2: Unscripted Replication}
-\label{sec:Exp2}
 

### 2016-01-25 experiment2/analysis/exp2Analysis.Rmd
Compare error rates across the two experiments; fix some bugs

--- 2015-03-29 analysis/keysarMouseTracking.Rmd
@@
 ```{r}
-setwd('~/Repos/psych254/replication_project/analysis/')
-mouseD = read.csv('fullSampleMouse.csv')
-errorD = read.csv('fullSampleError.csv')
-messageD = read.csv('fullSampleMessage.csv')
-str(mouseD)
+setwd('~/Repos/keysar_replication/experiment2/data/')
+preMouseDExp2 = read.csv('mouse/exp2Mouse.csv')
+preErrorDExp2 = read.csv('error/exp2Error.csv')
+preMessageDExp2 = read.csv('message/exp2Message.csv')
 library(tidyr)
@@
 library(ggplot2)
+library(lme4)
+library(lmerTest)
 ```
@@
 ```{r}
-nonCriticalMistakes <- errorD %>%
+nonNativeEnglishIDs <- c("0366-3", "8676-f")
+nonCriticalMistakes <- preErrorDExp2 %>%
   # only look at mistakes on noncritical (filler) items
@@
 
-bannedIDs <- nonCriticalMistakes$gameid
+mistakeIDs <- as.character(nonCriticalMistakes$gameid)
+bannedIDsExp2 <- unique(c(mistakeIDs, nonNativeEnglishIDs))
+numParticipantsExp2 <- (length(unique(preMouseDExp2$gameid)) -
+                        length(bannedIDsExp2))
+```
+
+Remove banned ids from data
+
+```{r}
+errorDExp2 <- preErrorDExp2 %>%
+  filter(!(gameid %in% bannedIDsExp2))
+mouseDExp2 <- preMouseDExp2 %>%
+  filter(!(gameid %in% bannedIDsExp2))
+messageDExp2 <- preMessageDExp2 %>%
+  filter(!(gameid %in% bannedIDsExp2))
 ```
@@
 ```{r}
-eD <- errorD %>%
-  filter(!(gameid %in% bannedIDs)) %>%
-  filter(critical == 1) %>%
+eDExp2 <- errorDExp2 %>%
   filter(attemptNum == 0) %>%
   group_by(gameid, condition) %>%
-  tally() %>%
-  mutate(cond2 = as.numeric(n >= 2)) %>%
+  summarize(n = sum(criticalError)) %>%
+  mutate(cond1 = as.numeric(n >= 1)) %>%
+  mutate(cond2 = as.numeric(n >= 2)) #
+
+errorTableExp2 <- eDExp2 %>%
   group_by(condition) %>%
-  summarise(atLeastOnce = length(n)/14, 
-            atLeastTwice = sum(cond2)/14)
-eD
+  summarise(atLeastOnce = sum(cond1)/numParticipantsExp2, 
+            atLeastTwice = sum(cond2)/numParticipantsExp2,
+            total = sum(n) / (4 * numParticipantsExp2))
+errorTableExp2
 
@@
 # where more than 60% of participants made errors.
-excludedED <- errorD %>%
-  filter(!(gameid %in% bannedIDs)) %>%
-  filter(critical == 1) %>%
-  filter(!(objectSet %in% c(2, 6, 7))) %>%
+excludedEDExp2 <- errorDExp2 %>%
   filter(attemptNum == 0) %>%
+  filter(!(objectSet %in% c(5, 6))) %>% # Just 5 & 6?
   group_by(gameid, condition) %>%
-  tally() %>%
+  summarize(n = sum(criticalError)) %>%
+  mutate(cond1 = as.numeric(n >= 1)) %>%
   mutate(cond2 = as.numeric(n >= 2)) %>%
   group_by(condition) %>%
-  summarise(atLeastOnce = length(n)/14, 
-            atLeastTwice = sum(cond2)/14)
-excludedED
+  summarise(atLeastOnce = sum(cond1)/numParticipantsExp2, 
+            atLeastTwice = sum(cond2)/numParticipantsExp2,
+            total = sum(n) / (4 * numParticipantsExp2))
+excludedEDExp2
 ```
@@
 ```{r}
-numPlayersPerCondition = mouseD %>% 
-  filter(!(gameid %in% bannedIDs)) %>% 
+numPlayersPerConditionExp2 = mouseDExp2 %>% 
   filter(critical == 1) %>% 
@@
 ```{r}
-itemInequalities <- errorD %>%
-  filter(!(gameid %in% bannedIDs)) %>%
-  filter(critical == 1) %>%
+itemInequalitiesExp2 <- errorDExp2 %>%
+  filter(condition == "exp") %>%
+  filter(criticalError == 1) %>%
   filter(attemptNum == 0) %>%
-  filter(condition == "exp") %>%
   group_by(objectSet) %>%
   tally() %>%
-  right_join(numPlayersPerCondition, by = c('objectSet')) %>%
-  transmute(errorCount = n, correctCount = total - n)
+  right_join(numPlayersPerConditionExp2, by = c('objectSet')) %>%
+  mutate(objectSet = objectSet,
+         errorCount = n, 
+         correctCount = total - n) %>%
+  mutate(errorRate = errorCount / (errorCount + correctCount)) %>%
+  select(objectSet, errorCount, correctCount, errorRate, total)
 
-t(itemInequalities)
-t <- chisq.test(itemInequalities)
-print(t)
-t$expected
-t$observed
+itemInequalitiesExp2[1,2:4] = c(0, itemInequalitiesExp2[1,]$total, 0)
+
+itemWiseTestExp2 <- chisq.test(itemInequalitiesExp2[-c(1,4,5)])
+print(itemWiseTestExp2)
 ```
 
-Set up mouse/message data
+Follow-up:
+
+Did error rates go down significantly?
 
 ```{r}
+exp1NumErrors = eDExp1 %>% 
+  filter(condition == "exp") %>% 
+  group_by(condition) %>%
+  summarize(numErrors = sum(n))
+exp2NumErrors = eDExp2 %>% 
+  filter(condition == "exp") %>% 
+  group_by(condition) %>%
+  summarize(numErrors = sum(n))
 
-mouData <- mouseD %>%
-  filter(!(gameid %in% bannedIDs)) %>%
+exp1TotalPossible = 4 * numParticipantsExp1
+exp2TotalPossible = 4 * numParticipantsExp2
+
+numErrors = c(exp1NumErrors$numErrors, exp2NumErrors$numErrors)
+totalPossible = c(exp1TotalPossible, exp2TotalPossible)
+prop.test(numErrors, totalPossible)
+
+# Diff from keysar (2003)?
+keysarNumErrors = 45
+keysarTotalPossible = 4 * 38
+numErrors = c(keysarNumErrors, exp2NumErrors$numErrors)
+totalPossible = c(keysarTotalPossible, exp2TotalPossible)
+prop.test(numErrors, totalPossible)
+```
+
+Set up message data
+-------------------
+
+Need to join the coder ratings...
+
+```{r}
+messageCoding1 <- read.csv("../analysis/messageCoding1.csv")
+messageCoding2 <- read.csv("../analysis/messageCoding2.csv")
+relevantProperties <- c("gameid", "attemptNum", "condition", "critical",
+                        "objectSet", "sender", "instructionNum")
+overinformativeCodes <- messageCoding1 %>% 
+  left_join(messageCoding2, by = relevantProperties) %>%
+  filter(!(gameid %in% bannedIDsExp2)) %>%
+  mutate(contents = contents.x, time = time.x) %>%
+  select(gameid, attemptNum, condition, critical, objectSet, sender, 
+         instructionNum, contents, time, targetCode1, distractCode1,
+         altCode1, overCode1, targetCode2, distractCode2, altCode2,
+         overCode2) %>%
+  mutate(avgAltCode = (altCode1 + altCode2)/2,
+         avgTargetCode = (targetCode1 + targetCode2)/2,
+         avgDistractCode = (distractCode1 + distractCode2)/2,
+         avgOverCode = (overCode1 + overCode2)/2)
+
+# How often are the descriptions more informative than orig?
+mean(overinformativeCodes$avgOverCode)
+
+# How often do the descriptions apply to the hidden object?
+mean(overinformativeCodes$avgDistractCode)
+
+# Does overinformativity perfectly avoid picking distractor?
+cor(overinformativeCodes$avgDistractCode, overinformativeCodes$avgOverCode)
+
+# Inter-rater reliability
+kappa2(cbind(overinformativeCodes$distractCode1, overinformativeCodes$distractCode2))
+kappa2(cbind(overinformativeCodes$overCode1, overinformativeCodes$overCode2))
+```
+
+Want to show whether precision helps avoid errors (mixed model?)
+
+First, we join the error data with the critical item coding data.
+
+```{r}
+criticalErrors = errorDExp2 %>% 
+  filter(attemptNum == 0)%>% 
+  filter(criticalError == 1) 
+itemWiseCodes <- overinformativeCodes %>% 
+  left_join(criticalErrors, by = c("gameid", "condition", "critical", 
+                                   "objectSet", "instructionNum",
+                                   "attemptNum")) %>%
+  select(gameid, objectSet, avgDistractCode, avgOverCode, 
+         criticalError) %>%
+  mutate(criticalError = ifelse(is.na(criticalError),0,1))
+
+mmRes <- glmer(criticalError ~ avgOverCode + avgDistractCode +  (1 | objectSet),
+      data = itemWiseCodes, family = binomial)
+
+summary(mmRes)
+```
+
+Set up mouse data
+
+```{r}
+mouDataExp2 <- mouseDExp2 %>%
   filter(attemptNum == 0)
 
-mD <- messageD %>% 
-  filter(!(gameid %in% bannedIDs)) %>%
+mDExp2 <- messageDExp2 %>% 
   filter(attemptNum == 0) %>%
@@
   mutate(messageStamp = first(time)) %>%
-  select(gameid, condition, attemptNum, instructionNum, critical, objectSet,sender,contents,messageStamp) %>%
+  select(gameid, condition, attemptNum, instructionNum, critical, 
+         objectSet,sender,contents,messageStamp) %>%
   distinct(gameid,contents)
 
-joined <- (right_join(mouData, mD, by = c("gameid", "condition", "objectSet", 
+joinedExp2 <- (right_join(mouDataExp2, mDExp2, by = c("gameid", "condition",
+                                                      "objectSet", 
                                          "instructionNum", "attemptNum", "critical")))
@@
 ```{r}
-dUnfilt <- joined %>%
+dUnfiltExp2 <- joinedExp2 %>%
   filter(critical == 1) %>%
@@
 
-d <- joined %>% 
+dExp2 <- joinedExp2 %>% 
   filter(critical == 1) %>% 
   group_by(objectSet, condition) %>%
-#  filter(time > messageStamp) #%>%
+  filter(time > messageStamp) %>%
   mutate(mouseY = 600 - mouseY) %>%
@@
 
-heatMapForLM <- d %>%
+heatMapForLMExp2 <- dExp2 %>%
   group_by(gameid, objectSet) %>%
@@
 
-heatMapForLM
+heatMapForLMExp2
 
-t1 <- lm(percent ~ condition * cellType, data = heatMapForLM)
+t1 <- lm(percent ~ condition * cellType, data = heatMapForLMExp2)
 summary(t1)
@@
 ```{r}
-heatMapForPlot <- d %>%
+heatMapForPlotExp2 <- dExp2 %>%
   group_by(gameid, objectSet) %>%
@@
 
-heatMapForPlot
+heatMapForPlotExp2
 
-g <- (ggplot(heatMapForPlot, aes(x = condition, y = percent, group = cellType,
-                                  color = cellType)) 
+g <- (ggplot(heatMapForPlotExp2, aes(x = condition, y = percent, 
+                                     group = cellType, color = cellType)) 
       + geom_line(aes(linetype = cellType), size = 2)
@@
                           ymin = percent - error), size =2 ,
-                      width = 0.1))
+                      width = 0.1)
+     + ylim(0, .45)
+      + ggtitle("Experiment 2 Hover-Time"))
 g
+ggsave("../../writing/cogsci-submission/images/exp2MouseTracking.pdf")
 ```
 
-Very noisy (unfinished) plot of all mouse-movements
+Compare mouse data from Exp1 w/ mouse data from Exp2:
+
+```{r}
+combinedHeatMap = rbind(heatMapForLMExp1 %>%
+                          mutate(exp = "Exp.1") %>% 
+                          filter(condition == "exp"),
+                        heatMapForLMExp2 %>% 
+                          mutate(exp = "Exp.2") %>%
+                          filter(condition == "exp"))
+combinedHeatMapForPlot <- combinedHeatMap %>%
+  group_by(cellType, exp) %>%
+  summarize(error = sem(percent),
+            percent = mean(percent))
+
+g <- (ggplot(combinedHeatMapForPlot, aes(x = cellType, y = percent, 
+                                  group = exp, color = exp)) 
+      + geom_line(aes(linetype = exp), size = 2)
+      + geom_errorbar(aes(ymax = percent + error, 
+                          ymin = percent - error), size =2 ,
+                          width = 0.1)
+      + ylim(0, .45)
+      + ggtitle("Experiment 2 Hover-Time"))
+g
+combinedRes = lm(percent ~ exp * cellType, data = combinedHeatMap)
+summary(combinedRes)
+```
 

### 2016-01-25 experiment1/analysis/exp1Analysis.Rmd
Fix some bugs in exp 1 analysis

--- 2015-03-29 analysis/keysarMouseTracking.Rmd
@@
 ```{r}
-setwd('~/Repos/psych254/replication_project/analysis/')
-mouseD = read.csv('fullSampleMouse.csv')
-errorD = read.csv('fullSampleError.csv')
-messageD = read.csv('fullSampleMessage.csv')
-str(mouseD)
+setwd('~/Repos/keysar_replication/experiment1/analysis/')
+preMouseDExp1 = read.csv('fullSampleMouse.csv')
+preErrorDExp1 = read.csv('fullSampleError.csv')
+preMessageDExp1 = read.csv('fullSampleMessage.csv')
 library(tidyr)
@@
 ```{r}
-nonCriticalMistakes <- errorD %>%
+nonCriticalMistakes <- preErrorDExp1 %>%
   # only look at mistakes on noncritical (filler) items
@@
 
-bannedIDs <- nonCriticalMistakes$gameid
+bannedIDsExp1 <- nonCriticalMistakes$gameid
+numParticipantsExp1 <- (length(unique(preMouseDExp1$gameid)) -
+                        length(bannedIDsExp1))
+
+```
+
+Remove banned ids from data
+
+```{r}
+errorDExp1 <- preErrorDExp1 %>%
+  filter(!(gameid %in% bannedIDsExp1))
+mouseDExp1 <- preMouseDExp1 %>%
+  filter(!(gameid %in% bannedIDsExp1))
+messageDExp1 <- preMessageDExp1 %>%
+  filter(!(gameid %in% bannedIDsExp1))
 ```
@@
 ```{r}
-eD <- errorD %>%
-  filter(!(gameid %in% bannedIDs)) %>%
-  filter(critical == 1) %>%
+eDExp1 <- errorDExp1 %>%
   filter(attemptNum == 0) %>%
   group_by(gameid, condition) %>%
-  tally() %>%
-  mutate(cond2 = as.numeric(n >= 2)) %>%
+  summarize(n = sum(criticalError)) %>%
+  mutate(cond1 = as.numeric(n >= 1)) %>%
+  mutate(cond2 = as.numeric(n >= 2))
+
+errorTableExp1 <- eDExp1 %>% 
   group_by(condition) %>%
-  summarise(atLeastOnce = length(n)/14, 
-            atLeastTwice = sum(cond2)/14)
-eD
+  summarise(atLeastOnce = sum(cond1)/numParticipantsExp1, 
+            atLeastTwice = sum(cond2)/numParticipantsExp1,
+            total = sum(n) / (4 * numParticipantsExp1))
+errorTableExp1
 
@@
 # where more than 60% of participants made errors.
-excludedED <- errorD %>%
-  filter(!(gameid %in% bannedIDs)) %>%
-  filter(critical == 1) %>%
-  filter(!(objectSet %in% c(2, 6, 7))) %>%
+excludedEDExp1 <- errorDExp1 %>%
   filter(attemptNum == 0) %>%
+  filter(!(objectSet %in% c(2, 6, 7))) %>% # Just 5 & 6?
   group_by(gameid, condition) %>%
-  tally() %>%
+  summarize(n = sum(criticalError)) %>%
+  mutate(cond1 = as.numeric(n >= 1)) %>%
   mutate(cond2 = as.numeric(n >= 2)) %>%
   group_by(condition) %>%
-  summarise(atLeastOnce = length(n)/14, 
-            atLeastTwice = sum(cond2)/14)
-excludedED
+  summarise(atLeastOnce = sum(cond1)/numParticipantsExp1, 
+            atLeastTwice = sum(cond2)/numParticipantsExp1,
+            total = sum(n) / (4 * numParticipantsExp1))
+excludedEDExp1
 ```
@@
 ```{r}
-numPlayersPerCondition = mouseD %>% 
-  filter(!(gameid %in% bannedIDs)) %>% 
+numPlayersPerConditionExp1 = mouseDExp1 %>% 
   filter(critical == 1) %>% 
@@
 ```{r}
-itemInequalities <- errorD %>%
-  filter(!(gameid %in% bannedIDs)) %>%
-  filter(critical == 1) %>%
+itemInequalitiesExp1 <- errorDExp1 %>%
+  filter(condition == "exp") %>%
+  filter(criticalError == 1) %>%
   filter(attemptNum == 0) %>%
-  filter(condition == "exp") %>%
   group_by(objectSet) %>%
   tally() %>%
-  right_join(numPlayersPerCondition, by = c('objectSet')) %>%
-  transmute(errorCount = n, correctCount = total - n)
+  right_join(numPlayersPerConditionExp1, by = c('objectSet')) %>%
+  mutate(objectSet = objectSet,
+         errorCount = n, 
+         correctCount = total - n) %>%
+  mutate(errorRate = errorCount / (errorCount + correctCount)) %>%
+  select(objectSet, errorCount, correctCount, errorRate, total)
+itemInequalitiesExp1[1,2:4] = c(0, itemInequalitiesExp1[1,]$total, 0)
 
-t(itemInequalities)
-t <- chisq.test(itemInequalities)
-print(t)
-t$expected
-t$observed
+itemWiseTestExp1 <- chisq.test(itemInequalitiesExp1[-c(1,4,5)])
+print(itemWiseTestExp1)
 ```
@@
 ```{r}
-
-mouData <- mouseD %>%
-  filter(!(gameid %in% bannedIDs)) %>%
+mouDataExp1 <- mouseDExp1 %>%
   filter(attemptNum == 0)
 
-mD <- messageD %>% 
-  filter(!(gameid %in% bannedIDs)) %>%
+mDExp1 <- messageDExp1 %>% 
   filter(attemptNum == 0) %>%
@@
 
-joined <- (right_join(mouData, mD, by = c("gameid", "condition", "objectSet", 
+joinedExp1 <- (right_join(mouDataExp1, mDExp1, by = c("gameid", "condition", "objectSet", 
                                          "instructionNum", "attemptNum", "critical")))
@@
 ```{r}
-dUnfilt <- joined %>%
+dUnfiltExp1 <- joinedExp1 %>%
   filter(critical == 1) %>%
@@
 
-d <- joined %>% 
+dExp1 <- joinedExp1 %>% 
   filter(critical == 1) %>% 
   group_by(objectSet, condition) %>%
-#  filter(time > messageStamp) #%>%
+  filter(time > messageStamp) %>%
   mutate(mouseY = 600 - mouseY) %>%
@@
 
-heatMapForLM <- d %>%
+heatMapForLMExp1 <- dExp1 %>%
   group_by(gameid, objectSet) %>%
@@
 
-heatMapForLM
+heatMapForLMExp1
 
-t1 <- lm(percent ~ condition * cellType, data = heatMapForLM)
+t1 <- lm(percent ~ condition * cellType, data = heatMapForLMExp1)
 summary(t1)
@@
 ```{r}
-heatMapForPlot <- d %>%
+heatMapForPlotExp1 <- dExp1 %>%
   group_by(gameid, objectSet) %>%
@@
 
-heatMapForPlot
+heatMapForPlotExp1
 
-g <- (ggplot(heatMapForPlot, aes(x = condition, y = percent, group = cellType,
+g <- (ggplot(heatMapForPlotExp1, aes(x = condition, y = percent, group = cellType,
                                   color = cellType)) 
@@
                           ymin = percent - error), size =2 ,
-                      width = 0.1))
+                      width = 0.1)
+      + ylim(0,.45)
+      + ggtitle("Experiment 1 Hover-Time"))
 g
+ggsave("../../writing/cogsci-submission/images/exp1MouseTracking.pdf")
 ```

### 2016-01-25 writing/cogsci-submission/cogsci_keysar.tex
Add more rigorous tests, fix table, add discussion

--- 2016-01-24 writing/cogsci-submission/cogsci_keysar.tex
@@
 \usepackage{textcomp}
+\usepackage{multirow}
 
@@
 & Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2  \\
-Experimental & 71 & 93 & 60 & 46 & 64 & 36 & 30 & 37 & 26\\
-Baseline & 0 & 14 & 14 & 0 & 0 & 0 & 0 & 3 & 4\\
+Experimental & 71 & 93 & 61 & 46 & 57 & 32 & 30 & 43 & 24\\
+Baseline        & 0   & 7   & 0   & 0   & 0   & 0   & 0   & 2   & 0\\
 \end{tabular}
@@
 \begin{center}
-\begin{tabular}{ m{1.5cm} | m{1.5cm} | m{1.5cm} |  r | m{1.6cm} | m{1.4cm} | m{1.3cm} | m{1.3cm} | m{1.5cm} |}
-& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
-instruction & ``glasses'' & ``bottom block'' & ``tape'' & ``large measuring cup'' & ``brush'' & ``eraser'' & ``small candle'' & ``mouse'' \\\hline
-target & sunglasses & block (3rd row) & cassette & medium cup & round hairbrush & board eraser & medium candle & computer mouse \\\hline
-hidden distractor & glasses case & block (4th row) & scotch-tape & large cup & flat \,\,\,hairbrush & pencil eraser & small candle & toy mouse \\\hline
- \# incorrect & 1 & 1 & 2 & 3 & 9 & 7 & 1 & 5 \\
- \# correct   & 11&14&10&13&7 &8   &12& 8 \\
+\begin{tabular}{ m{.75cm} | m{1.5cm} | m{1.5cm} | m{1.5cm} |  m{1.25cm} | m{1.6cm} | m{1.4cm} | m{1.3cm} | m{1.3cm} | m{1.5cm} |}
+& & Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
+& instruction & ``glasses'' & ``bottom block'' & ``tape'' & ``large measuring cup'' & ``brush'' & ``eraser'' & ``small candle'' & ``mouse'' \\\hline
+& target & sunglasses & block (3rd row) & cassette & medium cup & round hairbrush & board eraser & medium candle & computer mouse \\\hline
+& hidden distractor & glasses case & block (4th row) & scotch-tape & large cup & flat \,\,\,hairbrush & pencil eraser & small candle & toy mouse \\
+ \hline\hline
+\multirow{2}{.75cm}{Exp.~1}  & \# incorrect & 0 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
+                                               & \# correct & 6 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
+ \hline\hline
+\multirow{2}{.75cm}{Exp.~2} & \# incorrect & 0 & 1 & 1 & 3 & 9 & 7 & 1 & 5 \\
+                                              & \# correct   & 12&14&11&13&7 &8   &12& 8 \\
+
 \end{tabular}
-\caption{Item-wise error analysis for experiment 2}
-\label{table:exp2ItemWise}
-\end{center}
-\end{table*}
-
-\begin{table*}
-\begin{center}
-\begin{tabular}{ p{2cm} | r | r |  r | r | r | r | r | r |}
-& Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
-\# incorrect & 1 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
-\# correct & 5 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
-\end{tabular}
-\caption{Item-wise error analysis}
-\label{table:exp1ItemWise}
+\caption{Item-wise error rates}
+\label{table:ItemWise}
 \end{center}
@@
 \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-We find that 93\% of participants attempted to move a hidden object at least once in the Experiment condition, out of four possible items, compared to only 14\% in the baseline condition.
+We find that 93\% of participants (all but one) attempted to move the hidden distractor at least once in the Experimental condition, out of four possible items, compared to only 7\% (only one) in the baseline condition.
 This is similar to the effect observed by the authors in the original study, which found 71\% and 0\%.
@@
 %Next, we examine whether there are differences in error rates across items. 
-In Table \ref{table:exp1ItemWise}, we break down the pattern of errors by item. 
+In Table \ref{table:ItemWise}, we break down the pattern of errors by item. 
 %Because our randomization procedure undersampled the experimental condition for several items, a chi-squared test is inappropriate.
@@
 
-Error rates are reported in Table \ref{table:mainResults}, alongside the results from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our direct replication in Exp.~1.
-Note that error rates for baseline trials are the same as Exp.~1, confirming that participants tend to make more errors in the web version of the experiment.
-Error rates for experimental trials, however, are lower than the rates reported in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and in Exp.~1 \todo[inline]{rdh: need to run a comparison of estimated binomial proportions here}.
+Error rates are reported in Table \ref{table:mainResults}, alongside the results from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our scripted replication in Exp.~1.
+Participants never moved the hidden object in the baseline condition, and total error rates for experimental trials are significantly lower than the rates found in Exp.~1, $\chi^2(1) = 5.35, p = 0.02$.
 
-More dramatically, we find that patterns of errors in Exp.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 20.1, p = 0.005$.
+More dramatically, we find that patterns of errors in Exp.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
 Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the costly over-informative utterance remains ambiguous.
@@
 ``Hair brush'' is already a label at the sub-class level.
-If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 26\% to 12\% and the percentage of participants making at least one error drops from 60\% to 36\%.
-
-\todo[inline]{rdh: should merge item-wise analyses for Exp. 1 and Exp. 2 into a single table, and actually write out what the targets and distractors are for each items}
+If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 26\% to 10\% and the percentage of participants making at least one error drops from 60\% to 32\%.
 
@@
 
-\todo[inline]{rdh: make this analysis more rigorous by explicitly testing the size of the interaction?
-By testing the diff.
-in hover-time in target vs.
-distractor cells after restricting to experimental trials?}
+\todo[inline]{rdh: I don't think we have enough power to show a difference between mean hover-times in the two experiments -- the first experiment just didn't have a big enough sample (the psych254 class budget was a pretty hard constraint).
+Might be able to get more power by doing mixed model like below, or by calculating the times more precisely (currently, just count number of mouse movement events in time window, which probably adds a lot of noise).}
 
@@
 The coders determined (1) whether or not the label used by the speaker fit the hidden distractor and (2) whether or not the label used by the director was more or less precise than the scripted label used in Exp.~1.
-Inter-rater reliability was relatively high, with $r=0.82$ for distractor fit and $r=0.69$ for the precision comparison.
+Inter-rater reliability was relatively high, with Cohen's $\kappa=0.82$ for distractor fit and $\kappa=0.67$ for the precision comparison.
 Binary rater decisions were averaged to form a single distractor fit score for each message, and a single precision score.
 
-We found that the referring expressions speakers used to convey critical instructions only fit the distractor 53\% of the time, compared to the scripted message in Exp.~1, which fit the distractor 100\% of the time.
-Furthermore, 47\% of these referring expressions were rated to be more precise than the corresponding scripted instruction used in Exp.\~1.
-As expected, there's a strong negative relationship between the precision of the label and whether it might apply to the hidden distractor, $r = -0.73$.
-Finally, we test whether the speaker's choice of label accounts for variability in error rates across different items.
-We find that error rates are higher when the speaker's label fits the hidden distractor $b = 0.79, t(5) = 2.8, p = 0.36$ \todo[inline]{rdh: this isn't the right regression...
-I did it on the aggregated scores, and it should really be done on the raw subject-by-subject data.
-It's just a bit tricky to join the error data and the message data.}
+We found that the referring expressions speakers used to convey critical instructions only fit the distractor 53\% of the time, compared to the scripted message in Exp.~1, which (by definition) fit the distractor 100\% of the time.
+\todo[inline]{Except that `glasses' one, which doesn't really apply to a glasses case.
+I'm now pretty sure that this is the fault of the picture I picked (an \emph{open} glasses case rather than a closed one\dots which was dumb in retrospect) Maybe I should re-run analyses without that item, just in case?}
+Furthermore, 47\% of these referring expressions were rated to be more precise than the corresponding scripted instruction used in Exp.~1.
+\ndg{was remaining 53\% rated as scripted more precise or mostly `can't tell'?} As expected, there's a strong negative relationship between the precision of the label and whether it might apply to the hidden distractor, $r = -0.73$.
+Finally, we used a mixed effects logistic regression model to estimate the effect of distractor fit and precision on the probability of making an error on critical trials.
+We included random intercepts for the eight items.
+We find that participants are significantly more likely to make an error when the label applies to the distractor, $b = 3.15, z = 2.93, p = 0.003$, but that precision has no effect, $p = 0.96$.
+
+\begin{figure}[b!]
+\begin{center}
+\includegraphics[scale=.75]{images/exp1MouseTracking.pdf}
+\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
+\label{fig:exp1hover}
+\end{center}
+\end{figure}
 
@@
 \end{figure}
-\section{Exp.~2: Unscripted Replication}
-\label{sec:Exp2}
 
-\begin{figure}[b!]
-\begin{center}
-\includegraphics[scale=.75]{images/exp1MouseTracking.pdf}
-\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
-\label{fig:exp1hover}
-\end{center}
-\end{figure}
-\subsection{Discussion}
-
-\begin{itemize}
-\item people still try to move occluded objects, but they do it less, and that behavior is more confined to a few particularly difficult items.
-\item this happens because the scripted instructions we forced directors to use in Exp.~1 were unnatural and violated listener expectations of overinformativity
-\item the fact that the effect is attenuated in exp.
+%\begin{itemize}
+%\item people still try to move occluded objects, but they do it less, and that behavior is more confined to a few particularly difficult items.
+%\item this happens because the scripted instructions we forced directors to use in Exp.~1 were unnatural and violated listener expectations of overinformativity
+%\item the fact that the effect is attenuated in exp.
 2 is actually evidence of some theory of mind use (insofar as the director choosing to be overinformative, and the matcher expecting overinformativity constitutes ToM
-\end{itemize}
+%\end{itemize}
 
 \section{General Discussion}
+
+Removing constraints on the speaker produced a dramatic reduction in overall matcher errors and systematic variation in error rates across items.
+How did unscripted speaker behavior differ from scripted behavior such that these effects emerged?
+We found that unscripted speakers systematically tended to produce more precise, costly referring expressions than their scripted counterparts -- for example, ``the clear audiocassette'' when there was only one ``tape'' in the speaker's display.
+These precise labels were less likely to fit the hidden distractor, and therefore less likely to lead to a matcher error.
+
+A key point
 
@@
 
-Second, it's worth thinking about some ways our online version may differ from the original.
+% Second, it's worth thinking about some ways our online version may differ from the original.
 The fact that we didn't use a confederate didn't seem to be a problem, given our use of randomly scripted instructions.
@@
 
-\todo[inline]{rdh: still have some worries about whether people really understood the point of the black squares, even after the quiz (e.g.
-one person said ``This seems to work wether the space is black or not.'' Note that this would just inflate our errors, so our analysis here is conservative.)}
+\ndg{mention overinformativity in refexp somewhere in here...}
 
 One major concern is our graphical representation of occluded cells.
-We simply told players that items behind black cells were hidden to the director, and this was a critical question in the quiz that players had to pass before playing, so we know that they were aware of it.
-But they may forget part-way through the experiment, or don't fully internalize it.
+We told players that items behind black cells were hidden to the director, showed them side-by-side displays of the director and matcher views, and quizzed them on this detail in the short attention check that players had to pass before playing.
+But they may have forgotten part-way through the experiment, or failed to fully internalize it.
 In the in-lab version, there was an actual divider between players, so it would have been more salient.
-This would mean that some errors in our version game aren't due to a failure of perspective-taking, but instead because of a misunderstanding of what the other person's perspective is to begin with.
-However, this would only bias our estimates of errors toward supporting Keysar's results, making our analysis conservative if anything.
+We suspect that some subset of errors in our version game aren't due to a failure of perspective-taking, but instead because of a misunderstanding of what the other person's perspective is to begin with.
+For example, one matcher wrote, ``This seems to work wether [sic] the space is black or not.'' This suggests that they thought that the black squares were obstacles or invalid destinations.
+However, this source of errors is held constant across our two experiments, and it would only bias our estimates of errors toward supporting Keysar's results, making our analysis conservative.
+
+At a methodological level, the pair of experiments we report is a reminder that manipulations administered by a trained confederate can interact in unexpected ways with a participant's social and communicative expectations.
+Regardless of the experimental context, it's illuminating to see how real participants naturally interact, and
 
 \todo[inline]{rdh: methodological note that running multi-player control w/o confederate is important to know whether confederate behavior is valid and cooperative}
-
-\ndg{mention overinformativity in refexp somewhere in here...}
 

### 2016-01-26 writing/cogsci-submission/cogsci_keysar.tex
Add mturk rating analysis & fig into results

@@
 
+ \begin{figure}[b!]
+\begin{center}
+\includegraphics[scale=.75]{images/exp1MouseTracking.pdf}
+\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
+\label{fig:exp1hover}
+\end{center}
+\end{figure}
+
+\begin{figure}[b!]
+\begin{center}
+\includegraphics[scale=.75]{images/exp2MouseTracking.pdf}
+\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions.
+Error bars are 95\% confidence intervals.
+\todo[inline]{TODO: Collapse mouse-tracking plots to a single figure}}
+\label{fig:exp2hover}
+\end{center}
+\end{figure}
+
 \subsection{Results}
@@
 When we conducted a mouse-tracking analysis identical to the one reported for Exp.~1, we again found a significant interaction between condition and cell type on hover time, $b = -0.18, t(108) = -5.2, p < 0.001$.
-However, the effect is attenuated: in experimental trials, participants spend significantly more time hovering over the target than the distractor (see Figure \ref{fig:exp2hover}) whereas in Exp.~1, participants spent roughly the same amount of time hovering over each.
+However, the effect appears to be attenuated: in experimental trials, participants spend significantly more time hovering over the target than the distractor (see Figure \ref{fig:exp2hover}) whereas in Exp.~1, participants spent roughly the same amount of time hovering over each.
+Due to the relatively small sample size of Exp.~1, however, we do not have the power to show this effect statistically.
 
-\todo[inline]{rdh: I don't think we have enough power to show a difference between mean hover-times in the two experiments -- the first experiment just didn't have a big enough sample (the psych254 class budget was a pretty hard constraint).
-Might be able to get more power by doing mixed model like below, or by calculating the times more precisely (currently, just count number of mouse movement events in time window, which probably adds a lot of noise).}
+Next, we test whether these improvements in performance are in fact due to different speaker behavior.
+We recruited twenty raters on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects.
+Their responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.6$ (95\% CI = $[0.54, 0.66]$).
 
-Next, we test whether these reductions in error rate and hover-time interference are in fact due to different speaker behavior.
-Two coders were given the list of messages that directors sent to convey instructions for critical items, blind to condition.
-The coders determined (1) whether or not the label used by the speaker fit the hidden distractor and (2) whether or not the label used by the director was more or less precise than the scripted label used in Exp.~1.
-Inter-rater reliability was relatively high, with Cohen's $\kappa=0.82$ for distractor fit and $\kappa=0.67$ for the precision comparison.
-Binary rater decisions were averaged to form a single distractor fit score for each message, and a single precision score.
+In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating, $b = -0.33, t = 32, p < 0.001$.
+In Exp.~1, the scripted label fit the hidden distractor better than the target, but the ordering reversed in Exp.~2: the unconstrained labels fit the target much better and the hidden distractor much worse (see Fig.
+\ref{fig:fitnessInteraction}).
 
-We found that the referring expressions speakers used to convey critical instructions only fit the distractor 53\% of the time, compared to the scripted message in Exp.~1, which (by definition) fit the distractor 100\% of the time.
-\todo[inline]{Except that `glasses' one, which doesn't really apply to a glasses case.
-I'm now pretty sure that this is the fault of the picture I picked (an \emph{open} glasses case rather than a closed one\dots which was dumb in retrospect) Maybe I should re-run analyses without that item, just in case?}
-Furthermore, 47\% of these referring expressions were rated to be more precise than the corresponding scripted instruction used in Exp.~1.
-\ndg{was remaining 53\% rated as scripted more precise or mostly `can't tell'?} As expected, there's a strong negative relationship between the precision of the label and whether it might apply to the hidden distractor, $r = -0.73$.
-Finally, we used a mixed effects logistic regression model to estimate the effect of distractor fit and precision on the probability of making an error on critical trials.
-We included random intercepts for the eight items.
-We find that participants are significantly more likely to make an error when the label applies to the distractor, $b = 3.15, z = 2.93, p = 0.003$, but that precision has no effect, $p = 0.96$.
+Finally, we used a mixed effects logistic regression model to estimate the effect of distractor fit and speaker constraints on the probability of making an error on critical trials.
+We included random intercepts for game and item ids, and an interaction term for label fitness and label referent.
+We find a significant interaction of referent and fitness ratings on error rates: matchers are significantly more likely to make an error when the label fits the distractor more and the target less, $b =-1.39, z = -8.16, p < 0.001$.
+Additionally, participants were significantly less likely to produce errors in Exp.~2 than Exp.~1, $b = -2.10, z = -2.38, p = 0.017$.
 
-\begin{figure}[b!]
+\todo[inline]{what else should these mixed models do?
+Should I just use mean rating rather than every turker's ratings?
+Feels like using $20\times$ the degrees of freedom is wrong, somehow.
+Would be nice to see if they could account for item-wise effects explicitly somehow?
+Like the ``bottom block'' thing which was confusable in Exp.~1 but not Exp.~2.}
+
+\begin{figure}[t!]
 \begin{center}
-\includegraphics[scale=.75]{images/exp1MouseTracking.pdf}
-\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
-\label{fig:exp1hover}
-\end{center}
-\end{figure}
-
-\begin{figure}[b!]
-\begin{center}
-\includegraphics[scale=.75]{images/exp2MouseTracking.pdf}
-\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
-\label{fig:exp2hover}
+\includegraphics[scale=.18]{images/fitnessInteraction.jpeg}
+\caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
+\label{fig:fitnessInteraction}
 \end{center}
@@
 Removing constraints on the speaker produced a dramatic reduction in overall matcher errors and systematic variation in error rates across items.
-How did unscripted speaker behavior differ from scripted behavior such that these effects emerged?
 We found that unscripted speakers systematically tended to produce more precise, costly referring expressions than their scripted counterparts -- for example, ``the clear audiocassette'' when there was only one ``tape'' in the speaker's display.
 These precise labels were less likely to fit the hidden distractor, and therefore less likely to lead to a matcher error.
-
-A key point
 

### 2016-01-27 writing/cogsci-submission/cogsci_keysar.tex
Fill out discussion

@@
 \begin{center}
-\includegraphics[scale = .35]{images/overall_view.jpg}
+\includegraphics[scale = .5]{images/overall_view.png}
 \end{center}
 \vspace{-.25cm}
-\caption{Interface used in the reported experiments.
-Objects behind the black squares were hidden from the director.
-\todo[inline]{TODO: Use images from tape condition, add extra panel to show matcher view.}}
+\caption{Interface used in the reported experiments. Objects behind the black squares were hidden from the director.}
 \label{fig:interface}
@@
 
-In this paper we argue that the apparent failures this view is based on are in fact successes for sophisticated social reasoning. 
-In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners expect these more informative utterances, and interpret the actual utterances accordingly---resulting in ``errors''.
+In this paper we argue that the apparent failures used to support this view are in fact successes for sophisticated social reasoning.
+In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
 
 The argument of Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate.
-The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects.
-\todo[inline]{Refer to updated Fig.~1 with both director screen and matcher screen }
+The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects (see Fig.
+\ref{fig:interface}).
 The confederate played the role of `director,' giving instructions about how to move objects around a grid, and the participant played the role of `matcher,' attempting to follow these instructions.
-For example, the objects in one trial included a medium-sized measuring cup in one location and a smaller measuring cup in another location.
-The director gave an instruction like `move the large cup up one square,' referring to the medium-sized one.
-Some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
+For example, the objects in one trial included a cassette tape.
+The director gave an instruction like `move the tape up one square,' referring to the cassette.
+Critically, some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
 To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about the director's beliefs.
@@
 Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.
-%46\% of participants reached for the hidden object at least twice. 
 Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.
@@
 While these results are compelling, the paradigm has been criticized from several different angles.
-First, in all experiments, the privileged object was a better fit for the referring expression than the one in common ground (e.g.
-the largest measuring cup vs.
-the medium measuring cup for ``the large measuring cup"), making the privileged object \emph{a priori} more likely to be the referent.
-It would be fairer to compare two objects that fit the referring expression equally well \cite{HellerGrodnerTanenhaus08_Perspective}.
+First, \citeA{HellerGrodnerTanenhaus08_Perspective} has pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
+the hidden roll of tape vs.
+the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent.
+Although relative fit has not been empirically evaluated, it would be fairer to compare two objects that fit the referring expression equally well.
 Second, the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility \cite{HannaTanenhausTrueswell03_CommonGroundPerspective}.
@@
 In this paper, we offer an additional factor that helps account for Keysar's results. 
-Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations; given an utterance, a listener can then reason backward to the most plausible situation.
-This suggests that we consider whether the utterances used by the confederate in Keysar's critical conditions were actually what a speaker would be expected to say. 
-If not, then perhaps participants are making choices that are in fact consistent with correct pragmatic interpretation on the confederate's (uncooperative) utterance.
+Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations; given an utterance, a listener can then reason backward to the most plausible situation \cite{FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
+This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
+If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
 More precisely, when both players know that objects are occluded in the display, the speaker may tend to add additional precision to references in order to avoid confusion.
@@
 In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item.
-%=======
-%Some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
-To perform accurately, the matcher would need to apply theory of mind and reason about the director's beliefs.
-For example, if a much larger measuring cup were placed in an occluded slot and agents failed to account for the director's beliefs, they might interpret `the large cup' to mean this occluded cup even though the director couldn't possibly know about it.
-Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.
-46\% of participants reached for the hidden object at least twice.
-Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.
-%>>>>>>> 179ee1de2933b596af109cc10f157c31fdcea761
 
@@
 We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study.
-We replicated the original finding, but noted a tendency of directors to be ove-rinformative in unscripted trials.
+We replicated the original finding, but noted a tendency of directors to be overinformative in unscripted trials.
 We then ran the same experiment without scripting the directions for critical trials, observing unconstrained director behavior.
-We observe much greater precision in unconstrained director utterances and better performance of the matchers.
-This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectatuions about speaker behavior---that is, to theory of mind.
-
-\ndg{intro can probably be compressed some if needed.}
+We observe much greater precision in unconstrained director utterances, which match targets much better than distractors, and better performance of the matchers.
+This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.
 
@@
 The third section shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object.
-\todo[inline]{Big caption maybe unnecessary.
-Anything else to include in this table?
-Maybe Keysar et al, 2000, which reported \% of total cases, or our Exp.
-2 results if you remove the two most difficult items?
-There, errors drop quite a lot.}}
+}
 \label{table:mainResults}
@@
 \end{tabular}
-\caption{Item-wise error rates}
+\caption{Item-wise error rates for critical trials}
 \label{table:ItemWise}
@@
 
-Removing constraints on the speaker produced a dramatic reduction in overall matcher errors and systematic variation in error rates across items.
+Pragmatic language understanding requires sophisticated social reasoning.
+To interpret an utterance, a listener must consider the speaker's communicative goals and context.
+In both experiments, we found evidence of such considerations: when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors.
+This suggests that reference disambiguation was driven more by \emph{a priori} label fitness than consideration of the speaker's epistemic state, as Keysar and colleagues claim.
+
+However, this does not necessarily imply limitations on theory of mind.
+If listeners arrive at this interpretation strategy through pragmatic reasoning, then neglecting the speaker's epistemic state would be rational and also the result of theory of mind use.
+In Exp.~2, we found that such pragmatic interpretations were in fact justified: speakers naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor.
+Thus, the patterns of errors observed in Exp.~1 can be explained as an uncooperative confederate setting up certain pragmatic expectations through the task context, then deliberately flouting them in critical trials.
+% Removing constraints on the speaker produced a dramatic reduction in overall matcher errors and systematic variation in error rates across items.
 We found that unscripted speakers systematically tended to produce more precise, costly referring expressions than their scripted counterparts -- for example, ``the clear audiocassette'' when there was only one ``tape'' in the speaker's display.
@@
 
-We noticed several differences between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} that prevented a direct replication.
-The primary difference between our study and the original is the fact that it will be run online with participants connected via a virtual environment, instead of face-to-face in a room.
+It is worth noting several significant differences between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} that prevented a direct replication.
+The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room.
+We believe we addressed the major concern about exploring theory of mind in web experiments -- that participants do not truly believe they are interacting with another human -- by allowing instantaneous, responsive, real-time interaction.
+On the other hand, it is known that textual communication, as in our chat box, can differ from face-to-face verbal communication, and there was no way of addressing this.
+% Directors watched objects move through the actual moment-by-moment mouse trajectories that matchers used to drag them, and matchers saw the actual timing of the messages sent by the director.
+
 A related difference is our decision not to use a confederate.
-In general, studies using confederates are difficult to replicate exactly and are impractical to run online.
-We are aware that this change (while keeping scripted instructions, which were the primary justification for the confederate) may lead to additional variation across groups.
-It is also known that textual communication (via our instant messaging interface) can differ from verbal communication (as in the original study).
-Finally, the way people attempt to move things in our virtual environment may be different from how people attempt to move things in the lab, and we cannot eye-tracking data to back up our analysis, as the original paper did.
+While a confederate is useful for reducing variation across instances of the experiment, we see several downsides.
+Beyond the difficulties of conducting large-scale experiments with confederates on the web, it is difficult to exactly replicate all the subtleties of the confederate's behavior that might influence the .
+At a methodological level, the pair of experiments we report is a reminder that manipulations administered by a trained confederate can interact in unexpected ways with a participant's social and communicative expectations.
+Regardless of the experimental context, it's illuminating to see how real participants naturally interact, and
 
-% Second, it's worth thinking about some ways our online version may differ from the original.
-The fact that we didn't use a confederate didn't seem to be a problem, given our use of randomly scripted instructions.
-Similarly, because players could type messages directly to one another, and since the director could watch the matcher moving objects in real-time, it's also fair to say that participants truly believed they were playing with another human being.
-
-\ndg{mention overinformativity in refexp somewhere in here...}
-
-One major concern is our graphical representation of occluded cells.
+A final concern is our graphical representation of occluded cells.
 We told players that items behind black cells were hidden to the director, showed them side-by-side displays of the director and matcher views, and quizzed them on this detail in the short attention check that players had to pass before playing.
@@
 We suspect that some subset of errors in our version game aren't due to a failure of perspective-taking, but instead because of a misunderstanding of what the other person's perspective is to begin with.
-For example, one matcher wrote, ``This seems to work wether [sic] the space is black or not.'' This suggests that they thought that the black squares were obstacles or invalid destinations.
-However, this source of errors is held constant across our two experiments, and it would only bias our estimates of errors toward supporting Keysar's results, making our analysis conservative.
+However, this source of errors is held constant across our two experiments, and it would only bias our estimates of errors upward.
 
-At a methodological level, the pair of experiments we report is a reminder that manipulations administered by a trained confederate can interact in unexpected ways with a participant's social and communicative expectations.
-Regardless of the experimental context, it's illuminating to see how real participants naturally interact, and
+The Gricean maxim of quality dictates that cooperative speakers should make their utterance as informative as is required for current purposes, but no more so.
+If the referring expressions used in Exp.~1 uniquely pick out the target from the speaker's perspective, then why would unscripted speakers in Exp.~2 strive to be more informative than strictly necessary and why would listeners come to expect such ``overinformativity'' through social reasoning?
+Studies of overinformativity in referring expressions have uncovered a number of factors, such as the tendency to mention perceptually salient features to speed up the identification process  \cite{KoolenGattGoudbeekKrahmer11_Overspecification}.
+In Keysar's task, it seems particularly relevant that both participants knew about the presence of hidden objects, increasing uncertainty about possible distractors and placing additional pressure on the speaker to overspecify.
 
-\todo[inline]{rdh: methodological note that running multi-player control w/o confederate is important to know whether confederate behavior is valid and cooperative}
+It is interesting that the item-wise variability we observed in both experiments is at least partially due to the availability and effectiveness of more specific expressions.
+For example, no speaker in Exp.~2 produced ``the bottom block,'' which was used as the critical instruction in Exp.~1.
+Instead, they said ``the bluish block with a B'' or ``block with the blue writing,'' which relied on less confusable perceptual features and thereby decreased error rates.
+The items on which errors were most likely in Exp.~2 were difficult precisely because the overinformative utterance was \emph{still} ambiguous: even if the speaker uses ``hairbrush'' instead of the basic-level ``brush,'' for example, the distractor remains a more typical exemplar of the subcategory.
 

### 2016-01-27 experiment2/analysis/exp2Analysis.Rmd
Incorporate turk judges into analysis; bootstrap errors

@@
 ```{r}
-messageCoding1 <- read.csv("../analysis/messageCoding1.csv")
-messageCoding2 <- read.csv("../analysis/messageCoding2.csv")
-relevantProperties <- c("gameid", "attemptNum", "condition", "critical",
-                        "objectSet", "sender", "instructionNum")
-overinformativeCodes <- messageCoding1 %>% 
-  left_join(messageCoding2, by = relevantProperties) %>%
+library(boot)
+turkerRatings <- read.csv("../../normingExperiment/data/keysarNorming-trials.csv")
+relevantProperties <- c("objectSet", "label")
+
+criticalMessagesExp1 <- read.csv("../analysis/messagesExp1.csv") %>%
+  left_join(turkerRatings, by = relevantProperties) %>%
+  mutate(exp = "Exp.1")
+criticalMessagesExp2 <- read.csv("../analysis/messagesExp2.csv") %>%
+  right_join(turkerRatings, by = relevantProperties) %>%
+  mutate(exp = "Exp.2")
+
+overinformativeCodes <- rbind(criticalMessagesExp1, criticalMessagesExp2) %>%
   filter(!(gameid %in% bannedIDsExp2)) %>%
-  mutate(contents = contents.x, time = time.x) %>%
-  select(gameid, attemptNum, condition, critical, objectSet, sender, 
-         instructionNum, contents, time, targetCode1, distractCode1,
-         altCode1, overCode1, targetCode2, distractCode2, altCode2,
-         overCode2) %>%
-  mutate(avgAltCode = (altCode1 + altCode2)/2,
-         avgTargetCode = (targetCode1 + targetCode2)/2,
-         avgDistractCode = (distractCode1 + distractCode2)/2,
-         avgOverCode = (overCode1 + overCode2)/2)
+  filter(!(gameid %in% bannedIDsExp1))
 
-# How often are the descriptions more informative than orig?
-mean(overinformativeCodes$avgOverCode)
+summary(lmer(response ~ exp * referent + (1 | workerid) + (1 | objectSet), data = overinformativeCodes))
 
-# How often do the descriptions apply to the hidden object?
-mean(overinformativeCodes$avgDistractCode)
+stat = function(data, indices) {
+  d <- data[indices,]
+  applicability = d %>%   
+    group_by(exp, referent) %>%
+    summarize(m = mean(response))
+  print(applicability)
+  return(applicability$m)
+}
 
-# Does overinformativity perfectly avoid picking distractor?
-cor(overinformativeCodes$avgDistractCode, overinformativeCodes$avgOverCode)
+res = boot(data = overinformativeCodes, statistic = stat, R = 1000)
+Exp1DistrCI = boot.ci(res, type = "basic", index=1)
+Exp1targetCI = boot.ci(res, type = "basic", index=2)
+Exp2DistrCI = boot.ci(res, type = "basic", index=3)
+Exp2targetCI = boot.ci(res, type = "basic", index=4)
 
-# Inter-rater reliability
-kappa2(cbind(overinformativeCodes$distractCode1, overinformativeCodes$distractCode2))
-kappa2(cbind(overinformativeCodes$overCode1, overinformativeCodes$overCode2))
+plottingDF = data.frame(referent = c("distractor", "target", "distractor", "target"),
+                        exp = c("Exp.1", "Exp.1", "Exp.2", "Exp.2"),
+                        response = res$t0,
+                        lowerCI = c(Exp1DistrCI$basic[1,4], Exp1targetCI$basic[1,4],
+                                    Exp2DistrCI$basic[1,4], Exp2targetCI$basic[1,4]),
+                        upperCI = c(Exp1DistrCI$basic[1,5], Exp1targetCI$basic[1,5],
+                                    Exp2DistrCI$basic[1,5], Exp2targetCI$basic[1,5]))
+dodge <- position_dodge(width=0.9)
+ggplot(plottingDF, aes(x = exp, y = response, fill = referent)) +
+  geom_bar(stat = "identity", position = dodge) +
+  geom_errorbar(aes(ymin = lowerCI, ymax = upperCI), position = dodge, width = 0.25) +
+  theme_bw() +
+  ylab("mean fitness of label")
+ggsave("../../writing/cogsci-submission/images/fitnessInteraction.jpeg")
+
+# spread out data, so each row is a label, and each column is a turker
+wideRatings = turkerRatings %>% 
+  group_by(workerid, label, referent) %>% 
+  summarize(m = mean(response)) %>% 
+  spread(workerid, m)
+
+# Interrater reliability
+icc(wideRatings[,-c(1,2)],
+    model = "twoway", type = "agreement")
+
 ```
@@
 ```{r}
-criticalErrors = errorDExp2 %>% 
+criticalErrorsExp1 = errorDExp1 %>%
+  filter(attemptNum == 0) %>%
+  filter(criticalError == 1) %>%
+  mutate(exp = "Exp.1") %>%
+  select(gameid, objectSet, exp, criticalError)
+criticalErrorsExp2 = errorDExp2 %>% 
   filter(attemptNum == 0)%>% 
-  filter(criticalError == 1) 
+  filter(criticalError == 1) %>%
+  mutate(exp = "Exp.2") %>%
+  select(gameid, objectSet, exp, criticalError)
+
+criticalErrors = rbind(criticalErrorsExp1, criticalErrorsExp2)
+
 itemWiseCodes <- overinformativeCodes %>% 
-  left_join(criticalErrors, by = c("gameid", "condition", "critical", 
-                                   "objectSet", "instructionNum",
-                                   "attemptNum")) %>%
-  select(gameid, objectSet, avgDistractCode, avgOverCode, 
-         criticalError) %>%
+  left_join(criticalErrors, by = c("gameid", "objectSet", "exp")) %>%
   mutate(criticalError = ifelse(is.na(criticalError),0,1))
 
-mmRes <- glmer(criticalError ~ avgOverCode + avgDistractCode +  (1 | objectSet),
+mmRes <- glmer(criticalError ~ response * referent +  exp + (1 | objectSet),
       data = itemWiseCodes, family = binomial)

### 2016-01-28 experiment2/analysis/exp2Analysis.Rmd
Fix some joining bugs that create NAs

@@
 ```{r}
+setwd('~/Repos/keysar_replication/experiment1/analysis/')
+preMouseDExp1 = read.csv('fullSampleMouse.csv')
+preErrorDExp1 = read.csv('fullSampleError.csv')
+preMessageDExp1 = read.csv('fullSampleMessage.csv')
+
 setwd('~/Repos/keysar_replication/experiment2/data/')
@@
 ```{r}
+nonCriticalMistakes <- preErrorDExp1 %>%
+  # only look at mistakes on noncritical (filler) items
+  filter(critical != 1) %>%
+  # don't want to double-count people for messing the same thing up multiple times
+  filter(attemptNum == 0) %>% 
+  group_by(gameid) %>%
+  tally() %>%
+  # implement exclusion criteria of errors on >~10% of non-critical trials
+  filter(n >= 2)
+
+bannedIDsExp1 <- nonCriticalMistakes$gameid
+numParticipantsExp1 <- (length(unique(preMouseDExp1$gameid)) -
+                        length(bannedIDsExp1))
+
 nonNativeEnglishIDs <- c("0366-3", "8676-f")
@@
 ```{r}
+errorDExp1 <- preErrorDExp1 %>%
+  filter(!(gameid %in% bannedIDsExp1))
+mouseDExp1 <- preMouseDExp1 %>%
+  filter(!(gameid %in% bannedIDsExp1))
+messageDExp1 <- preMessageDExp1 %>%
+  filter(!(gameid %in% bannedIDsExp1))
+
 errorDExp2 <- preErrorDExp2 %>%
@@
 library(boot)
+library(irr)
 turkerRatings <- read.csv("../../normingExperiment/data/keysarNorming-trials.csv")
@@
 criticalMessagesExp2 <- read.csv("../analysis/messagesExp2.csv") %>%
-  right_join(turkerRatings, by = relevantProperties) %>%
+  left_join(turkerRatings, by = relevantProperties) %>%
   mutate(exp = "Exp.2")
@@
 
-summary(lmer(response ~ exp * referent + (1 | workerid) + (1 | objectSet), data = overinformativeCodes))
+View(overinformativeCodes)
+summary(lmer(response ~ exp * referent + (1 | workerid) + (1 | objectSet), 
+             data = overinformativeCodes))
 
@@
     summarize(m = mean(response))
-  print(applicability)
   return(applicability$m)
@@
   mutate(criticalError = ifelse(is.na(criticalError),0,1))
-
-mmRes <- glmer(criticalError ~ response * referent +  exp + (1 | objectSet),
+View(itemWiseCodes)
+mmRes <- glmer(criticalError ~ response * referent +  exp + (1 | objectSet) + (1 | gameid),
       data = itemWiseCodes, family = binomial)
-
+plot(allEffects(mmRes))
 summary(mmRes)
@@
 ```{r}
-dUnfiltExp2 <- joinedExp2 %>%
-  filter(critical == 1) %>%
-  group_by(objectSet)
-
 dExp2 <- joinedExp2 %>% 
@@
    filter(distractorY == begDistrY) 
-
+View(dExp2)
 ```
@@
                           mutate(exp = "Exp.1") %>% 
-                          filter(condition == "exp"),
+                          filter(cellType == "target"),
                         heatMapForLMExp2 %>% 
                           mutate(exp = "Exp.2") %>%
-                          filter(condition == "exp"))
+                          filter(cellType == "target"))
 combinedHeatMapForPlot <- combinedHeatMap %>%
-  group_by(cellType, exp) %>%
+  group_by(condition, exp) %>%
   summarize(error = sem(percent),
@@
 
-g <- (ggplot(combinedHeatMapForPlot, aes(x = cellType, y = percent, 
-                                  group = exp, color = exp)) 
-      + geom_line(aes(linetype = exp), size = 2)
+g <- (ggplot(combinedHeatMapForPlot, aes(x = exp, y = percent, 
+                                  group = condition, color = condition)) 
+      + geom_line(aes(linetype = condition), size = 2)
       + geom_errorbar(aes(ymax = percent + error, 

### 2016-01-28 writing/cogsci-submission/cogsci_keysar.tex
Update abstract, add some notes

@@
 \begin{abstract}
-?
+People are capable of accurately reasoning about the mental states of other people using theory of mind.
+An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take the speaker's perspective into account.
+This has been widely interpreted as a limitation on theory of mind use.
+In this paper we argue that these apparent failures are in fact successes.
+Through a minimal pair of replications comparing scripted vs.
+unscripted speakers, we show that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation.
+A listener who correctly reasons about the pragmatics of the task, thereby \emph{using} theory of mind, will rationally ignore whether an object is hidden to the speaker or not and will instead rely on the precision of the utterance to disambiguate.
+When an uncooperative confederate flouts these expectations, this strategy produces errors; when we allow participants to naturally interact, this strategy is justified and errors are reduced.
 
 \textbf{Keywords:} 
-?
+Theory of mind; social cognition; pragmatics
 \end{abstract}
@@
 the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent.
-Although relative fit has not been empirically evaluated, it would be fairer to compare two objects that fit the referring expression equally well.
+Although relative fit was not empirically evaluated prior to our study, it would generally be fairer to compare two objects that fit the referring expression equally well.
 Second, the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility \cite{HannaTanenhausTrueswell03_CommonGroundPerspective}.
@@
 Error bars are 95\% confidence intervals.
-\todo[inline]{TODO: Collapse mouse-tracking plots to a single figure}}
+\todo[inline]{TODO: Collapse mouse-tracking plots to a single figure showing some interaction}}
 \label{fig:exp2hover}
@@
 ``Hair brush'' is already a label at the sub-class level.
-If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 26\% to 10\% and the percentage of participants making at least one error drops from 60\% to 32\%.
+If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.
 
+\todo[inline]{rdh: I don't like these mouse-tracking analyses -- I feel like the way I'm computing hover-time is sloppy and might be skewing the results (i.e.
+counting number of `move' events rather than actually looking at start and end times).
+So I could spend some time tomorrow coding up the better way\dots}
 When we conducted a mouse-tracking analysis identical to the one reported for Exp.~1, we again found a significant interaction between condition and cell type on hover time, $b = -0.18, t(108) = -5.2, p < 0.001$.
@@
 
-In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating, $b = -0.33, t = 32, p < 0.001$.
+In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating, $b = 0.34, t = 36.4, p < 0.001$.
 In Exp.~1, the scripted label fit the hidden distractor better than the target, but the ordering reversed in Exp.~2: the unconstrained labels fit the target much better and the hidden distractor much worse (see Fig.
@@
 We included random intercepts for game and item ids, and an interaction term for label fitness and label referent.
-We find a significant interaction of referent and fitness ratings on error rates: matchers are significantly more likely to make an error when the label fits the distractor more and the target less, $b =-1.39, z = -8.16, p < 0.001$.
-Additionally, participants were significantly less likely to produce errors in Exp.~2 than Exp.~1, $b = -2.10, z = -2.38, p = 0.017$.
+We find a significant interaction of referent and fitness ratings on error rates: matchers are significantly more likely to make an error when the label fits the distractor more and the target less, $b =-1.39, z = -, p < 0.001$.
+Additionally, participants were significantly less likely to produce errors in Exp.~2 than Exp.~1, $b = -1.38, z = -2.41, p = 0.016$.
 
-\todo[inline]{what else should these mixed models do?
-Should I just use mean rating rather than every turker's ratings?
-Feels like using $20\times$ the degrees of freedom is wrong, somehow.
-Would be nice to see if they could account for item-wise effects explicitly somehow?
-Like the ``bottom block'' thing which was confusable in Exp.~1 but not Exp.~2.}
+\todo[inline]{rdh: I still get confused about which random effects to include\dots does it make sense to include random intercepts for game id and include experiment \# as a predictor when each id only participated in one experiment?}
 
@@
 While a confederate is useful for reducing variation across instances of the experiment, we see several downsides.
-Beyond the difficulties of conducting large-scale experiments with confederates on the web, it is difficult to exactly replicate all the subtleties of the confederate's behavior that might influence the .
-At a methodological level, the pair of experiments we report is a reminder that manipulations administered by a trained confederate can interact in unexpected ways with a participant's social and communicative expectations.
-Regardless of the experimental context, it's illuminating to see how real participants naturally interact, and
+Beyond the difficulties of conducting large-scale experiments with confederates on the web, it is difficult to exactly replicate all the subtleties of the confederate's behavior that might influence results.
+The pair of experiments we report is a reminder that manipulations administered by a trained confederate can interact in unexpected ways with a participant's social and communicative expectations.
+Regardless of the experimental context, it's illuminating to see how real participants naturally interact.
 
+\todo[inline]{rdh: can take out next paragraph if hurting for space...}
 A final concern is our graphical representation of occluded cells.

### 2016-01-28 writing/cogsci-submission/cogsci_keysar.tex
possible new title
some edits

@@
 
-\title{Pragmatics account for apparent limits on theory of mind use in adults}
+\title{Conversational expectations account for apparent limits on theory of mind use}
 
@@
 When an uncooperative confederate flouts these expectations, this strategy produces errors; when we allow participants to naturally interact, this strategy is justified and errors are reduced.
+\ndg{make a bit less technical}
 
@@
 
-\begin{figure*}[t!]
+\begin{figure*}[bt!]
 \begin{center}
@@
 While these results are compelling, the paradigm has been criticized from several different angles. 
-First, \citeA{HellerGrodnerTanenhaus08_Perspective} has pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
+First, \citeA{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
 the hidden roll of tape vs.
-the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent.
-Although relative fit was not empirically evaluated prior to our study, it would generally be fairer to compare two objects that fit the referring expression equally well.
+the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well.
+We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
+% Although relative fit was not empirically evaluated prior to our study,
 Second, the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility \cite{HannaTanenhausTrueswell03_CommonGroundPerspective}.
@@
 In this paper, we offer an additional factor that helps account for Keysar's results. 
-Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations; given an utterance, a listener can then reason backward to the most plausible situation \cite{FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
+Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations; given an utterance, a listener can then reason backward to the most plausible situation \cite{grice,clark,FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
 This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
@@
 For half the instructions, directors were free to communicate however they wished.
-For the other half, including all the critical instructions, their messages to the matcher were pre-scripted such that we could use the precise wording from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+For the other half, including all the critical instructions, their messages to the matcher were pre-scripted using the precise wording from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat window, and they would have to click `send' for the experiment to continue.)
@@
 This suggests that the dependent variable highlighted in the original study (i.e.
-``percentage of participants who moved the critical item at least once'') is somewhat problematic.
-It could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item.
+``percentage of participants who moved the critical item at least once'') is somewhat problematic: it could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item.
 Indeed, if we exclude the three `hard' items where over 60\% of participants in the experimental condition made errors, this dependent variable drops from 93\% to only 43\% of participants.
 %This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
+\ndg{say something about bad items maybe being ones where utterance fits distractor better?}
 
@@
 In baseline trials, people spent much more time in the target cell and much less time in the distractor cell than in the experimental condition (see Fig.
-\ref{fig:exp1hover}).\footnote{Note that this analysis includes participants who actually made errors, since the data is too sparse to exclude them.} %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
+\ref{fig:hover}).\footnote{Note that this analysis includes participants who actually made errors, since the data is too sparse to exclude them.} %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
 
@@
 If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells).
-That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative}, and thus led matchers astray. 
+That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
 We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
@@
 
+ \begin{figure}[bt]
+\begin{center}
+\includegraphics[scale=.35]{images/exp1MouseTracking.pdf}
+\includegraphics[scale=.35]{images/exp2MouseTracking.pdf}
+\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions for Expt.~1 (left) and Expt.~2 (right).
+Error bars are 95\% confidence intervals.
+\ndg{merge figs?}}
+\label{fig:hover}
+\end{center}
+\end{figure}
+
 \section{Exp.~2: Unscripted Replication}
@@
 
- \begin{figure}[b!]
-\begin{center}
-\includegraphics[scale=.75]{images/exp1MouseTracking.pdf}
-\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions. Error bars are 95\% confidence intervals.}
-\label{fig:exp1hover}
-\end{center}
-\end{figure}
-
-\begin{figure}[b!]
-\begin{center}
-\includegraphics[scale=.75]{images/exp2MouseTracking.pdf}
-\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions.
+%\begin{figure}[b!]
+%\begin{center}
+%\includegraphics[scale=.75]{images/exp2MouseTracking.pdf}
+%\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions.
 Error bars are 95\% confidence intervals.
 \todo[inline]{TODO: Collapse mouse-tracking plots to a single figure showing some interaction}}
-\label{fig:exp2hover}
-\end{center}
-\end{figure}
+%\label{fig:exp2hover}
+%\end{center}
+%\end{figure}
 
@@
 When we conducted a mouse-tracking analysis identical to the one reported for Exp.~1, we again found a significant interaction between condition and cell type on hover time, $b = -0.18, t(108) = -5.2, p < 0.001$.
-However, the effect appears to be attenuated: in experimental trials, participants spend significantly more time hovering over the target than the distractor (see Figure \ref{fig:exp2hover}) whereas in Exp.~1, participants spent roughly the same amount of time hovering over each.
+However, the effect appears to be attenuated: in experimental trials, participants spend significantly more time hovering over the target than the distractor (see Figure \ref{fig:hover}) whereas in Exp.~1, participants spent roughly the same amount of time hovering over each.
 Due to the relatively small sample size of Exp.~1, however, we do not have the power to show this effect statistically.
@@
 
-In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating, $b = 0.34, t = 36.4, p < 0.001$.
-In Exp.~1, the scripted label fit the hidden distractor better than the target, but the ordering reversed in Exp.~2: the unconstrained labels fit the target much better and the hidden distractor much worse (see Fig.
-\ref{fig:fitnessInteraction}).
+In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig.
+\ref{fig:fitnessInteraction}), $b = 0.34, t = 36.4, p < 0.001$.
+In Exp.~1, the scripted label fit the hidden distractor better than the target, but the ordering reversed in Exp.~2: the unconstrained labels fit the target much better and the hidden distractor much worse.
 

### 2016-01-28 writing/cogsci-submission/cogsci_keysar.tex
edits

@@
 %This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
-\ndg{say something about bad items maybe being ones where utterance fits distractor better?}
+\ndg{say something about bad items maybe being ones where utterance fits distractor better? (which we substantiate below..)}
 
@@
 In Exp.~1, the scripted label fit the hidden distractor better than the target, but the ordering reversed in Exp.~2: the unconstrained labels fit the target much better and the hidden distractor much worse.
+\ndg{i wonder if we can make it clearer why this is important?}
 
@@
 \todo[inline]{rdh: I still get confused about which random effects to include\dots does it make sense to include random intercepts for game id and include experiment \# as a predictor when each id only participated in one experiment?}
+
+\ndg{maybe do this in a few steps: first, does label fit explain item effects we noted in expt 1?
+second does it explain remaining errors in expt 2?
+then do a combined model.
+what about making a plot that shows the item-wise fits at least for expt 1?}
 
@@
 
-Pragmatic language understanding requires sophisticated social reasoning.
-To interpret an utterance, a listener must consider the speaker's communicative goals and context.
-In both experiments, we found evidence of such considerations: when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors.
-This suggests that reference disambiguation was driven more by \emph{a priori} label fitness than consideration of the speaker's epistemic state, as Keysar and colleagues claim.
+Pragmatic language understanding requires sophisticated social reasoning. To interpret an utterance, a listener must consider how a speaker is likely to behave in context.
+The rational use of theory of mind for a listener thus depends on her expectations about the speaker: If, in a particular context, she expects a speaker to provide sufficiently informative utterances, she may be justified in neglecting his epistemic state.
 
-However, this does not necessarily imply limitations on theory of mind.
-If listeners arrive at this interpretation strategy through pragmatic reasoning, then neglecting the speaker's epistemic state would be rational and also the result of theory of mind use.
-In Exp.~2, we found that such pragmatic interpretations were in fact justified: speakers naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor.
-Thus, the patterns of errors observed in Exp.~1 can be explained as an uncooperative confederate setting up certain pragmatic expectations through the task context, then deliberately flouting them in critical trials.
+%consider the speaker's communicative goals and context. 
+In both experiments, we found evidence that listeners neglect the speaker's epistemic state, as Keysar and colleagues claim.
+More specifically, we found that when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors.
+This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness.
+%
+However, this does not necessarily imply limitations on theory of mind---if speakers would naturally provide expressions which apply better to the target, then this behavior would be appropriate.
+% If listeners arrive at this interpretation strategy through pragmatic reasoning, then neglecting the speaker's epistemic state would be rational and also the result of theory of mind use. 
+In Exp.~2, we found that speakers did naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor.
+Thus, the patterns of errors observed in Exp.~1 can be explained as an uncooperative confederate---the experiment set up certain pragmatic expectations through the task context, then deliberately flouted them in critical trials.
 % Removing constraints on the speaker produced a dramatic reduction in overall matcher errors and systematic variation in error rates across items.
@@
 
-It is worth noting several significant differences between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} that prevented a direct replication.
+It is worth noting several significant differences between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+% that prevented a direct replication. 
 The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room.
@@
 While a confederate is useful for reducing variation across instances of the experiment, we see several downsides.
-Beyond the difficulties of conducting large-scale experiments with confederates on the web, it is difficult to exactly replicate all the subtleties of the confederate's behavior that might influence results.
+Beyond the difficulties of conducting large-scale experiments with confederates, it is difficult to exactly replicate all the subtleties of the confederate's behavior that might influence results.
 The pair of experiments we report is a reminder that manipulations administered by a trained confederate can interact in unexpected ways with a participant's social and communicative expectations.
-Regardless of the experimental context, it's illuminating to see how real participants naturally interact.
+Regardless of the experimental context, it is illuminating to see how real participants naturally interact.
 
-\todo[inline]{rdh: can take out next paragraph if hurting for space...}
-A final concern is our graphical representation of occluded cells.
+%\todo[inline]{rdh: can take out next paragraph if hurting for space...}
+%A final concern is our graphical representation of occluded cells.
 We told players that items behind black cells were hidden to the director, showed them side-by-side displays of the director and matcher views, and quizzed them on this detail in the short attention check that players had to pass before playing.
@@
 
+\ndg{there's an abrupt jump here. play with ordering?}
+
 The Gricean maxim of quality dictates that cooperative speakers should make their utterance as informative as is required for current purposes, but no more so.
-If the referring expressions used in Exp.~1 uniquely pick out the target from the speaker's perspective, then why would unscripted speakers in Exp.~2 strive to be more informative than strictly necessary and why would listeners come to expect such ``overinformativity'' through social reasoning?
+If the referring expressions used in Exp.~1 uniquely pick out the target from the speaker's perspective, then why would unscripted speakers in Exp.~2 strive to be more informative than strictly necessary and why would listeners come to expect such ``overinformativity''?
 Studies of overinformativity in referring expressions have uncovered a number of factors, such as the tendency to mention perceptually salient features to speed up the identification process  \cite{KoolenGattGoudbeekKrahmer11_Overspecification}.
-In Keysar's task, it seems particularly relevant that both participants knew about the presence of hidden objects, increasing uncertainty about possible distractors and placing additional pressure on the speaker to overspecify.
+In the current task, it seems particularly relevant that both participants knew about the presence of hidden objects, increasing uncertainty about possible distractors and placing additional pressure on the speaker to overspecify.
 
@@
 The items on which errors were most likely in Exp.~2 were difficult precisely because the overinformative utterance was \emph{still} ambiguous: even if the speaker uses ``hairbrush'' instead of the basic-level ``brush,'' for example, the distractor remains a more typical exemplar of the subcategory.
+
+\ndg{concluding paragraph. Overall, we both replicated the finding that listeners neglect the speakers epistemic state, and explained why this may be so (and why errors are uneven over items) ...
+}
 

### 2016-01-29 writing/cogsci-submission/cogsci_keysar.tex
rearranged conclusion a bit

@@
 Pragmatic language understanding requires sophisticated social reasoning. To interpret an utterance, a listener must consider how a speaker is likely to behave in context.
-The rational use of theory of mind for a listener thus depends on her expectations about the speaker: If, in a particular context, she expects a speaker to provide sufficiently informative utterances, she may be justified in neglecting his epistemic state.
+The rational use of theory of mind for a listener thus depends on her expectations about the speaker: If, in a particular context, she expects a speaker to provide sufficiently informative utterances, she may be justified in neglecting his epistemic state when resolving reference.
 
 %consider the speaker's communicative goals and context. 
-In both experiments, we found evidence that listeners neglect the speaker's epistemic state, as Keysar and colleagues claim.
-More specifically, we found that when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors.
+In our replication (Expt.~1), we found evidence that listeners neglect the speaker's epistemic state, as Keysar and colleagues claim.
+We also found that when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors. 
+Indeed, we found that the extremely heterogeneous pattern of errors across items was well explained by the relative fit of the utterance to target and distractor.
 This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness. 
 %
-However, this does not necessarily imply limitations on theory of mind---if speakers would naturally provide expressions which apply better to the target, then this behavior would be appropriate.
+However, this does not necessarily imply limitations on theory of mind---if speakers could be expected to naturally provide expressions which apply better to the target, then this behavior would be appropriate.
 % If listeners arrive at this interpretation strategy through pragmatic reasoning, then neglecting the speaker's epistemic state would be rational and also the result of theory of mind use. 
-In Exp.~2, we found that speakers did naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor.
-Thus, the patterns of errors observed in Exp.~1 can be explained as an uncooperative confederate---the experiment set up certain pragmatic expectations through the task context, then deliberately flouted them in critical trials.
+In Expt.~2, we found that speakers did naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor.
+For example, no speaker in Expt.~2 produced ``the bottom block,'' which was used as a critical instruction in Exp.~1.
+Instead, they said ``the bluish block with a B'' or ``block with the blue writing,'' which relied on less confusable perceptual features and thereby decreased error rates.
+Thus, the errors observed in Expt.~1 can be explained as a result of an uncooperative confederate (speaker)---the experiment set up certain pragmatic expectations through the task context, then deliberately flouted them in critical trials.
 % Removing constraints on the speaker produced a dramatic reduction in overall matcher errors and systematic variation in error rates across items.
@@
 These precise labels were less likely to fit the hidden distractor, and therefore less likely to lead to a matcher error.
+
+The Gricean maxim of quality dictates that cooperative speakers should make their utterance as informative as is required for current purposes, but no more so.
+If the referring expressions used in Expt.~1 uniquely pick out the target from the speaker's perspective, then why would unscripted speakers in Expt.~2 be more informative than this?
+%strictly necessary and why would listeners come to expect such ``overinformativity''? 
+One possibility is that awareness of the complex mismatch in epistemic state with the listener leads the speaker to provide extra information in an attempt to avoid miscommunication.
+Another possibility is that more general dynamics of language are in play.
+Studies of over-informativity in referring expressions have uncovered a general tendency of speakers to provide redundant information.
+This tendency can depend on a number of situational factors, such as the tendency to mention perceptually salient features to speed up the identification process  \cite{KoolenGattGoudbeekKrahmer11_Overspecification}.
+%In the current task, it seems particularly relevant that both participants knew about the presence of hidden objects, increasing uncertainty about possible distractors and placing additional pressure on the speaker to overspecify.
+The exact origin is a subject that must be followed up by future research, but it is clear that the tendency of speakers to produce highly informative referring expressions is useful to, and relied on by, listeners.
+Indeed, the remaining errors in Expt.~2 are at least partially due to the lack of availability of more specific expressions. 
+The items on which errors were most likely in Expt.~2 were difficult precisely because the natural utterance was \emph{still} ambiguous: even if the speaker uses ``hairbrush'' instead of the basic-level ``brush,'' for example, the distractor remains a more typical exemplar of the subcategory.
 
@@
 
-\ndg{there's an abrupt jump here. play with ordering?}
-
-The Gricean maxim of quality dictates that cooperative speakers should make their utterance as informative as is required for current purposes, but no more so.
-If the referring expressions used in Exp.~1 uniquely pick out the target from the speaker's perspective, then why would unscripted speakers in Exp.~2 strive to be more informative than strictly necessary and why would listeners come to expect such ``overinformativity''?
-Studies of overinformativity in referring expressions have uncovered a number of factors, such as the tendency to mention perceptually salient features to speed up the identification process  \cite{KoolenGattGoudbeekKrahmer11_Overspecification}.
-In the current task, it seems particularly relevant that both participants knew about the presence of hidden objects, increasing uncertainty about possible distractors and placing additional pressure on the speaker to overspecify.
-
-It is interesting that the item-wise variability we observed in both experiments is at least partially due to the availability and effectiveness of more specific expressions.
-For example, no speaker in Exp.~2 produced ``the bottom block,'' which was used as the critical instruction in Exp.~1.
-Instead, they said ``the bluish block with a B'' or ``block with the blue writing,'' which relied on less confusable perceptual features and thereby decreased error rates.
-The items on which errors were most likely in Exp.~2 were difficult precisely because the overinformative utterance was \emph{still} ambiguous: even if the speaker uses ``hairbrush'' instead of the basic-level ``brush,'' for example, the distractor remains a more typical exemplar of the subcategory.
-
-\ndg{concluding paragraph. Overall, we both replicated the finding that listeners neglect the speakers epistemic state, and explained why this may be so (and why errors are uneven over items) ...
+\ndg{concluding paragraph? Overall, we both replicated the finding that listeners neglect the speakers epistemic state, and explained why this may be so (and why errors are uneven over items) ...
+Or maybe we don't need anything else.
 }

### 2016-01-29 writing/cogsci-submission/cogsci_keysar.tex
Update results; tentative hovertime fig

--- 2016-01-28 writing/cogsci-submission/cogsci_keysar.tex
@@
 \definecolor{Green}{RGB}{10,200,100}
+\definecolor{Red}{RGB}{200,10,100}
+
 \newcommand{\ndg}[1]{\textcolor{Green}{[ndg: #1]}}
+\newcommand{\rdh}[1]{\textcolor{Red}{[rdh: #1]}}
 
@@
 
- \begin{figure}[bt]
+ \begin{figure}[bt!]
 \begin{center}
-\includegraphics[scale=.35]{images/exp1MouseTracking.pdf}
-\includegraphics[scale=.35]{images/exp2MouseTracking.pdf}
-\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions for Expt.~1 (left) and Expt.~2 (right).
-Error bars are 95\% confidence intervals.
-\ndg{merge figs?}}
+\includegraphics[scale=.8]{images/mousetracking.pdf}
+\caption{Mean percentage of time spent hovering over target cell in the experimental and baseline conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are 95\% confidence intervals.}
 \label{fig:hover}
@@
 
-Next, we test whether these improvements in performance are in fact due to different speaker behavior.
-We recruited twenty raters on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects.
+Next, we test whether these improvements in performance are in fact due to more informative speaker behavior.
+We recruited twenty judges on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects.
 Their responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.6$ (95\% CI = $[0.54, 0.66]$).
+In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig.
+\ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$.
+In Exp.~1, the scripted label fit the hidden distractor just as well as the target, but in Exp.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
+In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.
 
-In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig.
-\ref{fig:fitnessInteraction}), $b = 0.34, t = 36.4, p < 0.001$.
-In Exp.~1, the scripted label fit the hidden distractor better than the target, but the ordering reversed in Exp.~2: the unconstrained labels fit the target much better and the hidden distractor much worse.
-\ndg{i wonder if we can make it clearer why this is important?}
+Does differential informativity of scripted utterances account for the variability across items that we noted in Exp.~1?
+In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
+We find a correlation of $r = 0.76$ between these measures for Exp.~1, and a correlation of $r = 0.53$ for Exp.~2.
+The cross-experiment correlations of error rates from one experiment and label fitness ratios from the other, however, is only $r = 0.07$, suggesting that item-wise variability across the two experiments is driven by relative informativity of speaker utterances.
 
-Finally, we used a mixed effects logistic regression model to estimate the effect of distractor fit and speaker constraints on the probability of making an error on critical trials.
-We included random intercepts for game and item ids, and an interaction term for label fitness and label referent.
-We find a significant interaction of referent and fitness ratings on error rates: matchers are significantly more likely to make an error when the label fits the distractor more and the target less, $b =-1.39, z = -, p < 0.001$.
-Additionally, participants were significantly less likely to produce errors in Exp.~2 than Exp.~1, $b = -1.38, z = -2.41, p = 0.016$.
-
-\todo[inline]{rdh: I still get confused about which random effects to include\dots does it make sense to include random intercepts for game id and include experiment \# as a predictor when each id only participated in one experiment?}
-
-\ndg{maybe do this in a few steps: first, does label fit explain item effects we noted in expt 1?
-second does it explain remaining errors in expt 2?
-then do a combined model.
-what about making a plot that shows the item-wise fits at least for expt 1?}
+After separately establishing that matchers in Exp.~2 make fewer mistakes, that directors produce more informative utterances, and that informativity captures item-wise variability in error rates, we tested the link between these effects: does label informativity generally predict errors on critical trials?
+We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
+We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$.
+In a nested model comparison, we found that a model including item as a fixed effect provides a significantly better fit, $\chi^2(7) = 16.4, p = 0.022$.
 
@@
 \label{fig:fitnessInteraction}
+\end{center}
+\end{figure}
+
+\begin{figure}[t!]
+\begin{center}
+\includegraphics[scale=.1]{images/itemwisefitness.jpeg}
+\caption{Item-wise error rates compared with label fitness ratios. Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest density intervals for error rates.}
+\label{fig:itemLevel}
 \end{center}

### 2016-01-29 experiment2/analysis/exp2Analysis.Rmd
Update analysis script

@@
 
-Set up error data
+Filter out bad participants
 
@@
                         length(bannedIDsExp2))
-```
-
-Remove banned ids from data
-
-```{r}
 errorDExp1 <- preErrorDExp1 %>%
@@
 
-Follow-up:
-
-Did error rates go down significantly?
+Follow-up: Did error rates go down significantly in Exp. 2?
 
@@
 totalPossible = c(exp1TotalPossible, exp2TotalPossible)
-prop.test(numErrors, totalPossible)
-
-# Diff from keysar (2003)?
-keysarNumErrors = 45
-keysarTotalPossible = 4 * 38
-numErrors = c(keysarNumErrors, exp2NumErrors$numErrors)
-totalPossible = c(keysarTotalPossible, exp2TotalPossible)
 prop.test(numErrors, totalPossible)
@@
 
-overinformativeCodes <- rbind(criticalMessagesExp1, criticalMessagesExp2) %>%
+rawOverinformativeCodes <- rbind(criticalMessagesExp1, criticalMessagesExp2) %>%
   filter(!(gameid %in% bannedIDsExp2)) %>%
   filter(!(gameid %in% bannedIDsExp1))
+
+overinformativeCodes = rawOverinformativeCodes %>%
+  group_by(exp, label, objectSet, workerid, referent) %>%
+  summarize(response = mean(response))
 
@@
     model = "twoway", type = "agreement")
+```
 
+Compare item-wise error rates and fitness ratios
+
+```{r}
+library(binom)
+Exp1Errs = itemInequalitiesExp1 %>%
+  mutate(response = errorRate) %>%
+  mutate(condition = "error") %>%
+  do(mutate(., lower = binom.confint(.$errorCount, .$total, 
+                                     method = "bayes")$lower,
+               upper = binom.confint(.$errorCount, .$total, 
+                                     method = "bayes")$upper)) %>%
+  select(objectSet, response, condition, lower, upper)
+
+stat = function(data, indices) {
+  d <- data[indices,]
+  applicability = d %>%   
+    group_by(objectSet) %>%
+    summarize(m = .5 * mean(distractor)/mean(target))
+  return(applicability$m)
+}
+
+Exp1Fitness = overinformativeCodes %>%
+  filter(exp == "Exp.1") %>%
+  spread(referent, response)
+
+res = boot(data = Exp1Fitness, statistic = stat, R = 1000)
+Exp1PlottingDF = data.frame(response = rep(0, 8),
+                        upper = rep(0,8),
+                        lower = rep(0,8),
+                        objectSet = 1:8,
+                        condition = rep("fitness", 8))
+for (i in 1:8) {
+  ci = boot.ci(res, type = "basic", index=i)
+  Exp1PlottingDF[i,]$response = res$t0[i]
+  Exp1PlottingDF[i,]$lower = ci$basic[4]
+  Exp1PlottingDF[i,]$upper = ci$basic[5]
+}
+
+Exp1ItemwiseForPlot = rbind(Exp1PlottingDF, Exp1Errs)
+g <- (ggplot(Exp1ItemwiseForPlot, aes(x = objectSet, y = response, 
+                                      group = condition, fill = condition))
+  + geom_bar(stat = "identity", position = "dodge")
+  + geom_errorbar(aes(ymax = upper, ymin = lower), 
+                  position = dodge, width = 0.25)
+  + ylab("ratio (scaled)"))
+g
+ggsave("../../writing/cogsci-submission/images/itemWiseFitness.jpeg")
+```
+
+Same thing for Exp 2
+
+```{r}
+library(binom)
+Exp2Errs = itemInequalitiesExp2 %>%
+  mutate(response = errorRate) %>%
+  mutate(condition = "error") %>%
+  do(mutate(., lower = binom.confint(.$errorCount, .$total, 
+                                     method = "bayes")$lower,
+               upper = binom.confint(.$errorCount, .$total, 
+                                     method = "bayes")$upper)) %>%
+  select(objectSet, response, condition, lower, upper)
+
+stat = function(data, indices) {
+  d <- data[indices,]
+  applicability = d %>%   
+    group_by(objectSet) %>%
+    summarize(m = .25 * mean(distractor)/mean(target))
+  return(applicability$m)
+}
+
+Exp2Fitness = overinformativeCodes %>%
+  filter(exp == "Exp.2") %>%
+  spread(referent, response)
+
+res = boot(data = Exp2Fitness, statistic = stat, R = 1000)
+Exp2PlottingDF = data.frame(response = rep(0, 8),
+                        upper = rep(0,8),
+                        lower = rep(0,8),
+                        objectSet = 1:8,
+                        condition = rep("fitness", 8))
+for (i in 1:8) {
+  ci = boot.ci(res, type = "basic", index=i)
+  Exp2PlottingDF[i,]$response = res$t0[i]
+  Exp2PlottingDF[i,]$lower = ci$basic[4]
+  Exp2PlottingDF[i,]$upper = ci$basic[5]
+}
+
+Exp2ItemwiseForPlot = rbind(Exp2PlottingDF, Exp2Errs)
+g <- (ggplot(Exp2ItemwiseForPlot, aes(x = objectSet, y = response, 
+                                      group = condition, fill = condition))
+  + geom_bar(stat = "identity", position = "dodge")
+  + geom_errorbar(aes(ymax = upper, ymin = lower), 
+                  position = dodge , width = 0.25)
+  + ylab("measure (z-scored)"))
+g
+```
+
+```{r}
+# within exp. correlations high
+cor(Exp1Errs$response, Exp1PlottingDF$response)
+cor(Exp2Errs$response, Exp2PlottingDF$response)
+
+# Cross exp. correlations low
+cor(Exp2Errs$response, Exp1PlottingDF$response)
+cor(Exp1Errs$response, Exp2PlottingDF$response)
 ```
@@
 
-itemWiseCodes <- overinformativeCodes %>% 
+itemWiseCodes <- rawOverinformativeCodes %>% 
   left_join(criticalErrors, by = c("gameid", "objectSet", "exp")) %>%
-  mutate(criticalError = ifelse(is.na(criticalError),0,1))
+  mutate(criticalError = ifelse(is.na(criticalError),0,1)) %>%
+  group_by(gameid, objectSet, referent, criticalError, exp) %>%
+  summarize(meanFitness = mean(response)) %>%
+  spread(referent, meanFitness)
+library(lme4)
 View(itemWiseCodes)
-mmRes <- glmer(criticalError ~ response * referent +  exp + (1 | objectSet) + (1 | gameid),
+mmRes1 <- glmer(criticalError ~ scale(target) + scale(distractor) + (1 | gameid), data = itemWiseCodes, family = binomial)
+summary(mmRes1)
+mmRes2 <- glmer(criticalError ~ scale(target) + scale(distractor) + factor(objectSet) + (1 | gameid),
       data = itemWiseCodes, family = binomial)
-plot(allEffects(mmRes))
-summary(mmRes)
+
+anova(mmRes1, mmRes2)
+
+newItemWiseCodes = itemWiseCodes %>% 
+  mutate(modelPreds = exp(predict(mmRes1))/(exp(predict(mmRes1))+1)) %>%
+  group_by(exp, objectSet) %>%
+  summarize(empError = mean(criticalError),
+            modelError = mean(modelPreds))
+
+(ggplot(newItemWiseCodes, aes(x = empError, y = modelError, color = exp)) 
+   + geom_point()
+   + geom_smooth(method = lm))
+
+# Shoudln't worry too much about convergence errors: appear to be false positives
+# https://groups.google.com/forum/#!topic/predicts_project/gVfnVxCSYog
+
+gg <- mmRes2@optinfo$derivs$grad
+hh <- mmRes2@optinfo$derivs$Hessian
+vv <- sqrt(diag(solve(hh/2)))
+print(summary(abs(gg*vv)))
+
+relgrad <- with(mmRes2@optinfo$derivs,solve(Hessian,gradient))
+print(max(abs(relgrad)))
+
 ```
@@
 
-joinedExp2 <- (right_join(mouDataExp2, mDExp2, by = c("gameid", "condition",
-                                                      "objectSet", 
-                                         "instructionNum", "attemptNum", "critical")))
+joinedExp2 <- mouDataExp2 %>%
+  right_join(mDExp2, by = c("gameid", "condition", "objectSet", "instructionNum",
+                            "attemptNum", "critical")) %>%
+  left_join(errorDExp2, by = c("gameid", "condition", "objectSet", "attemptNum",
+                               "instructionNum", "critical"))
+View(joinedExp2)
 ```
@@
   group_by(condition, exp) %>%
-  summarize(error = sem(percent),
+  summarize(error = ci95(percent),
             percent = mean(percent))
@@
 g <- (ggplot(combinedHeatMapForPlot, aes(x = exp, y = percent, 
-                                  group = condition, color = condition)) 
-      + geom_line(aes(linetype = condition), size = 2)
+                                  group = condition, fill = condition)) 
+      + geom_bar(position = dodge, stat= "identity")
       + geom_errorbar(aes(ymax = percent + error, 
-                          ymin = percent - error), size =2 ,
-                          width = 0.1)
-      + ylim(0, .45)
-      + ggtitle("Experiment 2 Hover-Time"))
+                          ymin = percent - error), position = dodge,
+                          width = 0.25)
+      + ylim(0, .5)
+      + ggtitle("Decreased hover-time on target"))
 g
-combinedRes = lm(percent ~ exp * cellType, data = combinedHeatMap)
+combinedRes = lm(percent ~ exp * condition, data = combinedHeatMap)
 summary(combinedRes)

### 2016-01-29 writing/cogsci-submission/cogsci_keysar.tex
Minor revisions to abstract

@@
 \begin{abstract}
-People are capable of accurately reasoning about the mental states of other people using theory of mind.
-An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take the speaker's perspective into account.
-This has been widely interpreted as a limitation on theory of mind use.
+Theory of mind is a powerful cognitive ability: by the age of six, people are capable of accurately reasoning about the beliefs and desires underlying other people's behavior.
+An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take the speaker's beliefs into account.
+This has been widely interpreted as evidence for limitations on theory of mind use.
 In this paper we argue that these apparent failures are in fact successes.
+Pragmatic language understanding requires sophisticated social reasoning about the speaker's goals and context.
 Through a minimal pair of replications comparing scripted vs.
 unscripted speakers, we show that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation.
-A listener who correctly reasons about the pragmatics of the task, thereby \emph{using} theory of mind, will rationally ignore whether an object is hidden to the speaker or not and will instead rely on the precision of the utterance to disambiguate.
-When an uncooperative confederate flouts these expectations, this strategy produces errors; when we allow participants to naturally interact, this strategy is justified and errors are reduced.
+Listeners who are justified in expecting more informative utterances are thus tricked by confederates into making systematic patterns of errors.
+When we allow participants to naturally interact, their expectations are justified and errors are reduced.
 \ndg{make a bit less technical}

### 2016-01-29 writing/cogsci-submission/cogsci_keysar.tex
a few questions about analysis

@@
 
-More dramatically, we find that patterns of errors in Exp.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
-Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the costly over-informative utterance remains ambiguous.
+We find that patterns of errors in Exp.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
+Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the more precise utterances used remain ambiguous.
 For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
@@
 In a nested model comparison, we found that a model including item as a fixed effect provides a significantly better fit, $\chi^2(7) = 16.4, p = 0.022$.
+\ndg{remind me (and reader) what this model comparison means?}
 
@@
 \includegraphics[scale=.1]{images/itemwisefitness.jpeg}
-\caption{Item-wise error rates compared with label fitness ratios. Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest density intervals for error rates.}
+\caption{Item-wise error rates compared with label fitness ratios. Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest density intervals for error rates.
+\ndg{is this expt 1 or 2 data? say ratio is scaled to unit interval in order to compare visually to error rates?}
+}
 \label{fig:itemLevel}

### 2016-01-29 writing/cogsci-submission/cogsci_keysar.tex
Various edits; better presentation of mouse-tracking

@@
 unscripted speakers, we show that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation.
-Listeners who are justified in expecting more informative utterances are thus tricked by confederates into making systematic patterns of errors.
-When we allow participants to naturally interact, their expectations are justified and errors are reduced.
-\ndg{make a bit less technical}
+Listeners who pragmatically expect more informative utterances are thus tricked by confederates into making systematic patterns of errors.
+When we allow participants to naturally interact, we find that their expectations are justified and errors are reduced.
+This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.
 
@@
 \begin{center}
-\includegraphics[scale = .5]{images/overall_view.png}
+\includegraphics[scale = .45]{images/overall_view.png}
 \end{center}
@@
 To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about the director's beliefs.
-For example, imagine a much larger measuring cup were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `the large cup' to mean this occluded cup (which the director couldn't possibly know about).
+For example, imagine a roll of tape were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `tape' to mean the occluded roll of tape (which the director couldn't possibly know about).
 Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.
@@
 
-While these results are compelling, the paradigm has been criticized from several different angles. 
+While these results are compelling, the paradigm has been criticized from a couple different angles. 
 First, \citeA{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
@@
 Second, the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility \cite{HannaTanenhausTrueswell03_CommonGroundPerspective}.
-Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g.
-the largest measuring cup) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}---perhaps people would refer to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results.
+%Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g.
+the roll of tape) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}---perhaps people would refer to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results.
 %Note that none of these criticisms invalidate the paradoxical result that listeners move an occluded object, they just suggest an alternative explanation.
@@
 In this paper, we offer an additional factor that helps account for Keysar's results. 
-Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations; given an utterance, a listener can then reason backward to the most plausible situation \cite{grice,clark,FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
+Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations; given an utterance, a listener can then reason backward to the most plausible situation \cite{Grice75_LogicConversation,Clark96_UsingLanguage,FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
 This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
@@
 In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item.
-
-We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment.
-We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study.
-We replicated the original finding, but noted a tendency of directors to be overinformative in unscripted trials.
-We then ran the same experiment without scripting the directions for critical trials, observing unconstrained director behavior.
-We observe much greater precision in unconstrained director utterances, which match targets much better than distractors, and better performance of the matchers.
-This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.
-
 \begin{table*}
@@
 \end{table*}
+
+We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment.
+We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study.
+We replicated the original finding, but noted a tendency of directors to be overinformative in unscripted trials.
+We then ran the same experiment without scripting the directions for critical trials, observing unconstrained director behavior.
+We observe much greater precision in unconstrained director utterances, which match targets much better than distractors, and better performance of the matchers.
+This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.
 
@@
 \subsection{Results and discussion}
-
 In Table \ref{table:mainResults} we show the error rates on critical items, and compare to the data from
@@
 Still,
-We note that several conditions have much higher error rates than others -- for example, 75\% of participants in the experimental condition of item 6 made an error (the ``whiteboard eraser'' vs.
+We note that several items have much higher error rates than others -- for example, 75\% of participants in the experimental condition of item 6 made an error (the ``whiteboard eraser'' vs.
 the ``pencil eraser'') while only 17\% of participants in item 8 made an error (the ``computer mouse'' vs.
 the ``toy mouse'').
-This suggests that the dependent variable highlighted in the original study (i.e.
+Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.
+We empirically substantiate this observation in our results for Exp.~2 below.
+
+This item-wise variability suggests that the dependent variable highlighted in the original study (i.e.
 ``percentage of participants who moved the critical item at least once'') is somewhat problematic: it could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item.
@@
 %This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
-\ndg{say something about bad items maybe being ones where utterance fits distractor better? (which we substantiate below..)}
 
@@
 We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
-We found that there is an interaction between cell type and condition on the percent of hover time spent in that cell, $b = -0.21, t(11) = -3.1, p = 0.003$.
-In baseline trials, people spent much more time in the target cell and much less time in the distractor cell than in the experimental condition (see Fig.
-\ref{fig:hover}).\footnote{Note that this analysis includes participants who actually made errors, since the data is too sparse to exclude them.} %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
+We found that people tended to spend less total time hovering over the target cell on critical experimental trials than on baseline trials, $b = -.11, t(20) = -1.26, p = 0.224$, although due to an underpowered sample and imprecise measurements\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.}, this result was not significant (see Fig.
+\ref{fig:hover}).
+%accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
 
@@
 We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
-By the reasoning above we expect to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
+By the reasoning above we expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
 
@@
 
- \begin{figure}[bt!]
+ \begin{figure}[b!]
 \begin{center}
-\includegraphics[scale=.8]{images/mousetracking.pdf}
+\includegraphics[scale=.5]{images/mousetracking.pdf}
 \caption{Mean percentage of time spent hovering over target cell in the experimental and baseline conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are 95\% confidence intervals.}
@@
 
-%\begin{figure}[b!]
-%\begin{center}
-%\includegraphics[scale=.75]{images/exp2MouseTracking.pdf}
-%\caption{Mean percentage of time spent hovering over the distractor and target cells in the experimental and baseline conditions.
-Error bars are 95\% confidence intervals.
-\todo[inline]{TODO: Collapse mouse-tracking plots to a single figure showing some interaction}}
-%\label{fig:exp2hover}
-%\end{center}
-%\end{figure}
-
 \subsection{Results}
@@
 
-\todo[inline]{rdh: I don't like these mouse-tracking analyses -- I feel like the way I'm computing hover-time is sloppy and might be skewing the results (i.e.
-counting number of `move' events rather than actually looking at start and end times).
-So I could spend some time tomorrow coding up the better way\dots}
-When we conducted a mouse-tracking analysis identical to the one reported for Exp.~1, we again found a significant interaction between condition and cell type on hover time, $b = -0.18, t(108) = -5.2, p < 0.001$.
-However, the effect appears to be attenuated: in experimental trials, participants spend significantly more time hovering over the target than the distractor (see Figure \ref{fig:hover}) whereas in Exp.~1, participants spent roughly the same amount of time hovering over each.
-Due to the relatively small sample size of Exp.~1, however, we do not have the power to show this effect statistically.
+When we conducted a mouse-tracking analysis identical to the one reported for Exp.~1, we again found a significant decrease in total hover time over the target in experimental vs.
+baseline trials $b = 0.07, t(54) = -2.1, p = 0.045$, indicating that the presence of a hidden distractor still interfered with participants directly choosing the target.
+While the effect appears to be attenuated compared to Exp.~1 (see Figure \ref{fig:hover}), there is no significant interaction of experiment and condition on hover time $b = 0.04, t(74) = 0.45, p = 0.657$.
+Again, given the low sample size in Exp.~1, the noise from error, and the noise from counting samples instead of times, this result is limited.
+\begin{figure}[t!]
+\begin{center}
+\includegraphics[scale=.5]{images/itemwisefitness.pdf}
+\caption{Item-wise error rates compared with label fitness ratios, for Exp.~1.
+Fitness ratios are scaled to the unit interval in order to visually compare to error rates.
+Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest density intervals for error rates.}
+\label{fig:itemLevel}
+\end{center}
+\end{figure}
 
@@
 We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$.
-In a nested model comparison, we found that a model including item as a fixed effect provides a significantly better fit, $\chi^2(7) = 16.4, p = 0.022$.
-\ndg{remind me (and reader) what this model comparison means?}
+Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
 
@@
 \begin{center}
-\includegraphics[scale=.18]{images/fitnessInteraction.jpeg}
+\includegraphics[scale=.5]{images/fitnessInteraction.pdf}
 \caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
 \label{fig:fitnessInteraction}
-\end{center}
-\end{figure}
-
-\begin{figure}[t!]
-\begin{center}
-\includegraphics[scale=.1]{images/itemwisefitness.jpeg}
-\caption{Item-wise error rates compared with label fitness ratios. Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest density intervals for error rates.
-\ndg{is this expt 1 or 2 data? say ratio is scaled to unit interval in order to compare visually to error rates?}
-}
-\label{fig:itemLevel}
 \end{center}
@@
 
-\ndg{concluding paragraph? Overall, we both replicated the finding that listeners neglect the speakers epistemic state, and explained why this may be so (and why errors are uneven over items) ...
-Or maybe we don't need anything else.
-}
-
 \section{Acknowledgements}

### 2016-01-30 writing/cogsci-submission/cogsci_keysar.tex
Add some text to chatbox in Fig. 1

@@
 \begin{center}
-\includegraphics[scale = .45]{images/overall_view.png}
+\includegraphics[scale = .4]{images/overall_view.png}
 \end{center}
@@
 We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
-We found that people tended to spend less total time hovering over the target cell on critical experimental trials than on baseline trials, $b = -.11, t(20) = -1.26, p = 0.224$, although due to an underpowered sample and imprecise measurements\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.}, this result was not significant (see Fig.
+We found that people tended to spend less total time hovering over the target cell on critical experimental trials than on baseline trials, $b = -.11, t(20) = -1.26, p = 0.224$, although perhaps due to an underpowered sample and imprecise measurements\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.}, this result was not significant (see Fig.
 \ref{fig:hover}).

### 2016-01-30 writing/cogsci-submission/cogsci_keysar.tex
Refer to expts using consistent abbreviation

@@
 \hline
-& Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2 & Orig. & Exp. 1 & Exp. 2  \\
+& Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2  \\
 Experimental & 71 & 93 & 61 & 46 & 57 & 32 & 30 & 43 & 24\\
@@
 
-\section{Exp.~1: Scripted Replication}
+\section{Expt.~1: Scripted Replication}
 \label{sec:Exp1}
@@
  \hline\hline
-\multirow{2}{.75cm}{Exp.~1}  & \# incorrect & 0 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
+\multirow{2}{.75cm}{Expt.~1}  & \# incorrect & 0 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
                                                & \# correct & 6 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
  \hline\hline
-\multirow{2}{.75cm}{Exp.~2} & \# incorrect & 0 & 1 & 1 & 3 & 9 & 7 & 1 & 5 \\
+\multirow{2}{.75cm}{Expt.~2} & \# incorrect & 0 & 1 & 1 & 3 & 9 & 7 & 1 & 5 \\
                                               & \# correct   & 12&14&11&13&7 &8   &12& 8 \\
@@
 Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.
-We empirically substantiate this observation in our results for Exp.~2 below.
+We empirically substantiate this observation in our results for Expt.~2 below.
 
@@
 That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
-We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
+We tested this prediction in Expt.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
 By the reasoning above we expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
@@
 
-\section{Exp.~2: Unscripted Replication}
+\section{Expt.~2: Unscripted Replication}
 \label{sec:Exp2}
@@
 
-We recruited 64 participants (32 pairs) from Amazon Mechanical Turk, roughly doubling the sample size from Exp.~1.
+We recruited 64 participants (32 pairs) from Amazon Mechanical Turk, roughly doubling the sample size from Expt.~1.
 All participants were from the U.S.
@@
 
-Everything was the same as Exp.~1, except we did not use scripted messages for critical instructions.
+Everything was the same as Expt.~1, except we did not use scripted messages for critical instructions.
 
@@
 
-Error rates are reported in Table \ref{table:mainResults}, alongside the results from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our scripted replication in Exp.~1.
-Participants never moved the hidden object in the baseline condition, and total error rates for experimental trials are significantly lower than the rates found in Exp.~1, $\chi^2(1) = 5.35, p = 0.02$.
+Error rates are reported in Table \ref{table:mainResults}, alongside the results from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our scripted replication in Expt.~1.
+Participants never moved the hidden object in the baseline condition, and total error rates for experimental trials are significantly lower than the rates found in Expt.~1, $\chi^2(1) = 5.35, p = 0.02$.
 
-We find that patterns of errors in Exp.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
+We find that patterns of errors in Expt.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
 Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the more precise utterances used remain ambiguous.
@@
 
-When we conducted a mouse-tracking analysis identical to the one reported for Exp.~1, we again found a significant decrease in total hover time over the target in experimental vs.
+When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we again found a significant decrease in total hover time over the target in experimental vs.
 baseline trials $b = 0.07, t(54) = -2.1, p = 0.045$, indicating that the presence of a hidden distractor still interfered with participants directly choosing the target.
-While the effect appears to be attenuated compared to Exp.~1 (see Figure \ref{fig:hover}), there is no significant interaction of experiment and condition on hover time $b = 0.04, t(74) = 0.45, p = 0.657$.
-Again, given the low sample size in Exp.~1, the noise from error, and the noise from counting samples instead of times, this result is limited.
+While the effect appears to be attenuated compared to Expt.~1 (see Figure \ref{fig:hover}), there is no significant interaction of experiment and condition on hover time $b = 0.04, t(74) = 0.45, p = 0.657$.
+Again, given the low sample size in Expt.~1, the noise from error, and the noise from counting samples instead of times, this result is limited.
 \begin{figure}[t!]
@@
 \includegraphics[scale=.5]{images/itemwisefitness.pdf}
-\caption{Item-wise error rates compared with label fitness ratios, for Exp.~1.
+\caption{Item-wise error rates compared with label fitness ratios, for Expt.~1.
 Fitness ratios are scaled to the unit interval in order to visually compare to error rates.
@@
 \ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$.
-In Exp.~1, the scripted label fit the hidden distractor just as well as the target, but in Exp.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
+In Expt.~1, the scripted label fit the hidden distractor just as well as the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
 In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.
 
-Does differential informativity of scripted utterances account for the variability across items that we noted in Exp.~1?
+Does differential informativity of scripted utterances account for the variability across items that we noted in Expt.~1?
 In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
-We find a correlation of $r = 0.76$ between these measures for Exp.~1, and a correlation of $r = 0.53$ for Exp.~2.
+We find a correlation of $r = 0.76$ between these measures for Expt.~1, and a correlation of $r = 0.53$ for Expt.~2.
 The cross-experiment correlations of error rates from one experiment and label fitness ratios from the other, however, is only $r = 0.07$, suggesting that item-wise variability across the two experiments is driven by relative informativity of speaker utterances.
 
-After separately establishing that matchers in Exp.~2 make fewer mistakes, that directors produce more informative utterances, and that informativity captures item-wise variability in error rates, we tested the link between these effects: does label informativity generally predict errors on critical trials?
+After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors produce more informative utterances, and that informativity captures item-wise variability in error rates, we tested the link between these effects: does label informativity generally predict errors on critical trials?
 We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
@@
 In Expt.~2, we found that speakers did naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor.
-For example, no speaker in Expt.~2 produced ``the bottom block,'' which was used as a critical instruction in Exp.~1.
+For example, no speaker in Expt.~2 produced ``the bottom block,'' which was used as a critical instruction in Expt.~1.
 Instead, they said ``the bluish block with a B'' or ``block with the blue writing,'' which relied on less confusable perceptual features and thereby decreased error rates.
@@
 \small We're grateful to Boaz Keysar for providing select materials for our replication.
-Exp.~1 was originally conducted under the supervision of Michael Frank, with early input from Desmond Ong.
+Expt.~1 was originally conducted under the supervision of Michael Frank, with early input from Desmond Ong.
 This work was supported by ONR grants N00014-13-1-0788 and N00014-13-1-0287,  and a James S.

### 2016-01-30 experiment2/analysis/exp2Analysis.Rmd
Use mouse-tracking analysis more similar to keysar

@@
 
-View(overinformativeCodes)
 summary(lmer(response ~ exp * referent + (1 | workerid) + (1 | objectSet), 
@@
   geom_errorbar(aes(ymin = lowerCI, ymax = upperCI), position = dodge, width = 0.25) +
-  theme_bw() +
-  ylab("mean fitness of label")
-ggsave("../../writing/cogsci-submission/images/fitnessInteraction.jpeg")
+  theme_bw(base_size = 16) +
+  ylab("mean fitness of label") + 
+  ggtitle("Speaker Informativity")
+ggsave("../../writing/cogsci-submission/images/fitnessInteraction.pdf")
 
@@
   mutate(response = errorRate) %>%
-  mutate(condition = "error") %>%
+  mutate(measure = "error") %>%
   do(mutate(., lower = binom.confint(.$errorCount, .$total, 
@@
                         objectSet = 1:8,
-                        condition = rep("fitness", 8))
+                        measure = rep("fitness", 8))
 for (i in 1:8) {
@@
 g <- (ggplot(Exp1ItemwiseForPlot, aes(x = objectSet, y = response, 
-                                      group = condition, fill = condition))
+                                      group = measure, fill = measure))
   + geom_bar(stat = "identity", position = "dodge")
@@
                   position = dodge, width = 0.25)
+  + theme_bw(base_size = 16) 
+  + ggtitle("Exp.1 Item-Wise Error")
   + ylab("ratio (scaled)"))
 g
-ggsave("../../writing/cogsci-submission/images/itemWiseFitness.jpeg")
+ggsave("../../writing/cogsci-submission/images/itemWiseFitness.pdf")
 ```
@@
                   position = dodge , width = 0.25)
+  + theme_bw(base_size = 16)
   + ylab("measure (z-scored)"))
@@
   spread(referent, meanFitness)
+
 library(lme4)
-View(itemWiseCodes)
-mmRes1 <- glmer(criticalError ~ scale(target) + scale(distractor) + (1 | gameid), data = itemWiseCodes, family = binomial)
+mmRes1 <- glmer(criticalError ~ factor(objectSet) + (1 | gameid), data = itemWiseCodes, family = binomial)
 summary(mmRes1)
@@
 
-newItemWiseCodes = itemWiseCodes %>% 
-  mutate(modelPreds = exp(predict(mmRes1))/(exp(predict(mmRes1))+1)) %>%
-  group_by(exp, objectSet) %>%
-  summarize(empError = mean(criticalError),
-            modelError = mean(modelPreds))
-
-(ggplot(newItemWiseCodes, aes(x = empError, y = modelError, color = exp)) 
-   + geom_point()
-   + geom_smooth(method = lm))
+# newItemWiseCodes = itemWiseCodes %>% 
+#   mutate(modelPreds = exp(predict(mmRes1))/(exp(predict(mmRes1))+1)) %>%
+#   group_by(exp, objectSet) %>%
+#   summarize(empError = mean(criticalError),
+#             modelError = mean(modelPreds))
+# 
+# (ggplot(newItemWiseCodes, aes(x = empError, y = modelError, color = exp)) 
+#    + geom_point()
+#    + geom_smooth(method = lm))
 
@@
   right_join(mDExp2, by = c("gameid", "condition", "objectSet", "instructionNum",
-                            "attemptNum", "critical")) %>%
-  left_join(errorDExp2, by = c("gameid", "condition", "objectSet", "attemptNum",
-                               "instructionNum", "critical"))
-View(joinedExp2)
+                            "attemptNum", "critical")) #%>%
+#   left_join(errorDExp2, by = c("gameid", "condition", "objectSet", "attemptNum",
+#                                "instructionNum", "critical"))
 ```
@@
    filter(distractorY == begDistrY) 
-View(dExp2)
 ```
@@
 
-t1 <- lm(percent ~ condition * cellType, data = heatMapForLMExp2)
+t1 <- lm(percent ~ condition , data = heatMapForLMExp2 %>% filter(cellType == "target"))
 summary(t1)
@@
      + ylim(0, .45)
+     + theme_bw(base_size = 16)
       + ggtitle("Experiment 2 Hover-Time"))
@@
       + ylim(0, .5)
-      + ggtitle("Decreased hover-time on target"))
+      + theme_bw(base_size = 16)
+      + ggtitle("Target Hover-Time"))
 g
+ggsave("../../writing/cogsci-submission/images/mousetracking.pdf")
+
 combinedRes = lm(percent ~ exp * condition, data = combinedHeatMap)

### 2016-01-30 writing/cogsci-submission/cogsci_keysar.tex
Shrunk abstract, formatting

@@
 \begin{abstract}
-Theory of mind is a powerful cognitive ability: by the age of six, people are capable of accurately reasoning about the beliefs and desires underlying other people's behavior.
-An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take the speaker's beliefs into account.
-This has been widely interpreted as evidence for limitations on theory of mind use.
+Theory of mind is a powerful cognitive ability: by the age of six, people are capable of accurately reasoning about others' beliefs and desires.
+An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take a speaker's beliefs into account, revealing limitations on theory of mind.
 In this paper we argue that these apparent failures are in fact successes.
-Pragmatic language understanding requires sophisticated social reasoning about the speaker's goals and context.
 Through a minimal pair of replications comparing scripted vs.
-unscripted speakers, we show that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation.
-Listeners who pragmatically expect more informative utterances are thus tricked by confederates into making systematic patterns of errors.
-When we allow participants to naturally interact, we find that their expectations are justified and errors are reduced.
+unscripted speakers, we show that critical utterances used by Keysar and colleagues are uncooperative: they are less informative than what a speaker would actually produce in that situation.
+When we allow participants to naturally interact, we find that listener expectations are justified and errors are reduced.
 This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.
@@
 \begin{center}
-\includegraphics[scale = .4]{images/overall_view.png}
+\includegraphics[scale = .37]{images/overall_view.png}
 \end{center}
@@
 
-The argument of Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate.
+The argument offered by Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate.
 The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects (see Fig.
@@
 Critically, some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
-To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about the director's beliefs.
+To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about which objects were shared and which were private.
 For example, imagine a roll of tape were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `tape' to mean the occluded roll of tape (which the director couldn't possibly know about).
@@
 
-While these results are compelling, the paradigm has been criticized from a couple different angles. 
-First, \citeA{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
+While these results are compelling, the paradigm has been criticized from few different angles. 
+\citeA{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
 the hidden roll of tape vs.
@@
 % Although relative fit was not empirically evaluated prior to our study,
-Second, the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility \cite{HannaTanenhausTrueswell03_CommonGroundPerspective}.
+Moreover, \citeA{HannaTanenhausTrueswell03_CommonGroundPerspective} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
 %Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g.
@@
 In this paper, we offer an additional factor that helps account for Keysar's results. 
-Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations; given an utterance, a listener can then reason backward to the most plausible situation \cite{Grice75_LogicConversation,Clark96_UsingLanguage,FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
+Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations.
+Given an utterance, a listener can then reason backward to the most plausible situation \cite{Grice75_LogicConversation,Clark96_UsingLanguage,FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
 This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
@@
 \caption{Side-by-side comparison of error rates in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications.
-The first section shows the percentage of \emph{participants} attempting to move the occluded distractor at least one of the four possible cases.
-The second section shows the percentage attempting to move it at least twice.
+The first two sections show the percentage of \emph{participants} attempting to move the occluded distractor at least once, or twice, of the four possible cases.
 The third section shows the percentage of \emph{all experimental trials} that the participant actually tried to move the occluded object.
@@
 We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study.
-We replicated the original finding, but noted a tendency of directors to be overinformative in unscripted trials.
-We then ran the same experiment without scripting the directions for critical trials, observing unconstrained director behavior.
-We observe much greater precision in unconstrained director utterances, which match targets much better than distractors, and better performance of the matchers.
+We replicated the original finding, but noted a tendency of directors to be overinformative in unscripted filler trials.
+We then ran the same experiment without using any scripted instructions, observing unconstrained director utterances.
+We found much greater precision in unconstrained director utterances, which match targets much better than distractors, and better performance of the matchers.
 This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.
@@
 This paradigm differs from those used by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways.
-First, participants were not seated across from each other at a table: they each saw a views of the 4 x 4 grid on screen and communicated via a text box.
+First, participants were not seated across from each other at a table: they each saw a view of the 4 x 4 grid on their screen and communicated via a text box.
 Second, we did not use a trained confederate.
@@
 \begin{center}
-\begin{tabular}{ m{.75cm} | m{1.5cm} | m{1.5cm} | m{1.5cm} |  m{1.25cm} | m{1.6cm} | m{1.4cm} | m{1.3cm} | m{1.3cm} | m{1.5cm} |}
+\begin{tabular}{ m{1cm} | m{1.5cm} | m{1.5cm} | m{1.5cm} |  m{1.25cm} | m{1.6cm} | m{1.4cm} | m{1.1cm} | m{1.3cm} | m{1.3cm} |}
 & & Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
@@
 \end{tabular}
+\vspace{-.5cm}
 \caption{Item-wise error rates for critical trials}
+\vspace{-.25cm}
 \label{table:ItemWise}
@@
 If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt.
-We took the set of all mouse position samples in the decision window (collected approximately every 15ms), and measured what proportion were located inside the cell containing the target vs.
-the cell containing the distractor.
+We took the set of all mouse position samples in the decision window (collected approximately every 15ms), and measured what proportion were located inside the cell containing the target.
 We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
@@
 By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items.
-Informally, we noted a tendency toward, possibly unnecessary, precision in descriptions.
-Instead of ``move the stuffed animal down'', participants say ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants say ``move the red airplane to the right.''
+Informally, we noted a tendency toward additional, possibly unnecessary, precision in descriptions.
+Instead of ``move the stuffed animal down'', participants said ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.''
 Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions.
@@
 \includegraphics[scale=.5]{images/mousetracking.pdf}
+\vspace{-.5cm}
+
 \caption{Mean percentage of time spent hovering over target cell in the experimental and baseline conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are 95\% confidence intervals.}
@@
 For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
-``Hair brush'' is already a label at the sub-class level.
+Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' from Expt.~1, but the two objects were still confusable at the sub-class level.
 If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.
 
-When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we again found a significant decrease in total hover time over the target in experimental vs.
+When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we again found a decrease in total hover time over the target in experimental vs.
 baseline trials $b = 0.07, t(54) = -2.1, p = 0.045$, indicating that the presence of a hidden distractor still interfered with participants directly choosing the target.
-While the effect appears to be attenuated compared to Expt.~1 (see Figure \ref{fig:hover}), there is no significant interaction of experiment and condition on hover time $b = 0.04, t(74) = 0.45, p = 0.657$.
-Again, given the low sample size in Expt.~1, the noise from error, and the noise from counting samples instead of times, this result is limited.
+While the effect appears to be slightly attenuated compared to Expt.~1 (see Figure \ref{fig:hover}), there is no significant interaction of experiment and condition on hover time $b = 0.04, t(74) = 0.45, p = 0.657$.
+Again, given the low sample size in Expt.~1, the noise from error, and the noise from counting samples instead of differences in timestamps, this result is limited.
 \begin{figure}[t!]
@@
 \includegraphics[scale=.5]{images/itemwisefitness.pdf}
+\vspace{-.25cm}
 \caption{Item-wise error rates compared with label fitness ratios, for Expt.~1.
@@
 Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest density intervals for error rates.}
+\vspace{-.25cm}
 \label{fig:itemLevel}
@@
 \ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$.
-In Expt.~1, the scripted label fit the hidden distractor just as well as the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
+In Expt.~1, the scripted label fit the hidden distractor just as well or better than the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
 In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.
@@
 In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
-We find a correlation of $r = 0.76$ between these measures for Expt.~1, and a correlation of $r = 0.53$ for Expt.~2.
-The cross-experiment correlations of error rates from one experiment and label fitness ratios from the other, however, is only $r = 0.07$, suggesting that item-wise variability across the two experiments is driven by relative informativity of speaker utterances.
+We find a correlation of $r = 0.76$ between these two measures for Expt.~1, and a correlation of $r = 0.53$ for Expt.~2.
+Comparing error rates from one experiment with label fitness ratios from the other, however, yields much weaker correlations ($r = 0.07$), suggesting that item-wise variability across the two experiments is driven by relative informativity of speaker utterances.
 
-After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors produce more informative utterances, and that informativity captures item-wise variability in error rates, we tested the link between these effects: does label informativity generally predict errors on critical trials?
+After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials?
 We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
@@
 \includegraphics[scale=.5]{images/fitnessInteraction.pdf}
+\vspace{-.25cm}
 \caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
+\vspace{-.25cm}
 \label{fig:fitnessInteraction}
@@
 Indeed, we found that the extremely heterogeneous pattern of errors across items was well explained by the relative fit of the utterance to target and distractor.
-This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness. 
+This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness rather than consideration of occluded vs. mutual visible items. 
 %
@@
 It is worth noting several significant differences between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-% that prevented a direct replication. 
 The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room.
@@
 A related difference is our decision not to use a confederate.
-While a confederate is useful for reducing variation across instances of the experiment, we see several downsides.
+While confederates are useful for reducing variation across instances of the experiment, we see several downsides.
 Beyond the difficulties of conducting large-scale experiments with confederates, it is difficult to exactly replicate all the subtleties of the confederate's behavior that might influence results.

### 2016-05-12 analysis/cogSci_analysis.Rmd
Compile analysis script

--- 2016-01-30 experiment2/analysis/exp2Analysis.Rmd
@@
 ```{r}
+eDExp1 <- errorDExp1 %>%
+  filter(attemptNum == 0) %>%
+  group_by(gameid, condition) %>%
+  summarize(n = sum(criticalError)) %>%
+  mutate(cond1 = as.numeric(n >= 1)) %>%
+  mutate(cond2 = as.numeric(n >= 2))
+
 eDExp2 <- errorDExp2 %>%
@@
 ```{r}
+numPlayersPerConditionExp1 = mouseDExp1 %>% 
+  filter(critical == 1) %>% 
+  filter(attemptNum == 0) %>% 
+  group_by(gameid, condition, objectSet) %>% 
+  summarize() %>% 
+  group_by(condition, objectSet) %>% 
+  summarize(total = n()) %>% 
+  filter(condition == "exp")
+
 numPlayersPerConditionExp2 = mouseDExp2 %>% 
@@
 ```{r}
+itemInequalitiesExp1 <- errorDExp1 %>%
+  filter(condition == "exp") %>%
+  filter(criticalError == 1) %>%
+  filter(attemptNum == 0) %>%
+  group_by(objectSet) %>%
+  tally() %>%
+  right_join(numPlayersPerConditionExp1, by = c('objectSet')) %>%
+  mutate(objectSet = objectSet,
+         errorCount = n, 
+         correctCount = total - n) %>%
+  mutate(errorRate = errorCount / (errorCount + correctCount)) %>%
+  select(objectSet, errorCount, correctCount, errorRate, total)
+itemInequalitiesExp1[1,2:4] = c(0, itemInequalitiesExp1[1,]$total, 0)
+
+itemWiseTestExp1 <- chisq.test(itemInequalitiesExp1[-c(1,4,5)])
+
 itemInequalitiesExp2 <- errorDExp2 %>%
@@
                                      method = "bayes")$upper)) %>%
-  select(objectSet, response, condition, lower, upper)
+  select(objectSet, response, measure, lower, upper)
 
@@
 
-Same thing for Exp 2
+Alternatively, could make a scatter plot?
 
@@
 library(binom)
-Exp2Errs = itemInequalitiesExp2 %>%
-  mutate(response = errorRate) %>%
-  mutate(condition = "error") %>%
-  do(mutate(., lower = binom.confint(.$errorCount, .$total, 
+Exp1Errs = itemInequalitiesExp1 %>%
+  do(mutate(., errorLower = binom.confint(.$errorCount, .$total, 
                                      method = "bayes")$lower,
-               upper = binom.confint(.$errorCount, .$total, 
+               errorUpper = binom.confint(.$errorCount, .$total, 
                                      method = "bayes")$upper)) %>%
-  select(objectSet, response, condition, lower, upper)
+  select(objectSet, errorRate, errorLower, errorUpper)
 
@@
     group_by(objectSet) %>%
-    summarize(m = .25 * mean(distractor)/mean(target))
+    summarize(m = mean(distractor)/mean(target))
+  return(applicability$m)
+}
+
+Exp1Fitness = overinformativeCodes %>%
+  filter(exp == "Exp.1") %>%
+  spread(referent, response)
+
+res = boot(data = Exp1Fitness, statistic = stat, R = 1000)
+Exp1PlottingDF = data.frame(fitness = rep(0, 8),
+                        fitnessUpper = rep(0,8),
+                        fitnessLower = rep(0,8),
+                        objectSet = 1:8)
+for (i in 1:8) {
+  ci = boot.ci(res, type = "basic", index=i)
+  Exp1PlottingDF[i,]$fitness = res$t0[i]
+  Exp1PlottingDF[i,]$fitnessLower = ci$basic[4]
+  Exp1PlottingDF[i,]$fitnessUpper = ci$basic[5]
+}
+
+Exp1ItemwiseForPlot = left_join(Exp1Errs, Exp1PlottingDF, by = c("objectSet"))
+g <- (ggplot(Exp1ItemwiseForPlot, aes(x = fitness, y = errorRate))
+  + geom_point() 
+  + geom_errorbar(aes(ymax = errorUpper, ymin = errorLower), 
+                  width = 0.05)
+  + geom_smooth(method = "lm")
+  + theme_bw(base_size = 16) 
+  + ggtitle("Exp.1 Item-Wise Error")
+  + ylab("ratio (scaled)"))
+g
+#ggsave("../../writing/cogsci-submission/images/itemWiseFitness.pdf")
+```
+Same thing for Exp 2
+
+```{r}
+library(binom)
+binomStat = stat = function(data, indices) {
+  d <- data[indices,]
+  prop = d %>%   
+    group_by(objectSet) %>%
+    summarize(m = mean(distractor)/mean(target))
+  return(applicability$m)
+}
+
+Exp2Errs = itemInequalitiesExp2 %>%
+  do(mutate(., errorLower = binom.confint(.$errorCount, .$total, 
+                                     method = "bayes")$lower,
+               errorUpper = binom.confint(.$errorCount, .$total, 
+                                     method = "bayes")$upper)) %>%
+  select(objectSet, errorRate, errorLower, errorUpper)
+
+stat = function(data, indices) {
+  d <- data[indices,]
+  applicability = d %>%   
+    group_by(objectSet) %>%
+    summarize(m = mean(distractor)/mean(target))
   return(applicability$m)
@@
 res = boot(data = Exp2Fitness, statistic = stat, R = 1000)
-Exp2PlottingDF = data.frame(response = rep(0, 8),
-                        upper = rep(0,8),
-                        lower = rep(0,8),
-                        objectSet = 1:8,
-                        condition = rep("fitness", 8))
+
+Exp2PlottingDF = data.frame(fitness = rep(0, 8),
+                        fitnessUpper = rep(0,8),
+                        fitnessLower = rep(0,8),
+                        objectSet = 1:8)
 for (i in 1:8) {
   ci = boot.ci(res, type = "basic", index=i)
-  Exp2PlottingDF[i,]$response = res$t0[i]
-  Exp2PlottingDF[i,]$lower = ci$basic[4]
-  Exp2PlottingDF[i,]$upper = ci$basic[5]
+  Exp2PlottingDF[i,]$fitness = res$t0[i]
+  Exp2PlottingDF[i,]$fitnessLower = ci$basic[4]
+  Exp2PlottingDF[i,]$fitnessUpper = ci$basic[5]
 }
 
-Exp2ItemwiseForPlot = rbind(Exp2PlottingDF, Exp2Errs)
-g <- (ggplot(Exp2ItemwiseForPlot, aes(x = objectSet, y = response, 
-                                      group = condition, fill = condition))
-  + geom_bar(stat = "identity", position = "dodge")
-  + geom_errorbar(aes(ymax = upper, ymin = lower), 
-                  position = dodge , width = 0.25)
+Exp2ItemwiseForPlot = left_join(Exp2PlottingDF, Exp2Errs, by = c("objectSet")) %>%
+  mutate(experiment = "Exp. 2") %>%
+  rbind(Exp1ItemwiseForPlot %>% mutate(experiment = "Exp. 1"))
+
+g <- (ggplot(Exp2ItemwiseForPlot, aes(x = fitness, y = errorRate))
+  + geom_point(aes(color = experiment))
+  + geom_errorbarh(aes(xmax = fitnessUpper, xmin = fitnessLower,
+                       color = experiment), 
+                  height = 0.01)
+  + geom_errorbar(aes(ymax = errorUpper, ymin = errorLower,
+                     color = experiment), 
+                width = 0.05)
+  + geom_smooth(method = "lm",colour="black")
   + theme_bw(base_size = 16)
-  + ylab("measure (z-scored)"))
+  + theme(aspect.ratio=1)
+  + ylab("Error rate")
+  + xlab("Label fitness of distractor vs. target"))
 g
+ggsave("~/Repos/keysar_replication/writing/cogsci-revision/images/itemWiseFitness.pdf")
 ```
@@
 ```{r}
+summary(lm(errorRate ~ fitness * experiment, data = Exp2ItemwiseForPlot))
+
 # within exp. correlations high
@@
 ```
-
-```{r}
-# heatMapForPlot$new_labels = as.factor(sapply(X = heatMapForPlot$objectSet, 
-#                               FUN = function(v) {return(paste("Item", v))}))
-# 
-# g <- (ggplot(heatMapForPlot, aes(x = mouseX, y = mouseY, color = condition)) +
-#       geom_vline(xintercept = c(0, 150,300,450, 600)) +
-#       geom_hline(yintercept = c(0, 150,300,450, 600)) +
-#       geom_point() +
-#       geom_point(aes(x = begTargetX, y = begTargetY), size = 20,
-#                  shape = 4, color = "black", show_guide = FALSE) +
-#       geom_point(aes(x = begDistrX, y = begDistrY), size = 20,
-#                  shape = 4, color = "grey50", show_guide = FALSE) + 
-#       theme(panel.grid.major = element_blank(), 
-#             panel.grid.minor = element_blank()) +
-#       xlim(0, 600) +
-#       ylim(0, 600) +
-#       theme(aspect.ratio = 1) +
-#       facet_wrap(~ new_labels, nrow = 2))
-# g
-```

### 2016-05-12 experiment2/analysis/exp2Analysis.Rmd
Merged stashed changes

@@
 library(lmerTest)
+
+d <- d.b %>%
+  mutate(trait = sentence %in% c('Decisive', 'Competitive', 'Self-reliant', 
+                                 'Willing to take risks' , 'Ambitious', 'Daring' , 
+                                 'Adventurous' , 'Courageous')) %>%
+  mutate(traitType = factor(ifelse(trait == TRUE, "Masculine", "Feminine"))) %>%
+
 ```
@@
 ```{r}
+itemInequalitiesExp1 <- errorDExp1 %>%
+  filter(condition == "exp") %>%
+  filter(criticalError == 1) %>%
+  filter(attemptNum == 0) %>%
+  group_by(objectSet) %>%
+  tally() %>%
+  right_join(numPlayersPerConditionExp1, by = c('objectSet')) %>%
+  mutate(objectSet = objectSet,
+         errorCount = n, 
+         correctCount = total - n) %>%
+  mutate(errorRate = errorCount / (errorCount + correctCount)) %>%
+  select(objectSet, errorCount, correctCount, errorRate, total)
+itemInequalitiesExp1[1,2:4] = c(0, itemInequalitiesExp1[1,]$total, 0)
+
 itemInequalitiesExp2 <- errorDExp2 %>%
@@
 library(binom)
+
 Exp1Errs = itemInequalitiesExp1 %>%
@@
 Exp1ItemwiseForPlot = rbind(Exp1PlottingDF, Exp1Errs)
+
 g <- (ggplot(Exp1ItemwiseForPlot, aes(x = objectSet, y = response, 
@@
 ```{r}
+mouDataExp1 <- mouseDExp1 %>%
+  filter(attemptNum == 0)
+
+mDExp1 <- messageDExp1 %>% 
+  filter(attemptNum == 0) %>%
+  filter(sender == "director") %>%
+  group_by(gameid, objectSet, instructionNum) %>%
+  mutate(messageStamp = first(time)) %>%
+  select(gameid, condition, attemptNum, instructionNum, critical, 
+         objectSet,sender,contents,messageStamp) %>%
+  distinct(gameid,contents)
+
+joinedExp1 <- mouDataExp1 %>% 
+  right_join(mDExp1, by = c("gameid", "condition", "objectSet", 
+                             "instructionNum", "attemptNum", "critical"))
+
 mouDataExp2 <- mouseDExp2 %>%
@@
 joinedExp2 <- mouDataExp2 %>%
-  right_join(mDExp2, by = c("gameid", "condition", "objectSet", "instructionNum",
-                            "attemptNum", "critical")) #%>%
-#   left_join(errorDExp2, by = c("gameid", "condition", "objectSet", "attemptNum",
-#                                "instructionNum", "critical"))
+  left_join(mDExp2, by = c("gameid", "condition", "objectSet", "attemptNum",
+                               "instructionNum", "critical")) #%>%
 ```
@@
 ```{r}
+dExp1 <- joinedExp1 %>% 
+  filter(critical == 1) %>% 
+  group_by(objectSet, condition) %>%
+  filter(time > messageStamp) %>%
+  mutate(mouseY = 600 - mouseY) %>%
+  mutate(targetY = 600 - targetY) %>%
+  mutate(begTargetX = first(targetX),
+         begTargetY = first(targetY)) %>%
+  filter(targetX == begTargetX) %>%
+  filter(targetY == begTargetY) %>%
+  select(gameid, time, messageStamp, condition, 
+         objectSet, instructionNum, targetX, targetY, mouseX, mouseY) %>%
+  mutate(experiment = "Exp.1")
+
 dExp2 <- joinedExp2 %>% 
@@
   mutate(begTargetX = first(targetX),
-         begTargetY = first(targetY),
-         distractorX = as.numeric(levels(distractorX))[distractorX],
-         distractorY = 600 - as.numeric(levels(distractorY))[distractorY]) %>%
-   mutate(begDistrX = first(distractorX),
-          begDistrY = first(distractorY)) %>%
+         begTargetY = first(targetY)) %>%
    filter(targetX == begTargetX) %>%
    filter(targetY == begTargetY) %>%
-   filter(distractorX == begDistrX) %>%
-   filter(distractorY == begDistrY) 
+  select(gameid, time, messageStamp, condition, 
+         objectSet, instructionNum, targetX, targetY, mouseX, mouseY) %>%
+  mutate(experiment = "Exp.2")
 ```
@@
 
-heatMapForLMExp2 <- dExp2 %>%
-  group_by(gameid, objectSet) %>%
+intermediateHeatMap <- rbind(dExp1, dExp2) %>%
+  group_by(gameid, objectSet, condition, experiment) %>%
   mutate(inTargetSquare = as.numeric(mouseX > targetX - 75 
@@
                                      & mouseY < targetY + 75)) %>%
-  mutate(inDistractorSquare = as.numeric(mouseX > distractorX - 75
-                                         & mouseX < distractorX + 75
-                                         & mouseY > distractorY - 75
-                                         & mouseY < distractorY + 75)) %>%
-  group_by(gameid, condition) %>%
-  summarise(distractor = sum(inDistractorSquare) / n(),
-         target = sum(inTargetSquare) / n()) %>%
-  gather(cellType, percent, distractor, target)
+  filter(inTargetSquare == 1) %>%
+  summarise(messageTime = mean(messageStamp),
+            startTimeInTarget = first(time),
+            endTime = last(time))
 
-heatMapForLMExp2
+combinedHeatMap = intermediateHeatMap %>%
+    group_by(gameid, condition, experiment) %>%
+    summarise(targetHoverPercentage = mean((endTime - startTimeInTarget)/ 
+                                           (endTime - messageTime)))
+View(combinedHeatMap)
 
-t1 <- lm(percent ~ condition , data = heatMapForLMExp2 %>% filter(cellType == "target"))
-summary(t1)
+# Easier to report these as paired t-tests
+percentExp1 <- lmer(targetHoverPercentage ~ condition + (1 | gameid), 
+           data = combinedHeatMap %>% filter(experiment == "Exp.1"))
+summary(percentExp1)
+percentExp2 <- lmer(targetHoverPercentage ~ condition + (1 | gameid), 
+           data = heatMapForLMExp2 %>% filter(experiment == "Exp.2"))
+summary(percentExp2)
+percentExpVs <- lmer(targetHoverPercentage ~ condition * experiment + (1 | gameid),
+                     data = combinedHeatMap)
+summary(percentExpVs)
 ```
@@
 ```{r}
-heatMapForPlotExp2 <- dExp2 %>%
-  group_by(gameid, objectSet) %>%
-  mutate(inTargetSquare = as.numeric(mouseX > targetX - 75 
-                                     & mouseX < targetX + 75
-                                     & mouseY > targetY - 75
-                                    & mouseY < targetY + 75)) %>%
-  mutate(inDistractorSquare = as.numeric(mouseX > distractorX - 75
-                                         & mouseX < distractorX + 75
-                                         & mouseY > distractorY - 75
-                                         & mouseY < distractorY + 75)) %>%
-  group_by(gameid, condition) %>%
-  summarize(distractor = sum(inDistractorSquare) / n(),
-            target = sum(inTargetSquare) / n()) %>%
-  gather(cellType, percent, distractor, target) %>%
-  group_by(condition, cellType) %>%
-  summarize(error = sem(percent),
-            percent = mean(percent))
+stat = function(data, indices) {
+  d <- data[indices,]
+  output = d %>%
+    group_by(condition, experiment) %>%
+    summarise(targetHoverPercentage = mean((endTime - startTimeInTarget)/ 
+                                           (endTime - messageTime)))
+  return(output$targetHoverPercentage)
+}
 
-heatMapForPlotExp2
+heatMapBoot = boot(data = intermediateHeatMap, statistic = stat, R = 1000)
+Exp1baseCI = boot.ci(heatMapBoot, type = "basic", index=1)
+Exp2baseCI = boot.ci(heatMapBoot, type = "basic", index=2)
+Exp1expCI = boot.ci(heatMapBoot, type = "basic", index=3)
+Exp2expCI = boot.ci(heatMapBoot, type = "basic", index=4)
 
-g <- (ggplot(heatMapForPlotExp2, aes(x = condition, y = percent, 
-                                     group = cellType, color = cellType)) 
-      + geom_line(aes(linetype = cellType), size = 2)
-      + geom_errorbar(aes(ymax = percent + error, 
-                          ymin = percent - error), size =2 ,
-                      width = 0.1)
-     + ylim(0, .45)
-     + theme_bw(base_size = 16)
-      + ggtitle("Experiment 2 Hover-Time"))
-g
-ggsave("../../writing/cogsci-submission/images/exp2MouseTracking.pdf")
-```
+plottingDF = data.frame(condition = c("base", "base", "exp", "exp"),
+                        experiment = c("Exp.1", "Exp.2", "Exp.1", "Exp.2"),
+                        response = heatMapBoot$t0,
+                        lowerCI = c(Exp1baseCI$basic[1,4], Exp2baseCI$basic[1,4],
+                                    Exp1expCI$basic[1,4], Exp2expCI$basic[1,4]),
+                        upperCI = c(Exp1baseCI$basic[1,5], Exp2baseCI$basic[1,5],
+                                    Exp1expCI$basic[1,5], Exp2expCI$basic[1,5]))
 
-Compare mouse data from Exp1 w/ mouse data from Exp2:
+# heatMapForPlot = combinedHeatMap %>% 
+#   group_by(condition, experiment) %>%
+#   summarize(error = ci95(targetHoverPercentage),
+#             percent = mean(targetHoverPercentage))
 
-```{r}
-combinedHeatMap = rbind(heatMapForLMExp1 %>%
-                          mutate(exp = "Exp.1") %>% 
-                          filter(cellType == "target"),
-                        heatMapForLMExp2 %>% 
-                          mutate(exp = "Exp.2") %>%
-                          filter(cellType == "target"))
-combinedHeatMapForPlot <- combinedHeatMap %>%
-  group_by(condition, exp) %>%
-  summarize(error = ci95(percent),
-            percent = mean(percent))
-
-g <- (ggplot(combinedHeatMapForPlot, aes(x = exp, y = percent, 
+g <- (ggplot(plottingDF, aes(x = experiment, y = response, 
                                   group = condition, fill = condition)) 
       + geom_bar(position = dodge, stat= "identity")
-      + geom_errorbar(aes(ymax = percent + error, 
-                          ymin = percent - error), position = dodge,
+      + geom_errorbar(aes(ymax = upperCI, 
+                          ymin = lowerCI), position = dodge,
                           width = 0.25)
-      + ylim(0, .5)
+      + ylim(0, .4)
       + theme_bw(base_size = 16)
+      + ylab("% of total decision window")
       + ggtitle("Target Hover-Time"))
@@
 ggsave("../../writing/cogsci-submission/images/mousetracking.pdf")
-
-combinedRes = lm(percent ~ exp * condition, data = combinedHeatMap)
-summary(combinedRes)
 ```
-
-```{r}
-# heatMapForPlot$new_labels = as.factor(sapply(X = heatMapForPlot$objectSet, 
-#                               FUN = function(v) {return(paste("Item", v))}))
-# 
-# g <- (ggplot(heatMapForPlot, aes(x = mouseX, y = mouseY, color = condition)) +
-#       geom_vline(xintercept = c(0, 150,300,450, 600)) +
-#       geom_hline(yintercept = c(0, 150,300,450, 600)) +
-#       geom_point() +
-#       geom_point(aes(x = begTargetX, y = begTargetY), size = 20,
-#                  shape = 4, color = "black", show_guide = FALSE) +
-#       geom_point(aes(x = begDistrX, y = begDistrY), size = 20,
-#                  shape = 4, color = "grey50", show_guide = FALSE) + 
-#       theme(panel.grid.major = element_blank(), 
-#             panel.grid.minor = element_blank()) +
-#       xlim(0, 600) +
-#       ylim(0, 600) +
-#       theme(aspect.ratio = 1) +
-#       facet_wrap(~ new_labels, nrow = 2))
-# g
-```

### 2016-05-12 writing/cogsci-revision/cogsci_keysar.tex
Update revision to reflect original submission

--- 2016-01-30 writing/cogsci-submission/cogsci_keysar.tex
@@
 If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt.
-We took the set of all mouse position samples in the decision window (collected approximately every 15ms), and measured what proportion were located inside the cell containing the target.
+Within the decision window, we computed the total amount of time spent hovering over the cell containing the target and divided by the total length of the decision window to get a measure of the relative time spent considering the target.
 We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
-We found that people tended to spend less total time hovering over the target cell on critical experimental trials than on baseline trials, $b = -.11, t(20) = -1.26, p = 0.224$, although perhaps due to an underpowered sample and imprecise measurements\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.}, this result was not significant (see Fig.
-\ref{fig:hover}).
+A paired-samples $t$-test found that people tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig.
+\ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} .
 %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
@@
 \begin{center}
-\includegraphics[scale=.5]{images/mousetracking.pdf}
 \vspace{-.5cm}
-
-\caption{Mean percentage of time spent hovering over target cell in the experimental and baseline conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are 95\% confidence intervals.}
+\includegraphics[scale=.45]{images/mousetracking.pdf}
+\caption{Mean percentage of time spent hovering over target cell in the two conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are boostrapped 95\% confidence intervals.}
+\vspace{-.5cm}
 \label{fig:hover}
@@
 
-When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we again found a decrease in total hover time over the target in experimental vs.
-baseline trials $b = 0.07, t(54) = -2.1, p = 0.045$, indicating that the presence of a hidden distractor still interfered with participants directly choosing the target.
-While the effect appears to be slightly attenuated compared to Expt.~1 (see Figure \ref{fig:hover}), there is no significant interaction of experiment and condition on hover time $b = 0.04, t(74) = 0.45, p = 0.657$.
-Again, given the low sample size in Expt.~1, the noise from error, and the noise from counting samples instead of differences in timestamps, this result is limited.
+When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we found no significant decrease in target hover time between experimental and baseline trials $t(27) = 0.89, p = 0.38$.
+To directly test for differences in hover time patterns across the two experiments, we used a mixed-effects model with a random intercept for game ID and an interaction between condition and experiment on target hover time.
+We found a marginally significant interaction, $b = 0.09, t = 1.89, p = 0.066$ (see Figure \ref{fig:hover}), providing some evidence that the presence of a hidden distractor no longer interfered target selection in Expt~2.
+
 \begin{figure}[t!]
 \begin{center}
+\vspace{-.25cm}
 \includegraphics[scale=.5]{images/itemwisefitness.pdf}
-\vspace{-.25cm}
+\vspace{-.5cm}
 \caption{Item-wise error rates compared with label fitness ratios, for Expt.~1.
 Fitness ratios are scaled to the unit interval in order to visually compare to error rates.
-Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest density intervals for error rates.}
-\vspace{-.25cm}
+Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest posterior density intervals for error rates.}
+\vspace{-.5cm}
 \label{fig:itemLevel}
@@
 \begin{center}
+\vspace{-.25cm}
 \includegraphics[scale=.5]{images/fitnessInteraction.pdf}

### 2016-05-13 analysis/cogSci_analysis.Rmd
Update cogsci analysis

@@
 
-itemWiseTestExp1 <- chisq.test(itemInequalitiesExp1[-c(1,4,5)])
-
 itemInequalitiesExp2 <- errorDExp2 %>%
@@
   ggtitle("Speaker Informativity")
-ggsave("../../writing/cogsci-submission/images/fitnessInteraction.pdf")
+ggsave("../../writing/cogsci-revision/images/fitnessInteraction.pdf")
 
@@
 Compare item-wise error rates and fitness ratios
-
-```{r}
-library(binom)
-Exp1Errs = itemInequalitiesExp1 %>%
-  mutate(response = errorRate) %>%
-  mutate(measure = "error") %>%
-  do(mutate(., lower = binom.confint(.$errorCount, .$total, 
-                                     method = "bayes")$lower,
-               upper = binom.confint(.$errorCount, .$total, 
-                                     method = "bayes")$upper)) %>%
-  select(objectSet, response, measure, lower, upper)
-
-stat = function(data, indices) {
-  d <- data[indices,]
-  applicability = d %>%   
-    group_by(objectSet) %>%
-    summarize(m = .5 * mean(distractor)/mean(target))
-  return(applicability$m)
-}
-
-Exp1Fitness = overinformativeCodes %>%
-  filter(exp == "Exp.1") %>%
-  spread(referent, response)
-
-res = boot(data = Exp1Fitness, statistic = stat, R = 1000)
-Exp1PlottingDF = data.frame(response = rep(0, 8),
-                        upper = rep(0,8),
-                        lower = rep(0,8),
-                        objectSet = 1:8,
-                        measure = rep("fitness", 8))
-for (i in 1:8) {
-  ci = boot.ci(res, type = "basic", index=i)
-  Exp1PlottingDF[i,]$response = res$t0[i]
-  Exp1PlottingDF[i,]$lower = ci$basic[4]
-  Exp1PlottingDF[i,]$upper = ci$basic[5]
-}
-
-Exp1ItemwiseForPlot = rbind(Exp1PlottingDF, Exp1Errs)
-g <- (ggplot(Exp1ItemwiseForPlot, aes(x = objectSet, y = response, 
-                                      group = measure, fill = measure))
-  + geom_bar(stat = "identity", position = "dodge")
-  + geom_errorbar(aes(ymax = upper, ymin = lower), 
-                  position = dodge, width = 0.25)
-  + theme_bw(base_size = 16) 
-  + ggtitle("Exp.1 Item-Wise Error")
-  + ylab("ratio (scaled)"))
-g
-ggsave("../../writing/cogsci-submission/images/itemWiseFitness.pdf")
-```
-
-Alternatively, could make a scatter plot?
-
-```{r}
-library(binom)
-Exp1Errs = itemInequalitiesExp1 %>%
-  do(mutate(., errorLower = binom.confint(.$errorCount, .$total, 
-                                     method = "bayes")$lower,
-               errorUpper = binom.confint(.$errorCount, .$total, 
-                                     method = "bayes")$upper)) %>%
-  select(objectSet, errorRate, errorLower, errorUpper)
-
-stat = function(data, indices) {
-  d <- data[indices,]
-  applicability = d %>%   
-    group_by(objectSet) %>%
-    summarize(m = mean(distractor)/mean(target))
-  return(applicability$m)
-}
-
-Exp1Fitness = overinformativeCodes %>%
-  filter(exp == "Exp.1") %>%
-  spread(referent, response)
-
-res = boot(data = Exp1Fitness, statistic = stat, R = 1000)
-Exp1PlottingDF = data.frame(fitness = rep(0, 8),
-                        fitnessUpper = rep(0,8),
-                        fitnessLower = rep(0,8),
-                        objectSet = 1:8)
-for (i in 1:8) {
-  ci = boot.ci(res, type = "basic", index=i)
-  Exp1PlottingDF[i,]$fitness = res$t0[i]
-  Exp1PlottingDF[i,]$fitnessLower = ci$basic[4]
-  Exp1PlottingDF[i,]$fitnessUpper = ci$basic[5]
-}
-
-Exp1ItemwiseForPlot = left_join(Exp1Errs, Exp1PlottingDF, by = c("objectSet"))
-g <- (ggplot(Exp1ItemwiseForPlot, aes(x = fitness, y = errorRate))
-  + geom_point() 
-  + geom_errorbar(aes(ymax = errorUpper, ymin = errorLower), 
-                  width = 0.05)
-  + geom_smooth(method = "lm")
-  + theme_bw(base_size = 16) 
-  + ggtitle("Exp.1 Item-Wise Error")
-  + ylab("ratio (scaled)"))
-g
-#ggsave("../../writing/cogsci-submission/images/itemWiseFitness.pdf")
-```
-Same thing for Exp 2
 
@@
 
-res = boot(data = Exp2Fitness, statistic = stat, R = 1000)
+res = boot(data = Exp2Fitness, statistic = stat, R = 10000)
 
@@
   + ylab("Error rate")
-  + xlab("Label fitness of distractor vs. target"))
+  + xlab("Distractor fitness / target fitness"))
 g
@@
 ```{r}
-summary(lm(errorRate ~ fitness * experiment, data = Exp2ItemwiseForPlot))
-
-# within exp. correlations high
-cor(Exp1Errs$response, Exp1PlottingDF$response)
-cor(Exp2Errs$response, Exp2PlottingDF$response)
-
-# Cross exp. correlations low
-cor(Exp2Errs$response, Exp1PlottingDF$response)
-cor(Exp1Errs$response, Exp2PlottingDF$response)
+summary(lm(errorRate ~ fitness , data = Exp2ItemwiseForPlot))
 ```
@@
                                      & mouseY < targetY + 75)) %>%
-  mutate(inDistractorSquare = as.numeric(mouseX > distractorX - 75
-                                         & mouseX < distractorX + 75
-                                         & mouseY > distractorY - 75
-                                         & mouseY < distractorY + 75)) %>%
-  group_by(gameid, condition) %>%
-  summarise(distractor = sum(inDistractorSquare) / n(),
-         target = sum(inTargetSquare) / n()) %>%
-  gather(cellType, percent, distractor, target)
+  filter(inTargetSquare == 1) %>%
+  summarise(messageTime = mean(messageStamp),
+            startTimeInTarget = first(time),
+            endTime = last(time))
 
-heatMapForLMExp2
+combinedHeatMap = intermediateHeatMap %>%
+    group_by(gameid, condition, experiment) %>%
+    summarise(targetHoverPercentage = mean((endTime - startTimeInTarget)/ 
+                                           (endTime - messageTime)))
 
-t1 <- lm(percent ~ condition , data = heatMapForLMExp2 %>% filter(cellType == "target"))
-summary(t1)
+# Easier to report these as paired t-tests
+percentExp1 <- lmer(targetHoverPercentage ~ condition + (1 | gameid), 
+           data = combinedHeatMap %>% filter(experiment == "Exp.1"))
+summary(percentExp1)
+percentExp2 <- lmer(targetHoverPercentage ~ condition + (1 | gameid), 
+           data = heatMapForLMExp2 %>% filter(experiment == "Exp.2"))
+summary(percentExp2)
+percentExpVs <- lmer(targetHoverPercentage ~ condition * experiment + (1 | gameid),
+                     data = combinedHeatMap)
+summary(percentExpVs)
 ```
@@
 ```{r}
-heatMapForPlotExp2 <- dExp2 %>%
-  group_by(gameid, objectSet) %>%
-  mutate(inTargetSquare = as.numeric(mouseX > targetX - 75 
-                                     & mouseX < targetX + 75
-                                     & mouseY > targetY - 75
-                                    & mouseY < targetY + 75)) %>%
-  mutate(inDistractorSquare = as.numeric(mouseX > distractorX - 75
-                                         & mouseX < distractorX + 75
-                                         & mouseY > distractorY - 75
-                                         & mouseY < distractorY + 75)) %>%
-  group_by(gameid, condition) %>%
-  summarize(distractor = sum(inDistractorSquare) / n(),
-            target = sum(inTargetSquare) / n()) %>%
-  gather(cellType, percent, distractor, target) %>%
-  group_by(condition, cellType) %>%
-  summarize(error = sem(percent),
-            percent = mean(percent))
+stat = function(data, indices) {
+  d <- data[indices,]
+  output = d %>%
+    group_by(condition, experiment) %>%
+    summarise(targetHoverPercentage = mean((endTime - startTimeInTarget)/ 
+                                           (endTime - messageTime)))
+  return(output$targetHoverPercentage)
+}
 
-heatMapForPlotExp2
+heatMapBoot = boot(data = intermediateHeatMap, statistic = stat, R = 1000)
+Exp1baseCI = boot.ci(heatMapBoot, type = "basic", index=1)
+Exp2baseCI = boot.ci(heatMapBoot, type = "basic", index=2)
+Exp1expCI = boot.ci(heatMapBoot, type = "basic", index=3)
+Exp2expCI = boot.ci(heatMapBoot, type = "basic", index=4)
 
-g <- (ggplot(heatMapForPlotExp2, aes(x = condition, y = percent, 
-                                     group = cellType, color = cellType)) 
-      + geom_line(aes(linetype = cellType), size = 2)
-      + geom_errorbar(aes(ymax = percent + error, 
-                          ymin = percent - error), size =2 ,
-                      width = 0.1)
-     + ylim(0, .45)
-     + theme_bw(base_size = 16)
-      + ggtitle("Experiment 2 Hover-Time"))
-g
-ggsave("../../writing/cogsci-submission/images/exp2MouseTracking.pdf")
-```
+plottingDF = data.frame(condition = c("base", "base", "exp", "exp"),
+                        experiment = c("Exp.1", "Exp.2", "Exp.1", "Exp.2"),
+                        response = heatMapBoot$t0,
+                        lowerCI = c(Exp1baseCI$basic[1,4], Exp2baseCI$basic[1,4],
+                                    Exp1expCI$basic[1,4], Exp2expCI$basic[1,4]),
+                        upperCI = c(Exp1baseCI$basic[1,5], Exp2baseCI$basic[1,5],
+                                    Exp1expCI$basic[1,5], Exp2expCI$basic[1,5]))
 
-Compare mouse data from Exp1 w/ mouse data from Exp2:
+# heatMapForPlot = combinedHeatMap %>% 
+#   group_by(condition, experiment) %>%
+#   summarize(error = ci95(targetHoverPercentage),
+#             percent = mean(targetHoverPercentage))
 
-```{r}
-combinedHeatMap = rbind(heatMapForLMExp1 %>%
-                          mutate(exp = "Exp.1") %>% 
-                          filter(cellType == "target"),
-                        heatMapForLMExp2 %>% 
-                          mutate(exp = "Exp.2") %>%
-                          filter(cellType == "target"))
-combinedHeatMapForPlot <- combinedHeatMap %>%
-  group_by(condition, exp) %>%
-  summarize(error = ci95(percent),
-            percent = mean(percent))
-
-g <- (ggplot(combinedHeatMapForPlot, aes(x = exp, y = percent, 
+g <- (ggplot(plottingDF, aes(x = experiment, y = response, 
                                   group = condition, fill = condition)) 
       + geom_bar(position = dodge, stat= "identity")
-      + geom_errorbar(aes(ymax = percent + error, 
-                          ymin = percent - error), position = dodge,
+      + geom_errorbar(aes(ymax = upperCI, 
+                          ymin = lowerCI), position = dodge,
                           width = 0.25)
-      + ylim(0, .5)
+      + ylim(0, .4)
       + theme_bw(base_size = 16)
+      + ylab("% of total decision window")
       + ggtitle("Target Hover-Time"))
 g
-ggsave("../../writing/cogsci-submission/images/mousetracking.pdf")
-
-combinedRes = lm(percent ~ exp * condition, data = combinedHeatMap)
-summary(combinedRes)
+ggsave("../../writing/cogsci-revision/images/mousetracking.pdf")
 ```

### 2016-05-13 writing/cogsci-revision/cogsci_keysar.tex
Finish revisions

@@
 \end{center}
-\vspace{-.25cm}
+\vspace{-.5cm}
 \caption{Interface used in the reported experiments. Objects behind the black squares were hidden from the director.}
@@
 Among other things, this ability -- called \emph{theory of mind} \cite{PremackWoodruff78_ChimpanzeeToM} -- allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}.
-Children acquire this ability by at least age six \cite{WimmerPerner83_BeliefsAboutBeliefs, WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
+Children acquire this ability by at least age six \cite{WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
 %
@@
 \vspace{-.5cm}
-\includegraphics[scale=.45]{images/mousetracking.pdf}
+\includegraphics[scale=.5]{images/mousetracking.pdf}
 \caption{Mean percentage of time spent hovering over target cell in the two conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are boostrapped 95\% confidence intervals.}
@@
 \vspace{-.25cm}
-\includegraphics[scale=.5]{images/itemwisefitness.pdf}
+\includegraphics[scale=.5]{images/fitnessInteraction.pdf}
 \vspace{-.5cm}
-\caption{Item-wise error rates compared with label fitness ratios, for Expt.~1.
-Fitness ratios are scaled to the unit interval in order to visually compare to error rates.
-Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest posterior density intervals for error rates.}
+\caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
 \vspace{-.5cm}
-\label{fig:itemLevel}
+\label{fig:fitnessInteraction}
 \end{center}
@@
 In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
-We find a correlation of $r = 0.76$ between these two measures for Expt.~1, and a correlation of $r = 0.53$ for Expt.~2.
-Comparing error rates from one experiment with label fitness ratios from the other, however, yields much weaker correlations ($r = 0.07$), suggesting that item-wise variability across the two experiments is driven by relative informativity of speaker utterances.
+We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$.
+This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances.
 
@@
 Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
-
 \begin{figure}[t!]
@@
 \vspace{-.25cm}
-\includegraphics[scale=.5]{images/fitnessInteraction.pdf}
-\vspace{-.25cm}
-\caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
-\vspace{-.25cm}
-\label{fig:fitnessInteraction}
+\includegraphics[scale=.55]{images/itemwisefitness.pdf}
+\vspace{-.75cm}
+\caption{Item-wise error rates compared with label fitness ratios, across both experiments.
+Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest posterior density intervals for error rates.}
+\vspace{-.5cm}
+\label{fig:itemLevel}
 \end{center}
@@
 Instead, they said ``the bluish block with a B'' or ``block with the blue writing,'' which relied on less confusable perceptual features and thereby decreased error rates.
+While the stimuli
 Thus, the errors observed in Expt.~1 can be explained as a result of an uncooperative confederate (speaker)---the experiment set up certain pragmatic expectations through the task context, then deliberately flouted them in critical trials.
-% Removing constraints on the speaker produced a dramatic reduction in overall matcher errors and systematic variation in error rates across items.
-We found that unscripted speakers systematically tended to produce more precise, costly referring expressions than their scripted counterparts -- for example, ``the clear audiocassette'' when there was only one ``tape'' in the speaker's display.
-These precise labels were less likely to fit the hidden distractor, and therefore less likely to lead to a matcher error.
 
@@
 %In the current task, it seems particularly relevant that both participants knew about the presence of hidden objects, increasing uncertainty about possible distractors and placing additional pressure on the speaker to overspecify.
-The exact origin is a subject that must be followed up by future research, but it is clear that the tendency of speakers to produce highly informative referring expressions is useful to, and relied on by, listeners.
-Indeed, the remaining errors in Expt.~2 are at least partially due to the lack of availability of more specific expressions. 
-The items on which errors were most likely in Expt.~2 were difficult precisely because the natural utterance was \emph{still} ambiguous: even if the speaker uses ``hairbrush'' instead of the basic-level ``brush,'' for example, the distractor remains a more typical exemplar of the subcategory.
+The exact origin is a subject that must be followed up by future research \cite<e.g.>{GannBarr14_AudienceDesign}, but it is clear that the tendency of speakers to produce highly informative referring expressions is useful to, and relied on by, listeners.
+%Indeed, the remaining errors in Expt.~2 are at least partially due to the lack of availability of more specific expressions. 
+%The items on which errors were most likely in Expt.~2 were difficult precisely because the natural utterance was \emph{still} ambiguous: even if the speaker uses ``hairbrush'' instead of the basic-level ``brush,'' for example, the distractor remains a more typical exemplar of the subcategory.
+
+This finding is consistent with a recent proposal by \citeA{HellerParisienStevenson16_ProbabilisticWeighing} that referring expressions are interpreted by probabilistically integrating multiple sources of information: when conversational expectations lead participants to expect over-informative utterances, they (rationally) place relatively less weight on features of the environment determining common ground, such as shared perceptual access.
+Further work in a wider variety of tasks is necessary to pin down the various factors determining the relative weighting of these different sources of information, but we have argued that pragmatics play a crucial role.
 
@@
 We believe we addressed the major concern about exploring theory of mind in web experiments -- that participants do not truly believe they are interacting with another human -- by allowing instantaneous, responsive, real-time interaction.
-On the other hand, it is known that textual communication, as in our chat box, can differ from face-to-face verbal communication, and there was no way of addressing this.
+On the other hand, it is known that textual communication, as in our chat box, can differ from face-to-face verbal communication.
+Additionally, aspects of the interface such as the graphical representation of occluded cells may be less intuitive, or require more training, on the web than in the lab.
 % Directors watched objects move through the actual moment-by-moment mouse trajectories that matchers used to drag them, and matchers saw the actual timing of the messages sent by the director.
@@
 A related difference is our decision not to use a confederate.
-While confederates are useful for reducing variation across instances of the experiment, we see several downsides.
+While confederates are useful for reducing variation across instances of the experiment and delivering carefully targeted manipulations, their use may have unexpected consequences.
 Beyond the difficulties of conducting large-scale experiments with confederates, it is difficult to exactly replicate all the subtleties of the confederate's behavior that might influence results.

### 2016-07-30 analysis/cogSci_analysis.Rmd
Fix itemwise plot; make new fixed effects plots for logistic regression

@@
 ```{r}
-setwd('~/Repos/keysar_replication/experiment1/analysis/')
-preMouseDExp1 = read.csv('fullSampleMouse.csv')
-preErrorDExp1 = read.csv('fullSampleError.csv')
-preMessageDExp1 = read.csv('fullSampleMessage.csv')
+setwd('~/Repos/keysar_replication/experiment1/data/')
+preMouseDExp1 = read.csv('mouse/fullSampleMouse.csv')
+preErrorDExp1 = read.csv('error/fullSampleError.csv')
+preMessageDExp1 = read.csv('message/fullSampleMessage.csv')
 
@@
 library(lmerTest)
+library(binom)
 ```
@@
 
+errorTableExp1 <- eDExp1 %>%
+  group_by(condition) %>%
+  summarise(atLeastOnce = sum(cond1)/numParticipantsExp1, 
+            atLeastTwice = sum(cond2)/numParticipantsExp1,
+            total = sum(n) / (4 * numParticipantsExp1))
 errorTableExp2 <- eDExp2 %>%
@@
             total = sum(n) / (4 * numParticipantsExp2))
-errorTableExp2
 
@@
             total = sum(n) / (4 * numParticipantsExp2))
-excludedEDExp2
+
+errorTableExp1
+errorTableExp2
+
 ```
@@
   select(objectSet, errorCount, correctCount, errorRate, total)
+
+# Replace NAs with true values 
 itemInequalitiesExp1[1,2:4] = c(0, itemInequalitiesExp1[1,]$total, 0)
+
+itemWiseTestExp1 <- chisq.test(itemInequalitiesExp1 %>% select(errorCount,correctCount))
+print(itemWiseTestExp1)
 
@@
 
+# Replace NAs with true values 
 itemInequalitiesExp2[1,2:4] = c(0, itemInequalitiesExp2[1,]$total, 0)
 
-itemWiseTestExp2 <- chisq.test(itemInequalitiesExp2[-c(1,4,5)])
+itemWiseTestExp2 <- chisq.test(itemInequalitiesExp2 %>% select(errorCount,correctCount))
 print(itemWiseTestExp2)
@@
 ```{r}
+
 library(binom)
+
+Exp1Errs = itemInequalitiesExp1 %>%
+  do(mutate(., errorLower = binom.confint(.$errorCount, .$total, 
+                                     method = "bayes")$lower,
+               errorUpper = binom.confint(.$errorCount, .$total, 
+                                     method = "bayes")$upper)) %>%
+  select(objectSet, errorRate, errorLower, errorUpper)
+
+stat = function(data, indices) {
+  d <- data[indices,]
+  applicability = d %>%   
+    group_by(objectSet) %>%
+    summarize(m = mean(distractor)/mean(target))
+  return(applicability$m)
+}
+
+Exp1Fitness = overinformativeCodes %>%
+  filter(exp == "Exp.1") %>%
+  spread(referent, response)
+
+res = boot(data = Exp1Fitness, statistic = stat, R = 10000)
+Exp1PlottingDF = data.frame(fitness = rep(0, 8),
+                        fitnessUpper = rep(0,8),
+                        fitnessLower = rep(0,8),
+                        objectSet = 1:8)
+for (i in 1:8) {
+  ci = boot.ci(res, type = "basic", index=i)
+  Exp1PlottingDF[i,]$fitness = res$t0[i]
+  Exp1PlottingDF[i,]$fitnessLower = ci$basic[4]
+  Exp1PlottingDF[i,]$fitnessUpper = ci$basic[5]
+}
+
+Exp1ItemwiseForPlot = left_join(Exp1Errs, Exp1PlottingDF, by = c("objectSet"))
+
 binomStat = stat = function(data, indices) {
@@
 g <- (ggplot(Exp2ItemwiseForPlot, aes(x = fitness, y = errorRate))
-  + geom_point(aes(color = experiment))
+  + geom_point(aes(color = experiment), size = 2)
   + geom_errorbarh(aes(xmax = fitnessUpper, xmin = fitnessLower,
                        color = experiment), 
-                  height = 0.01)
+                  height = 0.01, alpha = .5)
   + geom_errorbar(aes(ymax = errorUpper, ymin = errorLower,
                      color = experiment), 
-                width = 0.05)
+                width = 0.05, alpha = .5)
   + geom_smooth(method = "lm",colour="black")
@@
   summarize(meanFitness = mean(response)) %>%
-  spread(referent, meanFitness)
+  spread(referent, meanFitness) %>%
+  ungroup() %>%
+  mutate(objectSetF = factor(objectSet)) %>%
+  mutate(criticalErrorF = factor(criticalError)) %>%
+  mutate(scaledTarget)
+```
 
+```{r}
 library(lme4)
-mmRes1 <- glmer(criticalError ~ factor(objectSet) + (1 | gameid), data = itemWiseCodes, family = binomial)
-summary(mmRes1)
-mmRes2 <- glmer(criticalError ~ scale(target) + scale(distractor) + factor(objectSet) + (1 | gameid),
+
+# Note: included objectSetF as *fixed-effect* in cogsci paper, e.g.
+# glmer(criticalError ~ objectSetF + (1 | gameid), data = itemWiseCodes, family = binomial)
+# This is cleaner and helps the model converge better, but doesn't qualitatively change results
+
+mmRes1 <- glmer(criticalError ~ (1 | gameid) + (1 | objectSetF), data = itemWiseCodes, family = binomial)
+mmRes2 <- glmer(criticalError ~ scale(target) + scale(distractor)
+                                + (1 | gameid) + (1 | objectSetF),
                 data = itemWiseCodes, family = binomial)
+anova(mmRes1, mmRes2)
+summary(mmRes2)
 
-anova(mmRes1, mmRes2)
+pl <- sjp.glmer(mmRes2, type = "ri.slope",facet.grid = FALSE, prnt.plot = F)
 
-# newItemWiseCodes = itemWiseCodes %>% 
-#   mutate(modelPreds = exp(predict(mmRes1))/(exp(predict(mmRes1))+1)) %>%
-#   group_by(exp, objectSet) %>%
-#   summarize(empError = mean(criticalError),
-#             modelError = mean(modelPreds))
-# 
-# (ggplot(newItemWiseCodes, aes(x = empError, y = modelError, color = exp)) 
-#    + geom_point()
-#    + geom_smooth(method = lm))
+targetPlot <- pl$plot[[3]]
+distractorPlot <- pl$plot[[4]]
+
+targetPlot+ 
+  xlab("label fitness for target (scaled)") +
+  ylab("predicted probability of error") +
+  ggtitle("target informativity") + 
+  labs(color="item")
+
+distractorPlot +
+  xlab("label fitness for distractor (scaled)") +
+  ylab("predicted probability of error") +
+  ggtitle("distractor informativity") + 
+  labs(color="item")
 
@@
 
-gg <- mmRes2@optinfo$derivs$grad
-hh <- mmRes2@optinfo$derivs$Hessian
-vv <- sqrt(diag(solve(hh/2)))
-print(summary(abs(gg*vv)))
-
-relgrad <- with(mmRes2@optinfo$derivs,solve(Hessian,gradient))
-print(max(abs(relgrad)))
+# gg <- mmRes2@optinfo$derivs$grad
+# hh <- mmRes2@optinfo$derivs$Hessian
+# vv <- sqrt(diag(solve(hh/2)))
+# print(summary(abs(gg*vv)))
+# 
+# relgrad <- with(mmRes2@optinfo$derivs,solve(Hessian,gradient))
+# print(max(abs(relgrad)))
 

### 2016-12-14 writing/psych285-final-paper/psych285_final.tex
Add final paper on modeling...

\documentclass[12pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{apacite}

%SetFonts

%SetFonts

\title{PSYCH 285: Final Paper}
\author{Robert Hawkins}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle

\section*{Introduction}

%% Broad point synthesizing a bunch of stuff we read about the false-belief task
If there is a single unifying message to take away from the vast literature on theory of mind, it is that the mental representations we use to reason about other minds are broader, more flexible, and more multi-faceted than any single experimental task can hope to pull apart. Traditional false-belief tasks  \cite<e.g.>{WimmerPerner83_BeliefsAboutBeliefs}, for instance, have been immensely useful for probing the ``layer'' of our representation that allows another's beliefs to be held alongside conflicting information about the true state of affairs. They also serve as useful litmus tests for drawing comparisons between human and non-human primates \cite{Krupenye16_GreatApesToM} or between younger and older children \cite{WellmanCrossWatson01_ToMMetaAnalysis}.

On the other hand, they are not particularly informative about other aspects of our representations, such as the conventions or norms we expect others to conform to \cite{Lewis69_Convention, RakoczyEtAl08_SourcesOfNormativity} or the tradeoffs between costs and rewards we expect others to rationally navigate \cite{JaraEttingerEtAl16_NaiveUtilityCalculus}. We have also become more aware of other aspects that are \emph{are} involved in task performance but were not originally intended to be. For instance, answering the critical question \emph{``Where will Sally look for her ball?''} may depend on reasoning about the experimenter's intentions when they draw attention to object locations or take the communicative action of asking a pedagogical question \cite{Strawson64_IntentionConvention, TomaselloCarpenter___Moll05_IntentionsCulturalCognition, WestraCarruthers17_PragmaticsToMScale}.

\section*{The Director-Matcher Task}
%% Introduce Keysar task by analogy to the false-belief task
Similar criticisms have recently been directed at the \emph{director-matcher task} used to argue for failures of theory of mind use \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}. While its history is shorter and less widely appreciated than the tradition of traditional false-belief tasks, it is the crux of one of the most influential arguments about adult theory of mind use to emerge over the past two decades: While adults are \emph{capable} of using deploying theory of mind, they do not always use it reliably. In other words, adults are ``reflexively mind-blind'' and only engage in perspective taking through an effortful process.

%% explain Keysar task
To demonstrate this, Keysar and colleagues placed several objects in a grid between two players. One player, a confederate, gives instructions to the other player about which objects to move. Critically, some cells of the grid are occluded so that only the matcher can see the object. They found that when an instruction was ambiguous between a hidden object and an object in common ground, nearly three out of four matches reached for the hidden one at least once \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Since a participant accurately taking the director's perspective would realize that the director couldn't possibly be referring to an object they don't know about, this was taken as a \emph{failure} of theory of mind. The same reference-game paradigm was used as a basis for a number of follow-up studies testing perspective-taking in children \cite{EpleyMorewedgeKeysar04_PerspectiveTakingChildren},  among friends vs. strangers \cite{SavitskyEtAl11_ClosenessCommunciation}, across Eastern and Western cultures \cite{WuKeysar07_Culture}, and in contexts of varying background information overlap \cite{WuKeysar07_InformationOverlap}.

%% Summarize specific criticisms of Keysar task
While the basic effect seems robust, there has been substantial disagreement over whether it does, in fact, reflect a \emph{failure}. Some of these counter-arguments point to methodological details.  \citeA{HannaTanenhausTrueswell03_CommonGroundPerspective} argued that common ground is usually established incrementally -- utterance by utterance -- over a conversation rather than through an artificially-imposed environmental constraint and also that the original task lacked a necessary control condition where both the target and distractor object are in common ground. Another methodological counter-argument from \citeA{HellerGrodnerTanenhaus08_Perspective} focused on the small set of items used and the wide variation in the relative fitness of the scripted word to the distractor versus the target (e.g. ``tape'' is more commonly used to refer to the hidden ``roll of tape'' than the shared ``cassette tape''). Finally, \citeA{RubioFernandez16_DirectorTaskAttention} point out that participants who initially adopt a selective-attention strategy where they ignore occluded cells can perform perfectly on all measures while \emph{not} using their theory of mind to consider the other's perspective at all. Together, these arguments make the case that the task is simply an unreliable test of theory of mind use, but do not make positive claims about how, when, or why theory of mind \emph{is} used.

%% present my results and pivoting to focus on pragmatics
A separate line of criticism considers the \emph{pragmatics} of the task in a way that reinterprets apparent failures on the task as successful uses of theory of mind \cite{RubioFernandez16_DirectorTaskAttention, HawkinsGoodman16_Keysar}. One key question hinges on the use of confederates in language-based tasks \cite{LockridgeBrennan02_SyntacticChoices, KuhlenBrennan13_LanguageInDialogue}. A confederate's behavior often differs dramatically from a naive subject's behavior in the same context. The relevant properties of behavior in dialogue -- latencies, prosody, eye gaze, lexical and syntactic alignment, choice of referring expression -- are still largely unknown to experimenters and difficult to explicitly control.  To the extent that their partner expects some properties of natural dialogue and hears something else, they may draw unexpected implicatures. Indeed, one conceptual replication of the task that used naive subjects in both roles found many more \emph{overinformative} referring expressions than the confederate's script contained \cite{HawkinsGoodman16_Keysar}. If the listener successfully used theory of mind reasoning to form expectations about how the speaker would behave in such as context, then the scripted utterances used by the confederate would naturally lead to errors. Using similar reasoning, \citeA{RubioFernandez16_DirectorTaskAttention} showed that listeners became suspicious of what was actually in common ground after hearing overinformative utterances, thus successfully using theory of mind reasoning.

\section*{A model of pragmatics in the director-matcher task}

The most successful computational models synthesize the thicket of contradictory empirical evidence and informal theory-building into a set of precisely-stated hypotheses that can be compared on the basis of quantitative predictions. The current state-of-the-art in modeling theory of mind reasoning has not addressed the underlying pragmatics of the task \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, KileyHamlinEtAl13_MentalisticCoreSocial}, and the current state-of-the-art in modeling pragmatics in language \cite{FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature, GoodmanFrank16_RSATiCS} has not been applied to standard theory of mind tasks despite explicitly framing language understanding as social reasoning. In this section, I formalize the director-matcher task inside a rational speech act (RSA) model, where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language. This allows us to test pragmatic accounts of task performance.

Previous theoretical advances in modeling pragmatics have typically hinged on `lifting' inference over some aspect of communication that was previously assumed to be fixed. For instance, \emph{manner implicatures} emerge from uncertainty over the true meanings of words in the lexicon \cite{BergenLevyGoodman16_LexicalUncertainty}, nonliteral constructions like \emph{hyperbole} and \emph{irony} emerge from uncertainty over the topic of discussion \cite{KaoWuBergenGoodman14_NonliteralNumberWords}, and \emph{prosody} and \emph{ellipsis} emerge from uncertainty (in the form of a noisy-channel model) over the various ways that a speaker's utterance could have been corrupted  \cite{BergenGoodman15_StrategicUseOfNoise}. Note that these are not ad hoc extensions; sources of uncertainty that are not strictly necessary to explain a particular effect are often omitted for simplicity, but they all coexist within the same model.

Pragmatic critiques of the director-matcher task point to an additional source of uncertainty that has not previously been explored: uncertainty over a communication partner's visual context. Considering this uncertainty, of course, \emph{is} theory of mind in action; our goal is to show how such a model built on social reasoning can produce the `failures' found by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} as well as the phenomena observed by both \citeA{HawkinsGoodman16_Keysar} and \citeA{RubioFernandez16_DirectorTaskAttention}. For simplicity, we consider the full $2\times 2$ grid containing a single occluded cell, which was used by \citeA{RubioFernandez16_DirectorTaskAttention}. We incorporate this source of uncertainty into both the speaker and listener models as follows:

\begin{enumerate}
\item A pragmatic speaker perceives an array of three objects $\mathcal{O}_S = \{o_i\}$, one of which is identified as a target $t$, but assumes that their partner sees an additional hidden object $o_h$ that is not in common ground. The speaker has uncertainty over the identity of this object, represented by a prior distribution $P(o_h)$ which they marginalize over when deciding what utterance to produce:
$$S_1(u | t, \mathcal{O}_p) \propto \sum_{o_h} L_0(t| \mathcal{O}_S \cup \{o_h\})P(o_h)P(u)$$

\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$. The listener, however, also has some uncertainty over whether the experimenter is deceptive, represented by a Bernoulli random variable $d$, such that they assign a small prior probability $P(d)$ to the possibility that $o_h$ is actually in common ground. Upon hearing an utterance $u$, they update both their beliefs about which object is the target and whether the experimenter is being deceptive:
$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
\end{enumerate}
Technical details are omitted here for brevity, but the full model is fully specified in WebPPL and available to run at \url{http://forestdb.org/models/keysar.html}.

We focus on two critical patterns of results produced by the pragmatic listener $L_2$. First, we consider how it responds to the scripted instruction ``fish'' when presented with a context containing two fish (a red one that is occluded and a blue one in common ground) as well as two other unrelated objects. While it is most likely to choose the red fish in common ground, it has a relatively high probability (43\%) of choosing the hidden red fish, an effect \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} claimed was due to failure of theory of mind. This decision emerges in the model for two primary reasons. For one, a speaker with uncertainty over context could in principle be communicating about any object, and if they believed the hidden object was a red fish, the utterance `fish' would have a high probability of being produced. To a lesser extent, context uncertainty makes the overinformative utterance `blue fish' more likely than `fish', hence the omission of the color modifier has the (subtle) implication of \emph{not} the blue fish. This speaker phenomenon, while subtle under the parameter regime of the current implementation, produces the dynamic observed in the unscripted replication by  \citeA{HawkinsGoodman16_Keysar}.

Second, we consider the listener's posterior over the \emph{deception} Bernoulli variable $d$. Just as \citeA{RubioFernandez16_DirectorTaskAttention} found that listeners were more likely to suspect deception after hearing an `overinformative' utterance like `blue fish' when only one fish was in common ground, so too does our pragmatic listener. The pragmatic speaker is relatively more likely to produce the additional color modifier if they see multiple fish than when they see a single fish and simply have uncertainty over the unknown object. Hence, our model simultaneously captures all three phenomena.

\section*{Discussion \& Conclusion}

Theory of mind use is intimately intertwined with language understanding; pragmatic reasoning is fundamentally social reasoning applied to communicative behavior. While traditional theory of mind tasks have always relied on language -- a critique that was first raised by developmental researchers \cite{BaillargeonScottHe10_FalseBeliefInfants} -- there is increasing recognition of the pragmatics inherent to these tasks \cite{WestraCarruthers17_PragmaticsToMScale}. In this paper, we turned this critical lens on the popular director-matcher task and synthesized several recent results under a single model of social reasoning under uncertainty about shared context.

Just as no single theory of mind task is sufficient to pull apart the many aspects of our mental representations, no single level of analysis is sufficient to interpret the empirical evidence \cite{Marr10_Vision}. We need mechanistic accounts that isolate distinct neural subsystems \cite{GallagherFrith03_fMRI_ToM, KosterHaleSaxe13_fMRI_ToM}; algorithmic accounts that focus on memory, attention, executive control, and other processing demands\cite{RubioFernandezGeurts12_FBFourthBirthday, ApperlyButterfill09_TwoSystems}; and computational accounts that ask functional question: what problem is theory of mind useful for solving and what information must be included in our representation in order to solve it  \citeA{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, Tomasello09_WhyWeCooperate, TomaselloVaish13_OriginsCooperation}. Moving ahead, each of these lenses will be critical to understanding exactly what aspects of our interlocutor's mind we represent in pragmatic reasoning, why other primates seem unable to engage in such reasoning, and how it develops across the lifespan.

\bibliography{psych285_final}
\bibliographystyle{apacite}

\end{document}

### 2017-07-15 analysis/cogSci_analysis.Rmd
Add a bunch of hard-coded value plots for talk

@@
 library(irr)
-turkerRatings <- read.csv("../../normingExperiment/data/keysarNorming-trials.csv")
+turkerRatings <- read.csv("../normingExperiment/data/keysarNorming-trials.csv")
 relevantProperties <- c("objectSet", "label")
 
-criticalMessagesExp1 <- read.csv("../analysis/messagesExp1.csv") %>%
+criticalMessagesExp1 <- read.csv("../experiment1/data/message/analysis/messagesExp1.csv") %>%
   left_join(turkerRatings, by = relevantProperties) %>%
@@
 ```
+
+# Vanity plots for talk
+
+```{r}
+library(ggplot2)
+library(ggthemes)
+d.model <- data.frame(x = rep(c('fish', 'blue fish'), 2),
+                      y = c(0.37,0.63,0.73, 0.27),
+                      source = c(rep('with context-uncertainty', 2),
+                                 rep('without context-uncertainty', 2)))
+ggplot(d.model, aes(x=x, y=y, fill = x)) +
+  geom_bar(stat='identity', position = 'dodge') +
+  theme_few() +
+  scale_fill_manual(values = c('red', 'grey')) +
+  facet_wrap(~ source) +
+  ylab("P(u | o)")
+```
+
+```{r}
+d.comp <- data.frame(x = rep(c('experimental', 'baseline'), 2),
+                      y = c(.30, 0, .43, .02),
+                      source = c(rep('original', 2),
+                                 rep('replication', 2)),
+                     error = c(1/sqrt(38*4),1/sqrt(38*4), 
+                               1/sqrt(14 * 4), 1/sqrt(14 * 4)))
+ggplot(d.comp, aes(x=x, y=y, fill = x)) +
+  geom_bar(stat='identity', position = 'dodge') +
+  theme_few() +
+  scale_fill_manual(values = c('red', 'grey')) +
+  geom_errorbar(aes(ymin = y - error, ymax = y + error), width = 0) +
+  facet_wrap(~ source) +
+  ylim(0, 1) +
+  ylab("% total errors")
+```
+
+```{r}
+d.comp <- data.frame(x = rep(c('experimental', 'baseline'), 3),
+                      y = c(.30, 0, .43, .02, .24, 0),
+                      source = ordered(c(rep('original', 2),
+                                 rep('Exp. 1', 2),
+                                 rep('Exp. 2', 2)),
+                                 levels = c('original', 'Exp. 1', 'Exp. 2')),
+                     error = c(1/sqrt(38*4),1/sqrt(38*4), 
+                               1/sqrt(14 * 4), 1/sqrt(14 * 4),
+                               1/sqrt(32*4), 1/sqrt(32*4)))
+ggplot(d.comp, aes(x=x, y=y, fill = x)) +
+  geom_bar(stat='identity', position = 'dodge') +
+  theme_few() +
+  scale_fill_manual(values = c('red', 'grey')) +
+  geom_errorbar(aes(ymin = y - error, ymax = y + error), width = 0) +
+  facet_wrap(~ source) +
+  ylim(0, 1) +
+  ylab("% total errors")
+```

### 2017-07-20 analysis/cogSci_analysis.Rmd
Fix data paths; add new model vanity plot

@@
 
-criticalMessagesExp1 <- read.csv("../experiment1/data/message/analysis/messagesExp1.csv") %>%
+criticalMessagesExp1 <- read.csv("./experiment1/data/processedLabelsExp1.csv") %>%
   left_join(turkerRatings, by = relevantProperties) %>%
   mutate(exp = "Exp.1")
-criticalMessagesExp2 <- read.csv("../analysis/messagesExp2.csv") %>%
+criticalMessagesExp2 <- read.csv("../experiment2/data/processedLabelsExp2.csv") %>%
   left_join(turkerRatings, by = relevantProperties) %>%
@@
 ```
+
+```{r}
+d.model <- data.frame(x = rep(c('red fish', 'blue fish'), 2),
+                      y = c(0.27, 0.73, 0, 1),
+                      source = c(rep('with speaker context-uncertainty', 2),
+                                 rep('without speaker context-uncertainty', 2)))
+ggplot(d.model, aes(x=x, y=y, fill = x)) +
+  geom_bar(stat='identity', position = 'dodge') +
+  theme_few() +
+  scale_fill_manual(values = c('red', 'blue')) +
+  facet_wrap(~ source) +
+  ylab("")
+```

### 2017-08-26 writing/journal_manuscript/pnas-submission.tex
Reformat writing for pnas; revise abstract/sig statement

--- 2016-05-13 writing/cogsci-revision/cogsci_keysar.tex
@@
-% arara: pdflatex
-% arara: bibtex
-% arara: pdflatex
-% arara: pdflatex
-% 
-% Annual Cognitive Science Conference
-% Sample LaTeX Paper -- Proceedings Format
-%
+\documentclass[9pt,twocolumn,twoside]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
 
-% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
-% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
-% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
-% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
-% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
-% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
-% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
-% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
-% Modified : Ken Forbus                              01/23/2004
-% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
-% Modified : Niels Taatgen (taatgen@cmu.edu)         10/24/2006
-% Modified : David Noelle (dnoelle@ucmerced.edu)     11/19/2014
+\templatetype{pnasresearcharticle} % Choose template 
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} = Template for a one-column mathematics article
+% {pnasinvited} = Template for a PNAS invited submission
 
-%% Change "letterpaper" in the following line to "a4paper" if you must.
-
-\documentclass[10pt,letterpaper]{article}
-
-\usepackage{hyperref}
-\usepackage{cogsci}
 \usepackage{pslatex}
@@
 \usepackage{graphicx}
-\usepackage{apacite}
 \usepackage{color}
@@
 
-\definecolor{Red}{RGB}{255,0,0}
-\newcommand{\red}[1]{\textcolor{Red}{#1}}
-
-\newcommand{\jd}[1]{\green{$^*$}\marginpar{\footnotesize{JD: \green{#1}}}}
-\definecolor{Green}{RGB}{10,200,100}
-\definecolor{Red}{RGB}{200,10,100}
-
-\newcommand{\ndg}[1]{\textcolor{Green}{[ndg: #1]}}
-\newcommand{\rdh}[1]{\textcolor{Red}{[rdh: #1]}}
-
-\newcommand{\subsubsubsection}[1]{{\em #1}}
-\newcommand{\eref}[1]{(\ref{#1})}
-\newcommand{\tableref}[1]{Table \ref{#1}}
-\newcommand{\figref}[1]{Figure \ref{#1}}
-\newcommand{\appref}[1]{Appendix \ref{#1}}
-\newcommand{\sectionref}[1]{Section \ref{#1}}
-
 \title{Conversational expectations account for apparent limits on theory of mind use}
 
-\author{{\large \bf Robert X.~D.~Hawkins, Noah D.~Goodman}\\
-  \{rxdh,ngoodman\}@stanford.edu\\
-  Department of Psychology, 450 Serra Mall \\
-  Stanford, CA 94305 USA}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Robert Hawkins}
+\author[a]{Hyowon Gweon} 
+\author[a]{Noah Goodman}
+
+\affil[a]{Department of Psychology, Stanford University}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Hawkins}
+
+% TODO: Currently a bit too long (and probably not 'significant' sounding enough) 
+\significancestatement{To communicate successfully, speakers must flexibly adjust the specificity of their utterances to the level demanded by the current context.
+In contexts where a speaker is unsure which objects their partner may be considering, it is helpful to provide additional disambiguating information: for example, saying "cassette tape" instead of the potentially ambiguous ``tape." %A listener who rationally expects such levels of specificity may, in turn, draw surprising inferences when they hear a less specific utterance than the situation demands.
+We argue that influential arguments for limits on Theory of Mind (ToM) in adults have neglected the nuanced pragmatics of the communication tasks they rest upon.
+A careful analysis of these pragmatics demands using a rational model of communication suggests a minimal pair of replications showing that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.}
+
+% Please include corresponding author, author contribution and author declaration information
+\authorcontributions{Please provide details of author contributions here.}
+\authordeclaration{Please declare any conflict of interest here.}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+\keywords{Pragmatics $|$ Theory of mind $|$ Social cognition $|$ Replication}
+
+\begin{abstract}
+Humans are capable of accurately reasoning about other minds, but how ubiquitously do we use this ability in everyday life?
+Recent theories of pragmatic language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication.
+To reconcile these points of view, we focus attention on the problem faced by the \emph{speaker.} Through a minimal pair of replications comparing scripted vs.
+unscripted speakers, we show that scripted confederates used in previous studies are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation.
+When we allow participants to naturally interact, we find that listener expectations are justified and errors are reduced.
+This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior--that is, to theory of mind.
+We formalize this argument in a computational model of pragmatic reasoning, and show that it provides a good fit with several recent findings.
+\end{abstract}
+
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 
@@
 
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalised the article contents.
+\verticaladjustment{-2pt}
+
 \maketitle
-
-\begin{abstract}
-Theory of mind is a powerful cognitive ability: by the age of six, people are capable of accurately reasoning about others' beliefs and desires.
-An influential series of language understanding experiments by Keysar and colleagues, however, showed that adults systematically failed to take a speaker's beliefs into account, revealing limitations on theory of mind.
-In this paper we argue that these apparent failures are in fact successes.
-Through a minimal pair of replications comparing scripted vs.
-unscripted speakers, we show that critical utterances used by Keysar and colleagues are uncooperative: they are less informative than what a speaker would actually produce in that situation.
-When we allow participants to naturally interact, we find that listener expectations are justified and errors are reduced.
-This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.
-
-\textbf{Keywords:} 
-Theory of mind; social cognition; pragmatics
-\end{abstract}
-
-\section{Introduction}
-\label{sec:intro}
-
-\begin{figure*}[bt!]
-\begin{center}
-\includegraphics[scale = .37]{images/overall_view.png}
-\end{center}
-\vspace{-.5cm}
-\caption{Interface used in the reported experiments. Objects behind the black squares were hidden from the director.}
-\label{fig:interface}
-\end{figure*}
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
@@
 For example, imagine a roll of tape were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `tape' to mean the occluded roll of tape (which the director couldn't possibly know about).
-Indeed, \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.
+Indeed, \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.
 Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.
@@
 While these results are compelling, the paradigm has been criticized from few different angles. 
-\citeA{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
+\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
 the hidden roll of tape vs.
@@
 % Although relative fit was not empirically evaluated prior to our study,
-Moreover, \citeA{HannaTanenhausTrueswell03_CommonGroundPerspective} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
+Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
 %Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g.
@@
 \end{tabular}
-\caption{Side-by-side comparison of error rates in \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications.
+\caption{Side-by-side comparison of error rates in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications.
 The first two sections show the percentage of \emph{participants} attempting to move the occluded distractor at least once, or twice, of the four possible cases.
@@
 
-We began by replicating \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment.
+We began by replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment.
 We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study.
@@
 
-\section{Expt.~1: Scripted Replication}
+%% mention rubio fernandez early?
+Using similar reasoning, \cite{RubioFernandez16_DirectorTaskAttention} showed that listeners became suspicious of what was actually in common ground after hearing overinformative utterances, thus successfully using theory of mind reasoning.
+
+\section*{Referring in an uncertain visual environment}
+
+In this section, I formalize the director-matcher task inside a rational speech act (RSA) model, where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language.
+This allows us to test pragmatic accounts of task performance.
+
+Previous theoretical advances in modeling pragmatics have typically hinged on `lifting' inference over some aspect of communication that was previously assumed to be fixed.
+For instance, \emph{manner implicatures} emerge from uncertainty over the true meanings of words in the lexicon \cite{BergenLevyGoodman16_LexicalUncertainty}, nonliteral constructions like \emph{hyperbole} and \emph{irony} emerge from uncertainty over the topic of discussion \cite{KaoWuBergenGoodman14_NonliteralNumberWords}, and \emph{prosody} and \emph{ellipsis} emerge from uncertainty (in the form of a noisy-channel model) over the various ways that a speaker's utterance could have been corrupted  \cite{BergenGoodman15_StrategicUseOfNoise}.
+Note that these are not ad hoc extensions; sources of uncertainty that are not strictly necessary to explain a particular effect are often omitted for simplicity, but they all coexist within the same model.
+
+Pragmatic critiques of the director-matcher task point to an additional source of uncertainty that has not previously been explored: uncertainty over a communication partner's visual context.
+Considering this uncertainty, of course, \emph{is} theory of mind in action; our goal is to show how such a model built on social reasoning can produce the `failures' found by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} as well as the phenomena observed by both \cite{HawkinsGoodman16_Keysar} and \cite{RubioFernandez16_DirectorTaskAttention}.
+For simplicity, we consider the full $2\times 2$ grid containing a single occluded cell, which was used by \cite{RubioFernandez16_DirectorTaskAttention}.
+We incorporate this source of uncertainty into both the speaker and listener models as follows:
+
+\begin{enumerate}
+\item A pragmatic speaker perceives an array of three objects $\mathcal{O}_S = \{o_i\}$, one of which is identified as a target $t$, but assumes that their partner sees an additional hidden object $o_h$ that is not in common ground.
+The speaker has uncertainty over the identity of this object, represented by a prior distribution $P(o_h)$ which they marginalize over when deciding what utterance to produce:
+$$S_1(u | t, \mathcal{O}_p) \propto \sum_{o_h} L_0(t| \mathcal{O}_S \cup \{o_h\})P(o_h)P(u)$$
+
+\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$.
+The listener, however, also has some uncertainty over whether the experimenter is deceptive, represented by a Bernoulli random variable $d$, such that they assign a small prior probability $P(d)$ to the possibility that $o_h$ is actually in common ground.
+Upon hearing an utterance$u$, they update both their beliefs about which object is the target and whether the experimenter is being deceptive:
+$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
+\end{enumerate}
+Technical details are omitted here for brevity, but the full model is fully specified in WebPPL and available to run at \url{http://forestdb.org/models/keysar.html}.
+
+We focus on two critical patterns of results produced by the pragmatic listener $L_2$.
+First, we consider how it responds to the scripted instruction ``fish'' when presented with a context containing two fish (a red one that is occluded and a blue one in common ground) as well as two other unrelated objects.
+While it is most likely to choose the red fish in common ground, it has a relatively high probability (43\%) of choosing the hidden red fish, an effect \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} claimed was due to failure of theory of mind.
+This decision emerges in the model for two primary reasons.
+For one, a speaker with uncertainty over context could in principle be communicating about any object, and if they believed the hidden object was a red fish, the utterance `fish' would have a high probability of being produced.
+To a lesser extent, context uncertainty makes the overinformative utterance `blue fish' more likely than `fish', hence the omission of the color modifier has the (subtle) implication of \emph{not} the blue fish.
+This speaker phenomenon, while subtle under the parameter regime of the current implementation, produces the dynamic observed in the unscripted replication by  \cite{HawkinsGoodman16_Keysar}.
+
+Second, we consider the listener's posterior over the \emph{deception} Bernoulli variable $d$.
+Just as \cite{RubioFernandez16_DirectorTaskAttention} found that listeners were more likely to suspect deception after hearing an `overinformative' utterance like `blue fish' when only one fish was in common ground, so too does our pragmatic listener.
+The pragmatic speaker is relatively more likely to produce the additional color modifier if they see multiple fish than when they see a single fish and simply have uncertainty over the unknown object.
+Hence, our model simultaneously captures all three phenomena.
+
+\section*{Expt.~1: Scripted Replication}
 \label{sec:Exp1}
 
-\subsection{Participants}
+\subsection*{Participants}
 
@@
 
-\subsection{Materials \& Procedures}
+\subsection*{Materials \& Procedures}
 
@@
 The rest of the objects were unrelated `fillers' placed in random locations.
-We used the same set of targets and occluded alternatives as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, but we were unable to obtain the filler objects from the original experiment and created our own.
+We used the same set of targets and occluded alternatives as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, but we were unable to obtain the filler objects from the original experiment and created our own.
 Before entering the game environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface.
@@
 For half the instructions, directors were free to communicate however they wished.
-For the other half, including all the critical instructions, their messages to the matcher were pre-scripted using the precise wording from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+For the other half, including all the critical instructions, their messages to the matcher were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat window, and they would have to click `send' for the experiment to continue.)
 
-We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 Each participant received half the items in the experimental condition and half in the baseline condition.
@@
 
-This paradigm differs from those used by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways.
+This paradigm differs from those used by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways.
 First, participants were not seated across from each other at a table: they each saw a view of the 4 x 4 grid on their screen and communicated via a text box.
@@
 We randomly assigned one of the players to the role of the instruction giver, and maintained the original wording by scripting a subset of their instructions.
-Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \citeA{KeysarBarr___Brauner00_TakingPerspective}.
+Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \cite{KeysarBarr___Brauner00_TakingPerspective}.
 
@@
 Except for instructions with scripted messages, the participants were free to use the chat box for questions, small talk, or whatever else.
-\begin{table*}
-\begin{center}
-\begin{tabular}{ m{1cm} | m{1.5cm} | m{1.5cm} | m{1.5cm} |  m{1.25cm} | m{1.6cm} | m{1.4cm} | m{1.1cm} | m{1.3cm} | m{1.3cm} |}
-& & Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
-& instruction & ``glasses'' & ``bottom block'' & ``tape'' & ``large measuring cup'' & ``brush'' & ``eraser'' & ``small candle'' & ``mouse'' \\\hline
-& target & sunglasses & block (3rd row) & cassette & medium cup & round hairbrush & board eraser & medium candle & computer mouse \\\hline
-& hidden distractor & glasses case & block (4th row) & scotch-tape & large cup & flat \,\,\,hairbrush & pencil eraser & small candle & toy mouse \\
- \hline\hline
-\multirow{2}{.75cm}{Expt.~1}  & \# incorrect & 0 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
-                                               & \# correct & 6 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
- \hline\hline
-\multirow{2}{.75cm}{Expt.~2} & \# incorrect & 0 & 1 & 1 & 3 & 9 & 7 & 1 & 5 \\
-                                              & \# correct   & 12&14&11&13&7 &8   &12& 8 \\
+% \begin{table*}
+% \begin{center}
+% \begin{tabular}{ m{1cm} | m{1.5cm} | m{1.5cm} | m{1.5cm} |  m{1.25cm} | m{1.6cm} | m{1.4cm} | m{1.1cm} | m{1.3cm} | m{1.3cm} |}
+% & & Item 1 & Item 2 & Item 3 & Item 4 & Item 5 & Item 6 & Item 7 & Item 8\\\hline
+% & instruction & ``glasses'' & ``bottom block'' & ``tape'' & ``large measuring cup'' & ``brush'' & ``eraser'' & ``small candle'' & ``mouse'' \\\hline
+% & target & sunglasses & block (3rd row) & cassette & medium cup & round hairbrush & board eraser & medium candle & computer mouse \\\hline
+% & hidden distractor & glasses case & block (4th row) & scotch-tape & large cup & flat \,\,\,hairbrush & pencil eraser & small candle & toy mouse \\
+%  \hline\hline
+% \multirow{2}{.75cm}{Expt.~1}  & \# incorrect & 0 & 5 & 1 & 3 & 2 &6 & 6 & 1 \\ 
+%                                                & \# correct & 6 & 3 & 1 & 8 & 2 & 2 & 4 & 6 \\
+%  \hline\hline
+% \multirow{2}{.75cm}{Expt.~2} & \# incorrect & 0 & 1 & 1 & 3 & 9 & 7 & 1 & 5 \\
+%                                               & \# correct   & 12&14&11&13&7 &8   &12& 8 \\
 
-\end{tabular}
-\vspace{-.5cm}
-\caption{Item-wise error rates for critical trials}
-\vspace{-.25cm}
-\label{table:ItemWise}
-\end{center}
-\end{table*}
+% \end{tabular}
+% \vspace{-.5cm}
+% \caption{Item-wise error rates for critical trials}
+% \vspace{-.25cm}
+% \label{table:ItemWise}
+% \end{center}
+% \end{table*}
 
-\subsection{Results and discussion}
+\subsection*{Results and discussion}
 In Table \ref{table:mainResults} we show the error rates on critical items, and compare to the data from
-\citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+\cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 We find that 93\% of participants (all but one) attempted to move the hidden distractor at least once in the Experimental condition, out of four possible items, compared to only 7\% (only one) in the baseline condition.
@@
 
-As a proxy for the eye-tracking analyses reported by \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
+As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
 We define the decision window as the span of time between the point when the matcher received their instruction message and when they started moving an object.
@@
 %
-%In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+%In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it.
@@
 
- \begin{figure}[b!]
-\begin{center}
-\vspace{-.5cm}
-\includegraphics[scale=.5]{images/mousetracking.pdf}
-\caption{Mean percentage of time spent hovering over target cell in the two conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are boostrapped 95\% confidence intervals.}
-\vspace{-.5cm}
-\label{fig:hover}
-\end{center}
-\end{figure}
+%  \begin{figure}[b!]
+% \begin{center}
+% \vspace{-.5cm}
+% \includegraphics[scale=.5]{images/mousetracking.pdf}
+% \caption{Mean percentage of time spent hovering over target cell in the two conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are boostrapped 95\% confidence intervals.}
+% \vspace{-.5cm}
+% \label{fig:hover}
+% \end{center}
+% \end{figure}
 
-\section{Expt.~2: Unscripted Replication}
+\section*{Expt.~2: Unscripted Replication}
 \label{sec:Exp2}
 
-\subsection{Participants}
+\subsection*{Participants}
 
@@
 
-\subsection{Results}
+\subsection*{Results}
 
-Error rates are reported in Table \ref{table:mainResults}, alongside the results from \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our scripted replication in Expt.~1.
+Error rates are reported in Table \ref{table:mainResults}, alongside the results from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our scripted replication in Expt.~1.
 Participants never moved the hidden object in the baseline condition, and total error rates for experimental trials are significantly lower than the rates found in Expt.~1, $\chi^2(1) = 5.35, p = 0.02$.
@@
 
-\begin{figure}[t!]
-\begin{center}
-\vspace{-.25cm}
-\includegraphics[scale=.5]{images/fitnessInteraction.pdf}
-\vspace{-.5cm}
-\caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
-\vspace{-.5cm}
-\label{fig:fitnessInteraction}
-\end{center}
-\end{figure}
+% \begin{figure}[t!]
+% \begin{center}
+% \vspace{-.25cm}
+% \includegraphics[scale=.5]{images/fitnessInteraction.pdf}
+% \vspace{-.5cm}
+% \caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
+% \vspace{-.5cm}
+% \label{fig:fitnessInteraction}
+% \end{center}
+% \end{figure}
 
@@
 Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
-\begin{figure}[t!]
-\begin{center}
-\vspace{-.25cm}
-\includegraphics[scale=.55]{images/itemwisefitness.pdf}
-\vspace{-.75cm}
-\caption{Item-wise error rates compared with label fitness ratios, across both experiments.
+% \begin{figure}[t!]
+% \begin{center}
+% \vspace{-.25cm}
+% \includegraphics[scale=.55]{images/itemwisefitness.pdf}
+% \vspace{-.75cm}
+% \caption{Item-wise error rates compared with label fitness ratios, across both experiments.
 Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest posterior density intervals for error rates.}
-\vspace{-.5cm}
-\label{fig:itemLevel}
-\end{center}
-\end{figure}
+% \vspace{-.5cm}
+% \label{fig:itemLevel}
+% \end{center}
+% \end{figure}
 
@@
 
-\section{General Discussion}
+\section*{General Discussion}
 
@@
 Instead, they said ``the bluish block with a B'' or ``block with the blue writing,'' which relied on less confusable perceptual features and thereby decreased error rates.
-While the stimuli
 Thus, the errors observed in Expt.~1 can be explained as a result of an uncooperative confederate (speaker)---the experiment set up certain pragmatic expectations through the task context, then deliberately flouted them in critical trials.
@@
 %In the current task, it seems particularly relevant that both participants knew about the presence of hidden objects, increasing uncertainty about possible distractors and placing additional pressure on the speaker to overspecify.
-The exact origin is a subject that must be followed up by future research \cite<e.g.>{GannBarr14_AudienceDesign}, but it is clear that the tendency of speakers to produce highly informative referring expressions is useful to, and relied on by, listeners.
+The exact origin is a subject that must be followed up by future research \cite[e.g.]{GannBarr14_AudienceDesign}, but it is clear that the tendency of speakers to produce highly informative referring expressions is useful to, and relied on by, listeners.
 %Indeed, the remaining errors in Expt.~2 are at least partially due to the lack of availability of more specific expressions. 
@@
 
-This finding is consistent with a recent proposal by \citeA{HellerParisienStevenson16_ProbabilisticWeighing} that referring expressions are interpreted by probabilistically integrating multiple sources of information: when conversational expectations lead participants to expect over-informative utterances, they (rationally) place relatively less weight on features of the environment determining common ground, such as shared perceptual access.
+This finding is consistent with a recent proposal by \cite{HellerParisienStevenson16_ProbabilisticWeighing} that referring expressions are interpreted by probabilistically integrating multiple sources of information: when conversational expectations lead participants to expect over-informative utterances, they (rationally) place relatively less weight on features of the environment determining common ground, such as shared perceptual access.
 Further work in a wider variety of tasks is necessary to pin down the various factors determining the relative weighting of these different sources of information, but we have argued that pragmatics play a crucial role.
 
-It is worth noting several significant differences between our study and \citeA{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+It is worth noting several significant differences between our study and \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room.
@@
 
-\section{Acknowledgements}
+\subsection*{Supporting Information (SI)}
 
-\small We're grateful to Boaz Keysar for providing select materials for our replication.
-Expt.~1 was originally conducted under the supervision of Michael Frank, with early input from Desmond Ong.
-This work was supported by ONR grants N00014-13-1-0788 and N00014-13-1-0287,  and a James S.
-McDonnell Foundation Scholar Award to NDG.
-RXDH was supported by the Stanford Graduate Fellowship and the National Science Foundation Graduate Research Fellowship under Grant No.
-DGE-114747.
+The main text of the paper must stand on its own without the SI.
+Refer to SI in the manuscript at an appropriate point in the text.
+Number supporting figures and tables starting with S1, S2, etc.
+Authors are limited to no more than 10 SI files, not including movie files.
+Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods.
+If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
+Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only.
 
-\bibliographystyle{apacite}
+\subsubsection*{SI Text}
 
-\setlength{\bibleftmargin}{.125in}
-\setlength{\bibindent}{-\bibleftmargin}
+Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).
 
-\bibliography{bibs}
+\subsubsection*{SI Figures}
+
+Provide a brief legend for each supporting figure after the supporting text.
+Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text.
+When saving TIFF files, use only LZW compression; do not use JPEG compression.
+Do not save figure numbers, legends, or author names as part of the image.
+Composite figures must be pre-assembled.
+
+\subsubsection*{3D Figures}
+
+Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.
+
+\subsubsection*{SI Tables}
+
+Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
+Do not use tabs or spaces to separate columns in Word tables.
+
+\subsubsection*{SI Datasets}
+
+Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
+
+\subsubsection*{SI Movies}
+
+Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file.
+All movies should be submitted at the desired reproduction size and length.
+Movies should be no more than 10 MB in size.
+
+\subsubsection*{Still images}
+
+Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
+
+\subsubsection*{Appendices}
+
+PNAS prefers that authors submit individual source files to ensure readability.
+If this is not possible, supply a single PDF file that contains all of the SI associated with the paper.
+This file type will be published in raw format and will not be edited or composed.
+
+\matmethods{Please describe your materials and methods here.
+This can be more than one paragraph, and may contain subsections and equations as required.
+Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.
+
+\subsection*{Subsection for Method}
+Example text for subsection.
+}
+
+\showmatmethods{} % Display the Materials and Methods section
+
+\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}
+
+\showacknow{} % Display the acknowledgments section
+
+% \pnasbreak splits and balances the columns before the references.
+% Uncomment \pnasbreak to view the references in the PNAS-style
+% If you see unexpected formatting errors, try commenting out \pnasbreak
+% as it can run into problems with floats and footnotes on the final page.
+%\pnasbreak
+
+% Bibliography
+\bibliography{pnas-sample}
 

### 2017-08-27 writing/journal_manuscript/pnas-submission.tex
Change bib name
Updated abstract

@@
 
-% TODO: Currently a bit too long (and probably not 'significant' sounding enough) 
-\significancestatement{To communicate successfully, speakers must flexibly adjust the specificity of their utterances to the level demanded by the current context.
-In contexts where a speaker is unsure which objects their partner may be considering, it is helpful to provide additional disambiguating information: for example, saying "cassette tape" instead of the potentially ambiguous ``tape." %A listener who rationally expects such levels of specificity may, in turn, draw surprising inferences when they hear a less specific utterance than the situation demands.
+% TODO: Currently a bit too long (and probably too in the weeds...) 
+\significancestatement{To communicate successfully, speakers must flexibly adapt %the specificity of 
+their utterances to the demands of the current context.
+In contexts  which other objects are in view of their partnerrying to refer to an object in  , it is helpful to provide additional disambiguating information: for example, saying "cassette tape" instead of the potentially ambiguous ``tape." %A listener who rationally expects such levels of specificity may, in turn, draw surprising inferences when they hear a less specific utterance than the situation demands.
 We argue that influential arguments for limits on Theory of Mind (ToM) in adults have neglected the nuanced pragmatics of the communication tasks they rest upon.
-A careful analysis of these pragmatics demands using a rational model of communication suggests a minimal pair of replications showing that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.}
+A careful analysis of these pragmatic demands using a rational model of communication suggests a minimal pair of replications which show that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.}
 
@@
 Humans are capable of accurately reasoning about other minds, but how ubiquitously do we use this ability in everyday life?
-Recent theories of pragmatic language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication.
-To reconcile these points of view, we focus attention on the problem faced by the \emph{speaker.} Through a minimal pair of replications comparing scripted vs.
-unscripted speakers, we show that scripted confederates used in previous studies are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation.
-When we allow participants to naturally interact, we find that listener expectations are justified and errors are reduced.
-This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior--that is, to theory of mind.
-We formalize this argument in a computational model of pragmatic reasoning, and show that it provides a good fit with several recent findings.
+Recent theories of language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication.
+To reconcile these points of view, we use a computational model of communication under uncertainty to analyze the pragmatic demands of a common theory of mind test.
+Through a minimal pair of replications comparing scripted vs.
+unscripted speakers, we show that scripted confederates used in previous studies are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation, thus violating the listener's rational social expectations.
+This ironically shows that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.
 \end{abstract}
@@
 Humans can accurately and intelligently reason about the mental states of other humans.
-Among other things, this ability -- called \emph{theory of mind} \cite{PremackWoodruff78_ChimpanzeeToM} -- allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}.
-Children acquire this ability by at least age six \cite{WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
+Theory of mind \cite{PremackWoodruff78_ChimpanzeeToM} allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}.
+%Children acquire this ability by at least age six \cite{WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
 %
@@
 Using similar reasoning, \cite{RubioFernandez16_DirectorTaskAttention} showed that listeners became suspicious of what was actually in common ground after hearing overinformative utterances, thus successfully using theory of mind reasoning.
+
+%% Mention Westra & Carruthers as another critique focusing on pragmatics?
 
@@
 % Bibliography
-\bibliography{pnas-sample}
+\bibliography{pnas-submission}
 

### 2017-09-17 writing/journal_manuscript/pnas-submission.tex
Add placeholder to talk about communication early
Work on intro

@@
 their utterances to the demands of the current context.
-In contexts  which other objects are in view of their partnerrying to refer to an object in  , it is helpful to provide additional disambiguating information: for example, saying "cassette tape" instead of the potentially ambiguous ``tape." %A listener who rationally expects such levels of specificity may, in turn, draw surprising inferences when they hear a less specific utterance than the situation demands.
+When they are unsure exactly which objects are in view of their partner, for example, they may provide additional disambiguating information when referring to a particular object: saying "cassette tape" instead of the potentially ambiguous ``tape." %A listener who rationally expects such levels of specificity may, in turn, draw surprising inferences when they hear a less specific utterance than the situation demands.
 We argue that influential arguments for limits on Theory of Mind (ToM) in adults have neglected the nuanced pragmatics of the communication tasks they rest upon.
-A careful analysis of these pragmatic demands using a rational model of communication suggests a minimal pair of replications which show that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.}
+A careful analysis of pragmatic demands using a rational model of communication suggests a minimal pair of experiments which show that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.}
 
@@
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Pragmatics $|$ Theory of mind $|$ Social cognition $|$ Replication}
+\keywords{Theory of mind $|$ Pragmatics $|$ Communication $|$ Social cognition $|$ Replication}
 
 \begin{abstract}
-Humans are capable of accurately reasoning about other minds, but how ubiquitously do we use this ability in everyday life?
+%Humans are capable of accurately reasoning about other minds, but how ubiquitously do we use this ability in everyday life? 
 Recent theories of language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication.
@@
 
-Humans can accurately and intelligently reason about the mental states of other humans.
-Theory of mind \cite{PremackWoodruff78_ChimpanzeeToM} allows us to infer the underlying beliefs and intentions that motivate others' actions, and to use these inferences to predict future actions \cite{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning}.
+Our success as a cooperative species depends upon our ability to accurately predict the behavior of other agents.
+The underlying mental representation we use to make such predictions --- our theory of mind \cite{PremackWoodruff78_ChimpanzeeToM} --- has been increasingly well-understood as a generative model of rational action \cite{KileyHamlinEtAl13_MentalisticCoreSocial, BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}.
+From a young age, we can invert this model to infer another agent's likely mental states given their past behavior, and simulate their future behavior to form sophisticated expectations about how they would act in a range of scenarios \cite{SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM}.
+Yet the extent to which theory of mind is deployed in \emph{communication}, an especially ubiquitous and high-stakes site for social interaction, remains controversial.
+A large body of evidence has demonstrated striking failures of perspective-taking, where listeners appear to be ``mind-blind,'' neglecting the visual access of their partner when interpreting their utterances \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
+
+While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, we argue in this paper that the rational models speakers and listeners use to reason about their conversation partners are in fact much richer.
+Communication abounds with social expectations.
+We expect our conversation partners to follow common conventions about the meanings of words \cite{Lewis69_Convention, CentolaBaronchelli15_ConventionEmergence}, to update these conventions in light of previous interactions \cite{BrennanClark96_ConceptualPactsConversation}, and to avoid saying things that are confusing or unnecessarily complicated in the current context \cite{Grice75_LogicConversation}.
+
+%The computational mechanisms underlying theory of mind have been increasingly well-understood as accurate predictions about another agent's future actions using a mental model of their beliefs and desires \cite{PremackWoodruff78_ChimpanzeeToM}, so too .
+
+%The underlying mechanisms that support our ability to communicate so flexibly and effortlessly?
+These expectations derive from in internal model of our partner's mind Making these predictions about our partner requires a model of their mind.
+rich models of someone's beliefs and desires from observing them act in the world, observe someone's behavior, we  infer the beliefs and desires
+
+% rational models of other minds to infer predict what our friends like (e.g.
+, Baker, Jara-Ettinger, etc), depends on a mental model of our conversation partner.
+to predict what they will say or how they will interpret our words.
+
+%speakers and listeners adapt their language use to their shared history and to the immediate context in rich, intricate ways \cite{Grice75_LogicConversation, Clark96_UsingLanguage}.
+For instance, in a context where multiple dogs may be salient to the listener, speakers introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \cite{BrennanClark96_ConceptualPactsConversation, GrafEtAl16_BasicLevel}.
+Listeners, for their part, draw ad-hoc implicatures based on expectations that speakers will give contextually sensitive descriptions \cite{BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature} and \dots
+
+%What underlying mechanisms support our ability to communicate so flexibly and effortlessly? 
 %Children acquire this ability by at least age six \cite{WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
 %
-While theory of mind use often appears to be automatic and effortless, Keysar and colleagues \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind} have argued that it is actually the opposite, even for adults: we are ``mindblind" by default and only overcome our egocentric biases through an effortful process of perspective-taking.
+
+% While theory of mind use often appears to be automatic and effortless, Keysar and colleagues \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind} have argued that it is actually the opposite, even for adults: we are ``mindblind" by default and only overcome our egocentric biases through an effortful process of perspective-taking.
 In other words, while adults are \emph{capable} of applying theory of mind reasoning, we do not always apply it reliably.
@@
 % Although relative fit was not empirically evaluated prior to our study,
-Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
+Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
 %Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g.
@@
 %% Mention Westra & Carruthers as another critique focusing on pragmatics?
+
+%% Mention neural coordination papers? e.g. Stephens et al 2010, Stolk et al, 2014?
 

### 2017-09-19 writing/journal_manuscript/pnas-submission.tex
Refocus intro; rearrange expts

@@
 
-Our success as a cooperative species depends upon our ability to accurately predict the behavior of other agents.
-The underlying mental representation we use to make such predictions --- our theory of mind \cite{PremackWoodruff78_ChimpanzeeToM} --- has been increasingly well-understood as a generative model of rational action \cite{KileyHamlinEtAl13_MentalisticCoreSocial, BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}.
-From a young age, we can invert this model to infer another agent's likely mental states given their past behavior, and simulate their future behavior to form sophisticated expectations about how they would act in a range of scenarios \cite{SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM}.
-Yet the extent to which theory of mind is deployed in \emph{communication}, an especially ubiquitous and high-stakes site for social interaction, remains controversial.
-A large body of evidence has demonstrated striking failures of perspective-taking, where listeners appear to be ``mind-blind,'' neglecting the visual access of their partner when interpreting their utterances \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
+Our success as a cooperative species depends upon our ability to make sense of other agents. 
+A natural solution to this problem is to maintain an internal predictive model of how minds work: how an agent's unobservable mental states give rise to observable actions. 
+With such a model in mind, people can make inferences about an agent's likely mental state given their past behavior and run simulations to form sophisticated expectations about their future behavior.
+This theoretical approach to social cognition, where theory of mind reasoning \cite{PremackWoodruff78_ChimpanzeeToM, WellmanCrossWatson01_ToMMetaAnalysis, SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM} is grounded in an explicit computational model of how other minds work, has been useful for understanding a number of experimental phenomena in both adults and children \cite{KileyHamlinEtAl13_MentalisticCoreSocial, PantelisEtAl14_IntentionInference, BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}.
+Yet our current understanding of theory of mind use in \emph{communication} remains limited.
+Past efforts have largely focused on a relatively narrow behavioral operationalization of theory of mind as literal perspective-taking about visual access \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
 
-While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, we argue in this paper that the rational models speakers and listeners use to reason about their conversation partners are in fact much richer.
-Communication abounds with social expectations.
-We expect our conversation partners to follow common conventions about the meanings of words \cite{Lewis69_Convention, CentolaBaronchelli15_ConventionEmergence}, to update these conventions in light of previous interactions \cite{BrennanClark96_ConceptualPactsConversation}, and to avoid saying things that are confusing or unnecessarily complicated in the current context \cite{Grice75_LogicConversation}.
+%While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
+We argue in this paper that the mental models speakers and listeners use to reason about their partners in communication are in fact much richer than previously considered.
+By formalizing and rigorously comparing the predicted behavior of speakers and listeners using different mental models of one another, we show that theory of mind is required to account for communicative behavior in reference games.
+This approach also allows us to understand the sophisticated pragmatics underlying behavior that, under more naive behavioral measures, appeared to show limits.
 
-%The computational mechanisms underlying theory of mind have been increasingly well-understood as accurate predictions about another agent's future actions using a mental model of their beliefs and desires \cite{PremackWoodruff78_ChimpanzeeToM}, so too .
+Communication abounds with pragmatic social expectations.
+We expect our conversation partners to follow common conventions about the meanings of words \cite{Lewis69_Convention, CentolaBaronchelli15_ConventionEmergence}, to update these conventions in light of previous interactions \cite{BrennanClark96_ConceptualPactsConversation}, and to avoid saying things that are confusing or unnecessarily complicated in the current context \cite{Grice75_LogicConversation}.For instance, in a context where multiple dogs may be salient to the listener, speakers introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \cite{BrennanClark96_ConceptualPactsConversation, GrafEtAl16_BasicLevel}.
+Listeners, for their part, draw ad-hoc implicatures based on expectations that speakers will give contextually sensitive descriptions \cite{BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}
 
 %The underlying mechanisms that support our ability to communicate so flexibly and effortlessly?
-These expectations derive from in internal model of our partner's mind Making these predictions about our partner requires a model of their mind.
-rich models of someone's beliefs and desires from observing them act in the world, observe someone's behavior, we  infer the beliefs and desires
-
-% rational models of other minds to infer predict what our friends like (e.g.
-, Baker, Jara-Ettinger, etc), depends on a mental model of our conversation partner.
-to predict what they will say or how they will interpret our words.
 
 %speakers and listeners adapt their language use to their shared history and to the immediate context in rich, intricate ways \cite{Grice75_LogicConversation, Clark96_UsingLanguage}.
-For instance, in a context where multiple dogs may be salient to the listener, speakers introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \cite{BrennanClark96_ConceptualPactsConversation, GrafEtAl16_BasicLevel}.
 Listeners, for their part, draw ad-hoc implicatures based on expectations that speakers will give contextually sensitive descriptions \cite{BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature} and \dots
-
-%What underlying mechanisms support our ability to communicate so flexibly and effortlessly? 
-%Children acquire this ability by at least age six \cite{WellmanCrossWatson01_ToMMetaAnalysis} and it serves as an important landmark in the developmental trajectory of intuitive theory use \cite{GopnikWellman12_ReconstructingConstructivism}.
-%
 
@@
 
-In this paper we argue that the apparent failures used to support this view are in fact successes for sophisticated social reasoning.
-In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
+% In this paper we argue that the apparent failures used to support this view are in fact successes for sophisticated social reasoning.
 
@@
 
-In this paper, we offer an additional factor that helps account for Keysar's results. 
+%In this paper, we offer an additional factor that helps account for Keysar's results. 
 Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations.
@@
 
-\section*{Expt.~1: Scripted Replication}
+\section*{Expt.~1: Unscripted Replication}
 \label{sec:Exp1}
+
+\subsection*{Participants}
+
+We recruited 64 participants (32 pairs) from Amazon Mechanical Turk, roughly doubling the sample size from Expt.~1.
+All participants were from the U.S.
+Three participants were excluded for making 2 or more errors on non-critical items, and one additional participant was excluded because they were not a native English speaker.
+
+\subsection*{Materials \& Procedures}
+
+Everything was the same as Expt.~1, except we did not use scripted messages for critical instructions.
+
+\subsection*{Results}
+
+Error rates are reported in Table \ref{table:mainResults}, alongside the results from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our scripted replication in Expt.~1.
+Participants never moved the hidden object in the baseline condition, and total error rates for experimental trials are significantly lower than the rates found in Expt.~1, $\chi^2(1) = 5.35, p = 0.02$.
+
+We find that patterns of errors in Expt.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
+Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the more precise utterances used remain ambiguous.
+For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
+Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' from Expt.~1, but the two objects were still confusable at the sub-class level.
+If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.
+
+When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we found no significant decrease in target hover time between experimental and baseline trials $t(27) = 0.89, p = 0.38$.
+To directly test for differences in hover time patterns across the two experiments, we used a mixed-effects model with a random intercept for game ID and an interaction between condition and experiment on target hover time.
+We found a marginally significant interaction, $b = 0.09, t = 1.89, p = 0.066$ (see Figure \ref{fig:hover}), providing some evidence that the presence of a hidden distractor no longer interfered target selection in Expt~2.
+
+% \begin{figure}[t!]
+% \begin{center}
+% \vspace{-.25cm}
+% \includegraphics[scale=.5]{images/fitnessInteraction.pdf}
+% \vspace{-.5cm}
+% \caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
+% \vspace{-.5cm}
+% \label{fig:fitnessInteraction}
+% \end{center}
+% \end{figure}
+
+Next, we test whether these improvements in performance are in fact due to more informative speaker behavior.
+We recruited twenty judges on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects.
+Their responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.6$ (95\% CI = $[0.54, 0.66]$).
+In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig.
+\ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$.
+In Expt.~1, the scripted label fit the hidden distractor just as well or better than the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
+In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.
+
+Does differential informativity of scripted utterances account for the variability across items that we noted in Expt.~1?
+In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
+We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$.
+This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances.
+
+After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials?
+We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
+We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$.
+Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
+% \begin{figure}[t!]
+% \begin{center}
+% \vspace{-.25cm}
+% \includegraphics[scale=.55]{images/itemwisefitness.pdf}
+% \vspace{-.75cm}
+% \caption{Item-wise error rates compared with label fitness ratios, across both experiments.
+Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest posterior density intervals for error rates.}
+% \vspace{-.5cm}
+% \label{fig:itemLevel}
+% \end{center}
+% \end{figure}
+
+%\begin{itemize}
+%\item people still try to move occluded objects, but they do it less, and that behavior is more confined to a few particularly difficult items.
+%\item this happens because the scripted instructions we forced directors to use in Exp.~1 were unnatural and violated listener expectations of overinformativity
+%\item the fact that the effect is attenuated in exp.
+2 is actually evidence of some theory of mind use (insofar as the director choosing to be overinformative, and the matcher expecting overinformativity constitutes ToM
+%\end{itemize}
+
+\section*{Expt.~2: Scripted Replication}
+\label{sec:Exp2}
+
+A large body of evidence has demonstrated striking failures of perspective-taking, where listeners appear to be ``mind-blind,'' neglecting the visual access of their partner when interpreting their utterances
+
+In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
 
@@
 
-\section*{Expt.~2: Unscripted Replication}
-\label{sec:Exp2}
-
-\subsection*{Participants}
-
-We recruited 64 participants (32 pairs) from Amazon Mechanical Turk, roughly doubling the sample size from Expt.~1.
-All participants were from the U.S.
-Three participants were excluded for making 2 or more errors on non-critical items, and one additional participant was excluded because they were not a native English speaker.
-
-\subsection*{Materials \& Procedures}
-
-Everything was the same as Expt.~1, except we did not use scripted messages for critical instructions.
-
-\subsection*{Results}
-
-Error rates are reported in Table \ref{table:mainResults}, alongside the results from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our scripted replication in Expt.~1.
-Participants never moved the hidden object in the baseline condition, and total error rates for experimental trials are significantly lower than the rates found in Expt.~1, $\chi^2(1) = 5.35, p = 0.02$.
-
-We find that patterns of errors in Expt.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
-Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the more precise utterances used remain ambiguous.
-For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
-Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' from Expt.~1, but the two objects were still confusable at the sub-class level.
-If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.
-
-When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we found no significant decrease in target hover time between experimental and baseline trials $t(27) = 0.89, p = 0.38$.
-To directly test for differences in hover time patterns across the two experiments, we used a mixed-effects model with a random intercept for game ID and an interaction between condition and experiment on target hover time.
-We found a marginally significant interaction, $b = 0.09, t = 1.89, p = 0.066$ (see Figure \ref{fig:hover}), providing some evidence that the presence of a hidden distractor no longer interfered target selection in Expt~2.
-
-% \begin{figure}[t!]
-% \begin{center}
-% \vspace{-.25cm}
-% \includegraphics[scale=.5]{images/fitnessInteraction.pdf}
-% \vspace{-.5cm}
-% \caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
-% \vspace{-.5cm}
-% \label{fig:fitnessInteraction}
-% \end{center}
-% \end{figure}
-
-Next, we test whether these improvements in performance are in fact due to more informative speaker behavior.
-We recruited twenty judges on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects.
-Their responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.6$ (95\% CI = $[0.54, 0.66]$).
-In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig.
-\ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$.
-In Expt.~1, the scripted label fit the hidden distractor just as well or better than the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
-In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.
-
-Does differential informativity of scripted utterances account for the variability across items that we noted in Expt.~1?
-In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
-We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$.
-This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances.
-
-After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials?
-We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
-We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$.
-Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
-% \begin{figure}[t!]
-% \begin{center}
-% \vspace{-.25cm}
-% \includegraphics[scale=.55]{images/itemwisefitness.pdf}
-% \vspace{-.75cm}
-% \caption{Item-wise error rates compared with label fitness ratios, across both experiments.
-Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest posterior density intervals for error rates.}
-% \vspace{-.5cm}
-% \label{fig:itemLevel}
-% \end{center}
-% \end{figure}
-
-%\begin{itemize}
-%\item people still try to move occluded objects, but they do it less, and that behavior is more confined to a few particularly difficult items.
-%\item this happens because the scripted instructions we forced directors to use in Exp.~1 were unnatural and violated listener expectations of overinformativity
-%\item the fact that the effect is attenuated in exp.
-2 is actually evidence of some theory of mind use (insofar as the director choosing to be overinformative, and the matcher expecting overinformativity constitutes ToM
-%\end{itemize}
-
 \section*{General Discussion}

### 2017-09-19 writing/journal_manuscript/pnas-submission.tex
Finish major revision of intro

@@
 
+% Characterize the computational approach to theory of mind we plan to take
+% Point out a gaping hole that could benefit from this approach
 Our success as a cooperative species depends upon our ability to make sense of other agents. 
-A natural solution to this problem is to maintain an internal predictive model of how minds work: how an agent's unobservable mental states give rise to observable actions. 
-With such a model in mind, people can make inferences about an agent's likely mental state given their past behavior and run simulations to form sophisticated expectations about their future behavior.
+A natural solution to this problem is to maintain an internal predictive model of how minds work: how an agent's unobservable beliefs and desires give rise to observable actions. 
+Armed with a computational theory of mind, people can make inferences about an agent's likely mental state given their past behavior and run simulations to form sophisticated expectations about their future behavior.
 This theoretical approach to social cognition, where theory of mind reasoning \cite{PremackWoodruff78_ChimpanzeeToM, WellmanCrossWatson01_ToMMetaAnalysis, SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM} is grounded in an explicit computational model of how other minds work, has been useful for understanding a number of experimental phenomena in both adults and children \cite{KileyHamlinEtAl13_MentalisticCoreSocial, PantelisEtAl14_IntentionInference, BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}.
 Yet our current understanding of theory of mind use in \emph{communication} remains limited. 
-Past efforts have largely focused on a relatively narrow behavioral operationalization of theory of mind as literal perspective-taking about visual access \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
+Theory of mind as applied within communication depends on an accurate model of what a speaker would say, or what a listener would think, in different situations. 
+Past efforts have instead largely focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
 
 %While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
-We argue in this paper that the mental models speakers and listeners use to reason about their partners in communication are in fact much richer than previously considered.
-By formalizing and rigorously comparing the predicted behavior of speakers and listeners using different mental models of one another, we show that theory of mind is required to account for communicative behavior in reference games.
-This approach also allows us to understand the sophisticated pragmatics underlying behavior that, under more naive behavioral measures, appeared to show limits.
+We argue in this paper that the mental models supporting theory of mind use in communication are much richer than past efforts have explored. 
+We consider a class of mental models representing not only a partner's visual access but also Gricean expectations of rationality and cooperativity \cite{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
+For example, suppose a friend wants to refer to a nearby Dalmatian. 
+%There are numerous descriptions that would all be literally true, but different expressions are preferred in different contexts.
+If there are no other dogs nearby, we may expect them to use a parsimonious basic-level term like ``dog.''
+When there is potential for confusion, however, we expect them to introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \cite{BrennanClark96_ConceptualPactsConversation, GrafEtAl16_BasicLevel}.
+When speakers violate these expectations, listeners draw inferences about the intended meaning in order to account for the violation: for example, XXX \cite{XXX, BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}.
 
-Communication abounds with pragmatic social expectations.
-We expect our conversation partners to follow common conventions about the meanings of words \cite{Lewis69_Convention, CentolaBaronchelli15_ConventionEmergence}, to update these conventions in light of previous interactions \cite{BrennanClark96_ConceptualPactsConversation}, and to avoid saying things that are confusing or unnecessarily complicated in the current context \cite{Grice75_LogicConversation}.For instance, in a context where multiple dogs may be salient to the listener, speakers introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \cite{BrennanClark96_ConceptualPactsConversation, GrafEtAl16_BasicLevel}.
-Listeners, for their part, draw ad-hoc implicatures based on expectations that speakers will give contextually sensitive descriptions \cite{BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}
+This intricate dance of expectations has recently been formalized in a probabilistic model of communication as recursive social reasoning.
+The resulting Rational Speech Act (RSA) framework has successfully described many puzzling pragmatic phenomena, including vagueness \cite{Lassiter}, hyperbole \cite{Kao}, irony \cite{}, metaphor, and overinformativity \cite{Judith}.
+In the simplest version of this framework, speakers judge the utility of different utterances proportionate to the probability of correct interpretation under an internal model of a listener.
+Listener, in turn, use their respective internal model of the speaker to reason backward to the most plausible intention \cite{Grice75_LogicConversation,Clark96_UsingLanguage,GoodmanFrank16_RSATiCSFrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
+
+Theory of mind mechanisms are therefore at the very heart of the Rational Speech Act framework, yet they have never been directly compared against egocentric or ``mindblind'' approaches under the same behavioral paradigm.
+By formalizing and rigorously comparing the predicted behavior of speakers and listeners using mental models of one another, we show that theory of mind is required to account for context-sensitive behavior.
+In addition to these novel predictions, this approach also allows us to understand the sophisticated pragmatics underlying behavior that, under more naive behavioral measures, apparently reflect failures of theory of mind.
+%Communication abounds with potential social expectations, and evaluating the extent to which theory of mind is deployed in communication requires considering these expectations \emph{en masse}.
+While egocentric views of perspective-taking make more clear-cut behavioral predictions, the predictions derived from a computational theory of mind are more varied and complicated but ultimately more accurate.
 
@@
 
-The argument offered by Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate.
+%The argument offered by Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate.
 The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects (see Fig.
@@
 The director gave an instruction like `move the tape up one square,' referring to the cassette.
-Critically, some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
+%Critically, some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
 To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about which objects were shared and which were private.
-For example, imagine a roll of tape were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `tape' to mean the occluded roll of tape (which the director couldn't possibly know about).
+%For example, imagine a roll of tape were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `tape' to mean the occluded roll of tape (which the director couldn't possibly know about).
 Indeed, \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.
-Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.
+%Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.
 
-While these results are compelling, the paradigm has been criticized from few different angles. 
-\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
-the hidden roll of tape vs.
-the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well.
-We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
-% Although relative fit was not empirically evaluated prior to our study,
-Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
-%Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g.
-the roll of tape) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}---perhaps people would refer to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results.
-%Note that none of these criticisms invalidate the paradoxical result that listeners move an occluded object, they just suggest an alternative explanation.
-Reference disambiguation is probabilistic: common ground and goodness of fit are weighed against one another.
-
-%In this paper, we offer an additional factor that helps account for Keysar's results. 
-Theory of mind as applied within language understanding depends on an accurate model of what a speaker would say in different situations.
-Given an utterance, a listener can then reason backward to the most plausible situation \cite{Grice75_LogicConversation,Clark96_UsingLanguage,FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}.
-This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
-If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
-More precisely, when both players know that objects are occluded in the display, the speaker may tend to add additional precision to references in order to avoid confusion.
+%This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
+%If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
+%More precisely, when both players know that objects are occluded in the display, the speaker may tend to add additional precision to references in order to avoid confusion.
 If the listener \emph{expects} the speaker to do this, they will pragmatically pick the \emph{a priori} more likely referent of the referring expression, which in critical trials will be the occluded object.
 In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item.
-\begin{table*}
-\begin{center}
-\begin{tabular}{ p{2cm} | r | r |  r || r | r | r || r | r | r}
-& \multicolumn{3}{c||}{\% attempted at least once} & \multicolumn{3}{c||}{\% attempted at least twice} & \multicolumn{3}{c}{\% of total cases}\\
-\hline
-& Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2  \\
-Experimental & 71 & 93 & 61 & 46 & 57 & 32 & 30 & 43 & 24\\
-Baseline        & 0   & 7   & 0   & 0   & 0   & 0   & 0   & 2   & 0\\
-\end{tabular}
-\caption{Side-by-side comparison of error rates in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications.
+%\begin{table*}
+%\begin{center}
+%\begin{tabular}{ p{2cm} | r | r |  r || r | r | r || r | r | r}
+%& \multicolumn{3}{c||}{\% attempted at least once} & \multicolumn{3}{c||}{\% attempted at least twice} & \multicolumn{3}{c}{\% of total cases}\\
+%\hline
+%& Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2 & Orig. & Expt. 1 & Expt. 2  \\
+%Experimental & 71 & 93 & 61 & 46 & 57 & 32 & 30 & 43 & 24\\
+%Baseline        & 0   & 7   & 0   & 0   & 0   & 0   & 0   & 2   & 0\\
+%\end{tabular}
+%\caption{Side-by-side comparison of error rates in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our two replications.
 The first two sections show the percentage of \emph{participants} attempting to move the occluded distractor at least once, or twice, of the four possible cases.
@@
 }
-\label{table:mainResults}
-\end{center}
-\end{table*}
-
-We began by replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment.
+%\label{table:mainResults}
+%\end{center}
+%\end{table*}
+%
+%We began by replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in a multi-player web experiment.
 We recruited participants to be both director and matcher (instead of using a confederate), but instructions for critical items, as well as a random subset of filler items, remained scripted as in the original study.
@@
 This minimal pair of experiments demonstrates that listener mistakes are at least partially due to the pragmatics of the task, ironically showing that apparent failures of theory of mind are in fact attributable to sophisticated expectations about speaker behavior---that is, to theory of mind.
-
-%% mention rubio fernandez early?
-Using similar reasoning, \cite{RubioFernandez16_DirectorTaskAttention} showed that listeners became suspicious of what was actually in common ground after hearing overinformative utterances, thus successfully using theory of mind reasoning.
-
-%% Mention Westra & Carruthers as another critique focusing on pragmatics?
-
-%% Mention neural coordination papers? e.g. Stephens et al 2010, Stolk et al, 2014?
 
@@
 
+% discuss other criticisms of Keysar paradigm?
+While these results are compelling, the paradigm has been criticized from few different angles. 
+\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
+the hidden roll of tape vs.
+the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well.
+We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
+% Although relative fit was not empirically evaluated prior to our study,
+Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
+%Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g.
+the roll of tape) was \emph{also} in common ground \cite{BrownSchmidtHanna11_IncrementalPerspectiveTaking}---perhaps people would refer to the critical item less in the privileged condition than in the full common ground condition, which would put a more positive spin on the results.
+%Note that none of these criticisms invalidate the paradoxical result that listeners move an occluded object, they just suggest an alternative explanation.
+Reference disambiguation is probabilistic: common ground and goodness of fit are weighed against one another.
+
 %\subsection{Discussion}
@@
 
+While We expect our conversation partners to follow common conventions about the meanings of words \cite{Lewis69_Convention, CentolaBaronchelli15_ConventionEmergence}, to update these conventions in light of previous interactions \cite{BrennanClark96_ConceptualPactsConversation}, and perhaps most importantly,
+
 It is worth noting several significant differences between our study and \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
@@
 Regardless of the experimental context, it is illuminating to see how real participants naturally interact.
+
+\todo[inline]{mention rubio fernandez early?}
+Using similar reasoning, \cite{RubioFernandez16_DirectorTaskAttention} showed that listeners became suspicious of what was actually in common ground after hearing overinformative utterances, thus successfully using theory of mind reasoning.
+
+\todo[inline]{Mention Westra \& Carruthers as another critique focusing on pragmatics?}
+
+\todo[inline]{Mention neural coordination papers? e.g. Stephens et al 2010, Stolk et al, 2014?}
 

### 2017-09-19 writing/journal_manuscript/pnas-submission.tex
Minor edits to intro

@@
 We argue in this paper that the mental models supporting theory of mind use in communication are much richer than past efforts have explored. 
-We consider a class of mental models representing not only a partner's visual access but also Gricean expectations of rationality and cooperativity \cite{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
+We consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \cite{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
 For example, suppose a friend wants to refer to a nearby Dalmatian. 
@@
 
-Theory of mind mechanisms are therefore at the very heart of the Rational Speech Act framework, yet they have never been directly compared against egocentric or ``mindblind'' approaches under the same behavioral paradigm.
-By formalizing and rigorously comparing the predicted behavior of speakers and listeners using mental models of one another, we show that theory of mind is required to account for context-sensitive behavior.
-In addition to these novel predictions, this approach also allows us to understand the sophisticated pragmatics underlying behavior that, under more naive behavioral measures, apparently reflect failures of theory of mind.
+%Since positive claims about theory of mind in communication concern the mental model we hold of our partner, these claims are best evaluated in a computational framework that formalizes the content of our representations and how we use them.
+We test a richer, more realistic theory of mind accounting not just for visual access but also for Gricean expectations of parsimony and context-appropriate informativity under uncertainty.
+Under this richer model, we find that behavior taken as evidence for limits on theory of mind reasoning is in fact precisely what is required by the pragmatics of the situation.
+
+Theory of mind mechanisms are therefore at the heart of the Rational Speech Act framework, yet they have never been directly compared against the predictions of egocentric or ``mindblind'' theories under the same behavioral paradigm.
+Since positive claims about theory of mind in communication concern the mental model we hold of our partner, these claims are best evaluated in a computational framework that formalizes the content of our representations and how we use them.
+By rigorously comparing the predicted behavior of speakers and listeners using different mental models, we show that theory of mind is required to account for context-sensitive behavior.
+In addition, we replicate effects that, under more naive behavioral measures, were interpreted as evidence of limits on theory of mind use, but under a richer computational model are shown to be rational consequences of an uncooperative confederate.
 %Communication abounds with potential social expectations, and evaluating the extent to which theory of mind is deployed in communication requires considering these expectations \emph{en masse}.

### 2017-09-27 modeling/refModule/json/makeNominalMeanings.js
Make json of lexicon

@@ -1,0 +1,67 @@
+var _ = require('lodash');
+var papaparse = require('papaparse');
+var fs = require('fs');
+
+var labelToObjName = function(label) {
+  var lowerCased = label.toLowerCase();
+  var noWhiteSpace = lowerCased.replace(/[^A-Z0-9]+/ig, "");
+  return noWhiteSpace == "mms" ? "mnms" : noWhiteSpace;
+};
+
+var typicality = function () {
+  var filename = ("../../data/modelTypicalities.csv");
+  var parseResult = papaparse.parse(fs.readFileSync(filename, 'utf8'),
+				    {header: true, skipEmptyLines : true});
+  this.data = parseResult.data;
+  //console.log("this.data:", this.data);
+};
+
+// typicality.prototype.clean = function() {
+//   this.data = _.map(this.data, function(row) {
+//     var newRow = _.clone(row);
+//     var wordsInLabel = row.label.split(' ');
+//     newRow.label = (_.includes(['a', 'an'], wordsInLabel[0]) ?
+// 		    wordsInLabel.slice(1).join(' ') :
+// 		    wordsInLabel.join(' '));
+//     return newRow;
+//   });
+// };
+
+typicality.prototype.getLabels = function() {
+  return _.uniq(_.map(this.data, 'label'));
+};
+
+typicality.prototype.getTypicality = function(label, obj) {
+  console.log(label)
+  console.log(obj)
+  var subset = _.filter(this.data, function(row) {
+    return row.object === obj & row.label === label;
+  });
+  console.log(subset)
+  return Number(subset[0]["m"]);
+};
+
+typicality.prototype.getPossibleReferents = function(label) {
+  return _.map(_.filter(this.data, function(row) {
+    return row.label === label;
+  }), 'object');
+};
+
+typicality.prototype.makeTree = function() {
+  this.tree = {};
+  var that = this;
+  _.each(this.labels, function(label) {
+    var cleanedLabel = labelToObjName(label);
+    that.tree[cleanedLabel] = {};
+    _.each(that.getPossibleReferents(label), function(object) {
+      that.tree[cleanedLabel][object] = that.getTypicality(label, object);
+    });
+  });
+};
+
+var t = new typicality();
+//t.clean();
+t.labels = t.getLabels();
+t.makeTree();
+
+fs.writeFileSync('keysar-meanings.json', JSON.stringify(t.tree, null, 2));

### 2017-09-27 modeling/refModule/ref.js
Add simplest model

@@ -1,0 +1,172 @@
+var _ = require('lodash');
+var fs = require('fs');
+var babyparse = require('babyparse');
+
+var powerset = function (set) {
+  if (set.length == 0)
+    return [[]];
+  else {
+    var rest = powerset(set.slice(1));
+    return rest.map(
+      function(element) {
+        return [set[0]].concat(element);
+      }).concat(rest);
+  }
+};
+
+// get all words that can be used to describe target
+var possibleUtts = function(target, lexicon) {
+  return _.keys(_.pickBy(lexicon, function(value, key) {
+    return _.has(value, target);
+  }));
+};
+
+var colors = ["yellow", "orange", "red", "pink", "green",
+              "purple", "white", "blue", "brown", "black"];
+
+var sizes = ["big", "small"];
+
+var types = ["fan", "tv", "desk", "couch", "desk", "chair", "couch"];
+
+var makeArr = function(n, v) {
+  return _.repeat(n, v);
+};
+
+var makeColorSizeLists = function(wordsOrObjects) {
+  var colorList = wordsOrObjects === 'words' ? colors.concat('') : colors;
+  var sizeList = wordsOrObjects === 'words' ? sizes.concat('') : sizes;
+  var typeList = wordsOrObjects === 'words' ? types.concat('thing') : types;
+
+  return _.flattenDepth(_.map(sizeList, function(size) {
+    return _.map(colorList, function(color) {
+      return _.map(typeList, function(type) {
+        return [size, color, type]
+      });
+    });
+  }), 2);
+};
+
+var colorSizeWordMeanings = function(params) {
+  return _.extend(
+    _.zipObject(colors, _.times(colors.length, _.constant(params.colorTyp))),
+    _.zipObject(sizes, _.times(sizes.length, _.constant(params.sizeTyp))),
+    _.zipObject(types, _.times(types.length, _.constant(1))),
+    {'thing' : 1}
+  );
+};
+
+var constructLexicon = function() {
+  return require('./json/keysar-meanings.json');
+}
+
+function readCSV(filename){
+  return babyparse.parse(fs.readFileSync(filename, 'utf8'),
+			 {header:true}).data;
+};
+
+function writeCSV(jsonCSV, filename){
+  fs.writeFileSync(filename, babyparse.unparse(jsonCSV) + '\n');
+}
+
+function appendCSV(jsonCSV, filename){
[…]

### 2017-09-27 analysis/cogSci_analysis.Rmd
Add script for reformatting speaker data to notebook

@@
 
-criticalMessagesExp1 <- read.csv("./experiment1/data/processedLabelsExp1.csv") %>%
+criticalMessagesExp1 <- read.csv("../experiment1/data/processedLabelsExp1.csv") %>%
   left_join(turkerRatings, by = relevantProperties) %>%
@@
 library(binom)
-
+library(ggthemes)
 Exp1Errs = itemInequalitiesExp1 %>%
@@
   + geom_smooth(method = "lm",colour="black")
-  + theme_bw(base_size = 16)
+  + theme_few(base_size = 16)
   + theme(aspect.ratio=1)
@@
 ```
+
+# Create input for BDA
+
+Create the informativities csv
+
+```{r}
+rbind(read_csv('../data/exp1SpeakerLabels.csv') %>% mutate(exp = 'scripted'),
+      read_csv('../data/exp2SpeakerLabels.csv') %>% mutate(exp = 'unscripted')) %>%
+  mutate(label = str_replace_all(label, fixed(' '), '')) %>%
+  mutate(label = str_replace_all(label, "[[:punct:]]", "")) %>% 
+  mutate(label = str_to_lower(label, locale = "en")) %>%
+  write_csv(path = '../modeling/bdaInput/speakerData.csv')
+```

### 2017-11-06 writing/journal_manuscript/pnas-submission.tex
Minor revisions to intro; begin writing up models

@@
 Armed with a computational theory of mind, people can make inferences about an agent's likely mental state given their past behavior and run simulations to form sophisticated expectations about their future behavior.
-This theoretical approach to social cognition, where theory of mind reasoning \cite{PremackWoodruff78_ChimpanzeeToM, WellmanCrossWatson01_ToMMetaAnalysis, SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM} is grounded in an explicit computational model of how other minds work, has been useful for understanding a number of experimental phenomena in both adults and children \cite{KileyHamlinEtAl13_MentalisticCoreSocial, PantelisEtAl14_IntentionInference, BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}.
+This theoretical approach to social cognition, where theory of mind reasoning \cite{PremackWoodruff78_ChimpanzeeToM, GopnikWellman92_TheoryTheory, WellmanCrossWatson01_ToMMetaAnalysis, SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM} is grounded in an explicit computational model of how other minds work, has been useful for understanding a number of experimental phenomena in both adults and children \cite{KileyHamlinEtAl13_MentalisticCoreSocial, PantelisEtAl14_IntentionInference, BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}.
 Yet our current understanding of theory of mind use in \emph{communication} remains limited. 
@@
 We argue in this paper that the mental models supporting theory of mind use in communication are much richer than past efforts have explored. 
-We consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \cite{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
+In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \cite{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
 For example, suppose a friend wants to refer to a nearby Dalmatian. 
@@
 If there are no other dogs nearby, we may expect them to use a parsimonious basic-level term like ``dog.''
-When there is potential for confusion, however, we expect them to introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \cite{BrennanClark96_ConceptualPactsConversation, GrafEtAl16_BasicLevel}.
+When there is potential for confusion, however, we expect them to consider our likely interpretation and introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \cite{BrennanClark96_ConceptualPactsConversation, VanDeemter16_ComputationalModelsOfReferring, GrafEtAl16_BasicLevel}.
 When speakers violate these expectations, listeners draw inferences about the intended meaning in order to account for the violation: for example, XXX \cite{XXX, BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}.
 
-This intricate dance of expectations has recently been formalized in a probabilistic model of communication as recursive social reasoning.
-The resulting Rational Speech Act (RSA) framework has successfully described many puzzling pragmatic phenomena, including vagueness \cite{Lassiter}, hyperbole \cite{Kao}, irony \cite{}, metaphor, and overinformativity \cite{Judith}.
+This intricate dance of expectations has been formalized in a probabilistic model of communication as recursive social reasoning.
+The Rational Speech Act (RSA) framework has successfully described many puzzling communicative phenomena, including vagueness \cite{Lassiter}, hyperbole \cite{Kao}, irony \cite{}, metaphor, and apparent over-informativeness \cite{Judith}.
 In the simplest version of this framework, speakers judge the utility of different utterances proportionate to the probability of correct interpretation under an internal model of a listener.
@@
 In addition, we replicate effects that, under more naive behavioral measures, were interpreted as evidence of limits on theory of mind use, but under a richer computational model are shown to be rational consequences of an uncooperative confederate.
-%Communication abounds with potential social expectations, and evaluating the extent to which theory of mind is deployed in communication requires considering these expectations \emph{en masse}.
-While egocentric views of perspective-taking make more clear-cut behavioral predictions, the predictions derived from a computational theory of mind are more varied and complicated but ultimately more accurate.
+Especially in asymmetric communicative environments where speakers are explicitly put in a state of uncertainty over which objects their partner is seeing, evaluating the extent to which theory of mind is deployed in communication requires flexible mental models that represent the many sources of social expectations.
+% While egocentric views of perspective-taking make more clear-cut behavioral predictions, the predictions derived from a computational theory of mind are more varied and complicated but ultimately more accurate.
 
@@
 
-\section*{Referring in an uncertain visual environment}
+\section*{Models of referring}
 
-In this section, I formalize the director-matcher task inside a rational speech act (RSA) model, where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language.
+\subsection*{Pure egocentrism}
+
+The simplest strategy for a speaker attempting to refer to a target object $o_i$ is to produce the simplest label that best fits $o_i$, effectively ignoring context and listener knowledge.
+\todo[inline]{rxdh: technically, mean soft-max here...} Similarly, the simplest strategy for a listener is to pick the object in their view that is best described by the utterance, ignoring the visual access or pragmatic goals of the speaker.
+These models are purely egocentric, in the sense that they do not engage in any social reasoning.
+
+\subsection*{Pure consideration of visual access}
+
+The most immediate model of social reasoning is consideration of the informativeness of a label in a shared context containing many objects.
+For a speaker, this means preferring labels that are good fits for the target but \emph{not} known distractors in common view.
+For instance, when referring to a dog in the shared context of other dogs, a speaker may opt to be more informative than simply saying `dog,' since it is ambiguous.
+For a listener, this means discounting objects that are not in shared context, regardless of how well the description fits.
+These are not available to the visual access of the speaker and thus not viable referents.
+
+\subsection*{Probabilistic weighing of perspectives}
+
+Rather than taking a purely egocentric perspective or purely considering a partner's visual access, speakers or listeners may be probabilistically blending these two perspectives.
+
+\subsection*{Gricean communication}
+
+\subsection*{Referring in a shared visual environment}
+
+We begin by considering the problem faced by a speaker who wants their partner to select a target object from a context containing other potentially similar objects.
+%  a simple probabilistic model of reference in the rational speech act (RSA) model.
+
+\subsection*{Referring under uncertainty}
+
+where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language.
 This allows us to test pragmatic accounts of task performance.
@@
 
-\subsection*{Participants}
-
-We recruited 64 participants (32 pairs) from Amazon Mechanical Turk, roughly doubling the sample size from Expt.~1.
-All participants were from the U.S.
-Three participants were excluded for making 2 or more errors on non-critical items, and one additional participant was excluded because they were not a native English speaker.
-
-\subsection*{Materials \& Procedures}
-
-Everything was the same as Expt.~1, except we did not use scripted messages for critical instructions.
+%The argument offered by Keysar and colleagues is based on an elegant experimental paradigm, where participants played a simple communication game with a confederate.
+The two players were placed on opposite sides of a $4 \times 4$ grid containing a set of everyday objects (see Fig.
+\ref{fig:interface}).
+The confederate played the role of `director,' giving instructions about how to move objects around a grid, and the participant played the role of `matcher,' attempting to follow these instructions.
+For example, the objects in one trial included a cassette tape.
+The director gave an instruction like `move the tape up one square,' referring to the cassette.
+%Critically, some objects were occluded such that only the matcher could see them, creating an asymmetry in the players' knowledge.
+To perform accurately on critical trials, the matcher would need to apply theory of mind to reason about which objects were shared and which were private.
+%For example, imagine a roll of tape were placed in an occluded slot: if a participant failed to account for the director's (partial) knowledge, she might interpret `tape' to mean the occluded roll of tape (which the director couldn't possibly know about).
+Indeed, \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} found that participants attempted to move the hidden item in 30\% of cases: 71\% of participants attempted to move this hidden item at least once (out of four critical cases) in the experimental condition, compared to 0\% in a control condition where there was no ambiguity over the referent.
+%Additionally, eye-tracking data showed that participants considered the hidden item more often and for longer in the experimental condition than the control condition.
 

### 2017-11-06 modeling/modelAnalysis.Rmd
Add simple outline of model analysis nb

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
library(tidyverse)
```

```{r}
```

```{r}
sumlogprob <- function(a, b) {
  if(a>b) {
    return(a+log1p(exp(b-a)))
  } else{
    return(b+log1p(exp(a-b)))
  }
}

params <- read_delim('./bdaOutput/testingParams.csv', delim = ',')

View(params)

modelOutput = params %>%
  filter(prob != -Inf) %>%
  group_by(model) %>%
  summarize(prob = reduce(logLikelihood, sumlogprob) - log(length(logLikelihood)))

print(modelOutput)
```

```{r}
params %>%
  group_by(model) %>%
  #arrange(logLikelihood) %>%
  filter(logLikelihood == max(logLikelihood))

?arrange
```

### 2017-11-13 analysis/cogSci_analysis.Rmd
Plot cassette informativity rating distributions

@@
 ```
+
+```{r}
+library(ggthemes)
+View(turkerRatings)
+
+d <- turkerRatings %>%
+  filter(objectSet == 3) %>%
+  group_by(label,referent) %>% 
+  mutate(avg = mean(response),
+            med = median(response))
+
+ggplot(d, aes(x = response)) +
+    geom_histogram(binwidth = .1) +
+    geom_vline(aes(xintercept=med)) +
+    facet_grid(label ~ referent) +
+    theme_few() +
+    theme(aspect.ratio = .25, text = element_text(size=8))
+
+ggsave(filename = '~/Downloads/cassette_informativity_ratings.png', height = 5)
+```

### 2017-11-13 modeling/refModule/ref.js
Revise model interface to look at item-level likelihoods

@@ -2,6 +2,25 @@ var _ = require('lodash');
 var fs = require('fs');
 var babyparse = require('babyparse');
 
+function _logsumexp(a) {
+  var m = Math.max.apply(null, a);
+  var sum = 0;
+  for (var i = 0; i < a.length; ++i) {
+    sum += (a[i] === -Infinity ? 0 : Math.exp(a[i] - m));
+  }
+  return m + Math.log(sum);
+}
+
+var getL0score = function(targetObj, utt, context, params) {
+  var similarities = params.lexicon[utt];
+  var a = [];
+  for(var i=0; i<context.length; i++){
+    a.append(params.simScale * similarities[context[i]]);
+  }
+  return (params.simScale * similarities[targetObj]
+	  - _logsumexp(a));
+};
+
 var powerset = function (set) {
   if (set.length == 0)
     return [[]];
@@ -93,7 +112,7 @@ var writeERP = function(erp, labels, filename, fixed) {
 };
 
 var supportWriter = function(s, p, handle) {
-  var sLst = _.pairs(s);
+  var sLst = _.toPairs(s);
   var l = sLst.length;
 
   for (var i = 0; i < l; i++) {
@@ -108,7 +127,7 @@ var bayesianErpWriter = function(erp, filePrefix) {
 				"value", "prob", "MCMCprob"] + '\n');
 
   var paramFile = fs.openSync(filePrefix + "Params.csv", 'w');
-  fs.writeSync(paramFile, ["parameter", "value", "MCMCprob"] + '\n');
+  fs.writeSync(paramFile, ["model", "alpha", "costWeight", "typWeight", "perspectiveWeight","omniscience", "logLikelihood", "prob"] + '\n');
 
   var supp = erp.support();
   supp.forEach(function(s) {

### 2017-11-13 modeling/modelAnalysis.Rmd
Fix numeric issue when there are -Infs

@@
 
-params <- read_delim('./bdaOutput/testingParams.csv', delim = ',')
+params <- read_csv('./bdaOutput/testingParams.csv', col_types = c('cnnnnnnd'))#, delim = ',')
 
@@
 modelOutput = params %>%
-  filter(prob != -Inf) %>%
+  #filter(logLikelihood != -Inf) %>%
   group_by(model) %>%
-  summarize(prob = reduce(logLikelihood, sumlogprob) - log(length(logLikelihood)))
+  summarize(logLikelihood = reduce(logLikelihood, sumlogprob) - log(length(logLikelihood)))
 
@@
 
-?arrange
 ```

### 2017-11-13 writing/journal_manuscript/pnas-submission.tex
Minor edits to introduction

@@
 
-% Characterize the computational approach to theory of mind we plan to take
-% Point out a gaping hole that could benefit from this approach
+%%% Characterize the computational approach to theory of mind we plan to take
+%%% Point out a gaping hole that could benefit from this approach
 Our success as a cooperative species depends upon our ability to make sense of other agents. 
-A natural solution to this problem is to maintain an internal predictive model of how minds work: how an agent's unobservable beliefs and desires give rise to observable actions. 
-Armed with a computational theory of mind, people can make inferences about an agent's likely mental state given their past behavior and run simulations to form sophisticated expectations about their future behavior.
-This theoretical approach to social cognition, where theory of mind reasoning \cite{PremackWoodruff78_ChimpanzeeToM, GopnikWellman92_TheoryTheory, WellmanCrossWatson01_ToMMetaAnalysis, SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM} is grounded in an explicit computational model of how other minds work, has been useful for understanding a number of experimental phenomena in both adults and children \cite{KileyHamlinEtAl13_MentalisticCoreSocial, PantelisEtAl14_IntentionInference, BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}.
+One solution is to maintain an internal predictive model of how minds work: how an agent's unobservable beliefs and desires give rise to observable actions. 
+Armed with a computational theory of mind, we can make inferences about an agent's likely mental state given their past behavior and form sophisticated expectations about their future behavior. 
+%This theoretical approach to social cognition, where theory of mind reasoning \cite{PremackWoodruff78_ChimpanzeeToM, GopnikWellman92_TheoryTheory, WellmanCrossWatson01_ToMMetaAnalysis, SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM} is grounded in an explicit computational model of how other minds work, has
+% The approach has been useful for understanding a number of experimental phenomena in both adults and children \cite{KileyHamlinEtAl13_MentalisticCoreSocial, PantelisEtAl14_IntentionInference, BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}.
 Yet our current understanding of theory of mind use in \emph{communication} remains limited. 
-Theory of mind as applied within communication depends on an accurate model of what a speaker would say, or what a listener would think, in different situations. 
-Past efforts have instead largely focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
+Just as making sense of an agent's physical movements in the world requires an accurate model of how beliefs and intentions translate into motor plans, making sense of language depends on an accurate model of what a speaker would say, or what a listener would think, in different situations.
 
 %While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
-We argue in this paper that the mental models supporting theory of mind use in communication are much richer than past efforts have explored. 
+Past efforts have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
+We argue in this paper that the mental models supporting theory of mind use in communication are much richer than previously explored. 
 In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \cite{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
@@
 These are not available to the visual access of the speaker and thus not viable referents.
+Note that
 

### 2018-02-23 writing/journal_manuscript/manuscript.tex
convert to open mind; revise a bit

--- 2017-11-13 writing/journal_manuscript/pnas-submission.tex
@@
-\documentclass[9pt,twocolumn,twoside]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
-% Note that the use of elements such as single-column equations
-% may affect the guide line number alignment.
+%% Documentclass:
+\documentclass[manuscript]{stjour}
 
-\templatetype{pnasresearcharticle} % Choose template 
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
+%% Or,
 
-\usepackage{pslatex}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{color}
+%% Manuscript, for double spaced, larger fonts
+% \documentclass[manuscript]{stjour}
+%% Only needed if you use `manuscript' option
+\journalname{Open Mind}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% For production only, not authors:
+%%\documentclass[OpenMind,finalfonts]{stjour}
+
+%%%%%%%%%%% Please supply information %%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% For Open Mind:
+%% Supplementary Materials:
+\supplementslinks{dx.doi.org/10.1098/rsif.2013.0969}
+
+%% If no conflicts, this command doesn't need to be used
+%% \conflictsofinterest{}
+
+%%%%%%%%%%% to be supplied by MIT Press, only %%%%%%%%%%%%%%%%%
+
+\citation{Niyogi, R. K., Breton, Y.-A., Solomon,
+R. B., Conover, K.,\\ Shizgal, P., Dayan, P. (2015).\\ 
+Optimal indolence: a normative microscopic approach to work and leisure. Open Mind 1(1):
+1−12.}
+
+\received{XXX}
+\accepted{XXX}
+\published{XXX}
+
+%% DOI address:
+\setdoi{10.1098/rsif.2013.0969}
+
+%%%%%%%% End MIT Press commands %%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% author definitions should be placed here:
+
+%% example definition
+\def\taupav{\tau_{\mathrm{Pav}}}
 \usepackage{todonotes}
-\usepackage{dsfont}
-\usepackage{array}
-\usepackage{textcomp}
-\usepackage{multirow}
+\begin{document}
 
-\title{Conversational expectations account for apparent limits on theory of mind use}
+\title{The pragmatics of perspective-taking}
+\subtitle{Gricean expectations account for apparent limits on theory of mind use}
 
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1]{Robert Hawkins}
-\author[a]{Hyowon Gweon} 
-\author[a]{Noah Goodman}
+%% If shortened title for running head is needed so that the article title can fit
+%%   in the running head, use [] argument, ie,
+%%
+%%   \title[Shortened Title for Running Head]{Title of Article}
+%%   \subtitle{Subtitle Here}
 
-\affil[a]{Department of Psychology, Stanford University}
+%% Since we use \affil{} in the argument of the \author command, we
+%% need to supply another version of the author names, without \affil{}
+%% to be used for running heads:
 
-% Please give the surname of the lead author for the running footer
-\leadauthor{Hawkins}
+\author[Hawkins, Gweon, Goodman]
+{Robert Hawkins\affil{1}, Hyowon Gweon\affil{1}, Noah Goodman\thanks{asdf}\affil{1}}
 
-% TODO: Currently a bit too long (and probably too in the weeds...) 
-\significancestatement{To communicate successfully, speakers must flexibly adapt %the specificity of 
-their utterances to the demands of the current context.
-When they are unsure exactly which objects are in view of their partner, for example, they may provide additional disambiguating information when referring to a particular object: saying "cassette tape" instead of the potentially ambiguous ``tape." %A listener who rationally expects such levels of specificity may, in turn, draw surprising inferences when they hear a less specific utterance than the situation demands.
-We argue that influential arguments for limits on Theory of Mind (ToM) in adults have neglected the nuanced pragmatics of the communication tasks they rest upon.
-A careful analysis of pragmatic demands using a rational model of communication suggests a minimal pair of experiments which show that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.}
+\affiliation{1}{Department of Psychology, Stanford University, Stanford, CA, US}
 
-% Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any conflict of interest here.}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}
+%ie.
+%\affiliation{1}{Gatsby Computational Neuroscience Unit, University
+%College London, London, United Kingdom}
 
-% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Theory of mind $|$ Pragmatics $|$ Communication $|$ Social cognition $|$ Replication}
+%\affiliation{2}{Another Department, Institution, City, Country}
+
+%ie
+%\affiliation{2}{Center for Studies in
+%Behavioral Neurobiology, Concordia University, Montreal, Quebec,
+%Canada}
+
+\correspondingauthor{Robert Hawkins}{rxdh@stanford.edu}
+
+% ie,
+%\correspondingauthor{Ritwik K. Niyogi}{ritwik.niyogi@gatsby.ucl.ac.uk}
+
+\keywords{theory of mind, pragmatics, interaction, communication, social cognition, replication}
+
+%ie
+%\keywords{work, leisure, normative, microscopic,  reinforcement learning, economics}
 
 \begin{abstract}
-%Humans are capable of accurately reasoning about other minds, but how ubiquitously do we use this ability in everyday life? 
-Recent theories of language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication.
+%To communicate successfully, speakers must flexibly adapt %the specificity of 
+%their utterances to the demands of the current context.
+When they are unsure exactly which objects are in view of their partner, for example, they may provide additional disambiguating information when referring to a particular object: saying "cassette tape" instead of the potentially ambiguous ``tape."
+Recent debates over theory of mind use have hinged on behavior in natural-language communication tasks but have largely neglected the nuanced pragmatics of these tasks.
+Under Gricean expectations of cooperativity, speakers ought to be more informative in contexts with high uncertainty about what is in the listener's field of view.
+If pragmatic reasoning is truly grounded in theory of mind,   A careful analysis of pragmatic demands using a rational model of communication suggests a minimal pair of experiments which show that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.
+%Recent theories of language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication.
 To reconcile these points of view, we use a computational model of communication under uncertainty to analyze the pragmatic demands of a common theory of mind test.
@@
 
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\begin{document}
-
-% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
-% You should only change this length when you've finalised the article contents.
-\verticaladjustment{-2pt}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
+\section{Introduction}\label{introduction}
 
@@
 
-\section*{Models of referring}
+In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
 
-\subsection*{Pure egocentrism}
+\subsection{Experiment~1: Direct replication}
+\label{sec:Exp1}
 
-The simplest strategy for a speaker attempting to refer to a target object $o_i$ is to produce the simplest label that best fits $o_i$, effectively ignoring context and listener knowledge.
-\todo[inline]{rxdh: technically, mean soft-max here...} Similarly, the simplest strategy for a listener is to pick the object in their view that is best described by the utterance, ignoring the visual access or pragmatic goals of the speaker.
-These models are purely egocentric, in the sense that they do not engage in any social reasoning.
+\subsection{Participants}
 
-\subsection*{Pure consideration of visual access}
+We recruited 34 participants (17 pairs) from Amazon Mechanical Turk. All participants were from the U.S. Three pairs were excluded for making 2 or more errors on non-critical items.
 
-The most immediate model of social reasoning is consideration of the informativeness of a label in a shared context containing many objects.
-For a speaker, this means preferring labels that are good fits for the target but \emph{not} known distractors in common view.
-For instance, when referring to a dog in the shared context of other dogs, a speaker may opt to be more informative than simply saying `dog,' since it is ambiguous.
-For a listener, this means discounting objects that are not in shared context, regardless of how well the description fits.
-These are not available to the visual access of the speaker and thus not viable referents.
-Note that
+\subsection{Materials \& Procedures}
 
-\subsection*{Probabilistic weighing of perspectives}
+Participants interacted in a real-time, multi-player environment on the web \cite{Hawkins15_RealTimeWebExperiments}. 
+%To deal with participants arriving at unpredictable times, we implemented a waiting room mechanism, where the first person to enter the game environment would see a screen saying ``Waiting for another player..." and the second person would trigger the game to start.
+Pairs of participants---assigned randomly to `director' and `matcher' roles---interacted with one another through a web interface, shown in Fig.
+\ref{fig:interface}.
+On the left side of the screen, participants could freely type messages to one another; on the right side the screen, players could view a set of objects placed in a 4 x 4 grid.
+Five of the grid cells were occluded from only the director's perspective, and the remaining 11 were visible to both matcher and director.
+Six or seven objects were displayed in the grid at a given time.
+One of these objects was a `target', such as a cassette tape, placed in an unoccluded cell such that both participants could see it.
+Another object, such as a roll of tape, was placed in an occluded slot such that it was only visible to the matcher.
+The rest of the objects were unrelated `fillers' placed in random locations.
+We used the same set of targets and occluded alternatives as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, but we were unable to obtain the filler objects from the original experiment and created our own.
+Before entering the game environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface.
+Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
 
-Rather than taking a purely egocentric perspective or purely considering a partner's visual access, speakers or listeners may be probabilistically blending these two perspectives.
+The experiment was composed of eight items, with each item using a different set of objects.
+Each item included one `critical pair' of objects, one of them the target and the other hidden, such as the cassette tape and the roll of tape.
+For each item, we gave the director a series of four instructions to move objects around, which were displayed as a series of arrows pointing from some object to an unoccupied cell.
+To collect clean mouse-tracking data, we began every instruction by asking the matcher to click a small circle in the center of the grid.
+After this small circle was clicked, the director was allowed to communicate the next instruction and we started recording from the matcher's mouse.
+One of the instructions was a `critical instruction,' which referred to the target object.
+For half the instructions, directors were free to communicate however they wished.
+For the other half, including all the critical instructions, their messages to the matcher were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat window, and they would have to click `send' for the experiment to continue.)
 
-\subsection*{Gricean communication}
+We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+Each participant received half the items in the experimental condition and half in the baseline condition.
+The assignment of items to conditions was randomized across participants, and the order of conditions was randomized under the constraint that the same condition would not be used on more than two consecutive items.
+All object sets, object placements, and corresponding instruction sets were the same for all participants.
 
-\subsection*{Referring in a shared visual environment}
-
-We begin by considering the problem faced by a speaker who wants their partner to select a target object from a context containing other potentially similar objects.
-%  a simple probabilistic model of reference in the rational speech act (RSA) model.
-
-\subsection*{Referring under uncertainty}
-
-where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language.
-This allows us to test pragmatic accounts of task performance.
-
-Previous theoretical advances in modeling pragmatics have typically hinged on `lifting' inference over some aspect of communication that was previously assumed to be fixed.
-For instance, \emph{manner implicatures} emerge from uncertainty over the true meanings of words in the lexicon \cite{BergenLevyGoodman16_LexicalUncertainty}, nonliteral constructions like \emph{hyperbole} and \emph{irony} emerge from uncertainty over the topic of discussion \cite{KaoWuBergenGoodman14_NonliteralNumberWords}, and \emph{prosody} and \emph{ellipsis} emerge from uncertainty (in the form of a noisy-channel model) over the various ways that a speaker's utterance could have been corrupted  \cite{BergenGoodman15_StrategicUseOfNoise}.
-Note that these are not ad hoc extensions; sources of uncertainty that are not strictly necessary to explain a particular effect are often omitted for simplicity, but they all coexist within the same model.
-
-Pragmatic critiques of the director-matcher task point to an additional source of uncertainty that has not previously been explored: uncertainty over a communication partner's visual context.
-Considering this uncertainty, of course, \emph{is} theory of mind in action; our goal is to show how such a model built on social reasoning can produce the `failures' found by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} as well as the phenomena observed by both \cite{HawkinsGoodman16_Keysar} and \cite{RubioFernandez16_DirectorTaskAttention}.
-For simplicity, we consider the full $2\times 2$ grid containing a single occluded cell, which was used by \cite{RubioFernandez16_DirectorTaskAttention}.
-We incorporate this source of uncertainty into both the speaker and listener models as follows:
-
-\begin{enumerate}
-\item A pragmatic speaker perceives an array of three objects $\mathcal{O}_S = \{o_i\}$, one of which is identified as a target $t$, but assumes that their partner sees an additional hidden object $o_h$ that is not in common ground.
-The speaker has uncertainty over the identity of this object, represented by a prior distribution $P(o_h)$ which they marginalize over when deciding what utterance to produce:
-$$S_1(u | t, \mathcal{O}_p) \propto \sum_{o_h} L_0(t| \mathcal{O}_S \cup \{o_h\})P(o_h)P(u)$$
-
-\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$.
-The listener, however, also has some uncertainty over whether the experimenter is deceptive, represented by a Bernoulli random variable $d$, such that they assign a small prior probability $P(d)$ to the possibility that $o_h$ is actually in common ground.
-Upon hearing an utterance$u$, they update both their beliefs about which object is the target and whether the experimenter is being deceptive:
-$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
-\end{enumerate}
-Technical details are omitted here for brevity, but the full model is fully specified in WebPPL and available to run at \url{http://forestdb.org/models/keysar.html}.
-
-We focus on two critical patterns of results produced by the pragmatic listener $L_2$.
-First, we consider how it responds to the scripted instruction ``fish'' when presented with a context containing two fish (a red one that is occluded and a blue one in common ground) as well as two other unrelated objects.
-While it is most likely to choose the red fish in common ground, it has a relatively high probability (43\%) of choosing the hidden red fish, an effect \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} claimed was due to failure of theory of mind.
-This decision emerges in the model for two primary reasons.
-For one, a speaker with uncertainty over context could in principle be communicating about any object, and if they believed the hidden object was a red fish, the utterance `fish' would have a high probability of being produced.
-To a lesser extent, context uncertainty makes the overinformative utterance `blue fish' more likely than `fish', hence the omission of the color modifier has the (subtle) implication of \emph{not} the blue fish.
-This speaker phenomenon, while subtle under the parameter regime of the current implementation, produces the dynamic observed in the unscripted replication by  \cite{HawkinsGoodman16_Keysar}.
-
-Second, we consider the listener's posterior over the \emph{deception} Bernoulli variable $d$.
-Just as \cite{RubioFernandez16_DirectorTaskAttention} found that listeners were more likely to suspect deception after hearing an `overinformative' utterance like `blue fish' when only one fish was in common ground, so too does our pragmatic listener.
-The pragmatic speaker is relatively more likely to produce the additional color modifier if they see multiple fish than when they see a single fish and simply have uncertainty over the unknown object.
-Hence, our model simultaneously captures all three phenomena.
-
-\section*{Expt.~1: Unscripted Replication}
-\label{sec:Exp1}
+This paradigm differs from those used by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways.
+First, participants were not seated across from each other at a table: they each saw a view of the 4 x 4 grid on their screen and communicated via a text box.
+Second, we did not use a trained confederate.
+We randomly assigned one of the players to the role of the instruction giver, and maintained the original wording by scripting a subset of their instructions.
+Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \cite{KeysarBarr___Brauner00_TakingPerspective}.
 
@@
 
-\subsection*{Results}
+\subsection{Results}
+
+In Table \ref{table:mainResults} we show the error rates on critical items, and compare to the data from
+\cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+We find that 93\% of participants (all but one) attempted to move the hidden distractor at least once in the Experimental condition, out of four possible items, compared to only 7\% (only one) in the baseline condition.
+This is similar to the effect observed by the authors in the original study, which found 71\% and 0\%.
+Our errors were larger across the board, perhaps due to the interface or the population, but the gap between the two conditions is roughly the same size.
+%Next, we examine whether there are differences in error rates across items. 
+In Table \ref{table:ItemWise}, we break down the pattern of errors by item. 
+%Because our randomization procedure undersampled the experimental condition for several items, a chi-squared test is inappropriate.
+Using Fisher's exact test, we cannot reject the null hypothesis of independence ($p = 0.13$).
+Still,
+We note that several items have much higher error rates than others -- for example, 75\% of participants in the experimental condition of item 6 made an error (the ``whiteboard eraser'' vs.
+the ``pencil eraser'') while only 17\% of participants in item 8 made an error (the ``computer mouse'' vs.
+the ``toy mouse'').
+Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.
+We empirically substantiate this observation in our results for Expt.~2 below.
+
+This item-wise variability suggests that the dependent variable highlighted in the original study (i.e.
+``percentage of participants who moved the critical item at least once'') is somewhat problematic: it could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item.
+Indeed, if we exclude the three `hard' items where over 60\% of participants in the experimental condition made errors, this dependent variable drops from 93\% to only 43\% of participants.
+%This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
+
+As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
+We define the decision window as the span of time between the point when the matcher received their instruction message and when they started moving an object.
+If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt.
+Within the decision window, we computed the total amount of time spent hovering over the cell containing the target and divided by the total length of the decision window to get a measure of the relative time spent considering the target.
+We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
+A paired-samples $t$-test found that people tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig.
+\ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} .
+%accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
+
+By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items.
+Informally, we noted a tendency toward additional, possibly unnecessary, precision in descriptions.
+Instead of ``move the stuffed animal down'', participants said ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.''
+Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions.
+If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells).
+That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
+We tested this prediction in Expt.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
+By the reasoning above we expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
+
+% discuss other criticisms of Keysar paradigm?
+While these results are compelling, the paradigm has been criticized from few different angles. 
+\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
+the hidden roll of tape vs.
+the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well.
+We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
+% Although relative fit was not empirically evaluated prior to our study,
+Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
 
@@
 
-\section*{Expt.~2: Scripted Replication}
-\label{sec:Exp2}
-
-A large body of evidence has demonstrated striking failures of perspective-taking, where listeners appear to be ``mind-blind,'' neglecting the visual access of their partner when interpreting their utterances
-
-In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
-
-\subsection*{Participants}
-
-We recruited 34 participants (17 pairs) from Amazon Mechanical Turk. All participants were from the U.S. Three pairs were excluded for making 2 or more errors on non-critical items.
-
-\subsection*{Materials \& Procedures}
-
-Participants interacted in a real-time, multi-player environment on the web \cite{Hawkins15_RealTimeWebExperiments}. 
-%To deal with participants arriving at unpredictable times, we implemented a waiting room mechanism, where the first person to enter the game environment would see a screen saying ``Waiting for another player..." and the second person would trigger the game to start.
-Pairs of participants---assigned randomly to `director' and `matcher' roles---interacted with one another through a web interface, shown in Fig.
-\ref{fig:interface}.
-On the left side of the screen, participants could freely type messages to one another; on the right side the screen, players could view a set of objects placed in a 4 x 4 grid.
-Five of the grid cells were occluded from only the director's perspective, and the remaining 11 were visible to both matcher and director.
-Six or seven objects were displayed in the grid at a given time.
-One of these objects was a `target', such as a cassette tape, placed in an unoccluded cell such that both participants could see it.
-Another object, such as a roll of tape, was placed in an occluded slot such that it was only visible to the matcher.
-The rest of the objects were unrelated `fillers' placed in random locations.
-We used the same set of targets and occluded alternatives as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, but we were unable to obtain the filler objects from the original experiment and created our own.
-Before entering the game environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface.
-Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
-
-The experiment was composed of eight items, with each item using a different set of objects.
-Each item included one `critical pair' of objects, one of them the target and the other hidden, such as the cassette tape and the roll of tape.
-For each item, we gave the director a series of four instructions to move objects around, which were displayed as a series of arrows pointing from some object to an unoccupied cell.
-To collect clean mouse-tracking data, we began every instruction by asking the matcher to click a small circle in the center of the grid.
-After this small circle was clicked, the director was allowed to communicate the next instruction and we started recording from the matcher's mouse.
-One of the instructions was a `critical instruction,' which referred to the target object.
-For half the instructions, directors were free to communicate however they wished.
-For the other half, including all the critical instructions, their messages to the matcher were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat window, and they would have to click `send' for the experiment to continue.)
-
-We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-Each participant received half the items in the experimental condition and half in the baseline condition.
-The assignment of items to conditions was randomized across participants, and the order of conditions was randomized under the constraint that the same condition would not be used on more than two consecutive items.
-All object sets, object placements, and corresponding instruction sets were the same for all participants.
-
-This paradigm differs from those used by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways.
-First, participants were not seated across from each other at a table: they each saw a view of the 4 x 4 grid on their screen and communicated via a text box.
-Second, we did not use a trained confederate.
-We randomly assigned one of the players to the role of the instruction giver, and maintained the original wording by scripting a subset of their instructions.
-Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \cite{KeysarBarr___Brauner00_TakingPerspective}.
+%A large body of evidence has demonstrated striking failures of perspective-taking, where listeners appear to be ``mind-blind,'' neglecting the visual access of their partner when interpreting their utterances
 
@@
 
-\subsection*{Results and discussion}
-In Table \ref{table:mainResults} we show the error rates on critical items, and compare to the data from
-\cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-We find that 93\% of participants (all but one) attempted to move the hidden distractor at least once in the Experimental condition, out of four possible items, compared to only 7\% (only one) in the baseline condition.
-This is similar to the effect observed by the authors in the original study, which found 71\% and 0\%.
-Our errors were larger across the board, perhaps due to the interface or the population, but the gap between the two conditions is roughly the same size.
-%Next, we examine whether there are differences in error rates across items. 
-In Table \ref{table:ItemWise}, we break down the pattern of errors by item. 
-%Because our randomization procedure undersampled the experimental condition for several items, a chi-squared test is inappropriate.
-Using Fisher's exact test, we cannot reject the null hypothesis of independence ($p = 0.13$).
-Still,
-We note that several items have much higher error rates than others -- for example, 75\% of participants in the experimental condition of item 6 made an error (the ``whiteboard eraser'' vs.
-the ``pencil eraser'') while only 17\% of participants in item 8 made an error (the ``computer mouse'' vs.
-the ``toy mouse'').
-Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.
-We empirically substantiate this observation in our results for Expt.~2 below.
-
-This item-wise variability suggests that the dependent variable highlighted in the original study (i.e.
-``percentage of participants who moved the critical item at least once'') is somewhat problematic: it could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item.
-Indeed, if we exclude the three `hard' items where over 60\% of participants in the experimental condition made errors, this dependent variable drops from 93\% to only 43\% of participants.
-%This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
-
-As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
-We define the decision window as the span of time between the point when the matcher received their instruction message and when they started moving an object.
-If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt.
-Within the decision window, we computed the total amount of time spent hovering over the cell containing the target and divided by the total length of the decision window to get a measure of the relative time spent considering the target.
-We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
-A paired-samples $t$-test found that people tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig.
-\ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} .
-%accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
-
-By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items.
-Informally, we noted a tendency toward additional, possibly unnecessary, precision in descriptions.
-Instead of ``move the stuffed animal down'', participants said ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.''
-Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions.
-If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells).
-That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
-We tested this prediction in Expt.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
-By the reasoning above we expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
-
-% discuss other criticisms of Keysar paradigm?
-While these results are compelling, the paradigm has been criticized from few different angles. 
-\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
-the hidden roll of tape vs.
-the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well.
-We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
-% Although relative fit was not empirically evaluated prior to our study,
-Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
 %Third, none of the experiments reported by Keysar and colleagues included a key comparison condition where the critical item (e.g.
@@
 
-\section*{General Discussion}
+\section{Evaluating models of referring}
+
+\subsection{Pure egocentrism}
+
+The simplest strategy for a speaker attempting to refer to a target object $o_i$ is to produce the simplest label that best fits $o_i$, effectively ignoring context and listener knowledge.
+\todo[inline]{rxdh: technically, mean soft-max here...} Similarly, the simplest strategy for a listener is to pick the object in their view that is best described by the utterance, ignoring the visual access or pragmatic goals of the speaker.
+These models are purely egocentric, in the sense that they do not engage in any social reasoning.
+
+\subsection{Pure consideration of visual access}
+
+The most immediate model of social reasoning is consideration of the informativeness of a label in a shared context containing many objects.
+For a speaker, this means preferring labels that are good fits for the target but \emph{not} known distractors in common view.
+For instance, when referring to a dog in the shared context of other dogs, a speaker may opt to be more informative than simply saying `dog,' since it is ambiguous.
+For a listener, this means discounting objects that are not in shared context, regardless of how well the description fits.
+These are not available to the visual access of the speaker and thus not viable referents.
+
+\subsection{Probabilistic weighing of perspectives}
+
+Rather than taking a purely egocentric perspective or purely considering a partner's visual access, speakers or listeners may be probabilistically integrating these two perspectives \cite{HellerParisienStevenson16_ProbabilisticWeighing}.
+% When conversational expectations lead participants to expect over-informative utterances, they (rationally) place relatively less weight on features of the environment determining common ground, such as shared perceptual access.
+
+\subsection{Gricean communication under uncertainty}
+
+We begin by considering the problem faced by a speaker who wants their partner to select a target object from a context containing other potentially similar objects.
+%  a simple probabilistic model of reference in the rational speech act (RSA) model.
+
+where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language.
+This allows us to test pragmatic accounts of task performance.
+
+Pragmatic critiques of the director-matcher task point to an additional source of uncertainty that has not previously been explored: uncertainty over a communication partner's visual context.
+Considering this uncertainty, of course, \emph{is} theory of mind in action; our goal is to show how such a model built on social reasoning can produce the `failures' found by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} as well as the phenomena observed by both \cite{HawkinsGoodman16_Keysar} and \cite{RubioFernandez16_DirectorTaskAttention}.
+For simplicity, we consider the full $2\times 2$ grid containing a single occluded cell, which was used by \cite{RubioFernandez16_DirectorTaskAttention}.
+We incorporate this source of uncertainty into both the speaker and listener models as follows:
+
+\begin{enumerate}
+\item A pragmatic speaker perceives an array of three objects $\mathcal{O}_S = \{o_i\}$, one of which is identified as a target $t$, but assumes that their partner sees an additional hidden object $o_h$ that is not in common ground.
+The speaker has uncertainty over the identity of this object, represented by a prior distribution $P(o_h)$ which they marginalize over when deciding what utterance to produce:
+$$S_1(u | t, \mathcal{O}_p) \propto \sum_{o_h} L_0(t| \mathcal{O}_S \cup \{o_h\})P(o_h)P(u)$$
+
+\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$.
+The listener, however, also has some uncertainty over whether the experimenter is deceptive, represented by a Bernoulli random variable $d$, such that they assign a small prior probability $P(d)$ to the possibility that $o_h$ is actually in common ground.
+Upon hearing an utterance$u$, they update both their beliefs about which object is the target and whether the experimenter is being deceptive:
+$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
+\end{enumerate}
+Technical details are omitted here for brevity, but the full model is fully specified in WebPPL and available to run at %\url{http://forestdb.org/models/keysar.html}.
+
+%We focus on two critical patterns of results produced by the pragmatic listener $L_2$.
+First, we consider how it responds to the scripted instruction ``fish'' when presented with a context containing two fish (a red one that is occluded and a blue one in common ground) as well as two other unrelated objects.
+While it is most likely to choose the red fish in common ground, it has a relatively high probability (43\%) of choosing the hidden red fish, an effect \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} claimed was due to failure of theory of mind.
+This decision emerges in the model for two primary reasons.
+For one, a speaker with uncertainty over context could in principle be communicating about any object, and if they believed the hidden object was a red fish, the utterance `fish' would have a high probability of being produced.
+To a lesser extent, context uncertainty makes the overinformative utterance `blue fish' more likely than `fish', hence the omission of the color modifier has the (subtle) implication of \emph{not} the blue fish.
+This speaker phenomenon, while subtle under the parameter regime of the current implementation, produces the dynamic observed in the unscripted replication by  \cite{HawkinsGoodman16_Keysar}.
+%
+%Second, we consider the listener's posterior over the \emph{deception} Bernoulli variable $d$.
+Just as \cite{RubioFernandez16_DirectorTaskAttention} found that listeners were more likely to suspect deception after hearing an `overinformative' utterance like `blue fish' when only one fish was in common ground, so too does our pragmatic listener.
+The pragmatic speaker is relatively more likely to produce the additional color modifier if they see multiple fish than when they see a single fish and simply have uncertainty over the unknown object.
+Hence, our model simultaneously captures all three phenomena.
+
+\section{General Discussion}
 
@@
 
-This finding is consistent with a recent proposal by \cite{HellerParisienStevenson16_ProbabilisticWeighing} that referring expressions are interpreted by probabilistically integrating multiple sources of information: when conversational expectations lead participants to expect over-informative utterances, they (rationally) place relatively less weight on features of the environment determining common ground, such as shared perceptual access.
 Further work in a wider variety of tasks is necessary to pin down the various factors determining the relative weighting of these different sources of information, but we have argued that pragmatics play a crucial role.
 
-While We expect our conversation partners to follow common conventions about the meanings of words \cite{Lewis69_Convention, CentolaBaronchelli15_ConventionEmergence}, to update these conventions in light of previous interactions \cite{BrennanClark96_ConceptualPactsConversation}, and perhaps most importantly,
+Previous theoretical advances in modeling pragmatics have typically hinged on `lifting' inference over some aspect of communication that was previously assumed to be fixed.
+For instance, \emph{manner implicatures} emerge from uncertainty over the true meanings of words in the lexicon \cite{BergenLevyGoodman16_LexicalUncertainty}, nonliteral constructions like \emph{hyperbole} and \emph{irony} emerge from uncertainty over the topic of discussion \cite{KaoWuBergenGoodman14_NonliteralNumberWords}, and \emph{prosody} and \emph{ellipsis} emerge from uncertainty (in the form of a noisy-channel model) over the various ways that a speaker's utterance could have been corrupted  \cite{BergenGoodman15_StrategicUseOfNoise}.
+Note that these are not ad hoc extensions; sources of uncertainty that are not strictly necessary to explain a particular effect are often omitted for simplicity, but they all coexist within the same model.
+
+% While We expect our conversation partners to follow common conventions about the meanings of words \cite{Lewis69_Convention, CentolaBaronchelli15_ConventionEmergence}, to update these conventions in light of previous interactions \cite{BrennanClark96_ConceptualPactsConversation}, and perhaps most importantly,
 
@@
 
-%\todo[inline]{rdh: can take out next paragraph if hurting for space...}
-%A final concern is our graphical representation of occluded cells.
-We told players that items behind black cells were hidden to the director, showed them side-by-side displays of the director and matcher views, and quizzed them on this detail in the short attention check that players had to pass before playing.
-But they may have forgotten part-way through the experiment, or failed to fully internalize it.
-In the in-lab version, there was an actual divider between players, so it would have been more salient.
-We suspect that some subset of errors in our version game aren't due to a failure of perspective-taking, but instead because of a misunderstanding of what the other person's perspective is to begin with.
-However, this source of errors is held constant across our two experiments, and it would only bias our estimates of errors upward.
+%
+%\section{Sample Section}
+%Text here. Text here. Text here. Text here.
+%Text here. Text here. Text here. Text here.
+%Text here. Text here. Text here. Text here.
+%Text here. Text here. Text here. Text here.
+%
+%\subsection{Sample Subsection}
+%Text here. Text here. Text here. Text here.
+%Text here. Text here. Text here. Text here.
+%Text here. Text here. Text here. Text here.
+%Text here. Text here. Text here. Text here.
+%
+%\subsubsection{Sample Subsubsection}
+%Text here. Text here. Text here. Text here.
+%Text here. Text here. Text here. Text here.
+%Text here. Text here. Text here. Text here.
+%Text here. Text here. Text here. Text here.
+%
+%\section{Sample equations}
+%\begin{equation}
+%\label{eq:rhoCHT}
+%\rho^{\pi}= \frac{RI + \mathbb{E}_{\pi([L,\tau_L]|\textrm{post})}
+%\left[C_L(\taupav+\tau_L) \right]   +
+%\displaystyle{\int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}
+%\Biggl[\/\sum_{n_{L|[\textrm{pre},w]}}C_L(\tau_L)
+%\Biggr]            }      {P +
+%\mathbb{E}_{\pi([L,\tau_L] |\textrm{post})}[\tau_{L}] +\taupav +
+%\displaystyle{ \int_{0}^{P}}{dw~ \mathbb{E}_{\pi_{w_L}}}   
+%\Biggl[\sum_{n_{L|[\textrm{pre},w]}}\tau_L\Biggr]  
+%}
+%\end{equation}
+%As long as
+%$RI - K_LP > 
+%\frac{1}{\beta}$
+%\begin{equation}
+%%\def\theequation{5.1}
+%\left.\begin{array}{lrcl}
+%&\rho^{\pi} &=&  \displaystyle\frac{\beta ( RI + K_L \taupav )-1} {\beta
+%(P+\taupav )}    \\[12pt]
+%\hbox{and}\hbox to .25in{\hfill}&\mathbb{E}[\tau_L | \text{post}] &=&\displaystyle \frac{P+\taupav}{\beta ( RI -
+%K_LP)-1}  
+%\label{eq:analytical_linear}
+%\end{array}\right\}\hbox to 1.25in{\hfill}
+%\end{equation} 
+%
+%\subsection{Simple code sample}
+%
+%\begin{code}
+%\begin{verbatim}
+%procedure bubbleSort( A : list of sortable items )
+%    n = length(A)
+%    repeat
+%       newn = 0
+%       for i = 1 to n-1 inclusive do
+%          if A[i-1] > A[i] then
+%             swap(A[i-1], A[i])
+%             newn = i
+%          end if
+%       end for
+%       n = newn
+%    until n = 0
+%end procedure
+%\end{verbatim}
+%\end{code}
+%
+%
+%
+%\subsection{Algorithm environment}
+%
+%%% \begin{algorithm} takes option [p][b][t][h],  or some combination, like \begin{figure}
+%%% See documentation for algorithmic.sty for more information on formatting algorithms.
+%
+%\begin{algorithm}[h]
+%\caption{A sample in an algorithm environment.}
+%\begin{algorithmic}
+%\If {$i\geq maxval$}
+%    \State $i\gets 0$
+%\Else
+%    \If {$i+k\leq maxval$}
+%        \State $i\gets i+k$
+%    \EndIf
+%\EndIf
+%\end{algorithmic}
+%\end{algorithm}
+%
+%
+%\section{Jargon definitions}
+%
+%\begin{wideboxedtext}
+%\begin{glossary}
+%\symbol{Term} Definition
+%\symbol{Term} Definition
+%%ie
+%%\symbol{$1/\lambda$} mean of exponential effective prior probability density for leisure time
+%%\symbol{CHT} Cumulative Handling Time 
+%\end{glossary}
+%\end{wideboxedtext}
+%
+%\begin{boxedtext}
+%\begin{glossary}
+%\symbol{Term} Definition
+%\symbol{Term} Definition
+%%ie
+%%\symbol{$1/\lambda$} mean of exponential effective prior probability density for leisure time
+%%\symbol{CHT} Cumulative Handling Time 
+%\end{glossary}
+%\end{boxedtext}
+%
+%\section{Itemized Lists}
+%
+%\subsection{Roman list:}
+%
+%\begin{enumerate}
+%\item[(i)] at high 
+%payoffs, subjects work almost continuously, engaging in little leisure
+%in between work bouts; 
+%\item[(ii)] at low payoffs, they 
+%engage in leisure all at once, in long bouts after working, rather
+%than distributing the same amount of leisure time into multiple short
+%leisure bouts; 
+%\item[(iii)] subjects work continuously for the entire price duration, as long as
+%the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
+%\item[(iv)] the duration of leisure bouts is variable.
+%\end{enumerate}
+%
+%\subsection{Numbered list:}
+%
+%\begin{enumerate}
+%\item at high 
+%payoffs, subjects work almost continuously, engaging in little leisure
+%inbetween work bouts; 
+%\item at low payoffs, they 
+%engage in leisure all at once, in long bouts after working, rather
+%than distributing the same amount of leisure time into multiple short
+%leisure bouts; 
+%\item subjects work continuously for the entire price duration, as long as
+%the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
+%\item the duration of leisure bouts is variable.
+%\end{enumerate}
+%
+%
+%\subsection{Bulleted list:}
+%
+%\begin{itemize}
+%\item at high 
+%payoffs, subjects work almost continuously, engaging in little leisure
+%inbetween work bouts; 
+%\item at low payoffs, they 
+%engage in leisure all at once, in long bouts after working, rather
+%than distributing the same amount of leisure time into multiple short
+%leisure bouts; 
+%\item subjects work continuously for the entire price duration, as long as
+%the price is not very long (as shown by an analysis conducted by Y-AB, to be published separately); %(\textbf{Figure \ref{fig:task_data}D}).
+%\item the duration of leisure bouts is variable.
+%\end{itemize}
+%
+%
+%\section{Sample citations}
+%For general information on the correct form for citations using
+%the APA 6 format, see the following sites:
+%\href{https://owl.english.purdue.edu/owl/resource/560/02/}
+%{APA 6, In-text citations, The Basics} and
+%\href{https://owl.english.purdue.edu/owl/resource/560/03/}
+%{APA 6, In-text citations}
+%
+%\section{Natbib citation mark up}
+%
+%\subsection{Single citations}
+%\noindent
+%\begin{tabular}{ll}
+%\bf Type&\bf Results\\
+%\hline
+%\verb+\citet{jon90}+&Jones et al. (1990)\\
+%\verb+\citet[chap. 2]{jon90}+&Jones et al. (1990, chap. 2)\\
+%    \verb+\citep{jon90}+	    &   	(Jones et al., 1990)\\
+%    \verb+\citep[chap. 2]{jon90}+ 	&    	(Jones et al., 1990, chap. 2)\\
+%    \verb+\citep[see][]{jon90}+ 	 &    	(see Jones et al., 1990)\\
+%    \verb+\citep[see][chap. 2]{jon90}+ 	&    	(see Jones et al., 1990, chap. 2)\\
+%    \verb+\citet*{jon90}+ 	    &    	Jones, Baker, and Williams (1990)\\
+%    \verb+\citep*{jon90}+	    &    	(Jones, Baker, and Williams,
+%    1990) \\
+%\end{tabular}
+%
+%For example, some citations from the OpenMindSample bibliography:
+%citet:\citet{anderson}, citep:\citep{antibayes}, and
+%cite*: \citet*{anderson}.
+%
+%\subsection{Multiple citations}
+%Multiple citations may be made by including more than one citation
+%key in the \verb+\cite+ command argument.
+%
+%\noindent
+%\begin{tabular}{ll}
+%\bf Type&\bf Results\\
+%\hline
+%\verb+\citet{jon90,jam91}+&Jones et al. (1990); James et al. (1991)\\
+%\verb+\citep{jon90,jam91}+&(Jones et al., 1990; James et al. 1991)\\
+%\verb+\citep{jon90,jon91}+&(Jones et al., 1990, 1991)\\
+%\verb+\citep{jon90a,jon90b}+&(Jones et al., 1990a,b)\\
+%\end{tabular}
+%
+%For example, multiple citations from the OpenMindSample bibliography:
+%citet:\citet{anderson,antibayes}, citep:\citep{anderson,antibayes}.
+%As you see, the citations are automatically hyperlinked to their
+%reference in the bibliography.
+%
+%\newpage
+%
+%\section{Sample figures}
+%
+%\begin{figure}[h] 
+%\centerline{\includegraphics[width=\textwidth]{Fig1}}
+%\caption{(Colour online) \textbf{Task and key features of the
+% data.} \\
+% A) Cumulative handling time (CHT) task. Grey bars denote work
+%(depressing a lever), white gaps show leisure. The subject must
+% accumulate work up to a total period of time called the
+%\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
+%intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
+%$2$s each time the price is attained, during which the lever is retracted so it cannot
+%work; not shown).
+%}
+%\label{fig:task_data}
+%\end{figure}
+%
+%\begin{figure}[ht] 
+%\widefigure{\fullpagewidth}{Fig1}
+%\caption{(Colour online) \textbf{Task and key features of the
+% data.} \\
+%A) Cumulative handling time (CHT) task. Grey bars denote work
+%(depressing a lever), white gaps show leisure. The subject must
+%accumulate work up to a total period of time called the
+%\emph{price} ($P$) in order to obtain a single reward (black dot) of subjective reward
+%intensity $RI$. The trial duration is $25\times \mathrm{price}$ (plus
+%$2$s each time the price is attained, during which the lever is retracted so it cannot
+%work; not shown).
+%}
+%\label{newfig:task_data}
+%\end{figure}
+%
+%\clearpage
+%\section{Sample tables}
+%
+%\begin{table}[!ht]
+%\caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
+%\label{tab:label}
+%\centering
+%\begin{tabular}{lc}
+%\hline
+% Run  & Time (min)  \\
+%\hline
+%  $l1$  & 260   \\
+%  $l2$  & 300   \\
+%  $l3$  & 340   \\
+%  $h1$  & 270   \\
+%  $h2$  & 250   \\
+%  $h3$  & 380   \\
+%  $r1$  & 370   \\
+%  $r2$  & 390   \\
+%\hline
+%\multicolumn{2}{l}{$^{a}$Table note text here.}
+%\end{tabular}
+%\end{table}
+%
+%\begin{table}[ht]
+%\widecaption{Sample table taken from [treu03]\label{tbl-1}}
+%\begin{widetable}
+%\advance\tabcolsep-1pt
+%\small
+%\begin{tabular}{ccrrccccccccc}
+%\hline
+%\bf 
+%POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
+%&\multicolumn1c{\bf Y} &\bf
+%RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
+%IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
+%IAP2 &\bf star &\bf E &\bf Comment\\
+%\hline
+%0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
+%21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
+%0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
+%0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
+%0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
+%0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
+%0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
+%0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
+%0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
+%22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
+%edge \\
+%\hline
+%\multicolumn{13}{l}{%
+%Table 2 is published in its entirety in the electronic
+%edition of the {\it Astrophysical Journal}.}\\[3pt]
+%\multicolumn{13}{l}{%
+%$^a$ Sample footnote for table 2.}\\[3pt]
+%\multicolumn{13}{l}{%
+%$^b$ Another sample footnote for table 2.}
+%\end{tabular}
+%\end{widetable}
+%\end{table}
+%
+%\begin{table}[p]
+%\rotatebox{90}{\vbox{\hsize=\textheight
+%\caption{Here is a caption for a table that is found in landscape
+%mode.}
+%\begin{tabular}{ccrrccccccccc}
+%\hline
+%\bf 
+%POS &\bf  chip &\multicolumn1c{\bf ID} &\multicolumn1c{\bf X}
+%&\multicolumn1c{\bf Y} &\bf
+%RA &\bf DEC &\bf IAU$\pm$ $\delta$ IAU &\bf
+%IAP1$\pm$ $\delta$ IAP1 &\bf IAP2 $\pm$ $\delta$
+%IAP2 &\bf star &\bf E &\bf Comment\\
+%\hline
+%0 & 2 & 1 & 1370.99 & 57.35\rlap{$^a$}    &   6.651120 &  17.131149 &
+%21.344$\pm$0.006\rlap{$^b$}  & 2 4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
+%0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2 3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
+%0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2 4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
+%0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2 5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
+%0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2 3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
+%0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2 6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
+%0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2 5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
+%0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 &
+%22.189$\pm$0.011  & 2 4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 &
+%edge \\
+%\hline
+%\multicolumn{13}{l}{%
+%Table 2 is published in its entirety in the electronic
+%edition of the {\it Astrophysical Journal}.}\\[3pt]
+%\multicolumn{13}{l}{%
+%$^a$ Sample footnote for table 2.}\\[3pt]
+%\multicolumn{13}{l}{%
+%$^b$ Another sample footnote for table 2.}
+%\end{tabular}
+%}}
+%\end{table}
+%\clearpage
+%
+%
+%\vglue 3in
+%Example of table continuing over pages:
+%
+%
+%\begin{center}
+%\begin{longtable}{ccc@{}}
+%\caption{ApJ costs from 1991 to 2013
+%\label{tab:table}} \\[2pt]
+%\hline
+%\bf Year & \bf Subscription & \bf Publication \\
+% & \bf cost &\bf charges\\
+% & \bf(\$) & \bf (\$/page)\\
+%\hline
+%\endfirsthead
+%
+%\multicolumn3c{Table \thetable, \it continued from previous page.}\\[6pt]
+%\multicolumn3c{ApJ costs from 1991 to 2013}\\[2pt]
+%\hline
+%\bf Year & \bf Subscription & \bf Publication \\
+% & \bf cost &\bf charges\\
+% & \bf(\$) & \bf (\$/page)\\
+%\hline
+%\endhead
+%\\\hline
+%\\[-8pt]
+%\multicolumn{3}{r}{\it Table continued on next page}\\ 
+%\endfoot
+%
+%\hline
+%\endlastfoot
+%
+%1991 & 600 & 100 \\
+%1992 & 650 & 105 \\
+%1993 & 550 & 103 \\
+%1994 & 450 & 110 \\
+%1995 & 410 & 112 \\
+%1996 & 400 & 114 \\
+%1997 & 525 & 115 \\
+%1998 & 590 & 116 \\
+%1999 & 575 & 115 \\
+%2000 & 450 & 103 \\
+%2001 & 490 &  90 \\
+%2002 & 500 &  88 \\
+%2003 & 450 &  90 \\
+%2004 & 460 &  88 \\
+%2005 & 440 &  79 \\
+%2006 & 350 &  77 \\
+%2007 & 325 &  70 \\
+%2008 & 320 &  65 \\
+%2009 & 190 &  68 \\
+%2010 & 280 &  70 \\
+%2011 & 275 &  68 \\
+%2012 & 150 &  56 \\
+%2013 & 140 &  55 \\
+%\end{longtable}
+%\end{center}
 
-\subsection*{Supporting Information (SI)}
+\section{Supportive Information}
+Here you enter further sources of information, if desired.
 
-The main text of the paper must stand on its own without the SI.
-Refer to SI in the manuscript at an appropriate point in the text.
-Number supporting figures and tables starting with S1, S2, etc.
-Authors are limited to no more than 10 SI files, not including movie files.
-Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods.
-If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
-Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only.
+%% A possible entry might be:
+% No supportive information is available at this time.
 
-\subsubsection*{SI Text}
+\acknowledgments
+Enter your acknowledgments here.
 
-Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).
+%% ie.,
 
-\subsubsection*{SI Figures}
+% The authors thank Laurence Aitchison for fruitful discussions.  RKN
+% and PD received funding from the Gatsby Charitable Foundation. Y-AB,
+% RBS, KC and PS received funding from Canadian Institutes of Health
+% Research grant $MOP74577$,
+% Fond de recherche Qu\'{e}bec - Sant\'{e} (Group grant to the Groupe 
+% de recherche en neurobiologie comportementale, Shimon Amir, P.I.), and
+% Concordia University Research Chair (Tier I).
 
-Provide a brief legend for each supporting figure after the supporting text.
-Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text.
-When saving TIFF files, use only LZW compression; do not use JPEG compression.
-Do not save figure numbers, legends, or author names as part of the image.
-Composite figures must be pre-assembled.
+\authorcontributions 
+Who helped formulate the project, who supplied data, analyses and
+experiments, etc.
 
-\subsubsection*{3D Figures}
+%% ie.
+%% Project was formulated by RKN, PD, PS,
+%% based on substantial data, analyses and experiments of Y-AB, KC, RS,
+%% PS. RKN, PD formalised the model, RKN implemented and ran the model;
+%% RKN analysed the molecular ethogram data; Y-AB formalised and
+%% implemented a CTMC model. All authors wrote the manuscript.
 
-Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.
+%%%%%%%%%%%%%%%%%%%%%%%
+%% The bibliography
 
-\subsubsection*{SI Tables}
+%% \nocite{*} is used here as a quick way to get every entry  in the .bib file to
+%% appear in the bibliography. Normally the bibliography is made using only
+%% the bibentries that you cite using \cite{}, or one of the Natbib citation
+%% entries, like \citep{}, \citet{} etc.
 
-Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
-Do not use tabs or spaces to separate columns in Word tables.
+%\nocite{*}
+\bibliography{refs}
 
-\subsubsection*{SI Datasets}
+\appendix
 
-Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
+\section{Sample Appendix Section}
+We derive the result in Eq. \eqref{eq:analytical_linear}. We consider a linear $C_L(\tau_L+\taupav)=K_L(\tau_L+\taupav)$, and
+make two further 
+simplifications: (i) the subject does not
+engage in leisure in the pre-reward state (and so works for the whole
+price when it works); and (ii) \emph{a priori}, arbitrarily long leisure durations are possible
+($\lambda=0$).
+Then the reward rate in Eq. \eqref{eq:rhoCHT} becomes
+\begin{equation}\label{eq:analyticalrho}
+\rho^{\pi}= \frac{\vrule height 10pt width0pt RI + K_L \{~ \mathbb{E}[ \tau_L | \textrm{post} ] + \taupav  \} }
+     {P +
+\mathbb{E}[ \tau_L | \textrm{post} ] +\taupav  } 
+\end{equation} 
+As discussed in the \emph{Results} section, the probability of engaging in instrumental leisure in the post-reward state is $\pi([L,\tau_L]
+~| \textrm{post}) = \exp\left[-\{\beta (\rho^\pi-K_L) 
+ \} \tau_L\right]$, which is an exponential distribution with
+mean 
+\begin{equation}
+\mathbb{E}[\tau_L | \textrm{post}]=\frac{1}{\beta (\rho^\pi-K_L) }
+\label{eq:analyticaltauL}
+\end{equation}  
+Re-arranging terms of this equation,
+\begin{equation}\label{eq:rhoversion2}
+\rho^{\pi}=\frac{1}{\beta ~\mathbb{E}[\tau_L | \textrm{post}]} +K_L 
+\end{equation} 
+Equating Eqs. \eqref{eq:analyticalrho} and \eqref{eq:rhoversion2} and solving for the mean instrumental leisure duration $\mathbb{E}[\tau_L | \textrm{post}]$, we derive
+\begin{equation}
+\mathbb{E}[\tau_L | \textrm{post}] = \frac{P+\taupav}{\beta ( RI - K_LP)-1} 
+\label{eq:solvedtauL}
+\end{equation}
+which is the second line of Eq.\eqref{eq:analytical_linear}.
+This is the mean instrumental leisure duration as long as  $RI - K_LP>1$, and  $\mathbb{E}[\tau_L | \textrm{post}] \rightarrow \infty$ otherwise.
+When the former condition holds, we may
+substitute Eq. \eqref{eq:solvedtauL} into Eq. \eqref{eq:analyticalrho} and solve for $\rho^{\pi}$
 
-\subsubsection*{SI Movies}
+\begin{figure}
+\widefigure{\fullpagewidth}{Fig1}
+\caption{Sample Appendix Caption. Here is a caption that might
+appear in an appendix. It is as wide as the full width of the page.}
+\end{figure}
 
-Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file.
-All movies should be submitted at the desired reproduction size and length.
-Movies should be no more than 10 MB in size.
+\newpage
 
-\subsubsection*{Still images}
+\section{Making Your Bibliography for an Open Mind Article}
+{\it Open Mind} uses the APA author-date  bibliography style,
+apacite.bst. For more
+information on apacite, for examples in how to make your .bib file and more, see:\\
+\href{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}
+{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}
 
-Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
+\noindent
+(In spite of the mention of apacite cite commands, please use only
+Natbib commands for in text citations, as shown above.)
 
-\subsubsection*{Appendices}
+\subsection{BibTeX}
+You will need to use BibTeX to form your bibliography; typing in the
+references would be
+a huge and unpleasant task. Look at the openmindsample.bbl file and you'll see why
+typing in the bibitems would be difficult.
 
-PNAS prefers that authors submit individual source files to ensure readability.
-If this is not possible, supply a single PDF file that contains all of the SI associated with the paper.
-This file type will be published in raw format and will not be edited or composed.
+For a good basic introduction to using BibTeX, see
+\href{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}
+{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}
 
-\matmethods{Please describe your materials and methods here.
-This can be more than one paragraph, and may contain subsections and equations as required.
-Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.
-
-\subsection*{Subsection for Method}
-Example text for subsection.
-}
-
-\showmatmethods{} % Display the Materials and Methods section
-
-\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}
-
-\showacknow{} % Display the acknowledgments section
-
-% \pnasbreak splits and balances the columns before the references.
-% Uncomment \pnasbreak to view the references in the PNAS-style
-% If you see unexpected formatting errors, try commenting out \pnasbreak
-% as it can run into problems with floats and footnotes on the final page.
-%\pnasbreak
-
-% Bibliography
-\bibliography{pnas-submission}
+When you use BibTeX, the form of the bibliography will be correct. You
+don't need to supply a bibliography style, since that is built into
+the stjour.cls file.
 

### 2018-03-15 experiments/final/experiment2/game.server.js
Fix some data output bugs

@@ -83,14 +83,13 @@ var dataOutput = function() {
 
   function getObjectLocs(objects) {
     return _.flatten(_.map(objects, o => {
-      return [o.name, o.speakerCoords.gridX, o.speakerCoords.gridY,
-	      o.listenerCoords.gridX, o.listenerCoords.gridY];
+      return [o.name, o.gridX, o.gridY];
     }));
   }
 
   function getObjectLocHeaderArray() {
     return _.flatten(_.map(_.range(1,5), i => {
-      return _.map(['name', 'speakerX', 'speakerY', 'listenerX', 'listenerY'], v => {
+      return _.map(['name', 'gridX', 'gridY'], v => {
 	return 'object' + i + v;
       });
     }));
@@ -105,68 +104,41 @@ var dataOutput = function() {
       assignmentId: client.assignmentid
     };
   };
-
-  var postTestWordOutput = function(client, message_data) {
-    var target = message_data[1];
-    var selections = message_data.slice(2);
-    var meaningHeader = _.map(client.game.objects, 'name');
-    var meaning = _.map(client.game.objects, obj => _.includes(selections, obj.name));
-    return _.extend(
-      commonOutput(client, message_data),
-      _.zipObject(meaningHeader, meaning), {
-	target,
-	finalRole: client.role
-      });	
-  };
   
-  var postTestObjectOutput = function(client, message_data) {
-    var target = message_data[1];
-    var selections = message_data.slice(2);
-    var meaningHeader = client.game.trialInfo.labels;
-    var meaning = _.map(client.game.trialInfo.labels,
-			label => _.includes(selections, label));
-    return _.extend(
-      commonOutput(client, message_data),
-      _.zipObject(meaningHeader, meaning), {
-	target,
-	finalRole: client.role
-      });	
-  };
-
   var clickedObjOutput = function(client, message_data) {
     var objects = client.game.trialInfo.currStim.objects;
     var intendedName = getIntendedTargetName(objects);
     var objLocations = _.zipObject(getObjectLocHeaderArray(), getObjectLocs(objects));
     return _.extend(
       commonOutput(client, message_data),
+      client.game.trialInfo.currContextType,
       objLocations, {
 	intendedName,
 	clickedName: message_data[1],
 	trialNum : client.game.state.roundNum + 1,	
 	correct: intendedName === message_data[1],
-	condition: client.game.condition,
-	contextType: client.game.trialInfo.currContextType
[…]

### 2018-03-15 writing/journal_manuscript/manuscript.tex
Add experiment outline P to intro; fiddle with methods

@@
 Our success as a cooperative species depends upon our ability to make sense of other agents. 
-One solution is to maintain an internal predictive model of how minds work: how an agent's unobservable beliefs and desires give rise to observable actions. 
-Armed with a computational theory of mind, we can make inferences about an agent's likely mental state given their past behavior and form sophisticated expectations about their future behavior. 
+One solution is to maintain an internal predictive model of other agents: how unobservable beliefs and desires give rise to observable actions. 
+Such a model allows us to make inferences about a social partner's likely internal state given their past behavior and form sophisticated expectations about their future behavior. 
 %This theoretical approach to social cognition, where theory of mind reasoning \cite{PremackWoodruff78_ChimpanzeeToM, GopnikWellman92_TheoryTheory, WellmanCrossWatson01_ToMMetaAnalysis, SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM} is grounded in an explicit computational model of how other minds work, has
@@
 %While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
-Past efforts have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
+Past efforts to evaluate theory of mind in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
 We argue in this paper that the mental models supporting theory of mind use in communication are much richer than previously explored. 
@@
 
-%This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
-%If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
-%More precisely, when both players know that objects are occluded in the display, the speaker may tend to add additional precision to references in order to avoid confusion.
+This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
+If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
+More precisely, when both players know that objects are occluded in the display, the speaker may tend to add additional precision to references in order to avoid confusion.
 If the listener \emph{expects} the speaker to do this, they will pragmatically pick the \emph{a priori} more likely referent of the referring expression, which in critical trials will be the occluded object.
@@
 In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
+In Experiment 1, we conduct a direct replication of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, manipulating whether the speaker is a confederate or naive participant.
+We find that unscripted speakers naturally produce more informative utterances than those in the confederate's script.
+In Experiment 2, we test the hypothesis that speakers systematically produce longer, more specific utterances than strictly required by their visible context when they know objects are occluded.
+Finally, in Experiment 3, we manipulate the listener's pragmatic beliefs about whether the speaker is indeed expected to produce such `over-informative' utterances.
+Taken together, these results suggest that apparent failures of perspective-taking are in fact driven by sophisticated pragmatic reasoning; in other words, by successful theory of mind use.
 
@@
 
-We recruited 34 participants (17 pairs) from Amazon Mechanical Turk. All participants were from the U.S. Three pairs were excluded for making 2 or more errors on non-critical items.
-
-\subsection{Materials \& Procedures}
-
-Participants interacted in a real-time, multi-player environment on the web \cite{Hawkins15_RealTimeWebExperiments}. 
+We recruited K participants (K/2 pairs) from Amazon Mechanical Turk to play an interactive communication game using the framework described in \cite{Hawkins15_RealTimeWebExperiments}.
+All participants were native English speakers.
+%Three pairs were excluded for making 2 or more errors on non-critical items.
+
+\subsection{Materials \& Procedure}
+
 %To deal with participants arriving at unpredictable times, we implemented a waiting room mechanism, where the first person to enter the game environment would see a screen saying ``Waiting for another player..." and the second person would trigger the game to start.
-Pairs of participants---assigned randomly to `director' and `matcher' roles---interacted with one another through a web interface, shown in Fig.
-\ref{fig:interface}.
-On the left side of the screen, participants could freely type messages to one another; on the right side the screen, players could view a set of objects placed in a 4 x 4 grid.
-Five of the grid cells were occluded from only the director's perspective, and the remaining 11 were visible to both matcher and director.
-Six or seven objects were displayed in the grid at a given time.
+Participants were randomly assigned to `director' and `matcher' roles and placed in a shared environment containing a chat box and a $4 \times 4$ grid of objects (see Fig.
+\ref{fig:interface}).
+Following \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, five of the grid cells were occluded from the director's perspective, and six or seven objects were displayed in the grid at a given time.
+On each trial, the director was shown an instruction indicating how one of the objects should be moved.
+They then used the chatbox to send a message to their partner  The experiment was composed of eight items, with each item using a different set of objects.
+Each item included one `critical pair' of objects, one of them the target and the other hidden, such as the cassette tape and the roll of tape.
+For each item, we gave the director a series of four instructions to move objects around, which were displayed as a series of arrows pointing from some object to an unoccupied cell.
+
+Participants could freely type messages to one another at any point.
+
 One of these objects was a `target', such as a cassette tape, placed in an unoccluded cell such that both participants could see it.
@@
 We used the same set of targets and occluded alternatives as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, but we were unable to obtain the filler objects from the original experiment and created our own.
+
 Before entering the game environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface.
-Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
-
-The experiment was composed of eight items, with each item using a different set of objects.
-Each item included one `critical pair' of objects, one of them the target and the other hidden, such as the cassette tape and the roll of tape.
-For each item, we gave the director a series of four instructions to move objects around, which were displayed as a series of arrows pointing from some object to an unoccupied cell.
+%Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
+
 To collect clean mouse-tracking data, we began every instruction by asking the matcher to click a small circle in the center of the grid.

### 2018-03-20 analysis/final_analysis.Rmd
Initial pilot analyses

---
title: "R Notebook"
output: html_notebook
---

# Experiment 2

## Import message data

```{r}
d_raw <- read_tsv('../data/final/experiment2/pilot0/chatMessage/pilot0_messages.csv')

# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
incompleteIDs <- unique((d %>% group_by(gameid) %>% 
                           filter(length(unique(trialNum)) != 24))$gameid)

# Some speakers relied purely on location in grid
pureLocation <- c('0180-a541e54c-5dea-4d06-9a21-9b842316eb59', '0101-0b659560-f05a-471a-89ac-60469e9c867f')
badGames <- c(incompleteIDs, pureLocation)

d <- d_raw %>%
  group_by(gameid, trialNum, occlusions, context) %>%
  summarize(text = paste0(text, collapse = ' ')) %>%   # Concatenate multiple utterance on same trial
  mutate(numRawWords = str_count(text, "\\S+")) %>%
  mutate(hidden = ifelse(occlusions == 'none', 'no', 'yes')) %>%
  filter(!(gameid %in% badGames))
```

## Check basic (no-hidden) pragmatic effect

```{r}
d %>%
  filter(hidden == 'no') %>%
  group_by(gameid, context) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(context, m) %>%
  mutate(diff = close - far) %>%
  ggplot(aes(x = far, y = close)) +
    geom_point() +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,12) +
    xlim(1,12) +
    theme(aspect.ratio = 1)

  summarize(m = mean(numRawWords), se = sd(numRawWords)/sqrt(length(numRawWords))) %>%
    spread(context, numRawWords)

```

## Message length collapsing across all subjects

```{r}
dodge <- position_dodge(width=0.9)

d %>%
  group_by(context, hidden) %>%
  summarize(m = mean(numRawWords), se = sd(numRawWords)/sqrt(length(numRawWords))) %>%
  ggplot(aes(x = hidden, y = m, fill = context)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), position = dodge, width = 0.1) +
    theme_bw()
```

## Scatter to look at within-subj effects

```{r}
d %>%
  group_by(gameid, context, hidden) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(hidden, m) %>%
  mutate(diff = yes - no) %>%
  ggplot(aes(x = no, y = yes)) +
    geom_point() +
    facet_wrap(~ context) +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,12) +
    ylab('# words (with occlusions)') +
    xlab('# words (without occlusions)') +
    ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
    xlim(1,12) +
    theme(aspect.ratio = 1)

ggsave('~/Downloads/scatter.pdf')
```

```{r}
d %>%
  group_by(gameid, context, hidden) %>%
  summarize(m = mean(numRawWords)) %>%
  spread(hidden, m) %>%
  mutate(diff = yes - no) %>%
  group_by(context) %>%
  summarize(m = mean(diff), se = sd(diff)/sqrt(length(diff))) %>%
  ggplot(aes(x = context, y = m)) +
    geom_bar(stat = 'identity') +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0) +
    theme_few() +
    ylim(-3, 3) +
    ylab('# words (occlusions - no occlusions)')

```

# Regression

helmert coding way (one way of slicing)

```{r}
library(ggthemes)
library(lme4)
library(lmerTest)

d.regression <- d %>% 
  unite(condition, context, hidden) %>%
  ungroup() %>%
  mutate(condition = ordered(condition, levels = c('far_no', 'close_no', 'far_yes', 'close_yes')))
contrasts(d.regression$condition) <- contr.helmert(4)
summary(lmer(numRawWords ~ condition + (1 + condition | gameid), data = d.regression))
```

interaction-y way

```{r}
summary(lmer(numRawWords ~ context * hidden + (1 + context + hidden | gameid), data = d))
```

### 2018-03-20 writing/journal_manuscript/manuscript.tex
Streamline Exp. 1; start Exp. 2

@@
 %While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
-Past efforts to evaluate theory of mind in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
+Past efforts to evaluate theory of mind in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \citep{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
 We argue in this paper that the mental models supporting theory of mind use in communication are much richer than previously explored. 
-In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \cite{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
+In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \citep{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
 For example, suppose a friend wants to refer to a nearby Dalmatian. 
@@
 If there are no other dogs nearby, we may expect them to use a parsimonious basic-level term like ``dog.''
-When there is potential for confusion, however, we expect them to consider our likely interpretation and introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \cite{BrennanClark96_ConceptualPactsConversation, VanDeemter16_ComputationalModelsOfReferring, GrafEtAl16_BasicLevel}.
+When there is potential for confusion, however, we expect them to consider our likely interpretation and introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \citep{BrennanClark96_ConceptualPactsConversation, VanDeemter16_ComputationalModelsOfReferring, GrafEtAl16_BasicLevel}.
 When speakers violate these expectations, listeners draw inferences about the intended meaning in order to account for the violation: for example, XXX \cite{XXX, BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}.
@@
 
+% Discuss other criticisms of Keysar paradigm?
+While these results are compelling, the paradigm has been criticized from few different angles. 
+\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
+the hidden roll of tape vs.
+the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well.
+We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
+Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
+
 This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
@@
 \ref{fig:interface}).
-Following \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, five of the grid cells were occluded from the director's perspective, and six or seven objects were displayed in the grid at a given time.
-On each trial, the director was shown an instruction indicating how one of the objects should be moved.
-They then used the chatbox to send a message to their partner  The experiment was composed of eight items, with each item using a different set of objects.
-Each item included one `critical pair' of objects, one of them the target and the other hidden, such as the cassette tape and the roll of tape.
-For each item, we gave the director a series of four instructions to move objects around, which were displayed as a series of arrows pointing from some object to an unoccupied cell.
-
-Participants could freely type messages to one another at any point.
-
-One of these objects was a `target', such as a cassette tape, placed in an unoccluded cell such that both participants could see it.
-Another object, such as a roll of tape, was placed in an occluded slot such that it was only visible to the matcher.
-The rest of the objects were unrelated `fillers' placed in random locations.
-We used the same set of targets and occluded alternatives as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, but we were unable to obtain the filler objects from the original experiment and created our own.
-
-Before entering the game environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface.
-%Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
-
-To collect clean mouse-tracking data, we began every instruction by asking the matcher to click a small circle in the center of the grid.
-After this small circle was clicked, the director was allowed to communicate the next instruction and we started recording from the matcher's mouse.
-One of the instructions was a `critical instruction,' which referred to the target object.
-For half the instructions, directors were free to communicate however they wished.
-For the other half, including all the critical instructions, their messages to the matcher were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat window, and they would have to click `send' for the experiment to continue.)
-
+Five of the grid cells were blocked from the director's view by a curtain, and six or seven objects were displayed in the grid at a given time.
+Trials were blocked into eight `items', each using a different set of objects.
+For each item, we gave the director a series of four instructions to move objects, which were displayed as a series of arrows pointing from some object to a neighboring unoccupied cell.
+They used a chat box to communicate the identity of this object and action to their partner, who proceeded to click and drag an object to complete the trial.
+
+We used a between-subject design to test whether the scripted confederate labels were in fact what cooperative speakers would say in context.
+In the `scripted' condition, directors were free to communicate however they wished for half of the instructions but for the other half, including all the critical instructions, their messages were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat box, and they would have to click `send' for the experiment to continue.) In the `unscripted' condition, directors were unrestricted and free to send whatever messages they deemed appropriate on all trials.
+
+In each block, one target belonged to a `critical pair' of objects, such as a visible cassette tape and a hidden roll of tape that could both plausibly be referred to as `the tape'.
 We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
@@
 
-This paradigm differs from those used by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways.
-First, participants were not seated across from each other at a table: they each saw a view of the 4 x 4 grid on their screen and communicated via a text box.
-Second, we did not use a trained confederate.
-We randomly assigned one of the players to the role of the instruction giver, and maintained the original wording by scripting a subset of their instructions.
-Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \cite{KeysarBarr___Brauner00_TakingPerspective}.
+%On half of the instruction sets (the `critical' condition), the critical distractor object was placed in an occluded slot such that it was only visible to the matcher; on the other half (the `baseline' condition), all hidden distractors were unrelated fillers.
+
+%Before entering the game environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface.
+%Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
+
+To collect clean mouse-tracking data, we asked the matcher to wait until the director sent a message; when the message was received, the matcher clicked a small circle in the center of the grid to continue the trial.
+We started recording from the matcher's mouse after this click.
 
@@
 
-In Table \ref{table:mainResults} we show the error rates on critical items, and compare to the data from
-\cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-We find that 93\% of participants (all but one) attempted to move the hidden distractor at least once in the Experimental condition, out of four possible items, compared to only 7\% (only one) in the baseline condition.
-This is similar to the effect observed by the authors in the original study, which found 71\% and 0\%.
+\subsubsection{Critical errors}
+
+In Figure \ref{table:mainResults} we compare error rates across scripted and unscripted conditions.
+Participants rarely moved the hidden object on baseline trials in either condition, and total error rates for critical trials are significantly higher when speakers used scripted utterances, $\chi^2(1) = 5.35, p = 0.02$.
+Error rates in the scripted condition were similar to the effects observed by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+%, which found 71\% and 0\%.
 Our errors were larger across the board, perhaps due to the interface or the population, but the gap between the two conditions is roughly the same size.
 %Next, we examine whether there are differences in error rates across items.
-In Table \ref{table:ItemWise}, we break down the pattern of errors by item. 
-%Because our randomization procedure undersampled the experimental condition for several items, a chi-squared test is inappropriate.
-Using Fisher's exact test, we cannot reject the null hypothesis of independence ($p = 0.13$).
-Still,
-We note that several items have much higher error rates than others -- for example, 75\% of participants in the experimental condition of item 6 made an error (the ``whiteboard eraser'' vs.
-the ``pencil eraser'') while only 17\% of participants in item 8 made an error (the ``computer mouse'' vs.
-the ``toy mouse'').
-Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.
-We empirically substantiate this observation in our results for Expt.~2 below.
-
-This item-wise variability suggests that the dependent variable highlighted in the original study (i.e.
+
+We also found substantial heterogeneity in errors rates across items (Table \ref{table:ItemWise}).
+Patterns of errors diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
+For example, 75\% of participants made an error on item 6 upon hearing the scripted instruction ``the eraser'' while only 17\% of participants in item 8 made an error upon hearing ``the mouse.'' Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.
+Looking more closely at these patterns, we see that the only items where errors are consistently made in the \emph{unscripted} condition are those where the more precise utterances used remain ambiguous.
+For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
+Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' but the two objects were still confusable at the sub-class level.
+If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total error rate in the unscripted condition drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.
+
+%This item-wise variability suggests that the dependent variable highlighted in the original study (i.e.
 ``percentage of participants who moved the critical item at least once'') is somewhat problematic: it could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item.
@@
 
+\subsubsection{Mouse-tracking}
+
 As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
-We define the decision window as the span of time between the point when the matcher received their instruction message and when they started moving an object.
+%We define the decision window as the span of time between  their instruction message and when they started moving an object.
 If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt.
-Within the decision window, we computed the total amount of time spent hovering over the cell containing the target and divided by the total length of the decision window to get a measure of the relative time spent considering the target.
-We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
-A paired-samples $t$-test found that people tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig.
-\ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} .
+Within the decision window, we computed the total proportion of time spent moving toward the target relative to the distractor.
+%We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
+%A paired-samples $t$-test found that matchers in the scripted condition tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig.
+\ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.%\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} .
 %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$
+
+%When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we found no significant decrease in target hover time between experimental and baseline trials $t(27) = 0.89, p = 0.38$.
+To directly test for differences in hover time patterns across the two experiments, we used a mixed-effects model with a random intercept for game ID and an interaction between condition and experiment on target hover time.
+We found a marginally significant interaction, $b = 0.09, t = 1.89, p = 0.066$ (see Figure \ref{fig:hover}), providing some evidence that the presence of a hidden distractor no longer interfered target selection in Expt~2.
+
+\subsubsection{Informativity}
 
@@
 That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
-We tested this prediction in Expt.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
-By the reasoning above we expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
-
-% discuss other criticisms of Keysar paradigm?
-While these results are compelling, the paradigm has been criticized from few different angles. 
-\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
-the hidden roll of tape vs.
-the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well.
-We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
-% Although relative fit was not empirically evaluated prior to our study,
-Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
-
-Error rates are reported in Table \ref{table:mainResults}, alongside the results from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our scripted replication in Expt.~1.
-Participants never moved the hidden object in the baseline condition, and total error rates for experimental trials are significantly lower than the rates found in Expt.~1, $\chi^2(1) = 5.35, p = 0.02$.
-
-We find that patterns of errors in Expt.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
-Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the more precise utterances used remain ambiguous.
-For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
-Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' from Expt.~1, but the two objects were still confusable at the sub-class level.
-If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.
-
-When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we found no significant decrease in target hover time between experimental and baseline trials $t(27) = 0.89, p = 0.38$.
-To directly test for differences in hover time patterns across the two experiments, we used a mixed-effects model with a random intercept for game ID and an interaction between condition and experiment on target hover time.
-We found a marginally significant interaction, $b = 0.09, t = 1.89, p = 0.066$ (see Figure \ref{fig:hover}), providing some evidence that the presence of a hidden distractor no longer interfered target selection in Expt~2.
+We expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
+
+Next, we test whether these improvements in performance are in fact due to more informative speaker behavior.
+We recruited twenty judges on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects.
+Their responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.6$ (95\% CI = $[0.54, 0.66]$).
+In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig.
+\ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$.
+In Expt.~1, the scripted label fit the hidden distractor just as well or better than the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
+In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.
+
+Does differential informativity of scripted utterances account for the variability across items that we noted in Expt.~1?
+In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
+We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$.
+This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances.
+
+After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials?
+We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
+We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$.
+Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
 
@@
 
-Next, we test whether these improvements in performance are in fact due to more informative speaker behavior.
-We recruited twenty judges on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects.
-Their responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.6$ (95\% CI = $[0.54, 0.66]$).
-In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig.
-\ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$.
-In Expt.~1, the scripted label fit the hidden distractor just as well or better than the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
-In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.
-
-Does differential informativity of scripted utterances account for the variability across items that we noted in Expt.~1?
-In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
-We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$.
-This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances.
-
-After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials?
-We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
-We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$.
-Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
 % \begin{figure}[t!]
@@
 % \end{figure}
-
-%\begin{itemize}
-%\item people still try to move occluded objects, but they do it less, and that behavior is more confined to a few particularly difficult items.
-%\item this happens because the scripted instructions we forced directors to use in Exp.~1 were unnatural and violated listener expectations of overinformativity
-%\item the fact that the effect is attenuated in exp.
-2 is actually evidence of some theory of mind use (insofar as the director choosing to be overinformative, and the matcher expecting overinformativity constitutes ToM
-%\end{itemize}
 
@@
 
-%\subsection{Discussion}
-%
-%In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+\subsection{Discussion}
+
+In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it.
@@
 Both of these results support Keysar's claim that adults do not reliably deploy theory of mind reasoning when it is called for.
-%
-%Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data. First, 
-%
-%our analysis of item-level differences shows that the dependent variable highlighted in the original study (i.e.
-``percentage of participants who moved the critical item at least once'') is somewhat problematic.
-It could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item.
-Indeed, if we exclude the three `hard' items where over 60\% of participants in the experimental condition made errors, this dependent variable drops from 93\% to only 43\% of participants.
-This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.
-
-%Second, because half of the instructions were unscripted, we were able to observe natural production of referential descriptions for a number of fillers.
+
+Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data. First,
+
+Second, because half of the instructions were unscripted, we were able to observe natural production of referential descriptions for a number of fillers.
 Informally, we noted a tendency toward over-informativity.
@@
 
-%  \begin{figure}[b!]
-% \begin{center}
-% \vspace{-.5cm}
-% \includegraphics[scale=.5]{images/mousetracking.pdf}
-% \caption{Mean percentage of time spent hovering over target cell in the two conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are boostrapped 95\% confidence intervals.}
-% \vspace{-.5cm}
-% \label{fig:hover}
-% \end{center}
-% \end{figure}
+  \begin{figure}[t!]
+ \begin{center}
+ \vspace{-.5cm}
+ \includegraphics[scale=.3]{figures/Exp2_design}
+ \caption{Within-pair factorial design used in Exp.~2.
+In the common condition, we reproduce a simple pragmatic reference contrast: when more than one star is present in common ground, speakers are expected to be more informative.
+On hidden versions of these condition, the same set of objects is visible, but some additional squares are occluded.
+If pragmatic directors represent uncertainty over what the matcher sees behind these squares, we predict that they will produce more informative utterances.}
+ \vspace{-.5cm}
+ \label{fig:hover}
+ \end{center}
+ \end{figure}
+
+\subsection{Experiment~2: Production in uncertain contexts}
+\label{sec:Exp2}
+
+\subsection{Participants}
+
+We recruited K participants (K/2 pairs) from Amazon Mechanical Turk. All participants were native English speakers. %Three pairs were excluded for making 2 or more errors on non-critical items.
+
+\subsection{Materials \& Procedure}
+
+To more directly test speaker behavior across different contexts, we moved to a more systematic stimulus domain.
+These objects varied along three discrete features: \emph{shape}, \emph{texture}, and \emph{color}.
+Each feature took four discrete values, giving a space of 64 possible objects.
+We placed these objects in a smaller $3 \times 3$ grid, and used a pure \emph{reference game} setup where listeners had to click the intended referent as opposed to following a movement instruction.
+
+To test how speakers flexibly modify their informativity in response to the uncertainty posed by
 
@@
 On the other hand, it is known that textual communication, as in our chat box, can differ from face-to-face verbal communication.
-Additionally, aspects of the interface such as the graphical representation of occluded cells may be less intuitive, or require more training, on the web than in the lab.
+Additionally, we omitted the procedural step of placing the hidden distractor in a paper bag, and some aspects of the interface such as the graphical representation of occluded cells may be less intuitive, or require more training, on the web than in the lab.
 % Directors watched objects move through the actual moment-by-moment mouse trajectories that matchers used to drag them, and matchers saw the actual timing of the messages sent by the director.

### 2018-03-21 writing/journal_manuscript/manuscript.tex
Revise abstract; streamline intro

@@
 When they are unsure exactly which objects are in view of their partner, for example, they may provide additional disambiguating information when referring to a particular object: saying "cassette tape" instead of the potentially ambiguous ``tape."
-Recent debates over theory of mind use have hinged on behavior in natural-language communication tasks but have largely neglected the nuanced pragmatics of these tasks.
-Under Gricean expectations of cooperativity, speakers ought to be more informative in contexts with high uncertainty about what is in the listener's field of view.
-If pragmatic reasoning is truly grounded in theory of mind,   A careful analysis of pragmatic demands using a rational model of communication suggests a minimal pair of experiments which show that apparent failures of visual perspective-taking are in fact successes for sophisticated social reasoning in context.
+Recent debates over adult theory of mind use have hinged on behavior in natural-language communication tasks but have largely neglected the nuanced pragmatics of these tasks.
+In particular, using occluded objects to create perspective asymmetries may have unexpected side effects under pragmatic reasoning.
+Cooperative speakers ought to flexibly adjust their utterances to be more informative in contexts with high uncertainty about what is in the listener's field of view.
+In a direct replication, we show that scripted confederates violate these expectations by using less informative labels than unscripted directors naturally use.
+In two additional experiments, we directly manipulate speaker and listener expectations to test each step of the pragmatic account.
+Hence, apparent failures of visual perspective-taking may in fact be successes for sophisticated theory of mind use in context.
 %Recent theories of language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication.
@@
 %%% Point out a gaping hole that could benefit from this approach
-Our success as a cooperative species depends upon our ability to make sense of other agents. 
-One solution is to maintain an internal predictive model of other agents: how unobservable beliefs and desires give rise to observable actions. 
-Such a model allows us to make inferences about a social partner's likely internal state given their past behavior and form sophisticated expectations about their future behavior. 
+%Our success as a cooperative species depends upon our ability to make sense of other agents. 
+%One solution is to maintain an internal predictive model of other agents: how unobservable beliefs and desires give rise to observable actions. 
+%Such a model allows us to make inferences about a social partner's likely internal state given their past behavior and form sophisticated expectations about their future behavior. 
 %This theoretical approach to social cognition, where theory of mind reasoning \cite{PremackWoodruff78_ChimpanzeeToM, GopnikWellman92_TheoryTheory, WellmanCrossWatson01_ToMMetaAnalysis, SetohScottBaillargeon16_FalseBelief, ScottBaillargeon17_EarlyToM} is grounded in an explicit computational model of how other minds work, has
-% The approach has been useful for understanding a number of experimental phenomena in both adults and children \cite{KileyHamlinEtAl13_MentalisticCoreSocial, PantelisEtAl14_IntentionInference, BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}.
-Yet our current understanding of theory of mind use in \emph{communication} remains limited. 
-Just as making sense of an agent's physical movements in the world requires an accurate model of how beliefs and intentions translate into motor plans, making sense of language depends on an accurate model of what a speaker would say, or what a listener would think, in different situations.
+% The approach has been useful for understanding a number of experimental phenomena in both adults and children \cite{KileyHamlinEtAl13_MentalisticCoreSocial, PantelisEtAl14_IntentionInference}. 
+%Yet our current understanding of theory of mind use in \emph{communication} remains limited.
+
+Our ability to successfully communicate and coordinate with social partners relies on some degree of contextual flexibility in production and interpretation.
+It may be technically correct to refer to a particular entity as a ``dog,'' but speakers tend to choose more specific or \emph{informative} expressions when trying to point out their dog at a busy dog park.
+Similarly, a hyperbolic statement like ``I waited a million years for a table'' tends to be interpreted as an expression of frustration at longer-than-expected wait times rather than a literal statement about the time elapsed \citep{KaoWuBergenGoodman14_NonliteralNumberWords}.
+%A doctor who uses a single Latinate term to tell their colleague about a condition may use many more words to tell a patient the same thing.
+% Our success as a cooperative species depends upon our ability to make sense of other agents.
+This systematic contextual variation could in principle be accommodated by internally caching a complex set of different meanings, but pragmatic accounts have argued for a more parsimonious explanation.
+Just as making sense of an agent's physical movements in the world requires an accurate model of how beliefs and intentions translate into motor plans \citep{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}, making sense of language depends on an accurate mental model of what a speaker would say, or what a listener would think, in different situations.
+In other words, flexible communication relies on a generic \emph{theory of mind} mechanism.
 
@@
 Past efforts to evaluate theory of mind in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \citep{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
-We argue in this paper that the mental models supporting theory of mind use in communication are much richer than previously explored. 
-In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \citep{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
-For example, suppose a friend wants to refer to a nearby Dalmatian. 
+We argue in this paper that the mental models supporting theory of mind use in communication are richer than previously explored. 
+In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \citep{FrankeJager16_ProbabilisticPragmatics, Grice75_LogicConversation,Clark96_UsingLanguage,GoodmanFrank16_RSATiCS, FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
+%For example, suppose a friend wants to refer to a nearby Dalmatian. 
 %There are numerous descriptions that would all be literally true, but different expressions are preferred in different contexts.
-If there are no other dogs nearby, we may expect them to use a parsimonious basic-level term like ``dog.''
-When there is potential for confusion, however, we expect them to consider our likely interpretation and introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \citep{BrennanClark96_ConceptualPactsConversation, VanDeemter16_ComputationalModelsOfReferring, GrafEtAl16_BasicLevel}.
-When speakers violate these expectations, listeners draw inferences about the intended meaning in order to account for the violation: for example, XXX \cite{XXX, BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}.
-
-This intricate dance of expectations has been formalized in a probabilistic model of communication as recursive social reasoning.
+%If there are no other dogs nearby, we may expect them to use a parsimonious basic-level term like ``dog.''
+%When there is potential for confusion, however, we expect them to consider our likely interpretation and introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \citep{BrennanClark96_ConceptualPactsConversation, VanDeemter16_ComputationalModelsOfReferring, GrafEtAl16_BasicLevel}.
+%When speakers violate these expectations, listeners draw inferences about the intended meaning in order to account for the violation: for example, XXX \cite{XXX, BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}.
+
+%This intricate dance of expectations has been formalized in a probabilistic model of communication as recursive social reasoning.
 The Rational Speech Act (RSA) framework has successfully described many puzzling communicative phenomena, including vagueness \cite{Lassiter}, hyperbole \cite{Kao}, irony \cite{}, metaphor, and apparent over-informativeness \cite{Judith}.
@@
 
-Theory of mind mechanisms are therefore at the heart of the Rational Speech Act framework, yet they have never been directly compared against the predictions of egocentric or ``mindblind'' theories under the same behavioral paradigm.
-Since positive claims about theory of mind in communication concern the mental model we hold of our partner, these claims are best evaluated in a computational framework that formalizes the content of our representations and how we use them.
-By rigorously comparing the predicted behavior of speakers and listeners using different mental models, we show that theory of mind is required to account for context-sensitive behavior. 
-In addition, we replicate effects that, under more naive behavioral measures, were interpreted as evidence of limits on theory of mind use, but under a richer computational model are shown to be rational consequences of an uncooperative confederate.
-Especially in asymmetric communicative environments where speakers are explicitly put in a state of uncertainty over which objects their partner is seeing, evaluating the extent to which theory of mind is deployed in communication requires flexible mental models that represent the many sources of social expectations.
+%Theory of mind mechanisms are the foundation of pragmatic accounts, especially as recently formalized in recursive reasoning models \cite{Grice75_LogicConversation,}, yet they have never been directly compared against the predictions of egocentric or ``mindblind'' theories under the same behavioral paradigm.
+%Since positive claims about theory of mind in communication concern the mental model we hold of our partner, these claims are best evaluated in a computational framework that formalizes the content of our representations and how we use them.
+%By rigorously comparing the predicted behavior of speakers and listeners using different mental models, we show that theory of mind is required to account for context-sensitive behavior. 
+%In addition, we replicate effects that, under more naive behavioral measures, were interpreted as evidence of limits on theory of mind use, but under a richer computational model are shown to be rational consequences of an uncooperative confederate.
+%Especially in asymmetric communicative environments where speakers are explicitly put in a state of uncertainty over which objects their partner is seeing, evaluating the extent to which theory of mind is deployed in communication requires flexible mental models that represent the many sources of social expectations.
 % While egocentric views of perspective-taking make more clear-cut behavioral predictions, the predictions derived from a computational theory of mind are more varied and complicated but ultimately more accurate.
@@
 % Discuss other criticisms of Keysar paradigm?
-While these results are compelling, the paradigm has been criticized from few different angles. 
-\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
+%While these results are compelling, the paradigm has been criticized from few different angles. 
+%\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
 the hidden roll of tape vs.
 the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well.
-We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
-Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
+%We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
+%Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
 
@@
 As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
-%We define the decision window as the span of time between  their instruction message and when they started moving an object.
+Mouse-position is highly correlated with gaze in web browsing \citep{ChenEtAl01_MouseCursor, SpiveyGrosjeanKnoblich05} and has been used  %We define the decision window as the span of time between  their instruction message and when they started moving an object.
 If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt.
@@
 %consider the speaker's communicative goals and context. 
-In our replication (Expt.~1), we found evidence that listeners neglect the speaker's epistemic state, as Keysar and colleagues claim.
+In our replication, we found evidence that listeners neglect the speaker's epistemic state, as Keysar and colleagues claim.
 We also found that when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors. 
@@
 
-Further work in a wider variety of tasks is necessary to pin down the various factors determining the relative weighting of these different sources of information, but we have argued that pragmatics play a crucial role.
-
-Previous theoretical advances in modeling pragmatics have typically hinged on `lifting' inference over some aspect of communication that was previously assumed to be fixed.
+%Further work in a wider variety of tasks is necessary to pin down the various factors determining the relative weighting of these different sources of information, but we have argued that pragmatics play a crucial role.
+
+%Previous theoretical advances in modeling pragmatics have typically hinged on `lifting' inference over some aspect of communication that was previously assumed to be fixed.
 For instance, \emph{manner implicatures} emerge from uncertainty over the true meanings of words in the lexicon \cite{BergenLevyGoodman16_LexicalUncertainty}, nonliteral constructions like \emph{hyperbole} and \emph{irony} emerge from uncertainty over the topic of discussion \cite{KaoWuBergenGoodman14_NonliteralNumberWords}, and \emph{prosody} and \emph{ellipsis} emerge from uncertainty (in the form of a noisy-channel model) over the various ways that a speaker's utterance could have been corrupted  \cite{BergenGoodman15_StrategicUseOfNoise}.
@@
 
-\appendix
-
-\section{Sample Appendix Section}
-We derive the result in Eq. \eqref{eq:analytical_linear}. We consider a linear $C_L(\tau_L+\taupav)=K_L(\tau_L+\taupav)$, and
-make two further 
-simplifications: (i) the subject does not
-engage in leisure in the pre-reward state (and so works for the whole
-price when it works); and (ii) \emph{a priori}, arbitrarily long leisure durations are possible
-($\lambda=0$).
-Then the reward rate in Eq. \eqref{eq:rhoCHT} becomes
-\begin{equation}\label{eq:analyticalrho}
-\rho^{\pi}= \frac{\vrule height 10pt width0pt RI + K_L \{~ \mathbb{E}[ \tau_L | \textrm{post} ] + \taupav  \} }
-     {P +
-\mathbb{E}[ \tau_L | \textrm{post} ] +\taupav  } 
-\end{equation} 
-As discussed in the \emph{Results} section, the probability of engaging in instrumental leisure in the post-reward state is $\pi([L,\tau_L]
-~| \textrm{post}) = \exp\left[-\{\beta (\rho^\pi-K_L) 
- \} \tau_L\right]$, which is an exponential distribution with
-mean 
-\begin{equation}
-\mathbb{E}[\tau_L | \textrm{post}]=\frac{1}{\beta (\rho^\pi-K_L) }
-\label{eq:analyticaltauL}
-\end{equation}  
-Re-arranging terms of this equation,
-\begin{equation}\label{eq:rhoversion2}
-\rho^{\pi}=\frac{1}{\beta ~\mathbb{E}[\tau_L | \textrm{post}]} +K_L 
-\end{equation} 
-Equating Eqs. \eqref{eq:analyticalrho} and \eqref{eq:rhoversion2} and solving for the mean instrumental leisure duration $\mathbb{E}[\tau_L | \textrm{post}]$, we derive
-\begin{equation}
-\mathbb{E}[\tau_L | \textrm{post}] = \frac{P+\taupav}{\beta ( RI - K_LP)-1} 
-\label{eq:solvedtauL}
-\end{equation}
-which is the second line of Eq.\eqref{eq:analytical_linear}.
+%\appendix
+
+%\section{Sample Appendix Section}
+%We derive the result in Eq. \eqref{eq:analytical_linear}. We consider a linear $C_L(\tau_L+\taupav)=K_L(\tau_L+\taupav)$, and
+%make two further 
+%simplifications: (i) the subject does not
+%engage in leisure in the pre-reward state (and so works for the whole
+%price when it works); and (ii) \emph{a priori}, arbitrarily long leisure durations are possible
+%($\lambda=0$).
+%Then the reward rate in Eq. \eqref{eq:rhoCHT} becomes
+%\begin{equation}\label{eq:analyticalrho}
+%\rho^{\pi}= \frac{\vrule height 10pt width0pt RI + K_L \{~ \mathbb{E}[ \tau_L | \textrm{post} ] + \taupav  \} }
+%     {P +
+%\mathbb{E}[ \tau_L | \textrm{post} ] +\taupav  } 
+%\end{equation} 
+%As discussed in the \emph{Results} section, the probability of engaging in instrumental leisure in the post-reward state is $\pi([L,\tau_L]
+%~| \textrm{post}) = \exp\left[-\{\beta (\rho^\pi-K_L) 
+% \} \tau_L\right]$, which is an exponential distribution with
+%mean 
+%\begin{equation}
+%\mathbb{E}[\tau_L | \textrm{post}]=\frac{1}{\beta (\rho^\pi-K_L) }
+%\label{eq:analyticaltauL}
+%\end{equation}  
+%Re-arranging terms of this equation,
+%\begin{equation}\label{eq:rhoversion2}
+%\rho^{\pi}=\frac{1}{\beta ~\mathbb{E}[\tau_L | \textrm{post}]} +K_L 
+%\end{equation} 
+%Equating Eqs. \eqref{eq:analyticalrho} and \eqref{eq:rhoversion2} and solving for the mean instrumental leisure duration $\mathbb{E}[\tau_L | \textrm{post}]$, we derive
+%\begin{equation}
+%\mathbb{E}[\tau_L | \textrm{post}] = \frac{P+\taupav}{\beta ( RI - K_LP)-1} 
+%\label{eq:solvedtauL}
+%\end{equation}
+%which is the second line of Eq.\eqref{eq:analytical_linear}.
 This is the mean instrumental leisure duration as long as  $RI - K_LP>1$, and  $\mathbb{E}[\tau_L | \textrm{post}] \rightarrow \infty$ otherwise.
 When the former condition holds, we may
-substitute Eq. \eqref{eq:solvedtauL} into Eq. \eqref{eq:analyticalrho} and solve for $\rho^{\pi}$
-
-\begin{figure}
-\widefigure{\fullpagewidth}{Fig1}
-\caption{Sample Appendix Caption. Here is a caption that might
-appear in an appendix. It is as wide as the full width of the page.}
-\end{figure}
-
-\newpage
-
-\section{Making Your Bibliography for an Open Mind Article}
-{\it Open Mind} uses the APA author-date  bibliography style,
-apacite.bst. For more
-information on apacite, for examples in how to make your .bib file and more, see:\\
-\href{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}
-{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}
-
-\noindent
-(In spite of the mention of apacite cite commands, please use only
-Natbib commands for in text citations, as shown above.)
-
-\subsection{BibTeX}
-You will need to use BibTeX to form your bibliography; typing in the
-references would be
-a huge and unpleasant task. Look at the openmindsample.bbl file and you'll see why
-typing in the bibitems would be difficult.
-
-For a good basic introduction to using BibTeX, see
-\href{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}
-{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}
-
-When you use BibTeX, the form of the bibliography will be correct. You
-don't need to supply a bibliography style, since that is built into
-the stjour.cls file.
+%substitute Eq. \eqref{eq:solvedtauL} into Eq. \eqref{eq:analyticalrho} and solve for $\rho^{\pi}$
+%
+%\begin{figure}
+%\widefigure{\fullpagewidth}{Fig1}
+%\caption{Sample Appendix Caption. Here is a caption that might
+%appear in an appendix. It is as wide as the full width of the page.}
+%\end{figure}
+
+%\newpage
+
+%\section{Making Your Bibliography for an Open Mind Article}
+%{\it Open Mind} uses the APA author-date  bibliography style,
+%apacite.bst. For more
+%information on apacite, for examples in how to make your .bib file and more, see:\\
+%\href{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}
+%{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}
+%
+%\noindent
+%(In spite of the mention of apacite cite commands, please use only
+%Natbib commands for in text citations, as shown above.)
+%
+%\subsection{BibTeX}
+%You will need to use BibTeX to form your bibliography; typing in the
+%references would be
+%a huge and unpleasant task. Look at the openmindsample.bbl file and you'll see why
+%typing in the bibitems would be difficult. 
+%
+%For a good basic introduction to using BibTeX, see
+%\href{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}
+%{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}
+%
+%When you use BibTeX, the form of the bibliography will be correct. You
+%don't need to supply a bibliography style, since that is built into
+%the stjour.cls file.
 

### 2018-03-22 writing/journal_manuscript/manuscript.tex
fiddle with wording

@@
 
-This suggests that we consider whether the utterances produced by the confederate in Keysar's critical conditions were actually what a speaker in that context would be expected to say. 
+This suggests that we consider whether the scripted utterances produced by  confederates in popular perspective-taking tasks were actually what a speaker in that context would be expected to say \citep{KuhlenBrennan13_LanguageInDialogue}.
 If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
-More precisely, when both players know that objects are occluded in the display, the speaker may tend to add additional precision to references in order to avoid confusion.
-If the listener \emph{expects} the speaker to do this, they will pragmatically pick the \emph{a priori} more likely referent of the referring expression, which in critical trials will be the occluded object.
+More precisely, when both players know that objects are occluded in the display, speakers add additional precision to references in order to avoid confusion. 
+If the listener justifiably expects the speaker to do so, they may \emph{rationally} adopt a strategy of choosing the \emph{a priori} more likely referent of the referring expression regardless of perspective.
 In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item. 
@@
 
-In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
+%In particular, we argue that critical utterances used by Keysar and colleagues are \emph{uncooperative}: they are less informative than what a speaker would actually produce in that situation; listeners who are sensitive to the pragmatics of the situation expect these more informative utterances and produce ``errors'' when their expectations are flouted.
 In Experiment 1, we conduct a direct replication of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, manipulating whether the speaker is a confederate or naive participant.
@@
 In Experiment 2, we test the hypothesis that speakers systematically produce longer, more specific utterances than strictly required by their visible context when they know objects are occluded.
-Finally, in Experiment 3, we manipulate the listener's pragmatic beliefs about whether the speaker is indeed expected to produce such `over-informative' utterances.
-Taken together, these results suggest that apparent failures of perspective-taking are in fact driven by sophisticated pragmatic reasoning; in other words, by successful theory of mind use.
+%Finally, in Experiment 3, we manipulate the listener's pragmatic beliefs about whether the speaker is indeed expected to produce such `over-informative' utterances.
+Taken together, these results suggest that apparent failures of perspective-taking are in fact attributable to sophisticated pragmatic expectations about speaker behavior---that is, to successful theory of mind use.
+
+ \begin{figure}[t!]
+ \begin{center}
+ \vspace{-.25cm}
+ \includegraphics[scale=.5]{figures/Exp1_design.png}
+ \vspace{-.5cm}
+ \caption{Screenshots of interface on a critical trial: a cassette tape is in view of both players, but a roll of tape is occluded from the speaker's view.}
+ \vspace{-.5cm}
+ \label{fig:exp1paradigm}
+ \end{center}
+ \end{figure}
 
@@
 Participants were randomly assigned to `director' and `matcher' roles and placed in a shared environment containing a chat box and a $4 \times 4$ grid of objects (see Fig.
-\ref{fig:interface}).
-Five of the grid cells were blocked from the director's view by a curtain, and six or seven objects were displayed in the grid at a given time.
+\ref{fig:exp1paradigm}).
 Trials were blocked into eight `items', each using a different set of objects.
-For each item, we gave the director a series of four instructions to move objects, which were displayed as a series of arrows pointing from some object to a neighboring unoccupied cell.
+For each item, we gave the director a series of four instructions, which were displayed as a series of arrows pointing from some object to a neighboring unoccupied cell.
 They used a chat box to communicate the identity of this object and action to their partner, who proceeded to click and drag an object to complete the trial.
@@
 We used a between-subject design to test whether the scripted confederate labels were in fact what cooperative speakers would say in context.
-In the `scripted' condition, directors were free to communicate however they wished for half of the instructions but for the other half, including all the critical instructions, their messages were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+Directors assigned to the `scripted' condition were free to communicate however they wished for half of the instructions.
+For the other half, including all the critical instructions, their messages were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
 For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat box, and they would have to click `send' for the experiment to continue.) In the `unscripted' condition, directors were unrestricted and free to send whatever messages they deemed appropriate on all trials.
@@
 The assignment of items to conditions was randomized across participants, and the order of conditions was randomized under the constraint that the same condition would not be used on more than two consecutive items.
-All object sets, object placements, and corresponding instruction sets were the same for all participants.
+Five of the grid cells were blocked from the director's view by a curtain, and six or seven objects were displayed in the grid at a given time.
+All object sets, object placements, and corresponding instruction sets were fixed across participants.
+
+In addition to collecting messages sent through the chat box and matcher errors, we collected mouse-tracking data in analogy to the eye-tracking common in these paradigms.
+To collect clean data, we asked the matcher to wait until the director sent a message; when the message was received, the matcher clicked a small circle in the center of the grid to show the objects and proceed with the trial.
+We recorded at 100Hz from the matcher's mouse in the decision window after this click.
 
@@
 %Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.
-
-To collect clean mouse-tracking data, we asked the matcher to wait until the director sent a message; when the message was received, the matcher clicked a small circle in the center of the grid to continue the trial.
-We started recording from the matcher's mouse after this click.
 
@@
 If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt.
-Within the decision window, we computed the total proportion of time spent moving toward the target relative to the distractor.
+% Within the decision window, we computed the total proportion of time spent moving toward the target relative to the distractor.
 %We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly.
@@
 Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions.
-If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells).
-That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
-We expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
+%If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells).
+%That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
+We therefore expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
 
@@
 
-Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data. First,
+%Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data. First,
 
@@
  \vspace{-.5cm}
- \includegraphics[scale=.3]{figures/Exp2_design}
+ \includegraphics[scale=.25]{figures/Exp2_design}
  \caption{Within-pair factorial design used in Exp.~2.
@@
 
-\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$.
+%\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$.
 The listener, however, also has some uncertainty over whether the experimenter is deceptive, represented by a Bernoulli random variable $d$, such that they assign a small prior probability $P(d)$ to the possibility that $o_h$ is actually in common ground.
 Upon hearing an utterance$u$, they update both their beliefs about which object is the target and whether the experimenter is being deceptive:
-$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
+%$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
 \end{enumerate}

### 2018-03-26 writing/journal_manuscript/manuscript.tex
Minor revisions

@@
 
-%consider the speaker's communicative goals and context. 
-In our replication, we found evidence that listeners neglect the speaker's epistemic state, as Keysar and colleagues claim.
-We also found that when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors. 
+% 
+In our replication, we found empirical support for the claim that listeners neglect the speaker's epistemic state in some contexts, as claimed by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}: when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors.
 Indeed, we found that the extremely heterogeneous pattern of errors across items was well explained by the relative fit of the utterance to target and distractor.
-This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness rather than consideration of occluded vs. mutual visible items. 
-%
-However, this does not necessarily imply limitations on theory of mind---if speakers could be expected to naturally provide expressions which apply better to the target, then this behavior would be appropriate.
+This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness rather than consideration of occluded vs. mutually visible items.
+
+However, this finding does \emph{not} necessarily imply limitations on theory of mind.
+Speakers systematically provide more informative utterances in contexts where views are uncertain, hence listeners with well-calibrated social expectations are justified in neglecting perspective.
+Rather than debating provide a principled, rational explanation of \emph{why} listeners may neglect speaker perspective in the first place.
+These reasons stem from more sophisticated pragmatic expectations than simple consideration of perspective, thereby opening a deeper avenue by which theory of mind use guides flexible communication.
+% if speakers could be expected to naturally provide expressions which apply better to the target, then this behavior would be appropriate.
 % If listeners arrive at this interpretation strategy through pragmatic reasoning, then neglecting the speaker's epistemic state would be rational and also the result of theory of mind use.
-In Expt.~2, we found that speakers did naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor.
-For example, no speaker in Expt.~2 produced ``the bottom block,'' which was used as a critical instruction in Expt.~1.
+
+%We found that speakers did naturally produce more precise, informative utterances than required and these unconstrained utterances fit the target significantly better than they fit the hidden distractor.
+%For example, no speaker in Expt.~2 produced ``the bottom block,'' which was used as a critical instruction in Expt.~1.
 Instead, they said ``the bluish block with a B'' or ``block with the blue writing,'' which relied on less confusable perceptual features and thereby decreased error rates.
-Thus, the errors observed in Expt.~1 can be explained as a result of an uncooperative confederate (speaker)---the experiment set up certain pragmatic expectations through the task context, then deliberately flouted them in critical trials.
+%Thus, the errors observed in Expt.~1 can be explained as a result of an uncooperative confederate (speaker)---the experiment set up certain pragmatic expectations through the task context, then deliberately flouted them in critical trials.
+
+%%%%%%%
+% Discussion points:
+
+% 1.
+Tying back to original expt (e.g.
+pitch Expt.
+3 idea).
+Listeners didn't know what's going on when they heard the dramatically under-informative `tape' -- one direction for future work is to directly manipulate expectations about speakers...
+
+% 2. Other reasons for over-informativity (e.g. Judith stuff)
+
+% 3. Relationship to Heller model, Rubio-Fernandez, other critiques of Keysar?
+
+% 4. 
+%%%%%%%
 
 The Gricean maxim of quality dictates that cooperative speakers should make their utterance as informative as is required for current purposes, but no more so.
-If the referring expressions used in Expt.~1 uniquely pick out the target from the speaker's perspective, then why would unscripted speakers in Expt.~2 be more informative than this?
+If the referring expressions used in by confederates uniquely pick out the target from the speaker's perspective, then why our unscripted speakers be more informative than this?
 %strictly necessary and why would listeners come to expect such ``overinformativity''? 
-One possibility is that awareness of the complex mismatch in epistemic state with the listener leads the speaker to provide extra information in an attempt to avoid miscommunication.
+One possibility, which we tested in Exp.
+2, is that awareness of the complex mismatch in epistemic state with the listener leads the speaker to provide extra information in an attempt to avoid miscommunication.
 Another possibility is that more general dynamics of language are in play.

### 2018-03-30 analysis/final_analysis.Rmd
Preliminary analyses of new data

@@
 ```{r}
-d_raw <- read_tsv('../data/final/experiment2/pilot0/chatMessage/pilot0_messages.csv')
+library(tidyverse)
+library(ggthemes)
+library(lme4)
+library(lmerTest)
+
+d_raw <- read_tsv('../data/final/experiment2/planned_sample/chatMessage/message.csv') %>%
+  filter(iterationName == 'planned_sample') %>%
+  filter(role == 'speaker')
 
 # nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
-incompleteIDs <- unique((d %>% group_by(gameid) %>% 
+incompleteIDs <- unique((d_raw %>% group_by(gameid) %>% 
                            filter(length(unique(trialNum)) != 24))$gameid)
@@
 ```{r}
-d %>%
+basic.d <- d %>%
   filter(hidden == 'no') %>%
@@
   spread(context, m) %>%
-  mutate(diff = close - far) %>%
-  ggplot(aes(x = far, y = close)) +
+  mutate(diff = close - far)
+
+ggplot(basic.d, aes(x = far, y = close)) +
     geom_point() +
@@
     theme(aspect.ratio = 1)
+```
 
-  summarize(m = mean(numRawWords), se = sd(numRawWords)/sqrt(length(numRawWords))) %>%
-    spread(context, numRawWords)
+Aggregated to difference scores:
 
+```{r}
+ggplot(basic.d, aes(x = diff)) +
+    geom_histogram(binwidth = 0.25) +
+    theme_bw() +
+    geom_vline(xintercept = 0) +
+    # ylim(1,12) +
+    xlim(-4,4) +
+    theme(aspect.ratio = 1)
+
+t.test(basic.d$diff)
 ```
@@
 ```{r}
-d %>%
+hidden.d <- d %>%
   group_by(gameid, context, hidden) %>%
@@
   spread(hidden, m) %>%
-  mutate(diff = yes - no) %>%
-  ggplot(aes(x = no, y = yes)) +
+  mutate(diff = yes - no)
+
+ggplot(hidden.d, aes(x = no, y = yes)) +
     geom_point() +
@@
     geom_abline(intercept = 0, slope = 1) +
-    ylim(1,12) +
+    ylim(1,15) +
     ylab('# words (with occlusions)') +
@@
     ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
-    xlim(1,12) +
+    xlim(1,15) +
     theme(aspect.ratio = 1)
@@
 
+Histogram
+
 ```{r}
-d %>%
-  group_by(gameid, context, hidden) %>%
-  summarize(m = mean(numRawWords)) %>%
-  spread(hidden, m) %>%
-  mutate(diff = yes - no) %>%
+ggplot(hidden.d %>% group_by(context) %>% mutate(m = mean(diff)), aes(x = diff)) +
+    geom_histogram(binwidth = 0.33) +
+    theme_bw() +
+    geom_vline(xintercept = 0, linetype = 2) +
+    geom_vline(aes(xintercept = m), color = 'red') +
+    facet_wrap(~ context) +
+    # ylim(1,12) +
+    xlim(-5,5) +
+    theme(aspect.ratio = 1)
+```
+
+```{r}
+hidden.d %>%
   group_by(context) %>%
@@
 
-# Regression
+# Stats
+
+Simple effect of hidden-ness in 'far' contexts
+
+```{r}
+summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
+             data = d %>% 
+                filter(context == 'far')))
+t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far'), var.equal = FALSE)
+```
+
+interaction-y way
+
+```{r}
+summary(lmer(numRawWords ~ context * hidden + (1 + context * hidden | gameid), data = d))
+```
 
@@
 ```{r}
-library(ggthemes)
-library(lme4)
-library(lmerTest)
-
 d.regression <- d %>% 
@@
 ```
-
-interaction-y way
-
-```{r}
-summary(lmer(numRawWords ~ context * hidden + (1 + context + hidden | gameid), data = d))
-```

### 2018-03-31 analysis/final_analysis.Rmd
Import subject info; implement exclusions

@@
 
+# General imports
+
+```{r}
+library(jsonlite)
+library(tidyverse)
+library(ggthemes)
+library(lme4)
+library(lmerTest)
+```
+
 # Experiment 2
+
+## Import turk survey data
+
+```{r}
+path <- "../data/final/experiment2/planned_sample/turk/"
+files = dir(path, pattern = "*.json")
+
+subjInfo = data.frame()
+
+for (f in files) {
+  jf = paste0(path, f)
+  jd = read_json(jf)
+
+  newRow = data.frame()
+  if(!is.null(jd$answers$subject_information)) {
+    # for(property in names(jd$answers$subject_information) {
+    #   if(!is.null(jd$answers$subject_information[property])) {
+    #     newRow[property] = jd$answers$subject_information[property]
+    #   } else 
+    #     newRow[property] = NA
+    # }
+    newRow = data.frame(gameid = jd$answers$subject_information$gameID,
+                        role = jd$answers$subject_information$role,
+                        gameLength = jd$answers$subject_information$totalLength,
+                        thinksHuman = ifelse(!is.null(jd$answers$subject_information$thinksHuman), jd$answers$subject_information$thinksHuman, NA),
+                        understandsInstructions = jd$answers$subject_information$confused,
+                        nativeEnglish = jd$answers$subject_information$nativeEnglish,
+                        ratePartner = jd$answers$subject_information$ratePartner,
+                        comments = jd$answers$subject_information$comments,
+                        strategy = jd$answers$subject_information$strategy)
+    subjInfo = bind_rows(subjInfo, newRow)
+  }
+}
+
+```
 
@@
 ```{r}
-library(tidyverse)
-library(ggthemes)
-library(lme4)
-library(lmerTest)
-
 d_raw <- read_tsv('../data/final/experiment2/planned_sample/chatMessage/message.csv') %>%
@@
                            filter(length(unique(trialNum)) != 24))$gameid)
+confused <- unique((subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
 
-# Some speakers relied purely on location in grid
-pureLocation <- c('0180-a541e54c-5dea-4d06-9a21-9b842316eb59', '0101-0b659560-f05a-471a-89ac-60469e9c867f')
-badGames <- c(incompleteIDs, pureLocation)
+# Some speakers relied on location in grid
+location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', '5684-f09de856-4d64-423c-b342-b30e7325eb0e', '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
+badGames <- c(incompleteIDs, pureLocation, confused)
 
@@
   spread(context, m) %>%
-  mutate(diff = close - far)
+  mutate(diff = close- far,
+         ratio = close / far)
 
@@
   spread(hidden, m) %>%
-  mutate(diff = yes - no)
+  mutate(diff = yes- no)
 
@@
     # ylim(1,12) +
-    xlim(-5,5) +
+    xlim(-6,6) +
     theme(aspect.ratio = 1)
+
+ggsave("~/Downloads/histograms.pdf")
 ```
@@
 
+```
+
+Clusters
+
+```{r}
+hidden.d %>% 
+  filter(context == 'far') %>% 
+  mutate(group = case_when(diff < -0 ~ 'anti',
+                           diff > 0 ~ 'pro',
+                           TRUE ~ 'all_the_things')) %>%
+  group_by(group) %>% tally()
 ```

### 2018-03-31 writing/journal_manuscript/manuscript.tex
Minor revisiosn to Exp. 2

@@
 In a direct replication, we show that scripted confederates violate these expectations by using less informative labels than unscripted directors naturally use.
-In two additional experiments, we directly manipulate speaker and listener expectations to test each step of the pragmatic account.
+In a follow-up experiment, we directly manipulate whether objects are hidden to test the pragmatic account of speaker informativity in uncertain environments.
 Hence, apparent failures of visual perspective-taking may in fact be successes for sophisticated theory of mind use in context.
@@
  \vspace{-.5cm}
- \label{fig:hover}
+ \label{fig:exp2design}
  \end{center}
@@
 
+In Experiment 2, we more directly test the hypothesis that speakers naturally produce more informative referring expressions in more uncertain contexts (e.g.
+contexts where objects are occluded such as those used in Director-Matcher tasks to test theory of mind use).
+
 \subsection{Participants}
 
-We recruited K participants (K/2 pairs) from Amazon Mechanical Turk. All participants were native English speakers. %Three pairs were excluded for making 2 or more errors on non-critical items.
+We recruited 200 participants (100 pairs) from Amazon Mechanical Turk. All participants were native English speakers.
 
@@
 
-To more directly test speaker behavior across different contexts, we moved to a more systematic stimulus domain.
-These objects varied along three discrete features: \emph{shape}, \emph{texture}, and \emph{color}.
+Stimuli were sampled from a more systematic domain than the idiosyncratic objects used in Exp.~1.
+They varied along three discrete features: \emph{shape}, \emph{texture}, and \emph{color}.
 Each feature took four discrete values, giving a space of 64 possible objects.
@@
 
+Our design is entirely within-pair.
+Each trial is sampled from a factorial 2 x 2 design manipulating (1) the presence or absence of occlusions (`hidden' vs.
+`no hidden'), and (2) the closeness of visible distractors to the target (`close' vs.
+`far'; see Fig.
+\ref{fig:exp2design}).
+On 'hidden' trials, two cells are covered in curtains such that only the matcher can see if there is an object in the cell; on 'no hidden' trials, all cells are in common ground.
+On 'far' trials, the target is the only object with a particular shape; on 'close' trials, there is also a visible distractor sharing the target's shape, differing only in color or texture.
+Each trial type appeared 6 times for a total of 24 trials, and the sequence of trials was pseudo-randomized such that no trial type appeared more than twice in each
+
  To test how speakers flexibly modify their informativity in response to the uncertainty posed by
+
+\subsection{Results}
+
+Our primary is the length (in words) of their naturally produced referring expressions in different contexts.
+
+As a secondary measure (intended for use in a cognitive model), we will also annotate each utterance by the number of object features mentioned (e.g. shape, texture, color).
 

### 2018-04-02 analysis/final_analysis.Rmd
Feature-based analyses

--- 2018-03-30 analysis/final_analysis.Rmd
@@
 
+# General imports
+
+```{r}
+library(jsonlite)
+library(tidyverse)
+library(ggthemes)
+library(lme4)
+library(lmerTest)
+```
+
 # Experiment 2
+
+## Import turk survey data
 
@@
 ```{r}
-library(tidyverse)
-library(ggthemes)
-library(lme4)
-library(lmerTest)
-
-d_raw <- read_tsv('../data/final/experiment2/planned_sample/chatMessage/message.csv') %>%
-  filter(iterationName == 'planned_sample') %>%
-  filter(role == 'speaker')
+d_annotated <- read_csv('../data/final/experiment2/planned_sample/chatMessage/messages_with_annotations.csv')
 
@@
                            filter(length(unique(trialNum)) != 24))$gameid)
+confused <- unique((subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
 
-# Some speakers relied purely on location in grid
-pureLocation <- c('0180-a541e54c-5dea-4d06-9a21-9b842316eb59', '0101-0b659560-f05a-471a-89ac-60469e9c867f')
-badGames <- c(incompleteIDs, pureLocation)
+# Some speakers relied on location in grid (TODO: WHAT ABOUT 'BELOW' PEOPLE?)
+location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
+                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
+                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
+                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
+# Some speakers played this weird taboo game where they gave riddles 
+tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
+              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
+              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
+badGames <- c(incompleteIDs, pureLocation, confused)
 
-d <- d_raw %>%
-  group_by(gameid, trialNum, occlusions, context) %>%
-  summarize(text = paste0(text, collapse = ' ')) %>%   # Concatenate multiple utterance on same trial
-  mutate(numRawWords = str_count(text, "\\S+")) %>%
-  mutate(hidden = ifelse(occlusions == 'none', 'no', 'yes')) %>%
+d <- d_annotated %>%
+  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
   filter(!(gameid %in% badGames))
@@
   spread(context, m) %>%
-  mutate(diff = close - far)
+  mutate(diff = close- far,
+         ratio = log(close / far))
 
@@
 ```{r}
-ggplot(basic.d, aes(x = diff)) +
+ggplot(basic.d, aes(x = ratio)) +
     geom_histogram(binwidth = 0.25) +
@@
 
-t.test(basic.d$diff)
+t.test(basic.d$ratio)
 ```
@@
   summarize(m = mean(numRawWords), se = sd(numRawWords)/sqrt(length(numRawWords))) %>%
-  ggplot(aes(x = hidden, y = m, fill = context)) +
+  ggplot(aes(x = context, y = m, fill = hidden)) +
     geom_bar(stat = 'identity', position = dodge) +
     geom_errorbar(aes(ymin = m - se, ymax = m + se), position = dodge, width = 0.1) +
-    theme_bw()
+    theme_few() +
+    ylab('mean # words') +
+    theme(aspect.ratio = .5)
+
+ggsave('~/Downloads/aggregated_means.pdf')
 ```
@@
 ## Scatter to look at within-subj effects
+
+Some people used exactly the same number of words on every trial...
+
+```{r}
+hidden.d %>% 
+  filter(context == 'far') %>% 
+  mutate(group = case_when(diff < -0 ~ 'anti',
+                           diff > 0 ~ 'pro',
+                           TRUE ~ 'all_the_things')) %>%
+  group_by(group) %>% tally()
+
+all_the_things_ers <- (hidden.d %>% filter(context =='far') %>% filter(diff == 0))$gameid
+```
 
@@
   spread(hidden, m) %>%
-  mutate(diff = yes - no)
+  mutate(diff = yes- no,
+         ratio = log(yes/no))
 
@@
 ```{r}
-ggplot(hidden.d %>% group_by(context) %>% mutate(m = mean(diff)), aes(x = diff)) +
+ggplot(hidden.d %>% filter(!(gameid %in% all_the_things_ers)) %>% group_by(context) %>% mutate(m = mean(diff)), aes(x = diff)) +
     geom_histogram(binwidth = 0.33) +
@@
              data = d %>% 
-                filter(context == 'far')))
-t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far'), var.equal = FALSE)
+               filter(context == 'far') %>%
+               filter(!(gameid %in% all_the_things_ers))
+               ))
+t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far') , var.equal = FALSE)
 ```
@@
 ```{r}
-summary(lmer(numRawWords ~ context * hidden + (1 + context * hidden | gameid), data = d))
+summary(lmer(numRawWords ~ context * hidden + (1 + context * hidden | gameid), data = d %>% filter(!(gameid %in% all_the_things_ers))))
 ```
@@
 ```
+
+## Use feature data instead
+
+```{r}
+d %>% 
+  group_by(context, occlusions) %>% 
+  summarize(color = mean(colorMention), shape = mean(shapeMention), texture = mean(textureMention))
+```
+
+Histograms of number of features mentioned in each condition
+
+```{r}
+ggplot(d, aes(x = numFeatures)) +
+  geom_histogram(binwidth = 1) +
+  facet_grid(context ~ occlusions) +
+  theme_few()
+```
+
+```{r}
+d %>% group_by(context, occlusions) %>% 
+  mutate(numTrials = length(text)) %>% 
+  group_by(context, occlusions, colorMention, textureMention) %>% 
+  summarize(n = length(text)/mean(numTrials)) %>%
+  ggplot(aes(x = colorMention, y = textureMention, fill = n)) +
+    geom_bin2d(stat = 'identity') +
+    stat_bin2d(geom = "text", aes(label = round(n, 2))) +
+    scale_fill_gradient(low = "white", high = "red") +
+    facet_grid(occlusions ~ context) +
+    theme_few()
+```
+
+Note: Random intercept doesn't converge (less variance in these numbers...)
+
+```{r}
+summary(lmer(numFeatures ~ context * occlusions + (1 + context + occlusions | gameid), data = d))
+```
+
+```{r}
+summary(glmer(colorMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
+```
+
+```{r}
+summary(glmer(textureMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
+```
+
+```{r}
+d %>% 
+  group_by(context, occlusions) %>% 
+  summarize(numFeatures = mean(numFeatures))
+```

### 2018-04-02 data/final/experiment2/planned_sample/chatMessage/preprocessMessages.R
Fix cogsci exp. file structure; preprocess messages

@@ -1,0 +1,31 @@
+library(tidyverse)
+library(stringi)
+library(quanteda)
+
+rawD <- read_tsv('./raw_message.csv')
+
+mentionedOutlineColor <- function(m) {
+  colors <- c("blue","green", "black", "red")
+  words <- tolower(stri_extract_all_words(m, simplify = TRUE))
+  return(length(intersect(colors, words)) > 0)
+}
+
+mentionedShape <- function(m) {
+  shapes <- c("star","square", "circle", "triangle")
+  words <- tolower(stri_extract_all_words(m, simplify = TRUE))
+  return(length(intersect(shapes, words)) > 0)
+}
+
+
+d <- rawD %>%
+  filter(role == 'speaker') %>%
+  group_by(gameid, trialNum, context, occlusions) %>%
+  summarize(text = paste0(text, collapse = ' ')) %>%   # Concatenate multiple utterance on same trial
+  mutate(numRawWords = str_count(text, "\\S+")) %>%
+  mutate(hidden = ifelse(occlusions == 'none', 'no', 'yes')) %>%
+  rowwise() %>%
+  mutate(colorMention = mentionedOutlineColor(text),
+         shapeMention = mentionedShape(text),
+         textureMention = '')
+
+write_csv(d, 'messages_no_annotations.csv')

### 2018-04-02 experiments/cogsci/experiment1/game.server.js

@@ -197,7 +197,7 @@ game_server.findGame = function(player) {
                 var d = new Date();
                 var start_time = d.getFullYear() + '-' + d.getMonth() + 1 + '-' + d.getDate() + '-' + d.getHours() + '-' + d.getMinutes() + '-' + d.getSeconds() + '-' + d.getMilliseconds()
               var name = start_time + '_' + game.id;
-	      var prefix = "../../../data/experiment1/";
+	      var prefix = "../../../data/cogsci/raw/experiment1/";
                 var mouse_f = prefix + "mouse/" + name + ".csv"
                 fs.writeFile(mouse_f, "gameid, time, condition, critical, objectSet, instructionNum, attemptNum, targetX, targetY, distractorX, distractorY, mouseX, mouseY\n", function (err) {if(err) throw err;})
                 game.gamecore.mouseDataStream = fs.createWriteStream(mouse_f, {'flags' : 'a'});

### 2018-04-03 experiments/final/experiment1/game.client.js
Fix end game message bug

@@ -56,6 +56,7 @@ function client_onserverupdate_received(data){
   globalGame.scriptedInstruction = data.scriptedInstruction;
   globalGame.attemptNum = data.attemptNum;
 
+  globalGame.roundNum = data.roundNum;
   globalGame.data = data.dataObj;
   globalGame.game_started = data.gs;
   globalGame.players_threshold = data.pt;

### 2018-04-03 experiments/final/experiment1/game.core.js

@@ -303,6 +303,7 @@ game_core.prototype.server_send_update = function(){
     pt : this.players_threshold,
     pc : this.player_count,
     dataObj  : this.data,
+    roundNum : this.roundNum,
     scriptedInstruction : this.scriptedInstruction,
     attemptNum : this.attemptNum,
     curr_dest : this.currentDestination,

### 2018-04-03 experiments/final/experiment1/game.client.js
Add bonus display & progress display

@@ -47,7 +47,6 @@ function client_onserverupdate_received(data){
 
   if(globalGame.instructions.length > 0) {
     globalGame.currTarget = globalGame.instructions[globalGame.instructionNum].split(' ')[0];
-    console.log(globalGame.currTarget);
     globalGame.critical = _.find(globalGame.objects, obj => {
       return obj.name == globalGame.currTarget;
     })['critical'] == 'target';  
@@ -57,10 +56,17 @@ function client_onserverupdate_received(data){
   globalGame.attemptNum = data.attemptNum;
 
   globalGame.roundNum = data.roundNum;
-  globalGame.data = data.dataObj;
+  if(!_.has(globalGame, 'data')) {
+    globalGame.data = data.dataObj;
+  }
+
   globalGame.game_started = data.gs;
   globalGame.players_threshold = data.pt;
   globalGame.player_count = data.pc;
+
+  $('#roundnumber').empty()
+    .append("Round: " + (globalGame.roundNum + 1) + "/" + globalGame.numRounds +
+	    '\nInstruction: ' + (globalGame.instructionNum + 1) + '/4');
 
   // Draw all this new stuff
   drawScreen(globalGame, globalGame.get_player(globalGame.my_id));
@@ -114,6 +120,12 @@ client_onMessage = function(data) {
       var cell = globalGame.getPixelFromCell({gridX : commands[3], gridY : commands[4]});
       // var targetName = globalGame.instructions[globalGame.instructionNum].split(' ')[0];
       setTimeout(() => drawFeedbackIcon(globalGame, type, cell), 200);
+      var scoreDiff = (type == 'correct' ?
+		       Math.max(0, globalGame.bonusAmt - globalGame.attemptNum)  : 0);
+      globalGame.data.subject_information.score += scoreDiff;
+      $('#score').empty()
+        .append("Bonus: $" + (globalGame.data.subject_information.score/100).toFixed(2));
+
       break;
     }
   } 
@@ -130,6 +142,11 @@ var customSetup = function(game) {
   // For mouse-tracking, matcher must wait until director sends message  
   game.socket.on('newRoundUpdate', function(data){
     $('#messages').empty();
+    if(game.roundNum + 2 > game.numRounds) {
+      $('#roundnumber').empty();
+      $('#instructs').empty();
+    } 
+
     if(globalGame.my_role == globalGame.playerRoleNames.role2) {
       var msg = 'Waiting for your partner to send a message...';
       globalGame.get_player(globalGame.my_id).message = msg;

### 2018-04-03 experiments/final/experiment1/game.core.js

@@ -33,6 +33,7 @@ var game_core = function(options){
   this.iterationName = 'pilot1';
   this.dataStore = ['csv', 'mongo'];
   this.anonymizeCSV = true;
+  this.bonusAmt = 1; // in cents
 
   this.players_threshold = 2;
   this.playerRoleNames = {
@@ -74,10 +75,9 @@ var game_core = function(options){
 
     this.data = {
       id : this.id,
-      trials : [],
-      catch_trials : [], system : {}, 
       subject_information : {
-	gameID: this.id, scale: 0
+	score: 0,
+	gameID: this.id
       }
     };
     this.server_send_update();
@@ -170,8 +170,8 @@ game_core.prototype.setScriptAndDir = function(instruction) {
 };
 
 game_core.prototype.newInstruction = function() {
-  console.log('sending new instruction');
   this.instructionNum += 1;
+  console.log('sending new instruction ' + this.instructionNum);
   var instruction = this.instructions[this.instructionNum]
   this.currTarget = instruction.split(' ')[0]; 
   this.setScriptAndDir(instruction)

### 2018-04-03 experiments/final/experiment2/game.core.js
Check in earlier change to Exp. 2 iteration name

@@ -34,7 +34,7 @@ var game_core = function(options){
   this.email = 'rxdh@stanford.edu';
   this.projectName = 'ToM';
   this.experimentName = 'speakerManipulation';
-  this.iterationName = 'pilot0';
+  this.iterationName = 'planned_sample';
   this.anonymizeCSV = true;
   this.bonusAmt = 1; // in cents
   

### 2018-04-03 experiments/final/store.js
Fix duplicate blocking, and allow requester to bypass!

@@ -108,14 +108,12 @@ function serve() {
       const query = request.body.query;
       const projection = request.body.projection;
 
-      // hardcoded for now (TODO: get list of collections in db)
-      var collectionList = ['ToM']//'sketchpad_basic','sketchpad_repeated',
-			    //'chatbox_basic', //'chairs_chatbox',
-			    //'artificialLanguage']; 
+      // hardcoded for now (TODO: get list of collections in db?)
+      var collectionList = ['speakerManipulation', 'replication'];
 
       function checkCollectionForHits(collectionName, query, projection, callback) {
         const collection = database.collection(collectionName);        
-        collection.find(query, projection).limit(1).toArray((err, items) => {          
+        collection.find(query, projection).limit(1).toArray((err, items) => { 
           callback(!_.isEmpty(items));
         });  
       }
@@ -141,8 +139,12 @@ function serve() {
         response.json(hits>0);
       }
 
-      checkEach(collectionList, checkCollectionForHits, query, projection, evaluateTally);
-
+      // Always let the requester test ;) 
+      if(query.workerId == 'A1BOIDKD33QSDK') {
+	response.json(false);
+      } else {
+	checkEach(collectionList, checkCollectionForHits, query, projection, evaluateTally);
+      }
     });
 
     app.post('/db/insert', (request, response) => {

### 2018-04-03 experiments/final/sharedUtils/sharedUtils.js
Fix database being used for duplicate checking

@@ -26,7 +26,7 @@ var handleInvalidID = function(req, res) {
 var checkPreviousParticipant = function(workerId, callback) {
   var p = {'workerId': workerId};
   var postData = {
-    dbname: '3dObjects',
+    dbname: 'ToM',
     query: p,
     projection: {'_id': 1}
   };

### 2018-04-03 analysis/final_analysis.Rmd
Bootstrap feature analyses; add test mouse-tracking analysis

@@
 
+```{r}
+subjInfo <- read_csv('../data/final/experiment2/planned_sample/turk/subjInfo.csv')
+```
+
 ## Import message data
@@
 # nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
-incompleteIDs <- unique((d_raw %>% group_by(gameid) %>% 
+incompleteIDs <- unique((d_annotated %>% group_by(gameid) %>% 
                            filter(length(unique(trialNum)) != 24))$gameid)
@@
   group_by(context, hidden) %>%
-  summarize(m = mean(numRawWords), se = sd(numRawWords)/sqrt(length(numRawWords))) %>%
-  ggplot(aes(x = context, y = m, fill = hidden)) +
+  tidyboot_mean(column = numRawWords) %>%
+  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
     geom_bar(stat = 'identity', position = dodge) +
-    geom_errorbar(aes(ymin = m - se, ymax = m + se), position = dodge, width = 0.1) +
+    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
     theme_few() +
     ylab('mean # words') +
-    theme(aspect.ratio = .5)
+    theme(aspect.ratio = .5) +
+    scale_fill_colorblind()
 
-ggsave('~/Downloads/aggregated_means.pdf')
+ggsave('~/Downloads/aggregated_means.pdf', width = 4, height = 2)
 ```
@@
 ```
+
+```{r}
+d.booted <- d %>% 
+  gather(feature, mentioned, shapeMention, colorMention, textureMention) %>%
+  group_by(context, hidden, feature) %>% 
+  tidyboot_mean(column = mentioned)
+
+ggplot(d.booted, aes(x = context, y = empirical_stat, fill = hidden)) + 
+  geom_bar(stat = 'identity', position = dodge) +
+  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
+  facet_wrap(~ feature) + 
+  theme_few() +
+  theme(aspect.ratio = 1) + 
+  ylab('% utterances with feature') +
+  scale_fill_colorblind()
+
+ggsave('~/Downloads/by-feature-analysis.pdf')
+```
+
+## Mouse-tracking
+
+should be really straightforward but might help catch a bug for Exp. 1?
+
+```{r}
+d.mouse <- read_tsv('../data/final/experiment2/planned_sample/updateMouse/mouse.csv') %>%
+  filter(!(gameid %in% badGames)) %>%
+  left_join(d_annotated)
+
+d.mouse %>%
+  filter(distractorDistance != 'none') %>%
+  mutate(distractorDistance = as.numeric(distractorDistance)) %>%
+  group_by(gameid, trialNum) %>%
+  mutate(initTime = first(localTime)) %>%
+  group_by(context, hidden, gameid, trialNum, targetDistance < 300, distractorDistance < 300)  %>%
+  summarize(firstTarget = first(localTime) - mean(initTime))
+  # filter(distractorDistance < 300)
+#  group_by()
+```

### 2018-04-05 writing/journal_manuscript/manuscript.tex
Pass through manuscript; start on results

@@
 When they are unsure exactly which objects are in view of their partner, for example, they may provide additional disambiguating information when referring to a particular object: saying "cassette tape" instead of the potentially ambiguous ``tape."
-Recent debates over adult theory of mind use have hinged on behavior in natural-language communication tasks but have largely neglected the nuanced pragmatics of these tasks.
+Recent debates over adult theory of mind use have hinged on natural-language communication tasks with knowledge asymmetries, but have largely neglected the nuanced pragmatics of these tasks.
+These pragmatic considerations .
 In particular, using occluded objects to create perspective asymmetries may have unexpected side effects under pragmatic reasoning.
-Cooperative speakers ought to flexibly adjust their utterances to be more informative in contexts with high uncertainty about what is in the listener's field of view.
+Cooperative speakers ought to flexibly adjust their utterances to be more informative uncertainty about what is in the listener's field of view.
 In a direct replication, we show that scripted confederates violate these expectations by using less informative labels than unscripted directors naturally use.
@@
 Hence, apparent failures of visual perspective-taking may in fact be successes for sophisticated theory of mind use in context.
+Language is well-tuned to deal with knowledge mismatch, and perspective neglect is just part of that.
 %Recent theories of language use as recursive social reasoning remain in tension with evidence that adult listeners apparently fail to take their partner's perspective in communication.
@@
 Our ability to successfully communicate and coordinate with social partners relies on some degree of contextual flexibility in production and interpretation.
-It may be technically correct to refer to a particular entity as a ``dog,'' but speakers tend to choose more specific or \emph{informative} expressions when trying to point out their dog at a busy dog park.
+When trying to refer to a particular dog at a busy dog park, it may be technically correct to call it a ``dog,'' but speakers tend to choose more specific or \emph{informative} expressions, like ``the little terrier with the blue collar'' \citep{VanDeemter16_ComputationalModelsOfReferring,GrafEtAl16_BasicLevel,BrennanClark96_ConceptualPactsConversation,DaleReiter95_GriceanRefs}.
 Similarly, a hyperbolic statement like ``I waited a million years for a table'' tends to be interpreted as an expression of frustration at longer-than-expected wait times rather than a literal statement about the time elapsed \citep{KaoWuBergenGoodman14_NonliteralNumberWords}.
@@
 % Our success as a cooperative species depends upon our ability to make sense of other agents.
-This systematic contextual variation could in principle be accommodated by internally caching a complex set of different meanings, but pragmatic accounts have argued for a more parsimonious explanation.
-Just as making sense of an agent's physical movements in the world requires an accurate model of how beliefs and intentions translate into motor plans \citep{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}, making sense of language depends on an accurate mental model of what a speaker would say, or what a listener would think, in different situations.
-In other words, flexible communication relies on a generic \emph{theory of mind} mechanism.
+This systematic contextual variation could in principle be accommodated by internally caching a complex set of different rules. 
+But a key insight from the study of pragmatics is that many such phenomena can be more parsimoniously explained as the consequence of shared social reasoning mechanisms \citep{Grice75_LogicConversation, GoodmanFrank16_RSATiCS}.
+Just as making sense of an agent's physical movements in the world requires an accurate model of how beliefs and intentions translate into motor plans \citep{BakerSaxeTenenbaum09_ActionUnderstandingInversePlanning, JaraEttingerEtAl16_NaiveUtilityCalculus, BakerEtAl17_ToMBayesian}, making sense of language depends on an accurate mental model of what a speaker would say, or what a listener would understand, in different situations.
+In other words, communication depends on a flexible \emph{theory of mind} mechanism.
 
 %While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
-Past efforts to evaluate theory of mind in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \citep{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
+Past efforts to examine theory of mind use in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's knowledge or visual access \citep{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}.
 We argue in this paper that the mental models supporting theory of mind use in communication are richer than previously explored. 
-In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \citep{FrankeJager16_ProbabilisticPragmatics, Grice75_LogicConversation,Clark96_UsingLanguage,GoodmanFrank16_RSATiCS, FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
+In particular, we consider a class of mental models representing not only a partner's knowledge or visual access but also Gricean expectations of parsimony and cooperativity \citep{FrankeJager16_ProbabilisticPragmatics, Grice75_LogicConversation,Clark96_UsingLanguage,GoodmanFrank16_RSATiCS, FrankGoodman12_PragmaticReasoningLanguageGames, GoodmanStuhlmuller13_KnowledgeImplicature}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context.
 %For example, suppose a friend wants to refer to a nearby Dalmatian. 
@@
 
-This suggests that we consider whether the scripted utterances produced by  confederates in popular perspective-taking tasks were actually what a speaker in that context would be expected to say \citep{KuhlenBrennan13_LanguageInDialogue}.
+This suggests that we consider whether the scripted utterances produced by  confederates in perspective-taking tasks were actually what a speaker in that context would be expected to say \citep{KuhlenBrennan13_LanguageInDialogue}.
 If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
-More precisely, when both players know that objects are occluded in the display, speakers add additional precision to references in order to avoid confusion. 
-If the listener justifiably expects the speaker to do so, they may \emph{rationally} adopt a strategy of choosing the \emph{a priori} more likely referent of the referring expression regardless of perspective.
-In other words, it is precisely \emph{because} the listener takes the speaker's mental state into consideration that they are tricked by an uncooperative confederate into choosing the wrong item. 
+More precisely, we hypothesize that when both players know that objects are occluded in context, pragmatic speakers may add additional precision to their references in order to avoid confusion. 
+If a listener reasoning in this way justifiably expects the speaker to be more informative in context, they may \emph{rationally} adopt a ``perspective-neglecting'' strategy.
+% of choosing the \emph{a priori} more likely referent of the referring expression regardless of occlusion.
+We propose that it is precisely \emph{because} the listener uses theory of mind to form appropriate pragmatic expectations about the speaker's language use that they can be tricked by an uncooperative confederate into misunderstanding.
 %\begin{table*}
@@
 In Experiment 1, we conduct a direct replication of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, manipulating whether the speaker is a confederate or naive participant.
-We find that unscripted speakers naturally produce more informative utterances than those in the confederate's script.
-In Experiment 2, we test the hypothesis that speakers systematically produce longer, more specific utterances than strictly required by their visible context when they know objects are occluded.
-%Finally, in Experiment 3, we manipulate the listener's pragmatic beliefs about whether the speaker is indeed expected to produce such `over-informative' utterances.
+We find that unscripted speakers naturally produce more informative utterances than those in the confederate's script, and that listeners consequently make fewer errors.
+In Experiment 2, we manipulate the presence or absence of occlusions in context to directly pinpoint the speaker phenomenon of occlusion sensitivity.
+We show for the first time that speakers systematically produce longer, more specific utterances than strictly required by their visible context when they have uncertainty about what their partner is seeing.
+Thus, listeners are rationally justified in ignoring perspective and relying on informative utterances.
 Taken together, these results suggest that apparent failures of perspective-taking are in fact attributable to sophisticated pragmatic expectations about speaker behavior---that is, to successful theory of mind use.
@@
  \vspace{-.25cm}
- \includegraphics[scale=.5]{figures/Exp1_design.png}
+ \includegraphics[scale=.38]{figures/Exp1_design.pdf}
  \vspace{-.5cm}
@@
 
-We recruited K participants (K/2 pairs) from Amazon Mechanical Turk to play an interactive communication game using the framework described in \cite{Hawkins15_RealTimeWebExperiments}.
-All participants were native English speakers.
-%Three pairs were excluded for making 2 or more errors on non-critical items.
+We recruited 350 participants (170 pairs) from Amazon Mechanical Turk to play an interactive communication game using the framework described in \cite{Hawkins15_RealTimeWebExperiments}.
+Following our preregistered exclusion criteria\footnote{Unless }, we removed k games with 2 or more errors on non-critical items or reported confusion in an exit survey and l games containing non-native english speakers, leaving X games in our final sample.
 
@@
 Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it.
-Furthermore, the matcher tended to spend more time hovering their cursor over the occluded cell, and less time hovering over the target object, in ambiguous experimental trials than in baseline trials with no ambiguity.
-Both of these results support Keysar's claim that adults do not reliably deploy theory of mind reasoning when it is called for.
-
-%Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data. First,
-
-%Second, because half of the instructions were unscripted, we were able to observe natural production of referential descriptions for a number of fillers.
-Informally, we noted a tendency toward over-informativity.
-Instead of ``move the stuffed animal down'', they'll say ``move the stuffed pa'nda bear down.'' Or, instead of saying ``move the plane to the right,'' they'll say ``move the red airplane to the right,'' even though there is only one plane.
-This suggested that the confederate's scripted instructions were in some sense uncooperative and counter to the matcher's expectations.
-Perhaps directors were taking the time to make more costly over-informative descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment.
-If the matcher reasoned in this way and expected the director to be overinformative, then they would be justified in picking the first object that meets the description rather than worrying about occluded cells.
-We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
-If speakers really are being overinformative, and listeners are correctly expecting this, we should see fewer errors after removing uncooperative scripts.
+Furthermore, even on trials where they moved the correct object, matchers were more likely to hover their cursor over the occluded cell.
+These results support the claim that listeners neglect the speaker's perspective.
+
+However, when interacting with a natural
 
@@
 
-\section{Evaluating models of referring}
-
-\subsection{Pure egocentrism}
-
-The simplest strategy for a speaker attempting to refer to a target object $o_i$ is to produce the simplest label that best fits $o_i$, effectively ignoring context and listener knowledge.
+%\section{Evaluating models of referring}
+%
+%\subsection{Pure egocentrism}
+%
+%The simplest strategy for a speaker attempting to refer to a target object $o_i$ is to produce the simplest label that best fits $o_i$, effectively ignoring context and listener knowledge.
 \todo[inline]{rxdh: technically, mean soft-max here...} Similarly, the simplest strategy for a listener is to pick the object in their view that is best described by the utterance, ignoring the visual access or pragmatic goals of the speaker.
 These models are purely egocentric, in the sense that they do not engage in any social reasoning.
-
-\subsection{Pure consideration of visual access}
-
-The most immediate model of social reasoning is consideration of the informativeness of a label in a shared context containing many objects.
+%
+%\subsection{Pure consideration of visual access}
+%
+%The most immediate model of social reasoning is consideration of the informativeness of a label in a shared context containing many objects.
 For a speaker, this means preferring labels that are good fits for the target but \emph{not} known distractors in common view.
@@
 These are not available to the visual access of the speaker and thus not viable referents.
-
-\subsection{Probabilistic weighing of perspectives}
-
-Rather than taking a purely egocentric perspective or purely considering a partner's visual access, speakers or listeners may be probabilistically integrating these two perspectives \cite{HellerParisienStevenson16_ProbabilisticWeighing}.
+%
+%\subsection{Probabilistic weighing of perspectives}
+%
+%Rather than taking a purely egocentric perspective or purely considering a partner's visual access, speakers or listeners may be probabilistically integrating these two perspectives \cite{HellerParisienStevenson16_ProbabilisticWeighing}.
 % When conversational expectations lead participants to expect over-informative utterances, they (rationally) place relatively less weight on features of the environment determining common ground, such as shared perceptual access.
-
-\subsection{Gricean communication under uncertainty}
-
-We begin by considering the problem faced by a speaker who wants their partner to select a target object from a context containing other potentially similar objects.
+%
+%\subsection{Gricean communication under uncertainty}
+%
+%We begin by considering the problem faced by a speaker who wants their partner to select a target object from a context containing other potentially similar objects.
 %  a simple probabilistic model of reference in the rational speech act (RSA) model.
-
-where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language.
+%
+%where speakers produce informative utterances by reasoning about a rational listener agent who updates their beliefs about the world according to the literal semantics of the language.
 This allows us to test pragmatic accounts of task performance.
-
-Pragmatic critiques of the director-matcher task point to an additional source of uncertainty that has not previously been explored: uncertainty over a communication partner's visual context.
+%
+%Pragmatic critiques of the director-matcher task point to an additional source of uncertainty that has not previously been explored: uncertainty over a communication partner's visual context.
 Considering this uncertainty, of course, \emph{is} theory of mind in action; our goal is to show how such a model built on social reasoning can produce the `failures' found by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} as well as the phenomena observed by both \cite{HawkinsGoodman16_Keysar} and \cite{RubioFernandez16_DirectorTaskAttention}.
@@
 We incorporate this source of uncertainty into both the speaker and listener models as follows:
-
-\begin{enumerate}
-\item A pragmatic speaker perceives an array of three objects $\mathcal{O}_S = \{o_i\}$, one of which is identified as a target $t$, but assumes that their partner sees an additional hidden object $o_h$ that is not in common ground.
+%
+%\begin{enumerate}
+%\item A pragmatic speaker perceives an array of three objects $\mathcal{O}_S = \{o_i\}$, one of which is identified as a target $t$, but assumes that their partner sees an additional hidden object $o_h$ that is not in common ground.
 The speaker has uncertainty over the identity of this object, represented by a prior distribution $P(o_h)$ which they marginalize over when deciding what utterance to produce:
-$$S_1(u | t, \mathcal{O}_p) \propto \sum_{o_h} L_0(t| \mathcal{O}_S \cup \{o_h\})P(o_h)P(u)$$
-
-%\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$.
+%$$S_1(u | t, \mathcal{O}_p) \propto \sum_{o_h} L_0(t| \mathcal{O}_S \cup \{o_h\})P(o_h)P(u)$$
+%
+%%\item A pragmatic listener, reasoning about this uncertain speaker, is presented with a full array of four objects $\mathcal{O}_L$, one of which is identified as hidden to the speaker, $o_h$.
 The listener, however, also has some uncertainty over whether the experimenter is deceptive, represented by a Bernoulli random variable $d$, such that they assign a small prior probability $P(d)$ to the possibility that $o_h$ is actually in common ground.
 Upon hearing an utterance$u$, they update both their beliefs about which object is the target and whether the experimenter is being deceptive:
-%$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
-\end{enumerate}
-Technical details are omitted here for brevity, but the full model is fully specified in WebPPL and available to run at %\url{http://forestdb.org/models/keysar.html}.
-
+%%$$L_2(t, d | u, \mathcal{O}_L) = P(t)\left(S(u | t, \mathcal{O}_L)P(d)\right)^d(S(u | t, \mathcal{O}_L - o_h)(1-P(d))^{1-d}$$
+%\end{enumerate}
+%Technical details are omitted here for brevity, but the full model is fully specified in WebPPL and available to run at %\url{http://forestdb.org/models/keysar.html}. 
+%
 %We focus on two critical patterns of results produced by the pragmatic listener $L_2$.
@@
 % If listeners arrive at this interpretation strategy through pragmatic reasoning, then neglecting the speaker's epistemic state would be rational and also the result of theory of mind use.
+
+%Second, because half of the instructions were unscripted, we were able to observe natural production of referential descriptions for a number of fillers.
+Informally, we noted a tendency toward over-informativity.
+Instead of ``move the stuffed animal down'', they'll say ``move the stuffed pa'nda bear down.'' Or, instead of saying ``move the plane to the right,'' they'll say ``move the red airplane to the right,'' even though there is only one plane.
+This suggested that the confederate's scripted instructions were in some sense uncooperative and counter to the matcher's expectations.
+Perhaps directors were taking the time to make more costly over-informative descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment.
+If the matcher reasoned in this way and expected the director to be overinformative, then they would be justified in picking the first object that meets the description rather than worrying about occluded cells.
+We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
+If speakers really are being overinformative, and listeners are correctly expecting this, we should see fewer errors after removing uncooperative scripts.
 

### 2018-04-05 writing/journal_manuscript/manuscript.tex
Update exp. 1 participants

@@
 
-We recruited 350 participants (170 pairs) from Amazon Mechanical Turk to play an interactive communication game using the framework described in \cite{Hawkins15_RealTimeWebExperiments}.
-Following our preregistered exclusion criteria\footnote{Unless }, we removed k games with 2 or more errors on non-critical items or reported confusion in an exit survey and l games containing non-native english speakers, leaving X games in our final sample.
+We recruited 368 participants (196 pairs) from Amazon Mechanical Turk to play an interactive communication game using the framework described in \cite{Hawkins15_RealTimeWebExperiments}.
+58 pairs were unable to complete the game due to a server outage.
+Following our preregistered exclusion criteria\footnote{Unless otherwise mentioned, all analyses and \dots are preregistered at \textrm{https://osf.io/qwkmp/}}, we also removed 13 games who reported confusion in an exit survey or made two or more errors on non-critical items and 2 additional games containing non-native English speakers.
+This left 127 pairs in our final sample.
 
@@
 %To deal with participants arriving at unpredictable times, we implemented a waiting room mechanism, where the first person to enter the game environment would see a screen saying ``Waiting for another player..." and the second person would trigger the game to start.
-Participants were randomly assigned to `director' and `matcher' roles and placed in a shared environment containing a chat box and a $4 \times 4$ grid of objects (see Fig.
+Participants in each pair were randomly assigned to `director' and `matcher' roles and placed in a shared environment containing a chat box and a $4 \times 4$ grid of objects (see Fig.
 \ref{fig:exp1paradigm}).

### 2018-04-05 analysis/final_analysis.Rmd
Print out some descriptive statistics

@@
 library(lmerTest)
+library(tidyboot)
+```
+
+# Experiment 1
+
+```{r}
+exp1_messages <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv') %>%
+  mutate(critical = ifelse(critical == 'True', 1, 0)) %>%
+  rename(messageTime = serverTime) %>%
+  filter(sender == 'director') %>%
+  group_by(gameid, trialNum, instructionNum, attemptNum, critical, condition, objectSet, targetObject) %>%
+  summarize(text = paste0(contents, collapse = ' '),
+            typingRT = sum(typingRT)) %>%   # Concatenate multiple utterance on same trial
+  mutate(numRawWords = str_count(text, "\\S+"))
+
+## Have to fix attemptNum bug (got recorded incorrectly)
+exp1_drops <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
+  mutate(critical = ifelse(critical == 'True', 1, 0),
+         correct = ifelse(correct == 'correct', 1, 0),
+         instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
+  group_by(gameid, instructionNum, trialNum) %>%
+  mutate(attemptNum = ifelse(attemptNum == 0, max(attemptNum), attemptNum - 1)) %>%
+  rename(dropTime = serverTime) %>%
+  select(-X1, -eventType)
+```
+
+```{r}
+exp1_subjInfo <- read_csv('../data/final/experiment1/turk/subjInfo.csv')
+incompleteIDs <- unique((exp1_messages %>% 
+                           group_by(gameid, trialNum, instructionNum) %>% 
+                           summarize(id = row_number()) %>% 
+                           group_by(gameid) %>% 
+                           tally() %>% 
+                           filter(n != 32))$gameid)
+
+confused <- unique((exp1_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
+nonNative <- unique((exp1_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)
+
+nonCriticalMistakes <- unique((exp1_drops %>%
+  # only look at mistakes on noncritical (filler) items
+  filter(critical != 1) %>%
+  # don't want to double-count people for messing the same thing up multiple times
+  filter(attemptNum == 1) %>% 
+  group_by(gameid, correct) %>%
+  tally() %>%
+  # implement exclusion criteria of errors on >~10% of non-critical trials
+  filter(n > 1))$gameid)
+
+exp1_badGames <- c(incompleteIDs, nonNative, confused, nonCriticalMistakes)
+
+exp1_d <- exp1_messages %>%
+  left_join(exp1_drops) %>% #, by = c('gameid', 'trialNum', 'instructionNum', 'trialType', 
+                      #         'targetObject', 'critical', 'condition', 'attemptNum')) %>%
+  filter(!(gameid %in% exp1_badGames))
+```
+
+## Descriptive stats
+
+How many recruited, how many excluded, for what reasons, etc?
+
+```{r}
+paste0('recruited ', length(exp1_subjInfo$gameid))
+paste0('total games ', length(unique(exp1_drops$gameid)))
+paste0('incomplete ', length(incompleteIDs))
+paste0('too many mistakes ', length(setdiff(union(confused, nonCriticalMistakes), incompleteIDs)))
+paste0('non-native english ', length(setdiff(nonNative, union(confused, union(nonCriticalMistakes, incompleteIDs)))))
+paste0('final sample ', length(unique(exp1_d$gameid)))
+```
+
+Happened to get slightly more people in scripted condition
+
+```{r}
+exp1_d %>% 
+  group_by(gameid, condition) %>% 
+  tally() %>% 
+  group_by(condition) %>% 
+  summarize(numInCondition = length(n))
+```
+
+## Listener Analyses
+
+First, look at error rates
+
+```{r}
+critTrials <- exp1_d %>% 
+  filter(attemptNum == 0) %>%
+  filter(critical == 1) %>%
+  filter(trialType == 'exp')
+
+critTrials %>%
+  group_by(gameid, condition) %>%
+  summarize(pctCriticalError = 1 - mean(correct)) %>%
+  ggplot(aes(x = pctCriticalError, fill = condition)) +
+    #geom_density(aes(y = ..density../sum(..density..)), alpha = 0.5, adjust = 1.25)
+    geom_histogram(aes(y = ..count../sum(..count..)), alpha = 0.5, position = 'dodge', binwidth = 0.125)  +
+    facet_grid(condition ~ .) +
+    theme_few()
+
+ggsave('~/Downloads/criticalErrors.pdf')
+```
+
+```{r}
+errorTable <- critTrials %>%
+  group_by(gameid, condition) %>%
+  summarize(numErrors = 4 - sum(correct),
+            numPossible = length(correct),
+         atLeastOnce = sum(correct) < 4,
+         atLeastTwice = sum(correct) < 3,
+          atLeastThree = sum(correct) < 2,
+         pctErrors = 1 - mean(correct)) %>%
+  group_by(condition) %>%
+  summarize(pctAtLeastOnce = mean(atLeastOnce),
+            pctAtLeastTwice = mean(atLeastTwice),
+            pctAtLeastThree = mean(atLeastThree),
+            avgPctErrors = mean(pctErrors),
+            totalNumErrors = sum(numErrors),
+            totalPossible = sum(numPossible))
+errorTable
+```
+
+### Statistical tests of error reduction
+
+```{r}
+prop.test(errorTable$totalNumErrors, errorTable$totalPossible)
+```
+
+### Control for item-level error rates...
+
+```{r}
+summary(glmer(correct ~ condition + (1|gameid) + (1+condition | objectSet), family = 'binomial', 
+              data = critTrials))
+```
+
+## Error heterogeneity across items...
+
+```{r}
+diffs <- critTrials %>% 
+  mutate(error = 1 - correct) %>%
+  group_by(objectSet, condition, targetObject) %>%
+  summarize(pctErrors = mean(error)) %>%
+  spread(condition, pctErrors) %>%
+  mutate(errorReduction = scripted - unscripted) %>%
+  gather(condition, pctErrors, scripted, unscripted)
+
+critTrials %>% 
+  left_join(diffs) %>%
+  mutate(error = 1 - correct) %>%
+  group_by(objectSet, condition, targetObject, errorReduction) %>%
+  tidyboot_mean(column = error) %>%
+  ggplot(aes(x = reorder(targetObject, -errorReduction), y = empirical_stat, color = condition, group = condition)) +
+    geom_point() +
+    geom_line(stat = 'identity', position = 'dodge') +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
+    theme_few() +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+    ggtitle('error heterogeneity across item') +
+    xlab('critical item') +
+    ylab('% error in condition')
+
+ggsave('~/Downloads/itemHeterogeneity.pdf')
+```
+
+```{r}
+exp1_d %>% 
+  filter(attemptNum == 0) %>%
+  filter(correct == 1) %>%
+  filter(critical == 1) %>% 
+  ungroup() %>%
+  filter(responseTime < median(responseTime, na.rm = T) +  3* sd(responseTime)) %>%
+  group_by(trialType, condition) %>%
+  mutate(m = median(responseTime)) %>%
+  ggplot(aes(x = responseTime)) +
+    geom_histogram() +
+    geom_vline(aes(xintercept = m))+
+    facet_wrap(trialType ~ condition) +
+    scale_x_log10() +
+    theme_few() 
+```
+
+```{r}
+summary(lmer(log(responseTime) ~ trialType * condition + (1 | gameid),
+     data = exp1_d %>% 
+        filter(attemptNum == 0) %>%
+        filter(correct == 1) %>%
+        filter(critical == 1) %>% 
+        ungroup() %>%
+        filter(responseTime < median(responseTime, na.rm = T) +  3* sd(responseTime))))
+```
+
+## Mouse-tracking
+
+Even on critical trials where they didn't make errors, how often did they 'consider' the distractor (i.e. hover over the distractor cell)
+
+```{r}
+d.onDistractor <- exp1_d %>% 
+  left_join(read_csv('../data/final/experiment1/mouseOverDistractor/distractorMouseFromMongo.csv') %>% 
+              mutate(critical = ifelse(critical == 'True', 1, 0))
+            ) %>%
+  select(-serverTime, -X1) %>%
+  filter(attemptNum == 0) %>%
+  filter(correct == 1) %>%
+  filter(critical == 1) %>%
+  group_by(gameid, trialNum) %>%
+  mutate(toggle = floor((row_number() - 1) / 2)) %>%
+  mutate(onOff = ifelse(is.na(onOff), 'off', onOff)) %>%
+  spread(onOff, timeElapsed) %>%
+  mutate(diff = ifelse(is.na(off), 0, off - on)) %>%
+  group_by(gameid,condition,objectSet, trialType) %>%
+  summarize(totalAmtTimeOnDistractor = sum(diff)) %>%
+  mutate(hovered = totalAmtTimeOnDistractor > 0) %>%
+  ungroup()
+
+dodge <- position_dodge(width=0.9)
+d.onDistractor %>%
+  group_by(condition, trialType) %>%
+  tidyboot_mean(column = hovered) %>%
+  ggplot(aes(x = condition, y=empirical_stat, fill = trialType)) +
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
+    theme_few() +
+    ylab('% trials hovering over distractor') +
+    theme(aspect.ratio = 0.5)
+
+ggsave('~/Downloads/hoverOverDistractor.pdf')
+```
+
+```{r}
+summary(glmer(hovered ~ trialType * condition + (1 + trialType|gameid) + (1 |objectSet), data = d.onDistractor, family = 'binomial'))
+```
+
+Note: because I stopped recording from the mouse as soon as someone starts dragging, it's technically possible for them to hover over the distractor object, click and start to drag it, but then drop it in place and go over to the target object.
+So they get it 'correct' even though it looks like their mouse never gets close...
+
+Note: a couple researcher degrees of freedom here: 
+* do we take the time from 'object reveal' or from first mouse move (i.e. subtract off the first time recorded NOT in target or distractor)? 
+* do we restrict to only 'correct' trials (i.e. to make an argument like "even when they got it right, they were still distracted") or include everything?
+* how big a radius around the target center should we consider a 'mouse-fixation' (obviously needs to be close enough to move it, but objects were different sizes...)
+
+```{r}
+# Touch people?
+touch <- c('0095-03bb67a7-ae89-402d-9f7f-c5898784e63f', '0093-b4e1ce5c-2e5c-462a-bfed-87fa81611139')
+targetRadius = 175
+d.mouse <- exp1_d %>% ungroup() %>%
+  left_join(read_csv('../data/final/experiment1/updateMouse/mouseFromMongo.csv') %>%
+                  mutate(critical = ifelse(critical == 'True', 1, 0))
+  ) %>%
+  filter(!(gameid %in% touch)) %>%
+  filter(critical == 1) %>% filter(attemptNum == 0) %>% filter(correct == 1) %>%
+  select(-critical, -attemptNum, -eventType, -iterationName) %>%
+  mutate(distractorDistance = as.numeric(distractorDistance)) %>%
+  group_by(condition, trialType, gameid, objectSet, onTarget = targetDistance < targetRadius) %>%
+  summarize(firstTime = first(timeFromReveal), 
+            lastTime = last(timeFromReveal)) %>%
+  group_by(condition, trialType, gameid, objectSet) %>%
+  summarize(initMouseMove = min(firstTime), 
+            initTargetHover = max(firstTime) - initMouseMove, 
+            finalTargetHover = max(lastTime) - initMouseMove) %>%
+  select(-initMouseMove) %>%
+  gather(metric, value, initTargetHover, finalTargetHover) %>%
+  #spread(onTarget, value) %>%
+  #mutate(timeFromFirstMouseMove = `TRUE` - `FALSE`) %>%
+  #ungroup() %>%
+  #filter(firstTime < mean(firstTime, na.rm = T) + 3 * sd(firstTime, na.rm = T)) %>%
+  group_by(condition, trialType, metric) %>%
+  tidyboot_mean(column = value)
+
+d.mouse
+ggplot(d.mouse, aes(x = trialType, y = empirical_stat, fill = metric)) +
+  geom_bar(stat= 'identity', position = dodge) +
+  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
+  facet_wrap(~ condition) +
+  theme_few()
 ```
@@
 ```{r}
-subjInfo <- read_csv('../data/final/experiment2/planned_sample/turk/subjInfo.csv')
+exp2_subjInfo <- read_csv('../data/final/experiment2/planned_sample/turk/subjInfo.csv')
 ```
@@
                            filter(length(unique(trialNum)) != 24))$gameid)
-confused <- unique((subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
+confused <- unique((exp2_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
 
@@
 ```
-
-## Mouse-tracking
-
-should be really straightforward but might help catch a bug for Exp. 1?
-
-```{r}
-d.mouse <- read_tsv('../data/final/experiment2/planned_sample/updateMouse/mouse.csv') %>%
-  filter(!(gameid %in% badGames)) %>%
-  left_join(d_annotated)
-
-d.mouse %>%
-  filter(distractorDistance != 'none') %>%
-  mutate(distractorDistance = as.numeric(distractorDistance)) %>%
-  group_by(gameid, trialNum) %>%
-  mutate(initTime = first(localTime)) %>%
-  group_by(context, hidden, gameid, trialNum, targetDistance < 300, distractorDistance < 300)  %>%
-  summarize(firstTarget = first(localTime) - mean(initTime))
-  # filter(distractorDistance < 300)
-#  group_by()
-```

### 2018-05-05 experiments/final/normingExperiment/experiment/js/experiment.js
Add norming experiment for new referring exps

@@ -1,0 +1,148 @@
+function make_slides(f) {
+  var q_first = null;
+  var slides = {};
+  slides.i0 = slide({
+    name : "i0",
+    start: function() {
+      exp.startT = Date.now();
+    }
+  });
+  
+  slides.q_instructions = slide({
+    name : "q_instructions",
+    button : function() {
+      exp.go(); //use exp.go() if and only if there is no "present" data.
+    }
+  });
+  
+  slides.rating = slide({
+    name: "rating",
+    present : _.shuffle(stimList),
+    present_handle : function(stim) {
+      // Set up scenario & instructions
+      this.trial_start = Date.now();
+      this.stim = stim;
+      this.init_sliders();
+      exp.sliderPost = {};
+      var contextsentence = "How well does <strong>"+this.stim.label+"</strong> apply to this object?";
+      if(stim.context == "isolated") {
+	var objimagehtml = '<img src="'+this.stim.object.url+'" style="height:190px;">';
+      } else {
+	var objimagehtml = '<img src="'+this.stim.object.contextURL+'" style="height:190px;">';
+      }
+      $("#contextsentence").html(contextsentence);
+      $("#objectimage").html(objimagehtml);
+      console.log(this);
+      $(".err1").hide();
+    },
+
+    button : function() {
+      if (exp.sliderPost > -1 && exp.sliderPost < 16) {
+	$(".err").hide();
+	this.log_responses();
+	_stream.apply(this); //use exp.go() if and only if there is no "present" data.
+      } else {
+	$(".err").show();
+      }
+    },
+    init_sliders : function() {
+      utils.make_slider("#single_slider", function(event, ui) {
+	exp.sliderPost = ui.value;
+      });
+    },
+    log_responses : function() {
+      exp.data_trials.push({
+        "label" : this.stim.label,
+        "slideNumber": exp.phase,
+        "response" : exp.sliderPost,
+	"objectSet" : this.stim.objectSet,
+	"referent" : this.stim.referent,
+	"context" : this.stim.context,
+	"rt" : Date.now() - _s.trial_start
+      });
+    },
+  });
+  
+  slides.subj_info =  slide({
+    name : "subj_info",
+    submit : function(e){
+      //if (e.preventDefault) e.preventDefault(); // I don't know what this means.
+      exp.subj_data = {
+        language : $("#language").val(),
+        enjoyment : $("#enjoyment").val(),
[…]

### 2018-05-05 experiments/final/normingExperiment/experiment/js/genObjectSet.js

@@ -1,0 +1,29 @@
+var fs = require('fs');
+var babyparse = require('papaparse');
+var _ = require('underscore');
+
+function readCSV(filename){
+    return babyparse.parse(fs.readFileSync(filename, 'utf8'),
+			   {header: true, skipEmptyLines : true}
+			  ).data;
+  };
+
+var criticalMessages = readCSV('../../../../../data/final/experiment1/refExpressions_standardized.csv');
+
+var unscriptedLabels = _.map(criticalMessages, function(row) {
+  return {text: row.text.trim(), objectSet : row.objectSet};
+});
+
+var scriptedLabels = [{text: "glasses", objectSet : 1},
+		      {text: "bottom block", objectSet : 2},
+		      {text: "tape",  objectSet : 3},
+		      {text: "large measuring cup", objectSet : 4},
+		      {text: "brush", objectSet : 5},
+		      {text: "eraser", objectSet : 6},
+		      {text: "small candle", objectSet : 7},
+		      {text: "mouse", objectSet : 8}];
+
+var uniqueLabels = _.unique(unscriptedLabels.concat(scriptedLabels),
+			    function(item, key, a) {  return item.text; });
+
+fs.writeFileSync("uniqueLabels.json", JSON.stringify(uniqueLabels, null, 4));

### 2018-05-05 experiments/final/normingExperiment/experiment/js/objectSet.js

@@ -1,0 +1,601 @@
+var stimList = function() {
+  var uniqueLabels = [
+    {
+      "label": "glasses",
+      "objectSet": "1"
+    },
+    {
+      "label": "black sunglasses",
+      "objectSet": "1"
+    },
+    {
+      "label": "sunglasses",
+      "objectSet": "1"
+    },
+    {
+      "label": "shades",
+      "objectSet": "1"
+    },
+    {
+      "label": "B block",
+      "objectSet": "2"
+    },
+    {
+      "label": "toy block",
+      "objectSet": "2"
+    },
+    {
+      "label": "blue block",
+      "objectSet": "2"
+    },
+    {
+      "label": "B cube",
+      "objectSet": "2"
+    },
+    {
+      "label": "child's blue block",
+      "objectSet": "2"
+    },
+    {
+      "label": "alphabet block",
+      "objectSet": "2"
+    },
+    {
+      "label": "B box",
+      "objectSet": "2"
+    },
+    {
+      "label": "block",
+      "objectSet": "2"
+    },
+    {
+      "label": "block with the blue writing",
+      "objectSet": "2"
+    },
+    {
+      "label": "B",
+      "objectSet": "2"
+    },
+    {
+      "label": "block with the B",
+      "objectSet": "2"
+    },
+    {
+      "label": "the bluish block with a B",
+      "objectSet": "2"
+    },
+    {
+      "label": "letter block",
+      "objectSet": "2"
+    },
+    {
+      "label": "block toy",
+      "objectSet": "2"
+    },
+    {
+      "label": "block",
+      "objectSet": "2"
+    },
+    {
+      "label": "blue B block",
+      "objectSet": "2"
+    },
+    {
+      "label": "tape",
+      "objectSet": "3"
+    },
+    {
+      "label": "cassette",
+      "objectSet": "3"
+    },
+    {
+      "label": "cassette tape",
+      "objectSet": "3"
+    },
+    {
+      "label": "clear audiocassette",
+      "objectSet": "3"
+    },
+    {
+      "label": "large measure cup",
+      "objectSet": "4"
+    },
+    {
+      "label": "pyrex measuring cup",
+      "objectSet": "4"
+    },
+    {
+      "label": "pyrex",
+      "objectSet": "4"
+    },
+    {
+      "label": "measuring cup",
+      "objectSet": "4"
+    },
+    {
+      "label": "big measuring cup",
[…]

### 2018-05-05 experiments/final/normingExperiment/experiment/js/stim_list.js

@@ -1,0 +1,69 @@
+var stim_list = {
+  "pets" :    [{"s" : "Mary is at a party making polite conversation. Susan asks: ",
+                  "q" : "'Do you have a dog?'",
+                  "a_name" : "Mary",
+                  "q_entity" : "dog",
+                  "tt": "dog_conv",
+                  "a" : ["I have a dog","I have a cat", "I have an animal","I have a dalmatian"]},
+                 {"s" : "Mary knows Susan's family is coming into town and she wants to impress them with a cute pet. Susan calls Mary and asks: ",
+                  "q" : "'Do you have a dog?'",            
+                  "a_name" : "Mary",                
+                  "q_entity" : "dog",
+                  "tt": "dog_pr_any",
+                  "a" : ["I have a dog","I have a cat", "I have an animal","I have a dalmatian"]},
+                 {"s" : "Susan is visiting Mary's house for dinner, and Mary knows she is allergic to dogs. Before coming over, Susan asks: ",
+                  "q" : "'Do you have a dog?'",
+                  "a_name" : "Mary",                 
+                  "q_entity" : "dog",
+                  "tt": "dog_info",
+                  "a" : ["I have a dog","I have a cat", "I have an animal","I have a dalmatian"]}],
+    "clothes":  [{"s" : "Bob's friends roped him into going to the opera with them and want to make sure he has the appropriate clothes to wear. His friend Mark asks: ",
+                  "q": "'Do you have a shirt?'",
+                  "a_name" : "Bob",    
+                  "q_entity" : "shirt", 
+                  "tt": "shirt_info",
+                  "a" : ["I have a shirt", "I have some pants", "I have some clothing", "I have a button-down shirt"]},
+                 {"s" : "Mark is drinking coffee at a coffee shop before going on a date and accidentally spills coffee all over his shirt. He calls his friend Bob who lives close by, explains the situation, and asks: ",
+                  "q": "'Do you have a shirt?'",                 
+                  "a_name" : "Bob",                     
+                  "q_entity" : "shirt",
+                  "tt": "shirt_pr_basic",
+                  "a" : ["I have a shirt", "I have some pants", "I have some clothing", "I have a button-down shirt"]},   
+                 {"s" : "Bob's school is collecting clothes to send to those in need and Bob wants to donate. Mark, who's working at the collection, asks him: ",
+                  "q": "'Do you have a shirt?'",                 
+                  "a_name" : "Bob",                     
+                  "q_entity" : "shirt",
[…]

### 2018-05-05 analysis/final_analysis.Rmd
Exclude buggy trials; look at ref exps

@@
 ```{r}
+raw <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv')
 exp1_messages <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv') %>%
@@
 
-## Have to fix attemptNum bug (got recorded incorrectly)
+## Have to post-hoc attemptNum and instructionNum bugs (got reset for next trial prematurely)
+## Also, remove trials where multi-drop bug happened (i.e. where clicking multiple times triggered multiple instruction resetting)
 exp1_drops <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
@@
          instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
+  group_by(gameid, targetObject) %>%
+  filter(length(unique(instructionNum)) < 2) %>%
   group_by(gameid, instructionNum, trialNum) %>%
@@
 
-exp1_badGames <- c(incompleteIDs, nonNative, confused, nonCriticalMistakes)
+buggyTrials <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
+  mutate(critical = ifelse(critical == 'True', 1, 0),
+       correct = ifelse(correct == 'correct', 1, 0),
+       instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
+  group_by(gameid, targetObject) %>%
+  mutate(numInstructions = length(unique(instructionNum))) %>%
+  filter(numInstructions > 1)
 
-exp1_d <- exp1_messages %>%
-  left_join(exp1_drops) %>% #, by = c('gameid', 'trialNum', 'instructionNum', 'trialType', 
+taboo <- c('4241-cda4a373-ffd7-41e3-ae60-82b9d377a455')
+onlyLocation <- c('3222-f68d2df5-8977-4ff2-9ee9-c60f8ddb61f1', '9457-4189cc9b-a82b-4db9-828d-bf90d1987aee')
+exp1_badGames <- c(incompleteIDs, nonNative, confused, nonCriticalMistakes, taboo, onlyLocation, unique(buggyTrials$gameid))
+
+exp1_d <- exp1_drops %>% 
+  left_join(exp1_messages) %>% #, by = c('gameid', 'trialNum', 'instructionNum', 'trialType', 
                       #         'targetObject', 'critical', 'condition', 'attemptNum')) %>%
@@
 
+And slightly uneven numbers of 'experimental' trials in each item...
+
+```{r}
+exp1_d %>% 
+  group_by(gameid, objectSet, trialType, condition) %>% 
+  tally() %>% 
+  group_by(condition, objectSet, trialType) %>% 
+  summarize(numInCondition = length(n))
+```
+
 ## Listener Analyses
@@
   group_by(gameid, condition) %>%
-  summarize(pctCriticalError = 1 - mean(correct)) %>%
+  summarize(pctCriticalError = 4 - sum(correct)) %>%
   ggplot(aes(x = pctCriticalError, fill = condition)) +
-    #geom_density(aes(y = ..density../sum(..density..)), alpha = 0.5, adjust = 1.25)
-    geom_histogram(aes(y = ..count../sum(..count..)), alpha = 0.5, position = 'dodge', binwidth = 0.125)  +
+    geom_histogram(alpha = 0.5, position = 'dodge', binwidth = 0.5)  +
     facet_grid(condition ~ .) +
-    theme_few()
+    theme_few() +
+    ylab('# pairs making k critical errors') +
+    xlab('# critical errors') + 
+    theme(aspect.ratio = 1, legend.position="none") +
+    scale_fill_colorblind()+
+    ggtitle('critical errors')
 
-ggsave('~/Downloads/criticalErrors.pdf')
+ggsave('../writing/journal_manuscript/figures/criticalErrors.pdf')
+```
+
+```{r}
+critTrials %>%
+  group_by(condition) %>%
+  mutate(error = 1 - correct) %>%
+  tidyboot_mean(column = error) %>%
+  ggplot(aes(x = condition, y =empirical_stat, fill = condition)) +
+    geom_bar(stat = 'identity', alpha = .5) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
+    ylab('% errors') +
+    scale_fill_colorblind()+
+    theme_few() +
+    theme(aspect.ratio = 1, legend.position="none")
+
+ggsave('../writing/journal_manuscript/figures/errorMeans.pdf', height = 3, width = 3)
 ```
@@
 
-ggsave('~/Downloads/itemHeterogeneity.pdf')
+ggsave('../writing/journal_manuscript/figures//itemHeterogeneity.pdf')
 ```
@@
   ggplot(aes(x = condition, y=empirical_stat, fill = trialType)) +
-    geom_bar(stat = 'identity', position = dodge) +
+    geom_bar(stat = 'identity', position = dodge, alpha =.5) +
     geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
@@
     ylab('% trials hovering over distractor') +
+    scale_fill_colorblind()+
     theme(aspect.ratio = 0.5)
 
-ggsave('~/Downloads/hoverOverDistractor.pdf')
+ggsave('../writing/journal_manuscript/figures//hoverOverDistractor.pdf', height = 3, width = 6)
 ```
@@
 
+## Prep informativity norming stims
+
+```{r}
+refExps <- exp1_d %>% 
+  filter(condition == 'unscripted') %>% 
+  filter(critical == 1) %>% 
+  filter(attemptNum == 0) %>% 
+  ungroup() %>%
+  select(gameid, objectSet, text)
+
+write_csv(refExps, './refExpressions_raw.csv')
+```
+
+For 6 of the 8 items, unscripted speakers had significantly more words in their referring expressions... (n.s. for the other 2, but idiosyncratic)
+
+```{r}
+refs <- read_csv('../data/final/experiment1/refExpressions.csv')
+scripted <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), 
+                       m = c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse'),
+                       numRawWords = c(1,2,1,3,1,1,2,1))
+refs %>% 
+  mutate(numRawWords = str_count(text, "\\S+")) %>%
+  group_by(objectSet) %>%
+  tidyboot_mean(numRawWords) %>%
+  left_join(scripted)
+```
+
+## TO-DO
+
+1. Subject-level analyses of informativity: could be a mixture of 'lazy' egocentric speakers and 'attentive' audience-sensitive speakers? 
+2. Sometimes the additional information is in movement instruction (i.e. "one square left toward a curtain" or "one square up below the comb")
+3.
+Change in informativity over time (are people already being informative on first round?
+are people 'learning' to be more informative over course of expt?
+or getting less informative because tired of task?)
+4. Just length of referring expression in words?
+
 # Experiment 2
@@
               '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
-badGames <- c(incompleteIDs, pureLocation, confused)
+badGames <- c(incompleteIDs, location_abusers, confused)
 
@@
     xlim(1,12) +
-    theme(aspect.ratio = 1)
+    theme(aspect.ratio = 1, text = element_text(size=20))
 ```
@@
 ```{r}
-ggplot(basic.d, aes(x = ratio)) +
-    geom_histogram(binwidth = 0.25) +
-    theme_bw() +
-    geom_vline(xintercept = 0) +
-    # ylim(1,12) +
-    xlim(-4,4) +
-    theme(aspect.ratio = 1)
+tidyboot_mean(basic.d %>%ungroup(), column = diff) %>%
+  ggplot(aes(x = 'close - far', y = empirical_stat)) +
+      geom_bar(stat ='identity', width = 0.25) +
+      geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
+      theme_few()+
+      geom_hline(yintercept = 0) +
+      #geom_vline(xintercept = 0) +
+      ylim(-1,1) +
+      # xlim(-4,4) +
+      ylab('# words used') +
+    theme(aspect.ratio = 1, text = element_text(size=20))
 
-t.test(basic.d$ratio)
+t.test(basic.d$diff)
 ```
@@
     ylab('mean # words') +
-    theme(aspect.ratio = .5) +
+    theme(aspect.ratio = 1, text = element_text(size=20)) +
     scale_fill_colorblind()
@@
     xlab('# words (without occlusions)') +
-    ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
+    #ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
     xlim(1,15) +
-    theme(aspect.ratio = 1)
+    theme(aspect.ratio = 1, text = element_text(size=20))
 
@@
 ```{r}
-ggplot(hidden.d %>% filter(!(gameid %in% all_the_things_ers)) %>% group_by(context) %>% mutate(m = mean(diff)), aes(x = diff)) +
+hidden.d %>% 
+  #filter(!(gameid %in% all_the_things_ers)) %>% 
+  group_by(context) %>% 
+  mutate(m = mean(diff)) %>%
+  ggplot(aes(x = diff)) +
       geom_histogram(binwidth = 0.33) +
@@
   ggplot(aes(x = context, y = m)) +
-    geom_bar(stat = 'identity') +
+    geom_bar(stat = 'identity', width = 0.25) +
     geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0) +
     theme_few() +
+    geom_hline(yintercept = 0) +
     ylim(-3, 3) +
+    theme(aspect.ratio = 1, text = element_text(size=20))  +
     ylab('# words (occlusions - no occlusions)')
@@
   theme_few() +
-  theme(aspect.ratio = 1) + 
+  theme(aspect.ratio = 1, text = element_text(size=20))  +
   ylab('% utterances with feature') +

### 2018-05-05 writing/journal_manuscript/manuscript.tex
Tweak expt 1 methods...

@@
 
+\subsection{Experiment~1: Direct replication}
+\label{sec:Exp1}
+
+\subsection{Participants}
+
+We recruited 368 participants (196 pairs) from Amazon Mechanical Turk to play an interactive communication game using the framework described in \cite{Hawkins15_RealTimeWebExperiments}.
+58 pairs were unable to complete the game due to a server outage.
+Following our preregistered exclusion criteria\footnote{Unless otherwise mentioned, all analyses and \dots are preregistered at \textrm{https://osf.io/qwkmp/}}, we also removed 13 games who reported confusion in an exit survey or made two or more errors on non-critical items and 2 additional games containing non-native English speakers.
+This left 127 pairs in our final sample.
+
 \begin{figure}[t!]
@@
 
-\subsection{Experiment~1: Direct replication}
-\label{sec:Exp1}
-
-\subsection{Participants}
-
-We recruited 368 participants (196 pairs) from Amazon Mechanical Turk to play an interactive communication game using the framework described in \cite{Hawkins15_RealTimeWebExperiments}.
-58 pairs were unable to complete the game due to a server outage.
-Following our preregistered exclusion criteria\footnote{Unless otherwise mentioned, all analyses and \dots are preregistered at \textrm{https://osf.io/qwkmp/}}, we also removed 13 games who reported confusion in an exit survey or made two or more errors on non-critical items and 2 additional games containing non-native English speakers.
-This left 127 pairs in our final sample.
-
 \subsection{Materials \& Procedure}
@@
 They used a chat box to communicate the identity of this object and action to their partner, who proceeded to click and drag an object to complete the trial.
-
-We used a between-subject design to test whether the scripted confederate labels were in fact what cooperative speakers would say in context.
+In each blocked set of objects, one target belonged to a `critical pair' of objects, such as a visible cassette tape and a hidden roll of tape that could both plausibly be referred to as `the tape'.
+
+We used a between-subject design to test whether the scripted labels used by confederates were in fact comparable to what cooperative speakers would say in context.
 Directors assigned to the `scripted' condition were free to communicate however they wished for half of the instructions.
@@
 
-In each block, one target belonged to a `critical pair' of objects, such as a visible cassette tape and a hidden roll of tape that could both plausibly be referred to as `the tape'.
 We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-Each participant received half the items in the experimental condition and half in the baseline condition.
+Each participant was shown half of the items in the experimental condition and half in the baseline condition.
 The assignment of items to conditions was randomized across participants, and the order of conditions was randomized under the constraint that the same condition would not be used on more than two consecutive items.
@@
 In addition to collecting messages sent through the chat box and matcher errors, we collected mouse-tracking data in analogy to the eye-tracking common in these paradigms.
-To collect clean data, we asked the matcher to wait until the director sent a message; when the message was received, the matcher clicked a small circle in the center of the grid to show the objects and proceed with the trial.
-We recorded at 100Hz from the matcher's mouse in the decision window after this click.
+To collect the cleanest possible data, we asked the matcher to wait until the director sent a message; when the message was received, the matcher clicked a small circle in the center of the grid to show the objects and proceed with the trial.
+We recorded at 100Hz from the matcher's mouse in the decision window after this click, until the point where they clicked and started to drag one of the objects.
 

### 2018-05-07 writing/journal_manuscript/manuscript.tex
Tighten up & fill in exp. 1 results

@@
 
-This suggests that we consider whether the scripted utterances produced by  confederates in perspective-taking tasks were actually what a speaker in that context would be expected to say \citep{KuhlenBrennan13_LanguageInDialogue}.
+This suggests that we consider whether the scripted utterances produced by  confederates in influential perspective-taking tasks were actually what a speaker in that context would be expected to say \citep{KuhlenBrennan13_LanguageInDialogue}.
 If not, then perhaps listeners are making choices that are in fact consistent with a correct pragmatic interpretation of the confederate's (uncooperative) utterance.
@@
 58 pairs were unable to complete the game due to a server outage.
-Following our preregistered exclusion criteria\footnote{Unless otherwise mentioned, all analyses and \dots are preregistered at \textrm{https://osf.io/qwkmp/}}, we also removed 13 games who reported confusion in an exit survey or made two or more errors on non-critical items and 2 additional games containing non-native English speakers.
+Following our preregistered exclusion criteria\footnote{Unless otherwise mentioned, all analyses and \dots are preregistered at \textrm{https://osf.io/qwkmp/}}, we also removed 13 games who reported confusion or made two or more errors on filler items and 2 additional games containing non-native English speakers.
 This left 127 pairs in our final sample.
+\todo[inline]{rxdh: also mention removal of 3 games only using location or using 'taboo' utterances.}
 
@@
 They used a chat box to communicate the identity of this object and action to their partner, who proceeded to click and drag an object to complete the trial.
-In each blocked set of objects, one target belonged to a `critical pair' of objects, such as a visible cassette tape and a hidden roll of tape that could both plausibly be referred to as `the tape'.
+In each blocked set of objects, one target belonged to a `critical pair' of objects, such as a visible cassette tape and a hidden roll of tape that could both plausibly be referred to as `the tape.'
 
@@
 For the other half, including all the critical instructions, their messages were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat box, and they would have to click `send' for the experiment to continue.) In the `unscripted' condition, directors were unrestricted and free to send whatever messages they deemed appropriate on all trials.
+For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square."\footnote{That is, in these conditions, the scripted message would automatically appear in the director's chat box, and they would have to click `send' for the experiment to continue.} In the `unscripted' condition, directors were unrestricted and free to send whatever messages they deemed appropriate on all trials.
 
@@
 
-\subsubsection{Critical errors}
-
-In Figure \ref{table:mainResults} we compare error rates across scripted and unscripted conditions.
-Participants rarely moved the hidden object on baseline trials in either condition, and total error rates for critical trials are significantly higher when speakers used scripted utterances, $\chi^2(1) = 5.35, p = 0.02$.
-Error rates in the scripted condition were similar to the effects observed by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-%, which found 71\% and 0\%.
-Our errors were larger across the board, perhaps due to the interface or the population, but the gap between the two conditions is roughly the same size.
-%Next, we examine whether there are differences in error rates across items.
-
-We also found substantial heterogeneity in errors rates across items (Table \ref{table:ItemWise}).
-Patterns of errors diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
-For example, 75\% of participants made an error on item 6 upon hearing the scripted instruction ``the eraser'' while only 17\% of participants in item 8 made an error upon hearing ``the mouse.'' Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.
-Looking more closely at these patterns, we see that the only items where errors are consistently made in the \emph{unscripted} condition are those where the more precise utterances used remain ambiguous.
-For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
-Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' but the two objects were still confusable at the sub-class level.
-If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total error rate in the unscripted condition drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.
+ \begin{figure}[t!]
+ \begin{center}
+ \vspace{-.25cm}
+ \includegraphics[scale=1]{figures/resultsFig_v1.pdf}
+ \vspace{-.5cm}
+ \caption{Listener errors with scripted and unscripted instructions. Error bars are bootstrapped 95\% confidence intervals.}
+ \vspace{-.5cm}
+ \label{fig:exp1results}
+ \end{center}
+ \end{figure}
+
+\subsubsection{Listener errors}
+
+Our scripted condition successfully replicated the results of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} with even stronger effects: listeners attempted to move the hidden object on approximately 50\% of critical trials.
+However, listeners attempted to move the hidden object less than half as often when their partner was able to produce natural referring expressions, $p_1 = 0.51, p_2 = 0.20, \chi^2(1) = 48.5, p < 0.001$ (Fig.
+\ref{fig:exp1results}A).
+We also found substantial heterogeneity in errors rates across items, which diverge significantly from a uniform distribution in both scripted ($\chi^2=55, p < 0.001$) and unscripted ($\chi^2=36, p <0.001$) conditions under a non-parametric $\chi^2$ test (Fig.
+\ref{fig:exp1results}B).
+Still, listeners in the unscripted condition made fewer errors for nearly every item, some more dramatically than others.
+A maximal mixed-effects logistic regression with random intercepts for each dyad and random slopes and intercepts for each item shows a significant effect of condition ($z = 2.856, p = 0.004$).
+%For items where participants
+%For example, 75\% of participants made an error on item 6 upon hearing the scripted instruction ``the eraser'' while only 17\% of participants in item 8 made an error upon hearing ``the mouse.''
 
@@
 
-\subsubsection{Mouse-tracking}
-
-As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis.
-Mouse-position is highly correlated with gaze in web browsing \citep{ChenEtAl01_MouseCursor, SpiveyGrosjeanKnoblich05} and has been used  %We define the decision window as the span of time between  their instruction message and when they started moving an object.
+Even if participants in the unscripted condition make fewer actual errors, they may still be \emph{considering} the hidden object just as often on trials where they go on to make correct responses.
+As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we therefore conducted a mouse-tracking analysis.
+Mouse position is highly correlated with gaze in web browsing \citep{ChenEtAl01_MouseCursor, SpiveyGrosjeanKnoblich05,RoddenEtAl08_EyeMouseCoordinationWebSearch} but precise timing can be difficult to measure.
+So we chose as our dependent variable the proportion of critical instructions that participants hovered over (or `fixated on') a hidden distractor.
+We found that listeners in the scripted condition hovered over the hidden distractor on over 30\% of critical trials relative to a baseline of 10\% when the hidden object was an unrelated filler, again replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+In the unscripted condition, however, there was no difference from the baseline, giving rise to a significant interaction in a maximal mixed-effect model with both dyad-level and object-level random effects ($z = -2.7, p = 0.006$; Fig.
+\ref{fig:exp1results}C).
+%predicted that even when participants went on to make a correct decision, they would be
+%We define the decision window as the span of time between  their instruction message and when they started moving an object.
 If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt.
@@
 
-\subsubsection{Informativity}
-
-By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items.
-Informally, we noted a tendency toward additional, possibly unnecessary, precision in descriptions.
-Instead of ``move the stuffed animal down'', participants said ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.''
-Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions.
+\subsubsection{Speaker informativity}
+
+Next, we test whether these improvements in listener performance are in fact due to more informative speaker behavior in the unscripted condition.
+%By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items.
+%First, on filler trials in both conditions, we noted a tendency toward additional, possibly unnecessary, precision in descriptions.
+For example, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.''
+%Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions.
 %If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells).
 %That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
-We therefore expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
-
-Next, we test whether these improvements in performance are in fact due to more informative speaker behavior.
-We recruited twenty judges on Amazon Mechanical Turk, who provided ratings for how well the 71 unique labels used by speakers across both experiments (including scripted labels) fit the target and hidden distractor objects.
-Their responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.6$ (95\% CI = $[0.54, 0.66]$).
+%We therefore expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
+We extracted the referring expressions generated by speakers in all critical trials and standardized spelling and grammar, yielding 122 unique labels after including scripted utterances.
+We then recruited XXX judges on Amazon Mechanical Turk to provide ratings for how well these labels fit the target and hidden distractor objects.
+\todo[inline]{Update with details once finalized} Responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of YYY (95\% CI = ZZZ).
 In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig.
-\ref{fig:fitnessInteraction}), $b = 0.32, t = 9, p < 0.001$.
-In Expt.~1, the scripted label fit the hidden distractor just as well or better than the target, but in Expt.~2, the unconstrained labels fit the target much better and the hidden distractor much worse.
-In other words, the scripted labels used in Keysar's studies were less informative than speakers normally produce in this scenario.
+\ref{fig:exp1results}D), $b = AAA, t = BBB, p < CCC$.
+The scripted label fit the hidden distractor just as well or better than the target, but unscripted labels fit the target much better and the hidden distractor much worse.
+In other words, the scripted labels used in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} were less informative than speakers normally produce in this scenario.
 
@@
 Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
-
-% \begin{figure}[t!]
-% \begin{center}
-% \vspace{-.25cm}
-% \includegraphics[scale=.5]{images/fitnessInteraction.pdf}
-% \vspace{-.5cm}
-% \caption{Mean fitness ratings provided by judges on Amazon Mechanical Turk. Error bars are bootstrapped 95\% confidence intervals.}
-% \vspace{-.5cm}
-% \label{fig:fitnessInteraction}
-% \end{center}
-% \end{figure}
 
@@
 These results support the claim that listeners neglect the speaker's perspective.
+
+Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.
+Looking more closely at these patterns, we see that the only items where errors are consistently made in the \emph{unscripted} condition are those where the more precise utterances used remain ambiguous.
+For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
+Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' but the two objects were still confusable at the sub-class level.
 

### 2018-05-08 experiments/final/normingExperiment/experiment/js/experiment.js
Only show stims in context

@@ -25,11 +25,7 @@ function make_slides(f) {
       this.init_sliders();
       exp.sliderPost = {};
       var contextsentence = "How well does <strong>"+this.stim.label+"</strong> apply to this object?";
-      if(stim.context == "isolated") {
-	var objimagehtml = '<img src="'+this.stim.object.url+'" style="height:190px;">';
-      } else {
-	var objimagehtml = '<img src="'+this.stim.object.contextURL+'" style="height:190px;">';
-      }
+      var objimagehtml = '<center><img src="'+this.stim.object.contextURL+'" style="height:500px;"></center>';
       $("#contextsentence").html(contextsentence);
       $("#objectimage").html(objimagehtml);
       console.log(this);

### 2018-05-08 experiments/final/normingExperiment/experiment/js/objectSet.js

@@ -1,318 +1,15 @@
 var stimList = function() {
-  var uniqueLabels = [
-    {
-      "label": "glasses",
-      "objectSet": "1"
-    },
-    {
-      "label": "black sunglasses",
-      "objectSet": "1"
-    },
-    {
-      "label": "sunglasses",
-      "objectSet": "1"
-    },
-    {
-      "label": "shades",
-      "objectSet": "1"
-    },
-    {
-      "label": "B block",
-      "objectSet": "2"
-    },
-    {
-      "label": "toy block",
-      "objectSet": "2"
-    },
-    {
-      "label": "blue block",
-      "objectSet": "2"
-    },
-    {
-      "label": "B cube",
-      "objectSet": "2"
-    },
-    {
-      "label": "child's blue block",
-      "objectSet": "2"
-    },
-    {
-      "label": "alphabet block",
-      "objectSet": "2"
-    },
-    {
-      "label": "B box",
-      "objectSet": "2"
-    },
-    {
-      "label": "block",
-      "objectSet": "2"
-    },
-    {
-      "label": "block with the blue writing",
-      "objectSet": "2"
-    },
-    {
-      "label": "B",
-      "objectSet": "2"
-    },
-    {
-      "label": "block with the B",
-      "objectSet": "2"
-    },
-    {
-      "label": "the bluish block with a B",
-      "objectSet": "2"
-    },
-    {
-      "label": "letter block",
-      "objectSet": "2"
-    },
-    {
-      "label": "block toy",
-      "objectSet": "2"
-    },
-    {
-      "label": "block",
-      "objectSet": "2"
-    },
-    {
-      "label": "blue B block",
-      "objectSet": "2"
-    },
-    {
-      "label": "tape",
-      "objectSet": "3"
-    },
-    {
-      "label": "cassette",
-      "objectSet": "3"
-    },
-    {
-      "label": "cassette tape",
-      "objectSet": "3"
-    },
-    {
-      "label": "clear audiocassette",
-      "objectSet": "3"
-    },
-    {
-      "label": "large measure cup",
-      "objectSet": "4"
-    },
-    {
-      "label": "pyrex measuring cup",
-      "objectSet": "4"
-    },
-    {
-      "label": "pyrex",
-      "objectSet": "4"
-    },
-    {
-      "label": "measuring cup",
-      "objectSet": "4"
-    },
-    {
-      "label": "big measuring cup",
[…]

### 2018-05-08 experiments/final/normingExperiment/experiment/js/stim_list.js

@@ -1,69 +1,0 @@
-var stim_list = {
-  "pets" :    [{"s" : "Mary is at a party making polite conversation. Susan asks: ",
-                  "q" : "'Do you have a dog?'",
-                  "a_name" : "Mary",
-                  "q_entity" : "dog",
-                  "tt": "dog_conv",
-                  "a" : ["I have a dog","I have a cat", "I have an animal","I have a dalmatian"]},
-                 {"s" : "Mary knows Susan's family is coming into town and she wants to impress them with a cute pet. Susan calls Mary and asks: ",
-                  "q" : "'Do you have a dog?'",            
-                  "a_name" : "Mary",                
-                  "q_entity" : "dog",
-                  "tt": "dog_pr_any",
-                  "a" : ["I have a dog","I have a cat", "I have an animal","I have a dalmatian"]},
-                 {"s" : "Susan is visiting Mary's house for dinner, and Mary knows she is allergic to dogs. Before coming over, Susan asks: ",
-                  "q" : "'Do you have a dog?'",
-                  "a_name" : "Mary",                 
-                  "q_entity" : "dog",
-                  "tt": "dog_info",
-                  "a" : ["I have a dog","I have a cat", "I have an animal","I have a dalmatian"]}],
-    "clothes":  [{"s" : "Bob's friends roped him into going to the opera with them and want to make sure he has the appropriate clothes to wear. His friend Mark asks: ",
-                  "q": "'Do you have a shirt?'",
-                  "a_name" : "Bob",    
-                  "q_entity" : "shirt", 
-                  "tt": "shirt_info",
-                  "a" : ["I have a shirt", "I have some pants", "I have some clothing", "I have a button-down shirt"]},
-                 {"s" : "Mark is drinking coffee at a coffee shop before going on a date and accidentally spills coffee all over his shirt. He calls his friend Bob who lives close by, explains the situation, and asks: ",
-                  "q": "'Do you have a shirt?'",                 
-                  "a_name" : "Bob",                     
-                  "q_entity" : "shirt",
-                  "tt": "shirt_pr_basic",
-                  "a" : ["I have a shirt", "I have some pants", "I have some clothing", "I have a button-down shirt"]},   
-                 {"s" : "Bob's school is collecting clothes to send to those in need and Bob wants to donate. Mark, who's working at the collection, asks him: ",
-                  "q": "'Do you have a shirt?'",                 
-                  "a_name" : "Bob",                     
-                  "q_entity" : "shirt",
[…]

### 2018-05-08 experiments/final/normingExperiment/experiment/js/experiment.js
Update to use uniqueLabels

@@ -24,11 +24,12 @@ function make_slides(f) {
       this.stim = stim;
       this.init_sliders();
       exp.sliderPost = {};
-      var contextsentence = "How well does <strong>"+this.stim.label+"</strong> apply to this object?";
+      console.log(this.stim);
+      var contextsentence = "How well does <strong>"+this.stim.text+"</strong> apply to this object?";
       if(stim.context == "isolated") {
-	var objimagehtml = '<img src="'+this.stim.object.url+'" style="height:190px;">';
+	var objimagehtml = '<img src="'+this.stim.object.url+'" style="transform: scale(2.5);  transform-origin: top;height:' + this.stim.object.height + "px;width:" + this.stim.object.width + 'px;">';
       } else {
-	var objimagehtml = '<img src="'+this.stim.object.contextURL+'" style="height:190px;">';
+	var objimagehtml = '<img src="'+this.stim.object.contextURL+'" style="height:300px;">';
       }
       $("#contextsentence").html(contextsentence);
       $("#objectimage").html(objimagehtml);

### 2018-05-08 experiments/final/normingExperiment/experiment/js/objectSet.js

@@ -1,294 +1,6 @@
+// Requires importing uniqueLabels.js before this in html
+
 var stimList = function() {
-  var uniqueLabels = [
-    {
-      "label": "glasses",
-      "objectSet": "1"
-    },
-    {
-      "label": "black sunglasses",
-      "objectSet": "1"
-    },
-    {
-      "label": "sunglasses",
-      "objectSet": "1"
-    },
-    {
-      "label": "shades",
-      "objectSet": "1"
-    },
-    {
-      "label": "B block",
-      "objectSet": "2"
-    },
-    {
-      "label": "toy block",
-      "objectSet": "2"
-    },
-    {
-      "label": "blue block",
-      "objectSet": "2"
-    },
-    {
-      "label": "B cube",
-      "objectSet": "2"
-    },
-    {
-      "label": "child's blue block",
-      "objectSet": "2"
-    },
-    {
-      "label": "alphabet block",
-      "objectSet": "2"
-    },
-    {
-      "label": "B box",
-      "objectSet": "2"
-    },
-    {
-      "label": "block",
-      "objectSet": "2"
-    },
-    {
-      "label": "block with the blue writing",
-      "objectSet": "2"
-    },
-    {
-      "label": "B",
-      "objectSet": "2"
-    },
-    {
-      "label": "block with the B",
-      "objectSet": "2"
-    },
-    {
-      "label": "the bluish block with a B",
-      "objectSet": "2"
-    },
-    {
-      "label": "letter block",
-      "objectSet": "2"
-    },
-    {
-      "label": "block toy",
-      "objectSet": "2"
-    },
-    {
-      "label": "block",
-      "objectSet": "2"
-    },
-    {
-      "label": "blue B block",
-      "objectSet": "2"
-    },
-    {
-      "label": "tape",
-      "objectSet": "3"
-    },
-    {
-      "label": "cassette",
-      "objectSet": "3"
-    },
-    {
-      "label": "cassette tape",
-      "objectSet": "3"
-    },
-    {
-      "label": "clear audiocassette",
-      "objectSet": "3"
-    },
-    {
-      "label": "large measure cup",
-      "objectSet": "4"
-    },
-    {
-      "label": "pyrex measuring cup",
-      "objectSet": "4"
-    },
-    {
-      "label": "pyrex",
-      "objectSet": "4"
-    },
-    {
-      "label": "measuring cup",
-      "objectSet": "4"
-    },
-    {
[…]

### 2018-05-08 experiments/final/normingExperiment/experiment/js/uniqueLabels.js
Add unique labels

@@ -1,0 +1,490 @@
+var uniqueLabels = [
+    {
+        "text": "candle",
+        "objectSet": "7"
+    },
+    {
+        "text": "cassette",
+        "objectSet": "3"
+    },
+    {
+        "text": "mouse",
+        "objectSet": "8"
+    },
+    {
+        "text": "dark sunglasses",
+        "objectSet": "1"
+    },
+    {
+        "text": "thing that looks like a brush",
+        "objectSet": "5"
+    },
+    {
+        "text": "glass measuring cup",
+        "objectSet": "4"
+    },
+    {
+        "text": "block with a blue 'B' showing",
+        "objectSet": "2"
+    },
+    {
+        "text": "expo dry erase eraser",
+        "objectSet": "6"
+    },
+    {
+        "text": "B block",
+        "objectSet": "2"
+    },
+    {
+        "text": "measuring cup",
+        "objectSet": "4"
+    },
+    {
+        "text": "hairbrush",
+        "objectSet": "5"
+    },
+    {
+        "text": "sunglasses",
+        "objectSet": "1"
+    },
+    {
+        "text": "eraser",
+        "objectSet": "6"
+    },
+    {
+        "text": "cassette tape",
+        "objectSet": "3"
+    },
+    {
+        "text": "whiteboard eraser",
+        "objectSet": "6"
+    },
+    {
+        "text": "child's B block",
+        "objectSet": "2"
+    },
+    {
+        "text": "computer mouse",
+        "objectSet": "8"
+    },
+    {
+        "text": "glass measuring cup with the handle",
+        "objectSet": "4"
+    },
+    {
+        "text": "radio cassette",
+        "objectSet": "3"
+    },
+    {
+        "text": "block with the blue letter B facing up",
+        "objectSet": "2"
+    },
+    {
+        "text": "red candle",
+        "objectSet": "7"
+    },
+    {
+        "text": "black sunglasses",
+        "objectSet": "1"
+    },
+    {
+        "text": "clear pyrex measuring cup",
+        "objectSet": "4"
+    },
+    {
+        "text": "expo whiteboard eraser",
+        "objectSet": "6"
+    },
+    {
+        "text": "round barrel hairbrush in the bottom row",
+        "objectSet": "5"
+    },
+    {
+        "text": "3rd row block",
+        "objectSet": "2"
+    },
+    {
+        "text": "Row 1 cassette tape",
+        "objectSet": "3"
+    },
+    {
[…]

### 2018-05-08 analysis/final_analysis.Rmd

@@
 
+## Test error homogeneity
+
+```{r}
+errorRates = critTrials %>% 
+  mutate(error = 1 - correct) %>%
+  group_by(objectSet, condition, targetObject) %>%
+  summarize(errors = sum(error), nonerrors = length(error) - errors) %>%
+  ungroup()
+
+scriptedErrors <- errorRates %>% 
+  filter(condition == 'scripted') %>%
+  select(errors, nonerrors)
+chisq.test(as.matrix(rbind(scriptedErrors$errors, scriptedErrors$nonerrors)), simulate.p.value = T)
+
+unscriptedErrors <- errorRates %>% 
+  filter(condition == 'unscripted') %>%
+  select(errors, nonerrors)
+chisq.test(as.matrix(rbind(unscriptedErrors$errors, unscriptedErrors$nonerrors)), simulate.p.value = T)
+```
+
 ```{r}
@@
   tidyboot_mean(column = hovered) %>%
-  ggplot(aes(x = condition, y=empirical_stat, fill = trialType)) +
-    geom_bar(stat = 'identity', position = dodge, alpha =.5) +
+  mutate(trial = ifelse(trialType == 'base', 'baseline', 'experimental')) %>%
+  ggplot(aes(x = condition, y=empirical_stat, fill = condition, group = trial)) +
+    geom_bar(aes(alpha = trial), stat = 'identity', position = dodge) +
     geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
     theme_few() +
+    scale_alpha_manual(values=c(.25, .75)) +
     ylab('% trials hovering over distractor') +
     scale_fill_colorblind()+
-    theme(aspect.ratio = 0.5)
+    guides(fill='none') +
+    theme(aspect.ratio = 1, legend.position = 'top')
 
-ggsave('../writing/journal_manuscript/figures//hoverOverDistractor.pdf', height = 3, width = 6)
+ggsave('../writing/journal_manuscript/figures/hoverOverDistractor.pdf', height = 3, width = 3)
 ```
@@
 ```{r}
-refs <- read_csv('../data/final/experiment1/refExpressions.csv')
-scripted <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), 
-                       m = c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse'),
-                       numRawWords = c(1,2,1,3,1,1,2,1))
+scripted_refs <- c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse')
+standardized_refs <- read_csv('../data/final/experiment1/refExpressions_standardized.csv')
+length(unique(c(standardized_refs$text, scripted_refs)))
+unscripted_refs <- read_csv('../data/final/experiment1/refExpressions.csv')
+scripted.df <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), m = scripted_refs, numRawWords = c(1,2,1,3,1,1,2,1))
 refs %>% 

### 2018-05-08 experiments/final/normingExperiment/experiment/js/experiment.js
Fix some display bugs

@@ -24,6 +24,7 @@ function make_slides(f) {
       this.stim = stim;
       this.init_sliders();
       exp.sliderPost = {};
+      console.log(stim);
       var contextsentence = "How well does <strong>"+this.stim.text+"</strong> apply to this object?";
       var objimagehtml = '<center><img src="'+this.stim.object.contextURL+'" style="height:500px;"></center>';
       $("#contextsentence").html(contextsentence);

### 2018-05-08 experiments/final/normingExperiment/experiment/js/objectSet.js

@@ -137,7 +137,7 @@ var stimList = function() {
     return _.map(referents, function(referent) {
       var relevantObject = criticalItems[Number(labelObj.objectSet) - 1][referent];
 
-      return {label: labelObj.text, objectSet : labelObj.objectSet,
+      return {text: labelObj.text, objectSet : labelObj.objectSet,
 	      referent: referent, object : relevantObject};
     });
   }));

### 2018-05-08 writing/journal_manuscript/manuscript.tex
Fill our rest of exp. 1 results, discussion

@@
 As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we therefore conducted a mouse-tracking analysis.
-Mouse position is highly correlated with gaze in web browsing \citep{ChenEtAl01_MouseCursor, SpiveyGrosjeanKnoblich05,RoddenEtAl08_EyeMouseCoordinationWebSearch} but precise timing can be difficult to measure.
-So we chose as our dependent variable the proportion of critical instructions that participants hovered over (or `fixated on') a hidden distractor.
-We found that listeners in the scripted condition hovered over the hidden distractor on over 30\% of critical trials relative to a baseline of 10\% when the hidden object was an unrelated filler, again replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-In the unscripted condition, however, there was no difference from the baseline, giving rise to a significant interaction in a maximal mixed-effect model with both dyad-level and object-level random effects ($z = -2.7, p = 0.006$; Fig.
+Mouse position is highly correlated with gaze in web browsing \citep{ChenEtAl01_MouseCursor, SpiveyGrosjeanKnoblich05,RoddenEtAl08_EyeMouseCoordinationWebSearch} but precise timing can be difficult and unreliable to measure.
+So we used as our dependent variable the proportion of critical instructions that participants hovered over (or `fixated on') a hidden distractor.
+\footnote{We preregistered as our dependent variable the mean \emph{amount of time} spent hovering over hidden distractor, but because this DV was exactly zero for the majority of trials and a large, highly variable number for the rest of trials, we decided post-hoc that the mean was not as meaningful as the binarized version.
+Still, regression results are unchanged when conducted on log-transformed hover times.}
+We found that listeners in the \emph{scripted} condition hovered over the hidden distractor on over 30\% of critical trials relative to a baseline of 10\% when the hidden object was an unrelated filler, again replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+In the unscripted condition, however, there was no difference from the baseline.
+We found a significant interaction in a maximal mixed-effect model with dyad-level and object-level random intercepts and a dyad-level random effect for the difference from baseline ($z = -2.8, p = 0.004$; Fig.
 \ref{fig:exp1results}C).
@@
 %We therefore expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
-We extracted the referring expressions generated by speakers in all critical trials and standardized spelling and grammar, yielding 122 unique labels after including scripted utterances.
+The simplest measure of speaker informativity is the raw number of   words used in referring expressions.
+Compared to the scripted referring expressions, speakers in the unscripted condition use significantly more words to refer to critical objects ($b = 0.48, t = 2.5, p < 0.001$ in a mixed-effects regression on difference scores with a fixed intercept and random intercepts for object and dyads.) However, this is a very coarse measure: the short ``pyrex glass'' may be more specific than ``large measuring glass'' despite using fewer words, and more words doesn't necessarily mean more information.
+
+\todo[inline]{Update following 2 paragraphs with details once finalized} 
+For a more explicit measure, we extracted the referring expressions generated by speakers in all critical trials and standardized spelling and grammar, yielding 122 unique labels after including scripted utterances.
 We then recruited XXX judges on Amazon Mechanical Turk to provide ratings for how well these labels fit the target and hidden distractor objects.
-\todo[inline]{Update with details once finalized} Responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of YYY (95\% CI = ZZZ).
-In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig.
-\ref{fig:exp1results}D), $b = AAA, t = BBB, p < CCC$.
+Responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of YYY (95\% CI = ZZZ).
+In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (Fig.
+\ref{fig:exp1results}D, $b = AAA, t = BBB, p < CCC$).
 The scripted label fit the hidden distractor just as well or better than the target, but unscripted labels fit the target much better and the hidden distractor much worse.
@@
 
-Does differential informativity of scripted utterances account for the variability across items that we noted in Expt.~1?
-In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
+To what extent does differential informativity of utterances account for the heterogeneity in error?
+%In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
 We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$.
 This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances.
-
-After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials?
 We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
-We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$.
-Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
+We found that participants are less likely to make errors when the target fit is higher, (\emph{coefficient STAT STAT STAT}) and more likely to make errors when the distractor fit is higher, (\emph{coefficient STAT STAT STAT}).
+Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = XXX, p < YYY$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
 
@@
 In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it.
-Furthermore, even on trials where they moved the correct object, matchers were more likely to hover their cursor over the occluded cell.
-These results support the claim that listeners neglect the speaker's perspective.
-
-Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.
-Looking more closely at these patterns, we see that the only items where errors are consistently made in the \emph{unscripted} condition are those where the more precise utterances used remain ambiguous.
-For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat.
-Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' but the two objects were still confusable at the sub-class level.
-
-However, when interacting with a natural
+Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it, even on trials where they moved the correct object, matchers were more likely to hover their cursor over the occluded cell.
+Furthermore, in a novel analysis of utterance informativity, we found that patterns of errors across items are well-predicted by relative fit to the target vs.
+distractor.
+Taken together, these results support the claim that listeners neglect the speaker's perspective.
+
+At the same time, however, our \emph{speaker} results suggest a deeper explanation for this phenomenon of perspective-neglect.
+On filler trials in the scripted condition and all trials in the unscripted condition, we noted a tendency toward additional, possibly unnecessary, precision in descriptions.
+For example, instead of the scripted ``move the brush one square to the right'' when there is only one brush in their display, speakers instead gave descriptions like ``move the hairbrush with two lines on the handle to the right.''
+Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment, thus increasing the chance of miscommunication from imprecise descriptions.
 
@@
 
-In Experiment 2, we more directly test the hypothesis that speakers naturally produce more informative referring expressions in more uncertain contexts (e.g.
-contexts where objects are occluded such as those used in Director-Matcher tasks to test theory of mind use).
+In Experiment 2, we more directly test the hypothesis that speakers naturally produce more informative referring expressions in more uncertain contexts where items are occluded.
+For this experiment, we move to a more systematic stimulus domain.
+The idiosyncracies we observed among the linguistic items used in Expt.
+1 make it difficult to normatively assess pragmatic increases in informativity for a given target.
+For example, when unscripted participants used an additional modifier (`cassette tape') to describe the target tape in one item, it happened to be highly informative relative to the distractor (a roll of tape).
+Adding a similar modifier (`hair brush') to describe the target in another item, however, was ineffective because both the target and distractor are hair brushes that are confusable at the subordinate level (one is round and one is flat).
+Unsurprisingly, then, for
 

### 2018-05-13 analysis/final_analysis.Rmd
Do a bunch of bootstrapping for Expt. 1 (messy)

@@
   filter(critical == 1) %>%
-  filter(trialType == 'exp')
+  filter(trialType == 'exp') %>%
+  mutate(error = 1 - correct)
 
@@
   group_by(condition) %>%
-  mutate(error = 1 - correct) %>%
   tidyboot_mean(column = error) %>%
@@
 
-### Control for item-level error rates...
+### Control for item-level and subject-level error rates...
 
@@
 ```{r}
-diffs <- critTrials %>% 
-  mutate(error = 1 - correct) %>%
+error_diffs <- critTrials %>% 
   group_by(objectSet, condition, targetObject) %>%
@@
   mutate(errorReduction = scripted - unscripted) %>%
-  gather(condition, pctErrors, scripted, unscripted)
+  gather(condition, pctErrors, scripted, unscripted) %>%
+  right_join(critTrials) %>%
+  group_by(objectSet, condition, targetObject, errorReduction) %>%
+  tidyboot_mean(column = error)
 
-critTrials %>% 
-  left_join(diffs) %>%
-  mutate(error = 1 - correct) %>%
-  group_by(objectSet, condition, targetObject, errorReduction) %>%
-  tidyboot_mean(column = error) %>%
-  ggplot(aes(x = reorder(targetObject, -errorReduction), y = empirical_stat, color = condition, group = condition)) +
+ggplot(error_diffs, aes(x = reorder(targetObject, -errorReduction), y = empirical_stat, color = condition, group = condition)) +
     geom_point() +
@@
 ```{r}
-summary(glmer(hovered ~ trialType * condition + (1 + trialType|gameid) + (1 |objectSet), data = d.onDistractor, family = 'binomial'))
+summary(glmer(hovered ~ trialType * condition + (1 + trialType |gameid) + (1 |objectSet), data = d.onDistractor, family = 'binomial'))
+```
+Note that this is still significant for our preregistered dv (raw amount of data) -- it's just less interpretable given the distribution...
+
+```{r}
+summary(lmer(totalAmtTimeOnDistractor ~ trialType * condition + (1 + trialType|gameid) + (1 |objectSet), 
+             data = d.onDistractor %>% mutate(totalAmtTimeOnDistractor = ifelse(totalAmtTimeOnDistractor == 0, 0, log(totalAmtTimeOnDistractor)))))
 ```
@@
 
-For 6 of the 8 items, unscripted speakers had significantly more words in their referring expressions... (n.s. for the other 2, but idiosyncratic)
+Look at all of them to make sure there aren't dumb duplicates...
 
 ```{r}
-scripted_refs <- c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse')
+scripted_words <- c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse')
+scripted_refs <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), text = scripted_words, stringsAsFactors=FALSE)
+
 standardized_refs <- read_csv('../data/final/experiment1/refExpressions_standardized.csv')
-length(unique(c(standardized_refs$text, scripted_refs)))
+cat('unique utterances: ', length(unique(c(scripted_refs, standardized_refs$text))))
+unscripted_utt_counts <- standardized_refs %>% 
+  group_by(objectSet,text) %>%
+  tally() %>%
+  mutate(condition = 'unscripted')
+```
+
+Unscripted speakers had significantly more words in their referring expressions across items... (note: if you just bootstrap cis for each item, it's significant for 6 of 8...)
+
+```{r}
 unscripted_refs <- read_csv('../data/final/experiment1/refExpressions.csv')
-scripted.df <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), m = scripted_refs, numRawWords = c(1,2,1,3,1,1,2,1))
-refs %>% 
-  mutate(numRawWords = str_count(text, "\\S+")) %>%
-  group_by(objectSet) %>%
-  tidyboot_mean(numRawWords) %>%
-  left_join(scripted)
+scripted.df <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), scriptedMessage = scripted_refs, scriptedNumWords = c(1,2,1,3,1,1,2,1))
+diff.df <- unscripted_refs %>% 
+  mutate(unscriptedNumWords = str_count(text, "\\S+")) %>%
+  left_join(scripted.df, by = c('objectSet')) %>%
+  mutate(diff = unscriptedNumWords - scriptedNumWords)
+
+library(lmerTest)
+summary(lmer(diff ~ 1 + (1 | objectSet) + (1 | gameid), data= diff.df))
+```
+
+## Analyze informativity ratings
+
+First identify 4 turkers who were responding randomly...
+
+```{r}
+raw_inf_ratings <- read_csv('../informativity_ratings.csv')
+# ggplot(raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)), aes(x = response, fill = referent)) +
+#   geom_density(alpha = 0.5) +
+#   facet_wrap(~ workerid, scales = 'free_y')
+# 
+# ggplot(raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)), aes(x = slideNumber, y = response, color = referent)) + 
+#   geom_point() +
+#   facet_wrap(~ workerid)
+
+raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)) %>%
+  group_by(workerid) %>%
+  summarize(medianRT = median(rt))
+
+bad_raters <- c('956e201e2604ead2f2fa81d0ef666626', '6559a18fae791987122350d706538e90', '2301c2d66270fecb3c16cd1593e4aa0f', 'dd0463e01b62413231a9218253908a91')#, 'deec4028cd9f327085a947bac0e19aab', '766e6aa74bc2c113e29531dab30d63e9')
+```
+
+```{r}
+library(irr)
+inf_ratings <- raw_inf_ratings %>%
+  filter(!(workerid %in% bad_raters))
+
+inf_diffs <- inf_ratings %>%
+  select(-slideNumber, -rt) %>%
+  group_by(workerid, label) %>%
+  spread(referent, response) %>%
+  mutate(diff = target - distractor) %>%
+  ungroup()
+
+wideRatings <- inf_diffs %>% select(label, diff, workerid) %>% spread(workerid, diff)
+icc(wideRatings[,-c(1)],
+    model = "twoway", type = "agreement")
+
+inf_diffs_collapsed <- inf_diffs %>%
+  group_by(label) %>%
+  summarize(distractor = mean(distractor),
+            target = mean(target),
+            paired_inf = mean(diff))
+```
+
+lmer mixed-model here is complicated because of the complex & partially crossed nature of the variance in the data (i.e.
+between judgements of raters, between object sets, between utterances chosen by different speakers)  TODO: double-bootstrapping thing where you first resample 'mean' ratings for each utterance, then you resample the distribution of utterances in a condition (i.e.
+within object set), then you take a weighted average to get your 'informativity'
+
+```{r}
+## Make a df with each utterance duplicated proportionally to the times it was said
+utt_data_tmp <- scripted_refs %>% mutate(n = 68) %>% mutate(condition = 'scripted') %>%
+  group_by(text) %>% 
+  slice(rep(1:n(), each = n)) %>%
+  mutate(gameid = as.character(row_number()))
+
+utt_data <- left_join(
+  inf_diffs %>% group_by(label) %>% summarize(m = mean(diff)),
+  utt_data_tmp %>% rename(label = text)
+) %>% filter(condition == 'scripted') %>% group_by(label, objectSet, condition) %>% summarize(m = mean(m)) %>%
+  right_join(critTrials %>% filter(condition == 'scripted'), by = c('objectSet', 'condition')) %>%
+  group_by(gameid, condition, objectSet, label) %>%
+  group_by(label, objectSet, condition, gameid) %>%
+  tally() %>%
+  select(-n) %>%
+  rename(text = label) %>%
+  bind_rows(standardized_refs %>% mutate(condition = 'unscripted', objectSet = as.numeric(objectSet)))
+```
+
+```{r}
+## In the inner loop, we want to sample a 'mean' rating for each utterance
+## using same sample of *raters* for all utterances... 
+sampleRatings = function(df, indices) {
+  grouped_df <- df %>%
+    group_by(label, objectSet) %>%
+    do(data.frame(diff = mean(.$diff[indices]),
+                  target = mean(.$target[indices]),
+                  distractor = mean(.$distractor[indices])))
+  return(grouped_df)
+}
+
+## In outer loop, we want to sample 'pairs' as a unit (with their 8 utterances)
+samplePairs = function(df) {
+  uniq_ids <- unique(df$gameid)
+  new_ids <- sample(uniq_ids, length(uniq_ids), replace = TRUE)
+  resampled_ids = data.frame(gameid = new_ids,
+                             sampleid = 1:length(new_ids), 
+                             stringsAsFactors = FALSE)
+  return(left_join(resampled_ids, df, by = c('gameid')))
+}
+
+## Finally, want to sample object sets per pair
+sampleObjects = function(df, indices) {
+  new_objects <- data.frame(objectSet = df$objectSet[indices])
+  return(left_join(new_objects, df, by = c('objectSet')))
+}
+# sampleObjects(data.frame(objectSet = c(1,2,3,4,5,6,7,8), utt = c('a', 'b', 'c', 'a', 'b', 'c', 'd', 'e')),
+#               obj_sets)
+```
+
+```{r}
+## do a multi-stage bootstrap...
+## 1. resample a set of raters to get a mean for each utterance...
+## 2. sample object set indices (e.g. 5,2,3,3,5,1,8,5)
+## 3. resample pairs within each condition and then resample utterances within pairs using object set indices
+boot_samples = replicate(1000, {
+  raters = sample(1:16,16,replace = TRUE)
+  obj_sets = sample(1:8, 8, replace = TRUE)
+  resampled_ratings = sampleRatings(inf_diffs, raters)
+
+  resampled_pairs = utt_data %>%
+    group_by(condition) %>%
+    do(samplePairs(.)) %>%
+    group_by(condition, sampleid) %>%
+    do(sampleObjects(., obj_sets)) %>%
+    rename(label = text)
+
+  combined <- left_join(resampled_pairs, resampled_ratings, by = 'label') %>%
+    gather(referent, value, target,distractor,diff) %>%
+    group_by(condition, referent) %>%
+    summarize(m = mean(value))
+  diff_of_diffs <- combined %>% filter(referent == 'diff') %>% spread(condition, m) %>% mutate(diff_of_diffs = unscripted - scripted)
+  return(c((combined %>% filter(referent != 'diff'))$m, diff_of_diffs$diff_of_diffs))
+})
+```
+
+```{r}
+# Look at diff of diffs (if excludes 0, significant)
+data.frame(test_stat = boot_samples[5,]) %>%
+  summarize(m = mean(test_stat),
+            ci_upper = sort(test_stat)[length(test_stat) * 0.025],
+            ci_lower = sort(test_stat)[length(test_stat) * 0.975]) 
+```
+
+```{r}
+boot_stats = data.frame(
+  scripted_distractor = boot_samples[1,],
+  scripted_target = boot_samples[2,],
+  unscripted_distractor = boot_samples[3,],
+  unscripted_target = boot_samples[4,]
+) %>%
+  gather(condition, value) %>%
+  group_by(condition) %>%
+  summarize(m = mean(value),
+            ci_upper = sort(value)[length(value) * 0.025],
+            ci_lower = sort(value)[length(value) * 0.975]) %>%
+  separate(condition, into = c('condition', 'referent'), by = '_')
+
+dodge <- position_dodge(width=0.9)
+ggplot(boot_stats, aes(x = condition, y = m, fill = condition, alpha = referent)) +
+  geom_bar(stat= 'identity', position = 'dodge') +
+  scale_alpha_manual(values=c(.25, .75)) +
+  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) + 
+  theme_few() +
+  scale_fill_colorblind() +
+  guides(fill = 'none') +
+  ylab('utterance informativity ratings') +
+  theme(aspect.ratio = 1)
+
+ggsave('../writing/journal_manuscript/figures/informativity.pdf')
+```
+
+### How well does informativity correlate with error?
+
+now bootstrap again but within items
+
+```{r}
+boot_objectsets = replicate(1000, {
+  raters = sample(1:16,16,replace = TRUE)
+  resampled_ratings = sampleRatings(inf_diffs, raters)
+
+  resampled_pairs = utt_data %>%
+    group_by(condition) %>%
+    do(samplePairs(.)) %>%
+    rename(label = text)
+
+  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
+    gather(referent, value, target,distractor,diff) %>%
+    group_by(condition, objectSet, referent) %>%
+    summarize(m = mean(value))
+
+  return((combined %>% filter(referent == 'diff'))$m)
+})
+```
+
+Kind of messy way to aggregate these, but so it goes... (replicate doesn't return a dataframe)
+
+(note that the 'outlier' here is the scripted 'candle' trial: raters thought it was *much* more likely to refer )
+
+```{r}
+boot_objectsets_stats = data.frame(
+  condition = c(rep('scripted', 8), rep('unscripted', 8)),
+  m_rating = rep(0,16),
+  ci_upper_rating = rep(0,16),
+  ci_lower_rating = rep(0,16),
+  objectSet = rep(1:8,2)
+)
+
+for (i in 1:16) {
+  boot_row = boot_objectsets[i,]
+  boot_objectsets_stats[i,]$m_rating = mean(boot_row)
+  boot_objectsets_stats[i,]$ci_upper_rating = sort(boot_row)[length(boot_row) * 0.025]
+  boot_objectsets_stats[i,]$ci_lower_rating = sort(boot_row)[length(boot_row) * 0.975]
+}
+
+left_join(boot_objectsets_stats, error_diffs, by = c('condition', 'objectSet')) %>%
+  ggplot(aes(x = m_rating, y = mean, color = condition, group = 1)) +
+    geom_point(size = 2, shape = 16) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), alpha = .3) +
+    geom_errorbarh(aes(xmax = ci_upper_rating, xmin = ci_lower_rating), alpha =.3) +
+    geom_smooth(linetype = 'dotted', method = 'lm', colour="black", se = F) +
+    theme_few() +
+    theme(aspect.ratio = 1, legend.position = 'top') +
+    scale_color_colorblind() +
+    xlab('informativity diff (target - distractor)') +
+    ylab('% critical error')
+
+ggsave('../writing/journal_manuscript/figures/informativity_predicts_errors.pdf',
+       height = 3, width = 3, useDingbats=FALSE)
+```
+
+```{r}
+boot_correlation = replicate(1000, {
+  raters = sample(1:16,16,replace = TRUE)
+  resampled_ratings = sampleRatings(inf_diffs, raters)
+
+  resampled_pairs = utt_data %>%
+    group_by(condition) %>%
+    do(samplePairs(.)) %>%
+    rename(label = text)
+
+  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
+    select(-target,-distractor,-label) %>%
+    inner_join(critTrials, by = c('gameid', 'condition', 'objectSet')) %>%
+    group_by(objectSet, condition) %>%
+    summarize(inf = mean(diff), error = mean(error))
+
+  return(cor(combined$inf, combined$error))
+})
+
+cat('r = ', mean(boot_correlation), 
+    ', 95% CI = [', sort(boot_correlation)[length(boot_correlation) * 0.025],
+    ', ', sort(boot_correlation)[length(boot_correlation) * 0.975],']')
 ```
@@
 dodge <- position_dodge(width=0.9)
+cbPalette = c(rgb(0.8, .8, .8), rgb(.8, .3, .3),  rgb(0, .6, .4), rgb(.4, .4, .4))
 
@@
     theme(aspect.ratio = 1, text = element_text(size=20)) +
-    scale_fill_colorblind()
+    scale_fill_manual(values=cbPalette) +
+    theme(aspect.ratio = 1)
 
-ggsave('~/Downloads/aggregated_means.pdf', width = 4, height = 2)
+ggsave('../writing/journal_manuscript/figures/num_words.pdf', width = 4, height = 3)
 ```
@@
 ```{r}
+hidden.d <- d %>%
+  group_by(gameid, context, hidden) %>%
+  summarize(m = mean(numRawWords)) %>%
+  spread(hidden, m) %>%
+  mutate(diff = yes- no,
+         ratio = log(yes/no))
+
 hidden.d %>% 
@@
 ```{r}
-hidden.d <- d %>%
-  group_by(gameid, context, hidden) %>%
-  summarize(m = mean(numRawWords)) %>%
-  spread(hidden, m) %>%
-  mutate(diff = yes- no,
-         ratio = log(yes/no))
 
@@
 
-ggsave('~/Downloads/scatter.pdf')
+ggsave('../writing/journal_manuscript/figures//scatter.pdf')
 ```
@@
   ylab('% utterances with feature') +
-  scale_fill_colorblind()
+  scale_fill_manual(values=cbPalette)
 
-ggsave('~/Downloads/by-feature-analysis.pdf')
+ggsave('../writing/journal_manuscript/figures/by-feature-analysis.pdf', 
+       height = 3, width = 9)
 ```

### 2018-05-13 writing/journal_manuscript/manuscript.tex
Finish up exp. 1; put all figures in

@@
  \vspace{-.25cm}
- \includegraphics[scale=1]{figures/resultsFig_v1.pdf}
+ \includegraphics[scale=1]{figures/Exp1_listener_results.pdf}
  \vspace{-.5cm}
- \caption{Listener errors with scripted and unscripted instructions. Error bars are bootstrapped 95\% confidence intervals.}
+ \caption{Listener results for Exp.~1.
+(A) Errors distributions with scripted and unscripted instructions -- participants in the unscripted condition made significantly fewer errors.
+(B) Even when made the correct response, listeners in the scripted condition were more likely to hover over the distractor relative to baseline while the unscripted condition shows no difference.
+Error bars are bootstrapped 95\% confidence intervals.}
  \vspace{-.5cm}
- \label{fig:exp1results}
+ \label{fig:exp1listener}
  \end{center}
@@
 However, listeners attempted to move the hidden object less than half as often when their partner was able to produce natural referring expressions, $p_1 = 0.51, p_2 = 0.20, \chi^2(1) = 48.5, p < 0.001$ (Fig.
-\ref{fig:exp1results}A).
-We also found substantial heterogeneity in errors rates across items, which diverge significantly from a uniform distribution in both scripted ($\chi^2=55, p < 0.001$) and unscripted ($\chi^2=36, p <0.001$) conditions under a non-parametric $\chi^2$ test (Fig.
-\ref{fig:exp1results}B).
+\ref{fig:exp1listener}A).
+We also found substantial heterogeneity in errors rates across items, which diverge significantly from a uniform distribution in both scripted ($\chi^2=55, p < 0.001$) and unscripted ($\chi^2=36, p <0.001$) conditions under a non-parametric $\chi^2$ test (see Supplemental Fig.
+\ref{fig:error_heterogeneity}).
 Still, listeners in the unscripted condition made fewer errors for nearly every item, some more dramatically than others.
@@
 We found a significant interaction in a maximal mixed-effect model with dyad-level and object-level random intercepts and a dyad-level random effect for the difference from baseline ($z = -2.8, p = 0.004$; Fig.
-\ref{fig:exp1results}C).
+\ref{fig:exp1listener}B).
 %predicted that even when participants went on to make a correct decision, they would be
@@
 \subsubsection{Speaker informativity}
+
+ \begin{figure}[t!]
+ \begin{center}
+ \vspace{-.25cm}
+ \includegraphics[scale=1]{figures/Exp1_speaker_results.pdf}
+ \vspace{-.5cm}
+ \caption{Speaker results for Exp.~1.
+(A) While scripted utterances were judged to fit target and distractor roughly equally, speakers in the unscripted condition naturally produced utterances that fit the target much better than the distractor.
+(B) The extent to which an utterance fits the target more than the distractor is highly predictive of error rates at an item-by-item level.
+All error bars are bootstrapped 95\% confidence intervals.}
+ \vspace{-.5cm}
+ \label{fig:exp1speaker}
+ \end{center}
+ \end{figure}
 
@@
 The simplest measure of speaker informativity is the raw number of   words used in referring expressions.
-Compared to the scripted referring expressions, speakers in the unscripted condition use significantly more words to refer to critical objects ($b = 0.48, t = 2.5, p < 0.001$ in a mixed-effects regression on difference scores with a fixed intercept and random intercepts for object and dyads.) However, this is a very coarse measure: the short ``pyrex glass'' may be more specific than ``large measuring glass'' despite using fewer words, and more words doesn't necessarily mean more information.
-
-\todo[inline]{Update following 2 paragraphs with details once finalized} 
-For a more explicit measure, we extracted the referring expressions generated by speakers in all critical trials and standardized spelling and grammar, yielding 122 unique labels after including scripted utterances.
-We then recruited XXX judges on Amazon Mechanical Turk to provide ratings for how well these labels fit the target and hidden distractor objects.
-Responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of YYY (95\% CI = ZZZ).
-In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (Fig.
-\ref{fig:exp1results}D, $b = AAA, t = BBB, p < CCC$).
-The scripted label fit the hidden distractor just as well or better than the target, but unscripted labels fit the target much better and the hidden distractor much worse.
-In other words, the scripted labels used in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} were less informative than speakers normally produce in this scenario.
-
-To what extent does differential informativity of utterances account for the heterogeneity in error?
-%In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit.
-We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$.
-This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances.
-We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
-We found that participants are less likely to make errors when the target fit is higher, (\emph{coefficient STAT STAT STAT}) and more likely to make errors when the distractor fit is higher, (\emph{coefficient STAT STAT STAT}).
+Compared to the scripted referring expressions, speakers in the unscripted condition use significantly more words to refer to critical objects ($b = 0.48, t = 2.5$ in a mixed-effects regression on difference scores with a fixed intercept and random intercepts for object and dyads.) However, this is a very coarse measure: the short ``pyrex glass'' may be more specific than ``large measuring glass'' despite using fewer words, and more words doesn't necessarily mean more information.
+\todo[inline]{We can remove the above analysis for space if necessary}
+
+For a more explicit measure, we extracted the referring expressions generated by speakers in all critical trials and standardized spelling and grammar, yielding 123 unique labels after including scripted utterances.
+We recruited 20 judges on Amazon Mechanical Turk to rate whether each label was true of the target and hidden distractor objects.
+They were shown objects in the context of the full grid (with no occlusions) such that they could feasibly judge spatial or relative references like ``bottom block.'' Responses were given on a slider with endpoints labeled ``strongly agree'' and ``strongly disagree.''  After excluding 4 judges for guessing with response times $< 1s$, inter-rater reliability was relatively high, with intra-class correlation coefficient of $0.54 (95\% CI = [0.47, 0.61])$.
+Because all judges rated all utterances for both target and distractor, we could compute the \emph{informativity} of an utterance as the difference score between target and distractor responses.
+
+Our primary measure of interest is the difference in informativity across conditions.
+Due to the complexity of the multiple partially crossed levels of variability in our data, standard mixed-effects models were not appropriate.
+Instead, we used a multi-stage, non-parametric bootstrap to compute 95\% confidence intervals.
+We found that speakers in the unscripted condition systematically produced utterances with higher informativity ratings than the scripted utterances ($m = 0.5$, 95\% bootstrapped CI = $[0.23, 0.78], p < .001$; see Fig.
+\ref{fig:exp1speaker}A).
+Scripted labels fit the hidden distractor just as well or better than the target, but unscripted labels fit the target much better and the hidden distractor much worse.
+In other words, the scripted labels used in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} were less informative than speakers normally produce in this context.
+
+Finally, we consider the link between our speaker and listener results.
+If listeners have a well-tuned pragmatic expectation about the speaker's likely level of informativity, they may rationally neglect the speaker's perspective and simply pick the object that fits the description best.
+To what extent does the informativity of speaker utterances account for heterogeneity in listener errors across items in the two conditions?
+As predicted by a perspective-negligence account, we found a strong negative correlation between informativity and error rates: listeners make fewer errors when utterances are a good fit for the target relative to the distractor ($\rho = -0.8$, bootstrapped 95\% CI $= [-0.89, -0.69]$; Fig.
+\ref{fig:exp1speaker}B).
+%  We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
+We found that participants are less likely to make errors when informativity is higher.
 Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = XXX, p < YYY$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
+\todo[inline]{rxdh: I think this bootstrapped correlation stat at the objectSet-condition level makes this point well enough for our purposes (it's a pretty huge effect...), but I could also do a fancier thing boostrapping logistic coefficients at the trial-by-trial level if necessary.
+}
+\todo[inline]{rxdh: also, is it worth having a short appendix with more details on the multi-stage bootstrapping procedure? it's very straightforward but takes a few sentences to explain\dots}
 
@@
 \subsection{Results}
-
-Our primary is the length (in words) of their naturally produced referring expressions in different contexts.
-
-As a secondary measure (intended for use in a cognitive model), we will also annotate each utterance by the number of object features mentioned (e.g. shape, texture, color).
+ \begin{figure}[t!]
+ \begin{center}
+ \vspace{-.25cm}
+ \includegraphics[scale=1.25]{figures/Exp2_results.pdf}
+ \vspace{-.5cm}
+ \caption{Results for Exp.~2.
+(A) Speakers used significantly more words when occlusions were present, and effect is even greater than the standard pragmatic effect of being in a close vs.
+far context (B) Utterances broken out by feature mentioned.
+All error bars are bootstrapped 95\% confidence intervals.}
+ \vspace{-.5cm}
+ \label{fig:exp2results}
+ \end{center}
+ \end{figure}
+
+Our primary is the length (in words) of their naturally produced referring expressions in different contexts. \ref{fig:exp2results}A
+
+As a secondary measure (intended for use in a cognitive model), we will also annotate each utterance by the number of object features mentioned (e.g. shape, texture, color). \ref{fig:exp2results}B
 

### 2018-05-14 analysis/final_analysis.Rmd
Add final expt. 2 figs & analyses

@@
 or getting less informative because tired of task?)
-4. Just length of referring expression in words?
 
@@
 confused <- unique((exp2_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
+nonNative <- unique((exp2_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)
 
@@
               '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
-badGames <- c(incompleteIDs, location_abusers, confused)
+badGames <- c(incompleteIDs, nonNative, location_abusers, confused, tabooers)
 
@@
   mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
-  filter(!(gameid %in% badGames))
+  filter(!(gameid %in% badGames)) %>%
+  mutate(occlusions = as.factor(occlusions),
+        context = as.factor(context))
 ```
 
-## Check basic (no-hidden) pragmatic effect
+## Descriptive stats
+
+How many recruited, how many excluded, for what reasons, etc?
+
+```{r}
+paste0('recruited ', length(exp2_subjInfo$gameid))
+paste0('total games ', length(unique(d_annotated$gameid)))
+paste0('incomplete ', length(incompleteIDs))
+paste0('confused ', length(setdiff(confused,incompleteIDs)))
+paste0('non-native english ', length(setdiff(nonNative, union(confused, incompleteIDs))))
+paste0('violated instructions ', length(setdiff(c(location_abusers, tabooers), union(nonNative, union(confused, incompleteIDs)))))
+paste0('final sample ', length(unique(d$gameid)))
+```
+
+## Quick listener analysis: how common are errors?
+
+Errors rare...
+
+```{r}
+exp2_clicks <- read_tsv('../data/final/experiment2/planned_sample/clickedObj/clickedObj.csv') %>%
+  mutate(correct = ifelse(correct == 'true', 1, 0)) %>%
+  filter(!(gameid %in% badGames))
+
+## Errors by condition
+tmp <- exp2_clicks %>%
+  group_by(context, occlusions) %>%
+  summarize(m = sum(correct), total = length(correct) - m)
+
+chisq.test(t(tmp[,3:4]))
+
+## Error distribution by pair
+exp2_clicks %>%
+  group_by(gameid) %>%
+  summarize(numErrors = 24- sum(correct)) %>%
+  group_by(numErrors) %>%
+  tally()
+```
+
+## Visualize basic (no-hidden) pragmatic effect
 
@@
          ratio = log(close / far))
+
+summary(lmer(numRawWords ~ context + (1 + context | gameid), d %>% filter(hidden == 'no')))
 
@@
 
-## Message length collapsing across all subjects
+## Visualize message length collapsing across all subjects
 
@@
 
-## Scatter to look at within-subj effects
+## Visualize individual differences
 
@@
          ratio = log(yes/no)) 
-
-hidden.d %>% 
-  filter(context == 'far') %>% 
-  mutate(group = case_when(diff < -0 ~ 'anti',
-                           diff > 0 ~ 'pro',
-                           TRUE ~ 'all_the_things')) %>%
-  group_by(group) %>% tally()
-
-all_the_things_ers <- (hidden.d %>% filter(context =='far') %>% filter(diff == 0))$gameid
 ```
@@
 ```{r}
-
 ggplot(hidden.d, aes(x = no, y = yes)) +
@@
 
+Look for population who mentioned all the things all the time...
+
+```{r}
+hidden.d %>% 
+  filter(context == 'far') %>% 
+  mutate(group = case_when(diff < -0 ~ 'anti',
+                           diff > 0 ~ 'pro',
+                           TRUE ~ 'all_the_things')) %>%
+  group_by(group) %>% tally()
+
+all_the_things_ers <- (hidden.d %>% filter(context =='far') %>% filter(diff == 0))$gameid
+```
+
 ```{r}
@@
 ```{r}
-summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
-             data = d %>% 
-               filter(context == 'far') %>%
-               filter(!(gameid %in% all_the_things_ers))
+# summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
+#              data = d %>% filter(context == 'far') 
 ))
-t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far') , var.equal = FALSE)
+#t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far') , var.equal = FALSE)
 ```
@@
 ```{r}
-summary(lmer(numRawWords ~ context * hidden + (1 + context * hidden | gameid), data = d %>% filter(!(gameid %in% all_the_things_ers))))
-```
+contrasts(d$context) <- contr.treatment(2, base = 2)
+contrasts(d$occlusions) <- contr.treatment(2, base = 2)
 
-helmert coding way (one way of slicing)
-
-```{r}
-d.regression <- d %>% 
-  unite(condition, context, hidden) %>%
-  ungroup() %>%
-  mutate(condition = ordered(condition, levels = c('far_no', 'close_no', 'far_yes', 'close_yes')))
-contrasts(d.regression$condition) <- contr.helmert(4)
-summary(lmer(numRawWords ~ condition + (1 + condition | gameid), data = d.regression))
+summary(lmer(numRawWords ~ context * occlusions + (1 + context * occlusions | gameid), data = d ))
 ```
@@
 
-Note: Random intercept doesn't converge (less variance in these numbers...)
+Note: Random interaction doesn't converge (less variance in these numbers...)
 
 ```{r}
-summary(lmer(numFeatures ~ context * occlusions + (1 + context + occlusions | gameid), data = d))
 ```
@@
 ```{r}
-summary(glmer(colorMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
+contrasts(d$context) <- contr.treatment(2, base = 2)
+contrasts(d$occlusions) <- contr.treatment(2, base = 2)
+summary(glmer(colorMention ~ occlusions * context + (1 + occlusions | gameid), 
+              family = 'binomial', data = d))
 ```
@@
 ```{r}
-summary(glmer(textureMention ~ occlusions + (1 + occlusions | gameid), family = 'binomial', data = d))
+summary(glmer(textureMention ~ occlusions * context + (1 + occlusions | gameid), 
+              family = 'binomial', data = d))
+summary(tmp2)
+fixef(tmp2)
+baseprob = fixef(tmp2)[1] 
+cat('no-occlusion:', exp(baseprob) / (1 + exp(baseprob)))
+occludedprob = fixef(tmp2)[1] + fixef(tmp2)[2]
+cat('with occlusion:', exp(occludedprob) / (1 + exp(occludedprob)))
 ```
@@
 
-ggplot(d.booted, aes(x = context, y = empirical_stat, fill = hidden)) + 
+ggplot(d.booted, aes(x = context, y = mean, fill = hidden)) + 
   geom_bar(stat = 'identity', position = dodge) +

### 2018-05-14 writing/journal_manuscript/manuscript.tex
Add expt. 2 results

@@
 
-Finally, we consider the link between our speaker and listener results.
-If listeners have a well-tuned pragmatic expectation about the speaker's likely level of informativity, they may rationally neglect the speaker's perspective and simply pick the object that fits the description best.
-To what extent does the informativity of speaker utterances account for heterogeneity in listener errors across items in the two conditions?
-As predicted by a perspective-negligence account, we found a strong negative correlation between informativity and error rates: listeners make fewer errors when utterances are a good fit for the target relative to the distractor ($\rho = -0.8$, bootstrapped 95\% CI $= [-0.89, -0.69]$; Fig.
+Finally, we consider the link between the speaker and listener behavior we observed across condition.
+To what extent does the informativity of speaker utterances account for the substantial variance across items we observed in listener errors in the two conditions?
+While a naive perspective-taking account would predict no relationship, a perspective-negligence account predicts that listeners simply pick the object that fits the description best regardless of whether it is hidden or not.
+Indeed, we found a strong negative correlation between informativity and error rates: listeners make fewer errors when utterances are a better fit for the target relative to the distractor ($\rho = -0.8$, bootstrapped 95\% CI $= [-0.89, -0.69]$; Fig.
 \ref{fig:exp1speaker}B).
@@
 
-At the same time, however, our \emph{speaker} results suggest a deeper explanation for this phenomenon of perspective-neglect.
-On filler trials in the scripted condition and all trials in the unscripted condition, we noted a tendency toward additional, possibly unnecessary, precision in descriptions.
-For example, instead of the scripted ``move the brush one square to the right'' when there is only one brush in their display, speakers instead gave descriptions like ``move the hairbrush with two lines on the handle to the right.''
-Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment, thus increasing the chance of miscommunication from imprecise descriptions.
+At the same time, however, our \emph{speaker} results suggest a deeper explanation for the phenomenon of perspective-neglect.
+We found a tendency toward additional, possibly unnecessary, informativity in speaker descriptions.
+For example, instead of the scripted ``move the \emph{brush} one square to the right'' which would be sufficient with only one brush in their display, speakers instead chose to give descriptions like ``move the \emph{hairbrush with two lines on the handle} to the right.''
+Perhaps speakers were taking the time to make more precise descriptions than required by their own immediate display because they believed it was contextually relevant. 
+In particular, speakers know that there additional objects of unknown identity in their partner's field of view, potentially increasing the risk of miscommunication from imprecise descriptions. 
+If natural speakers do adaptively produce more informative descriptions in context, then pragmatic listeners expecting speakers to do so may be justified in weighting considerations of the speaker's visual perspective less strongly.
+
+\subsection{Experiment~2: Production in uncertain contexts}
+\label{sec:Exp2}
+
+In Experiment 2, we more directly test the hypothesis that speakers adaptively increase the informativity of their referring expressions in more \emph{uncertain} contexts where items are occluded.
+The idiosyncracies we observed among the linguistic items used in Expt.
+1 make it difficult to normatively assess pragmatic `increases' in informativity for a given target.
+For example, when unscripted participants used an additional modifier (`cassette tape') to describe the target tape in one item, it happened to be highly informative relative to the distractor (a roll of tape).
+Adding a similar modifier (`hair brush') to describe the target in another item, however, was ineffective because both the target and distractor are hair brushes that are confusable at the subordinate level (one is round and one is flat).
+For this experiment, then, we move to a more systematic, feature-based stimulus domain and manipulate the presence or absence of occluders.
+
+\subsection{Participants}
+
+We recruited 204 participants (102 pairs) from Amazon Mechanical Turk.
+After we removed 7 games that disconnected part-way through and 12 additional games according to our pre-registered exclusion criteria (due to being non-native English speakers, reporting confusion about the instructions, or clearly violating the instructions), we were left with a sample of 84 full games.
+
+\subsection{Materials \& Procedure}
 
@@
  \vspace{-.5cm}
- \includegraphics[scale=.25]{figures/Exp2_design}
+ \includegraphics[scale=.35]{figures/Exp2_design.pdf}
  \caption{Within-pair factorial design used in Exp.~2.
@@
 
-\subsection{Experiment~2: Production in uncertain contexts}
-\label{sec:Exp2}
-
-In Experiment 2, we more directly test the hypothesis that speakers naturally produce more informative referring expressions in more uncertain contexts where items are occluded.
-For this experiment, we move to a more systematic stimulus domain.
-The idiosyncracies we observed among the linguistic items used in Expt.
-1 make it difficult to normatively assess pragmatic increases in informativity for a given target.
-For example, when unscripted participants used an additional modifier (`cassette tape') to describe the target tape in one item, it happened to be highly informative relative to the distractor (a roll of tape).
-Adding a similar modifier (`hair brush') to describe the target in another item, however, was ineffective because both the target and distractor are hair brushes that are confusable at the subordinate level (one is round and one is flat).
-Unsurprisingly, then, for
-
-\subsection{Participants}
-
-We recruited 200 participants (100 pairs) from Amazon Mechanical Turk. All participants were native English speakers.
-
-\subsection{Materials \& Procedure}
-
-Stimuli were sampled from a more systematic domain than the idiosyncratic objects used in Exp.~1.
-They varied along three discrete features: \emph{shape}, \emph{texture}, and \emph{color}.
+Stimuli varied along three discrete features: \emph{shape}, \emph{texture}, and \emph{color}.
 Each feature took four discrete values, giving a space of 64 possible objects.
@@
 Each trial is sampled from a factorial 2 x 2 design manipulating (1) the presence or absence of occlusions (`hidden' vs.
-`no hidden'), and (2) the closeness of visible distractors to the target (`close' vs.
+`no hidden'), and (2) the closeness of \emph{visible} distractors to the target (`close' vs.
 `far'; see Fig.
 \ref{fig:exp2design}).
-On 'hidden' trials, two cells are covered in curtains such that only the matcher can see if there is an object in the cell; on 'no hidden' trials, all cells are in common ground.
+On `hidden' trials, two cells were covered in curtains such that only the matcher can see if there is an object in the cell; on 'no hidden' trials, all cells are in common ground.
 On 'far' trials, the target is the only object with a particular shape; on 'close' trials, there is also a visible distractor sharing the target's shape, differing only in color or texture.
-Each trial type appeared 6 times for a total of 24 trials, and the sequence of trials was pseudo-randomized such that no trial type appeared more than twice in each
-
- To test how speakers flexibly modify their informativity in response to the uncertainty posed by
+Each trial type appeared 6 times for a total of 24 trials, and the sequence of trials was pseudo-randomized such that no trial type appeared more than twice in each block of eight trials.
 
 \subsection{Results}
+
  \begin{figure}[t!]
@@
 
-Our primary is the length (in words) of their naturally produced referring expressions in different contexts. \ref{fig:exp2results}A
-
-As a secondary measure (intended for use in a cognitive model), we will also annotate each utterance by the number of object features mentioned (e.g. shape, texture, color). \ref{fig:exp2results}B
+Listener errors were rare and unstructured: 88\% of listeners made one or fewer errors, and there was no significant difference in error rates across the four conditions ($\chi^2(3) = 1.23, p = 0.74$).
+Because this experiment was designed to investigate speaker pragmatics, we focus our attention on how speaker behavior adjusted flexibly from context to context.
+
+Our primary measure for speakers is the length (in words) of naturally produced referring expressions.
+We conducted a mixed-effect regression of context and occlusion on number of words with maximal random effect structure containing intercept, slopes, and interaction.
+First, we examined the \emph{simple} effect of close vs.
+far contexts in trials with no occlusions.
+We found that speakers used significantly more words on average when there was a distractor in context that shared the same shape as the target ($b = 0.56, t = 5.1, p < 0.001$; see Fig.
+\ref{fig:exp2results}A).
+This replicates the findings of classic studies with similar designs \cite[e.g] {BrennanClark96_ConceptualPactsConversation, GrafEtAl16_BasicLevel, MonroeEtAl17_ColorsInContext}.
+Next, we turn to the simple effect of occlusion in far contexts (which are most similar to the displays used in Expt.~1).
+The average speaker used 1.25 additional words when they knew their partner could potentially see additional objects ($t = 7.5, p < 0.001$).
+Finally, we found a significant interaction ($b = -0.49, t = 3.8, p <0.001$) where the effect of occlusion was larger in far contexts, likely indicating a ceiling on the level of informativity possible in our simple stimulus space.
+
+What are these additional words spent on describing?
+As a secondary analysis, we annotated each utterance by which object features were mentioned (shape, texture, color).
+Speakers nearly always mentioned shape  (e.g.
+`star', `triangle') as the head noun of their referring expression regardless of context, so differences in words across conditions must be due to providing additional information about color and texture.
+We ran separate mixed-effect logistic regressions for color and texture predicting mention from context; due to well-known convergence issues in logistic models, the maximum random effect structure supported by our data contains only speaker-level intercepts and slopes for occlusion.
+In both cases, we found simple effects of occlusion in far contexts ($b = 1.33, z = 2.9, p = 0.004$ for color; $b = 2.1, z = 5.7, p < 0.001$ for texture, see Fig.
+\ref{fig:exp2results}B).
+In other words, in displays like the one in Fig.
+\ref{fig:exp2results}A where the target was the only star, speakers were slightly more likely to produce the star's color---and substantially more likely to produce its texture---when there were occlusions present even though the visible objects were exactly the same.
+
+\todo[inline]{Say something about the small population of 10 or so speakers who adopted the strategy of mentioning every feature every time?}
 

### 2018-05-14 writing/journal_manuscript/manuscript.tex
Finish full draft

@@
 Recent debates over adult theory of mind use have hinged on natural-language communication tasks with knowledge asymmetries, but have largely neglected the nuanced pragmatics of these tasks.
-These pragmatic considerations .
 In particular, using occluded objects to create perspective asymmetries may have unexpected side effects under pragmatic reasoning.
@@
 %While these results are compelling, the paradigm has been criticized from few different angles. 
-%\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
+%\cite{} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g.
 the hidden roll of tape vs.
@@
 %We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
-%Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
+%Moreover,  argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.
 
@@
 This left 127 pairs in our final sample.
-\todo[inline]{rxdh: also mention removal of 3 games only using location or using 'taboo' utterances.}
+\todo[inline]{TODO: also mention removal of 3 games only using location or using 'taboo' utterances (i.e.
+cryptic clues).}
 
@@
 
+%While a naive perspective-taking account would predict no relationship, a perspective-negligence account predicts that listeners simply pick the object that fits the description best regardless of whether it is hidden or not.
 Finally, we consider the link between the speaker and listener behavior we observed across condition.
 To what extent does the informativity of speaker utterances account for the substantial variance across items we observed in listener errors in the two conditions?
-While a naive perspective-taking account would predict no relationship, a perspective-negligence account predicts that listeners simply pick the object that fits the description best regardless of whether it is hidden or not.
-Indeed, we found a strong negative correlation between informativity and error rates: listeners make fewer errors when utterances are a better fit for the target relative to the distractor ($\rho = -0.8$, bootstrapped 95\% CI $= [-0.89, -0.69]$; Fig.
+We found a strong negative correlation between informativity and error rates: listeners make fewer errors when utterances are a better fit for the target relative to the distractor ($\rho = -0.8$, bootstrapped 95\% CI $= [-0.89, -0.69]$; Fig.
 \ref{fig:exp1speaker}B).
+This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness rather than consideration of occluded vs.
+mutually visible items.
  %  We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID.
@@
 In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it, even on trials where they moved the correct object, matchers were more likely to hover their cursor over the occluded cell.
-Furthermore, in a novel analysis of utterance informativity, we found that patterns of errors across items are well-predicted by relative fit to the target vs.
+Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it.
+And even on trials where they moved the correct object, matchers were more likely to hover their cursor over the occluded cell.
+Furthermore, by considering utterance informativity, we found that patterns of errors across items are well-predicted by relative fit to the target vs.
 distractor.
-Taken together, these results support the claim that listeners neglect the speaker's perspective.
+Taken together, these results support the claim that listeners neglect the speaker's perspective in some contexts.
 
@@
 In particular, speakers know that there additional objects of unknown identity in their partner's field of view, potentially increasing the risk of miscommunication from imprecise descriptions. 
-If natural speakers do adaptively produce more informative descriptions in context, then pragmatic listeners expecting speakers to do so may be justified in weighting considerations of the speaker's visual perspective less strongly.
+If natural speakers do adaptively produce more informative descriptions in context, then pragmatic listeners rationally expecting speakers to do so may be justified in weighting considerations of the speaker's visual perspective less strongly.
 
@@
 On hidden versions of these condition, the same set of objects is visible, but some additional squares are occluded.
-If pragmatic directors represent uncertainty over what the matcher sees behind these squares, we predict that they will produce more informative utterances.}
+If pragmatic directors represent uncertainty over what the matcher sees behind these squares, we predict that they will produce more informative utterances.
+\todo[inline]{TODO: Use screenshots of real displays instead of cartoon.}}
  \vspace{-.5cm}
@@
 In other words, in displays like the one in Fig.
-\ref{fig:exp2results}A where the target was the only star, speakers were slightly more likely to produce the star's color---and substantially more likely to produce its texture---when there were occlusions present even though the visible objects were exactly the same.
+\ref{fig:exp2results}A where the target was the only `star', speakers were slightly more likely to produce the star's color---and substantially more likely to produce its texture---when there were occlusions present, even though the visible objects were exactly the same.
 
 \todo[inline]{Say something about the small population of 10 or so speakers who adopted the strategy of mentioning every feature every time?}
+
+\subsection{Discussion}
+
+We found in Expt.~1 that speakers naturally produced more informative utterances than would be necessary given the set of objects in their visual field.
+In Expt.~2 we directly tested the hypothesis that increased informativity is due to uncertainty about the listener's view imposed by the presence of occlusions.
+By manipulating occlusions, we found that speakers are not only context-sensitive in choosing referring expressions that are appropriately distinguishing in a shared context, but also \emph{occlusion-sensitive} in adaptively compensating for additional uncertainty by adding additional information.
+These forms of sensitivity are difficult to explain under an egocentric theory: speakers were spontaneously willing to spend additional time and keystrokes to give information that would be redundant given the set of objects visible in their own display.
 
@@
 
-Pragmatic language understanding requires sophisticated social reasoning. To interpret an utterance, a listener must consider how a speaker is likely to behave in context.
-The rational use of theory of mind for a listener thus depends on her expectations about the speaker: If, in a particular context, she expects a speaker to provide sufficiently informative utterances, she may be justified in neglecting his epistemic state when resolving reference.
-
-% 
-In our replication, we found empirical support for the claim that listeners neglect the speaker's epistemic state in some contexts, as claimed by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}: when the speaker's utterance fit the hidden distractor better than the target, matchers were more likely to make errors.
-Indeed, we found that the extremely heterogeneous pattern of errors across items was well explained by the relative fit of the utterance to target and distractor.
-This suggests that reference disambiguation was driven primarily by \emph{a priori} label fitness rather than consideration of occluded vs. mutually visible items.
-
-However, this finding does \emph{not} necessarily imply limitations on theory of mind.
+The rational use of theory of mind for a listener extends well beyond simple consideration of who sees what; it depends on her ability to reason holistically about the speaker's likely communicative behavior in context.
+%Pragmatic language understanding depends upon social reasoning.
+To interpret an utterance, a listener must consider how a speaker is likely to behave in context; to produce an utterance, a speaker must consider how the listener is likely to understand it.
+In this paper, we considered the hypothesis that successes in this broader Gricean application of theory of mind may ironically explain apparent failures to account for visual perspective. 
+In a direct replication, we found strong empirical support for the claim that listeners neglect the speaker's perspective in some contexts. 
+Rather than debating the existence of this phenomenon, then, we suggested an alternative, non-egocentric explanation of \emph{when} and \emph{why} listeners may neglect speaker perspective in the first place.
+In particular, while the scripted instructions in Exp.~1 systematically violated Gricean expectations of cooperativity, natural speakers did not.
+They appropriately increased the informativity of their referring expressions to account for occlusions in context.
+Our demonstrations of speaker occlusion-sensitivity thus provide justification for relying on speakers to be informative.
+
+% 1. Tying back to original expt (e.g. pitch Expt. 3 idea). 
+If visual perspective-taking is as effortful and cognitively demanding as claimed \cite{LinKeysarEpley10_ReflexivelyMindblind}, then pragmatic reasoning may play an important role in how listeners rationally allocate their resources during communication \cite{GriffithsLiederGoodman15_LevelsOfAnalysis}.
+Using theory of mind to form early expectations about how speakers are likely to behave in context allows listeners to adopt a more efficient strategy in the long-term.
+An important direction for future work, then, is to evaluate the causal role of pragmatic reasoning by directly manipulating the listener's Gricean expectations.
+We hypothesize that listeners would flexibly choose to rely more heavily on perspective information if they could not rely on their partner to be cooperative.
+In fact, one explanation for why error rates in our replication were higher than the original is the influence of the `filler' trials where speakers in the scripted condition were allowed to freely produce referring expressions.
+Many speakers were producing long, highly informative utterances in these filler trials, potentially setting up strong expectations for cooperativity that were then flouted in critical trials.
+
+%These reasons stem from more sophisticated pragmatic expectations than simple consideration of perspective, thereby opening a deeper avenue by which theory of mind use guides flexible communication.
+% if speakers could be expected to naturally provide expressions which apply better to the target, then this behavior would be appropriate.
+a listener expects a speaker to provide sufficiently informative utterances, she may be justified in neglecting his visual perspective when resolving reference.
+We predict that
+
+%However, this finding does \emph{not} necessarily imply limitations on theory of mind.
 Speakers systematically provide more informative utterances in contexts where views are uncertain, hence listeners with well-calibrated social expectations are justified in neglecting perspective.
-Rather than debating provide a principled, rational explanation of \emph{why} listeners may neglect speaker perspective in the first place.
-These reasons stem from more sophisticated pragmatic expectations than simple consideration of perspective, thereby opening a deeper avenue by which theory of mind use guides flexible communication.
+There is a long history of debate over both the methods and interpretation of the director-matcher task \cite[e.g.]{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects, HellerGrodnerTanenhaus08_Perspective, BrownSchmidtTanenhaus08_TargetedGame}.
+Our findings are particularly resonant with several recent results.
+\cite{RubioFernandez16_DirectorTaskAttention} secretly scrambled occlusions across the speaker's and listener's screens to create a mismatch in expectations about which objects are shared.
+Listeners were able to gradually infer this manipulation from the speaker's choice of referring expressions, implicating Gricean theory of mind reasoning about what is appropriate or inappropriate to say in context.
+Modeling work by \cite{HellerParisienStevenson16_ProbabilisticWeighing} (see also \cite{MozuraitisEtAl18_ProductionCombination}) has argued that speakers and listeners are probabilistically weighting their own egocentric visual perspective along with their partner's perspective.
+An open question raised by these studies is how that weighting parameter is determined in a given context.
+Our results suggest that Gricean reasoning may play an important role.
+
+%These reasons stem from more sophisticated pragmatic expectations than simple consideration of perspective, thereby opening a deeper avenue by which theory of mind use guides flexible communication. 
 % if speakers could be expected to naturally provide expressions which apply better to the target, then this behavior would be appropriate.
@@
 
-% 1.
-Tying back to original expt (e.g.
-pitch Expt.
-3 idea).
-Listeners didn't know what's going on when they heard the dramatically under-informative `tape' -- one direction for future work is to directly manipulate expectations about speakers...
-
 % 2. Other reasons for over-informativity (e.g. Judith stuff)
-
-% 3. Relationship to Heller model, Rubio-Fernandez, other critiques of Keysar?
-
-% 4. 
 %%%%%%%
 
-The Gricean maxim of quality dictates that cooperative speakers should make their utterance as informative as is required for current purposes, but no more so.
+%The Gricean maxim of quality dictates that cooperative speakers should make their utterance as informative as is required for current purposes, but no more so.
 If the referring expressions used in by confederates uniquely pick out the target from the speaker's perspective, then why our unscripted speakers be more informative than this?
 %strictly necessary and why would listeners come to expect such ``overinformativity''? 
-One possibility, which we tested in Exp.
+%One possibility, which we tested in Exp.
 2, is that awareness of the complex mismatch in epistemic state with the listener leads the speaker to provide extra information in an attempt to avoid miscommunication.
-Another possibility is that more general dynamics of language are in play.
+Finally, while we are not often in communication scenarios with such explicitly marked occluders, we are near constantly in a state of uncertainty over what is in our partner's visual context.
+The occlusion-sensitivity we observed in Exp.
+2 is representative of how pragmatic mechanisms help solve this problem.
 Studies of over-informativity in referring expressions have uncovered a general tendency of speakers to provide redundant information.
@@
 %In the current task, it seems particularly relevant that both participants knew about the presence of hidden objects, increasing uncertainty about possible distractors and placing additional pressure on the speaker to overspecify.
-The exact origin is a subject that must be followed up by future research \cite[e.g.]{GannBarr14_AudienceDesign}, but it is clear that the tendency of speakers to produce highly informative referring expressions is useful to, and relied on by, listeners.
+The exact origin of this behavior is a subject that must be followed up by future research \cite[e.g.]{GannBarr14_AudienceDesign}, but it is clear that the tendency of speakers to produce highly informative referring expressions is useful to, and relied on by, listeners.
 %Indeed, the remaining errors in Expt.~2 are at least partially due to the lack of availability of more specific expressions. 
@@
 
-It is worth noting several significant differences between our study and \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
-The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room.
+%It is worth noting several significant differences between our study and \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
+%The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room.
 We believe we addressed the major concern about exploring theory of mind in web experiments -- that participants do not truly believe they are interacting with another human -- by allowing instantaneous, responsive, real-time interaction.
@@
 
-A related difference is our decision not to use a confederate.
+%Finally, a key aspect of our approach was our decision not to use a confederate.
 While confederates are useful for reducing variation across instances of the experiment and delivering carefully targeted manipulations, their use may have unexpected consequences.
@@
 The pair of experiments we report is a reminder that manipulations administered by a trained confederate can interact in unexpected ways with a participant's social and communicative expectations.
-Regardless of the experimental context, it is illuminating to see how real participants naturally interact.
-
-\todo[inline]{mention rubio fernandez early?}
-Using similar reasoning, \cite{RubioFernandez16_DirectorTaskAttention} showed that listeners became suspicious of what was actually in common ground after hearing overinformative utterances, thus successfully using theory of mind reasoning.
-
-\todo[inline]{Mention Westra \& Carruthers as another critique focusing on pragmatics?}
-
-\todo[inline]{Mention neural coordination papers? e.g. Stephens et al 2010, Stolk et al, 2014?}
+
+\todo[inline]{End with something punchy?}
+
+%\todo[inline]{Mention neural coordination papers? e.g. Stephens et al 2010, Stolk et al, 2014?}
 

### 2018-06-27 modeling/refModule/ref.js
Update package.json and writing to file

@@ -114,24 +114,24 @@ var writeERP = function(erp, labels, filename, fixed) {
 var supportWriter = function(s, p, handle) {
   var sLst = _.toPairs(s);
   var l = sLst.length;
-
-  for (var i = 0; i < l; i++) {
-    fs.writeSync(handle, sLst[i].join(',')+','+p+'\n');
-  }
+  fs.writeSync(handle, sLst[0].join(',')+','+p+'\n');
+  // for (var i = 0; i < l; i++) {
+  //   fs.writeSync(handle, sLst[i].join(',')+','+p+'\n');
+  // }
 };
 
 // Note this is highly specific to a single type of erp
 var bayesianErpWriter = function(erp, filePrefix) {
   var predictiveFile = fs.openSync(filePrefix + "Predictives.csv", 'w');
-  fs.writeSync(predictiveFile, ["condition", "TargetColor","TargetType","Dist1Color","Dist1Type","Dist2Color","Dist2Type",
-				"value", "prob", "MCMCprob"] + '\n');
+  fs.writeSync(predictiveFile, ["close_hidden", "far_hidden","close_visible","far_visible",
+				"prob", "MCMCprob"] + '\n');
 
   var paramFile = fs.openSync(filePrefix + "Params.csv", 'w');
   fs.writeSync(paramFile, ["model", "alpha", "costWeight", "typWeight", "perspectiveWeight","omniscience", "logLikelihood", "prob"] + '\n');
 
   var supp = erp.support();
   supp.forEach(function(s) {
-    supportWriter(s.predictive, erp.score(s), predictiveFile);
+    supportWriter(s.predictives, erp.score(s), predictiveFile);
     supportWriter(s.params, erp.score(s), paramFile);
   });
   fs.closeSync(predictiveFile);

### 2018-06-27 analysis/final_analysis.Rmd
Add more model analyses, checking

@@
 or getting less informative because tired of task?)
+4. Order effects (i.e. are people more likely to use a modifier when they had to on the previous round?)
+5. Are people more informative when there are more objects?
 
@@
 
+Write out for BDA
+
+```{r}
+d %>% 
+  select(-numRawWords, -numFeatures, -text, -occlusions) %>%
+  write_csv('../modeling/bdaInput/bdaInput.csv')
+```
+
 ## Descriptive stats
@@
   tally()
+```
+
+Add missing column annotating whether critical distractor in close trials different from target in texture, color, or both...
+
+```{r}
+library(dplyr)
+bdaInput = exp2_clicks %>% 
+  mutate(names = substr(names,2,nchar(names)-1)) %>%
+  separate(names, sep = ',', into = c('item1','item2','item3', 'item4', 'item5'),
+           extra = "merge", fill = "left") %>%
+  select(-item5) %>%
+  gather(distractorNum, distractorName, item1:item4) %>%
+  separate(intendedName, into = c('targetTexture', 'targetColor', 'targetShape')) %>%
+  filter(!is.na(distractorName)) %>%
+  separate(distractorName, into = c('distractorTexture', 'distractorColor', 'distractorShape')) %>%
+  filter(distractorShape == targetShape) %>%
+  mutate(textureMatch = distractorTexture == targetTexture) %>%
+  mutate(colorMatch = distractorColor == targetColor) %>%
+  mutate(distractorType = ifelse(textureMatch, 'texture_shape', ifelse(colorMatch, 'color_shape', 'shape_only'))) %>%
+  select(gameid, trialNum, distractorType) %>%
+  right_join(exp2_clicks) %>%
+  mutate(distractorType = ifelse(is.na(distractorType), 'diff_shape', distractorType)) %>%
+  right_join(d) %>%
+  select(gameid, trialNum, distractorType, context, hidden, numDistractors, colorMention, shapeMention, textureMention)
+
+write_csv(bdaInput, '../modeling/bdaInput/bdaInput.csv')
+```
+
+Now that we have these distractor types... check if results differ across what the distractor is?
+
+```{r}
+bdaInput %>%
+  group_by(distractorType, hidden) %>%
+  tidyboot_mean(column = numFeatures) %>%
+  ggplot(aes(x = distractorType, y = mean, fill = hidden)) +
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0)
+```
+
+check if results depend on number of distractors?
+
+```{r}
+bdaInput %>%
+  mutate(numVisible= numDistractors - numObjsOccluded) %>%
+  group_by(numVisible, hidden) %>%
+  tidyboot_mean(column = numFeatures) %>%
+  ggplot(aes(x = numVisible, y = mean, fill = hidden)) +
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0)
 ```
@@
 d %>% 
-  group_by(context, occlusions) %>% 
-  summarize(numFeatures = mean(numFeatures))
+  group_by(context, hidden) %>% 
+  tidyboot_mean(column = numFeatures) %>%
+  ggplot(aes(x = context, y = mean, fill = hidden)) +
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0)
 ```
@@
 ```
+
+## Model-based analyses
+
+```{r}
+library(coda)
+
+estimate_mode <- function(s) {
+  d <- density(s)
+  return(d$x[which.max(d$y)])
+}
+HPDhi<- function(s){
+  m <- HPDinterval(mcmc(s))
+  return(m["var1","upper"])
+}
+HPDlo<- function(s){
+  m <- HPDinterval(mcmc(s))
+  return(m["var1","lower"])
+}
+
+rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
+      read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'egocentric model')) %>%
+  mutate(i = row_number()) %>%
+  gather(condition, value, close_hidden:far_visible) %>%
+  group_by(condition, source) %>%
+  summarize(mean = estimate_mode(value),
+            ci_lower = round(HPDlo(value), 3),
+            ci_upper = round(HPDhi(value), 3)) %>%
+  separate(condition, into = c('context', 'hidden')) %>%
+  mutate(hidden = ifelse(hidden == 'hidden', 'yes', 'no')) %>%
+  rbind(d %>% 
+          group_by(context, hidden) %>% 
+          tidyboot_mean(column = numFeatures) %>%
+          ungroup() %>%
+          mutate(context = as.character(context)) %>%
+          mutate(source = as.character('empirical data')) %>% 
+          select(-empirical_stat, -n)) %>%
+  ggplot(aes(x = context, fill = hidden, y = mean)) +
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
+    facet_wrap(~ source) +
+    theme_few() +
+    ylab('mean # features produced') +
+    theme(aspect.ratio = 1)
+```
+
+Examine param posteriors for occlusion-sensitive model
+
+```{r}
+read.csv('../modeling/bdaOutput/occlusionSensitivityParams.csv') %>%
+  gather(parameter, value, alpha:shapeCost) %>%
+  group_by(parameter) %>%
+  summarize(mode = estimate_mode(value),
+            md_lo = round(HPDlo(value), 3),
+            md_hi = round(HPDhi(value), 3))
+```
+
+log-likelihood plot
+
+```{r}
+
+```

### 2018-06-27 analysis/final_analysis.Rmd
Make plot with pre-reg dv

@@
   summarize(totalAmtTimeOnDistractor = sum(diff)) %>%
+  mutate(totalAmtTimeOnDistractor = ifelse(totalAmtTimeOnDistractor == 0, 0, log(totalAmtTimeOnDistractor))) %>%
   mutate(hovered = totalAmtTimeOnDistractor > 0) %>%
@@
   group_by(condition, trialType) %>%
-  tidyboot_mean(column = hovered) %>%
+  tidyboot_mean(column = totalAmtTimeOnDistractor) %>%
   mutate(trial = ifelse(trialType == 'base', 'baseline', 'experimental')) %>%
@@
     scale_alpha_manual(values=c(.25, .75)) +
-    ylab('% trials hovering over distractor') +
+    ylab('(log) ms hovering over distractor') +
     scale_fill_colorblind()+
@@
 
-ggsave('../writing/journal_manuscript/figures/hoverOverDistractor.pdf', height = 3, width = 3)
+ggsave('../../pragmatics_of_perspective_taking_tex/journal_manuscript/figures/hoverOverDistractor.pdf', height = 3, width = 3)
 ```
@@
 summary(lmer(totalAmtTimeOnDistractor ~ trialType * condition + (1 + trialType|gameid) + (1 |objectSet), 
-             data = d.onDistractor %>% mutate(totalAmtTimeOnDistractor = ifelse(totalAmtTimeOnDistractor == 0, 0, log(totalAmtTimeOnDistractor)))))
+             data = d.onDistractor))
 ```
@@
 ```{r}
-
-```
+library(tidyverse)
+library(ggthemes)
+rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityParams.csv'),
+      read.csv('../modeling/bdaOutput/egocentricParams.csv')) %>%
+  group_by(source) %>%
+  summarize(maxLikelihood = max(logLikelihood)) %>%
+  ggplot(aes(x = source, y = maxLikelihood)) +
+    geom_bar(stat = 'identity') +
+    theme_few() +
+    theme(aspect.ratio = 1)
+
+ggsave("../../pragmatics_of_perspective_taking_tex/journal_manuscript/figures/likelihoodPlot.pdf")
+```

### 2018-07-24 data/final/experiment2/turk/compile.R
reorganize data to match paper organization

@@ -1,0 +1,27 @@
+library(tidyverse)
+library(jsonlite)
+
+path <- "./"
+files = dir(path, pattern = "*.json")
+
+subjInfo = data.frame()
+
+for (f in files) {
+  jf = paste0(path, f)
+  jd = read_json(jf)
+  print(jf)
+  newRow = data.frame()
+  if(!is.null(jd$answers$subject_information)) {
+    newRow = data.frame(gameid = jd$answers$subject_information$gameID,
+                        role = jd$answers$subject_information$role,
+                        gameLength = jd$answers$subject_information$totalLength,
+                        thinksHuman = ifelse(!is.null(jd$answers$subject_information$thinksHuman), jd$answers$subject_information$thinksHuman, NA),
+                        understandsInstructions =  ifelse(!is.null(jd$answers$subject_information$confused), jd$answers$subject_information$confused, NA),
+                        nativeEnglish =  ifelse(!is.null(jd$answers$subject_information$nativeEnglish), jd$answers$subject_information$nativeEnglish, NA),
+                        ratePartner = ifelse(!is.null(jd$answers$subject_information$ratePartner), jd$answers$subject_information$ratePartner, NA),
+                        comments = jd$answers$subject_information$comments,
+                        strategy = jd$answers$subject_information$strategy)
+    subjInfo = bind_rows(subjInfo, newRow)
+  }
+}
+write_csv(subjInfo, './subjInfo.csv')

### 2018-07-24 analysis/final_analysis.Rmd
Clean up notebook

--- 2018-05-14 analysis/final_analysis.Rmd
@@
 library(tidyboot)
+library(coda)
+library(irr)
+
+estimate_mode <- function(s) {
+  d <- density(s)
+  return(d$x[which.max(d$y)])
+}
+HPDhi<- function(s){
+  m <- HPDinterval(mcmc(s))
+  return(m["var1","upper"])
+}
+HPDlo<- function(s){
+  m <- HPDinterval(mcmc(s))
+  return(m["var1","lower"])
+}
 ```
@@
 
-```{r}
-raw <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv')
-exp1_messages <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv') %>%
+## Setup data
+
+### Import turk survey data
+
+```{r}
+exp1_subjInfo <- read_csv('../data/final/experiment1/planned_sample/turk/subjInfo.csv')
+```
+
+### Import message data and implement preregistered exclusions
+
+```{r}
+d_annotated <- read_csv('../data/final/experiment1/planned_sample/chatMessage/messages_with_annotations.csv')
+
+# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
+incompleteIDs <- unique((d_annotated %>% group_by(gameid) %>% 
+                           filter(length(unique(trialNum)) != 24))$gameid)
+confused <- unique((exp1_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
+nonNative <- unique((exp1_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)
+
+# Some speakers violated instructions by relying fully on location in grid 
+location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
+                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
+                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
+                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
+# Some speakers violated instructions by playing this weird taboo game where they gave riddles 
+tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
+              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
+              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
+badGames <- c(incompleteIDs, nonNative, location_abusers, confused, tabooers)
+
+d <- d_annotated %>%
+  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
+  filter(!(gameid %in% badGames)) %>%
+  mutate(occlusions = as.factor(occlusions),
+        context = as.factor(context))
+```
+
+How many recruited, how many excluded, for what reasons, etc?
+
+```{r}
+paste0('recruited ', length(exp1_subjInfo$gameid))
+paste0('total games ', length(unique(d_annotated$gameid)))
+paste0('incomplete ', length(incompleteIDs))
+paste0('confused ', length(setdiff(confused,incompleteIDs)))
+paste0('non-native english ', length(setdiff(nonNative, union(confused, incompleteIDs))))
+paste0('violated instructions ', length(setdiff(c(location_abusers, tabooers), union(nonNative, union(confused, incompleteIDs)))))
+paste0('final sample ', length(unique(d$gameid)))
+```
+
+### Import click data
+
+```{r}
+exp1_clicks <- read_tsv('../data/final/experiment1/planned_sample/clickedObj/clickedObj.csv') %>%
+  mutate(correct = ifelse(correct == 'true', 1, 0)) %>%
+  filter(!(gameid %in% badGames))
+```
+
+### Write out for modeling & bayesian data analysis
+
+Add missing column annotating whether critical distractor in close trials different from target in texture, color, or both...
+
+```{r}
+bdaInput = exp1_clicks %>% 
+  mutate(names = substr(names,2,nchar(names)-1)) %>%
+  separate(names, sep = ',', into = c('item1','item2','item3', 'item4', 'item5'),
+           extra = "merge", fill = "left") %>%
+  select(-item5) %>%
+  gather(distractorNum, distractorName, item1:item4) %>%
+  separate(intendedName, into = c('targetTexture', 'targetColor', 'targetShape')) %>%
+  filter(!is.na(distractorName)) %>%
+  separate(distractorName, into = c('distractorTexture', 'distractorColor', 'distractorShape')) %>%
+  filter(distractorShape == targetShape) %>%
+  mutate(textureMatch = distractorTexture == targetTexture) %>%
+  mutate(colorMatch = distractorColor == targetColor) %>%
+  mutate(distractorType = ifelse(textureMatch, 'texture_shape', ifelse(colorMatch, 'color_shape', 'shape_only'))) %>%
+  select(gameid, trialNum, distractorType) %>%
+  right_join(exp1_clicks) %>%
+  mutate(distractorType = ifelse(is.na(distractorType), 'diff_shape', distractorType)) %>%
+  right_join(d)
+
+write_csv(bdaInput %>% select(gameid, trialNum, distractorType, context, hidden, 
+                              numDistractors, colorMention, shapeMention, textureMention),
+  '../modeling/bdaInput/bdaInput.csv')
+```
+
+## Speaker results
+
+### Visualize message length collapsing across all subjects
+
+```{r}
+dodge <- position_dodge(width=0.9)
+cbPalette = c(rgb(0.8, .8, .8), rgb(.8, .3, .3),  rgb(0, .6, .4), rgb(.4, .4, .4))
+
+d %>%
+  group_by(context, hidden) %>%
+  tidyboot_mean(column = numRawWords) %>%
+  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
+    theme_few() +
+    theme(aspect.ratio = 1, text = element_text(size=20)) +
+    scale_fill_manual(values=cbPalette) +
+    ylab('mean # words')
+
+ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/num_words.pdf', width = 4, height = 3)
+```
+
+### Full mixed-model for effects of condition
+
+```{r}
+contrasts(d$context) <- contr.treatment(2, base = 2)
+contrasts(d$occlusions) <- contr.treatment(2, base = 2)
+
+summary(lmer(numRawWords ~ context * occlusions + (1 + context * occlusions | gameid), data = d ))
+```
+
+### Use feature data instead
+
+```{r}
+d.booted <- d %>% 
+  gather(feature, mentioned, shapeMention, colorMention, textureMention) %>%
+  group_by(context, hidden, feature) %>% 
+  tidyboot_mean(column = mentioned)
+
+d.booted %>%
+  mutate(feature = factor(feature, levels = c('shapeMention', 'colorMention', 'textureMention'))) %>%
+  ggplot(aes(x = context, y = mean, fill = hidden)) + 
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
+    facet_wrap(~ feature) + 
+    theme_few() +
+    theme(aspect.ratio = 1, text = element_text(size=20))  +
+    ylab('% utterances with feature') +
+    scale_fill_manual(values=cbPalette)
+
+ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/by-feature-analysis.pdf', 
+       height = 3, width = 9)
+```
+
+Note: Random interaction doesn't converge for full model (less variance in these numbers...)
+
+```{r}
+contrasts(d$context) <- contr.treatment(2, base = 2)
+contrasts(d$occlusions) <- contr.treatment(2, base = 2)
+summary(glmer(colorMention ~ occlusions * context + (1 + occlusions | gameid), 
+              family = 'binomial', data = d))
+```
+
+```{r}
+summary(glmer(textureMention ~ occlusions * context + (1 + occlusions | gameid), 
+              family = 'binomial', data = d))
+```
+
+## Listener analyses
+
+### How common are errors?
+
+Quite rare...
+
+```{r}
+## Errors by condition
+numErrorsByCondition <- exp1_clicks %>%
+  group_by(context, occlusions) %>%
+  summarize(totalCorrect = sum(correct), totalErrors = length(correct) - totalCorrect)
+
+chisq.test(t(numErrorsByCondition[,3:4]))
+
+## Error distribution by pair
+exp1_clicks %>%
+  group_by(gameid) %>%
+  summarize(numErrors = 24- sum(correct)) %>%
+  group_by(numErrors) %>%
+  tally()
+```
+
+## Model-based analyses
+
+Look at pattern of predictives for each model, compared to empirical number of features mentioned
+
+```{r}
+rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
+      read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'occlusion-blind model')) %>%
+  mutate(i = row_number()) %>%
+  gather(condition, value, close_hidden:far_visible) %>%
+  group_by(condition, source) %>%
+  summarize(mean = estimate_mode(value),
+            ci_lower = round(HPDlo(value), 3),
+            ci_upper = round(HPDhi(value), 3)) %>%
+  separate(condition, into = c('context', 'hidden')) %>%
+  mutate(hidden = ifelse(hidden == 'hidden', 'yes', 'no')) %>%
+  rbind(d %>% 
+          group_by(context, hidden) %>% 
+          tidyboot_mean(column = numFeatures) %>%
+          ungroup() %>%
+          mutate(context = as.character(context)) %>%
+          mutate(source = as.character('empirical data')) %>% 
+          select(-empirical_stat, -n)) %>%
+  mutate(source = factor(source, levels = c('empirical data', 'occlusion-sensitive model', 
+                                            'occlusion-blind model'))) %>%
+  ggplot(aes(x = context, fill = hidden, y = mean)) +
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
+    facet_wrap(~ source) +
+    theme_few() +
+    theme(aspect.ratio = 1, text = element_text(size=20)) +
+    ylab('mean # features produced')
+
+ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/predictives.pdf',
+       height = 5, width = 9)
+```
+
+Get MAP estimates and CIs from param posteriors for occlusion-sensitive model
+
+```{r}
+params <- read.csv('../modeling/bdaOutput/occlusionSensitivityParams.csv')
+
+params %>%
+  gather(parameter, value, alpha:shapeCost) %>%
+  group_by(parameter) %>%
+  summarize(mode = estimate_mode(value),
+            md_lo = HPDlo(value),
+            md_hi = HPDhi(value))
+```
+
+Visualize param posteriors
+
+```{r}
+numSamples = 5000
+params.samples <- params[rep(row.names(params), exp(params$prob)*numSamples), ] %>%
+    gather(parameter, value, alpha:shapeCost)
+
+answererParamPosteriors = ggplot(params.samples, aes(x=log(value)))+
+    geom_density(aes(y=..density..),
+                 data =subset(params.samples, parameter == "alpha" ),
+                 adjust = 3, alpha=.2, fill="#FF6666")+
+    geom_density(aes(y=..density..),
+                 data=subset(params.samples, parameter == "textureCost"),
+                 adjust = 3, alpha=.2, fill="#FF6666")+
+    geom_density(aes(y=..density..),
+                 data=subset(params.samples, parameter == "colorCost"),
+                 adjust = 3, alpha=.2, fill="#FF6666")+
+    geom_density(aes(y=..density..),
+                 data=subset(params.samples, parameter == "shapeCost"),
+                 adjust = 3, alpha=.2, fill="#FF6666")+
+    xlab('log parameter value')+
+    ggtitle("Parameter Posteriors (5000 samples)") +
+    facet_grid(parameter ~ . , scales = 'free') +
+    theme_few()
+
+answererParamPosteriors
+ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/appendix/figures/paramPosterior.pdf',
+       height = 4, width = 6)
+```
+
+log-likelihood plot
+
+```{r}
+rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityParams.csv'),
+      read.csv('../modeling/bdaOutput/egocentricParams.csv')) %>%
+  group_by(source) %>%
+  summarize(maxLikelihood = max(logLikelihood)) %>%
+  ggplot(aes(x = source, y = maxLikelihood)) +
+    geom_bar(stat = 'identity') +
+    theme_few() +
+    coord_cartesian(ylim=c(0, -4000)) +
+    theme(aspect.ratio = 3)
+
+ggsave("../../pragmatics_of_perspective_taking_tex/pnas_format/appendix/figures/likelihoodPlot.pdf",
+       height = 6, width = 2)
+```
+
+## Additional checks & exploratory analyses:
+
+### check if speaker results differ across what the distractor is?
+
+```{r}
+bdaInput %>%
+  group_by(distractorType, hidden) %>%
+  tidyboot_mean(column = numFeatures) %>%
+  ggplot(aes(x = distractorType, y = mean, fill = hidden)) +
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0)
+```
+
+### check if speaker results depend on number of distractors?
+
+```{r}
+bdaInput %>%
+  mutate(numVisible= numDistractors - numObjsOccluded) %>%
+  group_by(numVisible, hidden) %>%
+  tidyboot_mean(column = numFeatures) %>%
+  ggplot(aes(x = numVisible, y = mean, fill = hidden)) +
+    geom_bar(stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0)
+```
+
+### Visualize basic (no-hidden) pragmatic effect
+
+```{r}
+basic.d <- d %>%
+  filter(hidden == 'no') %>%
+  group_by(gameid, context) %>%
+  summarize(m = mean(numRawWords)) %>%
+  spread(context, m) %>%
+  mutate(diff = close- far,
+         ratio = log(close / far))
+
+ggplot(basic.d, aes(x = far, y = close)) +
+    geom_point() +
+    theme_bw() +
+    geom_abline(intercept = 0, slope = 1) +
+    ylim(1,12) +
+    xlim(1,12) +
+    theme(aspect.ratio = 1, text = element_text(size=20))
+```
+
+Aggregated to difference scores:
+
+```{r}
+tidyboot_mean(basic.d %>%ungroup(), column = diff) %>%
+  ggplot(aes(x = 'close - far', y = empirical_stat)) +
+      geom_bar(stat ='identity', width = 0.25) +
+      geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
+      theme_few()+
+      geom_hline(yintercept = 0) +
+      #geom_vline(xintercept = 0) +
+      ylim(-1,1) +
+      # xlim(-4,4) +
+      ylab('# words used') +
+    theme(aspect.ratio = 1, text = element_text(size=20))
+
+t.test(basic.d$diff)
+```
+
+### Visualize individual differences in num features
+
+Some people used exactly the same number of words on every trial...
+
+```{r}
+hidden.d <- d %>%
+  group_by(gameid, context, hidden) %>%
+  summarize(m = mean(numFeatures)) %>%
+  spread(hidden, m) %>%
+  mutate(diff = yes- no,
+         ratio = log(yes/no)) 
+```
+
+```{r}
+ggplot(hidden.d, aes(x = no, y = yes)) +
+    geom_density_2d() +
+    facet_wrap(~ context) +
+    theme_bw() +
+    geom_abline(intercept = 0, slope = 1) +
+    ylim(0,3) +
+    ylab('# words (with occlusions)') +
+    xlab('# words (without occlusions)') +
+    #ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
+    xlim(0,3) +
+    theme(aspect.ratio = 1, text = element_text(size=20))
+
+```
+
+Histogram
+
+```{r}
+hidden.d %>% 
+  #filter(!(gameid %in% all_the_things_ers)) %>% 
+  group_by(context) %>% 
+  mutate(m = mean(diff)) %>%
+  ggplot(aes(x = diff)) +
+      geom_histogram(binwidth = 0.33) +
+      theme_bw() +
+      geom_vline(xintercept = 0, linetype = 2) +
+      geom_vline(aes(xintercept = m), color = 'red') +
+      facet_wrap(~ context) +
+      # ylim(1,12) +
+      xlim(-3,3) +
+      theme(aspect.ratio = 1)
+```
+
+```{r}
+hidden.d %>%
+  group_by(context) %>%
+  summarize(m = mean(diff), se = sd(diff)/sqrt(length(diff))) %>%
+  ggplot(aes(x = context, y = m)) +
+    geom_bar(stat = 'identity', width = 0.25) +
+    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0) +
+    theme_few() +
+    geom_hline(yintercept = 0) +
+    ylim(-3, 3) +
+    theme(aspect.ratio = 1, text = element_text(size=20))  +
+    ylab('# words (occlusions - no occlusions)')
+
+```
+
+Histograms of number of features mentioned in each condition
+
+```{r}
+ggplot(d, aes(x = numFeatures)) +
+  geom_histogram(binwidth = 1) +
+  facet_grid(context ~ occlusions) +
+  theme_few()
+```
+
+```{r}
+d %>% group_by(context, occlusions) %>% 
+  mutate(numTrials = length(text)) %>% 
+  group_by(context, occlusions, colorMention, textureMention) %>% 
+  summarize(n = length(text)/mean(numTrials)) %>%
+  ggplot(aes(x = colorMention, y = textureMention, fill = n)) +
+    geom_bin2d(stat = 'identity') +
+    stat_bin2d(geom = "text", aes(label = round(n, 2))) +
+    scale_fill_gradient(low = "white", high = "red") +
+    facet_grid(occlusions ~ context) +
+    theme_few()
+```
+
+# Experiment 2
+
+## Import messages and drops
+
+```{r}
+exp2_messages <- read_csv('../data/final/experiment2/chatMessage/messagesFromMongo.csv') %>%
   mutate(critical = ifelse(critical == 'True', 1, 0)) %>%
@@
 
-## Have to post-hoc attemptNum and instructionNum bugs (got reset for next trial prematurely)
+## Have to correct for post-hoc attemptNum and instructionNum bugs (got reset for next trial prematurely)
 ## Also, remove trials where multi-drop bug happened (i.e. where clicking multiple times triggered multiple instruction resetting)
-exp1_drops <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
+exp2_drops <- read_csv('../data/final/experiment2/drop/dropsFromMongo.csv') %>%
   mutate(critical = ifelse(critical == 'True', 1, 0),
@@
 
-```{r}
-exp1_subjInfo <- read_csv('../data/final/experiment1/turk/subjInfo.csv')
-incompleteIDs <- unique((exp1_messages %>% 
+## Implement exclusion criteria
+
+First, get error rates on filler items...
+
+```{r}
+fillerErrors <- exp2_drops %>%
+  #filter(critical != 1) %>%   # only look at mistakes on noncritical (filler) items
+  filter(attemptNum == 0) %>%   # don't want to double-count people for messing the same thing up multiple times
+  group_by(gameid, correct, critical, condition) %>%
+  tally() %>%
+  ungroup() %>%
+  mutate(correct = ifelse(correct == 1, 'no_mistake', 'mistake')) %>%
+  spread(correct, n, fill = 0) %>%
+  mutate(numErrors = mistake, errorRate = mistake/ (no_mistake + mistake) )
+
+nonCriticalMistakes = (fillerErrors %>%
+  filter(critical == 0) %>%
+  filter(numErrors >= 2))$gameid
+```
+
+And make exclusions based on subj info exit survey
+
+```{r}
+exp2_subjInfo <- read_csv('../data/final/experiment2/turk/subjInfo.csv')
+incompleteIDs <- unique((exp2_messages %>% 
                            group_by(gameid, trialNum, instructionNum) %>% 
@@
 
-confused <- unique((exp1_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
-nonNative <- unique((exp1_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)
-
-nonCriticalMistakes <- unique((exp1_drops %>%
-  # only look at mistakes on noncritical (filler) items
-  filter(critical != 1) %>%
-  # don't want to double-count people for messing the same thing up multiple times
-  filter(attemptNum == 1) %>% 
-  group_by(gameid, correct) %>%
-  tally() %>%
-  # implement exclusion criteria of errors on >~10% of non-critical trials
-  filter(n > 1))$gameid)
-
-buggyTrials <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
+confused <- unique((exp2_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
+nonNative <- unique((exp2_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)
+
+buggyTrials <- read_csv('../data/final/experiment2/drop/dropsFromMongo.csv') %>%
   mutate(critical = ifelse(critical == 'True', 1, 0),
@@
 onlyLocation <- c('3222-f68d2df5-8977-4ff2-9ee9-c60f8ddb61f1', '9457-4189cc9b-a82b-4db9-828d-bf90d1987aee')
-exp1_badGames <- c(incompleteIDs, nonNative, confused, nonCriticalMistakes, taboo, onlyLocation, unique(buggyTrials$gameid))
-
-exp1_d <- exp1_drops %>% 
-  left_join(exp1_messages) %>% #, by = c('gameid', 'trialNum', 'instructionNum', 'trialType', 
+exp2_badGames <- c(incompleteIDs, nonNative, nonCriticalMistakes, confused, taboo, onlyLocation, unique(buggyTrials$gameid))
+exp2_d <- exp2_drops %>% 
+  left_join(exp2_messages) %>% #, by = c('gameid', 'trialNum', 'instructionNum', 'trialType', 
                       #         'targetObject', 'critical', 'condition', 'attemptNum')) %>%
-  filter(!(gameid %in% exp1_badGames))
+  filter(!(gameid %in% exp2_badGames))
+
 ```
@@
 ```{r}
-paste0('recruited ', length(exp1_subjInfo$gameid))
-paste0('total games ', length(unique(exp1_drops$gameid)))
+paste0('total games ', length(unique(exp2_messages$gameid)))
 paste0('incomplete ', length(incompleteIDs))
-paste0('too many mistakes ', length(setdiff(union(confused, nonCriticalMistakes), incompleteIDs)))
-paste0('non-native english ', length(setdiff(nonNative, union(confused, union(nonCriticalMistakes, incompleteIDs)))))
-paste0('final sample ', length(unique(exp1_d$gameid)))
+paste0('too many mistakes or confused ', length(setdiff(unique(c(onlyLocation, taboo, unique(buggyTrials$gameid), nonCriticalMistakes, confused)), incompleteIDs)))
+paste0('non-native english ', length(setdiff(nonNative, unique(c(onlyLocation, taboo, unique(buggyTrials$gameid), confused, nonCriticalMistakes, incompleteIDs)))))
+paste0('final sample ', length(unique(exp2_d$gameid)))
 ```
@@
 ```{r}
-exp1_d %>% 
+exp2_d %>% 
   group_by(gameid, condition) %>% 
@@
 ```{r}
-exp1_d %>% 
+exp2_d %>% 
   group_by(gameid, objectSet, trialType, condition) %>% 
@@
 
-First, look at error rates
-
-```{r}
-critTrials <- exp1_d %>% 
+### Visualize & compare error rates across conditions
+
+```{r}
+critTrials <- exp2_d %>% 
   filter(attemptNum == 0) %>%
@@
   ggplot(aes(x = pctCriticalError, fill = condition)) +
-    geom_histogram(alpha = 0.5, position = 'dodge', binwidth = 0.5)  +
+    geom_histogram(alpha = 0.75, position = 'dodge', binwidth = 0.5)  +
     facet_grid(condition ~ .) +
@@
 
-ggsave('../writing/journal_manuscript/figures/criticalErrors.pdf')
-```
-
-```{r}
-critTrials %>%
-  group_by(condition) %>%
-  tidyboot_mean(column = error) %>%
-  ggplot(aes(x = condition, y =empirical_stat, fill = condition)) +
-    geom_bar(stat = 'identity', alpha = .5) +
-    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
-    ylab('% errors') +
-    scale_fill_colorblind()+
-    theme_few() +
-    theme(aspect.ratio = 1, legend.position="none")
-
-ggsave('../writing/journal_manuscript/figures/errorMeans.pdf', height = 3, width = 3)
+ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/criticalErrors.pdf',
+       height = 5, width = 2.5)
 ```
@@
   summarize(numErrors = 4 - sum(correct),
-            numPossible = length(correct),
+            numPossible = 4,
          atLeastOnce = sum(correct) < 4,
@@
             totalPossible = sum(numPossible))
+
 errorTable
-```
-
-### Statistical tests of error reduction
-
-```{r}
 prop.test(errorTable$totalNumErrors, errorTable$totalPossible)
 ```
-
-### Control for item-level and subject-level error rates...
 
@@
 
-## Error heterogeneity across items...
+### Mouse-tracking
+
+Even on critical trials where they didn't make errors, how often did they 'consider' the distractor (i.e. hover over the distractor cell)
+
+```{r}
+d.onDistractor <- exp2_d %>% 
+  left_join(read_csv('../data/final/experiment2/mouseOverDistractor/distractorMouseFromMongo.csv') %>% 
+              mutate(critical = ifelse(critical == 'True', 1, 0))
+            ) %>%
+  select(-serverTime, -X1) %>%
+  filter(attemptNum == 0) %>%
+  filter(correct == 1) %>%
+  filter(critical == 1) %>%
+  group_by(gameid, trialNum) %>%
+  mutate(toggle = floor((row_number() - 1) / 2)) %>%
+  mutate(onOff = ifelse(is.na(onOff), 'off', onOff)) %>%
+  spread(onOff, timeElapsed) %>%
+  mutate(diff = ifelse(is.na(off), 0, off - on)) %>%
+  group_by(gameid,condition,objectSet, trialType) %>%
+  summarize(totalAmtTimeOnDistractor = sum(diff)) %>%
+  mutate(totalAmtTimeOnDistractor = ifelse(totalAmtTimeOnDistractor == 0, 0, log(totalAmtTimeOnDistractor))) %>%
+  mutate(hovered = totalAmtTimeOnDistractor > 0) %>%
+  ungroup()
+
+dodge <- position_dodge(width=0.9)
+d.onDistractor %>%
+  group_by(condition, trialType) %>%
+  tidyboot_mean(column = totalAmtTimeOnDistractor) %>%
+  mutate(trial = ifelse(trialType == 'base', 'baseline', 'experimental')) %>%
+  ggplot(aes(x = condition, y=empirical_stat, fill = condition, group = trial)) +
+    geom_bar(aes(alpha = trial), stat = 'identity', position = dodge) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
+    theme_few() +
+    scale_alpha_manual(values=c(.25, .75)) +
+    ylab('(log) ms hovering over distractor') +
+    scale_fill_colorblind()+
+    guides(fill='none') +
+    theme(aspect.ratio = 2, legend.position = 'top')
+
+ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/hoverOverDistractor.pdf', height = 5, width = 2.5)
+```
+
+This is our pregistered DV
+
+```{r}
+summary(lmer(totalAmtTimeOnDistractor ~ trialType * condition + (1 +trialType|gameid) + (1 +trialType|objectSet), 
+             data = d.onDistractor))
+```
+
+This is the 'cleaned up' proportionality measure we report in the footnote
+
+```{r}
+summary(glmer(hovered ~ trialType * condition + (1|gameid) + (1+trialType|objectSet), data = d.onDistractor, family = 'binomial'))
+```
+## Speaker analyses
+
+### Prep informativity norming stims
+
+```{r}
+refExps <- exp2_d %>% 
+  filter(condition == 'unscripted') %>% 
+  filter(critical == 1) %>% 
+  filter(attemptNum == 0) %>% 
+  ungroup() %>%
+  select(gameid, objectSet, text)
+
+write_csv(refExps, './refExpressions_raw.csv')
+```
+
+Look at all of them to make sure there aren't dumb duplicates...
+
+```{r}
+scripted_words <- c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse')
+scripted_refs <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), text = scripted_words, stringsAsFactors=FALSE)
+
+standardized_refs <- read_csv('../data/final/experiment2/refExpressions_standardized.csv')
+cat('unique utterances: ', length(unique(c(scripted_refs, standardized_refs$text))))
+standardized_refs %>% 
+  group_by(objectSet,text) %>%
+  tally() %>%
+  mutate(condition = 'unscripted')
+```
+
+### Unscripted speakers had significantly more words in their referring expressions across items...
+
+note: if you just bootstrap CIs for each item, it's significant for 6 of 8...
+
+```{r}
+unscripted_refs <- read_csv('../data/final/experiment2/refExpressions.csv')
+scripted.df <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), scriptedMessage = scripted_refs, scriptedNumWords = c(1,2,1,3,1,1,2,1))
+diff.df <- unscripted_refs %>% 
+  mutate(unscriptedNumWords = str_count(text, "\\S+")) %>%
+  left_join(scripted.df, by = c('objectSet')) %>%
+  mutate(diff = unscriptedNumWords - scriptedNumWords) %>%
+  filter(!(gameid %in% exp2_badGames))
+
+summary(lmer(diff ~ 1 + (1 | objectSet) + (1 | gameid), data= diff.df))
+```
+
+### Extract informativity ratings & compute inter-rater agreement
+
+First identify 4 turkers who were responding randomly by looking at median RT and excluding those <1 second...
+
+```{r}
+raw_inf_ratings <- read_csv('../data/final/norming/informativity_ratings.csv')
+
+raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)) %>%
+  group_by(workerid) %>%
+  summarize(medianRT = median(rt))
+
+bad_raters <- c('956e201e2604ead2f2fa81d0ef666626', '6559a18fae791987122350d706538e90', 
+                '2301c2d66270fecb3c16cd1593e4aa0f', 'dd0463e01b62413231a9218253908a91')
+```
+
+```{r}
+inf_ratings <- raw_inf_ratings %>%
+  filter(!(workerid %in% bad_raters))
+
+inf_diffs <- inf_ratings %>%
+  select(-slideNumber, -rt) %>%
+  group_by(workerid, label) %>%
+  spread(referent, response) %>%
+  mutate(diff = target - distractor) %>%
+  ungroup()
+
+wideRatings <- inf_diffs %>% select(label, diff, workerid) %>% spread(workerid, diff)
+icc(wideRatings[,-c(1)],
+    model = "twoway", type = "agreement")
+
+inf_diffs_collapsed <- inf_diffs %>%
+  group_by(label) %>%
+  summarize(distractor = mean(distractor),
+            target = mean(target),
+            paired_inf = mean(diff))
+```
+
+lmer mixed-model here is complicated because of the complex & partially crossed nature of the variance in the data (i.e.
+between judgements of raters, between object sets, between utterances chosen by different speakers)
+
+```{r}
+## Make a df with each utterance duplicated proportionally to the times it was said
+utt_data_tmp <- scripted_refs %>% mutate(n = 68) %>% mutate(condition = 'scripted') %>%
+  group_by(text) %>% 
+  slice(rep(1:n(), each = n)) %>%
+  mutate(gameid = as.character(row_number()))
+
+utt_data <- left_join(
+  inf_diffs %>% group_by(label) %>% summarize(m = mean(diff)),
+  utt_data_tmp %>% rename(label = text)
+) %>% filter(condition == 'scripted') %>% group_by(label, objectSet, condition) %>% summarize(m = mean(m)) %>%
+  right_join(exp2_d %>% filter(attemptNum == 0) %>% filter(critical == 1) %>% filter(condition == 'scripted'), by = c('objectSet', 'condition')) %>%
+  group_by(label, objectSet, condition, gameid) %>%
+  tally() %>%
+  select(-n) %>%
+  rename(text = label) %>%
+  bind_rows(standardized_refs %>% mutate(condition = 'unscripted', objectSet = as.numeric(objectSet))) %>%
+  filter(!(gameid %in% exp2_badGames))
+```
+
+```{r}
+source('./final_analysis_bootstrap_helpers.R')
+## do a multi-stage bootstrap...
+## 1. resample a set of raters to get a mean for each utterance...
+## 2. sample object set indices (e.g. 5,2,3,3,5,1,8,5)
+## 3. resample pairs within each condition and then resample utterances within pairs using object set indices
+boot_samples = replicate(1000, {
+  raters = sample(1:16,16,replace = TRUE)
+  obj_sets = sample(1:8, 8, replace = TRUE)
+  resampled_ratings = sampleRatings(inf_diffs, raters)
+  resampled_pairs = utt_data %>%
+    group_by(condition) %>%
+    do(samplePairs(.)) %>%
+    group_by(condition, sampleid) %>%
+    do(sampleObjects(., obj_sets)) %>%
+    rename(label = text)
+  combined <- left_join(resampled_pairs, resampled_ratings, by = 'label') %>%
+    gather(referent, value, target,distractor,diff) %>%
+    group_by(condition, referent) %>%
+    summarize(m = mean(value))
+  diff_of_diffs <- combined %>%
+    filter(referent == 'diff') %>%
+    spread(condition, m) %>%
+    mutate(diff_of_diffs = unscripted - scripted)
+
+  return(c((combined %>% filter(referent != 'diff'))$m, diff_of_diffs$diff_of_diffs))
+})
+```
+
+```{r}
+# Look at diff of diffs (if excludes 0, significant)
+data.frame(test_stat = boot_samples[5,]) %>%
+  summarize(m = mean(test_stat),
+            ci_upper = sort(test_stat)[length(test_stat) * 0.025],
+            ci_lower = sort(test_stat)[length(test_stat) * 0.975]) 
+```
+
+```{r}
+boot_stats = data.frame(
+  scripted_distractor = boot_samples[1,],
+  scripted_target = boot_samples[2,],
+  unscripted_distractor = boot_samples[3,],
+  unscripted_target = boot_samples[4,]
+) %>%
+  gather(condition, value) %>%
+  group_by(condition) %>%
+  summarize(m = mean(value),
+            ci_upper = sort(value)[length(value) * 0.025],
+            ci_lower = sort(value)[length(value) * 0.975]) %>%
+  separate(condition, into = c('condition', 'referent'), by = '_')
+
+dodge <- position_dodge(width=0.9)
+ggplot(boot_stats, aes(x = condition, y = m, fill = condition, alpha = referent)) +
+  geom_bar(stat= 'identity', position = 'dodge') +
+  scale_alpha_manual(values=c(.25, .75)) +
+  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) + 
+  theme_few() +
+  scale_fill_colorblind() +
+  guides(fill = 'none') +
+  ylab('utterance informativity ratings') +
+  theme(aspect.ratio = 2, legend.position = 'top')
+
+ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/informativity.pdf',
+       height = 5, width = 2.5)
+```
+
+### How well does informativity correlate with error?
+
+now bootstrap again but within items
+
+```{r}
+boot_objectsets = replicate(1000, {
+  raters = sample(1:16,16,replace = TRUE)
+  resampled_ratings = sampleRatings(inf_diffs, raters)
+
+  resampled_pairs = utt_data %>%
+    group_by(condition) %>%
+    do(samplePairs(.)) %>%
+    rename(label = text)
+
+  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
+    gather(referent, value, target,distractor,diff) %>%
+    group_by(condition, objectSet, referent) %>%
+    summarize(m = mean(value))
+
+  return((combined %>% filter(referent == 'diff'))$m)
+})
+```
+
+Kind of messy way to aggregate these, but so it goes... (replicate doesn't return a dataframe)
+
+(note that the 'outlier' here is the scripted 'candle' trial: raters thought it was *much* more likely to refer )
+
+```{r}
+boot_objectsets_stats = data.frame(
+  condition = c(rep('scripted', 8), rep('unscripted', 8)),
+  m_rating = rep(0,16),
+  ci_upper_rating = rep(0,16),
+  ci_lower_rating = rep(0,16),
+  objectSet = rep(1:8,2)
+)
+
+for (i in 1:16) {
+  boot_row = boot_objectsets[i,]
+  boot_objectsets_stats[i,]$m_rating = mean(boot_row)
+  boot_objectsets_stats[i,]$ci_upper_rating = sort(boot_row)[length(boot_row) * 0.025]
+  boot_objectsets_stats[i,]$ci_lower_rating = sort(boot_row)[length(boot_row) * 0.975]
+}
+
+left_join(boot_objectsets_stats, error_diffs, by = c('condition', 'objectSet')) %>%
+  ggplot(aes(x = m_rating, y = mean, color = condition, group = 1)) +
+    geom_point(size = 2, shape = 16) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), alpha = .3) +
+    geom_errorbarh(aes(xmax = ci_upper_rating, xmin = ci_lower_rating), alpha =.3) +
+    geom_smooth(linetype = 'dotted', method = 'lm', colour="black", se = F) +
+    theme_few() +
+    theme(aspect.ratio = 1, legend.position = 'top') +
+    scale_color_colorblind() +
+    xlab('informativity diff (target - distractor)') +
+    ylab('% critical error')
+
+ggsave('../writing/journal_manuscript/figures/informativity_predicts_errors.pdf',
+       height = 3, width = 3, useDingbats=FALSE)
+```
+
+```{r}
+boot_correlation = replicate(1000, {
+  raters = sample(1:16,16,replace = TRUE)
+  resampled_ratings = sampleRatings(inf_diffs, raters)
+
+  resampled_pairs = utt_data %>%
+    group_by(condition) %>%
+    do(samplePairs(.)) %>%
+    rename(label = text)
+
+  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
+    select(-target,-distractor,-label) %>%
+    inner_join(critTrials, by = c('gameid', 'condition', 'objectSet')) %>%
+    group_by(objectSet, condition) %>%
+    summarize(inf = mean(diff), error = mean(error))
+
+  return(cor(combined$inf, combined$error))
+})
+
+cat('r = ', mean(boot_correlation), 
+    ', 95% CI = [', sort(boot_correlation)[length(boot_correlation) * 0.025],
+    ', ', sort(boot_correlation)[length(boot_correlation) * 0.975],']')
+```
+
+## Exploratory & checks
+
+### Error heterogeneity across items (in supplemental)
 
@@
 
-ggsave('../writing/journal_manuscript/figures//itemHeterogeneity.pdf')
-```
-
-## Test error homogeneity
+ggsave('../writing/journal_manuscript/figures/itemHeterogeneity.pdf')
+```
 
@@
 
-```{r}
-exp1_d %>% 
+### Is there relationship between errors on critical and errors on filler?
+
+```{r}
+fillerErrors %>%
+  select(-mistake, -no_mistake, -numErrors) %>%
+  spread(critical, errorRate) %>%
+  ggplot(aes(x = `0`, y = `1`)) +
+    geom_point() +
+    geom_smooth(method = 'lm')
+
+fillerErrors %>%
+  ungroup() %>%
+  group_by(critical, condition) %>%
+  tidyboot_mean(column = errorRate)
+```
+
+### Visualize difference in mean error rates...
+
+```{r}
+critTrials %>%
+  group_by(condition) %>%
+  tidyboot_mean(column = error) %>%
+  ggplot(aes(x = condition, y =empirical_stat, fill = condition)) +
+    geom_bar(stat = 'identity', alpha = .5) +
+    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
+    ylab('% errors') +
+    scale_fill_colorblind()+
+    theme_few() +
+    theme(aspect.ratio = 1, legend.position="none") 
+```
+
+### Examine reaction times (too imprecise)
+
+```{r}
+exp2_d %>% 
   filter(attemptNum == 0) %>%
@@
 summary(lmer(log(responseTime) ~ trialType * condition + (1 | gameid),
-     data = exp1_d %>% 
+     data = exp2_d %>% 
         filter(attemptNum == 0) %>%
@@
 
-## Mouse-tracking
-
-Even on critical trials where they didn't make errors, how often did they 'consider' the distractor (i.e. hover over the distractor cell)
-
-```{r}
-d.onDistractor <- exp1_d %>% 
-  left_join(read_csv('../data/final/experiment1/mouseOverDistractor/distractorMouseFromMongo.csv') %>% 
-              mutate(critical = ifelse(critical == 'True', 1, 0))
-            ) %>%
-  select(-serverTime, -X1) %>%
-  filter(attemptNum == 0) %>%
-  filter(correct == 1) %>%
-  filter(critical == 1) %>%
-  group_by(gameid, trialNum) %>%
-  mutate(toggle = floor((row_number() - 1) / 2)) %>%
-  mutate(onOff = ifelse(is.na(onOff), 'off', onOff)) %>%
-  spread(onOff, timeElapsed) %>%
-  mutate(diff = ifelse(is.na(off), 0, off - on)) %>%
-  group_by(gameid,condition,objectSet, trialType) %>%
-  summarize(totalAmtTimeOnDistractor = sum(diff)) %>%
-  mutate(hovered = totalAmtTimeOnDistractor > 0) %>%
-  ungroup()
-
-dodge <- position_dodge(width=0.9)
-d.onDistractor %>%
-  group_by(condition, trialType) %>%
-  tidyboot_mean(column = hovered) %>%
-  mutate(trial = ifelse(trialType == 'base', 'baseline', 'experimental')) %>%
-  ggplot(aes(x = condition, y=empirical_stat, fill = condition, group = trial)) +
-    geom_bar(aes(alpha = trial), stat = 'identity', position = dodge) +
-    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
-    theme_few() +
-    scale_alpha_manual(values=c(.25, .75)) +
-    ylab('% trials hovering over distractor') +
-    scale_fill_colorblind()+
-    guides(fill='none') +
-    theme(aspect.ratio = 1, legend.position = 'top')
-
-ggsave('../writing/journal_manuscript/figures/hoverOverDistractor.pdf', height = 3, width = 3)
-```
-
-```{r}
-summary(glmer(hovered ~ trialType * condition + (1 + trialType |gameid) + (1 |objectSet), data = d.onDistractor, family = 'binomial'))
-```
-Note that this is still significant for our preregistered dv (raw amount of data) -- it's just less interpretable given the distribution...
-
-```{r}
-summary(lmer(totalAmtTimeOnDistractor ~ trialType * condition + (1 + trialType|gameid) + (1 |objectSet), 
-             data = d.onDistractor %>% mutate(totalAmtTimeOnDistractor = ifelse(totalAmtTimeOnDistractor == 0, 0, log(totalAmtTimeOnDistractor)))))
-```
-
 Note: because I stopped recording from the mouse as soon as someone starts dragging, it's technically possible for them to hover over the distractor object, click and start to drag it, but then drop it in place and go over to the target object.
@@
 targetRadius = 175
-d.mouse <- exp1_d %>% ungroup() %>%
-  left_join(read_csv('../data/final/experiment1/updateMouse/mouseFromMongo.csv') %>%
+d.mouse <- exp2_d %>% ungroup() %>%
+  left_join(read_csv('../data/final/experiment2/updateMouse/mouseFromMongo.csv') %>%
                   mutate(critical = ifelse(critical == 'True', 1, 0))
@@
 
-## Prep informativity norming stims
-
-```{r}
-refExps <- exp1_d %>% 
-  filter(condition == 'unscripted') %>% 
-  filter(critical == 1) %>% 
-  filter(attemptNum == 0) %>% 
-  ungroup() %>%
-  select(gameid, objectSet, text)
-
-write_csv(refExps, './refExpressions_raw.csv')
-```
-
-Look at all of them to make sure there aren't dumb duplicates...
-
-```{r}
-scripted_words <- c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse')
-scripted_refs <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), text = scripted_words, stringsAsFactors=FALSE)
-
-standardized_refs <- read_csv('../data/final/experiment1/refExpressions_standardized.csv')
-cat('unique utterances: ', length(unique(c(scripted_refs, standardized_refs$text))))
-unscripted_utt_counts <- standardized_refs %>% 
-  group_by(objectSet,text) %>%
-  tally() %>%
-  mutate(condition = 'unscripted')
-```
-
-Unscripted speakers had significantly more words in their referring expressions across items... (note: if you just bootstrap cis for each item, it's significant for 6 of 8...)
-
-```{r}
-unscripted_refs <- read_csv('../data/final/experiment1/refExpressions.csv')
-scripted.df <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), scriptedMessage = scripted_refs, scriptedNumWords = c(1,2,1,3,1,1,2,1))
-diff.df <- unscripted_refs %>% 
-  mutate(unscriptedNumWords = str_count(text, "\\S+")) %>%
-  left_join(scripted.df, by = c('objectSet')) %>%
-  mutate(diff = unscriptedNumWords - scriptedNumWords)
-
-library(lmerTest)
-summary(lmer(diff ~ 1 + (1 | objectSet) + (1 | gameid), data= diff.df))
-```
-
-## Analyze informativity ratings
-
-First identify 4 turkers who were responding randomly...
-
-```{r}
-raw_inf_ratings <- read_csv('../informativity_ratings.csv')
-# ggplot(raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)), aes(x = response, fill = referent)) +
-#   geom_density(alpha = 0.5) +
-#   facet_wrap(~ workerid, scales = 'free_y')
-# 
-# ggplot(raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)), aes(x = slideNumber, y = response, color = referent)) + 
-#   geom_point() +
-#   facet_wrap(~ workerid)
-
-raw_inf_ratings %>% mutate(workerid = substr(workerid, 0, 6)) %>%
-  group_by(workerid) %>%
-  summarize(medianRT = median(rt))
-
-bad_raters <- c('956e201e2604ead2f2fa81d0ef666626', '6559a18fae791987122350d706538e90', '2301c2d66270fecb3c16cd1593e4aa0f', 'dd0463e01b62413231a9218253908a91')#, 'deec4028cd9f327085a947bac0e19aab', '766e6aa74bc2c113e29531dab30d63e9')
-```
-
-```{r}
-library(irr)
-inf_ratings <- raw_inf_ratings %>%
-  filter(!(workerid %in% bad_raters))
-
-inf_diffs <- inf_ratings %>%
-  select(-slideNumber, -rt) %>%
-  group_by(workerid, label) %>%
-  spread(referent, response) %>%
-  mutate(diff = target - distractor) %>%
-  ungroup()
-
-wideRatings <- inf_diffs %>% select(label, diff, workerid) %>% spread(workerid, diff)
-icc(wideRatings[,-c(1)],
-    model = "twoway", type = "agreement")
-
-inf_diffs_collapsed <- inf_diffs %>%
-  group_by(label) %>%
-  summarize(distractor = mean(distractor),
-            target = mean(target),
-            paired_inf = mean(diff))
-```
-
-lmer mixed-model here is complicated because of the complex & partially crossed nature of the variance in the data (i.e.
-between judgements of raters, between object sets, between utterances chosen by different speakers)  TODO: double-bootstrapping thing where you first resample 'mean' ratings for each utterance, then you resample the distribution of utterances in a condition (i.e.
-within object set), then you take a weighted average to get your 'informativity'
-
-```{r}
-## Make a df with each utterance duplicated proportionally to the times it was said
-utt_data_tmp <- scripted_refs %>% mutate(n = 68) %>% mutate(condition = 'scripted') %>%
-  group_by(text) %>% 
-  slice(rep(1:n(), each = n)) %>%
-  mutate(gameid = as.character(row_number()))
-
-utt_data <- left_join(
-  inf_diffs %>% group_by(label) %>% summarize(m = mean(diff)),
-  utt_data_tmp %>% rename(label = text)
-) %>% filter(condition == 'scripted') %>% group_by(label, objectSet, condition) %>% summarize(m = mean(m)) %>%
-  right_join(critTrials %>% filter(condition == 'scripted'), by = c('objectSet', 'condition')) %>%
-  group_by(gameid, condition, objectSet, label) %>%
-  group_by(label, objectSet, condition, gameid) %>%
-  tally() %>%
-  select(-n) %>%
-  rename(text = label) %>%
-  bind_rows(standardized_refs %>% mutate(condition = 'unscripted', objectSet = as.numeric(objectSet)))
-```
-
-```{r}
-## In the inner loop, we want to sample a 'mean' rating for each utterance
-## using same sample of *raters* for all utterances... 
-sampleRatings = function(df, indices) {
-  grouped_df <- df %>%
-    group_by(label, objectSet) %>%
-    do(data.frame(diff = mean(.$diff[indices]),
-                  target = mean(.$target[indices]),
-                  distractor = mean(.$distractor[indices])))
-  return(grouped_df)
-}
-
-## In outer loop, we want to sample 'pairs' as a unit (with their 8 utterances)
-samplePairs = function(df) {
-  uniq_ids <- unique(df$gameid)
-  new_ids <- sample(uniq_ids, length(uniq_ids), replace = TRUE)
-  resampled_ids = data.frame(gameid = new_ids,
-                             sampleid = 1:length(new_ids), 
-                             stringsAsFactors = FALSE)
-  return(left_join(resampled_ids, df, by = c('gameid')))
-}
-
-## Finally, want to sample object sets per pair
-sampleObjects = function(df, indices) {
-  new_objects <- data.frame(objectSet = df$objectSet[indices])
-  return(left_join(new_objects, df, by = c('objectSet')))
-}
-# sampleObjects(data.frame(objectSet = c(1,2,3,4,5,6,7,8), utt = c('a', 'b', 'c', 'a', 'b', 'c', 'd', 'e')),
-#               obj_sets)
-```
-
-```{r}
-## do a multi-stage bootstrap...
-## 1. resample a set of raters to get a mean for each utterance...
-## 2. sample object set indices (e.g. 5,2,3,3,5,1,8,5)
-## 3. resample pairs within each condition and then resample utterances within pairs using object set indices
-boot_samples = replicate(1000, {
-  raters = sample(1:16,16,replace = TRUE)
-  obj_sets = sample(1:8, 8, replace = TRUE)
-  resampled_ratings = sampleRatings(inf_diffs, raters)
-
-  resampled_pairs = utt_data %>%
-    group_by(condition) %>%
-    do(samplePairs(.)) %>%
-    group_by(condition, sampleid) %>%
-    do(sampleObjects(., obj_sets)) %>%
-    rename(label = text)
-
-  combined <- left_join(resampled_pairs, resampled_ratings, by = 'label') %>%
-    gather(referent, value, target,distractor,diff) %>%
-    group_by(condition, referent) %>%
-    summarize(m = mean(value))
-  diff_of_diffs <- combined %>% filter(referent == 'diff') %>% spread(condition, m) %>% mutate(diff_of_diffs = unscripted - scripted)
-  return(c((combined %>% filter(referent != 'diff'))$m, diff_of_diffs$diff_of_diffs))
-})
-```
-
-```{r}
-# Look at diff of diffs (if excludes 0, significant)
-data.frame(test_stat = boot_samples[5,]) %>%
-  summarize(m = mean(test_stat),
-            ci_upper = sort(test_stat)[length(test_stat) * 0.025],
-            ci_lower = sort(test_stat)[length(test_stat) * 0.975]) 
-```
-
-```{r}
-boot_stats = data.frame(
-  scripted_distractor = boot_samples[1,],
-  scripted_target = boot_samples[2,],
-  unscripted_distractor = boot_samples[3,],
-  unscripted_target = boot_samples[4,]
-) %>%
-  gather(condition, value) %>%
-  group_by(condition) %>%
-  summarize(m = mean(value),
-            ci_upper = sort(value)[length(value) * 0.025],
-            ci_lower = sort(value)[length(value) * 0.975]) %>%
-  separate(condition, into = c('condition', 'referent'), by = '_')
-
-dodge <- position_dodge(width=0.9)
-ggplot(boot_stats, aes(x = condition, y = m, fill = condition, alpha = referent)) +
-  geom_bar(stat= 'identity', position = 'dodge') +
-  scale_alpha_manual(values=c(.25, .75)) +
-  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) + 
-  theme_few() +
-  scale_fill_colorblind() +
-  guides(fill = 'none') +
-  ylab('utterance informativity ratings') +
-  theme(aspect.ratio = 1)
-
-ggsave('../writing/journal_manuscript/figures/informativity.pdf')
-```
-
-### How well does informativity correlate with error?
-
-now bootstrap again but within items
-
-```{r}
-boot_objectsets = replicate(1000, {
-  raters = sample(1:16,16,replace = TRUE)
-  resampled_ratings = sampleRatings(inf_diffs, raters)
-
-  resampled_pairs = utt_data %>%
-    group_by(condition) %>%
-    do(samplePairs(.)) %>%
-    rename(label = text)
-
-  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
-    gather(referent, value, target,distractor,diff) %>%
-    group_by(condition, objectSet, referent) %>%
-    summarize(m = mean(value))
-
-  return((combined %>% filter(referent == 'diff'))$m)
-})
-```
-
-Kind of messy way to aggregate these, but so it goes... (replicate doesn't return a dataframe)
-
-(note that the 'outlier' here is the scripted 'candle' trial: raters thought it was *much* more likely to refer )
-
-```{r}
-boot_objectsets_stats = data.frame(
-  condition = c(rep('scripted', 8), rep('unscripted', 8)),
-  m_rating = rep(0,16),
-  ci_upper_rating = rep(0,16),
-  ci_lower_rating = rep(0,16),
-  objectSet = rep(1:8,2)
-)
-
-for (i in 1:16) {
-  boot_row = boot_objectsets[i,]
-  boot_objectsets_stats[i,]$m_rating = mean(boot_row)
-  boot_objectsets_stats[i,]$ci_upper_rating = sort(boot_row)[length(boot_row) * 0.025]
-  boot_objectsets_stats[i,]$ci_lower_rating = sort(boot_row)[length(boot_row) * 0.975]
-}
-
-left_join(boot_objectsets_stats, error_diffs, by = c('condition', 'objectSet')) %>%
-  ggplot(aes(x = m_rating, y = mean, color = condition, group = 1)) +
-    geom_point(size = 2, shape = 16) +
-    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), alpha = .3) +
-    geom_errorbarh(aes(xmax = ci_upper_rating, xmin = ci_lower_rating), alpha =.3) +
-    geom_smooth(linetype = 'dotted', method = 'lm', colour="black", se = F) +
-    theme_few() +
-    theme(aspect.ratio = 1, legend.position = 'top') +
-    scale_color_colorblind() +
-    xlab('informativity diff (target - distractor)') +
-    ylab('% critical error')
-
-ggsave('../writing/journal_manuscript/figures/informativity_predicts_errors.pdf',
-       height = 3, width = 3, useDingbats=FALSE)
-```
-
-```{r}
-boot_correlation = replicate(1000, {
-  raters = sample(1:16,16,replace = TRUE)
-  resampled_ratings = sampleRatings(inf_diffs, raters)
-
-  resampled_pairs = utt_data %>%
-    group_by(condition) %>%
-    do(samplePairs(.)) %>%
-    rename(label = text)
-
-  combined <- left_join(resampled_pairs, resampled_ratings, by = c('label', 'objectSet')) %>%
-    select(-target,-distractor,-label) %>%
-    inner_join(critTrials, by = c('gameid', 'condition', 'objectSet')) %>%
-    group_by(objectSet, condition) %>%
-    summarize(inf = mean(diff), error = mean(error))
-
-  return(cor(combined$inf, combined$error))
-})
-
-cat('r = ', mean(boot_correlation), 
-    ', 95% CI = [', sort(boot_correlation)[length(boot_correlation) * 0.025],
-    ', ', sort(boot_correlation)[length(boot_correlation) * 0.975],']')
-```
-
 ## TO-DO
@@
 or getting less informative because tired of task?)
-
-# Experiment 2
-
-## Import turk survey data
-
-```{r}
-exp2_subjInfo <- read_csv('../data/final/experiment2/planned_sample/turk/subjInfo.csv')
-```
-
-## Import message data
-
-```{r}
-d_annotated <- read_csv('../data/final/experiment2/planned_sample/chatMessage/messages_with_annotations.csv')
-
-# nonNativeSpeakerIDs <- unique((tangramSubjInfo %>% filter(nativeEnglish != "yes"))$gameid)
-incompleteIDs <- unique((d_annotated %>% group_by(gameid) %>% 
-                           filter(length(unique(trialNum)) != 24))$gameid)
-confused <- unique((exp2_subjInfo %>% filter(understandsInstructions != 'yes'))$gameid)
-nonNative <- unique((exp2_subjInfo %>% filter(nativeEnglish != 'yes'))$gameid)
-
-# Some speakers relied on location in grid (TODO: WHAT ABOUT 'BELOW' PEOPLE?)
-location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
-                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
-                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
-                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
-# Some speakers played this weird taboo game where they gave riddles 
-tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
-              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
-              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
-badGames <- c(incompleteIDs, nonNative, location_abusers, confused, tabooers)
-
-d <- d_annotated %>%
-  mutate(numFeatures = colorMention + shapeMention + textureMention) %>%
-  filter(!(gameid %in% badGames)) %>%
-  mutate(occlusions = as.factor(occlusions),
-        context = as.factor(context))
-```
-
-## Descriptive stats
-
-How many recruited, how many excluded, for what reasons, etc?
-
-```{r}
-paste0('recruited ', length(exp2_subjInfo$gameid))
-paste0('total games ', length(unique(d_annotated$gameid)))
-paste0('incomplete ', length(incompleteIDs))
-paste0('confused ', length(setdiff(confused,incompleteIDs)))
-paste0('non-native english ', length(setdiff(nonNative, union(confused, incompleteIDs))))
-paste0('violated instructions ', length(setdiff(c(location_abusers, tabooers), union(nonNative, union(confused, incompleteIDs)))))
-paste0('final sample ', length(unique(d$gameid)))
-```
-
-## Quick listener analysis: how common are errors?
-
-Errors rare...
-
-```{r}
-exp2_clicks <- read_tsv('../data/final/experiment2/planned_sample/clickedObj/clickedObj.csv') %>%
-  mutate(correct = ifelse(correct == 'true', 1, 0)) %>%
-  filter(!(gameid %in% badGames))
-
-## Errors by condition
-tmp <- exp2_clicks %>%
-  group_by(context, occlusions) %>%
-  summarize(m = sum(correct), total = length(correct) - m)
-
-chisq.test(t(tmp[,3:4]))
-
-## Error distribution by pair
-exp2_clicks %>%
-  group_by(gameid) %>%
-  summarize(numErrors = 24- sum(correct)) %>%
-  group_by(numErrors) %>%
-  tally()
-```
-
-## Visualize basic (no-hidden) pragmatic effect
-
-```{r}
-basic.d <- d %>%
-  filter(hidden == 'no') %>%
-  group_by(gameid, context) %>%
-  summarize(m = mean(numRawWords)) %>%
-  spread(context, m) %>%
-  mutate(diff = close- far,
-         ratio = log(close / far))
-
-summary(lmer(numRawWords ~ context + (1 + context | gameid), d %>% filter(hidden == 'no')))
-
-ggplot(basic.d, aes(x = far, y = close)) +
-    geom_point() +
-    theme_bw() +
-    geom_abline(intercept = 0, slope = 1) +
-    ylim(1,12) +
-    xlim(1,12) +
-    theme(aspect.ratio = 1, text = element_text(size=20))
-```
-
-Aggregated to difference scores:
-
-```{r}
-tidyboot_mean(basic.d %>%ungroup(), column = diff) %>%
-  ggplot(aes(x = 'close - far', y = empirical_stat)) +
-      geom_bar(stat ='identity', width = 0.25) +
-      geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
-      theme_few()+
-      geom_hline(yintercept = 0) +
-      #geom_vline(xintercept = 0) +
-      ylim(-1,1) +
-      # xlim(-4,4) +
-      ylab('# words used') +
-    theme(aspect.ratio = 1, text = element_text(size=20))
-
-t.test(basic.d$diff)
-```
-
-## Visualize message length collapsing across all subjects
-
-```{r}
-dodge <- position_dodge(width=0.9)
-cbPalette = c(rgb(0.8, .8, .8), rgb(.8, .3, .3),  rgb(0, .6, .4), rgb(.4, .4, .4))
-
-d %>%
-  group_by(context, hidden) %>%
-  tidyboot_mean(column = numRawWords) %>%
-  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
-    geom_bar(stat = 'identity', position = dodge) +
-    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
-    theme_few() +
-    ylab('mean # words') +
-    theme(aspect.ratio = 1, text = element_text(size=20)) +
-    scale_fill_manual(values=cbPalette) +
-    theme(aspect.ratio = 1)
-
-ggsave('../writing/journal_manuscript/figures/num_words.pdf', width = 4, height = 3)
-```
-
-## Visualize individual differences
-
-Some people used exactly the same number of words on every trial...
-
-```{r}
-hidden.d <- d %>%
-  group_by(gameid, context, hidden) %>%
-  summarize(m = mean(numRawWords)) %>%
-  spread(hidden, m) %>%
-  mutate(diff = yes- no,
-         ratio = log(yes/no)) 
-```
-
-```{r}
-ggplot(hidden.d, aes(x = no, y = yes)) +
-    geom_point() +
-    facet_wrap(~ context) +
-    theme_bw() +
-    geom_abline(intercept = 0, slope = 1) +
-    ylim(1,15) +
-    ylab('# words (with occlusions)') +
-    xlab('# words (without occlusions)') +
-    #ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
-    xlim(1,15) +
-    theme(aspect.ratio = 1, text = element_text(size=20))
-
-ggsave('../writing/journal_manuscript/figures//scatter.pdf')
-```
-
-Histogram
-
-```{r}
-hidden.d %>% 
-  #filter(!(gameid %in% all_the_things_ers)) %>% 
-  group_by(context) %>% 
-  mutate(m = mean(diff)) %>%
-  ggplot(aes(x = diff)) +
-      geom_histogram(binwidth = 0.33) +
-      theme_bw() +
-      geom_vline(xintercept = 0, linetype = 2) +
-      geom_vline(aes(xintercept = m), color = 'red') +
-      facet_wrap(~ context) +
-      # ylim(1,12) +
-      xlim(-6,6) +
-      theme(aspect.ratio = 1)
-
-ggsave("~/Downloads/histograms.pdf")
-```
-
-Look for population who mentioned all the things all the time...
-
-```{r}
-hidden.d %>% 
-  filter(context == 'far') %>% 
-  mutate(group = case_when(diff < -0 ~ 'anti',
-                           diff > 0 ~ 'pro',
-                           TRUE ~ 'all_the_things')) %>%
-  group_by(group) %>% tally()
-
-all_the_things_ers <- (hidden.d %>% filter(context =='far') %>% filter(diff == 0))$gameid
-```
-
-```{r}
-hidden.d %>%
-  group_by(context) %>%
-  summarize(m = mean(diff), se = sd(diff)/sqrt(length(diff))) %>%
-  ggplot(aes(x = context, y = m)) +
-    geom_bar(stat = 'identity', width = 0.25) +
-    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0) +
-    theme_few() +
-    geom_hline(yintercept = 0) +
-    ylim(-3, 3) +
-    theme(aspect.ratio = 1, text = element_text(size=20))  +
-    ylab('# words (occlusions - no occlusions)')
-
-```
-
-# Stats
-
-Simple effect of hidden-ness in 'far' contexts
-
-```{r}
-# summary(lmer(numRawWords ~ hidden + (1 + hidden | gameid), 
-#              data = d %>% filter(context == 'far') 
-))
-#t.test(numRawWords ~ hidden, data = d %>% filter(context == 'far') , var.equal = FALSE)
-```
-
-interaction-y way
-
-```{r}
-contrasts(d$context) <- contr.treatment(2, base = 2)
-contrasts(d$occlusions) <- contr.treatment(2, base = 2)
-
-summary(lmer(numRawWords ~ context * occlusions + (1 + context * occlusions | gameid), data = d ))
-```
-
-## Use feature data instead
-
-```{r}
-d %>% 
-  group_by(context, occlusions) %>% 
-  summarize(color = mean(colorMention), shape = mean(shapeMention), texture = mean(textureMention))
-```
-
-Histograms of number of features mentioned in each condition
-
-```{r}
-ggplot(d, aes(x = numFeatures)) +
-  geom_histogram(binwidth = 1) +
-  facet_grid(context ~ occlusions) +
-  theme_few()
-```
-
-```{r}
-d %>% group_by(context, occlusions) %>% 
-  mutate(numTrials = length(text)) %>% 
-  group_by(context, occlusions, colorMention, textureMention) %>% 
-  summarize(n = length(text)/mean(numTrials)) %>%
-  ggplot(aes(x = colorMention, y = textureMention, fill = n)) +
-    geom_bin2d(stat = 'identity') +
-    stat_bin2d(geom = "text", aes(label = round(n, 2))) +
-    scale_fill_gradient(low = "white", high = "red") +
-    facet_grid(occlusions ~ context) +
-    theme_few()
-```
-
-Note: Random interaction doesn't converge (less variance in these numbers...)
-
-```{r}
-```
-
-```{r}
-contrasts(d$context) <- contr.treatment(2, base = 2)
-contrasts(d$occlusions) <- contr.treatment(2, base = 2)
-summary(glmer(colorMention ~ occlusions * context + (1 + occlusions | gameid), 
-              family = 'binomial', data = d))
-```
-
-```{r}
-summary(glmer(textureMention ~ occlusions * context + (1 + occlusions | gameid), 
-              family = 'binomial', data = d))
-summary(tmp2)
-fixef(tmp2)
-baseprob = fixef(tmp2)[1] 
-cat('no-occlusion:', exp(baseprob) / (1 + exp(baseprob)))
-occludedprob = fixef(tmp2)[1] + fixef(tmp2)[2]
-cat('with occlusion:', exp(occludedprob) / (1 + exp(occludedprob)))
-```
-
-```{r}
-d %>% 
-  group_by(context, occlusions) %>% 
-  summarize(numFeatures = mean(numFeatures))
-```
-
-```{r}
-d.booted <- d %>% 
-  gather(feature, mentioned, shapeMention, colorMention, textureMention) %>%
-  group_by(context, hidden, feature) %>% 
-  tidyboot_mean(column = mentioned)
-
-ggplot(d.booted, aes(x = context, y = mean, fill = hidden)) + 
-  geom_bar(stat = 'identity', position = dodge) +
-  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
-  facet_wrap(~ feature) + 
-  theme_few() +
-  theme(aspect.ratio = 1, text = element_text(size=20))  +
-  ylab('% utterances with feature') +
-  scale_fill_manual(values=cbPalette)
-
-ggsave('../writing/journal_manuscript/figures/by-feature-analysis.pdf', 
-       height = 3, width = 9)
-```
+4. Order effects (i.e. are people more likely to use a modifier when they had to on the previous round?)
+5. Are people more informative when there are more objects?

### 2018-07-24 analysis/final_analysis_bootstrap_helpers.R
move bootstrap helpers to script

@@ -1,0 +1,28 @@
+## In the inner loop, we want to sample a 'mean' rating for each utterance
+## using same sample of *raters* for all utterances... 
+sampleRatings = function(df, indices) {
+  grouped_df <- df %>%
+    group_by(label, objectSet) %>%
+    do(data.frame(diff = mean(.$diff[indices]),
+                  target = mean(.$target[indices]),
+                  distractor = mean(.$distractor[indices])))
+  return(grouped_df)
+}
+
+## In outer loop, we want to sample 'pairs' as a unit (with their 8 utterances)
+samplePairs = function(df) {
+  uniq_ids <- unique(df$gameid)
+  new_ids <- sample(uniq_ids, length(uniq_ids), replace = TRUE)
+  resampled_ids = data.frame(gameid = new_ids,
+                             sampleid = 1:length(new_ids), 
+                             stringsAsFactors = FALSE)
+  return(left_join(resampled_ids, df, by = c('gameid')))
+}
+
+## Finally, want to sample object sets per pair
+sampleObjects = function(df, indices) {
+  new_objects <- data.frame(objectSet = indices)
+  return(left_join(new_objects, df, by = c('objectSet')))
+}
+# sampleObjects(data.frame(objectSet = c(1,2,3,4,5,6,7,8), utt = c('a', 'b', 'c', 'a', 'b', 'c', 'd', 'e')),
+#               obj_sets)

### 2018-10-10 modeling/refModule/json/makeNominalMeanings.js
remove old stuff

@@ -1,67 +1,0 @@
-var _ = require('lodash');
-var papaparse = require('papaparse');
-var fs = require('fs');
-
-var labelToObjName = function(label) {
-  var lowerCased = label.toLowerCase();
-  var noWhiteSpace = lowerCased.replace(/[^A-Z0-9]+/ig, "");
-  return noWhiteSpace == "mms" ? "mnms" : noWhiteSpace;
-};
-
-var typicality = function () {
-  var filename = ("../../data/modelTypicalities.csv");
-  var parseResult = papaparse.parse(fs.readFileSync(filename, 'utf8'),
-				    {header: true, skipEmptyLines : true});
-  this.data = parseResult.data;
-  //console.log("this.data:", this.data);
-};
-
-// typicality.prototype.clean = function() {
-//   this.data = _.map(this.data, function(row) {
-//     var newRow = _.clone(row);
-//     var wordsInLabel = row.label.split(' ');
-//     newRow.label = (_.includes(['a', 'an'], wordsInLabel[0]) ?
-// 		    wordsInLabel.slice(1).join(' ') :
-// 		    wordsInLabel.join(' '));
-//     return newRow;
-//   });
-// };
-
-typicality.prototype.getLabels = function() {
-  return _.uniq(_.map(this.data, 'label'));
-};
-
-typicality.prototype.getTypicality = function(label, obj) {
-  console.log(label)
-  console.log(obj)
-  var subset = _.filter(this.data, function(row) {
-    return row.object === obj & row.label === label;
-  });
-  console.log(subset)
-  return Number(subset[0]["m"]);
-};
-
-typicality.prototype.getPossibleReferents = function(label) {
-  return _.map(_.filter(this.data, function(row) {
-    return row.label === label;
-  }), 'object');
-};
-
-typicality.prototype.makeTree = function() {
-  this.tree = {};
-  var that = this;
-  _.each(this.labels, function(label) {
-    var cleanedLabel = labelToObjName(label);
-    that.tree[cleanedLabel] = {};
-    _.each(that.getPossibleReferents(label), function(object) {
-      that.tree[cleanedLabel][object] = that.getTypicality(label, object);
-    });
-  });
-};
-
-var t = new typicality();
-//t.clean();
-t.labels = t.getLabels();
-t.makeTree();
-
-fs.writeFileSync('keysar-meanings.json', JSON.stringify(t.tree, null, 2));

### 2018-10-10 modeling/refModule/ref.js

@@ -1,190 +1,0 @@
-var _ = require('lodash');
-var fs = require('fs');
-var babyparse = require('babyparse');
-
-function _logsumexp(a) {
-  var m = Math.max.apply(null, a);
-  var sum = 0;
-  for (var i = 0; i < a.length; ++i) {
-    sum += (a[i] === -Infinity ? 0 : Math.exp(a[i] - m));
-  }
-  return m + Math.log(sum);
-}
-
-var getL0score = function(targetObj, utt, context, params) {
-  var similarities = params.lexicon[utt];
-  var a = [];
-  for(var i=0; i<context.length; i++){
-    a.append(params.simScale * similarities[context[i]]);
-  }
-  return (params.simScale * similarities[targetObj]
-	  - _logsumexp(a));
-};
-
-var powerset = function (set) {
-  if (set.length == 0)
-    return [[]];
-  else {
-    var rest = powerset(set.slice(1));
-    return rest.map(
-      function(element) {
-        return [set[0]].concat(element);
-      }).concat(rest);
-  }
-};
-
-// get all words that can be used to describe target
-var possibleUtts = function(target, lexicon) {
-  return _.keys(_.pickBy(lexicon, function(value, key) {
-    return _.has(value, target);
-  }));
-};
-
-// Note: assumes the only object considered by the listener is the true distractor...
-var possibleObjects = function(target) {
-  var objectSet = target.split('_')[1];
-  return [target, ['distractor', objectSet].join('_')];
-};
-
-var makeArr = function(n, v) {
-  return _.repeat(n, v);
-};
-
-var makeColorSizeLists = function(wordsOrObjects) {
-  var colorList = wordsOrObjects === 'words' ? colors.concat('') : colors;
-  var sizeList = wordsOrObjects === 'words' ? sizes.concat('') : sizes;
-  var typeList = wordsOrObjects === 'words' ? types.concat('thing') : types;
-
-  return _.flattenDepth(_.map(sizeList, function(size) {
-    return _.map(colorList, function(color) {
-      return _.map(typeList, function(type) {
-        return [size, color, type]
-      });
-    });
-  }), 2);
-};
-
-var colorSizeWordMeanings = function(params) {
-  return _.extend(
-    _.zipObject(colors, _.times(colors.length, _.constant(params.colorTyp))),
[…]

### 2018-10-10 analysis/final_analysis.Rmd
make scatter

@@
 Look at pattern of predictives for each model, compared to empirical number of features mentioned
-
-```{r}
-rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
-      read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'occlusion-blind model')) %>%
+```{r}
+dodge <- position_dodge(width=0.9)
+
+rbind(
+    read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
+    read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'occlusion-blind model')
+    ) %>%
   mutate(i = row_number()) %>%
-  gather(condition, value, close_hidden:far_visible) %>%
-  group_by(condition, source) %>%
+  gather(condition, value, -prob, -MCMCprob, -source, -i) %>%
+  separate(condition, sep = '_', into = c('context', 'numDistractors', 'distractorType', 'hidden')) %>%
+  group_by(context, hidden, source) %>%
   summarize(mean = estimate_mode(value),
@@
             ci_upper = round(HPDhi(value), 3)) %>%
-  separate(condition, into = c('context', 'hidden')) %>%
+  #separate(condition, into = c('context', 'hidden')) %>%
+  ungroup() %>%
   mutate(hidden = ifelse(hidden == 'hidden', 'yes', 'no')) %>%
@@
 ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/predictives.pdf',
+       height = 5, width = 9)
+```
+
+```{r}
+to.plot <- rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
+     read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'occlusion-blind model')
+     ) %>%
+  mutate(i = row_number()) %>%
+  gather(condition, value, -prob, -MCMCprob, -source, -i) %>%
+  group_by(condition, source) %>%
+  summarize(mean = estimate_mode(value),
+            ci_lower = round(HPDlo(value), 3),
+            ci_upper = round(HPDhi(value), 3)) %>%
+  separate(condition, sep = '_', into = c('context', 'numDistractors', 'distractorType', 'hidden')) %>%
+  mutate(hidden = ifelse(hidden == 'hidden', 'yes', 'no')) %>%
+  rbind(read_csv('../modeling/bdaInput/bdaInput.csv') %>% 
+          mutate(numFeatures = color + shape + texture) %>%
+          group_by(context, distractorType, numDistractors, hidden) %>% 
+          tidyboot_mean(column = numFeatures) %>%
+          ungroup() %>%
+          mutate(numDistractors = numDistractors - 2) %>%
+          mutate(context = as.character(context)) %>%
+          mutate(distractorType = gsub("_", "", distractorType)) %>%
+          mutate(source = as.character('empirical data')) %>% 
+          ungroup() %>%
+          select(-empirical_stat, -n)) %>%
+  mutate(source = factor(source, levels = c('empirical data', 'occlusion-sensitive model', 
+                                            'occlusion-blind model'))) %>%
+  unite(col = asdf, mean, ci_lower, ci_upper) %>%
+  spread(source, asdf) %>%
+  separate(`empirical data`, sep = '_', into = c('emp_mean', 'emp_lower', 'emp_upper'))%>%
+  separate(`occlusion-sensitive model`, sep = '_', into = c('occ_mean', 'occ_lower', 'occ_upper')) %>%
+  separate(`occlusion-blind model`, sep = '_', into = c('ego_mean', 'ego_lower', 'ego_upper')) %>%
+  mutate(emp_mean = as.numeric(emp_mean), emp_lower = as.numeric(emp_lower), emp_upper = as.numeric(emp_upper),
+         occ_mean = as.numeric(occ_mean), occ_lower = as.numeric(occ_lower), occ_upper = as.numeric(occ_upper),
+         ego_mean = as.numeric(ego_mean), ego_lower = as.numeric(ego_lower), ego_upper = as.numeric(ego_upper))
+
+ggplot(to.plot, aes(x = occ_mean, y = emp_mean)) +
+  geom_point() +#(stat = 'identity', position = dodge) +
+  geom_errorbar(aes(ymin = emp_lower, ymax = emp_upper), width = 0) +
+  #geom_errorbarh(aes(xmin = occ_lower, xmax = occ_upper), width = 0) +
+  geom_smooth(method = 'lm') +
+  theme_few() +
+  theme(aspect.ratio = 1, text = element_text(size=20)) +
+  xlab('predicted # features')+
+  ylab('empirical # features')
+
+ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/predictiveScatter.pdf',
+       height = 5, width = 9)
+```
+
+```{r}
+ggplot(to.plot, aes(x = ego_mean, y = emp_mean)) +
+  geom_point() +#(stat = 'identity', position = dodge) +
+  geom_errorbar(aes(ymin = emp_lower, ymax = emp_upper), width = 0) +
+  #geom_errorbarh(aes(xmin = ego_lower, xmax = ego_upper), width = 0) +
+  geom_smooth(method = 'lm') +
+  theme_few() +
+  theme(aspect.ratio = 1, text = element_text(size=20)) +
+  xlab('predicted # features')+
+  ylab('empirical # features') 
+ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/egocentricScatter.pdf',
        height = 5, width = 9)

### 2018-10-10 modeling/refModule/ref.js
Check in node modules

@@ -1,0 +1,198 @@
+var _ = require('lodash');
+var fs = require('fs');
+var babyparse = require('babyparse');
+
+function _logsumexp(a) {
+  var m = Math.max.apply(null, a);
+  var sum = 0;
+  for (var i = 0; i < a.length; ++i) {
+    sum += (a[i] === -Infinity ? 0 : Math.exp(a[i] - m));
+  }
+  return m + Math.log(sum);
+}
+
+var getL0score = function(targetObj, utt, context, params) {
+  var similarities = params.lexicon[utt];
+  var a = [];
+  for(var i=0; i<context.length; i++){
+    a.append(params.simScale * similarities[context[i]]);
+  }
+  return (params.simScale * similarities[targetObj]
+	  - _logsumexp(a));
+};
+
+var powerset = function (set) {
+  if (set.length == 0)
+    return [[]];
+  else {
+    var rest = powerset(set.slice(1));
+    return rest.map(
+      function(element) {
+        return [set[0]].concat(element);
+      }).concat(rest);
+  }
+};
+
+// get all words that can be used to describe target
+var possibleUtts = function(target, lexicon) {
+  return _.keys(_.pickBy(lexicon, function(value, key) {
+    return _.has(value, target);
+  }));
+};
+
+// Note: assumes the only object considered by the listener is the true distractor...
+var possibleObjects = function(target) {
+  var objectSet = target.split('_')[1];
+  return [target, ['distractor', objectSet].join('_')];
+};
+
+var makeArr = function(n, v) {
+  return _.repeat(n, v);
+};
+
+var makeColorSizeLists = function(wordsOrObjects) {
+  var colorList = wordsOrObjects === 'words' ? colors.concat('') : colors;
+  var sizeList = wordsOrObjects === 'words' ? sizes.concat('') : sizes;
+  var typeList = wordsOrObjects === 'words' ? types.concat('thing') : types;
+
+  return _.flattenDepth(_.map(sizeList, function(size) {
+    return _.map(colorList, function(color) {
+      return _.map(typeList, function(type) {
+        return [size, color, type]
+      });
+    });
+  }), 2);
+};
+
+var colorSizeWordMeanings = function(params) {
+  return _.extend(
+    _.zipObject(colors, _.times(colors.length, _.constant(params.colorTyp))),
[…]

### 2018-10-10 analysis/final_analysis.Rmd
analysis update

@@
 
-ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/predictiveScatter.pdf',
-       height = 5, width = 9)
+ggsave('../../pragmatics_of_perspective_taking_tex/cognition_format/figures/predictiveScatter.pdf',
+       useDingbats=FALSE )
 ```
@@
   ylab('empirical # features') 
-ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/egocentricScatter.pdf',
-       height = 5, width = 9)
+ggsave('../../pragmatics_of_perspective_taking_tex/cognition_format/figures/egocentricScatter.pdf',
+       useDingbats=FALSE)
+       #height = 5, width = 9)
 ```

### 2018-10-10 analysis/final_analysis.Rmd
Use mean instead of mode

--- 2018-07-24 analysis/final_analysis.Rmd
@@
 Look at pattern of predictives for each model, compared to empirical number of features mentioned
-
-```{r}
-rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
-      read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'occlusion-blind model')) %>%
+```{r}
+dodge <- position_dodge(width=0.9)
+
+rbind(
+    read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
+    read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'occlusion-blind model')
+    ) %>%
   mutate(i = row_number()) %>%
-  gather(condition, value, close_hidden:far_visible) %>%
-  group_by(condition, source) %>%
-  summarize(mean = estimate_mode(value),
+  gather(condition, value, -prob, -MCMCprob, -source, -i) %>%
+  separate(condition, sep = '_', into = c('context', 'numDistractors', 'distractorType', 'hidden')) %>%
+  #filter(context == 'far' & source == 'occlusion-blind model' & hidden == 'visible') %>%
+  group_by(context, hidden, source) %>%
+  summarize(mean = mean(value),
             ci_lower = round(HPDlo(value), 3),
             ci_upper = round(HPDhi(value), 3)) %>%
-  separate(condition, into = c('context', 'hidden')) %>%
+  #separate(condition, into = c('context', 'hidden')) %>%
+  ungroup() %>%
   mutate(hidden = ifelse(hidden == 'hidden', 'yes', 'no')) %>%
@@
     theme_few() +
+    theme(aspect.ratio = .57, text = element_text(size=20)) +
+    ylab('mean # features produced')
+
+ggsave('../../pragmatics_and_perspective_taking_tex/cognition_format/figures/predictives.pdf',
+       useDingbats=FALSE )
+```
+
+```{r}
+to.plot <- rbind(read.csv('../modeling/bdaOutput/occlusionSensitivityPredictives.csv') %>% mutate(source = 'occlusion-sensitive model'),
+     read.csv('../modeling/bdaOutput/egocentricPredictives.csv') %>% mutate(source = 'occlusion-blind model')
+     ) %>%
+  mutate(i = row_number()) %>%
+  gather(condition, value, -prob, -MCMCprob, -source, -i) %>%
+  group_by(condition, source) %>%
+  summarize(mean = mean(value),
+            ci_lower = round(HPDlo(value), 3),
+            ci_upper = round(HPDhi(value), 3)) %>%
+  separate(condition, sep = '_', into = c('context', 'numDistractors', 'distractorType', 'hidden')) %>%
+  mutate(hidden = ifelse(hidden == 'hidden', 'yes', 'no')) %>%
+  rbind(read_csv('../modeling/bdaInput/bdaInput.csv') %>% 
+          mutate(numFeatures = color + shape + texture) %>%
+          group_by(context, distractorType, numDistractors, hidden) %>% 
+          tidyboot_mean(column = numFeatures) %>%
+          ungroup() %>%
+          mutate(numDistractors = numDistractors - 2) %>%
+          mutate(context = as.character(context)) %>%
+          mutate(distractorType = gsub("_", "", distractorType)) %>%
+          mutate(source = as.character('empirical data')) %>% 
+          ungroup() %>%
+          select(-empirical_stat, -n)) %>%
+  mutate(source = factor(source, levels = c('empirical data', 'occlusion-sensitive model', 
+                                            'occlusion-blind model'))) %>%
+  unite(col = asdf, mean, ci_lower, ci_upper) %>%
+  spread(source, asdf) %>%
+  separate(`empirical data`, sep = '_', into = c('emp_mean', 'emp_lower', 'emp_upper'))%>%
+  separate(`occlusion-sensitive model`, sep = '_', into = c('occ_mean', 'occ_lower', 'occ_upper')) %>%
+  separate(`occlusion-blind model`, sep = '_', into = c('ego_mean', 'ego_lower', 'ego_upper')) %>%
+  mutate(emp_mean = as.numeric(emp_mean), emp_lower = as.numeric(emp_lower), emp_upper = as.numeric(emp_upper),
+         occ_mean = as.numeric(occ_mean), occ_lower = as.numeric(occ_lower), occ_upper = as.numeric(occ_upper),
+         ego_mean = as.numeric(ego_mean), ego_lower = as.numeric(ego_lower), ego_upper = as.numeric(ego_upper))
+
+ggplot(to.plot, aes(x = occ_mean, y = emp_mean)) +
+  geom_point() +#(stat = 'identity', position = dodge) +
+  geom_errorbar(aes(ymin = emp_lower, ymax = emp_upper), width = 0) +
+  #geom_errorbarh(aes(xmin = occ_lower, xmax = occ_upper), width = 0) +
+  geom_smooth(method = 'lm') +
+  theme_few() +
   theme(aspect.ratio = 1, text = element_text(size=20)) +
-    ylab('mean # features produced')
-
-ggsave('../../pragmatics_of_perspective_taking_tex/pnas_format/figures/predictives.pdf',
-       height = 5, width = 9)
+  xlab('predicted # features')+
+  ylab('empirical # features')
+
+ggsave('../../pragmatics_of_perspective_taking_tex/cognition_format/figures/predictiveScatter.pdf',
+       useDingbats=FALSE )
+```
+
+```{r}
+ggplot(to.plot, aes(x = ego_mean, y = emp_mean)) +
+  geom_point() +#(stat = 'identity', position = dodge) +
+  geom_errorbar(aes(ymin = emp_lower, ymax = emp_upper), width = 0) +
+  #geom_errorbarh(aes(xmin = ego_lower, xmax = ego_upper), width = 0) +
+  geom_smooth(method = 'lm') +
+  theme_few() +
+  theme(aspect.ratio = 1, text = element_text(size=20)) +
+  xlab('predicted # features')+
+  ylab('empirical # features') 
+ggsave('../../pragmatics_of_perspective_taking_tex/cognition_format/figures/egocentricScatter.pdf',
+       useDingbats=FALSE)
+       #height = 5, width = 9)
 ```

### 2018-10-16 analysis/final_analysis.Rmd
Update ylim

@@
     theme_few() +
+    ylim(0,3) +
     theme(aspect.ratio = .57, text = element_text(size=20)) +

### 2019-08-12 analysis/final_analysis.Rmd
fix error created by new read_csv behavior

@@
 library(coda)
-library(irr)
+#library(irr)
 
@@
 exp2_drops <- read_csv('../data/final/experiment2/drop/dropsFromMongo.csv') %>%
-  mutate(critical = ifelse(critical == 'True', 1, 0),
-         correct = ifelse(correct == 'correct', 1, 0),
-         instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
+  mutate(correct = correct == 'correct',
+         instructionNum = ifelse(correct == FALSE, instructionNum + 1, instructionNum)) %>%
   group_by(gameid, targetObject) %>%
-  filter(length(unique(instructionNum)) < 2) %>%
+  filter(length(unique(instructionNum)) == 1) %>% 
   group_by(gameid, instructionNum, trialNum) %>%
@@
 fillerErrors <- exp2_drops %>%
-  #filter(critical != 1) %>%   # only look at mistakes on noncritical (filler) items
+  filter(critical == FALSE) %>%   # only look at mistakes on noncritical (filler) items
   filter(attemptNum == 0) %>%   # don't want to double-count people for messing the same thing up multiple times
@@
   ungroup() %>%
-  mutate(correct = ifelse(correct == 1, 'no_mistake', 'mistake')) %>%
+  mutate(correct = ifelse(correct, 'no_mistake', 'mistake')) %>%
   spread(correct, n, fill = 0) %>%
@@
 nonCriticalMistakes = (fillerErrors %>%
-  filter(critical == 0) %>%
   filter(numErrors >= 2))$gameid
@@
                            group_by(gameid, trialNum, instructionNum) %>% 
-                           summarize(id = row_number()) %>% 
+                           summarize(nn = n()) %>% 
                            group_by(gameid) %>% 
@@
 buggyTrials <- read_csv('../data/final/experiment2/drop/dropsFromMongo.csv') %>%
-  mutate(critical = ifelse(critical == 'True', 1, 0),
-       correct = ifelse(correct == 'correct', 1, 0),
-       instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
+  mutate(correct = correct == 'correct',
+       instructionNum = ifelse(correct == FALSE, instructionNum + 1, instructionNum)) %>%
   group_by(gameid, targetObject) %>%

### 2019-08-14 analysis/final_analysis.Rmd
add analyses of chagne over ti e

@@
 
+Examine how listeners might be adapting
+
+```{r}
+critTrials %>% 
+  #filter(condition == 'scripted') %>%
+  group_by(gameid) %>% 
+  mutate(trialNum = row_number(),
+         lastRoundIncorrect = 1-lag(correct)) %>% 
+  group_by(trialNum) %>% 
+  summarize(prop = sum(correct)/length(correct)) %>%
+  mutate(error = ifelse(correct == 0, 'error', 'no error')) %>%
+  ggplot(aes(x= trialNum, y = n, color = lastRoundIncorrect)) +
+    geom_line()
+```
+
+```{r}
+library(brms)
+summary(brm(correct ~ trialNum + (1 + trialNum| gameid), family = 'bernoulli', 
+              data = critTrials %>%   
+                    group_by(gameid) %>%
+                    mutate(trialNum = row_number(),
+                           lastRoundIncorrect = 1-lag(correct)), control = list(adapt_delta = .95)))
+```
+
 ### Mouse-tracking

### 2020-02-17 modeling/run_RR_grid.sh
write out RR results

@@ -1,0 +1,3 @@
+#!/bin/bash#
+# parallel --bar --colsep ',' "sh ./run_grid_search.sh {1} {2}" :::: input/grid.csv
+webppl RR_simulations.wppl --require ./refModule/ --require webppl-csv -- --perspectiveCost $1 --chainNum $2

### 2020-02-17 modeling/run_dynamics_grid.sh

@@ -1,0 +1,3 @@
+#!/bin/bash#
+# parallel --bar --colsep ',' "sh ./run_dynamics_grid.sh {1} {2}" :::: input/dynamics_grid.csv
+webppl dynamics_simulations.wppl --require ./refModule/ --require webppl-csv -- --numDatapoints $1 --chainNum $2

### 2020-02-17 analysis/final_analysis.Rmd
make figs

--- 2019-08-12 analysis/final_analysis.Rmd
@@
 
-## Model-based analyses
+## Model-based analyses (Appendix)
+
+### Resource-rational simulations
+
+```{r}
+RR <- bind_rows(
+  read_csv('../modeling/output/RRlistenerOutput.csv') %>% mutate(src = 'listener'),
+  read_csv('../modeling/output/RRspeakerOutput.csv') %>% mutate(src ='speaker')
+)
+
+ggplot(RR, aes(x = weighting, y = prob, color = perspectiveCost, group = perspectiveCost)) +
+  geom_line() +
+  facet_wrap(~ src) +
+  theme_few() +
+  xlim(0,1) +
+  theme(aspect.ratio = 1)
+
+ggsave('../../pragmatics_and_perspective_taking_tex/cognitive_science_revision/figures/simulations.pdf')
+```
+
+### Dynamics simulations
+
+```{r}
+dynamics <- bind_rows(
+  read_csv('../modeling/output/dynamicsOutput.csv')
+)
+
+dynamics %>%
+  group_by(numDatapoints) %>%
+  filter(prob == max(prob)) %>%
+  ggplot(aes(x = numDatapoints, y = weighting)) +
+    geom_line() +
+    theme_few() +
+    xlim(0, 30) +
+    ylim(.5, 1) +
+    labs(x = 'length of interaction', y = 'optimal weighting') +
+    theme(aspect.ratio = 1)
+
+ggsave('../../pragmatics_and_perspective_taking_tex/cognitive_science_revision/figures/dynamics.pdf', width = 2, height = 2,
+       units = 'in')
+```
+
+### Model comparison
 
@@
   filter(sender == 'director') %>%
-  group_by(gameid, trialNum, instructionNum, attemptNum, critical, condition, objectSet, targetObject) %>%
+  group_by(gameid, trialNum, instructionNum, attemptNum, 
+           critical, condition, objectSet, targetObject) %>%
   summarize(text = paste0(contents, collapse = ' '),
@@
 ```{r}
-d.onDistractor <- exp2_d %>% 
-  left_join(read_csv('../data/final/experiment2/mouseOverDistractor/distractorMouseFromMongo.csv') %>% 
-              mutate(critical = ifelse(critical == 'True', 1, 0))
-            ) %>%
-  select(-serverTime, -X1) %>%
+d.onDistractor.correctOnly <- exp2_d %>% 
+  filter(correct) %>%
   filter(attemptNum == 0) %>%
-  filter(correct == 1) %>%
   filter(critical == 1) %>%
+  left_join(read_csv('../data/final/experiment2/mouseOverDistractor/distractorMouseFromMongo.csv'),
+            by = c('gameid', 'trialNum', 'objectSet', 'trialType', 'condition', 'attemptNum')) %>%
+  select(-serverTime, -X1, -ends_with('.y')) %>%
   group_by(gameid, trialNum) %>%
@@
   mutate(diff = ifelse(is.na(off), 0, off - on)) %>%
-  group_by(gameid,condition,objectSet, trialType) %>%
+  group_by(gameid,trialNum,condition,objectSet, trialType) %>%
   summarize(totalAmtTimeOnDistractor = sum(diff)) %>%
@@
 dodge <- position_dodge(width=0.9)
-d.onDistractor %>%
+d.onDistractor.correctOnly %>%
   group_by(condition, trialType) %>%
@@
 summary(lmer(totalAmtTimeOnDistractor ~ trialType * condition + (1 +trialType|gameid) + (1 +trialType|objectSet), 
-             data = d.onDistractor))
+             data = d.onDistractor.correctOnly))
 ```
@@
 ```{r}
-summary(glmer(hovered ~ trialType * condition + (1|gameid) + (1+trialType|objectSet), data = d.onDistractor, family = 'binomial'))
-```
+summary(glmer(hovered ~ trialType * condition + (1|gameid) + (1+trialType|objectSet), 
+              data = d.onDistractor.correctOnly, family = 'binomial'))
+```
+
+## Adaptation over time
+
+```{r}
+critTrials %>% 
+  group_by(gameid) %>% 
+  mutate(trialNum = row_number()) %>% 
+  group_by(trialNum) %>% 
+  summarize(prop = sum(correct)/length(correct)) %>%
+  #mutate(error = ifelse(correct == 0, 'error', 'no error')) %>%
+  ggplot(aes(x= trialNum, y = prop)) +
+    geom_point() +
+    geom_smooth(method = 'lm')
+```
+
+Mixed-effects model on raw error probability
+
+```{r}
+summary(glmer(correct ~ trialNum * condition + (1 | gameid), family = 'binomial',
+             data = critTrials %>%   
+                    group_by(gameid) %>%
+                    mutate(trialNum = row_number())))
+```
+
+Supplementary Bayesian version with more complete random effects
+
+```{r}
+library(brms)
+summary(brm(correct ~ trialNum * condition  + (1 + trialNum | gameid), family = 'bernoulli', 
+              data = critTrials %>%   
+                    group_by(gameid) %>%
+                    mutate(trialNum = row_number(),
+                           lastRoundIncorrect = 1-lag(correct)), control = list(adapt_delta = .95)))
+```
+
 ## Speaker analyses
@@
 5. Are people more informative when there are more objects?
+
+# Supplemental experiment
+
+```{r}
+library(tidyverse)
+d.supp.clicks <- read_tsv('../data/final/supplemental_exp/clickedObj/allClicks.csv')
+paste0(c('we have', length(unique(d.supp.clicks$gameid)), 'participants.'), collapse = ' ')
+
+# Following same criterion as above, we remove participants who make more than one
+# non-critical error (ignoring errors in the initial practice block)
+exclusions <- d.supp.clicks %>%
+  rowwise() %>%
+  filter(occlusions != 'critical'  || (occlusions == 'critical' && !clickedHiddenObj && !correct)) %>%
+  filter(trialNum > 4) %>%
+  group_by(gameid) %>%
+  summarize(nonCriticalErrors = sum(!correct)) %>%
+  filter(nonCriticalErrors > 1) %>%
+  pull(gameid)
+```
+
+```{r}
+d.supp.clicks %>%
+  filter(!(gameid %in% exclusions)) %>%
+  group_by(gameid) %>%
+  filter(occlusions == 'critical') %>%
+  mutate(instance = row_number()) %>%
+  filter(clickedHiddenObj) %>%
+  group_by(gameid) %>%
+  summarize(errorInstances = paste0(instance, collapse =  ',')) #%>%
+  # group_by(instance) %>%
+  # summarize(clickedHidden =sum(clickedHiddenObj))
+```
+
+```{r}
+critical_only <- read_tsv('../data/final/supplemental_exp/clickedObj/allClicks.csv') %>%
+  filter(occlusions =='critical') %>%
+  rowwise() %>%
+  mutate(shape_of_target = strsplit(intendedName, '_')[[1]][3]) %>%
+  rowwise() %>%
+  mutate(i = grep(shape_of_target, strsplit(names, ',')[[1]])[1]) %>%
+  mutate(criticalDistractorX = as.numeric(gsub('\\[', '', strsplit(gridXs, ',')[[1]][i])),
+         criticalDistractorY = as.numeric(gsub('\\[', '', strsplit(gridYs, ',')[[1]][i])),
+         criticalDistractorLeftBorder = 600 * (criticalDistractorX - 1),
+         criticalDistractorRightBorder = 600 * (criticalDistractorX),
+         criticalDistractorTopBorder = 600 * (criticalDistractorY-1),
+         criticalDistractorBottomBorder = 600 * (criticalDistractorY)) %>%
+  select(gameid, trialNum, criticalDistractorLeftBorder,criticalDistractorRightBorder,criticalDistractorTopBorder,criticalDistractorBottomBorder)
+```
+
+```{r}
+distractor <- left_join(
+  read_tsv('../data/final/supplemental_exp/updateMouse/mouse.csv'),
+   read_tsv('../data/final/supplemental_exp/clickedObj/allClicks.csv'),
+  by = c('trialNum', 'gameid')) %>%
+  filter(occlusions == 'critical') %>%
+  left_join(critical_only, by = c('gameid', 'trialNum')) %>%
+  rowwise() %>%
+  mutate(overDistractor = (
+    rawMouseX < criticalDistractorRightBorder && 
+    rawMouseX > criticalDistractorLeftBorder &&
+    rawMouseY < criticalDistractorBottomBorder && 
+    rawMouseY > criticalDistractorTopBorder
+    )
+  ) %>%
+  filter(!(gameid %in% exclusions)) %>%
+  group_by(gameid, trialNum) %>% 
+  summarize(hovered = sum(overDistractor) > 0) %>%
+  mutate(instance = row_number())
+
+distractor %>%
+  group_by(instance) %>%
+  tidyboot_mean(sum(hovered) / length(hovered)) %>%
+  ggplot(aes(x = instance, y = empirical_stat)) +
+    geom_point() +
+    #geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0) +
+    geom_smooth(method = 'lm', formula = y ~ poly(x, 2), se = F) +
+    theme_bw() +
+    ylab('% trials hovering over distractor') +
+    xlab('occurrence in trial sequence') +
+    theme(aspect.ratio = 1/2)
+
+ggsave('hoverReduction.pdf')
+library(lme4)
+library(tidyboot)
+library(lmerTest)
+summary(lmer(hovered ~ instance + (1 | gameid), data = distractor))
+```

### 2020-12-19 data/final/experiment1/planned_sample/chatMessage/preprocessMessages.R
only keep cogsci data

@@ -1,31 +1,0 @@
-library(tidyverse)
-library(stringi)
-library(quanteda)
-
-rawD <- read_tsv('./raw_message.csv')
-
-mentionedOutlineColor <- function(m) {
-  colors <- c("blue","green", "black", "red")
-  words <- tolower(stri_extract_all_words(m, simplify = TRUE))
-  return(length(intersect(colors, words)) > 0)
-}
-
-mentionedShape <- function(m) {
-  shapes <- c("star","square", "circle", "triangle")
-  words <- tolower(stri_extract_all_words(m, simplify = TRUE))
-  return(length(intersect(shapes, words)) > 0)
-}
-
-
-d <- rawD %>%
-  filter(role == 'speaker') %>%
-  group_by(gameid, trialNum, context, occlusions) %>%
-  summarize(text = paste0(text, collapse = ' ')) %>%   # Concatenate multiple utterance on same trial
-  mutate(numRawWords = str_count(text, "\\S+")) %>%
-  mutate(hidden = ifelse(occlusions == 'none', 'no', 'yes')) %>%
-  rowwise() %>%
-  mutate(colorMention = mentionedOutlineColor(text),
-         shapeMention = mentionedShape(text),
-         textureMention = '')
-
-write_csv(d, 'messages_no_annotations.csv')

### 2020-12-19 data/final/experiment1/planned_sample/turk/compile.R

@@ -1,27 +1,0 @@
-library(tidyverse)
-library(jsonlite)
-
-path <- "./"
-files = dir(path, pattern = "*.json")
-
-subjInfo = data.frame()
-
-for (f in files) {
-  jf = paste0(path, f)
-  jd = read_json(jf)
-  
-  newRow = data.frame()
-  if(!is.null(jd$answers$subject_information)) {
-    newRow = data.frame(gameid = jd$answers$subject_information$gameID,
-                        role = jd$answers$subject_information$role,
-                        gameLength = jd$answers$subject_information$totalLength,
-                        thinksHuman = ifelse(!is.null(jd$answers$subject_information$thinksHuman), jd$answers$subject_information$thinksHuman, NA),
-                        understandsInstructions = jd$answers$subject_information$confused,
-                        nativeEnglish = jd$answers$subject_information$nativeEnglish,
-                        ratePartner = jd$answers$subject_information$ratePartner,
-                        comments = jd$answers$subject_information$comments,
-                        strategy = jd$answers$subject_information$strategy)
-    subjInfo = bind_rows(subjInfo, newRow)
-  }
-}
-write_csv(subjInfo, './subjInfo.csv')

### 2020-12-19 data/final/experiment2/turk/compile.R

@@ -1,27 +1,0 @@
-library(tidyverse)
-library(jsonlite)
-
-path <- "./"
-files = dir(path, pattern = "*.json")
-
-subjInfo = data.frame()
-
-for (f in files) {
-  jf = paste0(path, f)
-  jd = read_json(jf)
-  print(jf)
-  newRow = data.frame()
-  if(!is.null(jd$answers$subject_information)) {
-    newRow = data.frame(gameid = jd$answers$subject_information$gameID,
-                        role = jd$answers$subject_information$role,
-                        gameLength = jd$answers$subject_information$totalLength,
-                        thinksHuman = ifelse(!is.null(jd$answers$subject_information$thinksHuman), jd$answers$subject_information$thinksHuman, NA),
-                        understandsInstructions =  ifelse(!is.null(jd$answers$subject_information$confused), jd$answers$subject_information$confused, NA),
-                        nativeEnglish =  ifelse(!is.null(jd$answers$subject_information$nativeEnglish), jd$answers$subject_information$nativeEnglish, NA),
-                        ratePartner = ifelse(!is.null(jd$answers$subject_information$ratePartner), jd$answers$subject_information$ratePartner, NA),
-                        comments = jd$answers$subject_information$comments,
-                        strategy = jd$answers$subject_information$strategy)
-    subjInfo = bind_rows(subjInfo, newRow)
-  }
-}
-write_csv(subjInfo, './subjInfo.csv')

### 2020-12-19 experiments/manuscript/app.js
only keep cogsci experiments

@@ -1,141 +1,0 @@
-/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 2013 Robert XD Hawkins
-
-    originally written for: http://buildnewgames.com/real-time-multiplayer/
-    
-    substantially modified for collective behavior experiments 
-
-    MIT Licensed.
-*/
-
-global.__base = __dirname + '/';
-
-var 
-    use_https     = true,
-    argv          = require('minimist')(process.argv.slice(2)),
-    https         = require('https'),
-    fs            = require('fs'),
-    app           = require('express')(),
-    _             = require('underscore'),
-    Server        = require('./sharedUtils/serverBase.js');
-
-var gameport;
-
-if(argv.gameport) {
-  gameport = argv.gameport;
-  console.log('using port ' + gameport);
-} else {
-  gameport = 8888;
-  console.log('no gameport specified: using 8888\nUse the --gameport flag to change');
-}
-
-if(argv.expname) {
-  var exp = argv.expname.replace(/\/$/, "");
-  var gameServer = new Server(exp);  
-} else {
-  throw new Error("missing arguments. Use --expname flag (e.g. 'node app.js --expname spatial')");
-}
-
-try {
-  var privateKey  = fs.readFileSync('/etc/apache2/ssl/stanford-cogsci.org.key'),
-      certificate = fs.readFileSync('/etc/apache2/ssl/stanford-cogsci.org.crt'),
-      intermed    = fs.readFileSync('/etc/apache2/ssl/intermediate.crt'),
-      options     = {key: privateKey, cert: certificate, ca: intermed},
-      server      = require('https').createServer(options,app).listen(gameport),
-      io          = require('socket.io')(server);
-} catch (err) {
-  console.log("cannot find SSL certificates; falling back to http");
-  var server      = app.listen(gameport),
-      io          = require('socket.io')(server);
-}
-
-var utils = require('./sharedUtils/sharedUtils.js');
-
-var global_player_set = {};
-
-// Log something so we know that server-side setup succeeded
-console.log("info  - socket.io started");
-console.log('\t :: Express :: Listening on port ' + gameport );
-
-app.get( '/*' , function( req, res ) {
-  var id = req.query.workerId;
-  if(!id || id === 'undefined') {
-    // If no worker id supplied (e.g. for demo), allow to continue
-    return utils.serveFile(req, res);
-  } else if(!valid_id(id)) {
-    // If invalid id, block them    
[…]

### 2020-12-19 experiments/manuscript/experiment1/game.server.js

@@ -1,167 +1,0 @@
-/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 2013 Robert XD Hawkins
-
-    written by : http://underscorediscovery.com
-    written for : http://buildnewgames.com/real-time-multiplayer/
-
-    modified for collective behavior experiments on Amazon Mechanical Turk
-
-    MIT Licensed.
-*/
-    var
-        fs    = require('fs'),
-        utils = require(__base + '/sharedUtils/sharedUtils.js');
-
-// This is the function where the server parses and acts on messages
-// sent from 'clients' aka the browsers of people playing the
-// game. For example, if someone clicks on the map, they send a packet
-// to the server (check the client_on_click function in game.client.js)
-// with the coordinates of the click, which this function reads and
-// applies.
-var onMessage = function(client,message) {
-  //Cut the message up into sub components
-  var message_parts = message.split('.');
-
-  //The first is always the type of message
-  var message_type = message_parts[0];
-
-  //Extract important variables
-  var gc = client.game;
-  var id = gc.id;
-  var all = gc.get_active_players();
-  var target = gc.get_player(client.userid);
-  var others = gc.get_others(client.userid);
-  switch(message_type) {
-
-  case 'chatMessage' :
-    if(client.game.player_count == 2 && !gc.paused) {
-      var msg = message_parts[1].replace(/~~~/g,'.');
-      _.map(all, function(p){
-	p.player.instance.emit( 'chatMessage', {user: client.userid, msg: msg});});
-    }
-    break;
-
-  case 'updateMouse' :
-    break;
-    
-  case 'clickedObj' :
-    // Write event to file
-    others[0].player.instance.send('s.feedback.' + message_parts[1]);
-    target.instance.send('s.feedback.' + message_parts[1]);
-
-    // Continue
-    gc.newRound(1500);
-    break;
-
-  case 'playerTyping' :
-    _.map(others, function(p) {
-      p.player.instance.emit( 'playerTyping', {typing: message_parts[1]});
-    });
-    break;
-
-  case 'h' : // Receive message when browser focus shifts
-    target.visible = message_parts[1];
-    break;
-  }
-};
-
-/*
-  Associates events in onMessage with callback returning json to be saved
-  {
-    <eventName>: (client, message_parts) => {<datajson>}
-  }
[…]

### 2020-12-19 experiments/manuscript/experiment1/makeObjectsJSON.R

@@ -1,11 +1,0 @@
-library(tidyverse)
-library(jsonlite)
-filenames <- list.files("./images/stimuli", pattern="*.png", full.names=TRUE)
-output = data.frame(url = filenames) %>% 
-  mutate(id = row_number()) %>%
-  separate(url, into = c('tmp', 'garbage'), remove = F, sep = '.png') %>%
-  separate(tmp, into = c('garbage2', 'name'), sep = './images/stimuli/') %>%
-  separate(name, into = c('texture', 'color', 'shape'), remove = F) %>%
-  select(-garbage, -garbage2)
-
-write_json(output, path = './images/objects.json')

### 2020-12-19 experiments/manuscript/experiment2/drawing.js

@@ -1,261 +1,0 @@
-function constructOcclusions() {
-  globalGame.occludedList = [[4,4], [2,2], [3,2], [1,3], [4,1]];
-  globalGame.occlusionImages = [];
-  globalGame.occlusionCount = globalGame.occludedList.length;
-  _.map(globalGame.occludedList, function(loc) {
-    var cell = globalGame.getPixelFromCell({gridX: loc[0], gridY: loc[1]});
-    var imgObj = new Image();
-    imgObj.onload = occlusionCounter;
-    imgObj.src = (globalGame.my_role == globalGame.playerRoleNames.role1 ?
-		  './stimuli/mystery.jpg' :
-		  './stimuli/mystery_noQ.jpg');
-    globalGame.occlusionImages.push(_.extend(cell, {img: imgObj}));
-  });
-};
-
-function containsCell(cellList, cell) {
-  return _.some(cellList, function(compCell) {
-    return _.isEqual(cell, [compCell.gridX, compCell.gridY]);
-  });
-};
-
-// common loader keeping track if loads
-function occlusionCounter() {
-  globalGame.occlusionCount--;
-  if (globalGame.occlusionCount === 0 && !globalGame.paused)
-    drawOcclusionImages();
-}
-
-// called when all images are loaded
-function drawOcclusionImages() {
-  for(var i = 0; i < globalGame.occlusionImages.length; i++) {
-    var obj = globalGame.occlusionImages[i];
-    globalGame.ctx.drawImage(obj.img, obj.upperLeftX, obj.upperLeftY,
-			     obj.width, obj.height);
-  }
-}
-
-// common loader keeping track if loads
-function objectCounter() {
-  globalGame.objectCount--;
-  if (globalGame.objectCount === 0 && !globalGame.paused)
-    drawObjects();
-}
-
-var drawGrid = function(game){
-  //size of canvas
-  var cw = game.viewport.width;
-  var ch = game.viewport.height;
-
-  //padding around grid
-  var p = 0;
-
-  //grid width and height
-  var bw = cw - (p*2) ;
-  var bh = ch - (p*2) ;
-
-  game.ctx.beginPath();
-
-  // vertical lines
-  for (var x = 0; x <= bw; x += Math.floor((cw - 2*p) / 4)) {
-    game.ctx.moveTo(0.5 + x + p, p);
[…]

### 2020-12-19 experiments/manuscript/experiment2/game.server.js

@@ -1,174 +1,0 @@
-/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 2013 Robert XD Hawkins
-    
-    written by : http://underscorediscovery.com
-    written for : http://buildnewgames.com/real-time-multiplayer/
-    
-    modified for collective behavior experiments on Amazon Mechanical Turk
-
-    MIT Licensed.
-*/
-
-//require('look').start()
-
-var utils       = require('../sharedUtils/sharedUtils.js'),
-    fs          = require('fs');
-	    
-var moveObject = function(client, i, x, y) {
-  var obj = client.game.objects[i];
-  var others = client.game.get_others(client.userid);
-  obj.trueX = parseInt(x);
-  obj.trueY = parseInt(y);
-  _.map(others, function(p) {
-    p.player.instance.emit('objMove', {i: i, x: x, y: y});
-  });
-};
-
-var onMessage = function(client,message) {
-  //Cut the message up into sub components
-  var message_parts = message.split('.');
-
-  //The first is always the type of message
-  var message_type = message_parts[0];
-
-  //Extract important variables
-  var all = client.game.get_active_players();
-  var target = client.game.get_player(client.userid);
-  var others = client.game.get_others(client.userid);
-  switch(message_type) {
-  case 'objMove' :    // Client is changing angle
-    moveObject(client, message_parts[1], message_parts[2], message_parts[3]);
-    break;
-
-  case 'drop' :
-    var type = message_parts[1];
-    moveObject(client, message_parts[2], message_parts[3], message_parts[4]);
-    var extraData = message_parts[5] + '.' + message_parts[6];
-    if(type == 'correct') {
-      client.game.attemptNum = 0;
-      _.map(all, function(p) {
-	p.player.instance.send("s.feedback.correct." + extraData);});
-    } else {
-      client.game.attemptNum += 1;
-      _.map(all, function(p) {p.player.instance.send("s.feedback.incorrect." + extraData);});
-      client.game.instructionNum -= 1;
-    }
-    setTimeout(() => client.game.newRound(), 3000);
-    break;
-
-  case 'chatMessage' :
-    // Update others
-    var msg = message_parts[1].replace(/-/g,'.');
-    _.map(all, function(p){
-      p.player.instance.emit( 'chatMessage', {user: client.userid, msg: msg});
-    });
-    break;
-
-  case 'updateMouse' :
-    break;
-
-  case 'mouseOverDistractor' :
-    break;
[…]

### 2020-12-19 experiments/manuscript/normingExperiment/_shared/js/utils.js

@@ -1,358 +1,0 @@
-var utils = {
-	showSlide: function(slidename) {
-			$(".slide").hide();
-			$(".fullslide").hide();
-			$("#" + slidename).show();
-	},
-
-	get_exp_length: function() {
-	  if (exp.structure && exp.slides) {
-	    var length = 0;
-	    for (var i=0; i<exp.structure.length; i++) {
-	      var a_slide = exp.slides[exp.structure[i]];
-	      if (a_slide.present) {
-		length += a_slide.present.length;
-	      } else {
-		length ++;
-	      }
-	    }
-	    return length - 1; //don't count the last slide
-	  }
-	},
-
-	/* eq_el_height and match_row_height adapted from:
-	 * http://stackoverflow.com/questions/15397966/how-to-get-table-cells-to-have-equal-height
-	 */
-	eq_el_height: function(els, min_or_max) {
-			els.each(function() {
-					$(this).height('auto');
-			});
-
-			var m = $(els[0]).height();
-
-			els.each(function() {
-					var h = $(this).height();
-
-					if (min_or_max === "max") {
-							m = h > m ? h : m;
-					} else {
-							m = h < m ? h : m;
-					}
-			});
-
-			els.each(function() {
-					$(this).height(m);
-			});
-	},
-
-	match_row_height: function(label, sublabel) {
-		$(label).each(function() {
-			utils.eq_el_height($(this).find(sublabel/*'tr'*/), "max");
-		});
-	},
-
-  make_slider: function(label, response_callback, on_mouse_up_callback) {
-		$(label).empty();
-		$(label).slider({
-			range : "min",
-			min : 0,
-			max : 1,
-			step: 0.01,
-			value : 0.5,
-			slide : response_callback,
-		  change : response_callback,
-		  stop : on_mouse_up_callback
-		});
-		$(label + ' .ui-slider-handle').hide();
-				$(label).mousedown(function(){
-			$(label + ' .ui-slider-handle').show();
-				$(label).css({"background":"#99D6EB"});
-				$(label + ' .ui-slider-handle').css({
-					"background":"#667D94",
-					"border-color": "#001F29"
-				});
-				});
-				$(label).css({"background":"#eee"});
-	}
-}
-/*
-		var keys = getUnique.apply(list_keys);
-		//var keys = _.unique(list_keys);
-		keys.push(['sid', 1]);
[…]

### 2020-12-19 experiments/manuscript/normingExperiment/experiment/js/experiment.js

@@ -1,141 +1,0 @@
-function make_slides(f) {
-  var q_first = null;
-  var slides = {};
-  slides.i0 = slide({
-    name : "i0",
-    start: function() {
-      exp.startT = Date.now();
-    }
-  });
-   
-  slides.rating = slide({
-    name: "rating",
-    present : _.shuffle(stimList),
-    present_handle : function(stim) {
-      // Set up scenario & instructions
-      this.trial_start = Date.now();
-      this.stim = stim;
-      this.init_sliders();
-      exp.sliderPost = {};
-      console.log(stim);
-      var contextsentence = "<strong>"+this.stim.text+"</strong>";
-      var objimagehtml = '<center><img src="'+this.stim.object.contextURL+'" style="height:500px;"></center>';
-      $("#contextsentence").html(contextsentence);
-      $("#objectimage").html(objimagehtml);
-      console.log(this);
-      //$(".err1").hide();
-    },
-
-    init_sliders : function() {
-      // Takes a callback for 'change' and a callback for 'mouseup'
-      var that = this;
-      utils.make_slider("#single_slider", function(event, ui) {
-	exp.sliderPost = ui.value;
-      }, function(event, ui) {
-	if (exp.sliderPost > -1 && exp.sliderPost < 16) {
-	  //$(".err").hide();
-	  that.log_responses();
-	  _stream.apply(that); //Use exp.go() if and only if there is no "present" data.
-	} else {
-	  //$(".err").show();
-	}
-      });
-    },
-    log_responses : function() {
-      exp.data_trials.push({
-        "label" : this.stim.text,
-        "slideNumber": exp.phase,
-        "response" : exp.sliderPost,
-	"objectSet" : this.stim.objectSet,
-	"referent" : this.stim.referent,
-	"context" : this.stim.context,
-	"rt" : Date.now() - _s.trial_start
-      });
-    },
-  });
-  
-  slides.subj_info =  slide({
-    name : "subj_info",
-    submit : function(e){
-      //if (e.preventDefault) e.preventDefault(); // I don't know what this means.
-      exp.subj_data = {
-        language : $("#language").val(),
-        enjoyment : $("#enjoyment").val(),
-        asses : $('input[name="assess"]:checked').val(),
-        age : $("#age").val(),
-        gender : $("#gender").val(),
-        education : $("#education").val(),
-        comments : $("#comments").val(),
-      };
[…]

### 2020-12-19 experiments/manuscript/normingExperiment/experiment/js/genObjectSet.js

@@ -1,29 +1,0 @@
-var fs = require('fs');
-var babyparse = require('papaparse');
-var _ = require('underscore');
-
-function readCSV(filename){
-    return babyparse.parse(fs.readFileSync(filename, 'utf8'),
-			   {header: true, skipEmptyLines : true}
-			  ).data;
-  };
-
-var criticalMessages = readCSV('../../../../../data/final/experiment1/refExpressions_standardized.csv');
-
-var unscriptedLabels = _.map(criticalMessages, function(row) {
-  return {text: row.text.trim(), objectSet : row.objectSet};
-});
-
-var scriptedLabels = [{text: "glasses", objectSet : 1},
-		      {text: "bottom block", objectSet : 2},
-		      {text: "tape",  objectSet : 3},
-		      {text: "large measuring cup", objectSet : 4},
-		      {text: "brush", objectSet : 5},
-		      {text: "eraser", objectSet : 6},
-		      {text: "small candle", objectSet : 7},
-		      {text: "mouse", objectSet : 8}];
-
-var uniqueLabels = _.unique(unscriptedLabels.concat(scriptedLabels),
-			    function(item, key, a) {  return item.text; });
-
-fs.writeFileSync("uniqueLabels.json", JSON.stringify(uniqueLabels, null, 4));

### 2020-12-19 experiments/manuscript/normingExperiment/experiment/js/objectSet.js

@@ -1,147 +1,0 @@
-// Requires importing uniqueLabels.js before this in html
-
-var stimList = function() {
-  // BLOCK 1
-  var sunGlasses = {
-    url: 'js/stimuli/sunGlasses.png', name: "sunGlasses", width: 130, height: 65,
-    instruction: "sunGlasses right", initialLoc : [4,3], critical : "target",
-    scriptedInstruction: "move the glasses one space to the right",
-    contextURL: 'js/stimuli/sunglassesContext.png'};
-  var glassesCase = {
-    url: 'js/stimuli/glassesCase.png', name: "glassesCase", width : 100, height: 107,
-    instruction: "", initialLoc: [3,1], critical : "distractor",
-    contextURL: 'js/stimuli/glassesCaseContext.png'};
-
-  // BLOCK 2
-  var middleBlock = {
-    url: 'js/stimuli/middleBlock.png', name: "middleBlock", width: 80, height: 80,
-    instruction: "middleBlock left", initialLoc: [3,4],
-    contextURL: 'js/stimuli/middleBlockContext.png',
-    scriptedInstruction: "move the bottom block one space to the left", critical : "target"}
-  var bottomBlock = {
-    url: 'js/stimuli/bottomBlock.png', name: "bottomBlock", width: 80, height: 80,
-    contextURL: 'js/stimuli/bottomBlockContext.png',
-    instruction: "", initialLoc: [4,4], critical : "distractor"}
-
-  // BLOCK 3
-  var cassetteTape = {
-    url: 'js/stimuli/cassetteTape.png', name: "cassetteTape",  width: 87.5, height: 55,
-    instruction: "cassetteTape down", initialLoc: [1,3], critical : "target",
-    contextURL: 'js/stimuli/cassetteTapeContext.png',
-    scriptedInstruction: "move the tape down by one space",}
-  var rollOfTape = {
-    url: 'js/stimuli/rollOfTape.jpg', name: "rollOfTape", width: 83, height: 83,
-    contextURL: 'js/stimuli/rollOfTapeContext.png',
-    instruction: "", initialLoc: [1,4], critical : "distractor"}
-
-  // BLOCK 4
-
-  var mediumMeasuringCup = {
-    url: 'js/stimuli/mediumMeasuringCup.png', name: "mediumMeasuringCup", width : 110,
-    height: 87, contextURL: 'js/stimuli/mediumCupContext.png',
-    instruction: "mediumMeasuringCup down", initialLoc : [3,4], critical : "target",
[…]

### 2020-12-19 experiments/manuscript/normingExperiment/experiment/js/uniqueLabels.js

@@ -1,494 +1,0 @@
-var uniqueLabels = [
-    {
-        "text": "candle",
-        "objectSet": "7"
-    },
-    {
-        "text": "cassette",
-        "objectSet": "3"
-    },
-    {
-        "text": "mouse",
-        "objectSet": "8"
-    },
-    {
-        "text": "dark sunglasses",
-        "objectSet": "1"
-    },
-    {
-        "text": "thing that looks like a brush",
-        "objectSet": "5"
-    },
-    {
-        "text": "glass measuring cup",
-        "objectSet": "4"
-    },
-    {
-        "text": "block with a blue B showing",
-        "objectSet": "2"
-    },
-    {
-        "text": "expo dry erase eraser",
-        "objectSet": "6"
-    },
-    {
-        "text": "B block",
-        "objectSet": "2"
-    },
-    {
-        "text": "measuring cup",
-        "objectSet": "4"
-    },
-    {
-        "text": "hairbrush",
-        "objectSet": "5"
-    },
-    {
-        "text": "sunglasses",
-        "objectSet": "1"
-    },
-    {
-        "text": "eraser",
-        "objectSet": "6"
-    },
-    {
-        "text": "cassette tape",
-        "objectSet": "3"
-    },
-    {
-        "text": "whiteboard eraser",
-        "objectSet": "6"
-    },
-    {
-        "text": "child's B block",
-        "objectSet": "2"
-    },
-    {
-        "text": "computer mouse",
-        "objectSet": "8"
-    },
-    {
-        "text": "glass measuring cup with handle",
-        "objectSet": "4"
-    },
-    {
-        "text": "radio cassette",
-        "objectSet": "3"
-    },
-    {
-        "text": "block with blue letter B facing up",
-        "objectSet": "2"
-    },
-    {
-        "text": "red candle",
-        "objectSet": "7"
-    },
-    {
-        "text": "black sunglasses",
-        "objectSet": "1"
-    },
-    {
-        "text": "clear pyrex measuring cup",
-        "objectSet": "4"
-    },
-    {
-        "text": "expo whiteboard eraser",
-        "objectSet": "6"
-    },
-    {
-        "text": "round barrel hairbrush in bottom row",
-        "objectSet": "5"
-    },
-    {
-        "text": "3rd row block",
-        "objectSet": "2"
-    },
-    {
-        "text": "Row 1 cassette tape",
-        "objectSet": "3"
-    },
-    {
[…]

### 2020-12-19 experiments/manuscript/sharedUtils/clientBase.js

@@ -1,269 +1,0 @@
-var visible;
-
-var getURLParams = function() {
-  var match,
-      pl     = /\+/g,  // Regex for replacing addition symbol with a space
-      search = /([^&=]+)=?([^&]*)/g,
-      decode = function (s) { return decodeURIComponent(s.replace(pl, " ")); },
-      query  = location.search.substring(1);
-
-  var urlParams = {};
-  while ((match = search.exec(query))) {
-    urlParams[decode(match[1])] = decode(match[2]);
-  }
-  return urlParams;
-};
-
-var ondisconnect = function(data) {
-  // Redirect to exit survey
-  console.log("server booted");
-  this.viewport.style.display="none";
-  var email = globalGame.email ? globalGame.email : '';
-  var failMsg = [
-    '<h3>Oops! It looks like your partner lost their connection!</h3>',
-    '<p> Completing this survey will submit your HIT so you will still receive full ',
-    'compensation.</p> <p>If you experience any problems, please email us (',
-    email, ')</p>'
-  ].join('');
-  var successMsg = [
-    "<h3>Thanks for participating in our experiment!</h3>",
-    "<p>Before you submit your HIT, we'd like to ask you a few questions.</p>"
-  ].join('');
-
-  console.log(globalGame.email);
-
-  if(globalGame.roundNum + 2 > globalGame.numRounds) { 
-    $('#exit_survey').prepend(successMsg);    
-  } else {
-    $('#exit_survey').prepend(failMsg); 
-  }
-
-  $('#exit_survey').show();
-  $('#main').hide();
-  $('#header').hide();
-  
-  $('#message_panel').hide();
-  $('#submitbutton').hide();
-  $('#roleLabel').hide();
-  $('#instructs').hide();  
-  $('#score').hide();
-
-  $('#post_test').hide();
-  $('#sketchpad').hide(); // this is from sketchpad experiment (jefan 4/23/17)
-  $('#loading').hide(); // this is from sketchpad experiment (jefan 4/23/17)
-};
-
-// The server responded that we are now in a game
-var onconnect = function(data) {
-  this.my_id = data.id;
-  this.players[0].id = this.my_id;
-  this.urlParams = getURLParams();
-  drawScreen(this, this.get_player(this.my_id));
-};
-
-// Associates callback functions corresponding to different socket messages
-var sharedSetup = function(game) {
-  //Store a local reference to our connection to the server
-  game.socket = io.connect({reconnection: false});
-
[…]

### 2020-12-19 experiments/manuscript/sharedUtils/serverBase.js

@@ -1,139 +1,0 @@
-var utils = require(__base + 'sharedUtils/sharedUtils.js');
-
-global.window = global.document = global;
-
-class ReferenceGameServer {
-  constructor(expName) {
-    this.expName = expName;
-    this.core = require([__base, expName, 'game.core.js'].join('/')).game_core;
-    this.player = require([__base, expName, 'game.core.js'].join('/')).game_player;
-    this.customServer = require([__base, expName, 'game.server.js'].join('/'));
-    this.setCustomEvents = this.customServer.setCustomEvents;
-    
-    // Track ongoing games
-    this.games = {};
-    this.game_count = 0;
-  }
-
-  // if game relies on asynchronous stim logic, need to wait until everything
-  // is fetched before starting game (otherwise race conditions)
-  startGame(game) {
-    console.log('starting game');
-    game.newRound(0);
-  }
-  /*
-    Writes data specified by experiment instance to csv and/or mongodb
-  */
-  writeData (client, eventType, message_parts) {
-    var output = this.customServer.dataOutput;
-    var game = client.game;
-    if(_.has(output, eventType)) {
-      var dataPoint = _.extend(output[eventType](client, message_parts), {eventType});
-      if(_.includes(game.dataStore, 'csv'))
-	utils.writeDataToCSV(game, dataPoint);
-      if(_.includes(game.dataStore, 'mongo'))
-	utils.writeDataToMongo(game, dataPoint); 
-    }
-  }
-
-  onMessage (client, message) {
-    var message_parts = message.split('.');
-    this.customServer.onMessage(client, message);
-    if(!_.isEmpty(client.game.dataStore)) {
-      this.writeData(client, message_parts[0], message_parts);
-    }
-  }
-
-  findGame (player) {
-    this.log('looking for a game. We have : ' + this.game_count);
-    var joined_a_game = false;
-    for (var gameid in this.games) {
-      var game = this.games[gameid];
-      if(game.player_count < game.players_threshold) {
-	// End search
-	joined_a_game = true;
-
-	// Add player to game
-	game.player_count++;
-	game.players.push({
-	  id: player.userid,
-	  instance: player,
-	  player: new this.player(game, player)
-	});
-
-	// Add game to player
-	player.game = game;
-	player.role = game.playerRoleNames.role1;
-	player.send('s.join.' + game.players.length + '.' + player.role);
-
-	// notify existing players that someone new is joining
-	_.map(game.get_others(player.userid), function(p){
[…]

### 2020-12-19 experiments/manuscript/sharedUtils/sharedUtils.js

@@ -1,282 +1,0 @@
-var _ = require('underscore');
-var fs = require('fs');
-var path = require('path');
-var converter = require("color-convert");
-var DeltaE = require('../node_modules/delta-e');
-var mkdirp = require('mkdirp');
-var sendPostRequest = require('request').post;
-
-var serveFile = function(req, res) {
-  var fileName = req.params[0];
-  if(fileName.includes("html"))
-     console.log('\t :: Express :: file requested: ' + fileName);
-  return res.sendFile(fileName, {root: __base}); 
-};
-
-var handleDuplicate = function(req, res) {
-  console.log("duplicate id: blocking request");
-  return res.redirect('/sharedUtils/duplicate.html');
-};
-
-var handleInvalidID = function(req, res) {
-  console.log("invalid id: blocking request");
-  return res.redirect('/sharedUtils/invalid.html');
-};
-
-var checkPreviousParticipant = function(workerId, callback) {
-  var p = {'workerId': workerId};
-  var postData = {
-    dbname: 'ToM',
-    query: p,
-    projection: {'_id': 1}
-  };
-  sendPostRequest(
-    'http://localhost:5000/db/exists',
-    {json: postData},
-    (error, res, body) => {
-      try {
-	if (!error && res.statusCode === 200) {
-	  console.log("success! Received data " + JSON.stringify(body));
-	  callback(body);
-	} else {
-	  throw `${error}`;
-	}
-      } catch (err) {
-	console.log(err);
-	console.log('no database; allowing participant to continue');
-	return callback(false);
-      }
-    }
-  );
-};
-
-var writeDataToCSV = function(game, _dataPoint) {
-  var dataPoint = _.clone(_dataPoint);  
-  var eventType = dataPoint.eventType;
-
-  // Omit sensitive data
-  if(game.anonymizeCSV) 
-    dataPoint = _.omit(dataPoint, ['workerId', 'assignmentId']);
-  
-  // Establish stream to file if it doesn't already exist
-  if(!_.has(game.streams, eventType))
-    establishStream(game, dataPoint);    
-
-  var line = _.values(dataPoint).join('\t') + "\n";
-  game.streams[eventType].write(line, err => {if(err) throw err;});
-};
-
-var writeDataToMongo = function(game, line) {
-  var postData = _.extend({
[…]

### 2020-12-19 experiments/manuscript/store.js

@@ -1,235 +1,0 @@
-'use strict';
-
-const _ = require('lodash');
-const bodyParser = require('body-parser');
-const express = require('express');
-const fs = require('fs');
-const mongodb = require('mongodb');
-const path = require('path');
-const sendPostRequest = require('request').post;
-const colors = require('colors/safe');
-
-const app = express();
-const MongoClient = mongodb.MongoClient;
-const port = 5000;
-const mongoCreds = require('./auth.json');
-const mongoURL = `mongodb://${mongoCreds.user}:${mongoCreds.password}@localhost:27017/`;
-const handlers = {};
-
-
-function makeMessage(text) {
-  return `${colors.blue('[store]')} ${text}`;
-}
-
-function log(text) {
-  console.log(makeMessage(text));
-}
-
-function error(text) {
-  console.error(makeMessage(text));
-}
-
-function failure(response, text) {
-  const message = makeMessage(text);
-  console.error(message);
-  return response.status(500).send(message);
-}
-
-function success(response, text) {
-  const message = makeMessage(text);
-  //console.log(message);
-  return response.send(message);
-}
-
-function mongoConnectWithRetry(delayInMilliseconds, callback) {
-  MongoClient.connect(mongoURL, (err, connection) => {
-    if (err) {
-      console.error(`Error connecting to MongoDB: ${err}`);
-      setTimeout(() => mongoConnectWithRetry(delayInMilliseconds, callback), delayInMilliseconds);
-    } else {
-      log('connected succesfully to mongodb');
-      callback(connection);
-    }
-  });
-}
-
-// Keep track of which games have used each stim
-// function recordStimUse(stimdb, gameid, idList) {
-//   _.forEach(idList, id => {
-//     stimdb.update({_id: id}, {
-//       $push : {games : gameid},
-//       $inc  : {numGames : 1}
-//     }, {multi: true}, function(err, items) {
-//       // do something when done?
-//     });
-//   });
-// }
-
-// // this function isn't currently being used, but might be once there are >1 databases
-// function getDatabaseList(connection, callback) {
-//    connection.admin().listDatabases(function(err, result) {
-//       if (err) {
-//         console.error('Error listing databases from this MongoDB connection');
-//       } else {
-//         log('success: got list of databases from this mongo connection')
-//         callback(result.databases);
-//       }
-//    })
-// }
-
-// // also not being used, but might be later to scale to new db/cols
-// function getCollectionList(db, callback) {
-//     db.listCollections().toArray(function(err, collections) {
[…]

### 2020-12-19 experiments/manuscript/supplemental_exp/drawing.js

@@ -1,217 +1,0 @@
-// drawing.js
-// This file contains functions to draw on the HTML5 canvas
-
-function containsCell(cellList, cell) {
-  return _.some(cellList, function(compCell) {
-    return _.isEqual(cell, [compCell.gridX, compCell.gridY]);
-  });
-};
-
-// called when all images are loaded
-function drawOcclusionImages() {
-  for(var i = 0; i < globalGame.occlusionImages.length; i++) {
-    var obj = globalGame.occlusionImages[i];
-    globalGame.ctx.drawImage(obj.img, obj.upperLeftX, obj.upperLeftY,
-			     obj.width, obj.height);
-  }
-  drawObjects(globalGame);
-}
-
-// common loader keeping track if loads
-function occlusionCounter() {
-  globalGame.occlusionCount--;
-  if (globalGame.occlusionCount === 0)
-    drawOcclusionImages();
-}
-
-// Draws a grid of cells on the canvas (evenly divided
-var drawGrid = function(game){
-    //size of canvas
-    var cw = game.viewport.width;
-    var ch = game.viewport.height;
-
-    //padding around grid
-    var p = game.cellPadding / 2;
-
-    //grid width and height
-    var bw = cw - (p*2) ;
-    var bh = ch - (p*2) ;
-
-    game.ctx.beginPath();
-
-    // vertical lines
-  for (var x = 0; x <= bw; x += Math.floor((cw - 2*p) / game.numHorizontalCells)) {
-        game.ctx.moveTo(0.5 + x + p, p);
-        game.ctx.lineTo(0.5 + x + p, bh + p);}
-
-    // horizontal lines
-    for (var x = 0; x <= bh; x += Math.floor((ch - 2*p) / game.numVerticalCells)) {
-        game.ctx.moveTo(p, 0.5 + x + p);
-        game.ctx.lineTo(bw + p, 0.5 + x + p);}
-
-  game.ctx.lineWidth = '5';
-  game.ctx.strokeStyle = "#000000";
-  game.ctx.stroke();
-};
-
-// Preload so that all images render at same time
-var loadImages = function(sources, callback) {
-  var numLoaded = 0;
-  var images = {};
-  _.forEach(sources, function(source) {
-    images[source.name] = new Image();
-    images[source.name].onload = function() {
-      if(++numLoaded >= sources.length) {
-	callback(images);
-      }
-    };
-    images[source.name].src = source.url;
-  });
-};
-
-// Loop through the object list and draw each one in its specified location
[…]

### 2020-12-19 experiments/manuscript/supplemental_exp/game.client.js

@@ -1,265 +1,0 @@
-//   Copyright (c) 2012 Sven "FuzzYspo0N" Bergström,
-//                   2013 Robert XD Hawkins
-
-//     written by : http://underscorediscovery.com
-//     written for : http://buildnewgames.com/real-time-multiplayer/
-
-//     modified for collective behavior experiments on Amazon Mechanical Turk
-
-//     MIT Licensed.
-
-// A window global for our game root variable.
-var globalGame = {};
-// Keeps track of whether player is paying attention...
-var incorrect;
-var dragging;
-var waiting;
-
-var client_onserverupdate_received = function(data){
-  globalGame.my_role = data.trialInfo.roles[globalGame.my_id];
-
-  // Update client versions of variables with data received from
-  // server_send_update function in game.core.js
-  //data refers to server information
-  if(data.players) {
-    _.map(_.zip(data.players, globalGame.players),function(z){
-      z[1].id = z[0].id;
-    });
-  }
-  
-  if (globalGame.roundNum != data.roundNum) {
-    globalGame.objects = data.trialInfo.currStim.objects;
-    globalGame.currUtterance = data.trialInfo.currUtterance;
-    globalGame.occlusions = data.trialInfo.currStim.occlusions;
-  };
-
-  globalGame.game_started = data.gs;
-  globalGame.players_threshold = data.pt;
-  globalGame.player_count = data.pc;
-  globalGame.roundNum = data.roundNum;
-  globalGame.roundStartTime = new Date();
-  globalGame.allObjects = data.allObjects;
-  
-  if(!_.has(globalGame, 'data')) {
-    globalGame.data = data.dataObj;
-  }
-
-  // Get rid of "waiting" screen if there are multiple players
-  $('#messages').empty();
-
-  // reset labels
-  // Update w/ role (can only move stuff if agent)
-  $('#roleLabel').empty().append("You are the " + globalGame.my_role + '.');
-
-  if(globalGame.my_role === globalGame.playerRoleNames.role1) {
-    globalGame.viewport.removeEventListener("click", mouseClickListener, false);
-    $('#instructs')
-      .empty()
-      .append("<p>Send a message through the chat box</p>" +
-	      "<p>to tell the listener which object is the target.</p>");
-  } else if(globalGame.my_role === globalGame.playerRoleNames.role2) {
-    $('#instructs')
-      .empty()
-      .append("<p>After you see the speaker's message, </p>" +
[…]

### 2020-12-19 experiments/manuscript/supplemental_exp/game.core.js

@@ -1,412 +1,0 @@
-/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström,
-                  2013 Robert XD Hawkins
-
- written by : http://underscorediscovery.com
-    written for : http://buildnewgames.com/real-time-multiplayer/
-
-    substantially modified for collective behavior experiments on the web
-    MIT Licensed.
-*/
-
-/*
-  The main game class. This gets created on both server and
-  client. Server creates one for each game that is hosted, and each
-  client creates one for itself to play the game. When you set a
-  variable, remember that it's only set in that instance.
-*/
-
-var has_require = typeof require !== 'undefined';
-
-if( typeof _ === 'undefined' ) {
-  if( has_require ) {
-    _ = require('lodash');
-    utils  = require(__base + 'sharedUtils/sharedUtils.js');
-    assert = require('assert');
-  }
-  else throw 'mymodule requires underscore, see http://underscorejs.org';
-}
-
-var game_core = function(options){
-  // Store a flag if we are the server instance
-  this.server = options.server ;
-
-  // Some config settings
-  this.email = 'rdhawkins@princeton.edu';
-  this.projectName = 'ToM';
-  this.experimentName = 'listenerManipulation';
-  this.iterationName = 'pilot1';
-  this.anonymizeCSV = true;
-  this.bonusAmt = 1; // in cents
-  
-  // save data to the following locations (allowed: 'csv', 'mongo')
-  this.dataStore = ['csv', 'mongo'];
-
-  // How many players in the game?
-  this.players_threshold = 1;
-  this.playerRoleNames = {
-    role2 : 'listener'
-  };
-
-  //Dimensions of world in pixels and numberof cells to be divided into;
-  this.numHorizontalCells = 3;
-  this.numVerticalCells = 3;
-  this.cellDimensions = {height : 600, width : 600}; // in pixels
-  this.cellPadding = 0;
-  this.world = {
-    height: 600 * this.numVerticalCells,
-    width: 600 * this.numHorizontalCells
-  };
-  
-  // Which round are we on (initialize at -1 so that first round is 0-indexed)
-  this.roundNum = -1;
-  this.numOcclusions = 2;
-  
-  // How many rounds do we want people to complete?
-  this.numRounds = 28;
-  this.feedbackDelay = 100;
-
-  // This will be populated with the tangram set
-  this.trialInfo = {roles: _.values(this.playerRoleNames)};
-
[…]

### 2020-12-19 experiments/manuscript/supplemental_exp/game.server.js

@@ -1,183 +1,0 @@
-/*  Copyright (c) 2012 Sven "FuzzYspo0N" Bergström, 2013 Robert XD Hawkins
-
-    written by : http://underscorediscovery.com
-    written for : http://buildnewgames.com/real-time-multiplayer/
-
-    modified for collective behavior experiments on Amazon Mechanical Turk
-
-    MIT Licensed.
-*/
-    var
-        fs    = require('fs'),
-        utils = require(__base + '/sharedUtils/sharedUtils.js');
-
-// This is the function where the server parses and acts on messages
-// sent from 'clients' aka the browsers of people playing the
-// game. For example, if someone clicks on the map, they send a packet
-// to the server (check the client_on_click function in game.client.js)
-// with the coordinates of the click, which this function reads and
-// applies.
-var onMessage = function(client,message) {
-  //Cut the message up into sub components
-  var message_parts = message.split('.');
-
-  //The first is always the type of message
-  var message_type = message_parts[0];
-
-  //Extract important variables
-  var gc = client.game;
-  var id = gc.id;
-  var all = gc.get_active_players();
-  var target = gc.get_player(client.userid);
-  var others = gc.get_others(client.userid);
-  switch(message_type) {
-
-  case 'chatMessage' :
-    var msg = message_parts[1].replace(/~~~/g,'.');
-    _.map(all, function(p){
-      p.player.instance.emit( 'chatMessage', {user: client.userid, msg: msg});});
-    break;
-
-  case 'updateMouse' :
-    break;
-    
-  case 'clickedObj' :
-    // Write event to file
-    target.instance.send('s.feedback.' + message_parts[1]);
-
-    // Continue
-    gc.newRound(1500);
-    break;
-
-  case 'playerTyping' :
-    _.map(others, function(p) {
-      p.player.instance.emit( 'playerTyping', {typing: message_parts[1]});
-    });
-    break;
-
-  case 'exitSurvey' :
-    break;
-    
-  case 'h' : // Receive message when browser focus shifts
-    target.visible = message_parts[1];
-    break;
-  }
-};
-
-/*
-  Associates events in onMessage with callback returning json to be saved
-  {
-    <eventName>: (client, message_parts) => {<datajson>}
-  }
-  Note: If no function provided for an event, no data will be written
-*/
[…]

### 2020-12-19 analysis/final_analysis_bootstrap_helpers.R
only keep cogsci analyses

@@ -1,28 +1,0 @@
-## In the inner loop, we want to sample a 'mean' rating for each utterance
-## using same sample of *raters* for all utterances... 
-sampleRatings = function(df, indices) {
-  grouped_df <- df %>%
-    group_by(label, objectSet) %>%
-    do(data.frame(diff = mean(.$diff[indices]),
-                  target = mean(.$target[indices]),
-                  distractor = mean(.$distractor[indices])))
-  return(grouped_df)
-}
-
-## In outer loop, we want to sample 'pairs' as a unit (with their 8 utterances)
-samplePairs = function(df) {
-  uniq_ids <- unique(df$gameid)
-  new_ids <- sample(uniq_ids, length(uniq_ids), replace = TRUE)
-  resampled_ids = data.frame(gameid = new_ids,
-                             sampleid = 1:length(new_ids), 
-                             stringsAsFactors = FALSE)
-  return(left_join(resampled_ids, df, by = c('gameid')))
-}
-
-## Finally, want to sample object sets per pair
-sampleObjects = function(df, indices) {
-  new_objects <- data.frame(objectSet = indices)
-  return(left_join(new_objects, df, by = c('objectSet')))
-}
-# sampleObjects(data.frame(objectSet = c(1,2,3,4,5,6,7,8), utt = c('a', 'b', 'c', 'a', 'b', 'c', 'd', 'e')),
-#               obj_sets)

### 2020-12-19 analysis/cogSci_analysis.Rmd
fix paths

--- 2017-09-27 analysis/cogSci_analysis.Rmd
@@
 ```{r}
-setwd('~/Repos/keysar_replication/experiment1/data/')
-preMouseDExp1 = read.csv('mouse/fullSampleMouse.csv')
-preErrorDExp1 = read.csv('error/fullSampleError.csv')
-preMessageDExp1 = read.csv('message/fullSampleMessage.csv')
-
-setwd('~/Repos/keysar_replication/experiment2/data/')
-preMouseDExp2 = read.csv('mouse/exp2Mouse.csv')
-preErrorDExp2 = read.csv('error/exp2Error.csv')
-preMessageDExp2 = read.csv('message/exp2Message.csv')
 library(tidyr)
@@
 library(binom)
+
+preMouseDExp1 = read.csv('../data/exp1Mouse.csv')
+preErrorDExp1 = read.csv('../data/exp1Error.csv')
+preMessageDExp1 = read.csv('../data/exp1Message.csv')
+
+preMouseDExp2 = read.csv('../data/exp2Mouse.csv')
+preErrorDExp2 = read.csv('../data/exp2Error.csv')
+preMessageDExp2 = read.csv('../data/exp2Message.csv')
 ```
@@
 library(irr)
-turkerRatings <- read.csv("../normingExperiment/data/keysarNorming-trials.csv")
+turkerRatings <- read.csv("../data/informativities.csv")
 relevantProperties <- c("objectSet", "label")
 
-criticalMessagesExp1 <- read.csv("../experiment1/data/processedLabelsExp1.csv") %>%
+criticalMessagesExp1 <- read.csv("../data/Exp1SpeakerLabels.csv") %>%
   left_join(turkerRatings, by = relevantProperties) %>%
   mutate(exp = "Exp.1")
-criticalMessagesExp2 <- read.csv("../experiment2/data/processedLabelsExp2.csv") %>%
+criticalMessagesExp2 <- read.csv("../data/Exp2SpeakerLabels.csv") %>%
   left_join(turkerRatings, by = relevantProperties) %>%
@@
 g
-ggsave("~/Repos/keysar_replication/writing/cogsci-revision/images/itemWiseFitness.pdf")
+ggsave("../../writing/cogsci-revision/images/itemWiseFitness.pdf")
 ```
@@
   labs(color="item")
-
-# Shoudln't worry too much about convergence errors: appear to be false positives
-# https://groups.google.com/forum/#!topic/predicts_project/gVfnVxCSYog
-
-# gg <- mmRes2@optinfo$derivs$grad
-# hh <- mmRes2@optinfo$derivs$Hessian
-# vv <- sqrt(diag(solve(hh/2)))
-# print(summary(abs(gg*vv)))
-# 
-# relgrad <- with(mmRes2@optinfo$derivs,solve(Hessian,gradient))
-# print(max(abs(relgrad)))
 
@@
 ```
-
-# Create input for BDA
-
-Create the informativities csv
-
-```{r}
-rbind(read_csv('../data/exp1SpeakerLabels.csv') %>% mutate(exp = 'scripted'),
-      read_csv('../data/exp2SpeakerLabels.csv') %>% mutate(exp = 'unscripted')) %>%
-  mutate(label = str_replace_all(label, fixed(' '), '')) %>%
-  mutate(label = str_replace_all(label, "[[:punct:]]", "")) %>% 
-  mutate(label = str_to_lower(label, locale = "en")) %>%
-  write_csv(path = '../modeling/bdaInput/speakerData.csv')
-```

### 2020-12-19 modeling/refModule/ref.js

@@ -1,206 +1,0 @@
-var _ = require('lodash');
-var fs = require('fs');
-var babyparse = require('babyparse');
-
-function _logsumexp(a) {
-  var m = Math.max.apply(null, a);
-  var sum = 0;
-  for (var i = 0; i < a.length; ++i) {
-    sum += (a[i] === -Infinity ? 0 : Math.exp(a[i] - m));
-  }
-  return m + Math.log(sum);
-}
-
-var getL0score = function(targetObj, utt, context, params) {
-  var similarities = params.lexicon[utt];
-  var a = [];
-  for(var i=0; i<context.length; i++){
-    a.append(params.simScale * similarities[context[i]]);
-  }
-  return (params.simScale * similarities[targetObj]
-	  - _logsumexp(a));
-};
-
-var powerset = function (set) {
-  if (set.length == 0)
-    return [[]];
-  else {
-    var rest = powerset(set.slice(1));
-    return rest.map(
-      function(element) {
-        return [set[0]].concat(element);
-      }).concat(rest);
-  }
-};
-
-// get all words that can be used to describe target
-var possibleUtts = function(target, lexicon) {
-  return _.keys(_.pickBy(lexicon, function(value, key) {
-    return _.has(value, target);
-  }));
-};
-
-// Note: assumes the only object considered by the listener is the true distractor...
-var possibleObjects = function(target) {
-  var objectSet = target.split('_')[1];
-  return [target, ['distractor', objectSet].join('_')];
-};
-
-var makeArr = function(n, v) {
-  return _.repeat(n, v);
-};
-
-var makeColorSizeLists = function(wordsOrObjects) {
-  var colorList = wordsOrObjects === 'words' ? colors.concat('') : colors;
-  var sizeList = wordsOrObjects === 'words' ? sizes.concat('') : sizes;
-  var typeList = wordsOrObjects === 'words' ? types.concat('thing') : types;
-
-  return _.flattenDepth(_.map(sizeList, function(size) {
-    return _.map(colorList, function(color) {
-      return _.map(typeList, function(type) {
-        return [size, color, type]
-      });
-    });
-  }), 2);
-};
-
-var colorSizeWordMeanings = function(params) {
-  return _.extend(
-    _.zipObject(colors, _.times(colors.length, _.constant(params.colorTyp))),
[…]

### 2020-12-19 modeling/run_AIS.sh

@@ -1,3 +1,0 @@
-#!/bin/bash#
-# parallel --bar --colsep ',' "sh ./run_AIS.sh {1}" :::: input/BF_grid.csv > "output/AIS2.txt"
-~/.webppl/webppl BDA.wppl --require ./refModule/ --require webppl-csv -- --model $1 --AIS true

### 2020-12-19 modeling/run_RR_grid.sh

@@ -1,3 +1,0 @@
-#!/bin/bash#
-# parallel --bar --colsep ',' "sh ./run_grid_search.sh {1} {2}" :::: input/grid.csv
-webppl RR_simulations.wppl --require ./refModule/ --require webppl-csv -- --perspectiveCost $1 --chainNum $2

### 2020-12-19 modeling/run_dynamics_grid.sh

@@ -1,3 +1,0 @@
-#!/bin/bash#
-# parallel --bar --colsep ',' "sh ./run_dynamics_grid.sh {1} {2}" :::: input/dynamics_grid.csv
-webppl dynamics_simulations.wppl --require ./refModule/ --require webppl-csv -- --numDatapoints $1 --chainNum $2
